### 2013-07-16 sections/Abstract.tex
initial commit.

Linear correlations between the spiking activity of pairs of neurons are among the most familiar and useful descriptive statistics of neural activity.
Correlation matrices from multineuronal recordings contain richer information than available from an equivalent number of isolated pairs:
The basic structure of population activity can manifest in various low-dimensional representations of the covariance matrix.  The optimal choice of such reduced representation  depends profoundly on how well it can capture regularities in the underlying data generating process. 
For example, presumed diffuse correlated input onto all neurons implies a low-rank parameterization such as used in  factor analysis whereas presumed strong pairwise linear associations imply sparse partial correlations.  
Composite  `sparse+low-rank partial correlation' parameterizations allow combining common correlated input with strong pairwise associations.

To infer the optimal low-dimensional representation of neural correlations in a specitifc neural microcircuit, we compared the performance of several estimators of the covariance matrix on the activity of 100--300 neurons in mouse visual cortex: sample covariance, covariance shrinkage, factor analysis, sparse partial correlations, and sparse+low-rank partial correlations. 
The instantaneous firing rates, binned at 200 ms, were inferred from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.   
As expected, covariance shrinkage reliably outperformed the sample covariance estimate.
Factor analysis-based estimates significantly outperformed covariance shrinkage.
Yet sparsified partial correlations with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage-based  estimators.
The superior performance of sparse partial correlation-based estimates suggests the relative importance of detailed network interactions over common input in this circuit.
The resulting sparse partial correlation matrix, we argue, serves as a better substrate for functional network analysis. 
In addition, the overall estimation improvement will propagate to any existing analyses that rely on the covariance matrix.

### 2013-07-16 sections/Discussion.tex
initial commit.

\hl{The discussion section is not ready for any feedback. Currently it's just a trashbin with clips from scrapped versions of the introduction. -- Dimitri}

Fundamental questions in systems neuroscience concern the relationship between the function of neural microcircuits and their cytoarchitecture: circuit topology, cell types, and patterns of synaptic connectivity.

Considerable progress has been made in sensory areas where the functional characterizations of cells are defined by their reproducible responses to external stimuli (reviewed in \citep{Reid:2012}).

Many key questions in neuroscience revolve around the relationship between the cytoarchitecture of neural microcircuits and the functional organization of their activity \citep{Reid:2012}.  Investigators in this line of inquiry aim to construct networks of functional associations within groups of neurons inferred from observations of their activity under a variety of conditions.  The inferred functional network  could then related to the synaptic connectivity patterns, cell types, cell properties, or the cells' spatial arrangment in the attempt to uncover organizational principles of neural computation. In addition, the network itself can serve as subtrate for subsequent graph-theoretical analysis \citep{Feldt:2011}

Advances in electrophysiology and optical imaging have enabled simultaneous recordings from dozens to hundreds or thousands of cells. 
In particular, recent advances in two-photon imaging of calcium signals have allowed in vivo recordings from nearly every neuron brain-wide in zebrafish \citep{Leung:2013,Ahrens:2013} and in a 3D volume of a few hundred microns in diameter in mouse neocortex \citep{Katona:2012,Cotton:2013}.    
Ambitious projects currently under way aim to record the spiking activity of large fractions of cells from entire circuits and systems in behaving animals \citep{Alivisatos:2012}.  Direct observations of the population activity  of entire circuits open new possibilities for incisive statistical descriptions of the functional connectivity in the circuit.

One general approach is to construct statistical models reflecting hypothesized models of neural circuit function, including temporal dynamics, nonlinear interactions between neurons, and background network activity \citep{Pillow:2008,Buonomano:2009,Yu:2009}. 
Another general approach is to search for alternative statistics that best describe the population activity without assuming a specific model.  Such evaluations rely on constructing surrogate datasets \citep{Okun:2012} or maximum-entropy distributions that reproduce the statistic of interest \citep{Ganmor:2011,Tkacik:2012} with subsequent testing of how well such constructs can reproduce other observed properties of population activity.

Such more sophisticated statistical descriptions are unlikely to entirely replace neural correlations as the initial descriptive statistic of population activity in experimental neuroscience. Neural correlations have several strengths that will ensure their continued prominence: (a) correlations are well established, familiar, and intuitive to most researchers in biological sciences, (b) correlations are easy to estimate as they are less susceptible to the curse of dimensionality than more sophisticated models, (c) correlations make relatively weak mechanistic assumptions (at the cost of not revealing many potentially important aspects), and (d) correlations serve as input into other models of population activity or decoding algorithms.

Linear correlations between pairs of neurons are among the most common and familiar descriptions of functional associations in networks and of the collective activity of neuronal populations.  For example, it is tempting to consider the network constructed from the highest correlations in the recorded population (e.g.~\cite{Malmersjo:2013}) for subsequent graph-theoretical analysis \citep{Feldt:2011}.  In functional genomics, networks constructed from highest correlations are called \emph{relevance networks}.

However, pairwise correlations unreliable proxies of direct functional association such as direct synaptic connectivity as they can arise due to a wide variety of physiological interactions including direct synaptic interactions, indirect chains of synaptic connections, common inputs, correlations in inputs, synchrony, fluctuations in global network states, and others \citep{Shadlen:1998,Ostojic:2009,Pernice:2011,Schneidman:2006}.

Other experimental evidence suggests that detailed interactions in local microcircuits have only weak effects on overall population activity, which is then best characterized by collective features such as global population dynamics or population synchrony \citep{Okun:2012,Tkacik:2012,Tkacik:2013}.  From this point of view, the imporant aspects of the correlation structure lie it its global aspects, such as the eigenspectrum, while the individual pairwise correlations are not particularly meaningful.

### 2013-07-16 sections/Introduction.tex
initial commit.

\hl{\tiny why correlations:}
Linear correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \citep{Cohen:2011}.   
For example, \emph{noise correlations}, i.e.~the correlations of stimulus response variability between pairs of neurons in sensory areas, have been of particular interest thanks, in part, to their profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011}.  
Lending further credence to their role as indicators of detailed and specific functional connectivity have been a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Smith:2013,Denman:2013}.

\hl{\tiny ambiguity of interpretation:}
Neural correlations do not come with ready or unambiguous mechanistic interpretations.   
Theoretical and simulation studies show that neural correlations at various temporal scales arise from multiple underlying mechanisms such as direct synaptic interactions,  common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009}

\hl{\tiny opportunities with multineuronal data:}
The neural correlation matrix from multineuronal recordings is greater than the sum of its parts: it can be transformed into other representations  that accentuate different aspects of population activity.  
For example, the eigenvalue decomposition of the covariance matrix, also known as principal component analysis (PCA), expresses shared activity components across the entire population;  common fluctuations of population activity of the entire circuit may be accurately represented by just a few principal components but may affect all correlation coefficients. 
In contrast, the inverse of the correlation matrix reveals the partial correlations between neuron pairs, which reflect their specific linear depedence, conditioned on the activity of all the other recorded cells;
a strong interaction between a pair of neurons may be expressly represented by a single partial correlation but its effect would propagate to multiple correlations and principal components.

\hl{\tiny motivate dimensionality reduction:}
 Covariance estimates from real high-dimensional data  are both noisy and redundant.  The amount of recorded data increases only linearly with population size whereas the number of free parameters in the covariance matrix increases quadratically; as a result, empirical covariance estimates for multidimensional data are  accumulate substantial redundancy and noise.

To reduce the estimation error, statistically efficient estimators are regularized.  One regularization approach is to shrink the estimate toward a less variable low-dimensional estimate \citep{Schafer:2005}.  Another regularization approach, often called \emph{sparsification}, reduces the dimensionality of the estimate by setting its least significant parameters to zero. 
Regularized estimates perform favorably on new samples of data compared to the full model, provided the amount of regularization is chosen carefully.

\hl{\tiny choosing between regularization schemes}
Estimator improvements due to regularization depend profoundly on the parameterization to which it is applied.  Sparsification of the eigenspectrum produces a low-rank estimate of the covariance matrix also known as truncated PCA \citep{Rothman:2008} or, allowing for additional independent variances, as factor analysis (FA) \citep{Fan:2008}.  Sparsification of the inverse covariance is known as covariance selection \citep{Dempster:1972,Friedman:2008} and is related to finding a graphical model in which a large fraction of neuron pairs are conditionally independent of each other.  Representations that are capable of concentrating the significant real effects in a small number of parameters will reap greatest benefits from sparsification over ones that distribute the real effects widelyacross many parameters.  Random noise tends to be broadly distributed across all parameters in any parameterization.

\hl{\tiny our approach and summar of findings} 
In this study, we compared the performance of several covariance estimation schemes for spatially compact groups of 150--300 neurons in layers 2-4 in mouse primary visual cortex during visual stimulation.   The performance of the estimators was evaluted by computing the mean squared error between the optimized covariance estimate fitted to training data and the non-regularized covariance matrix from a separate test data set.  Low-rank covariance estimates performed significantly better than shrinkage estimators, but estimators with sparse partial correlations were more efficient still. Typically, between 3 and 16\% of neuronal paris were connected by non-zero partial interactions.  Mixed sparse estimators with sparse partial correlations and low-rank components performed comparably.

As noted above, the optimal covariance estimator is domain-dependent. 
These specific findings may not spread to all neural circuits.
Improved accuracy estimation accuracy is not the main aim or the only benefit.  
The selection of the optimal regularization scheme in itself serves as an effective descriptor of the dominant correlation structures. The finding that a identifiable subsets of partial correlations were important to describe the partial correlation makes such pairs as prime candidates as direct functional interactions.

### 2013-07-16 sections/Theory.tex
initial commit.

\subsection*{Problem setup}
Let $x_1,\ldots,x_n \in \mathbb R^{p\times 1}$ denote a sample of $n$ observations of instantaneous activity (e.g.~spiking) rates of $p$ neurons averaged in time bins $i=1,\ldots,n$.

For simplicity, we assume that observations are sampled from a shared true distribution $x_1,\ldots,x_n \sim f$. 
\hl{\tiny explain: the derivations remain valid even without such a strong assumption.  We account for overdispersion due to temporal dependence and shuffle chunks of data to account for nonstationarities.}

For simplicity, we assume that the data have zero means.
We seek to estimate the true covariance matrix 
\begin{equation}
\Sigma=\mathbb E_{z\sim f}[ zz^\T]
\end{equation}

The purpose of the estimation procedure is to use the sample $x_1,\ldots,x_n$ to compute a covariance estimate $S$ that is as close as possible to $\Sigma$. 
The quality of the covariance estimate $S$ will be defined using a \emph{loss function} $\mathcal L\left(S, \Sigma\right)$, which is minimized when $S=\Sigma$.

In particular, we consider the mean-squared loss
\begin{equation}
\mathcal L_e( S, \Sigma) =  \| S - \Sigma\|^2_F 
\end{equation}
where $\| \cdot \|_F$ denotes the Frobenius norm;

and the multivariate normal log likelihood loss
\begin{equation}
\mathcal L_g(S,\Sigma) = \ln |S| +  \Tr(S^{-1} \Sigma)
\end{equation}
where $|S|$ denotes the determinant of matrix $S$.

\hl{\tiny list other possible loss functions and justify the choice of these two. Consider Stein's entropy loss and qudratic loss.   James:1961 showed failures of the MSE loss, but it continued to be widely used. Fan:2008 provides other arguments.}

The linearity of $\mathcal L_e(S,\Sigma)$ and $\mathcal L_g(S,\Sigma)$ with respect to the elements of $\Sigma$ allows unbiased empirical estimation of $\mathcal L$ from a validation sample $y_1,\ldots,y_m \sim f$ acquired independently of the original training sample $x_1,\ldots,x_n$ that was used to construct the estimate $S$.
\begin{equation}
\begin{split}
\mathcal L(S,\Sigma) & \equiv \mathbb E_{z\sim f} \left[ \mathcal L(S,zz^\T) \right] \\ 
&\approx \frac 1 m \sum\limits_{i=1}^m \mathcal L(S,y_i y_i^\T) = \mathcal L(S,\hat S_y)
\end{split}
\end{equation}
where $\hat S_y = \frac 1 m \sum\limits_{i=1}^m x_i x_i^\T$ is the empirical covariance of the validation dataset.

\hl{\tiny cross-validation}
In practice, neural recordings are limited.   Instead of validation with a dedicated validation dataset, $K$-fold cross-validation is used: The data are split into $K$ subsets of approximately equal size and each subset is used as the validation set with the covariance estimator trained on the remaning data.  The estimation of the loss from the $K$ tests is then averaged, which is justified, again, thanks to the linearity property of the chosen loss functions. 
\hl{\tiny exlain how the data are split into continuous chunks so that dependences between the training and validation datasets due to temporal correlations are minimized.  Coversely, each test set comprises multiple distirbuted chunks to reduce the effect of non-stationarities.}

\subsection*{Covariance estimators}
It has long been recognized that the empirical covariance 
\begin{equation}
\hat S = \frac 1 n \sum\limits_{i=1}^n x_i x_i^\T
\end{equation}
is statistically inadmissible, meaning that larger sample sizes are required to reduce the loss function to the same level as other, more statistically efficient covariance estimators.

Covariance estimators can be made more efficient by two broad approaches, which are not mutually exclusive: (1) deliberately bias the estimator in order to reduce its variability and (b) reduce the dimensionality of the estimator by choosing a specific parametric model.

\subsubsection*{Shrinkage estimators}
The most familiar example of the first approach is the linear shrinkage estimator, calculated is a linear mixture of the empirical covariance estimate $\hat S$ and another low-variance target estimator $T$:
\begin{equation}
S_{shrink} = (1-\lambda) \hat S + \lambda T
\end{equation}
A popular target covariance is the identity matrix scaled by the mean sample variance  $T = \frac 1 p \Tr(\hat S) I$.
The optimal value of shrinkage estimate with respect to $\mathcal L_e$ can be estimated analytically from the data \citep{Ledoit:2004,Schafer:2005}.

\subsubsection*{Dimensionality reduction}
The second broad approach is to impose a structure on the covariance estimate by reducing the number of free parameters in one of its parameterizations.

### 2013-07-23 frontiers.tex
converted to Frontiers' template

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just a template to use when submitting manuscripts to Frontiers, it is not mandatory to use frontiers.cls nor frontiers.tex  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\documentclass{frontiersENG} % for Engineering articles
\documentclass{frontiersSCNS} % for Science articles
%\documentclass{frontiersMED} % for Medicine articles

\usepackage{url,lineno}
\linenumbers

% Leave a blank line between paragraphs in stead of using \\

\copyrightyear{}
\pubyear{}

\def\journal{Bioengineering}%%% write here for which journal %%%
\def\DOI{}
\def\articleType{Research Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Sample {et~al.}} %use et al only if is more than 1 author
\def\Authors{First Author\,$^{1,*}$, Co-Author\,$^{2}$ and Co-Author\,$^2$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X \\
$^{2}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X  }
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{corresponding Author}
\def\corrAddress{Laboratory X, Institute X, Department X, Organization X, Street X, City X , State XX (only USA, Canada and Australia), Zip Code, X Country X}
\def\corrEmail{email@uni.edu}

% \color{FrontiersColor} Is the color used in the Journal name, in the title, and the names of the sections

\begin{document}
\onecolumn
\firstpage{1}

\title[Running Title]{Article Title}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\editor{}
\topic{Research Topic}

\maketitle
\begin{abstract}

\section{}
%As a primary goal, the abstract should render the general significance and conceptual advance of the work clearly accessible to a broad readership. References should not be cited in the abstract.
See the Summary Table at \\ \url{http://www.frontiersin.org/}\texttt{\journal}\url{/authorguidelines} \\for abstract requirement and length according to article type.

\tiny
 \keyFont{ \section{Keywords:} Text Text Text Text Text Text Text Text } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

\section{Introduction}

% For Original Research Articles, Clinical Trial Articles, and Technology Reports the introduction should be succinct, with no subheadings.
%
%For Clinical Case Studies the Introduction should include symptoms at presentation, physical exams and lab results.
%
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text. Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text. Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text.Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text.

%\begin{methods}
\section{Material \& Methods}

Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.  \cite{Neuro2013} might want to know about  text text text text Text Text Text Text  Text Text Text Text Text Text  Text Text. \citep{Gene2012} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.  \cite{Neurobot2013} might want to know about  text text text text

\subsection{Original Research Articles, Clinical Trial Articles, and Technology Reports}

For Original Research Articles, Clinical Trial Articles, and Technology Reports the following sections are mandatory:

\begin{itemize}
%for bulleted list, use itemize
\item Introduction: Succinct, with no subheadings.
\item Materials and Methods: This section may be divided by subheadings. This section should contain sufficient detail so that when read in conjunction with cited references, all procedures can be repeated.
\item Results: This section may be divided by subheadings. Footnotes should not be used and have to be transferred into the main text.
\item Discussion: This section may be divided by subheadings. Discussions should cover the key findings of the study: discuss any prior art related to the subject so to place the novelty of the discovery in the appropriate context; discuss the potential short-comings and limitations on their interpretations; discuss their integration into the current understanding of the problem and how this advances the current views; speculate on the future direction of the research and freely postulate theories that could be tested in the future.
\end{itemize}

Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.

\textbf{Figure 1.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures.}\label{fig:01}% Don't add the figures in the LaTeX files, please upload them when submitting the article. Frontiers will add the figures at the end of the provisional pdf.

\subsection{Clinical Case Studies}

For Clinical Case Studies the following sections are mandatory:

\begin{itemize}
%for bulleted list, use itemize
\item Introduction: Include symptoms at presentation, physical exams and lab results.
\item Background: This section may be divided by subheadings. Include history and review of similar cases.
\item Results: This section may be divided by subheadings. Include diagnosis and treatment.
\item Concluding Remarks
\end{itemize}

Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.

%\end{methods}

\section{Results}

Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript. Figures will then be automatically embedded at the bottom of the submitted manuscript. Kindly ensure that each table and figure is mentioned in the text and in numerical order. Permission must be obtained for use of copyrighted material from other sources (including the web). Please note that it is compulsory to follow figure instructions. Figures which are not according to the guidelines will cause substantial delay during the production process.

\begin{table}[!t]
\processtable{Resolution Requirements for the figures\label{Tab:01}}
{\begin{tabular}{lllll}\toprule
Image Type & Description & Format & Color Mode & Resolution\\\midrule
Line Art & An image composed of lines and text,  & TIFF, EPS, JPEG & RGB, Bitmap & 900 - 1200 dpi\\
           & which does not contain tonal or shaded areas.& & &\\
           Halftone & A continuous tone photograph, which contains no text. & TIFF, EPS, JPEG & RGB, Grayscale & 300 dpi\\
Combination & Image contains halftone + text or line art elements. & TIFF, EPS, JPEG & RGB,Grayscale & 600 - 900 dpi\\\botrule
\end{tabular}}{This is a footnote}
\end{table}

\begin{equation}
\sum x+ y =Z\label{eq:01}
\end{equation}

\textbf{Table\ref{Tab:01}} shows the resolution requirements for the figures. The figures must be legible:
\begin{enumerate}
\item The smallest visible text is no less than 8 points in height, when viewed at actual size.
\item Solid lines are not broken up.
\item Image areas are not pixelated or stair stepped.
\item Text is legible and of high quality.
\item Any lines in the graphic are no smaller than 2 points width.
\end{enumerate}

\textbf{Figure 2.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures.}\label{fig:02}

\section{Discussion}

Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text.
Additional Requirements:
\subsection{Corrections}

Minor corrections to published articles can be communicated to the Frontiers Production Office at production.office@frontiersin.org. If you need to communicate important changes to an article please submit a General Commentary. Submit the article with the title “Erratum: Original Title of Article”.

\subsection{Commentaries on Articles}

At the beginning of your manuscript provide the citation of the article commented on.

\subsection{Focused Reviews}

For Tier 2 invited Focused Reviews the sections Introduction, Material and Methods, Results, and Discussion are recommended. In addition the authors must submit a short biography of the corresponding author(s). This short biography has a maximum of 600 characters, including spaces.

A picture (5 x 5 cm, in *.tif or *.jpg, min 300 dpi) must be submitted along with the biography in the manuscript and separately during figure upload.
Focused Reviews highlight and explain key concepts of your work. Please highlight a minimum of four and a maximum of ten key concepts in bold in your manuscript and provide the definitions/explanations at the end of your manuscript under “Key Concepts”. Each definition has a maximum of 400 characters, including spaces.

\subsection{Human Search and Animal Research}

All experiments on live vertebrates or higher invertebrates must be performed in accordance with relevant institutional and national guidelines and regulations. In the manuscript, authors must identify the committee approving the experiments and must confirm that all experiments conform to the relevant regulatory standards. For manuscripts reporting experiments on human subjects, authors must identify the committee approving the experiments and must also include a statement confirming that informed consent was obtained from all subjects. In Original Research Articles and Clinical Trial Articles these statements should appear in the Materials and Methods section.

\subsection{Clinical Trial Registration}

Clinical trials should be registered in a public trials registry in order to become the object of a publication at Frontiers. Trials must be registered at or before the start of patient enrollment. A clinical trial is defined as"any research study that prospectively assigns human participants or groups of humans to one or more health-related interventions to evaluate the effects on health outcomes."(\url{www.who.int/ictrp/en}). A list of acceptable registries can be found at \url{www.who.int/ictrp/en and www.icmje.org}.

\subsection{Inclusion of Proteomics Data}

Authors should provide relevant information relating to how the peptide/protein matches were undertaken, including methods used to process and analyze data, false discovery rates (FDR) for large-scale studies and threshold or cut-off rates for peptide and protein matches. Further information could include software used, mass spectrometer type, sequence database and version, number of sequences in database, processing methods, mass tolerances used for matching, variable/fixed modifications, allowable missed cleavages, etc.

Authors should provide as supplementary material information used to identify proteins and/or peptides. This should include information such as accession numbers, observed mass (m/z), charge, delta mass, matched mass, peptide/protein scores, peptide modification, miscleavages, peptide sequence, match rank, matched species (for cross species matching), number of peptide matches, ambiguous protein/peptide matches should be indicated, etc.
For quantitative proteomics analyses authors should provide information to justify the statistical significance including biological replicates, statistical methods, estimates of uncertainty and the methods used for calculating error.

For peptide matches with biologically relevant post-translational modifications (PTM) and for any protein match that has occurred using a single mass spectrum, authors should include this information as raw data, annotated spectra or submit data to an online repository (recommended option).
Authors are encouraged to submit raw or matched data and 2-DE images to public proteomics repositories. Submission codes and/or links to data should be provided within the manuscript.

\subsection{Data Sharing}

Frontiers supports the policy of data sharing, and authors are advised to make freely available any materials and information described in their article, and any data relevant to the article (while not compromising confidentiality in the context of human-subject research) that may be reasonably requested by others for the purpose of academic and non-commercial research. In regards to deposition of data and data sharing through databases, Frontiers urges authors to comply with the current best practices within their discipline.

\section*{Disclosure/Conflict-of-Interest Statement}
%All relationships financial, commercial or otherwise that might be perceived by the academic community as representing a potential conflict of interest must be described. If no such relationship exists, authors will be asked to declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Acknowledgement}
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.

\paragraph{Funding\textcolon} Text Text Text Text Text Text  Text Text.

\section*{Supplemental Data}
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.

\bibliographystyle{frontiersinSCNS&ENG} % for Science and Engineering articles
%\bibliographystyle{frontiersinMED} % for Medicine articles
\bibliography{test}

\end{document}

### 2013-07-23 sections/Abstract.tex
converted to Frontiers' template

@@
 Linear correlations between the spiking activity of pairs of neurons are among the most familiar and useful descriptive statistics of neural activity.
-Correlation matrices from multineuronal recordings contain richer information than available from an equivalent number of isolated pairs:
-The basic structure of population activity can manifest in various low-dimensional representations of the covariance matrix.
-The optimal choice of such reduced representation  depends profoundly on how well it can capture regularities in the underlying data generating process.
-For example, presumed diffuse correlated input onto all neurons implies a low-rank parameterization such as used in  factor analysis whereas presumed strong pairwise linear associations imply sparse partial correlations.
-Composite  `sparse+low-rank partial correlation' parameterizations allow combining common correlated input with strong pairwise associations.
+Multineuronal recordings provide far richer information than the equivalent number of neuron pairs considered in isolation.
+For example, the full covariance matrix reveals the partial correlations between pairs of neurons or the common correlated activity involving the entire population.
+Usual estimates of covariance matrix parameters require large samples for convergence.
+Convergence can be improved by regularization, i.e.~by imposing some kind of structure.  
+The optimal choice of the regularization approach is largely an empirical question as it depends profoundly on how well the imposed structure can still capture the dominant underlying physiogical interactions.
+For example, presumed diffuse correlated input onto many neurons motivate low-rank parameterizations such as in factor analysis.  
+Conversely, presumed strong pairwise linear associations between neurons motivate regularization by sparsifying the inverse covariance matrix.
+Finally, `sparse+low-rank' inverse covariance parameterizations allow combining diffuse correlations and strong pairwise associations.
 
-To infer the optimal low-dimensional representation of neural correlations in a specitifc neural microcircuit, we compared the performance of several estimators of the covariance matrix on the activity of 100--300 neurons in mouse visual cortex: sample covariance, covariance shrinkage, factor analysis, sparse partial correlations, and sparse+low-rank partial correlations.
+To infer the optimal structure of neural correlations in a specitifc neural microcircuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, Ledoit-Wolf covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance.
 The instantaneous firing rates, binned at 200 ms, were inferred from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.   
-As expected, covariance shrinkage reliably outperformed the sample covariance estimate.
-Factor analysis-based estimates significantly outperformed covariance shrinkage.
-Yet sparsified partial correlations with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage-based  estimators.
-The superior performance of sparse partial correlation-based estimates suggests the relative importance of detailed network interactions over common input in this circuit.
-The resulting sparse partial correlation matrix, we argue, serves as a better substrate for functional network analysis. 
-In addition, the overall estimation improvement will propagate to any existing analyses that rely on the covariance matrix.
+Each covariance estimator was optimized and evaluated by cross-validation.
+Unsurprisingly, covariance shrinkage reliably outperformed the sample covariance estimate.
+In turn, factor analysis-based estimates significantly outperformed covariance shrinkage.
+Yet sparse inverse covariance with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage  estimators.
+The superior performance of the sparse inverse covariance estimator suggests the relative importance of detailed network interactions over common diffuse input in this circuit, although these results may vary in other circuits or in other brain states.

### 2013-07-23 sections/Introduction.tex
converted to Frontiers' template

@@
 For example, \emph{noise correlations}, i.e.~the correlations of stimulus response variability between pairs of neurons in sensory areas, have been of particular interest thanks, in part, to their profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011}.
-Lending further credence to their role as indicators of detailed and specific functional connectivity have been a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Smith:2013,Denman:2013}.
+Lending further credence to their role as indicators of detailed and specific functional organization have been a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Smith:2013,Denman:2013}.
 
@@
 Neural correlations do not come with ready or unambiguous mechanistic interpretations.   
-Theoretical and simulation studies show that neural correlations at various temporal scales arise from multiple underlying mechanisms such as direct synaptic interactions,  common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009}
+Theoretical and simulation studies show that neural correlations at various temporal scales may arise from any combination of underlying mechanisms such as direct synaptic interactions,  common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}
 
 \hl{\tiny opportunities with multineuronal data:}
-The neural correlation matrix from multineuronal recordings is greater than the sum of its parts: it can be transformed into other representations  that accentuate different aspects of population activity.
-For example, the eigenvalue decomposition of the covariance matrix, also known as principal component analysis (PCA), expresses shared activity components across the entire population;  common fluctuations of population activity of the entire circuit may be accurately represented by just a few principal components but may affect all correlation coefficients.
-In contrast, the inverse of the correlation matrix reveals the partial correlations between neuron pairs, which reflect their specific linear depedence, conditioned on the activity of all the other recorded cells;
-a strong interaction between a pair of neurons may be expressly represented by a single partial correlation but its effect would propagate to multiple correlations and principal components.
+Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \citep{Shadlen:1998,Zohary:1994}.
+Modern massively multineuronal recordings allow estimations of entire correlation matrices from large neuronal populations, which provide more information than the equivalent number of pairwise correlations assessed in isolation.
+After all, the correlation matrix is greater than the sum of its parts: it can be transformed into other representations  that accentuate different aspects of its structure and may suggest different mechanistic interpretations.
+For example, the eigenvalue decomposition, also known as the principal component analysis, of the covariance matrix expresses shared correlated activity components across the entire population;  common fluctuations of population activity of the entire circuit may be accurately represented by just a few principal components but will affect all correlation coefficients.
+In contrast, the off-diagonal elements of the inverse of the correlation matrix constitute scaled partial correlations between neuron pairs, which reflect their specific linear dependencies, after accounting for all other recorded cells;
+a strong interaction between a pair of neurons may be expressly represented by a single partial correlation but its effects propagate to multiple correlations and principal components.
 
-\hl{\tiny motivate dimensionality reduction:}
- Covariance estimates from real high-dimensional data  are both noisy and redundant.
-The amount of recorded data increases only linearly with population size whereas the number of free parameters in the covariance matrix increases quadratically; as a result, empirical covariance estimates for multidimensional data are  accumulate substantial redundancy and noise.
+\hl{\tiny motivate regularization} Estimation of parameters of the covariance matrix from data is inherently challenging, increasingly for very high-dimensional data.
+As the amount of data grows only linearly with population size, the number of free parameters in a complete representation of the covariance matrix increases quadratically, multiplying opportunities for spurious patterns to emerge among the correlations.
+Although the estimation error of each covariance coefficient, in isolation, is unaffected by the dimensionality of the data, the estimation error of other parameters of the covariance matrix increases with population size.
+For example, the large eigenvalues of in the sample covariance matrix are biased upward while small eigenvalues are biased downward, with both variance and bias increasingly large with increasing data dimensionality \citep{Hayes:1981}.
+Similarly, the coefficients of the inverse covariance matrix require  larger sample sizes to be estimated accurately in high-dimensional data.
 
-To reduce the estimation error, statistically efficient estimators are regularized.
-One regularization approach is to shrink the estimate toward a less variable low-dimensional estimate \citep{Schafer:2005}.
-Another regularization approach, often called \emph{sparsification}, reduces the dimensionality of the estimate by setting its least significant parameters to zero.
-Regularized estimates perform favorably on new samples of data compared to the full model, provided the amount of regularization is chosen carefully.
+\hl{\tiny forms of regularization} The convergence of the estimate of multivariate parameters to their true values can be improved through \emph{regularization},  Regularization is the deliberate biasing or \emph{shrinkage} of the empirical estimate toward a less variable \emph{target estimate} \citep{Schafer:2005} or the deliberate restriction of the estimate to a lower-dimensional space \citep{Dempster:1972,Fan:2008,Friedman:2008,Rothman:2008}.
+\hl{\tiny The two definitions overlap and are often equivalent, but not always, according to James, but I would like to see an example of shrinkage that is not, effectively, a low dimensional parameterization of the original problem.}
+One intuitive explanation of regularization is as the addition of some  prior knowledge about the likely forms of the covariance matrix.
+Curiously, though, regularization need not reflect any accurate knowledge about the nature of covariance matrices in general or in the particular domain.
+Some improvement arises due to \emph{Stein's phenomenon}, i.e.
+the favorable tradeoff between bias and variability \emph{anytime} a moderate bias toward a less variable target is introduced.
+However, the incorporation into the regularization scheme of real knowledge about the probable forms of covariance matrices for a specific situation will likely confer additional advantage and improve the convergence more substantially than an \emph{ad hoc} regularizer.
 
-\hl{\tiny choosing between regularization schemes}
-Estimator improvements due to regularization depend profoundly on the parameterization to which it is applied.
+%Estimator improvements due to regularization depend profoundly on the parameterization to which it is applied.
 Sparsification of the eigenspectrum produces a low-rank estimate of the covariance matrix also known as truncated PCA \citep{Rothman:2008} or, allowing for additional independent variances, as factor analysis (FA) \citep{Fan:2008}.
@@
 
-\hl{\tiny our approach and summar of findings} 
+Two questions must be answered in the choice of the regularization scheme: what kind of regularization and how much regularization.
+The answer to the first question is domain-specific and is often motivated by theories of the origins of correlations in the specific subject domain.
+
+\hl{\tiny approach and summary of findings} 
 In this study, we compared the performance of several covariance estimation schemes for spatially compact groups of 150--300 neurons in layers 2-4 in mouse primary visual cortex during visual stimulation.

### 2013-07-23 sections/Theory.tex
converted to Frontiers' template

@@
-\subsection*{Problem setup}
+\subsection{Problem setup}
 Let $x_1,\ldots,x_n \in \mathbb R^{p\times 1}$ denote a sample of $n$ observations of instantaneous activity (e.g.~spiking) rates of $p$ neurons averaged in time bins $i=1,\ldots,n$.
@@
 For simplicity, we assume that observations are sampled from a shared true distribution $x_1,\ldots,x_n \sim f$. 
-\hl{\tiny explain: the derivations remain valid even without such a strong assumption.
-We account for overdispersion due to temporal dependence and shuffle chunks of data to account for nonstationarities.}
+\hl{\tiny explain: the derivations remain valid even without such a strong assumption.}
 
@@
 
-In particular, we consider the mean-squared loss
+In particular, we consider the sum of squared-error loss
 \begin{equation}
@@
 
-and the multivariate normal log likelihood loss
+Another useful loss denoted here as $\mathcal L_g$ is based  on the log likelihood of the multivarate normal distribution 
+$$\ln L(\Sigma \cond S) = - \tfrac 1 2 k\ln(2\pi) - \tfrac 1 2 \ln |S| - \tfrac 1 2 \Tr(S^{-1}\Sigma)$$ 
+where $|S|$ denotes the determinant of matrix $S$.
+
+The loss function is shifted so that $\mathcal L_g(\Sigma,\Sigma) = 0$:
 \begin{equation}
-\mathcal L_g(S,\Sigma) = \ln |S| +  \Tr(S^{-1} \Sigma)
+\begin{split}
+\mathcal L_g(S,\Sigma) =&  -2 \left[\ln L(\Sigma\cond S) - \ln L(\Sigma\cond \Sigma)\right]
+\\=& \Tr(S^{-1}\Sigma) -\ln |S^{-1}\Sigma| - p
+\end{split}
 \end{equation}
-where $|S|$ denotes the determinant of matrix $S$.
+
+Stein's loss can also be interpreted as the Kullback-Leibler divergence between the multivarate normal with covariance $\Sigma$ and another with $S$:
+\begin{equation}
+\mathcal L_{s} = 2 D_{KL} (\mathcal N(0,\Sigma) \| S)
+\end{equation}
 

### 2013-07-24 sections/Abstract.tex

@@
 Multineuronal recordings provide far richer information than the equivalent number of neuron pairs considered in isolation.
-For example, the full covariance matrix reveals the partial correlations between pairs of neurons or the common correlated activity involving the entire population.
-Usual estimates of covariance matrix parameters require large samples for convergence.
-Convergence can be improved by regularization, i.e.~by imposing some kind of structure.  
-The optimal choice of the regularization approach is largely an empirical question as it depends profoundly on how well the imposed structure can still capture the dominant underlying physiogical interactions.
-For example, presumed diffuse correlated input onto many neurons motivate low-rank parameterizations such as in factor analysis.  
-Conversely, presumed strong pairwise linear associations between neurons motivate regularization by sparsifying the inverse covariance matrix.
-Finally, `sparse+low-rank' inverse covariance parameterizations allow combining diffuse correlations and strong pairwise associations.
+For example, the full covariance matrix reveals the partial correlations between pairs of neurons, as well as correlated activity across the entire population.
+Covariance matrix estimates require large samples for convergence.
+Convergence  can be improved by regularization, \emph{i.e.}~by imposing some kind of structure.
+The optimal regularization must be determined empirically since its performance depends on how the imposed structure captures the underlying interactions.
+For example, we can use low-rank parameterizations of the covariance matrix to account for correlated input onto many neurons.
+Conversely, if correlations are a result of a small number of  pairwise linear interactions between the observed neurons, we can sparsify the inverse covariance matrix.
+We can also use a `sparse+low rank’ inverse covariance representation to account for common fluctuations and pairwise interactions in the recorded population.
 
-To infer the optimal structure of neural correlations in a specitifc neural microcircuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, Ledoit-Wolf covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance.
-The instantaneous firing rates, binned at 200 ms, were inferred from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.   
+To select the optimal structure of neural correlations in a local neural circuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, Ledoit-Wolf covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance.
+We inferred instantaneous firing rates in 200 ms bins from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.
 Each covariance estimator was optimized and evaluated by cross-validation.
-Unsurprisingly, covariance shrinkage reliably outperformed the sample covariance estimate.
+As expected, covariance shrinkage reliably outperformed the sample covariance estimate.
 In turn, factor analysis-based estimates significantly outperformed covariance shrinkage.
 Yet sparse inverse covariance with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage estimators.
-The superior performance of the sparse inverse covariance estimator suggests the relative importance of detailed network interactions over common diffuse input in this circuit, although these results may vary in other circuits or in other brain states.
+The superior performance of the sparse inverse covariance estimator suggests the relative importance of detailed network interactions over common diffuse input in the circuit we studied.

### 2013-07-24 sections/Introduction.tex

@@
-\hl{\tiny why correlations:}
-Linear correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \citep{Cohen:2011}.
-For example, \emph{noise correlations}, i.e.~the correlations of stimulus response variability between pairs of neurons in sensory areas, have been of particular interest thanks, in part, to their profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011}.
-Lending further credence to their role as indicators of detailed and specific functional organization have been a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Smith:2013,Denman:2013}.
+\hl{\tiny why correlations:} Linear correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \citep{Cohen:2011}.
+For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been of particular interest.
+Noise correlations can have profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011}, and have been interpreted to indicate detailed and specific functional organization.
+Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Smith:2013,Denman:2013}.
 
-\hl{\tiny ambiguity of interpretation:}
-Neural correlations do not come with ready or unambiguous mechanistic interpretations.   
-Theoretical and simulation studies show that neural correlations at various temporal scales may arise from any combination of underlying mechanisms such as direct synaptic interactions,  common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}
+\hl{\tiny ambiguity of interpretation:} Neural correlations do not come with ready or unambiguous mechanistic interpretations.
+Theoretical and simulation studies show that neural correlations at various temporal scales may arise from any combination of underlying mechanisms.
+These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}
 
-\hl{\tiny opportunities with multineuronal data:}
-Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \citep{Shadlen:1998,Zohary:1994}.
-Modern massively multineuronal recordings allow estimations of entire correlation matrices from large neuronal populations, which provide more information than the equivalent number of pairwise correlations assessed in isolation.
-After all, the correlation matrix is greater than the sum of its parts: it can be transformed into other representations  that accentuate different aspects of its structure and may suggest different mechanistic interpretations.
-For example, the eigenvalue decomposition, also known as the principal component analysis, of the covariance matrix expresses shared correlated activity components across the entire population;  common fluctuations of population activity of the entire circuit may be accurately represented by just a few principal components but will affect all correlation coefficients.
-In contrast, the off-diagonal elements of the inverse of the correlation matrix constitute scaled partial correlations between neuron pairs, which reflect their specific linear dependencies, after accounting for all other recorded cells;
-a strong interaction between a pair of neurons may be expressly represented by a single partial correlation but its effects propagate to multiple correlations and principal components.
+\hl{\tiny opportunities with multineuronal data:} Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \citep{Shadlen:1998,Zohary:1994}.
+Multineuronal recordings allow estimation of covariance matrices of larger populations.
+Such estimates provide more information than the equivalent number of pairwise correlations assessed in isolation.
+Indeed, the correlation matrix is greater than the sum of its parts: it can be transformed into other representations that accentuate different aspects of the correlation structure and may suggest different mechanistic interpretations.
+For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients.
+In contrast, the off-diagonal elements of the inverse of the correlation matrix constitute scaled partial correlations between neuron pairs, which reflect their specific linear dependencies, after accounting for the activity of all the other recorded cells; a strong interaction between a pair of neurons may be expressed by a single partial correlation but its effects may propagate to multiple correlations and eigenvalues.
 
-\hl{\tiny motivate regularization} Estimation of parameters of the covariance matrix from data is inherently challenging, increasingly for very high-dimensional data.
+\hl{\tiny motivate regularization} Estimation of parameters of the covariance matrix from data is inherently challenging, increasingly so for very high-dimensional data.
 As the amount of data grows only linearly with population size, the number of free parameters in a complete representation of the covariance matrix increases quadratically, multiplying opportunities for spurious patterns to emerge among the correlations.
@@
 \hl{\tiny forms of regularization} The convergence of the estimate of multivariate parameters to their true values can be improved through \emph{regularization},  Regularization is the deliberate biasing or \emph{shrinkage} of the empirical estimate toward a less variable \emph{target estimate} \citep{Schafer:2005} or the deliberate restriction of the estimate to a lower-dimensional space \citep{Dempster:1972,Fan:2008,Friedman:2008,Rothman:2008}.
-\hl{\tiny The two definitions overlap and are often equivalent, but not always, according to James, but I would like to see an example of shrinkage that is not, effectively, a low dimensional parameterization of the original problem.}
-One intuitive explanation of regularization is as the addition of some  prior knowledge about the likely forms of the covariance matrix.
-Curiously, though, regularization need not reflect any accurate knowledge about the nature of covariance matrices in general or in the particular domain.
-Some improvement arises due to \emph{Stein's phenomenon}, i.e.
-the favorable tradeoff between bias and variability \emph{anytime} a moderate bias toward a less variable target is introduced.
-However, the incorporation into the regularization scheme of real knowledge about the probable forms of covariance matrices for a specific situation will likely confer additional advantage and improve the convergence more substantially than an \emph{ad hoc} regularizer.
+A common intuitive explanation for the effect of regularization is that regularization adds some prior knowledge about the likely forms of the covariance matrix.
+Curiously, though, regularization need not reflect any plausible knowledge of the nature of covariance matrices in general or in the particular domain.
+Convergence can improve due to a purely statistical effect, sometimes known as \emph{Stein's phenomenon}, \emph{i.e.}~the favorable tradeoff between bias and variability, which arises \emph{anytime} a moderate bias toward a less variable target is mixed into an unbiased estimate.
+Therefore, an estimator that is shown to converge faster on specific types of data does not necessarily contain within itself some wisdom of the statistical regularities in the data.
+However, an estimator that does happen to accommodate the statistical structure of the data in a specific domain, it will gain additional advantage over an \emph{ad hoc} regularizer for that domain.
+Thus the optimal selection of the estimator for a specific problem with unknown statistical structure becomes an empirical question.
+And, if a clear winner emerges from a number of strong candidates, its regularization scheme can server as an effective statistical description of the covariance structure.
 
-%Estimator improvements due to regularization depend profoundly on the parameterization to which it is applied.
-Sparsification of the eigenspectrum produces a low-rank estimate of the covariance matrix also known as truncated PCA \citep{Rothman:2008} or, allowing for additional independent variances, as factor analysis (FA) \citep{Fan:2008}.
-Sparsification of the inverse covariance is known as covariance selection \citep{Dempster:1972,Friedman:2008} and is related to finding a graphical model in which a large fraction of neuron pairs are conditionally independent of each other.
-Representations that are capable of concentrating the significant real effects in a small number of parameters will reap greatest benefits from sparsification over ones that distribute the real effects widelyacross many parameters.
-Random noise tends to be broadly distributed across all parameters in any parameterization.
-
-Two questions must be answered in the choice of the regularization scheme: what kind of regularization and how much regularization.
-The answer to the first question is domain-specific and is often motivated by theories of the origins of correlations in the specific subject domain.
-
-\hl{\tiny approach and summary of findings} 
-In this study, we compared the performance of several covariance estimation schemes for spatially compact groups of 150--300 neurons in layers 2-4 in mouse primary visual cortex during visual stimulation.
-The performance of the estimators was evaluted by computing the mean squared error between the optimized covariance estimate fitted to training data and the non-regularized covariance matrix from a separate test data set.
+\hl{\tiny approach and summary of findings} In this study, we compared the performance of several covariance estimators for spatially compact groups of 150--300 neurons in layers 2--4 in mouse primary visual cortex during visual stimulation.
+The performance of the estimators was evaluated by computing the mean squared error between the optimized covariance estimate fitted to training data and the non-regularized covariance matrix from a separate test data set.
 Low-rank covariance estimates performed significantly better than shrinkage estimators, but estimators with sparse partial correlations were more efficient still.
-Typically, between 3 and 16\% of neuronal paris were connected by non-zero partial interactions.
+Typically, between 3 and 16\% of neuronal pairs were connected by non-zero partial interactions.
 Mixed sparse estimators with sparse partial correlations and low-rank components performed comparably.
 
-As noted above, the optimal covariance estimator is domain-dependent. 
-These specific findings may not spread to all neural circuits.
+\hl{\tiny summary of conlusions} As noted above, the optimal covariance estimator is domain-dependent. These specific findings may not spread to all neural circuits.
 Improved accuracy estimation accuracy is not the main aim or the only benefit.  
 The selection of the optimal regularization scheme in itself serves as an effective descriptor of the dominant correlation structures.
-The finding that a identifiable subsets of partial correlations were important to describe the partial correlation makes such pairs as prime candidates as direct functional interactions.
+The finding that identifiable subsets of partial correlations were important to describe the partial correlation makes such pairs as prime candidates as direct functional interactions.

### 2013-07-24 sections/Abstract.tex
expanded the theory section.

@@
-Linear correlations between the spiking activity of pairs of neurons are among the most familiar and useful descriptive statistics of neural activity.
-Multineuronal recordings provide far richer information than the equivalent number of neuron pairs considered in isolation.
-For example, the full covariance matrix reveals the partial correlations between pairs of neurons, as well as correlated activity across the entire population.
-Covariance matrix estimates require large samples for convergence.
-Convergence  can be improved by regularization, \emph{i.e.}~by imposing some kind of structure.
-The optimal regularization must be determined empirically since its performance depends on how the imposed structure captures the underlying interactions.
-For example, we can use low-rank parameterizations of the covariance matrix to account for correlated input onto many neurons.
-Conversely, if correlations are a result of a small number of  pairwise linear interactions between the observed neurons, we can sparsify the inverse covariance matrix.
-We can also use a `sparse+low rank’ inverse covariance representation to account for common fluctuations and pairwise interactions in the recorded population.
+Correlations between the spiking activity of pairs of neurons are among the most familiar and useful descriptive statistics of neural activity.
+Multineuronal recordings provide richer information than the equivalent number of neuron pairs.
+For example, the full covariance matrix reveals correlated activity across the entire population as well as partial correlations between pairs.
+Estimation of covariance matrices from finite recordings can be improved by regularization, \emph{i.e.}\;by imposing some kind of structure.
+Optimal regularization must be determined empirically since its performance depends on how well the imposed structure captures the underlying regularities.
+For example, we can use low-rank parameterizations of the covariance matrix to account for common activity fluctuations across the recorded population.
+Conversely, if correlations are strongly influenced by a small fraction of possible pairwise linear associations between the observed neurons, we can impose sparsity on the inverse covariance matrix.
+We can also use a `sparse+low rank’ inverse covariance representation to account for both common fluctuations and pairwise interactions.
 

### 2013-07-24 sections/Introduction.tex
expanded the theory section.

@@
 Curiously, though, regularization need not reflect any plausible knowledge of the nature of covariance matrices in general or in the particular domain.
-Convergence can improve due to a purely statistical effect, sometimes known as \emph{Stein's phenomenon}, \emph{i.e.}~the favorable tradeoff between bias and variability, which arises \emph{anytime} a moderate bias toward a less variable target is mixed into an unbiased estimate.
+Convergence can improve due to a purely statistical effect, sometimes known as \emph{Stein's phenomenon}, \emph{i.e.}\;the favorable tradeoff between bias and variability, which arises \emph{anytime} a moderate bias toward a less variable target is mixed into an unbiased estimate.
 Therefore, an estimator that is shown to converge faster on specific types of data does not necessarily contain within itself some wisdom of the statistical regularities in the data.

### 2013-07-24 sections/Theory.tex
expanded the theory section.

--- 2013-07-16 sections/Theory.tex
@@
-\subsection*{Problem setup}
-Let $x_1,\ldots,x_n \in \mathbb R^{p\times 1}$ denote a sample of $n$ observations of instantaneous activity (e.g.~spiking) rates of $p$ neurons averaged in time bins $i=1,\ldots,n$.
+\subsection{Problem setup}
+Let $x_1,\ldots,x_n$ denote a sample of observations of instantaneous spiking rates of $p$ neurons in time bins $i=1,\ldots,n$.
+Each obseravation $x_i$ is a $p\times 1$ vector of spike counts.
+We assume that observations independent, identically distributed, i.e.\;all observations are drawn from a common true distribution $x_1,\ldots,x_n \sim f$.
+However, this assumptions can be relaxed and corrected for as discussed in this section.
+We make no assumption about the exact form of distribution $f$.
 
-For simplicity, we assume that observations are sampled from a shared true distribution $x_1,\ldots,x_n \sim f$. 
-\hl{\tiny explain: the derivations remain valid even without such a strong assumption.
-We account for overdispersion due to temporal dependence and shuffle chunks of data to account for nonstationarities.}
+Then the true  covariance matrix $\Sigma$ is defined as
+\begin{equation}
+\Sigma =\mathbb E_z[ (z-\mathbb E_{z\sim f} [z])(z-\mathbb E_z [z])^\T]
+\end{equation}
+where $\mathbb E_z[g(z)]$ denotes the expected value of $g(z)$ when $z$ is distributed according to $f(z)$.
 
-For simplicity, we assume that the data have zero means.
-We seek to estimate the true covariance matrix 
+The purpose of the covariance estimation procedure is to obtain an estimate $\hat \Sigma$ that is as close as possible to $\Sigma$ given the finite sample  $x_1,\ldots,x_n$.
+We define a scalar \emph{loss function} $\mathcal L\left(\hat \Sigma, \Sigma\right)$ to measure the distance between $\hat \Sigma$ and $\Sigma$.
+The loss function must have minimum at $\hat \Sigma=\Sigma$.
+
+In particular, we consider the sum-of-squared-error loss
 \begin{equation}
-\Sigma=\mathbb E_{z\sim f}[ zz^\T]
+\mathcal L_e(\hat \Sigma, \Sigma) =  \| \hat \Sigma - \Sigma\|^2_F 
+\end{equation}
+where $\| A \|_F = \sqrt{\Tr(AA^\T)}$ denotes the Frobenius matrix norm; $\Tr(\cdot)$ denotes the matrix trace.
+
+We also consider the normal-theory-log-likelihood loss
+\begin{equation}
+\mathcal L_g(\hat \Sigma,\Sigma) = \ln |\hat \Sigma|  + \Tr(\hat \Sigma^{-1}\Sigma)
+\end{equation}
+where $|A|$ denotes the determinant of matrix $A$.
+
+Note that both $\mathcal L_e(\hat\Sigma,\Sigma)$ and $\mathcal L_g(\hat\Sigma,\Sigma)$ attain their minima when $\hat\Sigma = \Sigma$.
+$\mathcal L_g$ is derived as the negative log likelihood function (scaled and shifted) when $f$ is a multivariate normal distribution.
+However, using either of these loss functions does not amount to assuming a specific form of $f$.
+Other loss functions have been defended in statistical literature \citep{James:1961,Fan:2008}.
+
+The loss function can be estimated based on another estimate $\hat\Sigma_Y$ of the covariance matrix obtained from an independent \emph{validation} sample $Y = (y_1,\ldots,y_m)\sim f^m$.
+
+Both $\mathcal L_e(\hat\Sigma,\Sigma)$ and $\mathcal L_g(\hat\Sigma,\Sigma)$ are linear with respect to $\Sigma$ in the sense that
+\begin{equation}
+\mathcal L\left(\hat \Sigma,\alpha S_1 + (1-\alpha)S_2\right) 
+\equiv 
+\alpha\mathcal L(\hat\Sigma,S_1) + (1-\alpha)\mathcal L(\hat\Sigma,S_2)
+\end{equation}
+If we ensure that  $\hat\Sigma_Y$ is unbiased, i.e.\;$\mathbb E_Y\left[\hat \Sigma_Y\right]=\Sigma$, then this linearity implies that 
+\begin{equation}
+\mathbb E_Y \left[\mathcal L(\hat\Sigma,\hat\Sigma_Y)\right]
+=
+\mathcal L\left(\hat\Sigma,\mathbb E_Y \left[\hat\Sigma_Y\right]\right)
+= 
+\mathcal L(\hat\Sigma,\Sigma)
 \end{equation}
 
-The purpose of the estimation procedure is to use the sample $x_1,\ldots,x_n$ to compute a covariance estimate $S$ that is as close as possible to $\Sigma$. 
-The quality of the covariance estimate $S$ will be defined using a \emph{loss function} $\mathcal L\left(S, \Sigma\right)$, which is minimized when $S=\Sigma$.
+In other words, the empirical loss $\mathcal L(\hat\Sigma,\hat\Sigma_Y)$ is an ubiased estimator of true loss $\mathcal L(\hat\Sigma,\Sigma)$ provided that $\hat\Sigma_Y$ is unbiased and $\hat\Sigma$ and $\hat\Sigma_Y$ are estimated from independent samples.
+Note that $\hat\Sigma$ does not need to be unbiased.
 
-In particular, we consider the mean-squared loss
-\begin{equation}
-\mathcal L_e( S, \Sigma) =  \| S - \Sigma\|^2_F 
-\end{equation}
-where $\| \cdot \|_F$ denotes the Frobenius norm;
-
-and the multivariate normal log likelihood loss
-\begin{equation}
-\mathcal L_g(S,\Sigma) = \ln |S| +  \Tr(S^{-1} \Sigma)
-\end{equation}
-where $|S|$ denotes the determinant of matrix $S$.
-
-\hl{\tiny list other possible loss functions and justify the choice of these two.
-Consider Stein's entropy loss and qudratic loss.
-James:1961 showed failures of the MSE loss, but it continued to be widely used.
-Fan:2008 provides other arguments.}
-
-The linearity of $\mathcal L_e(S,\Sigma)$ and $\mathcal L_g(S,\Sigma)$ with respect to the elements of $\Sigma$ allows unbiased empirical estimation of $\mathcal L$ from a validation sample $y_1,\ldots,y_m \sim f$ acquired independently of the original training sample $x_1,\ldots,x_n$ that was used to construct the estimate $S$.
-\begin{equation}
-\begin{split}
-\mathcal L(S,\Sigma) & \equiv \mathbb E_{z\sim f} \left[ \mathcal L(S,zz^\T) \right] \\ 
-&\approx \frac 1 m \sum\limits_{i=1}^m \mathcal L(S,y_i y_i^\T) = \mathcal L(S,\hat S_y)
-\end{split}
-\end{equation}
-where $\hat S_y = \frac 1 m \sum\limits_{i=1}^m x_i x_i^\T$ is the empirical covariance of the validation dataset.
-
-\hl{\tiny cross-validation}
-In practice, neural recordings are limited.
-Instead of validation with a dedicated validation dataset, $K$-fold cross-validation is used: The data are split into $K$ subsets of approximately equal size and each subset is used as the validation set with the covariance estimator trained on the remaning data.
-The estimation of the loss from the $K$ tests is then averaged, which is justified, again, thanks to the linearity property of the chosen loss functions.
-\hl{\tiny exlain how the data are split into continuous chunks so that dependences between the training and validation datasets due to temporal correlations are minimized.
-Coversely, each test set comprises multiple distirbuted chunks to reduce the effect of non-stationarities.}
+In practice, recording a separate independent sample may not be sensible.
+Instead, the recorded data a split into a training sample and validation sample.
+One popular approach is $K$-fold validation, in which the data are split into $K$ roughly equal-sized subsets.
+Each subset then successively serves as the validation  sample while the rest of the data make up the training sample.
+This procedure yields $k$ estimates of the loss, although not quite independent.
 

### 2013-07-25 sections/Introduction.tex
improved introduction, added figure captions.

--- 2013-07-24 sections/Introduction.tex
@@
-\hl{\tiny why correlations:} Linear correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \citep{Cohen:2011}.
+Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \citep{Cohen:2011}.
 For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been of particular interest.
 Noise correlations can have profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011}, and have been interpreted to indicate detailed and specific functional organization.
-Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Smith:2013,Denman:2013}.
+Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Ko:2011,Smith:2013,Denman:2013}.
 
-\hl{\tiny ambiguity of interpretation:} Neural correlations do not come with ready or unambiguous mechanistic interpretations.
+Neural correlations do not come with ready or unambiguous mechanistic interpretations.
 Theoretical and simulation studies show that neural correlations at various temporal scales may arise from any combination of underlying mechanisms.
-These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}
+These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
 
-\hl{\tiny opportunities with multineuronal data:} Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \citep{Shadlen:1998,Zohary:1994}.
-Multineuronal recordings allow estimation of covariance matrices of larger populations.
+Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \citep{Shadlen:1998,Zohary:1994}.
+Multineuronal recordings allow estimation of covariance matrices of large populations of neurons.
 Such estimates provide more information than the equivalent number of pairwise correlations assessed in isolation.
@@
 In contrast, the off-diagonal elements of the inverse of the correlation matrix constitute scaled partial correlations between neuron pairs, which reflect their specific linear dependencies, after accounting for the activity of all the other recorded cells; a strong interaction between a pair of neurons may be expressed by a single partial correlation but its effects may propagate to multiple correlations and eigenvalues.
+The inverse of the covariance matrix plays an important role in decoding schemes such as linear discriminant analysis, for example.
+The mutliple  representations of the covariance matrix with their alternative interpretations add both complications and opportunities into the search for fundamental regularities in neural population activity.
 
-\hl{\tiny motivate regularization} Estimation of parameters of the covariance matrix from data is inherently challenging, increasingly so for very high-dimensional data.
-As the amount of data grows only linearly with population size, the number of free parameters in a complete representation of the covariance matrix increases quadratically, multiplying opportunities for spurious patterns to emerge among the correlations.
-Although the estimation error of each covariance coefficient, in isolation, is unaffected by the dimensionality of the data, the estimation error of other parameters of the covariance matrix increases with population size.
-For example, the large eigenvalues of in the sample covariance matrix are biased upward while small eigenvalues are biased downward, with both variance and bias increasingly large with increasing data dimensionality \citep{Hayes:1981}.
-Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately in high-dimensional data.
+As the number of recorded cells increases, the usual estimations of parameters of the covariance matrix become increasingly noisy and biased.
+Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically.
+For example, the large eigenvalues from the sample covariance matrix are biased upward while small eigenvalues are biased downward \citep{Hayes:1981}.
+Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.
 
-\hl{\tiny forms of regularization} The convergence of the estimate of multivariate parameters to their true values can be improved through \emph{regularization},  Regularization is the deliberate biasing or \emph{shrinkage} of the empirical estimate toward a less variable \emph{target estimate} \citep{Schafer:2005} or the deliberate restriction of the estimate to a lower-dimensional space \citep{Dempster:1972,Fan:2008,Friedman:2008,Rothman:2008}.
-A common intuitive explanation for the effect of regularization is that regularization adds some prior knowledge about the likely forms of the covariance matrix.
+In this study, we pursue two goals: 
+\begin{enumerate}[A.\;\;]
+	\item Devise good estimators of neural covariance matrices in recordings from large populations of neurons.
+	\item Aid interpretation of population activity  by isolating the most essential features or structure of the neural covariance structure.
+\end{enumerate}
+
+These two goals are inextricably connected and are best addressed together.
+Interpretation of neural correlations can be aided by approximating the measured covariance matrix by a simplified, reduced form.
+For example, \cite{Malmersjo:2013} analyze the spatial distribution and the graph-theoretical measures of only the highest, most significant correlations.
+
+Estimation of parameters of the covariance matrix from data is inherently challenging, increasingly so for very high-dimensional data.
+The efficiency of an estimator relates to the rate of its convergence to the true value of the covariance matrix as the function of sample size.
+The usual estimator, the sample covariance matrix, has been shown to be substantially less efficient than a number of other estimators.
+The efficiency of an estimator can be improved through \emph{regularization}, i.e.\;the deliberate biasing of the estimate toward a more simplified approximate representation \citep{Bickel:2006,Bickel:2008,Ledoit:2004,Schafer:2005}.
+A common intuitive explanation for the effect of regularization is that of adding some prior knowledge about the likely forms of the covariance matrix.
 Curiously, though, regularization need not reflect any plausible knowledge of the nature of covariance matrices in general or in the particular domain.
@@
 
-\hl{\tiny approach and summary of findings} In this study, we compared the performance of several covariance estimators for spatially compact groups of 150--300 neurons in layers 2--4 in mouse primary visual cortex during visual stimulation.
+In this study, we compared the performance of several covariance estimators for spatially compact groups of 150--300 neurons in layers 2--4 in mouse primary visual cortex during visual stimulation.
 The performance of the estimators was evaluated by computing the mean squared error between the optimized covariance estimate fitted to training data and the non-regularized covariance matrix from a separate test data set.
@@
 
-\hl{\tiny summary of conlusions} As noted above, the optimal covariance estimator is domain-dependent. These specific findings may not spread to all neural circuits.
+As noted above, the optimal covariance estimator is domain-dependent. These specific findings may not spread to all neural circuits.
 Improved accuracy estimation accuracy is not the main aim or the only benefit.  

### 2013-07-26 sections/Abstract.tex
introduction improvements

@@
 Estimation of covariance matrices from finite recordings can be improved by regularization, \emph{i.e.}\;by imposing some kind of structure.
-Optimal regularization must be determined empirically since its performance depends on how well the imposed structure captures the underlying regularities.
+Optimal regularization must be determined empirically since its performance depends on how well the imposed structure matches the underlying regularities.
 For example, we can use low-rank parameterizations of the covariance matrix to account for common activity fluctuations across the recorded population.
@@
 
-To select the optimal structure of neural correlations in a local neural circuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, Ledoit-Wolf covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance.
+To select the optimal structure of neural correlations in a local neural circuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance.
 We inferred instantaneous firing rates in 200 ms bins from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.

### 2013-07-26 sections/Introduction.tex
introduction improvements

@@
 
-As the number of recorded cells increases, the usual estimations of parameters of the covariance matrix become increasingly noisy and biased.
+With large numbers of recorded cells, the usual estimations of the covariance matrix from finite recordings become ill-conditioned.
 Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically.
-For example, the large eigenvalues from the sample covariance matrix are biased upward while small eigenvalues are biased downward \citep{Hayes:1981}.
+For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \citep{Hayes:1981}.
 Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.
@@
 In this study, we pursue two goals: 
-\begin{enumerate}[A.\;\;]
-	\item Devise good estimators of neural covariance matrices in recordings from large populations of neurons.
-	\item Aid interpretation of population activity  by isolating the most essential features or structure of the neural covariance structure.
+\begin{enumerate}[\qquad A.\;\;]
+	\item Devise better estimators of neural covariance matrices in recordings from large populations of neurons.
+	\item Aid interpretation of population activity  by isolating the essential structure of the neural covariance matrix.
 \end{enumerate}
 
-These two goals are inextricably connected and are best addressed together.
-Interpretation of neural correlations can be aided by approximating the measured covariance matrix by a simplified, reduced form.
-For example, \cite{Malmersjo:2013} analyze the spatial distribution and the graph-theoretical measures of only the highest, most significant correlations.
+The two goals are inextricably connected as they both benefit from reduced, simplified approximations of the covariance matrix.
 
-Estimation of parameters of the covariance matrix from data is inherently challenging, increasingly so for very high-dimensional data.
-The efficiency of an estimator relates to the rate of its convergence to the true value of the covariance matrix as the function of sample size.
-The usual estimator, the sample covariance matrix, has been shown to be substantially less efficient than a number of other estimators.
-The efficiency of an estimator can be improved through \emph{regularization}, i.e.\;the deliberate biasing of the estimate toward a more simplified approximate representation \citep{Bickel:2006,Bickel:2008,Ledoit:2004,Schafer:2005}.
-A common intuitive explanation for the effect of regularization is that of adding some prior knowledge about the likely forms of the covariance matrix.
-Curiously, though, regularization need not reflect any plausible knowledge of the nature of covariance matrices in general or in the particular domain.
-Convergence can improve due to a purely statistical effect, sometimes known as \emph{Stein's phenomenon}, \emph{i.e.}~the favorable tradeoff between bias and variability, which arises \emph{anytime} a moderate bias toward a less variable target is mixed into an unbiased estimate.
-Therefore, an estimator that is shown to converge faster on specific types of data does not necessarily contain within itself some wisdom of the statistical regularities in the data.
-However, an estimator that does happen to accommodate the statistical structure of the data in a specific domain, it will gain additional advantage over an \emph{ad hoc} regularizer for that domain.
-Thus the optimal selection of the estimator for a specific problem with unknown statistical structure becomes an empirical question.
-And, if a clear winner emerges from a number of strong candidates, its regularization scheme can server as an effective statistical description of the covariance structure.
+In estimation theory, the overall error of an estimate can be reduced by \emph{regularization}, which is the deliberate biasing of the solution toward a simplified, low-dimensional (`sparse') \emph{target estimate} \citep{Schafer:2005,Bickel:2006}.
+Regularized estimators allow reaching a favorable tradeoff between bias and variance in order to minimize the total error.
+Strikingly, some improvement can be produced even when the target estimate is chosen arbitrarily, which is sometimes described as ``Stein's phenomenon'' after its discoverer \cite{Stein:1956}.
+However, when a sparse target estimate does happen to match the dominant features of the true covariance matrix, it will induce relatively little bias while retaining low variance.
+For example, some covariance estimators have been motivated by a specific model of the underlying process such as single-factor models of the stock market, for example \citep{Ledoit:2003}.
+Since the true covariance structure is not known in practice, the performance of estimators is evaluated by how well their performance generalized to new data, outside the training sample by cross validation, for example.
 
-In this study, we compared the performance of several covariance estimators for spatially compact groups of 150--300 neurons in layers 2--4 in mouse primary visual cortex during visual stimulation.
+The above consideration served as the basis of our approach.
+We proposed several alternative candidate estimators of the neural covriance matrix in a given circuit, each motivated by a different hypothesis for the origin of neural correlations.
+When a consistent winner emerges from cross-validated performance comparisons, the sparse estimate used in its regularization can be interpreted to reflect a simplified correlation structure supported by the emprical data.
+
+We compared the performance of several covariance estimators for spatially compact groups of 150--300 neurons in layers 2--4 in mouse primary visual cortex during visual stimulation.
 The performance of the estimators was evaluated by computing the mean squared error between the optimized covariance estimate fitted to training data and the non-regularized covariance matrix from a separate test data set.
@@
 Mixed sparse estimators with sparse partial correlations and low-rank components performed comparably.
-
-As noted above, the optimal covariance estimator is domain-dependent. These specific findings may not spread to all neural circuits.
-Improved accuracy estimation accuracy is not the main aim or the only benefit.  
-The selection of the optimal regularization scheme in itself serves as an effective descriptor of the dominant correlation structures.
-The finding that identifiable subsets of partial correlations were important to describe the partial correlation makes such pairs as prime candidates as direct functional interactions.

### 2013-07-28 sections/Introduction.tex
expanded the section on the theory of regularization.

@@
 
-In this study, we pursue two goals: 
+With this study, we pursue two goals: 
 \begin{enumerate}[\qquad A.\;\;]
@@
 
-The two goals are inextricably connected as they both benefit from reduced, simplified approximations of the covariance matrix.
-
-In estimation theory, the overall error of an estimate can be reduced by \emph{regularization}, which is the deliberate biasing of the solution toward a simplified, low-dimensional (`sparse') \emph{target estimate} \citep{Schafer:2005,Bickel:2006}.
+To select the best covariance matrix estimate for a specific neural circuit, we evaluate the performance of four estimators using different regularization schemes, each motivated by  a different hypothesis for the origin of neural correlations.
+\emph{Regularization} is the deliberate biasing the solution toward a simplified, low-dimensional (`sparse') \emph{target estimate} \citep{Schafer:2005,Bickel:2006}.
 Regularized estimators allow reaching a favorable tradeoff between bias and variance in order to minimize the total error.
@@
 
-The above consideration served as the basis of our approach.
-We proposed several alternative candidate estimators of the neural covriance matrix in a given circuit, each motivated by a different hypothesis for the origin of neural correlations.
-When a consistent winner emerges from cross-validated performance comparisons, the sparse estimate used in its regularization can be interpreted to reflect a simplified correlation structure supported by the emprical data.
-
 We compared the performance of several covariance estimators for spatially compact groups of 150--300 neurons in layers 2--4 in mouse primary visual cortex during visual stimulation.

### 2013-07-28 sections/Theory.tex
improved introduction, added figure captions.
expanded the section on the theory of regularization.

@@
-\subsection{Problem setup}
-Let $x_1,\ldots,x_n$ denote a sample of observations of instantaneous spiking rates of $p$ neurons in time bins $i=1,\ldots,n$.
-Each obseravation $x_i$ is a $p\times 1$ vector of spike counts.
-We assume that observations independent, identically distributed, i.e.\;all observations are drawn from a common true distribution $x_1,\ldots,x_n \sim f$.
-However, this assumptions can be relaxed and corrected for as discussed in this section.
-We make no assumption about the exact form of distribution $f$.
+\subsection{Covariance estimation}
 
-Then the true  covariance matrix $\Sigma$ is defined as
+\subsubsection{Problem setup}
+
+Let $x_i \in \mathcal X, i=1\ldots,x_n$ denote a sample of consequtive observations of population activity of $p$ neurons in time bins $i$.  
+The activity of each neuron is the real-valued firing rate, thus  $\mathcal X \in \mathbb R^{p\times 1}$.  
+We assume that observations are drawn from some true cumulative probability distribution $F: \mathcal X \mapsto R$.
+Then let $\hat F_n$ denote the empirical distribution function obtained from the observed sample, which itself is a random variable.
 \begin{equation}
-\Sigma =\mathbb E_z[ (z-\mathbb E_{z\sim f} [z])(z-\mathbb E_z [z])^\T]
+\hat F_n(x) = \sum\limits_{i=1}^n \mathbf{1}(x > x_i)
 \end{equation}
-where $\mathbb E_z[g(z)]$ denotes the expected value of $g(z)$ when $z$ is distributed according to $f(z)$.
+where $\mathbf 1(x_i \le x)$ is the indicator function which equals 1 when all elements of $x$ are greater than the corresponding elements of $x_i$ and 0 otherwise.
 
-The purpose of the covariance estimation procedure is to obtain an estimate $\hat \Sigma$ that is as close as possible to $\Sigma$ given the finite sample  $x_1,\ldots,x_n$.
-We define a scalar \emph{loss function} $\mathcal L\left(\hat \Sigma, \Sigma\right)$ to measure the distance between $\hat \Sigma$ and $\Sigma$.
-The loss function must have minimum at $\hat \Sigma=\Sigma$.
+The observations need not be independently and identically sampled from $F$.
+The only assumption is that the data generating process is ergodic, \emph{i.e.}\;that $\max_x \left[\hat F_n(x) - F(x)\right] \to 0$ as the sample size $n$ increases.
+Then each $x_i$ can be thought to be drawn from $F$ when not conditioned on nearby observations.
 
-In particular, we consider the sum-of-squared-error loss
+The true covariance matrix $\Sigma \in \Theta$ is defined as a function of the true probability distribution
 \begin{equation}
-\mathcal L_e(\hat \Sigma, \Sigma) =  \| \hat \Sigma - \Sigma\|^2_F 
+\Sigma(F) = \int\limits_{z\in\mathcal X} (z - \bar x)(z - \bar x)^\T \dif F(z)
+\quad\mbox{where}\;
+\bar x = \int\limits_{z\in\mathcal X} z \dif F(z)
 \end{equation}
-where $\| A \|_F = \sqrt{\Tr(AA^\T)}$ denotes the Frobenius matrix norm; $\Tr(\cdot)$ denotes the matrix trace.
+The \emph{plug-in estimator} is obtained by plugging the empirical distribution in place of the true distribution
+\begin{equation}
+\begin{split}
+\Sigma(\hat F_n) = \int\limits_{z\in\mathcal X} (z - \bar x)(z - \bar x)^\T \dif \hat F_n(z) 
+\equiv \frac  1 n\sum\limits_{i=1}^n (x_i-\bar x)(x_i - \bar x)
+\\
+\mbox{where}\;\bar x = 
+\int\limits_{z\in \mathcal X}z \dif \hat F_n(z) \equiv \frac 1 n\sum\limits_{i=1}^n x_i
+\end{split}
+\end{equation}
+The plugin estimator is known to be biased toward zero because the estimate of the mean $\bar x$ absorbs a fraction of the variance.
 
-We also consider the normal-theory-log-likelihood loss
+The usual unbiased estimate, the sample covariance matrix, rescales the plug-in estimator compensates for the bias
 \begin{equation}
-\mathcal L_g(\hat \Sigma,\Sigma) = \ln |\hat \Sigma|  + \Tr(\hat \Sigma^{-1}\Sigma)
+\hat\Sigma(\hat F_n) = \frac n {n - d} \Sigma(\hat F_n) \equiv \frac 1 {n-d} \sum\limits_{i=1}^n (x_i - \bar x) (x_i - \bar x)^\T
 \end{equation}
-where $|A|$ denotes the determinant of matrix $A$.
+When observations are independent, then $d=1$.
+But when nearby samples are correlated, covariances become underestimated.
+Then we can use $d=cn$ where $c>1$ is called the \emph{overdispersion coefficient}, which may be estimated from data.
 
-Note that both $\mathcal L_e(\hat\Sigma,\Sigma)$ and $\mathcal L_g(\hat\Sigma,\Sigma)$ attain their minima when $\hat\Sigma = \Sigma$.
-$\mathcal L_g$ is derived as the negative log likelihood function (scaled and shifted) when $f$ is a multivariate normal distribution.
-However, using either of these loss functions does not amount to assuming a specific form of $f$.
-Other loss functions have been defended in statistical literature \citep{James:1961,Fan:2008}.
+\subsubsection{Definition of regularization}
+Now let's define a sequence of approximate biased estimates $\hat\Sigma_\gamma(\hat F_n)$ indexed by the \emph{regularization intensity}  $\gamma$ such that $\hat\Sigma_{\gamma=0}(\hat F_n)\equiv \hat\Sigma(\hat F_n)$, which become increasingly biased as $\gamma$ increases.
 
-The loss function can be estimated based on another estimate $\hat\Sigma_Y$ of the covariance matrix obtained from an independent \emph{validation} sample $Y = (y_1,\ldots,y_m)\sim f^m$.
+To evaluate the estimate, let's define the \emph{loss function} $\mathcal L : \Theta \times \Theta \mapsto \mathbb R$, which is minimized when its two arguments are equal.
+Then \emph{excess loss} $\ell(\hat S, S) = \mathcal L(\hat S, S) - \mathcal L(S,S)$ assumes zero at the minimum.
+
+The overall error, known as \emph{estimator risk} is expressed as
+\begin{equation}
+r_\gamma = \mathbb E_F\left[ \ell\left(\Sigma_\gamma(P_n),\Sigma\right) \right]
+\end{equation}
+
+The expected value of the approximate estimate 
+$\Sigma_\gamma^{(n)} = \mathbb E_F\left[\hat \Sigma_\gamma(\hat F_n)\right]$ 
+is biased with the amout of bias or \emph{approximaton error} expressed as  
+\begin{equation}
+b_\gamma^2 = \ell \left( \Sigma_\gamma^{(n)},\Sigma\right)
+\end{equation}
+
+Then the \emph{estimation error} 
+\begin{equation}
+\varepsilon_\gamma^2 = \mathbb E_F \left[ \ell\left(\hat \Sigma_\gamma(\hat F_n), \Sigma_\gamma^{(n)}\right) \right]
+\end{equation}
+
+When the loss function is the sum of squared error (Euclidean norm squared), which, for matricies, is called the Frobenius norm, $\mathcal L(\hat S,S) = \mathcal L_e(\hat S,S) = \|\hat S-S\|_F^2 = \Tr\left((\hat S-S)(\hat S-S)^\T\right)$, the estimation error is called \emph{variance} and the overall error is expressed as the the sum
+\begin{equation}
+r_\gamma =  b_\gamma^2 + \varepsilon_\gamma^2
+\end{equation}
+
+With other loss functions, the overall error is also an increasing function of both kinds of errors, but not necessarily a simple sum.
+
+One other popular loss function for covariance estimation arises from the theory of normal distributions.
+Under the assumption of normality  with zero mean, the log likelihood of the covariance matrix $\Sigma$ given the independent sample $z_1,\ldots,z_m$, with $S = \tfrac 1 m \sum\limits_{i=1}^m z_i z_i^\T$,
+\begin{equation}
+L\left(\Sigma \mid S\right) = -\tfrac m 2 \ln(2\pi) - \tfrac m 2 \ln \det \Sigma - \tfrac m 2 \Tr(\Sigma^{-1} S)
+\end{equation}
+Then the Gaussian loss function $\mathcal L_g$ is constructed by rescaling the normal log likelihood and dropping the constant term
+\begin{equation}
+\begin{split}
+\mathcal L_g(\hat S,S) = &-\frac 2 m L\left(S \mid \hat S \right) - \ln(2\pi) 
+\\ 
+\equiv & \ln \det \hat S + \Tr(\hat S^{-1} S) 
+\end{split}
+\end{equation}
+The corresponding excess loss 
+\begin{equation}
+\ell_g(\hat S,S) = \mathcal L_g(\hat S,S) - \mathcal L_g(S,S) 
+= -\ln \det \left(\hat S^{-1} S\right) + \Tr\left(\hat S^{-1}S\right) - p
+\end{equation}
+is known as \emph{entropy loss} \citep{James:1961}.
+
+Despite the fact that entropy loss is inspired by the normal theory, the choice of a loss function is not equivalent to assuming a specific form of $F$.
+The loss function only measures the distance between the estimate and the true value of a parameter of the true distribution but not the distance between distributions themselves.
+Minimizing the loss function improves the estimate under any true distribution.
+
+The task of regulazation comprises the choice of the \emph{regularization path} $\hat \Sigma_\gamma(\hat F_n)$ and the optimal location $\hat\gamma = \gamma(\hat F_n)$ on that path in oder to minimize the risk $r_\gamma$.
+By definition $\hat\Sigma_{\gamma=0}(\hat F_n)$ is an unbiased estimator with $b_{\gamma=0}^2 = 0$.
 
@@
 \end{equation}
-
 In other words, the empirical loss $\mathcal L(\hat\Sigma,\hat\Sigma_Y)$ is an ubiased estimator of true loss $\mathcal L(\hat\Sigma,\Sigma)$ provided that $\hat\Sigma_Y$ is unbiased and $\hat\Sigma$ and $\hat\Sigma_Y$ are estimated from independent samples.
@@
 Instead, the recorded data a split into a training sample and validation sample.
-One popular approach is $K$-fold validation, in which the data are split into $K$ roughly equal-sized subsets.
+One popular approach is $K$-fold \emph{cross-validation}, in which the data are split into $K$ roughly equal-sized subsets.
 Each subset then successively serves as the validation  sample while the rest of the data make up the training sample.

### 2013-07-29 sections/Abstract.tex
Expanded Covariance Estimation Theory section.

@@
-Correlations between the spiking activity of pairs of neurons are among the most familiar and useful descriptive statistics of neural activity.
+Correlations between the spiking activity of pairs of neurons are among the most familiar descriptive statistics of neural activity.
 Multineuronal recordings provide richer information than the equivalent number of neuron pairs.
 For example, the full covariance matrix reveals correlated activity across the entire population as well as partial correlations between pairs.
-Estimation of covariance matrices from finite recordings can be improved by regularization, \emph{i.e.}\;by imposing some kind of structure.
-Optimal regularization must be determined empirically since its performance depends on how well the imposed structure matches the underlying regularities.
-For example, we can use low-rank parameterizations of the covariance matrix to account for common activity fluctuations across the recorded population.
-Conversely, if correlations are strongly influenced by a small fraction of possible pairwise linear associations between the observed neurons, we can impose sparsity on the inverse covariance matrix.
+Estimation of covariance matrices can be improved by regularization, \emph{i.e.}\;by imposing some kind of structure.
+Optimal regularization must be determined empirically since its effect depends on how closely the imposed structure matches the underlying regularities.
+For example, we can use low-rank parameterizations of the covariance matrix to account for common fluctuations across the recorded population.
+Conversely, if correlations are strongly influenced by a small fraction of pairwise linear associations between the observed neurons, we can impose sparsity on the inverse covariance matrix.
 We can also use a `sparse+low rank’ inverse covariance representation to account for both common fluctuations and pairwise interactions.

### 2013-07-29 sections/Acquisition.tex
Expanded Covariance Estimation Theory section.

\subsection{Data acquisition}
Describe the two-photon experiment, refer to Fig.\;1

### 2013-07-29 sections/Theory.tex
Expanded Covariance Estimation Theory section.

@@
 Let $x_i \in \mathcal X, i=1\ldots,x_n$ denote a sample of consequtive observations of population activity of $p$ neurons in time bins $i$.  
-The activity of each neuron is the real-valued firing rate, thus  $\mathcal X \in \mathbb R^{p\times 1}$.  
+The activity of each neuron is the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
 We assume that observations are drawn from some true cumulative probability distribution $F: \mathcal X \mapsto R$.
@@
 \begin{equation}
-\hat F_n(x) = \sum\limits_{i=1}^n \mathbf{1}(x > x_i)
+\hat F_n(x) = \frac 1 n \sum\limits_{i=1}^n \mathbf{1}(x \ge x_i)
 \end{equation}
-where $\mathbf 1(x_i \le x)$ is the indicator function which equals 1 when all elements of $x$ are greater than the corresponding elements of $x_i$ and 0 otherwise.
+where $\mathbf 1(x \ge x_i)$ is the indicator function which equals 1 when all elements of $x$ are greater than the corresponding elements of $x_i$ and 0 otherwise.
 
@@
 \begin{equation}
-\Sigma(F) = \int\limits_{z\in\mathcal X} (z - \bar x)(z - \bar x)^\T \dif F(z)
+\Sigma = \int\limits_{z\in\mathcal X} (z - \bar x)(z - \bar x)^\T \dif F(z)
 \quad\mbox{where}\;
@@
 \end{equation}
-The \emph{plug-in estimator} is obtained by plugging the empirical distribution in place of the true distribution
+The \emph{plug-in estimator} $\hat\Sigma$ is obtained by plugging the empirical distribution $F_n$ in place of the true distribution $F$
 \begin{equation}
 \begin{split}
-\Sigma(\hat F_n) = \int\limits_{z\in\mathcal X} (z - \bar x)(z - \bar x)^\T \dif \hat F_n(z) 
+\hat\Sigma = \int\limits_{z\in\mathcal X} (z - \bar x)(z - \bar x)^\T \dif \hat F_n(z) 
 \equiv \frac  1 n\sum\limits_{i=1}^n (x_i-\bar x)(x_i - \bar x)
@@
 \end{equation}
-The plugin estimator is known to be biased toward zero because the estimate of the mean $\bar x$ absorbs a fraction of the variance.
+The plug-in estimator is known to be biased toward zero because the estimate of the mean $\bar x$ absorbs a fraction of the variances.
 
-The usual unbiased estimate, the sample covariance matrix, rescales the plug-in estimator compensates for the bias
+The usual unbiased estimate, the sample covariance matrix, rescales the plug-in estimator to compensate for the bias
 \begin{equation}
-\hat\Sigma(\hat F_n) = \frac n {n - d} \Sigma(\hat F_n) \equiv \frac 1 {n-d} \sum\limits_{i=1}^n (x_i - \bar x) (x_i - \bar x)^\T
+\hat\Sigma_0 = \frac n {n - d} \hat\Sigma \equiv \frac 1 {n-d} \sum\limits_{i=1}^n (x_i - \bar x) (x_i - \bar x)^\T
 \end{equation}
 When observations are independent, then $d=1$.
-But when nearby samples are correlated, covariances become underestimated.
-Then we can use $d=cn$ where $c>1$ is called the \emph{overdispersion coefficient}, which may be estimated from data.
+But when nearby samples are correlated, the estimate of the mean absorbs more of the variance, becomes \emph{overdispersed}.
+As a result the covariances become underestimated.
+Then we can use $d=cn$ where $c>1$ is called the \emph{overdispersion coefficient}, which may be estimated from the data.
 
-\subsubsection{Definition of regularization}
-Now let's define a sequence of approximate biased estimates $\hat\Sigma_\gamma(\hat F_n)$ indexed by the \emph{regularization intensity}  $\gamma$ such that $\hat\Sigma_{\gamma=0}(\hat F_n)\equiv \hat\Sigma(\hat F_n)$, which become increasingly biased as $\gamma$ increases.
-
-To evaluate the estimate, let's define the \emph{loss function} $\mathcal L : \Theta \times \Theta \mapsto \mathbb R$, which is minimized when its two arguments are equal.
+\subsubsection{Bias-variance tradeoff}
+Let $\hat\Sigma_\gamma$ be a family of approximate (biased) covariance estimates indexed by $\gamma$.
+To evaluate each estimate, let's define the \emph{loss function} $\mathcal L : \Theta \times \Theta \mapsto \mathbb R$ such that $\mathcal L(\hat S,S)$ is minimized when $\hat S = S$.
 Then \emph{excess loss} $\ell(\hat S, S) = \mathcal L(\hat S, S) - \mathcal L(S,S)$ assumes zero at the minimum.
 
-The overall error, known as \emph{estimator risk} is expressed as
+The overall error or \emph{estimator risk} 
 \begin{equation}
-r_\gamma = \mathbb E_F\left[ \ell\left(\Sigma_\gamma(P_n),\Sigma\right) \right]
+r_\gamma = \mathbb E\left[ \ell\left(\hat\Sigma_\gamma,\Sigma\right) \right]
+\end{equation}
+can be decomposed into \emph{approximation error} or \emph{bias}   
+\begin{equation}
+b_\gamma^2 = \ell \left( \mathbb E\left[\hat \Sigma_\gamma\right],\Sigma\right)
+\end{equation}
+and \emph{estimation error} or \emph{variance}
+\begin{equation}
+\varepsilon_\gamma^2 = \mathbb E \left[ \ell\left(\hat \Sigma_\gamma, 
+\mathbb E\left[\hat \Sigma_\gamma\right]\right) \right]
 \end{equation}
 
-The expected value of the approximate estimate 
-$\Sigma_\gamma^{(n)} = \mathbb E_F\left[\hat \Sigma_\gamma(\hat F_n)\right]$ 
-is biased with the amout of bias or \emph{approximaton error} expressed as  
+When the loss function is the error sum of squares, which, for matrices, is called the Frobenius norm, 
 \begin{equation}
-b_\gamma^2 = \ell \left( \Sigma_\gamma^{(n)},\Sigma\right)
+\mathcal L_e(\hat S,S) = \|\hat S-S\|_F^2 = \Tr\left((\hat S-S)(\hat S-S)^\T\right)
 \end{equation}
-
-Then the \emph{estimation error} 
-\begin{equation}
-\varepsilon_\gamma^2 = \mathbb E_F \left[ \ell\left(\hat \Sigma_\gamma(\hat F_n), \Sigma_\gamma^{(n)}\right) \right]
-\end{equation}
-
-When the loss function is the sum of squared error (Euclidean norm squared), which, for matricies, is called the Frobenius norm, $\mathcal L(\hat S,S) = \mathcal L_e(\hat S,S) = \|\hat S-S\|_F^2 = \Tr\left((\hat S-S)(\hat S-S)^\T\right)$, the estimation error is called \emph{variance} and the overall error is expressed as the the sum
+the bias-variance decomposition is the simple sum
 \begin{equation}
@@
 
-With other loss functions, the overall error is also an increasing function of both kinds of errors, but not necessarily a simple sum.
+With other loss functions, the risk is also an increasing function of the two components, although not generally the sum.
 
-One other popular loss function for covariance estimation arises from the theory of normal distributions.
+Regularization is the game of minimizing risk $r_\gamma$ by reducing estimation error $\varepsilon_\gamma$ at the cost of increased bias $b_\gamma$, the \emph{bias-variance tradeoff}.  
+We follow the convention that $\hat\Sigma_{\gamma=0}$ is the unbiased estimator $\hat\Sigma_0$, \emph{i.e.}\;$b_{\gamma=0}^2 = 0$.
+Then as $\gamma$ increases, $\hat\Sigma_\gamma$ less variable but increasingly biased.
+Thus $\hat\Sigma_\gamma$ defines the \emph{regularization path}, for which \emph{regularization intensity} $\gamma$ specifies the location along the path, which may be real-valued or discrete, or even a vector to steer the bias toward multiple targets.
+
+\subsubsection{Evaluation of estimators}
+Another popular loss  function for covariance estimation arises from the theory of normal distributions.
 Under the assumption of normality  with zero mean, the log likelihood of the covariance matrix $\Sigma$ given the independent sample $z_1,\ldots,z_m$, with $S = \tfrac 1 m \sum\limits_{i=1}^m z_i z_i^\T$,
@@
 
-Despite the fact that entropy loss is inspired by the normal theory, the choice of a loss function is not equivalent to assuming a specific form of $F$.
+Despite the fact that entropy loss is derived from normal theory, the choice of a loss function is not equivalent to assuming a specific form of $F$.
 The loss function only measures the distance between the estimate and the true value of a parameter of the true distribution but not the distance between distributions themselves.
@@
 
-The task of regulazation comprises the choice of the \emph{regularization path} $\hat \Sigma_\gamma(\hat F_n)$ and the optimal location $\hat\gamma = \gamma(\hat F_n)$ on that path in oder to minimize the risk $r_\gamma$.
-By definition $\hat\Sigma_{\gamma=0}(\hat F_n)$ is an unbiased estimator with $b_{\gamma=0}^2 = 0$.
+In practice, the true covariance matrix $\Sigma$ is not available and the estimator risk must be estimated from data.
+This requires a separate \emph{validation} empirical distribution $\hat F_m^\prime$ of $m$ observations sampled from $F$ independently of the \emph{training} distribution $\hat F_n$.
+The we can produce an independent unbiased covariance estimate $\hat \Sigma_0(\hat F_m^\prime)$.
 
-Both $\mathcal L_e(\hat\Sigma,\Sigma)$ and $\mathcal L_g(\hat\Sigma,\Sigma)$ are linear with respect to $\Sigma$ in the sense that
+Then the \emph{empirical risk} is 
 \begin{equation}
-\mathcal L\left(\hat \Sigma,\alpha S_1 + (1-\alpha)S_2\right) 
+\hat r_\gamma = \mathcal L\left(\hat\Sigma_\gamma,\hat\Sigma_0(\hat F_m^\prime)\right) 
+\end{equation}
+
+ The two loss functions $\mathcal L_e(\hat S,S)$ and $\mathcal L_g(\hat S,S)$ are particularly suitable for empirical risk estimation thanks to their linearity with respect to $S$ in the sense that 
+\begin{equation}
+\mathcal L\left(\hat S,\alpha S_1 + (1-\alpha)S_2\right) 
 \equiv 
-\alpha\mathcal L(\hat\Sigma,S_1) + (1-\alpha)\mathcal L(\hat\Sigma,S_2)
+\alpha\mathcal L(\hat\Sigma,S_1) + (1-\alpha)\mathcal L(\hat S,S_2)
 \end{equation}
-If we ensure that  $\hat\Sigma_Y$ is unbiased, i.e.\;$\mathbb E_Y\left[\hat \Sigma_Y\right]=\Sigma$, then this linearity implies that 
+which allows bringing the expectation inside 
 \begin{equation}
-\mathbb E_Y \left[\mathcal L(\hat\Sigma,\hat\Sigma_Y)\right]
+\mathbb E\left[ \mathcal L\left(\hat\Sigma_r, \hat\Sigma_0(\hat F_m^\prime)\right) \right] 
 =
-\mathcal L\left(\hat\Sigma,\mathbb E_Y \left[\hat\Sigma_Y\right]\right)
+\mathbb E\left[ \mathcal L\left(\hat\Sigma_\gamma, \mathbb E\left[\Sigma_0(\hat F_m^\prime)\right]\right) \right] 
 =
-\mathcal L(\hat\Sigma,\Sigma)
+\mathbb E\left[ \mathcal L\left(\hat\Sigma_\gamma, \Sigma\right) \right] 
 \end{equation}
-In other words, the empirical loss $\mathcal L(\hat\Sigma,\hat\Sigma_Y)$ is an ubiased estimator of true loss $\mathcal L(\hat\Sigma,\Sigma)$ provided that $\hat\Sigma_Y$ is unbiased and $\hat\Sigma$ and $\hat\Sigma_Y$ are estimated from independent samples.
-Note that $\hat\Sigma$ does not need to be unbiased.
+Then the expectation of the empirical risk is 
+\begin{equation}
+\mathbb E\left[\hat r_\gamma\right] 
+= \mathbb E\left[\mathcal L(\hat\Sigma_\gamma,\Sigma)\right]
+= r_\gamma + \mathcal L(\Sigma,\Sigma)
+\end{equation}
+This means, that our goal of minimizing the true loss $r_\gamma$ can be replaced with minimizing  the empirical loss $\hat r_\gamma$.
 
-In practice, recording a separate independent sample may not be sensible.
+In practice, recording a separate independent validation sample may not be sensible.
 Instead, the recorded data a split into a training sample and validation sample.
-One popular approach is $K$-fold \emph{cross-validation}, in which the data are split into $K$ roughly equal-sized subsets.
-Each subset then successively serves as the validation  sample while the rest of the data make up the training sample.
-This procedure yields $k$ estimates of the loss, although not quite independent.
+In $K$-fold \emph{cross-validation} the data are split into $K$ roughly equal-sized subsets.
+Each subset then successively serves to construct the validation distribution $\hat F_m^\prime$  while the rest of the data is used to construct the training distribution $\hat F_n$.
+This procedure produces $K$ estimates of the empirical risk, which are then averaged together to produce a better, averaged, estimate of the true risk.
 
-\subsection*{Covariance estimators}
-It has long been recognized that the empirical covariance 
+Different regularization schemes can then be compared by comparing cross-validated empirical losses.
+
+\subsubsection{Constructing a regularized estimator}
+The task of designing a regularized estimators consists of defining a sequence of approximations $\hat\Sigma_\gamma$, called \emph{regularization path} and the optimal location on that path, \emph{regularization intensity} $\hat\gamma$ as  functions of the empirical distribution.
+
+There are two basic approaches to design a family of approximations of a parameter that reduce the estimation error: \emph{dimensionality reduction} and \emph{shrinkage}.
+
+Dimensionality reduction projects the estimate onto a low-dimensional manifold (subspace) within the $p(p+1)/2$-dimensional space of covariance matrices $\Theta$.
+The variance in the projection is likely to be lower (if the manifold is smooth).
+For example, factor analysis with $\nu$ factors projects (w.r.t.\;$\mathcal L_g$) $\hat\Sigma_0$ onto the low-dimensional manifold of covariance matrices representable by factor models.
+Thus factor analysis can be used as a regularized covariance estimator \citep{Fan:2008} with $\gamma$ controlling the number of factors (or, alternatively, the nuclear norm).
+Another dimensionality reduction technique is \emph{covariance selection} \citep{Dempster:1972}, wherein a fraction of inverse covariance coefficients are constrained to zero and the closest (w.r.t.\;$\mathcal L_g$) solution to $\hat\Sigma_0$ is found.
+
+Covariance shrinkage \citep{Schafer:2005,Ledoit:2004} uses a fixed, low-dimensional \emph{target estimate} $T$ and constructs the estimator as the linear mixture, with $\gamma \in [0,1]$,
 \begin{equation}
-\hat S = \frac 1 n \sum\limits_{i=1}^n x_i x_i^\T
+\hat\Sigma_\gamma = \hat\Sigma_0  + \gamma (T-\hat\Sigma_0) 
 \end{equation}
-is statistically inadmissible, meaning that larger sample sizes are required to reduce the loss function to the same level as other, more statistically efficient covariance estimators.
+ The target can be a diagonal matrix with sample variances, an identity matrix scaled to the mean variance, a single factor model, or any other low-dimensional estimate.
 
-Covariance estimators can be made more efficient by two broad approaches, which are not mutually exclusive: (1) deliberately bias the estimator in order to reduce its variability and (b) reduce the dimensionality of the estimator by choosing a specific parametric model.
+An effective hybrid of dimensionality reduction and shrinkage can be obtained by using $L_1$-regularization techniques.
+For example, graphical lasso \citep{Meinshausen:2006,Friedman:2008}, shrinks the inverse of the covariance toward zero as
+\begin{equation}
+\hat\Sigma_\gamma = \arg\min\limits_S \mathcal L_g(S,\hat \Sigma_0) + \gamma \| S^{-1} \|_1
+\end{equation}
+This approach results in forcing many of the coefficients in the inverse of the estimate to zero, or ``sparsification'' of the the inverse.
 
-\subsubsection*{Shrinkage estimators}
-The most familiar example of the first approach is the linear shrinkage estimator, calculated is a linear mixture of the empirical covariance estimate $\hat S$ and another low-variance target estimator $T$:
-\begin{equation}
-S_{shrink} = (1-\lambda) \hat S + \lambda T
-\end{equation}
-A popular target covariance is the identity matrix scaled by the mean sample variance  $T = \frac 1 p \Tr(\hat S) I$.
-The optimal value of shrinkage estimate with respect to $\mathcal L_e$ can be estimated analytically from the data \citep{Ledoit:2004,Schafer:2005}.
+Many regularization schemes combine multiple targets or multiple approaches, in which case $\gamma$ can become a vector of regularization parameters \citep{Schafer:2005,Fan:2011,Ma:2013} to be optimized in parallel.
 
-\subsubsection*{Dimensionality reduction}
-The second broad approach is to impose a structure on the covariance estimate by reducing the number of free parameters in one of its parameterizations.
+\subsubsection{Optimization of the regularization parameter}
+Analytic solutions when exist \citep{Ledoit:2004,Schafer:2005}, nested cross-validation otherwise.
+
+\subsubsection{Interpretation of revealed structure}
+Sparseness (dimensionality reduction) are key, but must be supported by data, not emph{ad hoc}. \citep{Fan:2006,Malmersjo:2013}.
+
+\subsubsection{Out choices of regularized covariance estimates}
+\begin{enumerate}[\qquad 1.\;\;]
+\item sample covariance (benchmark)
+\item covariance shrinkage 
+\item factor model
+\item graphical lasso
+\item latent-variable graphical lasso
+\end{enumerate}

### 2013-07-29 sections/Results.tex
Added figure captions.

\textbf{Figure 3.}{ 
The fitting performance of regularized covariance estimators on two-photon recording. 
}\label{fig:03}

\textbf{Figure 4.}{ 
Fitting performance of regularized covariance estimators on 24 datasets two-photon recordinngs.
}\label{fig:04}

\textbf{Figure 5.}{
The depiction of the 3D arrangement of the non-zero partial correlations between pairs of neurons.
}\label{fig:04}

### 2013-07-29 sections/Simulation.tex
Added figure captions.

\subsection{Simulation}
To illustrate,

\textbf{Figure 2.}{
Simulation results. 
\textbf{A}: A $100 \times 100$ covariance matrix without a low-dimensional structure, a sample covariance with n=2000, distributions of true and sample correlation coefficients, eigenspectra of true and sample covariances.
\textbf{B}: A similar covariance matrix with 8 factors.
\textbf{C}: A similar covariance matrix with sparse inverse. 
\textbf{D}: A similar covariance matrix with sparse + low-rank inverse with 3 latent units.
\textbf{E}: Risk convergence rates
}\label{fig:02}

### 2013-07-29 sections/Theory.tex
Added figure captions.

@@
 
-\subsubsection{Problem setup}
+\paragraph{Problem setup:}
 
@@
 
-\subsubsection{Bias-variance tradeoff}
+\paragraph{Variance-bias tradeoff:}
 Let $\hat\Sigma_\gamma$ be a family of approximate (biased) covariance estimates indexed by $\gamma$.
@@
 
-With other loss functions, the risk is also an increasing function of the two components, although not generally the sum.
+With other loss functions, the risk is also an increasing function of the two components, although is not generally the sum.
 
@@
 
-\subsubsection{Evaluation of estimators}
+\paragraph{Evaluation of estimators:}
 Another popular loss  function for covariance estimation arises from the theory of normal distributions.
@@
 Despite the fact that entropy loss is derived from normal theory, the choice of a loss function is not equivalent to assuming a specific form of $F$.
-The loss function only measures the distance between the estimate and the true value of a parameter of the true distribution but not the distance between distributions themselves.
-Minimizing the loss function improves the estimate under any true distribution.
+The loss function measures the discrepancy between distribution parameters rather than the distance between distributions.
+\TODO{explain Bregman divergence?}
 
@@
 \equiv 
-\alpha\mathcal L(\hat\Sigma,S_1) + (1-\alpha)\mathcal L(\hat S,S_2)
+\alpha\mathcal L(\hat S,S_1) + (1-\alpha)\mathcal L(\hat S,S_2)
 \end{equation}
@@
 
-\subsubsection{Constructing a regularized estimator}
+\paragraph{Constructing a regularized estimator:}
 The task of designing a regularized estimators consists of defining a sequence of approximations $\hat\Sigma_\gamma$, called \emph{regularization path} and the optimal location on that path, \emph{regularization intensity} $\hat\gamma$ as  functions of the empirical distribution.
@@
 
-\subsubsection{Optimization of the regularization parameter}
+\paragraph{Optimization of the regularization parameter:}
 Analytic solutions when exist \citep{Ledoit:2004,Schafer:2005}, nested cross-validation otherwise.
 
-\subsubsection{Interpretation of revealed structure}
+\paragraph{Interpretation of revealed structure:}
 Sparseness (dimensionality reduction) are key, but must be supported by data, not emph{ad hoc}. \citep{Fan:2006,Malmersjo:2013}.
 
-\subsubsection{Out choices of regularized covariance estimates}
+\paragraph{Out choices of regularized covariance estimates:}
 \begin{enumerate}[\qquad 1.\;\;]
 \item sample covariance (benchmark)
-\item covariance shrinkage 
-\item factor model
+\item linear shrinkage toward identity 
+\item linear shrinkage toward a single-factor model.
+\item multifactor
 \item graphical lasso

### 2013-07-29 sections/Abstract.tex
Added figure captions.
minor changes

@@
-Correlations between the spiking activity of pairs of neurons are among the most familiar descriptive statistics of neural activity.
+Correlations \TODO{we assume readers understand that Pearson linear correlations are meant here} between the spiking activity of pairs of neurons are among the most familiar descriptive statistics of neural activity.
 Multineuronal recordings provide richer information than the equivalent number of neuron pairs.
-For example, the full covariance matrix reveals correlated activity across the entire population as well as partial correlations between pairs.
+For example, the full covariance \TODO{we assume readers understand the relationship between covariance and correlation} matrix reveals correlated activity across the entire population as well as partial correlations between pairs.
 Estimation of covariance matrices can be improved by regularization, \emph{i.e.}\;by imposing some kind of structure.
@@
 Conversely, if correlations are strongly influenced by a small fraction of pairwise linear associations between the observed neurons, we can impose sparsity on the inverse covariance matrix.
-We can also use a `sparse+low rank’ inverse covariance representation to account for both common fluctuations and pairwise interactions.
+We can also use a `sparse+low rank' inverse covariance representation to account for both common fluctuations and pairwise interactions.
 

### 2013-07-29 sections/Acquisition.tex
Added figure captions.
minor changes

@@
 \subsection{Data acquisition}
-Describe the two-photon experiment, refer to Fig.\;1
+Describe the two-photon experiment, refer to Fig.\;1.
+
+\textbf{Figure 1.}{
+Experimental setup and data acquisition. 
+\textbf{A}: Mouse in two-photon microscope, visual stimulation, 3D scanning of cells.
+\textbf{B}: Calcium signals
+\textbf{C}: Deconvolution and binning.
+}\label{fig:01}

### 2013-07-29 sections/Theory.tex
minor changes
minor comments

@@
 Regularization is the game of minimizing risk $r_\gamma$ by reducing estimation error $\varepsilon_\gamma$ at the cost of increased bias $b_\gamma$, the \emph{bias-variance tradeoff}.
+\TODO{This point can be untuitive.
+For example, \cite{Varoquaux:2012} misrepresents the effect of regularization by stating that regularization removes the upward bias of correlations in the sample covariance matrix.
+In reality, regularization should start with an unbiased estimator and bias it toward a low-variance target so that the estimation risk is reduced.
+Another important point is ``Stein's paradox'': regularization does not improve the estimate of each covariance.
+In fact, it makes it worse.
+Regularization only improves the risk when measured on the entire matrix.
+Regularization should not be used to improve the estimate of average correlations or the distribution of correlations.}
 We follow the convention that $\hat\Sigma_{\gamma=0}$ is the unbiased estimator $\hat\Sigma_0$, \emph{i.e.}\;$b_{\gamma=0}^2 = 0$.
@@
 
-\paragraph{Out choices of regularized covariance estimates:}
+\paragraph{Our choices of regularized covariance estimates:}
 \begin{enumerate}[\qquad 1.\;\;]

### 2013-07-30 sections/Abstract_KJ.tex
Kreso's review comments.

Linear correlations between the spiking activity of pairs of neurons are among the most familiar and useful descriptive statistics of neural activity.
Correlation matrices from multineuronal recordings contain richer information than available from an equivalent number of isolated pairs:
The basic structure of population activity can manifest in various low-dimensional representations of the covariance matrix.  \Kcomment{The previous may not be clear to someone who has not had experience with this before.  What low-dimensional representations? What basic structures?  Maybe follow this sentence with an example?  I think being more concrete, even at the expense of generality, will bring the point across more clearly.  You give examples below, so maybe just reorder.}  The optimal choice of such reduced representation  depends profoundly on how well it can capture regularities in the underlying data generating process. \Kcomment{Which reduced representation?  I understand that the two sentences are related, but ``low-dimensional representation'' and ``reduction'' are not necessarily equivalent.}
For example, diffuse correlated input onto all neurons implies a low-rank parameterization such as used in  factor analysis whereas presumed strong pairwise linear associations imply sparse partial correlations.  \Kcomment{I am no sure that the second part is true.  Couldn't we have strong all-to-all pairwise correlations without sparse partial correlations?  }
Composite  `sparse+low-rank partial correlation' parameterizations allow combining common correlated input with strong pairwise associations.  \Kcomment{This last sentence is a jump to a new idea (combining the two structures), which is important, but just is a hanger on here.}

To infer the optimal low-dimensional representation of neural correlations in a specitifc neural microcircuit, we compared the performance of several estimators of the covariance matrix on the activity of 100--300 neurons in mouse visual cortex: sample covariance, covariance shrinkage, factor analysis, sparse partial correlations, and sparse+low-rank partial correlations. 
The instantaneous firing rates, binned at 200 ms, were inferred from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.   
As expected, covariance shrinkage reliably outperformed the sample covariance estimate. \Kcomment{Why ``as expected''?  }
Factor analysis-based estimates significantly outperformed covariance shrinkage. \Kcomment{This was not completely expected.  Do you want to say something else?}
Yet sparsified partial correlations with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage-based  estimators. \Kcomment{Is it clear where this approach fits in the list you give above?}
The superior performance of sparse partial correlation-based estimates suggests the relative importance of detailed network interactions over common input in this circuit.
The resulting sparse partial correlation matrix, we argue, serves as a better substrate for functional network analysis. 
In addition, the overall estimation improvement will propagate to any existing analyses that rely on the covariance matrix.
\Kcomment{Such as?}

### 2013-07-30 sections/Introduction_KJ.tex
Kreso's review comments.

\hl{\tiny why correlations:}
Linear correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most widely used and familiar descriptive statistics of neural population activity \citep{Cohen:2011}.  \Kcomment{Need some references to the original literature here.  The earliest I know are papers by Perkel, et al (1967).  The first paper I know where they have recorded from multiple neurons simultaneously is Clark and Gerstein,  Simultaneous studies
of firing patterns in several neurons, Science (1964).  This may be something to cite when you talk about the actual problem -- although in this paper they did not use correlation as a measure of dependence.  There is also a 1964 
RAND Corporation report entitled ``Detection of Functional Interactions Among Neurons: A technique using repetitive presentations of stimuli'' from 1964.  I've only seen it cited, and have no idea how to get it.}
For example, \emph{noise correlations}, i.e.~the correlations of stimulus response variability between pairs of neurons in sensory areas, have been of particular interest thanks, in part, to their profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011}.  
Lending further credence to their role as indicators of detailed and specific functional connectivity \Kcomment{What is functional connectivity?  Isn't it essentially what we uncover using a method like estimating linear correlations?  I think we should be explicit about this.  The previous phrasing implies that there is some concrete ``functional connectivity'' that we are trying to uncover.} have been a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Smith:2013,Denman:2013}.  \Kcomment{The previous sentence is too long. I think it contains two main ideas:  one is relation between functional connectivity and 
correlations, second, correlations and circuit physiology.  They are related, but could be separated. Also, pair discovery
with appropriate reference.}

\hl{\tiny ambiguity of interpretation:}
Neural correlations do not come with ready or unambiguous mechanistic interpretations.   
Theoretical and simulation studies show that neural correlations at various temporal scales arise from multiple underlying mechanisms such as direct synaptic interactions,  common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009} \Kcomment{also add ref to Rosenbaum and Josi\'{c} ``Mechanisms that modulate the transfer of spiking correlations. I think it is relevant.}

\hl{\tiny opportunities with multineuronal data:}
The neural correlation matrix obtained from multineuronal recordings is greater than the sum of its parts: it can be transformed into other representations  that accentuate different aspects of population activity.  
For example, the eigenvalue decomposition of the covariance matrix, also known as principal component analysis (PCA), expresses shared activity components across the entire population;  common fluctuations in the population activity of the entire circuit may be accurately represented by just a few principal components but may affect all correlation coefficients. 
In contrast, the inverse of the correlation matrix reveals the \emph{partial correlations} between neuron pairs.  Such partial correlations reflect specific pairwise linear dependence, conditioned on the activity of all other recorded cells. 
\Kcomment{It may be overkill, but I would add another sentence here to re-emphasize this point.  It is one that
many will probably miss, that is, they will not get the importance of the fact that this is \emph{conditioned} on the 
activity of other cells.}
A strong interaction between a pair of neurons may be expressly represented by a single partial correlation, 
but the effect of such interaction can propagate and affect other pairwise correlations, as well as principal components.

\hl{\tiny motivate dimensionality reduction:}
 Covariance estimates from real high-dimensional data  are both noisy and redundant.  The amount of recorded data increases only linearly with population size whereas the number of free parameters in the covariance matrix increases quadratically.  As a result, empirical covariance estimates for multidimensional data  accumulate substantial redundancy and noise.  \Kcomment{The last sentence in this paragraph says the same thing as the first sentence.  You may want to change this.}

To reduce the estimation error, statistically efficient estimators are regularized.  One regularization approach is to shrink the \Kcomment{Which estimate?  The point estimate?}  estimate toward a less variable low-dimensional estimate \citep{Schafer:2005}.  Another regularization approach, often called \emph{sparsification}, reduces the dimensionality of the estimate by setting its least significant parameters to zero. 
Regularized estimates perform favorably on new samples of data compared to the full model, provided the amount of regularization is chosen carefully.

\hl{\tiny choosing between regularization schemes}
Estimator improvements due to regularization depend profoundly on the parameterization to which it is applied.  
\Kcomment{I would argue that more than that, it depends on the structure of the underlying data. In fact, I don't really
understand the previous statement.}
Sparsification \Kcomment{Will it be clear to this audience that sparsification is the same as setting some terms (eigenvalues, entries in a matrix) equal to zero?} of the eigenspectrum produces a low-rank estimate of the covariance matrix also known as truncated PCA \citep{Rothman:2008} \Kcomment{This is grammatically incorrect. The estimate cannot be known as the PCA, \emph{i.e.} the estimate is not an analysis.} or, allowing for additional independent variances, as factor analysis (FA) \citep{Fan:2008}. \Kcomment{This last part of the sentence is obscure.  What ``additional variances''?}  Sparsification of the inverse covariance is known as covariance selection \citep{Dempster:1972,Friedman:2008} and is related to finding a graphical model in which a large fraction of neuron pairs are conditionally independent of each other.  Representations that are capable of concentrating the significant real effects in a small number of parameters will reap greatest benefits from sparsification over ones that distribute the real effects widely across many parameters.  Random noise tends to be broadly distributed across all parameters in any parameterization. \Kcomment{Underlying all this is the idea of Occam's Razor, and I think you should at least mention
this. You do not need to talk about this from a Bayesian viewpoint, but you can say that you are avoiding overfitting
by choosing a smaller set of parameters.}

\hl{\tiny our approach and summar of findings} 
In this study, we compared the performance of several covariance estimation schemes for spatially compact \Kcomment{This is important, and you should explain a bit more about what you mean by spatially compact.  This could
mean different things.} groups of 150--300 neurons in layers 2-4 in mouse primary visual cortex during visual stimulation.   The performance of the estimators was evaluated by computing the mean squared error between the optimized covariance estimate fitted to training data and the non-regularized covariance matrix from a separate test data set.  Low-rank covariance estimates performed significantly better than shrinkage estimators, but estimators with sparse partial correlations were more efficient still. Typically, between 3 and 16\% of neuronal paris were connected by non-zero partial interactions.  Mixed sparse estimators with sparse partial correlations and low-rank components performed comparably. 
\Kcomment{Will you mention relation to physiology in the Discussion?}

As noted above, the optimal covariance estimator is domain-dependent. 
These specific findings may not spread to all neural circuits. \Kcomment{Do you mean that they may not hold 
for circuits in different areas of cortex?}
Improved accuracy estimation accuracy is not the main aim or the only benefit.  \Kcomment{Then what is -- I do not 
understand how you can have what you describe in what follows without an accurate estimate of partial correlations. 
Indeed, wouldn't an increase in accuracy help here, with the caveats that we discussed many times?}
The selection of the optimal regularization scheme in itself serves as an effective descriptor of the dominant correlation structures. The finding that a identifiable subsets of partial correlations were important to describe the partial correlation makes such pairs as prime candidates as direct functional interactions.

\Kcomment{Several things are not discussed here -- I think it is important that we choose models, and 
estimation approaches according to what we think is true about the physiology.  The approaches
that you suggest above do have this connection -- FA because coherent drive may be the most important,
sparsifying the inverse,  because connectivity is sparse.   Thus there is a good reason to choose these
approaches, and it is not only because this is what statisticians give us. Also, what are the
shortcomings of linear correlations, \emph{i.e.} how can nonlinear interactions  change the interpretation.?}

### 2013-07-30 sections/Theory.tex
Kreso's review comments.

@@
 Let $x_i \in \mathcal X, i=1\ldots,x_n$ denote a sample of consequtive observations of population activity of $p$ neurons in time bins $i$.  
-The activity of each neuron is the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
+The activity of each neuron is the real-valued firing rate \TODO{Need to define rate more precisely}, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
 We assume that observations are drawn from some true cumulative probability distribution $F: \mathcal X \mapsto R$.
@@
 where $\mathbf 1(x \ge x_i)$ is the indicator function which equals 1 when all elements of $x$ are greater than the corresponding elements of $x_i$ and 0 otherwise.
+\TODO{Are you using cumulative distributions to avoid binning?}
 
@@
 Then \emph{excess loss} $\ell(\hat S, S) = \mathcal L(\hat S, S) - \mathcal L(S,S)$ assumes zero at the minimum.
+\TODO{This is too abstract -- need to motivate with some examples maybe}
 

### 2013-07-30 sections/Theory_KJ.tex
Kreso's review comments.

\subsection*{Problem setup}
Let $x_1,\ldots,x_n \in \mathbb R^{p\times 1}$ denote a sample of $n$ observations of instantaneous activity (e.g.~spiking) rates of $p$ neurons averaged in time bins $i=1,\ldots,n$.

For simplicity, we assume that observations are sampled from a shared true distribution $x_1,\ldots,x_n \sim f$. 
\hl{\tiny explain: the derivations remain valid even without such a strong assumption.  We account for overdispersion due to temporal dependence and shuffle chunks of data to account for nonstationarities.} 
\Kcomment{Do the samples have to be independent?  I assume not, but does this complicate some
of the explanation below?}

For simplicity, we assume that the data have zero means. \Kcomment{This is confusing.  I think you are assuming
that the true marginals have zero means.  The sample means will not be zero, right?  I'd say that data having zero
means sample means are zero.}
We seek to estimate the true covariance matrix 
\begin{equation}
\Sigma=\mathbb E_{z\sim f}[ zz^\T].
\end{equation}

The purpose of the estimation procedure is to use the sample $x_1,\ldots,x_n$ to compute a covariance estimate $S$ that is as close as possible to $\Sigma$. 
The quality of the covariance estimate $S$ will be defined using a \emph{loss function} $\mathcal L\left(S, \Sigma\right)$, which is minimized when $S=\Sigma$.

In particular, we consider the mean-squared loss
\begin{equation}
\mathcal L_e( S, \Sigma) =  \| S - \Sigma\|^2_F 
\end{equation}
where $\| \cdot \|_F$ denotes the Frobenius norm;

and the multivariate normal log likelihood loss
\begin{equation}
\mathcal L_g(S,\Sigma) = \ln |S| +  \Tr(S^{-1} \Sigma)
\end{equation}
where $|S|$ denotes the determinant of matrix $S$. \Kcomment{You may want to explain both of these.
Again, keep the audience in mind.  First they may not know what the Frobenius norm is.  Second,
they may appreciate a reminder of the log likelihood loss. You can give the above equation (3) as a
special case then.}

\hl{\tiny list other possible loss functions and justify the choice of these two. Consider Stein's entropy loss and qudratic loss.   James:1961 showed failures of the MSE loss, but it continued to be widely used. Fan:2008 provides other arguments.}

The linearity of $\mathcal L_e(S,\Sigma)$ and $\mathcal L_g(S,\Sigma)$ with respect to the elements of $\Sigma$ allows unbiased empirical estimation of $\mathcal L$ from a validation sample $y_1,\ldots,y_m \sim f$ acquired independently of the original training sample $x_1,\ldots,x_n$ that was used to construct the estimate $S$.
\begin{equation}
\begin{split}
\mathcal L(S,\Sigma) & \equiv \mathbb E_{z\sim f} \left[ \mathcal L(S,zz^\T) \right] \\ 
&\approx \frac 1 m \sum\limits_{i=1}^m \mathcal L(S,y_i y_i^\T) = \mathcal L(S,\hat S_y)
\end{split}
\end{equation}
where $\hat S_y = \frac 1 m \sum\limits_{i=1}^m x_i x_i^\T$ is the empirical covariance of the validation dataset. 
\Kcomment{I think the previous is too terse.  I think you should explain the validation sample.}

\hl{\tiny cross-validation}
In practice, neural recordings are limited.   Instead of validation with a dedicated validation dataset, $K$-fold cross-validation is used: The data are split into $K$ subsets of approximately equal size and each subset is used as the validation set with the covariance estimator trained on the remaning data.  The estimation of the loss from the $K$ tests is then averaged, which is justified, again, thanks to the linearity property of the chosen loss functions. 
\hl{\tiny exlain how the data are split into continuous chunks so that dependences between the training and validation datasets due to temporal correlations are minimized.  Coversely, each test set comprises multiple distirbuted chunks to reduce the effect of non-stationarities.}

 \Kcomment{Will you explain double cross-validation?}

\subsection*{Covariance estimators}
It has long been recognized that the empirical covariance 
\begin{equation}
\hat S = \frac 1 n \sum\limits_{i=1}^n x_i x_i^\T
\end{equation}
is statistically inadmissible, meaning that larger sample sizes are required to reduce the loss function to the same level as other, more statistically efficient covariance estimators. \Kcomment{I am not quite sure what this implies. Inadmissible 
seems to mean that you can never use it.  Is this what you want to say?}

Covariance estimators can be made more efficient using two broad approaches, which are not mutually exclusive: (1) deliberately bias the estimator in order to reduce its variability and (b) reduce the dimensionality of the estimator by choosing a specific parametric model.

\subsubsection*{Shrinkage estimators}
The most familiar example of the first approach is the linear shrinkage estimator, calculated is a linear mixture of the empirical covariance estimate $\hat S$ and another low-variance target estimator $T$:
\begin{equation}
S_{shrink} = (1-\lambda) \hat S + \lambda T
\end{equation}
A popular target covariance is the identity matrix scaled by the mean sample variance  $T = \frac 1 p \Tr(\hat S) I$.
The optimal value of shrinkage estimate with respect to $\mathcal L_e$ can be estimated analytically from the data \citep{Ledoit:2004,Schafer:2005}.

\subsubsection*{Dimensionality reduction}
The second broad approach is to impose a structure on the covariance estimate by reducing the number of free parameters in one of its parameterizations.

### 2013-07-31 sections/Theory.tex
small changes

@@
 where $\mathbf 1(x \ge x_i)$ is the indicator function which equals 1 when all elements of $x$ are greater than the corresponding elements of $x_i$ and 0 otherwise.
-\TODO{Are you using cumulative distributions to avoid binning?}
+\TODO{KJ: Are you using cumulative distributions to avoid binning?
+DY: Yes.
+Many theorems such as Glivenko-Catelli Theorem and Skorohod's Representation are proven using  cumulative representation.
+The empirical distribution is nearly always represented in its cumulative form.
+Expressing convergence is requires binning or using cumulative; the latter is simpler.}
 
@@
 \end{equation}
-the bias-variance decomposition is the simple sum
+the variance-bias decomposition is the simple sum
 \begin{equation}
@@
 We follow the convention that $\hat\Sigma_{\gamma=0}$ is the unbiased estimator $\hat\Sigma_0$, \emph{i.e.}\;$b_{\gamma=0}^2 = 0$. 
-Then as $\gamma$ increases, $\hat\Sigma_\gamma$ less variable but increasingly biased.
-Thus $\hat\Sigma_\gamma$ defines the \emph{regularization path}, for which \emph{regularization intensity} $\gamma$ specifies the location along the path, which may be real-valued or discrete, or even a vector to steer the bias toward multiple targets.
+We define a sequence of approximations $\hat\Sigma_\gamma$ that become less variable as as $\gamma$ increases, which is usually accompanies by increased bias.
+Then $\hat\Sigma_\gamma$ defines the \emph{regularization path}, for which \emph{regularization intensity} $\gamma$ specifies the location along the path, which may be real-valued or discrete, or even a vector to steer the bias toward multiple targets.
+
+An unbiased estimate does not guarantee lack of bias in nonlinear functions of that estimate: $\mathbb E\left[\hat\Sigma_0\right]=0 \centernot\implies  E\left[\varphi(\hat\Sigma_0\right)]=0$, where $\varphi$ is a nonlinear function.
 

### 2013-08-01 sections/Theory.tex
minor

@@
-\subsection{Covariance estimation}
+\subsection{Theory and notation}
 
@@
 
-An unbiased estimate does not guarantee lack of bias in nonlinear functions of that estimate: $\mathbb E\left[\hat\Sigma_0\right]=0 \centernot\implies  E\left[\varphi(\hat\Sigma_0\right)]=0$, where $\varphi$ is a nonlinear function.
+An unbiased estimate does not guarantee lack of bias in nonlinear functions of that estimate: $\mathbb E\left[\hat\Sigma_0\right]=\Sigma \centernot\implies  E\left[\varphi(\hat\Sigma_0)\right]=\varphi(\Sigma)$, where $\varphi$ is a nonlinear function.
+For example, if $u_{\max}(S)$ is the largest eigevalue of square matrix $S$, then $\mathbb E\left[u_{\max}(\hat\Sigma_0)\right] > u_{\max}(\Sigma)$.
+Some regularization schemes have been designed to correct for the eigenspectrum bias  at the cost of adding bias to covariance coefficients \citep{Ledoit:2004}.
 
@@
 
-\paragraph{Constructing a regularized estimator:}
+\subsection{Regularized covariance estimates}
 The task of designing a regularized estimators consists of defining a sequence of approximations $\hat\Sigma_\gamma$, called \emph{regularization path} and the optimal location on that path, \emph{regularization intensity} $\hat\gamma$ as  functions of the empirical distribution.

### 2013-08-07 sections/Results.tex
added Figure 2: regularization target as graphical models.

@@
-\textbf{Figure 3.}{ 
+\textbf{Figure 4.}{ 
 The fitting performance of regularized covariance estimators on two-photon recording. 
-}\label{fig:03}
-
-\textbf{Figure 4.}{ 
-Fitting performance of regularized covariance estimators on 24 datasets two-photon recordinngs.
 }\label{fig:04}
@@
 \textbf{Figure 5.}{ 
+Fitting performance of regularized covariance estimators on 24 datasets two-photon recordinngs.
+}\label{fig:05}
+
+\textbf{Figure 6.}{
 The depiction of the 3D arrangement of the non-zero partial correlations between pairs of neurons.
-}\label{fig:04}
+}\label{fig:06}

### 2013-08-07 sections/Theory.tex
added Figure 2: regularization target as graphical models.

@@
 
-\subsection{Regularized covariance estimates}
+\subsection{Regularization of covariance estimates}
 The task of designing a regularized estimators consists of defining a sequence of approximations $\hat\Sigma_\gamma$, called \emph{regularization path} and the optimal location on that path, \emph{regularization intensity} $\hat\gamma$ as  functions of the empirical distribution.
@@
 \end{enumerate}
+
+\textbf{Figure 2.}{
+Gaussian Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
+\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
+\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent. 
+\textbf{C}: The Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from pairwise interactions between 
+\textbf{D}: A similar covariance matrix with sparse + low-rank inverse with 3 latent units.
+}\label{fig:02}

### 2013-08-08 sections/Theory.tex
updated the Theory section.

@@
-\subsection{Theory and notation}
+\subsection{Overview}
+Given a finite sample, the \emph{sample covariance matrix} $\hat\Sigma_0$ is the usual estimator of the \emph{true covariance matrix} $\Sigma$ of the underlying random process.
+$\hat\Sigma_0$ is unbiased but, with finite sample sizes, not as close to $\Sigma$ as possible.
+More \emph{efficient} estimators can be constructed by \emph{regularization}, which is the deliberate biasing of the usual estmate toward a different, low-dimensional parametric \emph{target estimate} $\hat T$.
+Target estimates are biased approximations because they cannot represent an arbitrary $\Sigma$, but they also have lower random variability.
+Regularization reduces the overall error by striking a favorable balance between the variance  (\emph{estimation error}) and the bias \emph{(approximation error)} of the estimate.
 
-\paragraph{Problem setup:}
+Regularization requires making the following three choices: 
+\begin{enumerate}[  1. ]
+\item define a set of low-dimensional target estimates
+\item given the data, select the optimal target estimate from the set 
+\item given the data, select the optimal amount of biasing toward the target estimate
+\end{enumerate}
+The first choice, the selection of the general form of target estimates, must be made \emph{a priori}.
+Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of the unbiased estimator.
+The perplexing fact that estimates can be improved by biasing toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.
+
+\subsection{Probabilistic framework}
 
@@
 The observations need not be independently and identically sampled from $F$.
-The only assumption is that the data generating process is ergodic, \emph{i.e.}\;that $\max_x \left[\hat F_n(x) - F(x)\right] \to 0$ as the sample size $n$ increases.
+The only assumption is that the data generating process is ergodic, \emph{i.e.}\;that $\max\limits_x \left[\hat F_n(x) - F(x)\right] \to 0$ as the sample size $n$ increases.
 Then each $x_i$ can be thought to be drawn from $F$ when not conditioned on nearby observations.
@@
 
-\paragraph{Variance-bias tradeoff:}
+\subsection{Measures of error}
 Let $\hat\Sigma_\gamma$ be a family of approximate (biased) covariance estimates indexed by $\gamma$.
@@
 
-\textbf{Figure 2.}{
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{figures/Figure2.pdf}
+\caption{
+%\textbf{Figure 2.}{
 Gaussian Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
@@
 \textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent. 
-\textbf{C}: The Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from pairwise interactions between 
-\textbf{D}: A similar covariance matrix with sparse + low-rank inverse with 3 latent units.
+\textbf{C}: In the Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from sparse pairwise linear interactions between visible units. 
+\textbf{D}: In the Gaussian graphical model with latent units, correlations arise  between pairs of nodes 
 }\label{fig:02}
+%}
+\end{figure}

### 2013-08-09 sections/Simulation.tex
added Figure 2: regularization target as graphical models.
Added figure with simulation results.

@@
-\subsection{Simulation}
-To illustrate,
+To illustrate
 
-\textbf{Figure 2.}{
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/Figure3.pdf}
+\caption{
+%\textbf{Figure 3.}{
 Simulation results.
-\textbf{A}: A $100 \times 100$ covariance matrix without a low-dimensional structure, a sample covariance with n=2000, distributions of true and sample correlation coefficients, eigenspectra of true and sample covariances.
-\textbf{B}: A similar covariance matrix with 8 factors.
-\textbf{C}: A similar covariance matrix with sparse inverse. 
-\textbf{D}: A similar covariance matrix with sparse + low-rank inverse with 3 latent units.
-\textbf{E}: Risk convergence rates
-}\label{fig:02}
+\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
+}\label{fig:03}
+%}
+\end{figure}

### 2013-08-10 sections/Introduction.tex
minor changes
minor

@@
-Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \citep{Cohen:2011}.
+Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \citep{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Renart:2010}.
 For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been of particular interest.
-Noise correlations can have profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011}, and have been interpreted to indicate detailed and specific functional organization.
-Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Ko:2011,Smith:2013,Denman:2013}.
+Noise correlations can have profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}, and have been interpreted to indicate detailed and specific functional organization.
+Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \citep{Smith:2008,Denman:2013}, their synaptic connectivity and stimulus tuning similarity \citep{Kohn:2005,Ko:2011}, cortical layer specificity \citep{Hansen:2012,Smith:2013}, cell-type specificity (?), progressive changes in development and in learning \citep{Golshani:2009}, changes due to sensory stimulation and global brain states \citep{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
 

### 2013-08-10 sections/Theory.tex
minor

@@
 
-Regularization requires making the following three choices: 
+A regularized estimate requires the following three choices: 
 \begin{enumerate}[  1. ]
@@
 \subsection{Probabilistic framework}
-
+\TODO{This section is not strictly necessary. It helps relate probabilistic models to our approach, which does not rely on an explicity probabilistic model.}
 Let $x_i \in \mathcal X, i=1\ldots,x_n$ denote a sample of consequtive observations of population activity of $p$ neurons in time bins $i$.  
@@
 %\textbf{Figure 2.}{
-Gaussian Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
+Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
 \textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 

### 2013-08-13 sections/Introduction.tex
minor

@@
 
-Neural correlations do not come with ready or unambiguous mechanistic interpretations.
-Theoretical and simulation studies show that neural correlations at various temporal scales may arise from any combination of underlying mechanisms.
+However, neural correlations do not come with ready or unambiguous mechanistic interpretations.
+Theoretical and simulation studies have shown that neural correlations may arise at various temporal scales from combinations of multiple underlying mechanisms.
 These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
@@
 
-With this study, we pursue two goals: 
+With this study, we pursue two related goals: 
 \begin{enumerate}[\qquad A.\;\;]
-	\item Devise better estimators of neural covariance matrices in recordings from large populations of neurons.
+	\item Devise more efficient estimators of neural covariance matrices in recordings from large populations of neurons.
 	\item Aid interpretation of population activity  by isolating the essential structure of the neural covariance matrix.

### 2013-08-13 sections/Theory.tex
minor

@@
 \subsection{Overview}
-Given a finite sample, the \emph{sample covariance matrix} $\hat\Sigma_0$ is the usual estimator of the \emph{true covariance matrix} $\Sigma$ of the underlying random process.
+Given the sample $x_1,\ldots,x_n$ of $p\times 1$ observations, the usual estimator of the true covariance matrix $\Sigma$ is  the sample covariance matrix:
+\begin{equation}
+\hat\Sigma_0 =  \frac 1 {n-d} \sum\limits_{i=1}^n (x_i - \bar x) (x_i - \bar x)^\T
+\quad\mbox{where}\;
+\bar x = \frac 1 n \sum\limits_{i=1}^n 
+\end{equation}
+where the \emph{index of disperson} $d=1$ when observations are independent or $d>1$ when nearby samples are correlated; the value of $d$ can be estimated from the data.
+
 $\hat\Sigma_0$ is unbiased but, with finite sample sizes, not as close to $\Sigma$ as possible.
-More \emph{efficient} estimators can be constructed by \emph{regularization}, which is the deliberate biasing of the usual estmate toward a different, low-dimensional parametric \emph{target estimate} $\hat T$.
-Target estimates are biased approximations because they cannot represent an arbitrary $\Sigma$, but they also have lower random variability.
+More \emph{efficient} estimators can be constructed by \emph{regularization}, which is the deliberate biasing of the usual estmate toward a less variable \emph{target estimate} $\hat T$.
+Target estimates are low-dimensional approximations cannot represent an arbitrary $\Sigma$, but they also have lower random variability.
 Regularization reduces the overall error by striking a favorable balance between the variance  (\emph{estimation error}) and the bias \emph{(approximation error)} of the estimate.
+
+\subsection{Target estimates}
+A good target estimate is one that has low dimensionality to reduce its variance while still allowing to accurately match typical covariance matrices in the given domain.
 
@@
 Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of the unbiased estimator.
-The perplexing fact that estimates can be improved by biasing toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.
+The perplexing fact that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.
 
@@
 
-The usual unbiased estimate, the sample covariance matrix, rescales the plug-in estimator to compensate for the bias
-\begin{equation}
-\hat\Sigma_0 = \frac n {n - d} \hat\Sigma \equiv \frac 1 {n-d} \sum\limits_{i=1}^n (x_i - \bar x) (x_i - \bar x)^\T
-\end{equation}
-When observations are independent, then $d=1$.
-But when nearby samples are correlated, the estimate of the mean absorbs more of the variance, becomes \emph{overdispersed}.
-As a result the covariances become underestimated.
-Then we can use $d=cn$ where $c>1$ is called the \emph{overdispersion coefficient}, which may be estimated from the data.
-
 \subsection{Measures of error}
@@
 
-When the loss function is the error sum of squares, which, for matrices, is called the Frobenius norm, 
+When the loss function is the mean squared error, which, for matrices, is called the normalized, squared Frobenius norm: 
 \begin{equation}
-\mathcal L_e(\hat S,S) = \|\hat S-S\|_F^2 = \Tr\left((\hat S-S)(\hat S-S)^\T\right)
+\mathcal L_e(\hat S,S) =\tfrac 1 n \|\hat S-S\|_F^2 = \tfrac 1 n \Tr\left((\hat S-S)(\hat S-S)^\T\right)
 \end{equation}

### 2013-08-14 sections/Theory.tex
edited the Theory section

@@
-\subsection{Overview}
-Given the sample $x_1,\ldots,x_n$ of $p\times 1$ observations, the usual estimator of the true covariance matrix $\Sigma$ is  the sample covariance matrix:
-\begin{equation}
-\hat\Sigma_0 =  \frac 1 {n-d} \sum\limits_{i=1}^n (x_i - \bar x) (x_i - \bar x)^\T
-\quad\mbox{where}\;
-\bar x = \frac 1 n \sum\limits_{i=1}^n 
-\end{equation}
-where the \emph{index of disperson} $d=1$ when observations are independent or $d>1$ when nearby samples are correlated; the value of $d$ can be estimated from the data.
+\subsection{Basic concepts}
+Let $x_i \in \mathcal X, i=1\ldots,n$ denote a sample of consecutive observations of population activity of $p$ neurons in time bins $i$.  
+The activity of each neuron is denoted by the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
+For example, in our applicaiton, the firing rate is estimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x_i$ to take on negative values.
 
-$\hat\Sigma_0$ is unbiased but, with finite sample sizes, not as close to $\Sigma$ as possible.
-More \emph{efficient} estimators can be constructed by \emph{regularization}, which is the deliberate biasing of the usual estmate toward a less variable \emph{target estimate} $\hat T$.
-Target estimates are low-dimensional approximations cannot represent an arbitrary $\Sigma$, but they also have lower random variability.
-Regularization reduces the overall error by striking a favorable balance between the variance  (\emph{estimation error}) and the bias \emph{(approximation error)} of the estimate.
+We do not assume that the observations $x_i$ are independently distributed.
+Rather, the data generating process is assumed to be \emph{ergodic}, \emph{i.e.}\;described by a \emph{true distirbution} $F: \mathcal X \mapsto  \mathbb [0, 1]$ (cumulative) over long periods of time such that the \emph{empirical distribution} $\hat F_n$ from a given sample will convege to $F$ with increasing sample size $n$.
 
-\subsection{Target estimates}
-A good target estimate is one that has low dimensionality to reduce its variance while still allowing to accurately match typical covariance matrices in the given domain.
-
-A regularized estimate requires the following three choices: 
-\begin{enumerate}[  1. ]
-\item define a set of low-dimensional target estimates
-\item given the data, select the optimal target estimate from the set 
-\item given the data, select the optimal amount of biasing toward the target estimate
-\end{enumerate}
-The first choice, the selection of the general form of target estimates, must be made \emph{a priori}.
-Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of the unbiased estimator.
-The perplexing fact that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.
-
-\subsection{Probabilistic framework}
-\TODO{This section is not strictly necessary. It helps relate probabilistic models to our approach, which does not rely on an explicity probabilistic model.}
-Let $x_i \in \mathcal X, i=1\ldots,x_n$ denote a sample of consequtive observations of population activity of $p$ neurons in time bins $i$.  
-The activity of each neuron is the real-valued firing rate \TODO{Need to define rate more precisely}, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
-We assume that observations are drawn from some true cumulative probability distribution $F: \mathcal X \mapsto R$.
-Then let $\hat F_n$ denote the empirical distribution function obtained from the observed sample, which itself is a random variable.
+Formally, the emprical distribution is defined as \TODO{this may be unecessary, but I leave it for now.}
 \begin{equation}
@@
 
-The observations need not be independently and identically sampled from $F$.
-The only assumption is that the data generating process is ergodic, \emph{i.e.}\;that $\max\limits_x \left[\hat F_n(x) - F(x)\right] \to 0$ as the sample size $n$ increases.
-Then each $x_i$ can be thought to be drawn from $F$ when not conditioned on nearby observations.
+And, for an ergodic process, $\max\limits_{x\in\mathcal X} \left|\hat F_n(x) - F(x)\right| \to 0$ as $n$ increases.
 
-The true covariance matrix $\Sigma \in \Theta$ is defined as a function of the true probability distribution
+The true covariance matrix $\Sigma \in \Theta$ is defined as a function of $F$:
 \begin{equation}
-\Sigma = \int\limits_{z\in\mathcal X} (z - \bar x)(z - \bar x)^\T \dif F(z)
+\Sigma = \int\limits_{z\in\mathcal X} (z - \mu)(z - \mu)^\T \dif F(z)
 \quad\mbox{where}\;
-\bar x = \int\limits_{z\in\mathcal X} z \dif F(z)
+\mu = \int\limits_{z\in\mathcal X} z \dif F(z)
 \end{equation}
-The \emph{plug-in estimator} $\hat\Sigma$ is obtained by plugging the empirical distribution $F_n$ in place of the true distribution $F$
+The domain $\Theta$ is the set of all positive-definite $p\times p$ matrices, which is a cone in $\mathbb R^{p\times(p+1)/2}$.
+
+But the usual estimator of the true covariance matrix $\Sigma$ is  the sample covariance matrix:
 \begin{equation}
-\begin{split}
-\hat\Sigma = \int\limits_{z\in\mathcal X} (z - \bar x)(z - \bar x)^\T \dif \hat F_n(z) 
-\equiv \frac  1 n\sum\limits_{i=1}^n (x_i-\bar x)(x_i - \bar x)
-\\
-\mbox{where}\;\bar x = 
-\int\limits_{z\in \mathcal X}z \dif \hat F_n(z) \equiv \frac 1 n\sum\limits_{i=1}^n x_i
-\end{split}
+\hat\Sigma_0 =  \frac 1 {n-c} \sum\limits_{i=1}^n (x_i -\hat\mu) (x_i - \hat\mu)^\T
+\quad\mbox{where}\;
+\hat\mu = \frac 1 n \sum\limits_{i=1}^n x_i
 \end{equation}
-The plug-in estimator is known to be biased toward zero because the estimate of the mean $\bar x$ absorbs a fraction of the variances.
+where $c$ is the sum of temporal correlations \TODO{there must be a better description of this}.
+For independent observations, $c=1$;  $c>1$ when nearby samples are correlated; the value of $c$ can be estimated from the data.
 
-\subsection{Measures of error}
+$\hat\Sigma_0$ is designed to be unbiased such that $\mathbb E\left[\hat\Sigma_0\right]=\Sigma$.
+Here and below, $\mathbb E[\cdot]$ denotes the expected value under the  unknown true distribution $F$.
+All variables with a hat (\emph{e.g.}\;$\hat \Sigma_0,\hat \mu$) are functions of the empirical distribution $\hat F_n$, which itself is a random variable.
+
+Nonlinear functions of an unbiased estimate can be biased: $\mathbb E\left[\hat\Sigma_0\right]=\Sigma \centernot\implies  E\left[\varphi(\hat\Sigma_0)\right]=\varphi(\Sigma)$.
+For example, if $u_{\max}(S)$ is the largest eigevalue of square matrix $S$, then $\mathbb E\left[u_{\max}(\hat\Sigma_0)\right] > u_{\max}(\Sigma)$.
+Some estimators have been designed to correct for the eigenspectrum bias at the cost of adding bias to covariance coefficients \citep{Ledoit:2004}.
+
+\subsection{Loss function}
+The optimization of a covariance matrix estimate $\hat\Sigma$ is performed with respect to a \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which expresses the discrepancy \TODO{right word?} between $\hat\Sigma$ and $\Sigma$ and attains its minimum when $\hat\Sigma=\Sigma$.
+Then \emph{excess loss}  
+\begin{equation}
+\ell(\hat\Sigma,\Sigma) = \mathcal L(\hat\Sigma,\Sigma)-\mathcal L(\Sigma,\Sigma)
+\end{equation}
+assumes zero at its minimum.
+
+A particularly useful loss function is the mean squared error (MSE), which is proportional to the square of the Frobenius  norm $\|\cdot\|_F$ of the difference between the matrices: 
+\begin{equation}
+\mathcal L_e(\hat\Sigma,\Sigma) =\frac 1 p \|\hat\Sigma-\Sigma\|_F^2 = \frac 1 p \Tr\left((\hat \Sigma-\Sigma)(\hat\Sigma-\Sigma)^\T\right)
+\end{equation}
+The MSE is its own excess loss: $\ell_e(\hat\Sigma,\Sigma) \equiv \mathcal L_e(\hat\Sigma,\Sigma)$.
+
+The Gaussian loss function $\mathcal L_g$  arises from the theory of normal distributions.
+When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{i=1}^n x_i x_i^\T$ is
+\begin{equation}
+L\left(\Sigma \mid \hat\Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
+\end{equation}
+Then $\mathcal L_g$ is constructed by rescaling $L\left(\Sigma \mid \hat\Sigma\right)$ and dropping the constant term:
+\begin{equation}
+\mathcal L_g(\hat\Sigma,\Sigma) 
+=  -\frac 2 {pn} L\left(\Sigma \mid \hat\Sigma \right) - \frac 1 p \ln(2\pi) 
+\equiv  \frac 1 p\left(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}) \right) 
+\end{equation}
+The corresponding excess loss 
+\begin{equation}
+\ell_g(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) - \mathcal L_g(\Sigma,\Sigma)  
+= \frac 1 p \left(-\ln \det (\hat \Sigma^{-1} \Sigma) + \Tr(\hat \Sigma^{-1}\Sigma)\right) - 1
+\end{equation}
+is known as \emph{entropy loss} \citep{James:1961}.
+
+Despite the fact that entropy loss is derived from normal theory, the choice of a loss function is not equivalent to assuming a specific form of $F$.
+The loss function measures the discrepancy between distribution parameters rather than the distance between distributions.
+\TODO{explain Bregman divergence?}
+
+\subsection{Emprical loss}
+
+\subsection{Variance/bias decomposition}
+the variance-bias decomposition is the simple sum
+\begin{equation}
+r_\gamma =  b_\gamma^2 + \varepsilon_\gamma^2
+\end{equation}
+
 Let $\hat\Sigma_\gamma$ be a family of approximate (biased) covariance estimates indexed by $\gamma$.
@@
 
-When the loss function is the mean squared error, which, for matrices, is called the normalized, squared Frobenius norm: 
-\begin{equation}
-\mathcal L_e(\hat S,S) =\tfrac 1 n \|\hat S-S\|_F^2 = \tfrac 1 n \Tr\left((\hat S-S)(\hat S-S)^\T\right)
-\end{equation}
-the variance-bias decomposition is the simple sum
-\begin{equation}
-r_\gamma =  b_\gamma^2 + \varepsilon_\gamma^2
-\end{equation}
+With other loss functions, the risk is also an increasing function of the two components, although is not generally the sum.
 
-With other loss functions, the risk is also an increasing function of the two components, although is not generally the sum.
+\subsection{Target estimates}
+A good target estimate is one that has low dimensionality to reduce its variance while still allowing to accurately match typical covariance matrices in the given domain.
+
+A regularized estimate requires the following three choices: 
+\begin{enumerate}[  1. ]
+\item define a set of low-dimensional target estimates
+\item given the data, select the optimal target estimate from the set 
+\item given the data, select the optimal amount of biasing toward the target estimate
+\end{enumerate}
+The first choice, the selection of the general form of target estimates, must be made \emph{a priori}.
+Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of the unbiased estimator.
+The perplexing fact that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.
+
+$\hat\Sigma_0$ is unbiased but, with finite sample sizes, not as close to $\Sigma$ as possible.
+More \emph{efficient} estimators can be constructed by \emph{regularization}, which is the deliberate biasing of the usual estmate toward a less variable \emph{target estimate} $\hat T$.
+Target estimates are low-dimensional approximations cannot represent an arbitrary $\Sigma$, but they also have lower random variability.
+Regularization reduces the overall error by striking a favorable balance between the variance  (\emph{estimation error}) and the bias \emph{(approximation error)} of the estimate.
 
@@
 
-An unbiased estimate does not guarantee lack of bias in nonlinear functions of that estimate: $\mathbb E\left[\hat\Sigma_0\right]=\Sigma \centernot\implies  E\left[\varphi(\hat\Sigma_0)\right]=\varphi(\Sigma)$, where $\varphi$ is a nonlinear function.
-For example, if $u_{\max}(S)$ is the largest eigevalue of square matrix $S$, then $\mathbb E\left[u_{\max}(\hat\Sigma_0)\right] > u_{\max}(\Sigma)$.
-Some regularization schemes have been designed to correct for the eigenspectrum bias  at the cost of adding bias to covariance coefficients \citep{Ledoit:2004}.
-
 \paragraph{Evaluation of estimators:}
-Another popular loss  function for covariance estimation arises from the theory of normal distributions.
-Under the assumption of normality  with zero mean, the log likelihood of the covariance matrix $\Sigma$ given the independent sample $z_1,\ldots,z_m$, with $S = \tfrac 1 m \sum\limits_{i=1}^m z_i z_i^\T$,
-\begin{equation}
-L\left(\Sigma \mid S\right) = -\tfrac m 2 \ln(2\pi) - \tfrac m 2 \ln \det \Sigma - \tfrac m 2 \Tr(\Sigma^{-1} S)
-\end{equation}
-Then the Gaussian loss function $\mathcal L_g$ is constructed by rescaling the normal log likelihood and dropping the constant term
-\begin{equation}
-\begin{split}
-\mathcal L_g(\hat S,S) = &-\frac 2 m L\left(S \mid \hat S \right) - \ln(2\pi) 
-\\ 
-\equiv & \ln \det \hat S + \Tr(\hat S^{-1} S) 
-\end{split}
-\end{equation}
-The corresponding excess loss 
-\begin{equation}
-\ell_g(\hat S,S) = \mathcal L_g(\hat S,S) - \mathcal L_g(S,S) 
-= -\ln \det \left(\hat S^{-1} S\right) + \Tr\left(\hat S^{-1}S\right) - p
-\end{equation}
-is known as \emph{entropy loss} \citep{James:1961}.
-
-Despite the fact that entropy loss is derived from normal theory, the choice of a loss function is not equivalent to assuming a specific form of $F$.
-The loss function measures the discrepancy between distribution parameters rather than the distance between distributions.
-\TODO{explain Bregman divergence?}
 

### 2013-08-15 sections/Theory.tex
expanded the Theory section

@@
 $\hat\Sigma_0$ is designed to be unbiased such that $\mathbb E\left[\hat\Sigma_0\right]=\Sigma$.
-Here and below, $\mathbb E[\cdot]$ denotes the expected value under the  unknown true distribution $F$.
+Here and througout, $\mathbb E[\cdot]$ denotes the expected value under the  unknown true distribution $F$.
 All variables with a hat (\emph{e.g.}\;$\hat \Sigma_0,\hat \mu$) are functions of the empirical distribution $\hat F_n$, which itself is a random variable.
@@
 For example, if $u_{\max}(S)$ is the largest eigevalue of square matrix $S$, then $\mathbb E\left[u_{\max}(\hat\Sigma_0)\right] > u_{\max}(\Sigma)$.
-Some estimators have been designed to correct for the eigenspectrum bias at the cost of adding bias to covariance coefficients \citep{Ledoit:2004}.
+Some covariance matrix estimators are designed to correct for the eigenspectrum bias at the cost of adding bias to covariance coefficients \citep{Ledoit:2004}.
 
 \subsection{Loss function}
-The optimization of a covariance matrix estimate $\hat\Sigma$ is performed with respect to a \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which expresses the discrepancy \TODO{right word?} between $\hat\Sigma$ and $\Sigma$ and attains its minimum when $\hat\Sigma=\Sigma$.
+The optimization of a covariance matrix estimate $\hat\Sigma$ is performed with respect to a \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which expresses the discrepancy between $\hat\Sigma$ and $\Sigma$ and attains its minimum when $\hat\Sigma=\Sigma$.
 Then \emph{excess loss}  
@@
 
-The Gaussian loss function $\mathcal L_g$  arises from the theory of normal distributions.
+The Gaussian loss function $\mathcal L_g$  arises from the theory of multivariate normal distributions.
 When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{i=1}^n x_i x_i^\T$ is
@@
 Despite the fact that entropy loss is derived from normal theory, the choice of a loss function is not equivalent to assuming a specific form of $F$.
-The loss function measures the discrepancy between distribution parameters rather than the distance between distributions.
+The loss function measures the discrepancy between distribution parameters rather than the distance between the distributions themselves.
 \TODO{explain Bregman divergence?}
 
-\subsection{Emprical loss}
+The expected value of excess loss is the \emph{estimator risk}:
+\begin{equation}\label{eq:risk}
+r = \mathbb E\left[\ell(\hat\Sigma,\Sigma)\right]
+\end{equation}
+The estimator risk is the primary quality criterion for covariance estimation.
+Estimator $\hat\Sigma_a$ is considered more \emph{efficient} than estimator $\hat\Sigma_b$ if $\hat\Sigma_a$ has lower risk for the given sample than $\hat\Sigma_b$.
+\TODO{explain that the risk depends not only on $\hat\Sigma$ but also on the distribution of $\Sigma$ in the specific domain.}
 
-\subsection{Variance/bias decomposition}
-the variance-bias decomposition is the simple sum
+\subsection{Cross validation}
+In practice, the true covariance matrix $\Sigma$ is not accessible and the estimator risk $r$ must be estimated from data.
+This requires a separate \emph{validation} empirical distribution $\hat F_m^\prime$ of $m$ observations sampled from $F$ independently of the \emph{training} distribution $\hat F_n$.
+Then let  $\hat \Sigma_0^\prime$ be the sample covariance obtained from $\hat F_m^\prime$.
+
+Then the \emph{empirical loss} of $\hat\Sigma$ is $\mathcal L(\hat\Sigma,\hat\Sigma_0)$ and its expectation is the \emph{empirical risk}  
 \begin{equation}
-r_\gamma =  b_\gamma^2 + \varepsilon_\gamma^2
+\hat r = \mathbb E\left[\mathcal L(\hat\Sigma,\hat\Sigma_0^\prime)\right]
 \end{equation}
 
-Let $\hat\Sigma_\gamma$ be a family of approximate (biased) covariance estimates indexed by $\gamma$.
-To evaluate each estimate, let's define the \emph{loss function} $\mathcal L : \Theta \times \Theta \mapsto \mathbb R$ such that $\mathcal L(\hat S,S)$ is minimized when $\hat S = S$.
-Then \emph{excess loss} $\ell(\hat S, S) = \mathcal L(\hat S, S) - \mathcal L(S,S)$ assumes zero at the minimum.
-\TODO{This is too abstract -- need to motivate with some examples maybe}
+ The two loss functions $\mathcal L_e(\hat\Sigma,\Sigma)$ and $\mathcal L_g(\hat\Sigma,\Sigma)$ are particularly suitable for risk estimation thanks to their linearity with respect to $\Sigma$ in the sense that
+\begin{equation}
+\mathcal L\left(\hat\Sigma,\alpha S_1 + (1-\alpha)S_2\right) 
+\equiv 
+\alpha\mathcal L(\hat \Sigma,S_1) + (1-\alpha)\mathcal L(\hat \Sigma,S_2)
+\end{equation}
+which allows bringing the expectation inside the empirical loss function:
+\begin{equation}
+\hat r = 
+\mathbb E\left[ \mathcal L(\hat\Sigma, \hat\Sigma_0^\prime) \right] 
+=
+\mathbb E\left[ \mathcal L\left(\hat\Sigma, \mathbb E\left[\hat\Sigma_0^\prime\right]\right) \right] 
+=
+\mathbb E\left[ \mathcal L(\hat\Sigma, \Sigma) \right] 
+= 
+r + \mathcal L(\Sigma,\Sigma)
+\end{equation}
+This means, that the empirical loss $\mathcal L(\hat\Sigma,\hat\Sigma_0^\prime)$ is an unbiased estimate of the estimator risk $r$ (up to a constant offset): minimization of the empirical loss implies minimization of the true estimator risk.
 
-The overall error or \emph{estimator risk} 
+Recording a separate independent validation sample is usually not sensible.
+Instead, the recorded data are split into a training sample and validation sample.
+In $K$-fold \emph{cross-validation} the data are split into $K$ roughly equal-sized subsets.
+Each subset then successively serve as the testing sample while the remainder of the data serves as the training sample.
+This procedure produces $K$ estimates of the  risk, which are then averaged together to produce a better averaged estimate.
+
+\subsection{Bias/variance decomposition}
+The estimator risk (\autoref{eq:risk}) can be decomposed into \emph{approximation error} or \emph{bias}   
 \begin{equation}
-r_\gamma = \mathbb E\left[ \ell\left(\hat\Sigma_\gamma,\Sigma\right) \right]
-\end{equation}
-can be decomposed into \emph{approximation error} or \emph{bias}   
-\begin{equation}
-b_\gamma^2 = \ell \left( \mathbb E\left[\hat \Sigma_\gamma\right],\Sigma\right)
+b^2 = \ell \left( \mathbb E[\hat \Sigma],\Sigma\right)
 \end{equation}
@@
 \begin{equation}
-\varepsilon_\gamma^2 = \mathbb E \left[ \ell\left(\hat \Sigma_\gamma, 
-\mathbb E\left[\hat \Sigma_\gamma\right]\right) \right]
+\varepsilon^2 = \mathbb E \left[ \ell\left(\hat \Sigma, 
+\mathbb E[\hat \Sigma]\right) \right]
 \end{equation}
 
-With other loss functions, the risk is also an increasing function of the two components, although is not generally the sum.
+Under the MSE loss function $\mathcal L_e$, the decomposition is the simple sum $r = b^2 + \varepsilon^2$. 
+Under other loss functions, $r$ is an increasing function of both $b^2$ and $\varepsilon^2$ although not generally a simple sum.  
+\TODO{As far as I know, the terms `bias' and `variance' are only applicable under a quadratic loss function  like $\mathcal L_e$.
+Under other loss functions, we may need to use approximation/estimation error.}
 
-\subsection{Target estimates}
+The sample covariance matrix $\hat\Sigma_0$ has zero bias but high variance. Other estimators may be biased but have lower estimation error and potentially lower risk.
+
+\subsection{Regularized estimators}
+
+\emph{Regularization} is the deliberate biasing of the unbiased estimate toward a less variable low-dimensional \emph{target estimate} $\hat T$ in order to strike a favorable balance between bias and variance and minimize the estimator risk.
+A regularized covariance matrix estimator must solve two problems: (a) choose and fit the target covariate estimate $\hat T$ and (b) mix $\hat\Sigma_0$ with $\hat T$ by the optimal amount.
+Regularized covariance matrix estimates can be generally expressed as
+\begin{equation}
+\hat\Sigma_{d,\lambda} = mix(\Sigma_0,\hat T_{\hat d},\hat\lambda) 
+\end{equation}
+where $\hat d$ indicates the choice of the target estimate from the family of target estimates $T_d$ and $\hat \lambda$ regulates the amount biasing of $\hat\Sigma_0$ toward $\hat T$.
+Both $\hat d$ and $\hat \lambda$ must be calculated from the training sample.
+
 A good target estimate is one that has low dimensionality to reduce its variance while still allowing to accurately match typical covariance matrices in the given domain.
@@
 \paragraph{Evaluation of estimators:}
-
-In practice, the true covariance matrix $\Sigma$ is not available and the estimator risk must be estimated from data.
-This requires a separate \emph{validation} empirical distribution $\hat F_m^\prime$ of $m$ observations sampled from $F$ independently of the \emph{training} distribution $\hat F_n$.
-The we can produce an independent unbiased covariance estimate $\hat \Sigma_0(\hat F_m^\prime)$.
-
-Then the \emph{empirical risk} is 
-\begin{equation}
-\hat r_\gamma = \mathcal L\left(\hat\Sigma_\gamma,\hat\Sigma_0(\hat F_m^\prime)\right) 
-\end{equation}
-
- The two loss functions $\mathcal L_e(\hat S,S)$ and $\mathcal L_g(\hat S,S)$ are particularly suitable for empirical risk estimation thanks to their linearity with respect to $S$ in the sense that 
-\begin{equation}
-\mathcal L\left(\hat S,\alpha S_1 + (1-\alpha)S_2\right) 
-\equiv 
-\alpha\mathcal L(\hat S,S_1) + (1-\alpha)\mathcal L(\hat S,S_2)
-\end{equation}
-which allows bringing the expectation inside 
-\begin{equation}
-\mathbb E\left[ \mathcal L\left(\hat\Sigma_r, \hat\Sigma_0(\hat F_m^\prime)\right) \right] 
-=
-\mathbb E\left[ \mathcal L\left(\hat\Sigma_\gamma, \mathbb E\left[\Sigma_0(\hat F_m^\prime)\right]\right) \right] 
-=
-\mathbb E\left[ \mathcal L\left(\hat\Sigma_\gamma, \Sigma\right) \right] 
-\end{equation}
-Then the expectation of the empirical risk is 
-\begin{equation}
-\mathbb E\left[\hat r_\gamma\right] 
-= \mathbb E\left[\mathcal L(\hat\Sigma_\gamma,\Sigma)\right]
-= r_\gamma + \mathcal L(\Sigma,\Sigma)
-\end{equation}
-This means, that our goal of minimizing the true loss $r_\gamma$ can be replaced with minimizing  the empirical loss $\hat r_\gamma$.
-
-In practice, recording a separate independent validation sample may not be sensible.
-Instead, the recorded data a split into a training sample and validation sample.
-In $K$-fold \emph{cross-validation} the data are split into $K$ roughly equal-sized subsets.
-Each subset then successively serves to construct the validation distribution $\hat F_m^\prime$  while the rest of the data is used to construct the training distribution $\hat F_n$.
-This procedure produces $K$ estimates of the empirical risk, which are then averaged together to produce a better, averaged, estimate of the true risk.
-
-Different regularization schemes can then be compared by comparing cross-validated empirical losses.
 

### 2013-08-15 sections/Theory.tex
minor

@@
 
-\subsection{Loss function}
+\subsection{Loss functions and risk}
 The optimization of a covariance matrix estimate $\hat\Sigma$ is performed with respect to a \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which expresses the discrepancy between $\hat\Sigma$ and $\Sigma$ and attains its minimum when $\hat\Sigma=\Sigma$.
@@
 
-\subsection{Regularized estimators}
+\subsection{Regularization}
 
-\emph{Regularization} is the deliberate biasing of the unbiased estimate toward a less variable low-dimensional \emph{target estimate} $\hat T$ in order to strike a favorable balance between bias and variance and minimize the estimator risk.
+\emph{Regularization} is the deliberate biasing of the unbiased estimate toward a less variable low-dimensional \emph{target estimate} $\hat T$ in order to minimize the estimator risk by striking a favorable balance between bias and variance.
 A regularized covariance matrix estimator must solve two problems: (a) choose and fit the target covariate estimate $\hat T$ and (b) mix $\hat\Sigma_0$ with $\hat T$ by the optimal amount. 
@@
 \end{equation}
-where $\hat d$ indicates the choice of the target estimate from the family of target estimates $T_d$ and $\hat \lambda$ regulates the amount biasing of $\hat\Sigma_0$ toward $\hat T$.
+where $\hat d$ indicates the choice of the target estimate from the family of target estimates $\hat T_d$.
+$mix(\cdot,\cdot,\hat\lambda)$ is the mixing function of $\hat\Sigma_0$ and $\hat T_{\hat d}$ in proportion controlled by $\hat \lambda$.
 Both $\hat d$ and $\hat \lambda$ must be calculated from the training sample.
+
+A number of regularization schemes have been developed, which differ by their target estimates and mixing functions.  In this study, we evaluate four regularized estimators.
 

### 2013-08-15 sections/Theory.tex
edited the Theory section

@@
 
-\emph{Regularization} is the deliberate biasing of the unbiased estimate toward a less variable low-dimensional \emph{target estimate} $\hat T$ in order to minimize the estimator risk by striking a favorable balance between bias and variance.
+\emph{Regularization} is the deliberate biasing (`shrinkage') of the unbiased estimate toward a less variable low-dimensional \emph{target estimate} $\hat T$ in order to minimize the estimator risk by striking a favorable balance between bias and variance.
 A regularized covariance matrix estimator must solve two problems: (a) choose and fit the target covariate estimate $\hat T$ and (b) mix $\hat\Sigma_0$ with $\hat T$ by the optimal amount. 
@@
 
-A number of regularization schemes have been developed, which differ by their target estimates and mixing functions.  In this study, we evaluate four regularized estimators.
+A number of regularization schemes have been developed, which differ by their target estimates and mixing functions.
+Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of $\hat\Sigma_0$.
+The perplexing phenomenon that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.
+\TODO{This point can be untuitive.
+For example, \cite{Varoquaux:2012} misrepresents the effect of regularization by stating that regularization removes the upward bias of correlations in the sample covariance matrix.
+In reality, regularization should start with an unbiased estimator and bias it toward a low-variance target so that the estimation risk is reduced.} However, if a family of target estimates can capture the underlying regularities in the covariance structure of a specific system with a small number of parameters, biasing the estimate toward these targets ought to produce a greater reduction in estimator risk.
+
+In this study, we evaluate four regularized estimators whose target estimates correspond to the graphical models in Figure \ref{fig:02}: diagonal (A), multifactor (B), sparse partial correlations (C), and sparse  partial correlations with latent units (D).
+The following describe each regularization scheme separately.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{figures/Figure2.pdf}
+\caption{
+Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
+\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
+\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent. 
+\textbf{C}: In the Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from sparse pairwise linear interactions between visible units. 
+\textbf{D}: In the Gaussian graphical model with latent units, correlations arise  between pairs of nodes 
+}\label{fig:02}
+\end{figure}
+
+\subsection{Estimator A: Shrinkage toward a diagonal}
+The most popular covariance regularization schemes use diagonal target estimates and linear shrinkage toward the target.  For example, the target could be the identity matrix:  
+\begin{equation}
+\hat T = \hat v I
+\end{equation}
+where $\hat v = \frac 1 p \sum\limits_{i=1}^p(\hat\Sigma_0)_{ii}$ is the mean sample variance.
+
+Alternatively, the target could contain the sample variances:
+\begin{equation}
+\hat T= \hat\Sigma_0 \circ I 
+\end{equation}
+where $\circ$ is the entrywise matrix product.
+
+Finally, the target could be a linear mixture of the common variance and independent variance targets
+\begin{equation}
+\hat T_\eta = (1-\eta)(\hat\Sigma_0 \circ I) + \eta\hat v I
+\end{equation}
+
+The regularized estimator is the linear mixture of $\hat\Sigma_0$ and $\hat T_\eta$:
+\begin{equation}
+\hat\Sigma_{\eta,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda \hat T_\eta 
+\end{equation}
+
+\subsection{Estimator B: Shrinkage toward a factor model}
+
+\subsection{Estimator C: Sparse partial correlations}
+
+\subsection{Estimator D: Sparse partial correlations with latent units}
 
@@
 \end{enumerate}
-The first choice, the selection of the general form of target estimates, must be made \emph{a priori}.
-Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of the unbiased estimator.
-The perplexing fact that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.
 
@@
 Regularization is the game of minimizing risk $r_\gamma$ by reducing estimation error $\varepsilon_\gamma$ at the cost of increased bias $b_\gamma$, the \emph{bias-variance tradeoff}.  
-\TODO{This point can be untuitive.
-For example, \cite{Varoquaux:2012} misrepresents the effect of regularization by stating that regularization removes the upward bias of correlations in the sample covariance matrix.
-In reality, regularization should start with an unbiased estimator and bias it toward a low-variance target so that the estimation risk is reduced.
-Another important point is ``Stein's paradox'': regularization does not improve the estimate of each covariance.
-In fact, it makes it worse.
-Regularization only improves the risk when measured on the entire matrix.
-Regularization should not be used to improve the estimate of average correlations or the distribution of correlations.}
 We follow the convention that $\hat\Sigma_{\gamma=0}$ is the unbiased estimator $\hat\Sigma_0$, \emph{i.e.}\;$b_{\gamma=0}^2 = 0$. 
@@
 Then $\hat\Sigma_\gamma$ defines the \emph{regularization path}, for which \emph{regularization intensity} $\gamma$ specifies the location along the path, which may be real-valued or discrete, or even a vector to steer the bias toward multiple targets.
-
-\paragraph{Evaluation of estimators:}
 
@@
 Sparseness (dimensionality reduction) are key, but must be supported by data, not emph{ad hoc}. \citep{Fan:2006,Malmersjo:2013}.
-
-\paragraph{Our choices of regularized covariance estimates:}
-\begin{enumerate}[\qquad 1.\;\;]
-\item sample covariance (benchmark)
-\item linear shrinkage toward identity 
-\item linear shrinkage toward a single-factor model.
-\item multifactor
-\item graphical lasso
-\item latent-variable graphical lasso
-\end{enumerate}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{figures/Figure2.pdf}
-\caption{
-%\textbf{Figure 2.}{
-Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
-\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
-\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent. 
-\textbf{C}: In the Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from sparse pairwise linear interactions between visible units. 
-\textbf{D}: In the Gaussian graphical model with latent units, correlations arise  between pairs of nodes 
-}\label{fig:02}
-%}
-\end{figure}

### 2013-08-16 sections/Theory.tex
updated figure3

@@
 The activity of each neuron is denoted by the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
-For example, in our applicaiton, the firing rate is estimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x_i$ to take on negative values.
+For example, in our applicaiton, the firing rate is examplestimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x_i$ to take on negative values.
 
@@
 For independent observations, $c=1$;  $c>1$ when nearby samples are correlated; the value of $c$ can be estimated from the data.
+\TODO{Under the Gaussian loss, $\mathcal L_g$ (\autoref{eq:GaussLoss}) the scaling is irrelevant, so we will not spend any time talking about the scaling.
+But it would be important if we used $\mathcal L_e$.}
 
@@
 A particularly useful loss function is the mean squared error (MSE), which is proportional to the square of the Frobenius  norm $\|\cdot\|_F$ of the difference between the matrices: 
-\begin{equation}
+\begin{equation}\label{eq:MSE}
 \mathcal L_e(\hat\Sigma,\Sigma) =\frac 1 p \|\hat\Sigma-\Sigma\|_F^2 = \frac 1 p \Tr\left((\hat \Sigma-\Sigma)(\hat\Sigma-\Sigma)^\T\right)
@@
 Then $\mathcal L_g$ is constructed by rescaling $L\left(\Sigma \mid \hat\Sigma\right)$ and dropping the constant term:
-\begin{equation}
+\begin{equation}\label{eq:GaussLoss}
 \mathcal L_g(\hat\Sigma,\Sigma) 
@@
 
-Under the MSE loss function $\mathcal L_e$, the decomposition is the simple sum $r = b^2 + \varepsilon^2$. 
+Under the MSE loss function $\mathcal L_e$ (\autoref{eq:MSE}), the decomposition is the simple sum $r = b^2 + \varepsilon^2$. 
 Under other loss functions, $r$ is an increasing function of both $b^2$ and $\varepsilon^2$ although not generally a simple sum.  
@@
 
-\begin{figure}
+\begin{figure}[htp]
 \centering
-\includegraphics[width=0.6\textwidth]{figures/Figure2.pdf}
+\includegraphics[width=0.5\textwidth]{figures/Figure2.pdf}
 \caption{
@@
 
-\subsection{Estimator A: Shrinkage toward a diagonal}
+\subsection{Estimator A: Shrinkage toward diagonal}
 The most popular covariance regularization schemes use diagonal target estimates and linear shrinkage toward the target.  For example, the target could be the identity matrix:  
@@
 \begin{equation}
-\hat T_\eta = (1-\eta)(\hat\Sigma_0 \circ I) + \eta\hat v I
+\hat T_{\hat\eta} = (1-\hat\eta)(\hat\Sigma_0 \circ I) + \hat\eta\hat v I
 \end{equation}
 
-The regularized estimator is the linear mixture of $\hat\Sigma_0$ and $\hat T_\eta$:
+The regularized estimator is the linear mixture of $\hat\Sigma_0$ and $\hat T_{\hat\eta}$:
 \begin{equation}
-\hat\Sigma_{\eta,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda \hat T_\eta 
+\hat\Sigma_{\hat\eta,\hat\lambda} = (1-\hat\lambda)\hat\Sigma_0 + \hat\lambda \hat T_{\hat\eta} 
 \end{equation}
 
+The hyperparameters $\hat\eta$ and $\hat\lambda$ must be estimated from the data.
+Under the MSE loss function $\mathcal L_e$ (\autoref{eq:MSE}), the optimal values of $\hat\eta$ and $\hat\lambda$ can be estimated analytically \citep{Ledoit:2004,Schafer:2005,Schaefer:2010}.
+However, these estimates are no longer optimal under the Guassian loss $\mathcal L_g$.
+\TODO{I have tested this and there was a substantial difference in both synthetic and empirical data}  When an analytical solution for optimal hyperparameters is not available, the optimal values can be found by cross validation, entirely within the the training sample.
+\TODO{explain nested cross-validation in more detail?}
+
+\TODO{Describe how the empirical loss is convex in the hyperparameters in this case}
+
 \subsection{Estimator B: Shrinkage toward a factor model}
+In estimator B, the target is a factor model, composed as the sum of a low-rank component $\hat L_{\hat d}\hat L_{\hat d}^\T$ and a diagonal matrix $\hat \Psi$:
+\begin{equation}
+\hat T_{\hat d} = \hat L_{\hat d} \hat L_{\hat d}^\T + \hat\Psi
+\end{equation}
+Here $\hat L_{\hat d}$ is an $p\times\hat d$ matrix and $\hat\Psi$ is diagonal.
+
+If $F$ were assume to be multivariate normal \TODO{or other distributions that are defined by linear dependencies}, the factor model is corresponds to the graphical model (\autoref{fig:02}B) where the activity of all neurons are dependent only $\hat d$ latent units.
+This suggest a mechanistic interpretation in which the interactions between neurons are insignificant compared to the influence of inputs outside the recorded circuit.
+
+Just as in Estimator A, Estimator B is allowed to only partially commit to the low-dimensional target: the overall estimate is the linear mixture of the sample covariance and the target
+%\begin{equation}
+%\hat\Sigma_{\hat d,\hat\lambda} = (1-\hat\lambda)\hat\Sigma_0 + \hat  
+%\end{equation}
+
+\TODO{Check \citep{Fan:2011,Fan:2006}}
 
 \subsection{Estimator C: Sparse partial correlations}
+\citep{Meinshausen:2006,Friedman:2008}
 
 \subsection{Estimator D: Sparse partial correlations with latent units}
-
-A good target estimate is one that has low dimensionality to reduce its variance while still allowing to accurately match typical covariance matrices in the given domain.
-
-A regularized estimate requires the following three choices: 
-\begin{enumerate}[  1. ]
-\item define a set of low-dimensional target estimates
-\item given the data, select the optimal target estimate from the set 
-\item given the data, select the optimal amount of biasing toward the target estimate
-\end{enumerate}
-
-$\hat\Sigma_0$ is unbiased but, with finite sample sizes, not as close to $\Sigma$ as possible.
-More \emph{efficient} estimators can be constructed by \emph{regularization}, which is the deliberate biasing of the usual estmate toward a less variable \emph{target estimate} $\hat T$.
-Target estimates are low-dimensional approximations cannot represent an arbitrary $\Sigma$, but they also have lower random variability.
-Regularization reduces the overall error by striking a favorable balance between the variance  (\emph{estimation error}) and the bias \emph{(approximation error)} of the estimate.
-
-Regularization is the game of minimizing risk $r_\gamma$ by reducing estimation error $\varepsilon_\gamma$ at the cost of increased bias $b_\gamma$, the \emph{bias-variance tradeoff}.  
-We follow the convention that $\hat\Sigma_{\gamma=0}$ is the unbiased estimator $\hat\Sigma_0$, \emph{i.e.}\;$b_{\gamma=0}^2 = 0$. 
-We define a sequence of approximations $\hat\Sigma_\gamma$ that become less variable as as $\gamma$ increases, which is usually accompanies by increased bias.
-Then $\hat\Sigma_\gamma$ defines the \emph{regularization path}, for which \emph{regularization intensity} $\gamma$ specifies the location along the path, which may be real-valued or discrete, or even a vector to steer the bias toward multiple targets.
-
-\subsection{Regularization of covariance estimates}
-The task of designing a regularized estimators consists of defining a sequence of approximations $\hat\Sigma_\gamma$, called \emph{regularization path} and the optimal location on that path, \emph{regularization intensity} $\hat\gamma$ as  functions of the empirical distribution.
-
-There are two basic approaches to design a family of approximations of a parameter that reduce the estimation error: \emph{dimensionality reduction} and \emph{shrinkage}.
-
-Dimensionality reduction projects the estimate onto a low-dimensional manifold (subspace) within the $p(p+1)/2$-dimensional space of covariance matrices $\Theta$.
-The variance in the projection is likely to be lower (if the manifold is smooth).
-For example, factor analysis with $\nu$ factors projects (w.r.t.\;$\mathcal L_g$) $\hat\Sigma_0$ onto the low-dimensional manifold of covariance matrices representable by factor models.
-Thus factor analysis can be used as a regularized covariance estimator \citep{Fan:2008} with $\gamma$ controlling the number of factors (or, alternatively, the nuclear norm).
-Another dimensionality reduction technique is \emph{covariance selection} \citep{Dempster:1972}, wherein a fraction of inverse covariance coefficients are constrained to zero and the closest (w.r.t.\;$\mathcal L_g$) solution to $\hat\Sigma_0$ is found.
-
-Covariance shrinkage \citep{Schafer:2005,Ledoit:2004} uses a fixed, low-dimensional \emph{target estimate} $T$ and constructs the estimator as the linear mixture, with $\gamma \in [0,1]$,
-\begin{equation}
-\hat\Sigma_\gamma = \hat\Sigma_0  + \gamma (T-\hat\Sigma_0) 
-\end{equation}
- The target can be a diagonal matrix with sample variances, an identity matrix scaled to the mean variance, a single factor model, or any other low-dimensional estimate.
-
-An effective hybrid of dimensionality reduction and shrinkage can be obtained by using $L_1$-regularization techniques.
-For example, graphical lasso \citep{Meinshausen:2006,Friedman:2008}, shrinks the inverse of the covariance toward zero as
-\begin{equation}
-\hat\Sigma_\gamma = \arg\min\limits_S \mathcal L_g(S,\hat \Sigma_0) + \gamma \| S^{-1} \|_1
-\end{equation}
-This approach results in forcing many of the coefficients in the inverse of the estimate to zero, or ``sparsification'' of the the inverse.
-
-Many regularization schemes combine multiple targets or multiple approaches, in which case $\gamma$ can become a vector of regularization parameters \citep{Schafer:2005,Fan:2011,Ma:2013} to be optimized in parallel.
-
-\paragraph{Optimization of the regularization parameter:}
-Analytic solutions when exist \citep{Ledoit:2004,Schafer:2005}, nested cross-validation otherwise.
-
-\paragraph{Interpretation of revealed structure:}
-Sparseness (dimensionality reduction) are key, but must be supported by data, not emph{ad hoc}. \citep{Fan:2006,Malmersjo:2013}.
+\citep{Ma:2013}

### 2013-08-17 sections/Estimators.tex
expanded the section on covariance estimators

--- 2013-08-16 sections/Theory.tex
@@
-\subsection{Basic concepts}
-Let $x_i \in \mathcal X, i=1\ldots,n$ denote a sample of consecutive observations of population activity of $p$ neurons in time bins $i$.  
-The activity of each neuron is denoted by the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
-For example, in our applicaiton, the firing rate is examplestimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x_i$ to take on negative values.
-
-We do not assume that the observations $x_i$ are independently distributed.
-Rather, the data generating process is assumed to be \emph{ergodic}, \emph{i.e.}\;described by a \emph{true distirbution} $F: \mathcal X \mapsto  \mathbb [0, 1]$ (cumulative) over long periods of time such that the \emph{empirical distribution} $\hat F_n$ from a given sample will convege to $F$ with increasing sample size $n$.
-
-Formally, the emprical distribution is defined as \TODO{this may be unecessary, but I leave it for now.}
-\begin{equation}
-\hat F_n(x) = \frac 1 n \sum\limits_{i=1}^n \mathbf{1}(x \ge x_i)
-\end{equation}
-where $\mathbf 1(x \ge x_i)$ is the indicator function which equals 1 when all elements of $x$ are greater than the corresponding elements of $x_i$ and 0 otherwise.
-\TODO{KJ: Are you using cumulative distributions to avoid binning?
-DY: Yes.
-Many theorems such as Glivenko-Catelli Theorem and Skorohod's Representation are proven using  cumulative representation.
-The empirical distribution is nearly always represented in its cumulative form.
-Expressing convergence is requires binning or using cumulative; the latter is simpler.}
-
-And, for an ergodic process, $\max\limits_{x\in\mathcal X} \left|\hat F_n(x) - F(x)\right| \to 0$ as $n$ increases.
-
-The true covariance matrix $\Sigma \in \Theta$ is defined as a function of $F$:
-\begin{equation}
-\Sigma = \int\limits_{z\in\mathcal X} (z - \mu)(z - \mu)^\T \dif F(z)
-\quad\mbox{where}\;
-\mu = \int\limits_{z\in\mathcal X} z \dif F(z)
-\end{equation}
-The domain $\Theta$ is the set of all positive-definite $p\times p$ matrices, which is a cone in $\mathbb R^{p\times(p+1)/2}$.
-
-But the usual estimator of the true covariance matrix $\Sigma$ is  the sample covariance matrix:
-\begin{equation}
-\hat\Sigma_0 =  \frac 1 {n-c} \sum\limits_{i=1}^n (x_i -\hat\mu) (x_i - \hat\mu)^\T
-\quad\mbox{where}\;
-\hat\mu = \frac 1 n \sum\limits_{i=1}^n x_i
-\end{equation}
-where $c$ is the sum of temporal correlations \TODO{there must be a better description of this}.
-For independent observations, $c=1$;  $c>1$ when nearby samples are correlated; the value of $c$ can be estimated from the data.
-\TODO{Under the Gaussian loss, $\mathcal L_g$ (\autoref{eq:GaussLoss}) the scaling is irrelevant, so we will not spend any time talking about the scaling.
-But it would be important if we used $\mathcal L_e$.}
-
-$\hat\Sigma_0$ is designed to be unbiased such that $\mathbb E\left[\hat\Sigma_0\right]=\Sigma$.
-Here and througout, $\mathbb E[\cdot]$ denotes the expected value under the  unknown true distribution $F$.
-All variables with a hat (\emph{e.g.}\;$\hat \Sigma_0,\hat \mu$) are functions of the empirical distribution $\hat F_n$, which itself is a random variable.
-
-Nonlinear functions of an unbiased estimate can be biased: $\mathbb E\left[\hat\Sigma_0\right]=\Sigma \centernot\implies  E\left[\varphi(\hat\Sigma_0)\right]=\varphi(\Sigma)$.
-For example, if $u_{\max}(S)$ is the largest eigevalue of square matrix $S$, then $\mathbb E\left[u_{\max}(\hat\Sigma_0)\right] > u_{\max}(\Sigma)$.
-Some covariance matrix estimators are designed to correct for the eigenspectrum bias at the cost of adding bias to covariance coefficients \citep{Ledoit:2004}.
-
-\subsection{Loss functions and risk}
-The optimization of a covariance matrix estimate $\hat\Sigma$ is performed with respect to a \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which expresses the discrepancy between $\hat\Sigma$ and $\Sigma$ and attains its minimum when $\hat\Sigma=\Sigma$.
-Then \emph{excess loss}  
-\begin{equation}
-\ell(\hat\Sigma,\Sigma) = \mathcal L(\hat\Sigma,\Sigma)-\mathcal L(\Sigma,\Sigma)
-\end{equation}
-assumes zero at its minimum.
-
-A particularly useful loss function is the mean squared error (MSE), which is proportional to the square of the Frobenius  norm $\|\cdot\|_F$ of the difference between the matrices: 
-\begin{equation}\label{eq:MSE}
-\mathcal L_e(\hat\Sigma,\Sigma) =\frac 1 p \|\hat\Sigma-\Sigma\|_F^2 = \frac 1 p \Tr\left((\hat \Sigma-\Sigma)(\hat\Sigma-\Sigma)^\T\right)
-\end{equation}
-The MSE is its own excess loss: $\ell_e(\hat\Sigma,\Sigma) \equiv \mathcal L_e(\hat\Sigma,\Sigma)$.
-
-The Gaussian loss function $\mathcal L_g$  arises from the theory of multivariate normal distributions.
-When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{i=1}^n x_i x_i^\T$ is
-\begin{equation}
-L\left(\Sigma \mid \hat\Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
-\end{equation}
-Then $\mathcal L_g$ is constructed by rescaling $L\left(\Sigma \mid \hat\Sigma\right)$ and dropping the constant term:
-\begin{equation}\label{eq:GaussLoss}
-\mathcal L_g(\hat\Sigma,\Sigma) 
-=  -\frac 2 {pn} L\left(\Sigma \mid \hat\Sigma \right) - \frac 1 p \ln(2\pi) 
-\equiv  \frac 1 p\left(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}) \right) 
-\end{equation}
-The corresponding excess loss 
-\begin{equation}
-\ell_g(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) - \mathcal L_g(\Sigma,\Sigma)  
-= \frac 1 p \left(-\ln \det (\hat \Sigma^{-1} \Sigma) + \Tr(\hat \Sigma^{-1}\Sigma)\right) - 1
-\end{equation}
-is known as \emph{entropy loss} \citep{James:1961}.
-
-Despite the fact that entropy loss is derived from normal theory, the choice of a loss function is not equivalent to assuming a specific form of $F$.
-The loss function measures the discrepancy between distribution parameters rather than the distance between the distributions themselves.
-\TODO{explain Bregman divergence?}
-
-The expected value of excess loss is the \emph{estimator risk}:
-\begin{equation}\label{eq:risk}
-r = \mathbb E\left[\ell(\hat\Sigma,\Sigma)\right]
-\end{equation}
-The estimator risk is the primary quality criterion for covariance estimation.
-Estimator $\hat\Sigma_a$ is considered more \emph{efficient} than estimator $\hat\Sigma_b$ if $\hat\Sigma_a$ has lower risk for the given sample than $\hat\Sigma_b$.
-\TODO{explain that the risk depends not only on $\hat\Sigma$ but also on the distribution of $\Sigma$ in the specific domain.}
-
-\subsection{Cross validation}
-In practice, the true covariance matrix $\Sigma$ is not accessible and the estimator risk $r$ must be estimated from data.
-This requires a separate \emph{validation} empirical distribution $\hat F_m^\prime$ of $m$ observations sampled from $F$ independently of the \emph{training} distribution $\hat F_n$.
-Then let  $\hat \Sigma_0^\prime$ be the sample covariance obtained from $\hat F_m^\prime$.
-
-Then the \emph{empirical loss} of $\hat\Sigma$ is $\mathcal L(\hat\Sigma,\hat\Sigma_0)$ and its expectation is the \emph{empirical risk}  
-\begin{equation}
-\hat r = \mathbb E\left[\mathcal L(\hat\Sigma,\hat\Sigma_0^\prime)\right]
-\end{equation}
-
- The two loss functions $\mathcal L_e(\hat\Sigma,\Sigma)$ and $\mathcal L_g(\hat\Sigma,\Sigma)$ are particularly suitable for risk estimation thanks to their linearity with respect to $\Sigma$ in the sense that
-\begin{equation}
-\mathcal L\left(\hat\Sigma,\alpha S_1 + (1-\alpha)S_2\right) 
-\equiv 
-\alpha\mathcal L(\hat \Sigma,S_1) + (1-\alpha)\mathcal L(\hat \Sigma,S_2)
-\end{equation}
-which allows bringing the expectation inside the empirical loss function:
-\begin{equation}
-\hat r = 
-\mathbb E\left[ \mathcal L(\hat\Sigma, \hat\Sigma_0^\prime) \right] 
-=
-\mathbb E\left[ \mathcal L\left(\hat\Sigma, \mathbb E\left[\hat\Sigma_0^\prime\right]\right) \right] 
-=
-\mathbb E\left[ \mathcal L(\hat\Sigma, \Sigma) \right] 
-= 
-r + \mathcal L(\Sigma,\Sigma)
-\end{equation}
-This means, that the empirical loss $\mathcal L(\hat\Sigma,\hat\Sigma_0^\prime)$ is an unbiased estimate of the estimator risk $r$ (up to a constant offset): minimization of the empirical loss implies minimization of the true estimator risk.
-
-Recording a separate independent validation sample is usually not sensible.
-Instead, the recorded data are split into a training sample and validation sample.
-In $K$-fold \emph{cross-validation} the data are split into $K$ roughly equal-sized subsets.
-Each subset then successively serve as the testing sample while the remainder of the data serves as the training sample.
-This procedure produces $K$ estimates of the  risk, which are then averaged together to produce a better averaged estimate.
-
-\subsection{Bias/variance decomposition}
-The estimator risk (\autoref{eq:risk}) can be decomposed into \emph{approximation error} or \emph{bias}   
-\begin{equation}
-b^2 = \ell \left( \mathbb E[\hat \Sigma],\Sigma\right)
-\end{equation}
-and \emph{estimation error} or \emph{variance}
-\begin{equation}
-\varepsilon^2 = \mathbb E \left[ \ell\left(\hat \Sigma, 
-\mathbb E[\hat \Sigma]\right) \right]
-\end{equation}
-
-Under the MSE loss function $\mathcal L_e$ (\autoref{eq:MSE}), the decomposition is the simple sum $r = b^2 + \varepsilon^2$. 
-Under other loss functions, $r$ is an increasing function of both $b^2$ and $\varepsilon^2$ although not generally a simple sum.  
-\TODO{As far as I know, the terms `bias' and `variance' are only applicable under a quadratic loss function  like $\mathcal L_e$.
-Under other loss functions, we may need to use approximation/estimation error.}
-
-The sample covariance matrix $\hat\Sigma_0$ has zero bias but high variance. Other estimators may be biased but have lower estimation error and potentially lower risk.
-
-\subsection{Regularization}
-
-\emph{Regularization} is the deliberate biasing (`shrinkage') of the unbiased estimate toward a less variable low-dimensional \emph{target estimate} $\hat T$ in order to minimize the estimator risk by striking a favorable balance between bias and variance.
-A regularized covariance matrix estimator must solve two problems: (a) choose and fit the target covariate estimate $\hat T$ and (b) mix $\hat\Sigma_0$ with $\hat T$ by the optimal amount. 
-Regularized covariance matrix estimates can be generally expressed as 
-\begin{equation}
-\hat\Sigma_{d,\lambda} = mix(\Sigma_0,\hat T_{\hat d},\hat\lambda) 
-\end{equation}
-where $\hat d$ indicates the choice of the target estimate from the family of target estimates $\hat T_d$.
-$mix(\cdot,\cdot,\hat\lambda)$ is the mixing function of $\hat\Sigma_0$ and $\hat T_{\hat d}$ in proportion controlled by $\hat \lambda$.
-Both $\hat d$ and $\hat \lambda$ must be calculated from the training sample.
-
-A number of regularization schemes have been developed, which differ by their target estimates and mixing functions.
-Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of $\hat\Sigma_0$.
-The perplexing phenomenon that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.
-\TODO{This point can be untuitive.
-For example, \cite{Varoquaux:2012} misrepresents the effect of regularization by stating that regularization removes the upward bias of correlations in the sample covariance matrix.
-In reality, regularization should start with an unbiased estimator and bias it toward a low-variance target so that the estimation risk is reduced.} However, if a family of target estimates can capture the underlying regularities in the covariance structure of a specific system with a small number of parameters, biasing the estimate toward these targets ought to produce a greater reduction in estimator risk.
-
-In this study, we evaluate four regularized estimators whose target estimates correspond to the graphical models in Figure \ref{fig:02}: diagonal (A), multifactor (B), sparse partial correlations (C), and sparse  partial correlations with latent units (D).
-The following describe each regularization scheme separately.
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/Figure2.pdf}
-\caption{
-Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
-\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
-\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent. 
-\textbf{C}: In the Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from sparse pairwise linear interactions between visible units. 
-\textbf{D}: In the Gaussian graphical model with latent units, correlations arise  between pairs of nodes 
-}\label{fig:02}
-\end{figure}
-
 \subsection{Estimator A: Shrinkage toward diagonal}
@@
 
-Just as in Estimator A, Estimator B is allowed to only partially commit to the low-dimensional target: the overall estimate is the linear mixture of the sample covariance and the target
-%\begin{equation}
-%\hat\Sigma_{\hat d,\hat\lambda} = (1-\hat\lambda)\hat\Sigma_0 + \hat  
-%\end{equation}
+Just as in Estimator A, Estimator B is allowed to commit to the low-dimensional target only partially: the overall estimate is the linear mixture of the sample covariance and the target
+\begin{equation}
+\hat\Sigma_{\hat d,\hat\lambda} = (1-\hat\lambda)\hat\Sigma_0 + \hat  T_{\hat d}
+\end{equation}
 
-\TODO{Check \citep{Fan:2011,Fan:2006}}
+\TODO{Check \citep{Ledoit:2003,Fan:2011,Fan:2006}}
 
-\subsection{Estimator C: Sparse partial correlations}
-\citep{Meinshausen:2006,Friedman:2008}
+\subsection{Estimator C: Sparse inverse}
+Assuming that $x$ is distributed normally, the inverse of the covariance matrix or \emph{precision matrix} $K=\Sigma^{-1}$ has special significance: zeros in the precision matrix indicate conditional independence between the corresponding pairs.
+To see this, let $x=(x_1,\ldots,x_p)^\T \sim g\left(x\cond K\right)$
+\begin{equation}
+g\left(x\cond K\right) = \frac 1 {Z(K)} \exp\left(-\frac 1 2 \sum\limits_{i=1}^p\sum\limits_{j=1}^p  K_{ij} x_i x_j\right)
+\end{equation} 
+If $K_{12}\equiv K_{21} = 0$, then the joint distribution of $x_1$ and $x_2$, conditioned on $x_3=a_3,\ldots,x_p=a_p$ can be decomposed as product of independent distributions of $x_1$ and $x_2$: 
+\begin{equation}
+\begin{split}
+g\left( x_1, x_2 \cond K, x_3=a_3,\ldots,x_p=a_p\right) &=
+\frac 1 {Z(K)} \exp\left(-\frac 1 2 \sum\limits_{i=3}^p\sum\limits_{j=3}^p  K_{ij} a_i a_j \right)\times
+\\
+ &  \exp\left( -\frac 1 2\left( K_{11} x_1^2 +  x_1\sum\limits_{i=3}^p K_{1i}a_i \right)  \right)
+\\
+ &  \exp\left( -\frac 1 2\left( K_{22} x_2^2 +  x_2\sum\limits_{i=3}^p K_{2i}a_i \right)  \right)
+\end{split}
+\end{equation}
 
-\subsection{Estimator D: Sparse partial correlations with latent units}
+\citep{Dempster:1972,Meinshausen:2006,Friedman:2008}
+
+\subsection{Estimator D: Sparse inverse with latent units}
 \citep{Ma:2013}

### 2013-08-17 sections/Theory.tex
expanded the section on covariance estimators

@@
 \subsection{Basic concepts}
-Let $x_i \in \mathcal X, i=1\ldots,n$ denote a sample of consecutive observations of population activity of $p$ neurons in time bins $i$.  
+Let $x(t) \in \mathcal X, t=1\ldots,n$ denote a sample of consecutive observations of population activity of $p$ neurons in time bins $t$.  
 The activity of each neuron is denoted by the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
-For example, in our applicaiton, the firing rate is examplestimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x_i$ to take on negative values.
+For example, in our applicaiton, the firing rate is estimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x(t)$ to take on negative and non-integer values.
 
-We do not assume that the observations $x_i$ are independently distributed.
+We do not assume that the observations $x(t)$ are independently distributed.
 Rather, the data generating process is assumed to be \emph{ergodic}, \emph{i.e.}\;described by a \emph{true distirbution} $F: \mathcal X \mapsto  \mathbb [0, 1]$ (cumulative) over long periods of time such that the \emph{empirical distribution} $\hat F_n$ from a given sample will convege to $F$ with increasing sample size $n$.
@@
 \begin{equation}
-\hat F_n(x) = \frac 1 n \sum\limits_{i=1}^n \mathbf{1}(x \ge x_i)
+\hat F_n(x) = \frac 1 n \sum\limits_{t=1}^n \mathbf{1}(x \ge x(t))
 \end{equation}
-where $\mathbf 1(x \ge x_i)$ is the indicator function which equals 1 when all elements of $x$ are greater than the corresponding elements of $x_i$ and 0 otherwise.
+where $\mathbf 1(x \ge x(t))$ is the indicator function which equals 1 when all elements of $x$ are greater than the corresponding elements of $x(t)$ and 0 otherwise.
 \TODO{KJ: Are you using cumulative distributions to avoid binning?
@@
 \begin{equation}
-\hat\Sigma_0 =  \frac 1 {n-c} \sum\limits_{i=1}^n (x_i -\hat\mu) (x_i - \hat\mu)^\T
+\hat\Sigma_0 =  \frac 1 {n-c} \sum\limits_{t=1}^n (x(t) -\hat\mu) (x(t) - \hat\mu)^\T
 \quad\mbox{where}\;
-\hat\mu = \frac 1 n \sum\limits_{i=1}^n x_i
+\hat\mu = \frac 1 n \sum\limits_{t=1}^n x(t)
 \end{equation}
@@
 \TODO{Under the Gaussian loss, $\mathcal L_g$ (\autoref{eq:GaussLoss}) the scaling is irrelevant, so we will not spend any time talking about the scaling.
-But it would be important if we used $\mathcal L_e$.}
+But it would be important if we used $\mathcal L_e$ (\autoref{eq:MSE}).}
 
@@
 
-\subsection{Loss functions and risk}
+\subsection{Loss and risk}
 The optimization of a covariance matrix estimate $\hat\Sigma$ is performed with respect to a \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which expresses the discrepancy between $\hat\Sigma$ and $\Sigma$ and attains its minimum when $\hat\Sigma=\Sigma$.
@@
 
-A particularly useful loss function is the mean squared error (MSE), which is proportional to the square of the Frobenius  norm $\|\cdot\|_F$ of the difference between the matrices: 
+A particularly useful loss function is the mean squared error (MSE), which is proportional to the square of the Frobenius  norm $\|\cdot\|_F$ of the difference between the two matrices: 
 \begin{equation}\label{eq:MSE}
@@
 \end{equation}
-The MSE is its own excess loss: $\ell_e(\hat\Sigma,\Sigma) \equiv \mathcal L_e(\hat\Sigma,\Sigma)$.
+Since $\mathcal L_e(\Sigma,\Sigma) \equiv 0$, the MSE is its own excess loss: $\ell_e(\hat\Sigma,\Sigma) \equiv \mathcal L_e(\hat\Sigma,\Sigma)$.
 
 The Gaussian loss function $\mathcal L_g$  arises from the theory of multivariate normal distributions.
-When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{i=1}^n x_i x_i^\T$ is
+When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{t=1}^n x(t) x(t)^\T$ is
 \begin{equation}
@@
 \end{equation}
-is known as \emph{entropy loss} \citep{James:1961}.
+is known as \emph{entropy loss} \citep{James:1961}. \TODO{The order of the operands is reversed. Check that it's okay.}
 
@@
 \end{figure}
-
-\subsection{Estimator A: Shrinkage toward diagonal}
-The most popular covariance regularization schemes use diagonal target estimates and linear shrinkage toward the target.  For example, the target could be the identity matrix:  
-\begin{equation}
-\hat T = \hat v I
-\end{equation}
-where $\hat v = \frac 1 p \sum\limits_{i=1}^p(\hat\Sigma_0)_{ii}$ is the mean sample variance.
-
-Alternatively, the target could contain the sample variances:
-\begin{equation}
-\hat T= \hat\Sigma_0 \circ I 
-\end{equation}
-where $\circ$ is the entrywise matrix product.
-
-Finally, the target could be a linear mixture of the common variance and independent variance targets
-\begin{equation}
-\hat T_{\hat\eta} = (1-\hat\eta)(\hat\Sigma_0 \circ I) + \hat\eta\hat v I
-\end{equation}
-
-The regularized estimator is the linear mixture of $\hat\Sigma_0$ and $\hat T_{\hat\eta}$:
-\begin{equation}
-\hat\Sigma_{\hat\eta,\hat\lambda} = (1-\hat\lambda)\hat\Sigma_0 + \hat\lambda \hat T_{\hat\eta} 
-\end{equation}
-
-The hyperparameters $\hat\eta$ and $\hat\lambda$ must be estimated from the data.
-Under the MSE loss function $\mathcal L_e$ (\autoref{eq:MSE}), the optimal values of $\hat\eta$ and $\hat\lambda$ can be estimated analytically \citep{Ledoit:2004,Schafer:2005,Schaefer:2010}.
-However, these estimates are no longer optimal under the Guassian loss $\mathcal L_g$.
-\TODO{I have tested this and there was a substantial difference in both synthetic and empirical data}  When an analytical solution for optimal hyperparameters is not available, the optimal values can be found by cross validation, entirely within the the training sample.
-\TODO{explain nested cross-validation in more detail?}
-
-\TODO{Describe how the empirical loss is convex in the hyperparameters in this case}
-
-\subsection{Estimator B: Shrinkage toward a factor model}
-In estimator B, the target is a factor model, composed as the sum of a low-rank component $\hat L_{\hat d}\hat L_{\hat d}^\T$ and a diagonal matrix $\hat \Psi$:
-\begin{equation}
-\hat T_{\hat d} = \hat L_{\hat d} \hat L_{\hat d}^\T + \hat\Psi
-\end{equation}
-Here $\hat L_{\hat d}$ is an $p\times\hat d$ matrix and $\hat\Psi$ is diagonal.
-
-If $F$ were assume to be multivariate normal \TODO{or other distributions that are defined by linear dependencies}, the factor model is corresponds to the graphical model (\autoref{fig:02}B) where the activity of all neurons are dependent only $\hat d$ latent units.
-This suggest a mechanistic interpretation in which the interactions between neurons are insignificant compared to the influence of inputs outside the recorded circuit.
-
-Just as in Estimator A, Estimator B is allowed to only partially commit to the low-dimensional target: the overall estimate is the linear mixture of the sample covariance and the target
-%\begin{equation}
-%\hat\Sigma_{\hat d,\hat\lambda} = (1-\hat\lambda)\hat\Sigma_0 + \hat  
-%\end{equation}
-
-\TODO{Check \citep{Fan:2011,Fan:2006}}
-
-\subsection{Estimator C: Sparse partial correlations}
-\citep{Meinshausen:2006,Friedman:2008}
-
-\subsection{Estimator D: Sparse partial correlations with latent units}
-\citep{Ma:2013}

### 2013-08-19 plos.tex
changed the format to PLOS computational biology

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

\DeclareMathOperator{\Tr}{tr}
\newcommand{\mcond}{\,\middle\vert\,}
\newcommand{\cond}{\,\vert\,}
\newcommand{\T}{\intercal}
\newcommand{\TODO}[1]{\emph{\tiny\color{blue}$\langle\langle$#1$\rangle\rangle$}}
\newcommand*\dif{\mathop{}\,d}

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
Improved estimates of neural correlations suggest detailed interactions in visual cortex
}
% Insert Author names, affiliations and corresponding author email.
\\
Dimitri Yatsenko$^{1,\ast}$, 
Alexander Ecker$^{2,1}$,
Emmanouil Foudarakis$^{1}$,
R.~James Cotton$^{1}$,
Kre\v{s}imir Josi\'{c}$^{3}$,
Andreas S.~Tolias$^{1}$
\\
\bf{1} Department of Neuroscience, Baylor College of Medicine, Houston, TX, USA
\\
\bf{2} Max Planck Institute of Biological Cybernetics, T\"ubingen, Germany
\\
\bf{3} Department of Mathematics and Department of Biology and Biochemistry, University of Houston, Houston, TX, USA
\\
$\ast$ E-mail: yatsenko@cns.bcm.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
\input{sections/Abstract.tex}

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}
\input{sections/Introduction.tex}

% Results and Discussion can be combined.
\section*{Results}
\input{sections/Results.tex}

\subsection*{Subsection 1}

\subsection*{Subsection 2}

\section*{Discussion}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
\subsection*{Theory of covariance estimation}
\input{sections/Theory.tex}
\input{sections/Estimators.tex}
\input{sections/Simulation.tex}

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

%\section*{References}
% The bibtex filename
\bibliography{references.bib}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\end{document}

### 2013-08-19 sections/Estimators.tex
changed the format to PLOS computational biology

@@
-\subsection{Estimator A: Shrinkage toward diagonal}
+\subsection*{Estimator A: Shrinkage toward diagonal}
 The most popular covariance regularization schemes use diagonal target estimates and linear shrinkage toward the target.  For example, the target could be the identity matrix:  
@@
 The hyperparameters $\hat\eta$ and $\hat\lambda$ must be estimated from the data.
-Under the MSE loss function $\mathcal L_e$ (\autoref{eq:MSE}), the optimal values of $\hat\eta$ and $\hat\lambda$ can be estimated analytically \citep{Ledoit:2004,Schafer:2005,Schaefer:2010}.
+Under the MSE loss function $\mathcal L_e$ (\ref{eq:MSE}), the optimal values of $\hat\eta$ and $\hat\lambda$ can be estimated analytically \cite{Ledoit:2004,Schafer:2005,Schaefer:2010}.
 However, these estimates are no longer optimal under the Guassian loss $\mathcal L_g$.
@@
 
-\subsection{Estimator B: Shrinkage toward a factor model}
+\subsection*{Estimator B: Shrinkage toward a factor model}
 In estimator B, the target is a factor model, composed as the sum of a low-rank component $\hat L_{\hat d}\hat L_{\hat d}^\T$ and a diagonal matrix $\hat \Psi$:
@@
 
-If $F$ were assume to be multivariate normal \TODO{or other distributions that are defined by linear dependencies}, the factor model is corresponds to the graphical model (\autoref{fig:02}B) where the activity of all neurons are dependent only $\hat d$ latent units.
+If $F$ were assume to be multivariate normal \TODO{or other distributions that are defined by linear dependencies}, the factor model is corresponds to the graphical model (\ref{fig:02}B) where the activity of all neurons are dependent only $\hat d$ latent units.
 This suggest a mechanistic interpretation in which the interactions between neurons are insignificant compared to the influence of inputs outside the recorded circuit.
@@
 
-\TODO{Check \citep{Ledoit:2003,Fan:2011,Fan:2006}}
+\TODO{Check \cite{Ledoit:2003,Fan:2011,Fan:2006}}
 
-\subsection{Estimator C: Sparse inverse}
+\subsection*{Estimator C: Sparse inverse}
 Assuming that $x$ is distributed normally, the inverse of the covariance matrix or \emph{precision matrix} $K=\Sigma^{-1}$ has special significance: zeros in the precision matrix indicate conditional independence between the corresponding pairs.
@@
 
-\citep{Dempster:1972,Meinshausen:2006,Friedman:2008}
+\cite{Dempster:1972,Meinshausen:2006,Friedman:2008}
 
-\subsection{Estimator D: Sparse inverse with latent units}
-\citep{Ma:2013}
+\subsection*{Estimator D: Sparse inverse with latent units}
+\cite{Ma:2013}

### 2013-08-19 sections/Introduction.tex
changed the format to PLOS computational biology

@@
-Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \citep{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Renart:2010}.
+Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Renart:2010}.
 For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been of particular interest.
-Noise correlations can have profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}, and have been interpreted to indicate detailed and specific functional organization.
-Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \citep{Smith:2008,Denman:2013}, their synaptic connectivity and stimulus tuning similarity \citep{Kohn:2005,Ko:2011}, cortical layer specificity \citep{Hansen:2012,Smith:2013}, cell-type specificity (?), progressive changes in development and in learning \citep{Golshani:2009}, changes due to sensory stimulation and global brain states \citep{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
+Noise correlations can have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}, and have been interpreted to indicate detailed and specific functional organization.
+Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity and stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, cell-type specificity (?), progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
 
@@
 Theoretical and simulation studies have shown that neural correlations may arise at various temporal scales from combinations of multiple underlying mechanisms.
-These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
+These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \cite{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
 
-Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \citep{Shadlen:1998,Zohary:1994}.
+Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \cite{Shadlen:1998,Zohary:1994}.
 Multineuronal recordings allow estimation of covariance matrices of large populations of neurons.
@@
 Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically.
-For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \citep{Hayes:1981}.
+For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \cite{Hayes:1981}.
 Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.
@@
 With this study, we pursue two related goals: 
-\begin{enumerate}[\qquad A.\;\;]
+\begin{enumerate}
 	\item Devise more efficient estimators of neural covariance matrices in recordings from large populations of neurons.
@@
 To select the best covariance matrix estimate for a specific neural circuit, we evaluate the performance of four estimators using different regularization schemes, each motivated by  a different hypothesis for the origin of neural correlations.
-\emph{Regularization} is the deliberate biasing the solution toward a simplified, low-dimensional (`sparse') \emph{target estimate} \citep{Schafer:2005,Bickel:2006}.
+\emph{Regularization} is the deliberate biasing the solution toward a simplified, low-dimensional (`sparse') \emph{target estimate} \cite{Schafer:2005,Bickel:2006}.
 Regularized estimators allow reaching a favorable tradeoff between bias and variance in order to minimize the total error.
@@
 However, when a sparse target estimate does happen to match the dominant features of the true covariance matrix, it will induce relatively little bias while retaining low variance.
-For example, some covariance estimators have been motivated by a specific model of the underlying process such as single-factor models of the stock market, for example \citep{Ledoit:2003}.
+For example, some covariance estimators have been motivated by a specific model of the underlying process such as single-factor models of the stock market, for example \cite{Ledoit:2003}.
 Since the true covariance structure is not known in practice, the performance of estimators is evaluated by how well their performance generalized to new data, outside the training sample by cross validation, for example.

### 2013-08-19 sections/Theory.tex
changed the format to PLOS computational biology

@@
-\subsection{Basic concepts}
+\subsection*{Basic concepts}
 Let $x(t) \in \mathcal X, t=1\ldots,n$ denote a sample of consecutive observations of population activity of $p$ neurons in time bins $t$.  
@@
 For independent observations, $c=1$;  $c>1$ when nearby samples are correlated; the value of $c$ can be estimated from the data.
-\TODO{Under the Gaussian loss, $\mathcal L_g$ (\autoref{eq:GaussLoss}) the scaling is irrelevant, so we will not spend any time talking about the scaling.
-But it would be important if we used $\mathcal L_e$ (\autoref{eq:MSE}).}
+\TODO{Under the Gaussian loss, $\mathcal L_g$ (\ref{eq:GaussLoss}) the scaling is irrelevant, so we will not spend any time talking about the scaling.
+But it would be important if we used $\mathcal L_e$ (\ref{eq:MSE}).}
 
@@
 
-Nonlinear functions of an unbiased estimate can be biased: $\mathbb E\left[\hat\Sigma_0\right]=\Sigma \centernot\implies  E\left[\varphi(\hat\Sigma_0)\right]=\varphi(\Sigma)$.
+%Nonlinear functions of an unbiased estimate can be biased: $\mathbb E\left[\hat\Sigma_0\right]=\Sigma \centernot\implies  E\left[\varphi(\hat\Sigma_0)\right]=\varphi(\Sigma)$.
 For example, if $u_{\max}(S)$ is the largest eigevalue of square matrix $S$, then $\mathbb E\left[u_{\max}(\hat\Sigma_0)\right] > u_{\max}(\Sigma)$.
-Some covariance matrix estimators are designed to correct for the eigenspectrum bias at the cost of adding bias to covariance coefficients \citep{Ledoit:2004}.
+Some covariance matrix estimators are designed to correct for the eigenspectrum bias at the cost  of adding bias to covariance coefficients \cite{Ledoit:2004}.
 
-\subsection{Loss and risk}
+\subsection*{Loss and risk}
 The optimization of a covariance matrix estimate $\hat\Sigma$ is performed with respect to a \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which expresses the discrepancy between $\hat\Sigma$ and $\Sigma$ and attains its minimum when $\hat\Sigma=\Sigma$.
@@
 \end{equation}
-is known as \emph{entropy loss} \citep{James:1961}. \TODO{The order of the operands is reversed. Check that it's okay.}
+is known as \emph{entropy loss} \cite{James:1961}. \TODO{The order of the operands is reversed. Check that it's okay.}
 
@@
 
-\subsection{Cross validation}
+\subsection*{Cross validation}
 In practice, the true covariance matrix $\Sigma$ is not accessible and the estimator risk $r$ must be estimated from data.
@@
 
-\subsection{Bias/variance decomposition}
-The estimator risk (\autoref{eq:risk}) can be decomposed into \emph{approximation error} or \emph{bias}   
+\subsection*{Bias/variance decomposition}
+The estimator risk (\ref{eq:risk}) can be decomposed into \emph{approximation error} or \emph{bias}   
 \begin{equation}
@@
 
-Under the MSE loss function $\mathcal L_e$ (\autoref{eq:MSE}), the decomposition is the simple sum $r = b^2 + \varepsilon^2$. 
+Under the MSE loss function $\mathcal L_e$ (\ref{eq:MSE}), the decomposition is the simple sum $r = b^2 + \varepsilon^2$. 
 Under other loss functions, $r$ is an increasing function of both $b^2$ and $\varepsilon^2$ although not generally a simple sum.  
@@
 
-\subsection{Regularization}
+\subsection*{Regularization}
 
@@
 Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of $\hat\Sigma_0$.
-The perplexing phenomenon that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \citep{Efron:1977}.
+The perplexing phenomenon that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \cite{Efron:1977}.
 \TODO{This point can be untuitive.

### 2013-09-06 sections/Results.tex
included figure5.

@@
-\textbf{Figure 4.}{ 
-The fitting performance of regularized covariance estimators on two-photon recording. 
+\begin{figure}[htp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/placeholder.pdf}
+\caption{
+Cross-validated performance of covariance estimators in one two-photon recording.
 }\label{fig:04}
+\end{figure}
 
-\textbf{Figure 5.}{ 
-Fitting performance of regularized covariance estimators on 24 datasets two-photon recordinngs.
+\begin{figure}[htp]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/Figure5.pdf}
+\caption{
+Cross-validated comparisons of the five covariance estimators on 31 two-photon recordings. 
 }\label{fig:05}
+\end{figure}
 
-\textbf{Figure 6.}{
+\begin{figure}[htp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/placeholder}
+\caption{
 The depiction of the 3D arrangement of the non-zero partial correlations between pairs of neurons.
 }\label{fig:06}
+\end{figure}

### 2013-09-23 sections/Summary.tex
re-arranged sections according to PLOS guidelines

We estimated the covariance matrices of the calcium activity of populations of neurons in the primary visual cortex.  
In the effort to improve the estimate, we imposed several types of sparse structure of the covariance matrix and, using cross-validation, compared the performance for resulting estimators.

### 2013-09-23 sections/Estimators.tex
Kreso's comments

@@
 Under the MSE loss function $\mathcal L_e$ (\ref{eq:MSE}), the optimal values of $\hat\eta$ and $\hat\lambda$ can be estimated analytically \cite{Ledoit:2004,Schafer:2005,Schaefer:2010}.
-However, these estimates are no longer optimal under the Guassian loss $\mathcal L_g$.
+However, these estimates are no longer optimal under the Gaussian loss $\mathcal L_g$.
 \TODO{I have tested this and there was a substantial difference in both synthetic and empirical data}  When an analytical solution for optimal hyperparameters is not available, the optimal values can be found by cross validation, entirely within the the training sample.
-\TODO{explain nested cross-validation in more detail?}
+\TODO{explain nested cross-validation in more detail?} \TODO{KJ: Yes - see above.}
+
+\TODO{KJ: I am confused.  This suggests that you are getting different results using the two different loss functions.
+However, there was no argument yet that either gives us more relevant information about the underlying distribution
+if we do not assume Gaussianity, for instance.  I would argue that the Gaussian loss is better, but this is simply 
+because the Euclidean metric is not appropriate on the cone of positive semidefinite variances.  As the 
+discussion we have been having with Alex demonstrates, this is a point that needs to be explained clearly.}
 
@@
 \subsection*{Estimator B: Shrinkage toward a factor model}
-In estimator B, the target is a factor model, composed as the sum of a low-rank component $\hat L_{\hat d}\hat L_{\hat d}^\T$ and a diagonal matrix $\hat \Psi$:
+For estimator B, the target is a factor model, composed as the sum of a low-rank component $\hat L_{\hat d}\hat L_{\hat d}^\T$ and a diagonal matrix $\hat \Psi$:
 \begin{equation}
@@
 
-Just as in Estimator A, Estimator B is allowed to commit to the low-dimensional target only partially: the overall estimate is the linear mixture of the sample covariance and the target
+Just as with Estimator A, Estimator B is allowed to commit to the low-dimensional target only partially: the overall estimate is the linear mixture of the sample covariance and the target
 \begin{equation}

### 2013-09-23 sections/Theory.tex
minor corrections
Kreso's comments

@@
 The activity of each neuron is denoted by the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
-For example, in our applicaiton, the firing rate is estimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x(t)$ to take on negative and non-integer values.
+For example, in our application, the firing rate is estimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x(t)$ to take on negative and non-integer values.
 
 We do not assume that the observations $x(t)$ are independently distributed.
-Rather, the data generating process is assumed to be \emph{ergodic}, \emph{i.e.}\;described by a \emph{true distirbution} $F: \mathcal X \mapsto  \mathbb [0, 1]$ (cumulative) over long periods of time such that the \emph{empirical distribution} $\hat F_n$ from a given sample will convege to $F$ with increasing sample size $n$.
+Rather, the data generating process is assumed to be \emph{ergodic}, \emph{i.e.}\;described by a \emph{true distribution} $F: \mathcal X \mapsto  \mathbb [0, 1]$ (cumulative) over long periods of time such that the \emph{empirical distribution} $\hat F_n$ from a given sample will convege to $F$ with increasing sample size $n$.
 
@@
 
-And, for an ergodic process, $\max\limits_{x\in\mathcal X} \left|\hat F_n(x) - F(x)\right| \to 0$ as $n$ increases.
+For an ergodic process, $\max\limits_{x\in\mathcal X} \left|\hat F_n(x) - F(x)\right| \to 0$ as $n$ increases.
+\TODO{If you decide to leave this in, provide a reference to a standard book in the field.}
 
@@
 \end{equation}
-where $c$ is the sum of temporal correlations \TODO{there must be a better description of this}.
+where $c$ is the sum of temporal correlations \TODO{there must be a better description of this} \TODO{Which?
+I think the equation is OK.
+Do you mean $c$?
+This is standard, and again, I suggest just providing a reference.}.
 For independent observations, $c=1$;  $c>1$ when nearby samples are correlated; the value of $c$ can be estimated from the data.
@@
 \begin{equation}
-L\left(\Sigma \mid \hat\Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
+L\left(\Sigma \mid \Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
 \end{equation}
@@
 \mathcal L_g(\hat\Sigma,\Sigma) 
-=  -\frac 2 {pn} L\left(\Sigma \mid \hat\Sigma \right) - \frac 1 p \ln(2\pi) 
-\equiv  \frac 1 p\left(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}) \right) 
+=  -\frac 2 {pn} L\left(\hat\Sigma \mid \Sigma \right) - \frac 1 p \ln(2\pi) 
+\equiv  \frac 1 p\left(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma) \right) 
 \end{equation}
+\TODO{Note that the switch from $L\left(\Sigma \mid \hat\Sigma \right)$ to $L\left(\hat\Sigma \mid \Sigma \right)$ is on purpose, which will become clear in the next few steps.}
 The corresponding excess loss 

### 2013-09-23 sections/Results.tex
Added Author Summary

--- 2013-07-29 sections/Results.tex
@@
-\textbf{Figure 3.}{ 
-The fitting performance of regularized covariance estimators on two-photon recording. 
-}\label{fig:03}
+\subsection*{Data acquisition}
+We recorded the calcium activity  of dense populations of neurons in the supragranular layers of mouse primary visual cortex using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}   while presenting numerious repetitions of full-field drifting gratings (Fig.
+1A and 1B).
+This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue (200 x 200 x 100 $\mu$m) in layers 2/3 and 4.
+Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.
+1C and 1D).
+The average stimulus response was subtracted to remove stimulus covariance and the sample noise covariance matrix was computed (Fig.
+1E).
 
-\textbf{Figure 4.}{ 
-Fitting performance of regularized covariance estimators on 24 datasets two-photon recordinngs.
-}\label{fig:04}
+\subsection*{Covariance estimation}
+We aim to estimate the true covariance matrix $\Sigma = \mathbb E [(x-\mu)(x-\mu)^\T]$ , where $\mathbb E[\cdot]$ denotes expectation, $x$ denotes the $p\times 1$ vector of real-valued instanteous firing rates of $p$ neurons in a bin of duration $\Delta t$ and $\mu = \mathbb E[x]$.
+For more rigourous notation, definitions, and derivations, see Appendix.
 
-\textbf{Figure 5.}{
-The depiction of the 3D arrangement of the non-zero partial correlations between pairs of neurons.
-}\label{fig:04}
+The usual estimator of $\Sigma$ is the sample covariance matrix
+\begin{equation}
+\hat \Sigma_0 = \frac 1 {n-c} \sum\limits_{t=1}^T (x(t)-\mu)(x(t)-\mu)^\T 
+\end{equation}
+where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals.
+Bessel's correction $c$ makes the estimate unbiased such that $\mathbb E[\Sigma_0] = \Sigma$.
+When observations $x(t)$ can be assumed to be independent then $c=1$; but when observations are correlated $c>1$ may be estimated from the data.
+
+\subsubsection*{Regularization}
+Although $\hat\Sigma_0$ is unbiased, it is not necessarily as close as possible to $\Sigma$.
+Closer estimates can be produced by admitting a moderate amount of bias (\emph{approximation error}) in order to reduce the variance (\emph{estimation error}) of the estimate.
+
+Regularization is the deliberate biasing of the unbiased estimate toward a low-dimensional target estimate in order to strike a favorable balance between bias and variance and minimize the overall expected error.
+\cite{Bickel:2006,
+
+A great variety of regularized covariance estimators can be conceived, which consistently outperform the sample covariance estimate.
+Regularized estimators differ from each other by the following aspects:
+\begin{itemize}
+\item the choice of the family of the target estimate, 
+\item the choice of the specific member of the family of target estimates
+\item the choice of the path from the unbiased estimate toward the chosen target
+\item the distance along that path
+\end{itemize}
+Steps 1 and 2 produce a low-dimensional estimate. Steps 3 and 4 are often called shrinkage and allow deviations from (or distribution about) the low-dimensional estimates.
+
+We evaluated four regularized estimators which we will denote as A, B, C, and D.  Their target estimates are graphically depicted in Figure 2. 
+The green spheres represent the observed neurons. The light-colored balls represent latent units and the edges connecting them represent conditional dependencies.
+
+\subsubsection*{Estimate A: Shrinkage toward the independent model}
+In A, the target is a family of diagonal matrices 
+\begin{equation}
+\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I
+\end{equation}
+where $\circ$ is the entrywise product (Hadamard product).
+When $\alpha=0$, $\hat T_\alpha$ has the average variance along the diagonal.
+When $ \alpha=1$, $ \hat T_\alpha$ has empirical variances on the diagonal.
+
+If dependencies between neurons are strictly linear, then a diagonal covariance matrix represents the case where neurons are independent, with no interactions.
+
+The overall estimate $\hat\Sigma$ is found by linear mixing of $\hat\Sigma_0$ and $ \hat T_\alpha$ controlled by the mixing proportion $\lambda$:
+\begin{equation}
+\hat\Sigma = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha 
+\end{equation}
+
+The hyperparameters $ \alpha$ and $ \lambda$ can be optimally selected from the training data by nested cross-validation.
+
+\subsubsection*{Estimate B: Shrinkage toward }
+In estimator B, the target is the multifactor model $ \hat T_d$ with $ d$ latent units.
+When all dependencies are linear, factor models correspond to mutually independent neurons each influenced by $ d$ latent units.
+(This is  the conventional factor analysis).
+
+Just as in A, the overall estimate is found by linear mixing:
+\begin{equation}
+\hat\Sigma = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d
+\end{equation}
+
+\subsubsection*{Estimate C}
+In estimate C, the target $ \hat S_\alpha$ is a matrix whose inverse is <em>sparse</em>, with a large fraction of off-diagonal elements fixed at zero.
+The coefficient $ \alpha$ controls the sparsity in $ \hat S_\alpha$.
+
+In models with only linear dependencies, zeros in the inverse covariance indicate conditional independence.
+Thus the target estimate is depicted as the graphical model with sparse interneuronal interactions.
+
+In practice the target $\hat S_\alpha$ is found by an $ L_1$-norm optimization technique, which combines dimensionality reduction and shrinkage into a single computationally efficient algorithm.
+
+Just as in the previous estimators, the optimal value of the regularization parameter $ \alpha$ is determined from the data by nested cross-validation.
+
+\subsubsection*{Estimate D}
+In estimate D, the target $\hat R_{d,\alpha}$ is the matrix whose inverse is the sum of a sparse matrix $\hat S_\alpha$ and a low-rank matrix $\hat L_d$ of rank $d$.
+\begin{equation}
+\hat R_{d,\alpha} = (\hat S_\alpha + \hat L_d)^{-1}
+\end{equation}
+
+If all dependencies are linear, the estimate D describes a group of neurons where each interacts with a small number of latent units and a small fraction of the observed neurons.
+
+\subsection*{Sparse inverse covariance with a low-rank component dominates other estimators}
+I computed noise covariances of population activity from 31 sites.
+In each site, I compared the cross-validated performance of the covariance matrix using the Gaussian log-likelihood loss function between the trained estimate $ \hat\Sigma$ and the testing sample covariance $ \Sigma_0^\prime$:
+\begin{equation}
+\mathcal L(\hat\Sigma,\hat\Sigma_0^\prime) = 
+\frac 1 p\left( \ln \det \hat \Sigma + \mbox{tr}(\hat \Sigma^{-1}\hat\Sigma_0^\prime) \right) 
+\end{equation}
+
+I obtained the following in figure 3.
+
+This is the full matrix of all comparisons of the estimators.
+The plots represents the histograms of the differences of the loss function between the pair of estimators for all 31 sites.
+When the differences are positive, the first estimate in the title outperforms the second estimate.
+
+Of particular interest is the first row, which shows that all regularized estimates  outperformed the sample covariance estimate.
+Even more informative is the last column, which shows that estimate D (sparse+lowrank) significantly outperformed all other estimates.

### 2013-09-23 sections/Summary.tex
Added Author Summary

@@
-We estimated the covariance matrices of the calcium activity of populations of neurons in the primary visual cortex.  
-In the effort to improve the estimate, we imposed several types of sparse structure of the covariance matrix and, using cross-validation, compared the performance for resulting estimators.
+Covariance matrices of the activity of populations of neurons are useful descriptions of the functional organization of neural circuits with implications for stimulus coding and neural circuit architecture.
+Estimates of the covariance matrix can be dramatically improved by \emph{regularization}, i.e.\;the biasing of the usual estimate toward one of several possible low-dimensional approximations, the \emph{target estimate}.
+However, the amount of improvement depends on how closely the selected approximation reflects the dominant interactions in the circuit.
+In simulation, we demonstrated that no regularized estimate was universally superior to others: each performed best when the target estimate was the closest match to the structure of the data generating process.
+We then evaluated covariance matrix estimates on the calcium activity recorded \emph{in vivo} from large populations of closely spaced neurons  in mouse primary visual cortex, after subtracting the average stimulus response.
+Using cross-validation, we found that the best estimate was produced when the covariance matrix was approximated as that of a sparse Gaussian graphical models with a small number of latent units. 
+Thanks to the paucity of their parameters, these optimally estimated covariance matrices also help interpretation of the covariance structure and allow relating the functional organization of the neural circuit to its cytoarchitecture.

### 2013-09-24 plos.tex
included figure5.
minor corrections
Kreso's comments
Added Author Summary
minor edits

@@
 \newcommand{\T}{\intercal}
-\newcommand{\TODO}[1]{\emph{\tiny\color{blue}$\langle\langle$#1$\rangle\rangle$}}
+\newcommand{\TODO}[1]{\emph{\small\color{blue}$\langle\langle$#1$\rangle\rangle$}}
 \newcommand*\dif{\mathop{}\,d}
@@
 Dimitri Yatsenko$^{1,\ast}$, 
-Alexander Ecker$^{2,1}$,
+Kre\v{s}imir Josi\'{c}$^{2}$,
+Alexander Ecker$^{3,1}$,
 Emmanouil Foudarakis$^{1}$,
 R.~James Cotton$^{1}$,
-Kre\v{s}imir Josi\'{c}$^{3}$,
 Andreas S.~Tolias$^{1}$
@@
 \section*{Author Summary}
+\input{sections/Summary.tex}
 
@@
 
-\subsection*{Subsection 1}
-
-\subsection*{Subsection 2}
-
 \section*{Discussion}
+%\input{sections/Discussion.tex}
 
@@
 \section*{Materials and Methods}
-\subsection*{Theory of covariance estimation}
-\input{sections/Theory.tex}
-\input{sections/Estimators.tex}
-\input{sections/Simulation.tex}
 
@@
 \section*{Figure Legends}
-%\begin{figure}[!ht]
-%\begin{center}
-%%\includegraphics[width=4in]{figure_name.2.eps}
-%\end{center}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
-%should be left justified, as specified by the options to the caption 
-%package.
-%}
-%\label{Figure_label}
-%\end{figure}
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=4in]{figures/Figure1.pdf}
+\end{center}
+\caption{
+{\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\tt A.} Visual stimuli comprising full-field  
+}
+\label{Figure_label}
+\end{figure}
 
-\section*{Tables}
+%\section*{Tables}
 %\begin{table}[!ht]

### 2013-09-24 sections/Results.tex
minor edits

@@
 \begin{equation}
-\hat \Sigma_0 = \frac 1 {n-c} \sum\limits_{t=1}^T (x(t)-\mu)(x(t)-\mu)^\T 
+\hat \Sigma_0 = \frac 1 {n-c} \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
 \end{equation}
 where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals.
-Bessel's correction $c$ makes the estimate unbiased such that $\mathbb E[\Sigma_0] = \Sigma$.
+Bessel's correction $c$ makes the estimate unbiased such that $\mathbb E[\hat\Sigma_0] = \Sigma$.
 When observations $x(t)$ can be assumed to be independent then $c=1$; but when observations are correlated $c>1$ may be estimated from the data.
 
-\subsubsection*{Regularization}
-Although $\hat\Sigma_0$ is unbiased, it is not necessarily as close as possible to $\Sigma$.
+\subsubsection*{Regularized estimates}
+Although $\hat\Sigma_0$ is unbiased, it is not as close as possible to $\Sigma$.
 Closer estimates can be produced by admitting a moderate amount of bias (\emph{approximation error}) in order to reduce the variance (\emph{estimation error}) of the estimate.
-
-Regularization is the deliberate biasing of the unbiased estimate toward a low-dimensional target estimate in order to strike a favorable balance between bias and variance and minimize the overall expected error.
-\cite{Bickel:2006,
+\emph{Regularization} is the deliberate biasing of the estimate toward a low-dimensional target estimate in order to strike a favorable balance between bias and variance and minimize the overall expected error \cite{Bickel:2006,Ledoit:2004}.
 

### 2013-09-25 sections/Figure1.tex
minor

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{figures/Figure1.pdf}
\end{center}
\caption{
{\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\tt A.} Visual stimuli comprising brief (500 ms) presentatios of full-field drifting gratings separated by blank screens. {\tt B.} Two-photon fast 3D imaging of calcium signals in an awake mouse. {\tt C.} Deconvolved calcium signals. {\tt D.} The sample covariance matrix of neuronal calcium signals in 200 ms time bins. 
}
\label{Figure_label}
\end{figure}

### 2013-09-25 sections/Results.tex
minor

@@
 1A and 1B).
-This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue (200 x 200 x 100 $\mu$m) in layers 2/3 and 4.
+This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
 Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.
@@
 
+\input{sections/Figure1.tex}
+
 \subsection*{Covariance estimation}
-We aim to estimate the true covariance matrix $\Sigma = \mathbb E [(x-\mu)(x-\mu)^\T]$ , where $\mathbb E[\cdot]$ denotes expectation, $x$ denotes the $p\times 1$ vector of real-valued instanteous firing rates of $p$ neurons in a bin of duration $\Delta t$ and $\mu = \mathbb E[x]$.
+We aim to estimate the true covariance matrix $\Sigma = \mathbb E [(x-\mu)(x-\mu)^\T]$ , where $\mathbb E[\cdot]$ denotes expectation \TODO{under the true model}, $x$ denotes the $p\times 1$ vector of real-valued instanteous firing rates of $p$ neurons in a bin of duration $\Delta t$ and $\mu = \mathbb E[x]$.
 For more rigourous notation, definitions, and derivations, see Appendix.

### 2013-09-25 sections/Figure1.tex
ongoing edits

@@
-\begin{figure}[!ht]
+\begin{figure}[htp]
 \begin{center}
-\includegraphics[width=4in]{figures/Figure1.pdf}
+\includegraphics[width=2.5in]{figures/Figure1.pdf}
 \end{center}
 \caption{
-{\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\tt A.} Visual stimuli comprising brief (500 ms) presentatios of full-field drifting gratings separated by blank screens.
-{\tt B.} Two-photon fast 3D imaging of calcium signals in an awake mouse.
-{\tt C.} Deconvolved calcium signals.
-{\tt D.} The sample covariance matrix of neuronal calcium signals in 200 ms time bins.
+{\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\bf A.} Visual stimuli comprising brief (500 ms) presentatios of full-field drifting gratings separated by blank screens.
+{\bf B.} Two-photon fast 3D imaging of calcium signals in an awake mouse.
+{\bf C.} Deconvolved calcium signals.
+{\bf D.} The sample covariance matrix of neuronal calcium signals in 200 ms time bins.
 }

### 2013-09-25 sections/Figure2.tex
ongoing edits

--- 2013-08-19 sections/Theory.tex
@@
-\subsection*{Basic concepts}
-Let $x(t) \in \mathcal X, t=1\ldots,n$ denote a sample of consecutive observations of population activity of $p$ neurons in time bins $t$.  
-The activity of each neuron is denoted by the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
-For example, in our applicaiton, the firing rate is estimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x(t)$ to take on negative and non-integer values.
-
-We do not assume that the observations $x(t)$ are independently distributed.
-Rather, the data generating process is assumed to be \emph{ergodic}, \emph{i.e.}\;described by a \emph{true distirbution} $F: \mathcal X \mapsto  \mathbb [0, 1]$ (cumulative) over long periods of time such that the \emph{empirical distribution} $\hat F_n$ from a given sample will convege to $F$ with increasing sample size $n$.
-
-Formally, the emprical distribution is defined as \TODO{this may be unecessary, but I leave it for now.}
-\begin{equation}
-\hat F_n(x) = \frac 1 n \sum\limits_{t=1}^n \mathbf{1}(x \ge x(t))
-\end{equation}
-where $\mathbf 1(x \ge x(t))$ is the indicator function which equals 1 when all elements of $x$ are greater than the corresponding elements of $x(t)$ and 0 otherwise.
-\TODO{KJ: Are you using cumulative distributions to avoid binning?
-DY: Yes.
-Many theorems such as Glivenko-Catelli Theorem and Skorohod's Representation are proven using  cumulative representation.
-The empirical distribution is nearly always represented in its cumulative form.
-Expressing convergence is requires binning or using cumulative; the latter is simpler.}
-
-And, for an ergodic process, $\max\limits_{x\in\mathcal X} \left|\hat F_n(x) - F(x)\right| \to 0$ as $n$ increases.
-
-The true covariance matrix $\Sigma \in \Theta$ is defined as a function of $F$:
-\begin{equation}
-\Sigma = \int\limits_{z\in\mathcal X} (z - \mu)(z - \mu)^\T \dif F(z)
-\quad\mbox{where}\;
-\mu = \int\limits_{z\in\mathcal X} z \dif F(z)
-\end{equation}
-The domain $\Theta$ is the set of all positive-definite $p\times p$ matrices, which is a cone in $\mathbb R^{p\times(p+1)/2}$.
-
-But the usual estimator of the true covariance matrix $\Sigma$ is  the sample covariance matrix:
-\begin{equation}
-\hat\Sigma_0 =  \frac 1 {n-c} \sum\limits_{t=1}^n (x(t) -\hat\mu) (x(t) - \hat\mu)^\T
-\quad\mbox{where}\;
-\hat\mu = \frac 1 n \sum\limits_{t=1}^n x(t)
-\end{equation}
-where $c$ is the sum of temporal correlations \TODO{there must be a better description of this}.
-For independent observations, $c=1$;  $c>1$ when nearby samples are correlated; the value of $c$ can be estimated from the data.
-\TODO{Under the Gaussian loss, $\mathcal L_g$ (\ref{eq:GaussLoss}) the scaling is irrelevant, so we will not spend any time talking about the scaling.
-But it would be important if we used $\mathcal L_e$ (\ref{eq:MSE}).}
-
-$\hat\Sigma_0$ is designed to be unbiased such that $\mathbb E\left[\hat\Sigma_0\right]=\Sigma$.
-Here and througout, $\mathbb E[\cdot]$ denotes the expected value under the  unknown true distribution $F$.
-All variables with a hat (\emph{e.g.}\;$\hat \Sigma_0,\hat \mu$) are functions of the empirical distribution $\hat F_n$, which itself is a random variable.
-
-%Nonlinear functions of an unbiased estimate can be biased: $\mathbb E\left[\hat\Sigma_0\right]=\Sigma \centernot\implies  E\left[\varphi(\hat\Sigma_0)\right]=\varphi(\Sigma)$.
-For example, if $u_{\max}(S)$ is the largest eigevalue of square matrix $S$, then $\mathbb E\left[u_{\max}(\hat\Sigma_0)\right] > u_{\max}(\Sigma)$.
-Some covariance matrix estimators are designed to correct for the eigenspectrum bias at the cost  of adding bias to covariance coefficients \cite{Ledoit:2004}.
-
-\subsection*{Loss and risk}
-The optimization of a covariance matrix estimate $\hat\Sigma$ is performed with respect to a \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which expresses the discrepancy between $\hat\Sigma$ and $\Sigma$ and attains its minimum when $\hat\Sigma=\Sigma$.
-Then \emph{excess loss}  
-\begin{equation}
-\ell(\hat\Sigma,\Sigma) = \mathcal L(\hat\Sigma,\Sigma)-\mathcal L(\Sigma,\Sigma)
-\end{equation}
-assumes zero at its minimum.
-
-A particularly useful loss function is the mean squared error (MSE), which is proportional to the square of the Frobenius  norm $\|\cdot\|_F$ of the difference between the two matrices: 
-\begin{equation}\label{eq:MSE}
-\mathcal L_e(\hat\Sigma,\Sigma) =\frac 1 p \|\hat\Sigma-\Sigma\|_F^2 = \frac 1 p \Tr\left((\hat \Sigma-\Sigma)(\hat\Sigma-\Sigma)^\T\right)
-\end{equation}
-Since $\mathcal L_e(\Sigma,\Sigma) \equiv 0$, the MSE is its own excess loss: $\ell_e(\hat\Sigma,\Sigma) \equiv \mathcal L_e(\hat\Sigma,\Sigma)$.
-
-The Gaussian loss function $\mathcal L_g$  arises from the theory of multivariate normal distributions.
-When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{t=1}^n x(t) x(t)^\T$ is
-\begin{equation}
-L\left(\Sigma \mid \hat\Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
-\end{equation}
-Then $\mathcal L_g$ is constructed by rescaling $L\left(\Sigma \mid \hat\Sigma\right)$ and dropping the constant term:
-\begin{equation}\label{eq:GaussLoss}
-\mathcal L_g(\hat\Sigma,\Sigma) 
-=  -\frac 2 {pn} L\left(\Sigma \mid \hat\Sigma \right) - \frac 1 p \ln(2\pi) 
-\equiv  \frac 1 p\left(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}) \right) 
-\end{equation}
-The corresponding excess loss 
-\begin{equation}
-\ell_g(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) - \mathcal L_g(\Sigma,\Sigma)  
-= \frac 1 p \left(-\ln \det (\hat \Sigma^{-1} \Sigma) + \Tr(\hat \Sigma^{-1}\Sigma)\right) - 1
-\end{equation}
-is known as \emph{entropy loss} \cite{James:1961}. \TODO{The order of the operands is reversed. Check that it's okay.}
-
-Despite the fact that entropy loss is derived from normal theory, the choice of a loss function is not equivalent to assuming a specific form of $F$.
-The loss function measures the discrepancy between distribution parameters rather than the distance between the distributions themselves.
-\TODO{explain Bregman divergence?}
-
-The expected value of excess loss is the \emph{estimator risk}:
-\begin{equation}\label{eq:risk}
-r = \mathbb E\left[\ell(\hat\Sigma,\Sigma)\right]
-\end{equation}
-The estimator risk is the primary quality criterion for covariance estimation.
-Estimator $\hat\Sigma_a$ is considered more \emph{efficient} than estimator $\hat\Sigma_b$ if $\hat\Sigma_a$ has lower risk for the given sample than $\hat\Sigma_b$.
-\TODO{explain that the risk depends not only on $\hat\Sigma$ but also on the distribution of $\Sigma$ in the specific domain.}
-
-\subsection*{Cross validation}
-In practice, the true covariance matrix $\Sigma$ is not accessible and the estimator risk $r$ must be estimated from data.
-This requires a separate \emph{validation} empirical distribution $\hat F_m^\prime$ of $m$ observations sampled from $F$ independently of the \emph{training} distribution $\hat F_n$.
-Then let  $\hat \Sigma_0^\prime$ be the sample covariance obtained from $\hat F_m^\prime$.
-
-Then the \emph{empirical loss} of $\hat\Sigma$ is $\mathcal L(\hat\Sigma,\hat\Sigma_0)$ and its expectation is the \emph{empirical risk}  
-\begin{equation}
-\hat r = \mathbb E\left[\mathcal L(\hat\Sigma,\hat\Sigma_0^\prime)\right]
-\end{equation}
-
- The two loss functions $\mathcal L_e(\hat\Sigma,\Sigma)$ and $\mathcal L_g(\hat\Sigma,\Sigma)$ are particularly suitable for risk estimation thanks to their linearity with respect to $\Sigma$ in the sense that
-\begin{equation}
-\mathcal L\left(\hat\Sigma,\alpha S_1 + (1-\alpha)S_2\right) 
-\equiv 
-\alpha\mathcal L(\hat \Sigma,S_1) + (1-\alpha)\mathcal L(\hat \Sigma,S_2)
-\end{equation}
-which allows bringing the expectation inside the empirical loss function:
-\begin{equation}
-\hat r = 
-\mathbb E\left[ \mathcal L(\hat\Sigma, \hat\Sigma_0^\prime) \right] 
-=
-\mathbb E\left[ \mathcal L\left(\hat\Sigma, \mathbb E\left[\hat\Sigma_0^\prime\right]\right) \right] 
-=
-\mathbb E\left[ \mathcal L(\hat\Sigma, \Sigma) \right] 
-= 
-r + \mathcal L(\Sigma,\Sigma)
-\end{equation}
-This means, that the empirical loss $\mathcal L(\hat\Sigma,\hat\Sigma_0^\prime)$ is an unbiased estimate of the estimator risk $r$ (up to a constant offset): minimization of the empirical loss implies minimization of the true estimator risk.
-
-Recording a separate independent validation sample is usually not sensible.
-Instead, the recorded data are split into a training sample and validation sample.
-In $K$-fold \emph{cross-validation} the data are split into $K$ roughly equal-sized subsets.
-Each subset then successively serve as the testing sample while the remainder of the data serves as the training sample.
-This procedure produces $K$ estimates of the  risk, which are then averaged together to produce a better averaged estimate.
-
-\subsection*{Bias/variance decomposition}
-The estimator risk (\ref{eq:risk}) can be decomposed into \emph{approximation error} or \emph{bias}   
-\begin{equation}
-b^2 = \ell \left( \mathbb E[\hat \Sigma],\Sigma\right)
-\end{equation}
-and \emph{estimation error} or \emph{variance}
-\begin{equation}
-\varepsilon^2 = \mathbb E \left[ \ell\left(\hat \Sigma, 
-\mathbb E[\hat \Sigma]\right) \right]
-\end{equation}
-
-Under the MSE loss function $\mathcal L_e$ (\ref{eq:MSE}), the decomposition is the simple sum $r = b^2 + \varepsilon^2$. 
-Under other loss functions, $r$ is an increasing function of both $b^2$ and $\varepsilon^2$ although not generally a simple sum.  
-\TODO{As far as I know, the terms `bias' and `variance' are only applicable under a quadratic loss function  like $\mathcal L_e$.
-Under other loss functions, we may need to use approximation/estimation error.}
-
-The sample covariance matrix $\hat\Sigma_0$ has zero bias but high variance. Other estimators may be biased but have lower estimation error and potentially lower risk.
-
-\subsection*{Regularization}
-
-\emph{Regularization} is the deliberate biasing (`shrinkage') of the unbiased estimate toward a less variable low-dimensional \emph{target estimate} $\hat T$ in order to minimize the estimator risk by striking a favorable balance between bias and variance.
-A regularized covariance matrix estimator must solve two problems: (a) choose and fit the target covariate estimate $\hat T$ and (b) mix $\hat\Sigma_0$ with $\hat T$ by the optimal amount. 
-Regularized covariance matrix estimates can be generally expressed as 
-\begin{equation}
-\hat\Sigma_{d,\lambda} = mix(\Sigma_0,\hat T_{\hat d},\hat\lambda) 
-\end{equation}
-where $\hat d$ indicates the choice of the target estimate from the family of target estimates $\hat T_d$.
-$mix(\cdot,\cdot,\hat\lambda)$ is the mixing function of $\hat\Sigma_0$ and $\hat T_{\hat d}$ in proportion controlled by $\hat \lambda$.
-Both $\hat d$ and $\hat \lambda$ must be calculated from the training sample.
-
-A number of regularization schemes have been developed, which differ by their target estimates and mixing functions.
-Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of $\hat\Sigma_0$.
-The perplexing phenomenon that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \cite{Efron:1977}.
-\TODO{This point can be untuitive.
-For example, \cite{Varoquaux:2012} misrepresents the effect of regularization by stating that regularization removes the upward bias of correlations in the sample covariance matrix.
-In reality, regularization should start with an unbiased estimator and bias it toward a low-variance target so that the estimation risk is reduced.} However, if a family of target estimates can capture the underlying regularities in the covariance structure of a specific system with a small number of parameters, biasing the estimate toward these targets ought to produce a greater reduction in estimator risk.
-
-In this study, we evaluate four regularized estimators whose target estimates correspond to the graphical models in Figure \ref{fig:02}: diagonal (A), multifactor (B), sparse partial correlations (C), and sparse  partial correlations with latent units (D).
-The following describe each regularization scheme separately.
-
 \begin{figure}[htp]

### 2013-09-25 sections/Results.tex
ongoing edits

@@
 \subsection*{Data acquisition}
-We recorded the calcium activity  of dense populations of neurons in the supragranular layers of mouse primary visual cortex using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}   while presenting numerious repetitions of full-field drifting gratings (Fig.
-1A and 1B).
+We recorded the calcium activity  of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}.
+Numerious repetitions of full-field drifting gratings (Fig.
+1A and 1B) were presented to the eye contralateral to the imaged site.
 This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
@@
 
-\subsection*{Covariance estimation}
+\subsection*{Sample covariance matrix}
 We aim to estimate the true covariance matrix $\Sigma = \mathbb E [(x-\mu)(x-\mu)^\T]$ , where $\mathbb E[\cdot]$ denotes expectation \TODO{under the true model}, $x$ denotes the $p\times 1$ vector of real-valued instanteous firing rates of $p$ neurons in a bin of duration $\Delta t$ and $\mu = \mathbb E[x]$.
@@
 
-\subsubsection*{Regularized estimates}
-Although $\hat\Sigma_0$ is unbiased, it is not as close as possible to $\Sigma$.
-Closer estimates can be produced by admitting a moderate amount of bias (\emph{approximation error}) in order to reduce the variance (\emph{estimation error}) of the estimate.
-\emph{Regularization} is the deliberate biasing of the estimate toward a low-dimensional target estimate in order to strike a favorable balance between bias and variance and minimize the overall expected error \cite{Bickel:2006,Ledoit:2004}.
+\subsection*{Regularized estimates of the covariance matrix}
+To assess the quality of a covariance matrix estimate $\hat\Sigma$, we must define a real-valued \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which attains its minimum when $\hat\Sigma = \Sigma$.
+For the purposes of this study, we adopted the \emph{Gaussian loss} function:
+\begin{equation}
+\mathcal L(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) = \frac 1 p(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma))
+\end{equation}
+which resembles the negative multivariate normal log-likelihood (see Appendix).
 
-A great variety of regularized covariance estimators can be conceived, which consistently outperform the sample covariance estimate.
-Regularized estimators differ from each other by the following aspects:
-\begin{itemize}
-\item the choice of the family of the target estimate, 
-\item the choice of the specific member of the family of target estimates
-\item the choice of the path from the unbiased estimate toward the chosen target
-\item the distance along that path
-\end{itemize}
-Steps 1 and 2 produce a low-dimensional estimate. Steps 3 and 4 are often called shrinkage and allow deviations from (or distribution about) the low-dimensional estimates.
+Although the sample covariance matrix is unbiased: $\mathbb E[\hat\Sigma_0]=\Sigma$, it is not as close to $\Sigma$ as possible, on average.
+We must seek  to minimize the expected value of the loss function, known as \emph{estimator risk}
+\begin{equation}
+r = \mathbb E[\mathcal L(\hat\Sigma, \Sigma)]
+\end{equation}
 
-We evaluated four regularized estimators which we will denote as A, B, C, and D.  Their target estimates are graphically depicted in Figure 2. 
-The green spheres represent the observed neurons. The light-colored balls represent latent units and the edges connecting them represent conditional dependencies.
+Estimator risk can be reduced by \emph{regularization}.
+Regularization is the deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
+Resulting regularized estimators are biased but  less variable.
+Regularization procedures must strike favorable balance between bias and variance so that the estimator risk is minimized.
+The terms \emph{bias} and \emph{variance} are applicable only the loss function is mean-squared-error.
+The more general terms are \emph{approximation error} and \emph{estimation error})
+
+Numerous regularized covariance matrix estimates have been devised.
+They start with the sample covariance matrix but differ in the choice of the target estimate and the shrinkage path and intensity.
+Some regularization schemes focus on the optimal choice of the low-dimensional target estimate from a large family of target estimates and do not require shrinkages.
+Others have a fairly simple target estimate but instead adjust the shrinkage intensity to reduce the estimation error.
+
+A regularized estimate of $\Sigma$ can be constructed by either (a) selecting an optimal low-dimensional approximation of $\hat\Sigma_0$ from a given family of approximations and/or (b) biasing or \emph{shrinkage} of $\hat\Sigma_0$ toward a given low-dimensional target estimate by an optimal amount.
+Both dimensionality reduction and shrinkage approaches reduce the variance of the solution while introducing bias.
+
+We evaluated four regularized estimators which we will denote as A, B, C, and D.
+Their target estimates correspond to covariance matrices of Gaussian graphical models depicted in Figure 2.
+Here the green spheres represent the recorded neurons.
+The light-colored balls represent latent units of the graphical model and the edges connecting them represent conditional dependencies.
+
+\input{sections/Figure2.tex}
+
+A \emph{graphical model} is a multivarate probability distribution with a specified graph of conditional dependencies between its variables \cite{Koller:2009}.
+When such a distribution is a multivariate normal distribution, the model becomes a \emph{Gaussian graphical model} or, equivalently, a \emph{Gaussian Markov Random Field}.
+Gaussian graphical models have a straightforward relationship to their covariance matrix $\Sigma$:  Zero elements in the inverse of $\Sigma$ indicate conditional independence between the corresponding pair of variables.
+
+Fitting a graphical model to data involves two tasks: (1) the selection of the set of non-zero covariances or \emph{covariance selection} \cite{Dempster:1972} and (b) the fitting of the non-zero elements.
 
 \subsubsection*{Estimate A: Shrinkage toward the independent model}
-In A, the target is a family of diagonal matrices 
+In estimate A, the target is a family of diagonal matrices 
 \begin{equation}
@@
 \subsubsection*{Estimate C}
-In estimate C, the target $ \hat S_\alpha$ is a matrix whose inverse is <em>sparse</em>, with a large fraction of off-diagonal elements fixed at zero.
+In estimate C, the target $ \hat S_\alpha$ is a matrix whose inverse is \emph{sparse}, with a large fraction of off-diagonal elements fixed at zero.
 The coefficient $ \alpha$ controls the sparsity in $ \hat S_\alpha$.

### 2013-09-25 sections/Theory.tex
ongoing edits

@@
 \begin{equation}
-L\left(\Sigma \mid \Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
+L\left(\Sigma \mid \hat\Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
 \end{equation}
@@
 The sample covariance matrix $\hat\Sigma_0$ has zero bias but high variance. Other estimators may be biased but have lower estimation error and potentially lower risk.
-
-\subsection*{Regularization}
-
-\emph{Regularization} is the deliberate biasing (`shrinkage') of the unbiased estimate toward a less variable low-dimensional \emph{target estimate} $\hat T$ in order to minimize the estimator risk by striking a favorable balance between bias and variance.
-A regularized covariance matrix estimator must solve two problems: (a) choose and fit the target covariate estimate $\hat T$ and (b) mix $\hat\Sigma_0$ with $\hat T$ by the optimal amount. 
-Regularized covariance matrix estimates can be generally expressed as 
-\begin{equation}
-\hat\Sigma_{d,\lambda} = mix(\Sigma_0,\hat T_{\hat d},\hat\lambda) 
-\end{equation}
-where $\hat d$ indicates the choice of the target estimate from the family of target estimates $\hat T_d$.
-$mix(\cdot,\cdot,\hat\lambda)$ is the mixing function of $\hat\Sigma_0$ and $\hat T_{\hat d}$ in proportion controlled by $\hat \lambda$.
-Both $\hat d$ and $\hat \lambda$ must be calculated from the training sample.
-
-A number of regularization schemes have been developed, which differ by their target estimates and mixing functions.
-Curiously, \emph{some} improvement can be produced with an arbitrary target estimate as long as its variance is lower than that of $\hat\Sigma_0$.
-The perplexing phenomenon that estimates can be improved by a bias toward an arbitrary less variable target is known as \emph{Stein's paradox} \cite{Efron:1977}.
-\TODO{This point can be untuitive.
-For example, \cite{Varoquaux:2012} misrepresents the effect of regularization by stating that regularization removes the upward bias of correlations in the sample covariance matrix.
-In reality, regularization should start with an unbiased estimator and bias it toward a low-variance target so that the estimation risk is reduced.} However, if a family of target estimates can capture the underlying regularities in the covariance structure of a specific system with a small number of parameters, biasing the estimate toward these targets ought to produce a greater reduction in estimator risk.
-
-In this study, we evaluate four regularized estimators whose target estimates correspond to the graphical models in Figure \ref{fig:02}: diagonal (A), multifactor (B), sparse partial correlations (C), and sparse  partial correlations with latent units (D).
-The following describe each regularization scheme separately.
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/Figure2.pdf}
-\caption{
-Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
-\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
-\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent. 
-\textbf{C}: In the Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from sparse pairwise linear interactions between visible units. 
-\textbf{D}: In the Gaussian graphical model with latent units, correlations arise  between pairs of nodes 
-}\label{fig:02}
-\end{figure}

### 2013-09-26 sections/Results.tex

@@
 To assess the quality of a covariance matrix estimate $\hat\Sigma$, we must define a real-valued \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which attains its minimum when $\hat\Sigma = \Sigma$.
-For the purposes of this study, we adopted the \emph{Gaussian loss} function:
+For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function (see Appendix for details):
 \begin{equation}
-\mathcal L(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) = \frac 1 p(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma))
+\mathcal L(\hat\Sigma,\Sigma) = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
 \end{equation}
-which resembles the negative multivariate normal log-likelihood (see Appendix).
 
-Although the sample covariance matrix is unbiased: $\mathbb E[\hat\Sigma_0]=\Sigma$, it is not as close to $\Sigma$ as possible, on average.
-We must seek  to minimize the expected value of the loss function, known as \emph{estimator risk}
+We seek an estimate of the covariance matrix that minimizes the expected value of loss, referred to as \emph{estimator risk}: 
 \begin{equation}
@@
 \end{equation}
+
+The sample covariance matrix is unbiased: it minimizes the bias $\mathcal L (\mathbb E[\hat\Sigma],\Sigma)$ but it does
 

### 2013-09-28 sections/Results.tex
ongoing changes in the Results section

@@
 This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
-Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.
-1C and 1D).
-The average stimulus response was subtracted to remove stimulus covariance and the sample noise covariance matrix was computed (Fig.
-1E).
+Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.\;1C and 1D).
+The average stimulus response was subtracted to remove stimulus covariance and the sample noise covariance matrix was computed (Fig.\;1E).
 
@@
 
-\subsection*{Sample covariance matrix}
-We aim to estimate the true covariance matrix $\Sigma = \mathbb E [(x-\mu)(x-\mu)^\T]$ , where $\mathbb E[\cdot]$ denotes expectation \TODO{under the true model}, $x$ denotes the $p\times 1$ vector of real-valued instanteous firing rates of $p$ neurons in a bin of duration $\Delta t$ and $\mu = \mathbb E[x]$.
-For more rigourous notation, definitions, and derivations, see Appendix.
+\subsection*{The sample covariance matrix}
+We aim to estimate the true covariance matrix $\Sigma = \E{(x-\mu)(x-\mu)^\T}$ of the instantaneous activity vector $x$ of a population of $p$ neurons.
+Here $\E{\cdot}$ denotes expectation \TODO{for the true data generating process} and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.
+
+For more rigorous notation, definitions, and derivations, see Appendix.
 
@@
 where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals.
-Bessel's correction $c$ makes the estimate unbiased such that $\mathbb E[\hat\Sigma_0] = \Sigma$.
+Bessel's correction $c$ makes the estimate unbiased such that $\E{\hat\Sigma_0} = \Sigma$.
 When observations $x(t)$ can be assumed to be independent then $c=1$; but when observations are correlated $c>1$ may be estimated from the data.
@@
 \subsection*{Regularized estimates of the covariance matrix}
-To assess the quality of a covariance matrix estimate $\hat\Sigma$, we must define a real-valued \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which attains its minimum when $\hat\Sigma = \Sigma$.
-For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function (see Appendix for details):
+The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.
+The loss function quantifies the discrepancy between $\hat\Sigma$ and $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.
+
+For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function:
+\begin{equation}\label{eq:loss}
+\loss{\hat\Sigma,\Sigma} = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
+\end{equation}
+This choice is motivated by mathematical convenience and, we assert, the conclusions of the study are unlikely to change drastically under other popular loss functions such as quadratic loss or Stein's entropy loss.
+
+The aim of our project is produce covariance matrix estimates that minimize the expected loss.  The expected value of the loss function of an estimator is known as its \emph{risk}: 
 \begin{equation}
-\mathcal L(\hat\Sigma,\Sigma) = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
+r = \E{\loss{\hat\Sigma, \Sigma}}
 \end{equation}
 
-We seek an estimate of the covariance matrix that minimizes the expected value of loss, referred to as \emph{estimator risk}: 
+Under many loss functions\footnote{
+The strict equality in Eq.~\ref{eq:bias-variance} does not hold for our choice of the loss function in Eq.~\ref{eq:loss}.
+However, the equality does hold for its close cousin, Stein's entropy loss,  $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$, which only differs by the order of its arguments and a constant offset.
+This does not present a major problem because we minimize the loss directly, without assessing the two error components individually.},
+the estimator's risk can be decomposed as the sum
+\begin{equation}\label{eq:bias-variance}
+r = b + \varepsilon
+\end{equation}
+of \emph{approximation error} (``bias'')
 \begin{equation}
-r = \mathbb E[\mathcal L(\hat\Sigma, \Sigma)]
+b = \loss{\bar\Sigma,\Sigma}
 \end{equation}
+and \emph{estimation error} (``variance'') 
+\begin{equation}
+\varepsilon = \E{\loss{\hat\Sigma, \bar\Sigma}}
+\end{equation}
+where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.
 
-The sample covariance matrix is unbiased: it minimizes the bias $\mathcal L (\mathbb E[\hat\Sigma],\Sigma)$ but it does
+An unbiased estimator, such as $\hat\Sigma_0$, makes $\bar\Sigma=\Sigma$ and thus minimizes the approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
 
-Estimator risk can be reduced by \emph{regularization}.
+The overall error, or risk, of the estimator can be minimized by \emph{regularization}.
 Regularization is the deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
-Resulting regularized estimators are biased but  less variable.
-Regularization procedures must strike favorable balance between bias and variance so that the estimator risk is minimized.
-The terms \emph{bias} and \emph{variance} are applicable only the loss function is mean-squared-error.
-The more general terms are \emph{approximation error} and \emph{estimation error})
+Regularization procedures solve the bias-variance tradeoff problem and produce biased but less variable estimates that, when performed optimally, allow minimizing the estimator risk.
 

### 2013-09-29 sections/Results.tex
minor edits

@@
 
-\subsection*{Regularized estimates of the covariance matrix}
+\subsection*{Evaluation of covariance matrix estimates}
 The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.
-The loss function quantifies the discrepancy between $\hat\Sigma$ and $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.
+The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.
 
@@
 \end{equation}
-This choice is motivated by mathematical convenience and, we assert, the conclusions of the study are unlikely to change drastically under other popular loss functions such as quadratic loss or Stein's entropy loss.
+This choice is motivated by mathematical convenience.
+Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$, Stein's entropy loss, and quadratic loss \cite{James:1961,Ledoit:2004,Schafer:2005,Fan:2008}.
+We assert that the main conclusions of our study are unlikely to change drastically other well behaved loss functions.
 
-The aim of our project is produce covariance matrix estimates that minimize the expected loss.  The expected value of the loss function of an estimator is known as its \emph{risk}: 
+The aim of our project is produce covariance matrix estimates that minimize the expected loss.  The expected loss of an estimator is known as its \emph{risk}: 
 \begin{equation}
@@
 
+In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through validation: 
+Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$. Then \emph{empirical loss} is 
+\begin{equation}
+\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
+\end{equation}
+ and its expected value or \emph{empirical risk} is
+\begin{equation}\label{eq:empiricalRisk}
+\hat r = \E{\hat\ell} = \mathbb E_{\hat\Sigma} \left[ \mathbb E_{\hat\Sigma_0^\prime} \left[ {\loss{\hat\Sigma,\hat\Sigma_0^\prime}} \right] \right]
+\end{equation}
+Because the chosen loss function is linear in its second argument in the sense that
+\begin{equation}
+\loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
+\end{equation}
+the expection on the second argument in Eq.~\ref{eq:empiricalRisk} may be taken inside the loss function:
+\begin{equation}
+\hat r = \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}  = \E{\loss{\hat\Sigma,\Sigma}} = r
+\end{equation}
+Thus the empirical loss $\loss{\hat \Sigma,\hat \Sigma_0^\prime}$ serves as an unbiased estimate of risk $r$.
+
+Because the loss function is equivalent to the negative normal log likelihood, the above derivation led us to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes the normal log likelihood of the validation dataset.
+
+\subsection*{Regularization}
 Under many loss functions\footnote{
-The strict equality in Eq.~\ref{eq:bias-variance} does not hold for our choice of the loss function in Eq.~\ref{eq:loss}.
-However, the equality does hold for its close cousin, Stein's entropy loss,  $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$, which only differs by the order of its arguments and a constant offset.
-This does not present a major problem because we minimize the loss directly, without assessing the two error components individually.},
+The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}.
+However, the equality does hold for its close cousin, Stein's \emph{entropy loss},  which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$.
+This defficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually.
+The bias-variance decomposition is presented here to motivate the use of regularization.},
 the estimator's risk can be decomposed as the sum
@@
 \end{equation}
-of \emph{approximation error} (``bias'')
+of \emph{approximation error} (``bias'' or systematic error)
 \begin{equation}
@@
 
-An unbiased estimator, such as $\hat\Sigma_0$, makes $\bar\Sigma=\Sigma$ and thus minimizes the approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
+The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thus minimizes  approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
 
-The overall error, or risk, of the estimator can be minimized by \emph{regularization}.
+The estimator risk can be reduced by \emph{regularization}.
 Regularization is the deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
-Regularization procedures solve the bias-variance tradeoff problem and produce biased but less variable estimates that, when performed optimally, allow minimizing the estimator risk.
+A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimates aiming to minimize the estimator's risk.
+Many popular covariance estimators use single pre-determined target estimator and optimize the shrinkage intensity.
+Others emphasize the optimal selection of the low-dimensional target estimate from a family of estimates.
+Yet others effectively combine shrinkage and selection.
 
-Numerous regularized covariance matrix estimates have been devised.
-They start with the sample covariance matrix but differ in the choice of the target estimate and the shrinkage path and intensity.
-Some regularization schemes focus on the optimal choice of the low-dimensional target estimate from a large family of target estimates and do not require shrinkages.
-Others have a fairly simple target estimate but instead adjust the shrinkage intensity to reduce the estimation error.
-
-A regularized estimate of $\Sigma$ can be constructed by either (a) selecting an optimal low-dimensional approximation of $\hat\Sigma_0$ from a given family of approximations and/or (b) biasing or \emph{shrinkage} of $\hat\Sigma_0$ toward a given low-dimensional target estimate by an optimal amount.
-Both dimensionality reduction and shrinkage approaches reduce the variance of the solution while introducing bias.
-
-We evaluated four regularized estimators which we will denote as A, B, C, and D.
+We evaluated four regularized estimators denoted here as A, B, C, and D.
 Their target estimates correspond to covariance matrices of Gaussian graphical models depicted in Figure 2.
-Here the green spheres represent the recorded neurons.
-The light-colored balls represent latent units of the graphical model and the edges connecting them represent conditional dependencies.
+The green spheres represent the recorded neurons, whereas the light-colored spheres represent latent units of the graphical model and the edges connecting them represent conditional dependencies.
 
@@
 A \emph{graphical model} is a multivarate probability distribution with a specified graph of conditional dependencies between its variables \cite{Koller:2009}.
-When such a distribution is a multivariate normal distribution, the model becomes a \emph{Gaussian graphical model} or, equivalently, a \emph{Gaussian Markov Random Field}.
+When this distribution is a multivariate normal distribution, the model becomes a \emph{Gaussian graphical model} or, equivalently, a \emph{Gaussian Markov Random Field}.
 Gaussian graphical models have a straightforward relationship to their covariance matrix $\Sigma$:  Zero elements in the inverse of $\Sigma$ indicate conditional independence between the corresponding pair of variables.
@@
 
-\subsubsection*{Estimate A: Shrinkage toward the independent model}
-In estimate A, the target is a family of diagonal matrices 
+\emph{Estimate A} is shrinkage estimator toward an independent model.  
 \begin{equation}

### 2013-09-30 sections/Results.tex
worked on the description of estimators.

@@
 
-The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thus minimizes  approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
+The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thereby minimizes the approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
 
@@
 A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimates aiming to minimize the estimator's risk.
-Many popular covariance estimators use single pre-determined target estimator and optimize the shrinkage intensity.
-Others emphasize the optimal selection of the low-dimensional target estimate from a family of estimates.
-Yet others effectively combine shrinkage and selection.
+Various regularization schemes focus on the dimensionality reduction part \TODO{rephrase} by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.
+Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}.
+Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.
 
 We evaluated four regularized estimators denoted here as A, B, C, and D.
-Their target estimates correspond to covariance matrices of Gaussian graphical models depicted in Figure 2.
-The green spheres represent the recorded neurons, whereas the light-colored spheres represent latent units of the graphical model and the edges connecting them represent conditional dependencies.
+Their target estimates correspond to covariance matrices of Gaussian graphical models depicted in Figure 2 where green spheres represent the recorded neurons,  light-colored spheres represent latent units of the graphical model, and edges connecting them represent conditional dependencies or interactions.
 
@@
 A \emph{graphical model} is a multivarate probability distribution with a specified graph of conditional dependencies between its variables \cite{Koller:2009}.
-When this distribution is a multivariate normal distribution, the model becomes a \emph{Gaussian graphical model} or, equivalently, a \emph{Gaussian Markov Random Field}.
+When this distribution is a multivariate normal distribution, the model becomes a \emph{Gaussian graphical model} (GGM) or, equivalently, a \emph{Gaussian Markov Random Field}.
 Gaussian graphical models have a straightforward relationship to their covariance matrix $\Sigma$:  Zero elements in the inverse of $\Sigma$ indicate conditional independence between the corresponding pair of variables.
@@
 
-\emph{Estimate A} is shrinkage estimator toward an independent model.  
+\subsubsection*{Estimator A: Shrinkage toward identity.}
+In \emph{estimator A}, the sample covariance matrix $\hat\Sigma_0$ is shrunk toward the independent model.  The covariance matrix of a GGM in which all units are independent will be diagonal.  
+For example, we  could leave the sample variances on the diagonal and zero out the off-diagonal products, making the target estimate 
+\begin{equation}\label{eq:sampleVariance}
+\hat T= \hat\Sigma_0\circ I
+\end{equation}
+where $\circ$ denotes the entrywise matrix product (Hadamard product). 
+Alternatively, the target could have equal variances, all equal to the average sample variance in the population $v = \frac 1 p \Tr(\hat\Sigma_0)$.
+\begin{equation}\label{eq:equalVariance}
+\hat T=\frac 1 p \Tr(\hat\Sigma_0) I
+\end{equation}
+The equal-variance target in Eq.~\ref{eq:equalVariance} has only one degree of freedom and thus lower estimation error but higher bias than the target with sample variances in Eq.~\ref{eq:sampleVariance}.
+It is not immediately clear which target is better in a particular application.
+We therefore use the linear mixture of the two targets paramaterized by the mixing coefficient $\alpha\in[0,1]$:
 \begin{equation}
@@
 \end{equation}
-where $\circ$ is the entrywise product (Hadamard product).
-When $\alpha=0$, $\hat T_\alpha$ has the average variance along the diagonal.
-When $ \alpha=1$, $ \hat T_\alpha$ has empirical variances on the diagonal.
+Then the estimator $\hat\Sigma_{\alpha,\lambda}$ is the linear shrinkage of the sample covariance matrix $\hat\Sigma_0$ toward $\hat T_\alpha$ controlled by the mixing proportion $\lambda\in[0,1]$:
+\begin{equation}
+\hat\Sigma_{\alpha,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha 
+\end{equation}
+This covariance matrix estimator is implemented by R's {\tt corpcor} package \cite{Schaefer:2010} and enjoys widespread use in many applications \cite{Schafer:2005}.
 
-If dependencies between neurons are strictly linear, then a diagonal covariance matrix represents the case where neurons are independent, with no interactions.
+We estimated the optimal values of hyperparameters $ \alpha$ and $ \lambda$  by nested cross-validation within the training dataset. \TODO{more detail?}
 
-The overall estimate $\hat\Sigma$ is found by linear mixing of $\hat\Sigma_0$ and $ \hat T_\alpha$ controlled by the mixing proportion $\lambda$:
+\subsubsection*{Estimate B: Shrinkage toward a factor model}
+\emph{Estimator B} shrinks the sample covariance matrix $\hat\Sigma_0$ toward a target estimated by \emph{factor analysis}.
+For multivariate normal distributions, factor analysis estimates a graphical model in which the observed units do not interact directly but are influenced by a small number of common inputs.
+
+The target estimate
 \begin{equation}
-\hat\Sigma = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha 
+T_d = L_dL_d^\T + \Psi
 \end{equation}
+where $L_d$ is the $p\times d$ \emph{loading matrix} and $\Psi$ is a diagonal matrix. 
+where $(L_d,\Psi) = \arg\min\limits_{\dot L_d,\dot\Psi} \loss{\hat\Sigma_0,\dot L_d \dot L_d^\T + \dot \Psi}$
 
-The hyperparameters $ \alpha$ and $ \lambda$ can be optimally selected from the training data by nested cross-validation.
-
-\subsubsection*{Estimate B: Shrinkage toward }
-In estimator B, the target is the multifactor model $ \hat T_d$ with $ d$ latent units.
+ $ \hat T_d$ with $ d$ latent units.
 When all dependencies are linear, factor models correspond to mutually independent neurons each influenced by $ d$ latent units.

### 2013-09-30 sections/Discussion.tex
Added figure captions.
minor changes
updated estimator descriptions.

@@
-\hl{The discussion section is not ready for any feedback. Currently it's just a trashbin with clips from scrapped versions of the introduction. -- Dimitri}
+\TODO{The discussion section is not yet ready for feedback. Currently it's a collection of scraps from earlier versions of the introduction.}
 
@@
 
-Considerable progress has been made in sensory areas where the functional characterizations of cells are defined by their reproducible responses to external stimuli (reviewed in \citep{Reid:2012}).
+Considerable progress has been made in sensory areas where the functional characterizations of cells are defined by their reproducible responses to external stimuli (reviewed in \cite{Reid:2012}).
 
-Many key questions in neuroscience revolve around the relationship between the cytoarchitecture of neural microcircuits and the functional organization of their activity \citep{Reid:2012}.
+Many key questions in neuroscience revolve around the relationship between the cytoarchitecture of neural microcircuits and the functional organization of their activity \cite{Reid:2012}.
 Investigators in this line of inquiry aim to construct networks of functional associations within groups of neurons inferred from observations of their activity under a variety of conditions.
 The inferred functional network  could then related to the synaptic connectivity patterns, cell types, cell properties, or the cells' spatial arrangment in the attempt to uncover organizational principles of neural computation.
-In addition, the network itself can serve as subtrate for subsequent graph-theoretical analysis \citep{Feldt:2011}
+In addition, the network itself can serve as subtrate for subsequent graph-theoretical analysis \cite{Feldt:2011}
 
 Advances in electrophysiology and optical imaging have enabled simultaneous recordings from dozens to hundreds or thousands of cells. 
-In particular, recent advances in two-photon imaging of calcium signals have allowed in vivo recordings from nearly every neuron brain-wide in zebrafish \citep{Leung:2013,Ahrens:2013} and in a 3D volume of a few hundred microns in diameter in mouse neocortex \citep{Katona:2012,Cotton:2013}.
-Ambitious projects currently under way aim to record the spiking activity of large fractions of cells from entire circuits and systems in behaving animals \citep{Alivisatos:2012}.
+In particular, recent advances in two-photon imaging of calcium signals have allowed in vivo recordings from nearly every neuron brain-wide in zebrafish \cite{Leung:2013,Ahrens:2013} and in a 3D volume of a few hundred microns in diameter in mouse neocortex \cite{Katona:2012,Cotton:2013}.
+Ambitious projects currently under way aim to record the spiking activity of large fractions of cells from entire circuits and systems in behaving animals \cite{Alivisatos:2012}.
 Direct observations of the population activity  of entire circuits open new possibilities for incisive statistical descriptions of the functional connectivity in the circuit.
 
-One general approach is to construct statistical models reflecting hypothesized models of neural circuit function, including temporal dynamics, nonlinear interactions between neurons, and background network activity \citep{Pillow:2008,Buonomano:2009,Yu:2009}.
+One general approach is to construct statistical models reflecting hypothesized models of neural circuit function, including temporal dynamics, nonlinear interactions between neurons, and background network activity \cite{Pillow:2008,Buonomano:2009,Yu:2009}.
 Another general approach is to search for alternative statistics that best describe the population activity without assuming a specific model.
-Such evaluations rely on constructing surrogate datasets \citep{Okun:2012} or maximum-entropy distributions that reproduce the statistic of interest \citep{Ganmor:2011,Tkacik:2012} with subsequent testing of how well such constructs can reproduce other observed properties of population activity.
+Such evaluations rely on constructing surrogate datasets \cite{Okun:2012} or maximum-entropy distributions that reproduce the statistic of interest \cite{Ganmor:2011,Tkacik:2012} with subsequent testing of how well such constructs can reproduce other observed properties of population activity.
 
@@
 Linear correlations between pairs of neurons are among the most common and familiar descriptions of functional associations in networks and of the collective activity of neuronal populations.
-For example, it is tempting to consider the network constructed from the highest correlations in the recorded population (e.g.~\cite{Malmersjo:2013}) for subsequent graph-theoretical analysis \citep{Feldt:2011}.
+For example, it is tempting to consider the network constructed from the highest correlations in the recorded population (e.g.~\cite{Malmersjo:2013}) for subsequent graph-theoretical analysis \cite{Feldt:2011}.
 In functional genomics, networks constructed from highest correlations are called \emph{relevance networks}.
 
-However, pairwise correlations unreliable proxies of direct functional association such as direct synaptic connectivity as they can arise due to a wide variety of physiological interactions including direct synaptic interactions, indirect chains of synaptic connections, common inputs, correlations in inputs, synchrony, fluctuations in global network states, and others \citep{Shadlen:1998,Ostojic:2009,Pernice:2011,Schneidman:2006}.
+However, pairwise correlations unreliable proxies of direct functional association such as direct synaptic connectivity as they can arise due to a wide variety of physiological interactions including direct synaptic interactions, indirect chains of synaptic connections, common inputs, correlations in inputs, synchrony, fluctuations in global network states, and others \cite{Shadlen:1998,Ostojic:2009,Pernice:2011,Schneidman:2006}.
 
-Other experimental evidence suggests that detailed interactions in local microcircuits have only weak effects on overall population activity, which is then best characterized by collective features such as global population dynamics or population synchrony \citep{Okun:2012,Tkacik:2012,Tkacik:2013}.
+Other experimental evidence suggests that detailed interactions in local microcircuits have only weak effects on overall population activity, which is then best characterized by collective features such as global population dynamics or population synchrony \cite{Okun:2012,Tkacik:2012,Tkacik:2013}.
 From this point of view, the imporant aspects of the correlation structure lie it its global aspects, such as the eigenspectrum, while the individual pairwise correlations are not particularly meaningful.

### 2013-09-30 sections/Estimators.tex
updated estimator descriptions.

--- 2013-08-19 sections/Estimators.tex
@@
-\subsection*{Estimator A: Shrinkage toward diagonal}
-The most popular covariance regularization schemes use diagonal target estimates and linear shrinkage toward the target.  For example, the target could be the identity matrix:  
+We evaluated four regularized estimators denoted here as A, B, C, and D.
+Their target estimates correspond to covariance matrices of Gaussian graphical models depicted in Figure 2 where green spheres represent the recorded neurons,  light-colored spheres represent latent units of the graphical model, and edges connecting them represent conditional dependencies or interactions.
+
+\input{sections/Figure2.tex}
+
+A \emph{graphical model} is a multivarate probability distribution with a specified graph of conditional dependencies between its variables \cite{Koller:2009}.
+When this distribution is a multivariate normal distribution, the model becomes a \emph{Gaussian graphical model} (GGM) or, equivalently, a \emph{Gaussian Markov Random Field}.
+Gaussian graphical models have a straightforward relationship to their covariance matrix $\Sigma$:  Zero elements in the inverse of $\Sigma$ indicate conditional independence between the corresponding pair of variables.
+
+Fitting a graphical model to data involves two tasks: (1) the selection of the set of non-zero covariances or \emph{covariance selection} \cite{Dempster:1972} and (b) the fitting of the non-zero elements.
+
+\subsubsection*{Estimator A: Shrinkage toward diagonal.}
+In \emph{estimator A}, the sample covariance matrix $\hat\Sigma_0$ is shrunk toward the independent model.  The covariance matrix of a GGM in which all units are independent will be diagonal.  
+For example, we  could leave the sample variances on the diagonal and zero out the off-diagonal products, making the target estimate 
+\begin{equation}\label{eq:sampleVariance}
+\hat T= \hat\Sigma_0\circ I
+\end{equation}
+where $\circ$ denotes the entrywise matrix product (Hadamard product). 
+Alternatively, the target could have equal variances, all equal to the average sample variance in the population $v = \frac 1 p \Tr(\hat\Sigma_0)$.
+\begin{equation}\label{eq:equalVariance}
+\hat T=\frac 1 p \Tr(\hat\Sigma_0) I
+\end{equation}
+The equal-variance target in Eq.~\ref{eq:equalVariance} has only one degree of freedom and thus lower estimation error but higher bias than the target with sample variances in Eq.~\ref{eq:sampleVariance}.
+It is not immediately clear which target is better in a particular application.
+We therefore use the linear mixture of the two targets paramaterized by the mixing coefficient $\alpha\in[0,1]$:
 \begin{equation}
-\hat T = \hat v I
+\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I
 \end{equation}
-where $\hat v = \frac 1 p \sum\limits_{i=1}^p(\hat\Sigma_0)_{ii}$ is the mean sample variance.
+Then the estimator $\hat\Sigma_{\alpha,\lambda}$ is the linear shrinkage of the sample covariance matrix $\hat\Sigma_0$ toward $\hat T_\alpha$ controlled by the mixing proportion $\lambda\in[0,1]$:
+\begin{equation}
+\hat\Sigma_{\alpha,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha 
+\end{equation}
+This covariance matrix estimator is implemented by R's {\tt corpcor} package \cite{Schaefer:2010} and enjoys widespread use in many applications \cite{Schafer:2005}.
 
-Alternatively, the target could contain the sample variances:
+We estimated the optimal values of hyperparameters $ \alpha$ and $ \lambda$  by  cross-validation within the training dataset.
+\TODO{more detail?} \TODO{discuss analytic optimization of shrinkage intensities.}
+
+\subsubsection*{Estimate B: Shrinkage toward a factor model}
+\emph{Estimator B} shrinks the sample covariance matrix $\hat\Sigma_0$ toward a target estimated by \emph{factor analysis} with $d$ factors.
+For multivariate normal distributions, factor analysis estimates a graphical model in which the observed units do not interact directly but are influenced by $d$ common inputs.
+
+The target estimate is then 
 \begin{equation}
-\hat T= \hat\Sigma_0 \circ I 
+\hat T_d = \hat L_d\hat L_d^\T + \hat \Psi
 \end{equation}
-where $\circ$ is the entrywise matrix product.
+where $\hat L_d$ is the $p\times d$ \emph{loading matrix} and $\hat \Psi$ is the $p\times p$ diagonal matrix of specific variances.
+The value of $\hat T\_d$ is found by minimizing $\loss{\hat T_d, \hat\Sigma_0}$.
 
-Finally, the target could be a linear mixture of the common variance and independent variance targets
+Just as in estimator A, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$.
 \begin{equation}
-\hat T_{\hat\eta} = (1-\hat\eta)(\hat\Sigma_0 \circ I) + \hat\eta\hat v I
+\hat\Sigma_{d,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d
+\end{equation}
+and the optimal values of $d$ and $\lambda$ are found by cross-validation within the training set.
+
+Factor analysis has a rich history of testing hypotheses about the dominance of a few common inputs \cite{findagoodone}.
+In neuroscience, the hypothesis that the population activity in small local circuits is largely driven by a few latent factors has been investigated \cite{Yu:2009,Ecker:2013}.
+Estimation of covariance matrices by shrinkage toward a factor model has been used in finance for portfolio risk assessment \cite{Ledoit:2003,Fan:2008}.
+
+\subsubsection*{Estimator C: Spare inverse}
+\emph{Estimator C} produces a low-dimensional estimate by finding an approximation of $\hat\Sigma_0$ that has many zeros in its inverse.
+This approximation problem is known as \emph{covariance selection} \cite{Dempster:1972,Meinshausen:2006}.
+
+For Gaussian Graphical Models, zeros in the inverse covariance matrix indicate statisitcal independence between the corresponding pairs.
+Thus the inverse covariance matrix is also the adjacency matrix of the graphical model, indicating which pairs are conditionally dependent.
+\TODO{discuss how quickly (or slowly) this breaks down with deviations from gaussianity, e.g.~MaxEnt models, Ising models.} \TODO{Also, motivate this from the perpsective of linear regression as in \cite{Varoquaux:2012}.} \TODO{Relate to partial pairwise correlations.} Regularization of covariance matrix estimation by sparsifying the inverse of the estimate has been using in function genomics \cite{Schafer:2005,other} and fMRI functional connectivity \cite{Varoquaux:2012}.
+
+The covariance matrix estimate is then 
+\begin{equation}
+\hat\Sigma_\rho = \left(\arg\min\limits_S \loss{S^{-1},\hat\Sigma_0}  \right)^{-1} \quad \mbox{such that} \quad  \|S_\rho\|_0 = \rho
+\end{equation}
+where $S_\rho$ is a sparse matrix with $\rho$ non-zero coefficients. $\|S\|_0$ denotes the number of non-zero elements of matrix $S$.
+Solving this approximation in this form is computationally challenging.
+However, replacing the $L_0$ norm with the $L_1$ norm makes the approximation problem convex  while still allowing to uncover the true zero structure \cite{Friedman:2008,Donoho:2000}.
+The resulting algorithm, known as \emph{graphical lasso} \cite{Friedman:2008} finds the estimator as
+\begin{equation}
+\hat\Sigma_\lambda = \left(\arg\min\limits_S \loss{S^{-1},\hat\Sigma_0} + \lambda \|S\|_1\right)^{-1}
+\end{equation}
+where $\lambda \|S\|_1$ is the $L_1$ norm of the matrix. \TODO{clarify}.
+
+Just as in the previous estimators, the optimal value of the regularization parameter $\lambda$ is found from training data by cross-validation.
+
+\subsubsection*{Estimator D: Sparse inverse + low-rank}
+\emph{Estimator D} produces a low-dimensional estimate by finding an apprxomation of $\hat\Sigma_0$ approximating the sample covariance matrix such that its inverse is the sum of a spase matrix and a low-rank matrix.
+
+In estimate D, the target $\hat R_{d,\alpha}$ is the matrix whose inverse is the sum of a sparse matrix $\hat S_\alpha$ and a low-rank matrix $\hat L_d$ of rank $d$.
+\begin{equation}
+\hat R_{d,\alpha} = (\hat S_\alpha + \hat L_d)^{-1}
 \end{equation}
 
-The regularized estimator is the linear mixture of $\hat\Sigma_0$ and $\hat T_{\hat\eta}$:
-\begin{equation}
-\hat\Sigma_{\hat\eta,\hat\lambda} = (1-\hat\lambda)\hat\Sigma_0 + \hat\lambda \hat T_{\hat\eta} 
-\end{equation}
+If all dependencies are linear, the estimate D describes a group of neurons where each interacts with a small number of latent units and a small fraction of the observed neurons.
 
-The hyperparameters $\hat\eta$ and $\hat\lambda$ must be estimated from the data.
-Under the MSE loss function $\mathcal L_e$ (\ref{eq:MSE}), the optimal values of $\hat\eta$ and $\hat\lambda$ can be estimated analytically \cite{Ledoit:2004,Schafer:2005,Schaefer:2010}.
-However, these estimates are no longer optimal under the Guassian loss $\mathcal L_g$.
-\TODO{I have tested this and there was a substantial difference in both synthetic and empirical data}  When an analytical solution for optimal hyperparameters is not available, the optimal values can be found by cross validation, entirely within the the training sample.
-\TODO{explain nested cross-validation in more detail?}
-
-\TODO{Describe how the empirical loss is convex in the hyperparameters in this case}
-
-\subsection*{Estimator B: Shrinkage toward a factor model}
-In estimator B, the target is a factor model, composed as the sum of a low-rank component $\hat L_{\hat d}\hat L_{\hat d}^\T$ and a diagonal matrix $\hat \Psi$:
-\begin{equation}
-\hat T_{\hat d} = \hat L_{\hat d} \hat L_{\hat d}^\T + \hat\Psi
-\end{equation}
-Here $\hat L_{\hat d}$ is an $p\times\hat d$ matrix and $\hat\Psi$ is diagonal.
-
-If $F$ were assume to be multivariate normal \TODO{or other distributions that are defined by linear dependencies}, the factor model is corresponds to the graphical model (\ref{fig:02}B) where the activity of all neurons are dependent only $\hat d$ latent units.
-This suggest a mechanistic interpretation in which the interactions between neurons are insignificant compared to the influence of inputs outside the recorded circuit.
-
-Just as in Estimator A, Estimator B is allowed to commit to the low-dimensional target only partially: the overall estimate is the linear mixture of the sample covariance and the target
-\begin{equation}
-\hat\Sigma_{\hat d,\hat\lambda} = (1-\hat\lambda)\hat\Sigma_0 + \hat  T_{\hat d}
-\end{equation}
-
-\TODO{Check \cite{Ledoit:2003,Fan:2011,Fan:2006}}
-
-\subsection*{Estimator C: Sparse inverse}
-Assuming that $x$ is distributed normally, the inverse of the covariance matrix or \emph{precision matrix} $K=\Sigma^{-1}$ has special significance: zeros in the precision matrix indicate conditional independence between the corresponding pairs.
-To see this, let $x=(x_1,\ldots,x_p)^\T \sim g\left(x\cond K\right)$
-\begin{equation}
-g\left(x\cond K\right) = \frac 1 {Z(K)} \exp\left(-\frac 1 2 \sum\limits_{i=1}^p\sum\limits_{j=1}^p  K_{ij} x_i x_j\right)
-\end{equation} 
-If $K_{12}\equiv K_{21} = 0$, then the joint distribution of $x_1$ and $x_2$, conditioned on $x_3=a_3,\ldots,x_p=a_p$ can be decomposed as product of independent distributions of $x_1$ and $x_2$: 
-\begin{equation}
-\begin{split}
-g\left( x_1, x_2 \cond K, x_3=a_3,\ldots,x_p=a_p\right) &=
-\frac 1 {Z(K)} \exp\left(-\frac 1 2 \sum\limits_{i=3}^p\sum\limits_{j=3}^p  K_{ij} a_i a_j \right)\times
-\\
- &  \exp\left( -\frac 1 2\left( K_{11} x_1^2 +  x_1\sum\limits_{i=3}^p K_{1i}a_i \right)  \right)
-\\
- &  \exp\left( -\frac 1 2\left( K_{22} x_2^2 +  x_2\sum\limits_{i=3}^p K_{2i}a_i \right)  \right)
-\end{split}
-\end{equation}
-
-\cite{Dempster:1972,Meinshausen:2006,Friedman:2008}
-
-\subsection*{Estimator D: Sparse inverse with latent units}
 \cite{Ma:2013}

### 2013-09-30 sections/Figure4.tex
updated estimator descriptions.

\begin{figure}[htp]
\centering
\includegraphics[width=1.0\textwidth]{figures/Figure4.pdf}
\caption{Estimator D dominates over the other estimators in estimation of noise covariances in dense recordings from primary visual cortex.
}\label{fig:04}
\end{figure}

### 2013-09-30 sections/Figure5.tex
updated estimator descriptions.

--- 2013-09-25 sections/Figure2.tex
@@
 \textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
-\textbf{C}: In the Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from sparse pairwise linear interactions between visible units.
-\textbf{D}: In the Gaussian graphical model with latent units, correlations arise  between pairs of nodes
+\textbf{C}: Graphical model with conditional dependencies between a specified subset of pairs of observed neurons, no latent units. 
+\textbf{D}: Graphical model with conditional dependencies between a specified subset of pairs of observed neurons and dense interactions with a few latent units.
 }\label{fig:02}

### 2013-09-30 sections/Results.tex
updated estimator descriptions.

--- 2013-09-29 sections/Results.tex
@@
 \begin{equation}
-\hat \Sigma_0 = \frac 1 {n-c} \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
+\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
 \end{equation}
-where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals.
-Bessel's correction $c$ makes the estimate unbiased such that $\E{\hat\Sigma_0} = \Sigma$.
-When observations $x(t)$ can be assumed to be independent then $c=1$; but when observations are correlated $c>1$ may be estimated from the data.
+where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals; $\nu$ is the number of degrees of freedom.
+For independent observations $\nu=n-1$ because estimation of the mean $\mu$ accounts for one degree of freedom.
+When observations are correlated, as is the case with calcium signals, $\nu < n-1$ and may be estimated from the signal.
+
+The sample covariance matrix is constructed to be unbiased, such that $\E{\hat\Sigma_0} = \Sigma$.
 
@@
 A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimates aiming to minimize the estimator's risk.
-Many popular covariance estimators use single pre-determined target estimator and optimize the shrinkage intensity.
-Others emphasize the optimal selection of the low-dimensional target estimate from a family of estimates.
-Yet others effectively combine shrinkage and selection.
+Various regularization schemes focus on the dimensionality reduction part \TODO{rephrase} by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.
+Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}.
+Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.
 
-We evaluated four regularized estimators denoted here as A, B, C, and D.
-Their target estimates correspond to covariance matrices of Gaussian graphical models depicted in Figure 2.
-The green spheres represent the recorded neurons, whereas the light-colored spheres represent latent units of the graphical model and the edges connecting them represent conditional dependencies.
+\input{sections/Estimators.tex}
 
-\input{sections/Figure2.tex}
+\subsection*{In simulation, estimator selection `recognizes' the covariance structure}
+\input{sections/Figure3.tex}
 
-A \emph{graphical model} is a multivarate probability distribution with a specified graph of conditional dependencies between its variables \cite{Koller:2009}.
-When this distribution is a multivariate normal distribution, the model becomes a \emph{Gaussian graphical model} or, equivalently, a \emph{Gaussian Markov Random Field}.
-Gaussian graphical models have a straightforward relationship to their covariance matrix $\Sigma$:  Zero elements in the inverse of $\Sigma$ indicate conditional independence between the corresponding pair of variables.
-
-Fitting a graphical model to data involves two tasks: (1) the selection of the set of non-zero covariances or \emph{covariance selection} \cite{Dempster:1972} and (b) the fitting of the non-zero elements.
-
-\emph{Estimate A} is shrinkage estimator toward an independent model.  
-\begin{equation}
-\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I
-\end{equation}
-where $\circ$ is the entrywise product (Hadamard product).
-When $\alpha=0$, $\hat T_\alpha$ has the average variance along the diagonal.
-When $ \alpha=1$, $ \hat T_\alpha$ has empirical variances on the diagonal.
-
-If dependencies between neurons are strictly linear, then a diagonal covariance matrix represents the case where neurons are independent, with no interactions.
-
-The overall estimate $\hat\Sigma$ is found by linear mixing of $\hat\Sigma_0$ and $ \hat T_\alpha$ controlled by the mixing proportion $\lambda$:
-\begin{equation}
-\hat\Sigma = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha 
-\end{equation}
-
-The hyperparameters $ \alpha$ and $ \lambda$ can be optimally selected from the training data by nested cross-validation.
-
-\subsubsection*{Estimate B: Shrinkage toward }
-In estimator B, the target is the multifactor model $ \hat T_d$ with $ d$ latent units.
-When all dependencies are linear, factor models correspond to mutually independent neurons each influenced by $ d$ latent units.
-(This is  the conventional factor analysis).
-
-Just as in A, the overall estimate is found by linear mixing:
-\begin{equation}
-\hat\Sigma = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d
-\end{equation}
-
-\subsubsection*{Estimate C}
-In estimate C, the target $ \hat S_\alpha$ is a matrix whose inverse is \emph{sparse}, with a large fraction of off-diagonal elements fixed at zero.
-The coefficient $ \alpha$ controls the sparsity in $ \hat S_\alpha$.
-
-In models with only linear dependencies, zeros in the inverse covariance indicate conditional independence.
-Thus the target estimate is depicted as the graphical model with sparse interneuronal interactions.
-
-In practice the target $\hat S_\alpha$ is found by an $ L_1$-norm optimization technique, which combines dimensionality reduction and shrinkage into a single computationally efficient algorithm.
-
-Just as in the previous estimators, the optimal value of the regularization parameter $ \alpha$ is determined from the data by nested cross-validation.
-
-\subsubsection*{Estimate D}
-In estimate D, the target $\hat R_{d,\alpha}$ is the matrix whose inverse is the sum of a sparse matrix $\hat S_\alpha$ and a low-rank matrix $\hat L_d$ of rank $d$.
-\begin{equation}
-\hat R_{d,\alpha} = (\hat S_\alpha + \hat L_d)^{-1}
-\end{equation}
-
-If all dependencies are linear, the estimate D describes a group of neurons where each interacts with a small number of latent units and a small fraction of the observed neurons.
-
-\subsection*{Sparse inverse covariance with a low-rank component dominates other estimators}
-I computed noise covariances of population activity from 31 sites.
-In each site, I compared the cross-validated performance of the covariance matrix using the Gaussian log-likelihood loss function between the trained estimate $ \hat\Sigma$ and the testing sample covariance $ \Sigma_0^\prime$:
-\begin{equation}
-\mathcal L(\hat\Sigma,\hat\Sigma_0^\prime) = 
-\frac 1 p\left( \ln \det \hat \Sigma + \mbox{tr}(\hat \Sigma^{-1}\hat\Sigma_0^\prime) \right) 
-\end{equation}
-
-I obtained the following in figure 3.
-
+\subsection*{In visual cortex, Estimator D dominates}
+\input{sections/Figure4.tex}
 This is the full matrix of all comparisons of the estimators.

### 2013-09-30 sections/Estimators.tex
minor

@@
 Thus the inverse covariance matrix is also the adjacency matrix of the graphical model, indicating which pairs are conditionally dependent.
-\TODO{discuss how quickly (or slowly) this breaks down with deviations from gaussianity, e.g.~MaxEnt models, Ising models.} \TODO{Also, motivate this from the perpsective of linear regression as in \cite{Varoquaux:2012}.} \TODO{Relate to partial pairwise correlations.} Regularization of covariance matrix estimation by sparsifying the inverse of the estimate has been using in function genomics \cite{Schafer:2005,other} and fMRI functional connectivity \cite{Varoquaux:2012}.
+\TODO{discuss how quickly (or slowly) this breaks down with deviations from gaussianity, e.g.~MaxEnt models, Ising models.} \TODO{Also, motivate this from the perpsective of linear regression as in \cite{Varoquaux:2012}.} \TODO{Also relate to partial pairwise correlations.} Regularization of covariance matrix estimation by sparsifying the inverse of the estimate has been using in function genomics \cite{Schafer:2005,other} and fMRI functional connectivity \cite{Varoquaux:2012}.
 

### 2013-10-01 sections/Estimators.tex
completed estimator description.

--- 2013-09-30 sections/Estimators.tex
@@
 \begin{equation}
-\hat T_d = \hat L_d\hat L_d^\T + \hat \Psi
+\hat T_d = \hat L_d + \hat \Psi
 \end{equation}
-where $\hat L_d$ is the $p\times d$ \emph{loading matrix} and $\hat \Psi$ is the $p\times p$ diagonal matrix of specific variances.
-The value of $\hat T\_d$ is found by minimizing $\loss{\hat T_d, \hat\Sigma_0}$.
+where $\hat L_d$ is a $p\times p$ matrix of rank $d$ and $\hat \Psi$ is a $p\times p$ diagonal matrix. The value of $\hat T_d$ is found by minimizing $\loss{\hat T_d, \hat\Sigma_0}$.
 
@@
 \emph{Estimator C} produces a low-dimensional estimate by finding an approximation of $\hat\Sigma_0$ that has many zeros in its inverse.
-This approximation problem is known as \emph{covariance selection} \cite{Dempster:1972,Meinshausen:2006}.
+This approximation problem is known as \emph{covariance selection} \cite{Dempster:1972}.
 
 For Gaussian Graphical Models, zeros in the inverse covariance matrix indicate statisitcal independence between the corresponding pairs.
-Thus the inverse covariance matrix is also the adjacency matrix of the graphical model, indicating which pairs are conditionally dependent.
-\TODO{discuss how quickly (or slowly) this breaks down with deviations from gaussianity, e.g.~MaxEnt models, Ising models.} \TODO{Also, motivate this from the perpsective of linear regression as in \cite{Varoquaux:2012}.} \TODO{Relate to partial pairwise correlations.} Regularization of covariance matrix estimation by sparsifying the inverse of the estimate has been using in function genomics \cite{Schafer:2005,other} and fMRI functional connectivity \cite{Varoquaux:2012}.
+Thus the inverse covariance matrix $S=-\Sigma^{-1}$ is also the adjacency matrix of the GGM, indicating which pairs are conditionally dependent.
+\TODO{discuss how quickly (or slowly) this breaks down with deviations from gaussianity, e.g.~MaxEnt models, Ising models.} \TODO{Also, motivate this from the perpsective of linear regression as in \cite{Varoquaux:2012}.} The off-diagonal elements of $S$ are proportional to partial pairwise correlations between corresponding pairs of neurons.
+\TODO{Expand.} Regularization of covariance matrix estimation by sparsifying the inverse of the estimate has been using in function genomics \cite{Schafer:2005,other} and fMRI functional connectivity \cite{Varoquaux:2012}.
 
@@
 \begin{equation}
-\hat\Sigma_\rho = \left(\arg\min\limits_S \loss{S^{-1},\hat\Sigma_0}  \right)^{-1} \quad \mbox{such that} \quad  \|S_\rho\|_0 = \rho
+-\hat\Sigma_\rho^{-1} =\arg\min\limits_S \loss{-S^{-1},\hat\Sigma_0}  \quad \mbox{such that} \quad  \|S_\rho\|_0 = \rho
 \end{equation}
@@
 Solving this approximation in this form is computationally challenging.
-However, replacing the $L_0$ norm with the $L_1$ norm makes the approximation problem convex  while still allowing to uncover the true zero structure \cite{Friedman:2008,Donoho:2000}.
-The resulting algorithm, known as \emph{graphical lasso} \cite{Friedman:2008} finds the estimator as
+The problem can be made computationally tractable by replacing the $L_0$ norm with the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.
+The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Banerjee:2008,Friedman:2008}, produces the following estimator
 \begin{equation}
-\hat\Sigma_\lambda = \left(\arg\min\limits_S \loss{S^{-1},\hat\Sigma_0} + \lambda \|S\|_1\right)^{-1}
+-\hat\Sigma_\lambda^{-1} = \arg\min\limits_S \loss{-S^{-1},\hat\Sigma_0} + \lambda \|S\|_1
 \end{equation}
@@
 
-In estimate D, the target $\hat R_{d,\alpha}$ is the matrix whose inverse is the sum of a sparse matrix $\hat S_\alpha$ and a low-rank matrix $\hat L_d$ of rank $d$.
+The inverse $S=-\Sigma^{-1}$ of the covariance matrix of a GGM constitutes the graph of conditional dependencies (interactions) between the units.
+With the assumption of sparse organization of interactions in the network, $S$ is assumed to be sparse.
+When the observed units are part of a larger GGM that includes other latent units, let the negative inverse $S^*$ of the joint distribution of both observed and latent units be composed of four partitions
 \begin{equation}
-\hat R_{d,\alpha} = (\hat S_\alpha + \hat L_d)^{-1}
+S^* = 
+\begin{pmatrix}
+S_{11} & S_{12} \\
+S_{12}^\T & S_{22} 
+\end{pmatrix}
+\end{equation}
+where partition $S_{11}$ describes the interactions between the visible units, $S_{12}$ describe the interactions between visible and latent units, and $S_{22}$ describes the interactions between the latent units.
+Then the covariance matrix of between the observed unit is 
+\begin{equation}
+\Sigma = -\left(S_{11} - S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
+\end{equation} 
+The rank of matrix $S_{12}S_{22}^{-1}S_{12}$ equals the number of latent units.
+Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $- S_{12}S_{22}^{-1}S_{12}$ with on low-rank matrix $L_d=- S_{12}S_{22}^{-1}S_{12}$ with rank $d$.
+
+Then we obtain estimator $\hat\Sigma_{d,\rho}$, parameterized by the sparsity parameter $\rho$ and  the number of latent units $d$:
+\begin{equation}
+\begin{split}
+-\hat\Sigma_{d,\rho}^{-1} &= \hat S_\rho + \hat L_d
+\\
+(\hat S_\rho, \hat L_d) &= \arg\min\limits_{S,L} \loss{-(S+L),\Sigma_0} \quad \mbox{such that} \quad \|S\|_0=\rho \quad \mbox{and} \quad \rank(L) = d
+\end{split}
 \end{equation}
 
-If all dependencies are linear, the estimate D describes a group of neurons where each interacts with a small number of latent units and a small fraction of the observed neurons.
-
-\cite{Ma:2013}
+Again, in this formulation, the problem is computationally intractible.
+The problem can be made convex by replacing the $L_0$ norm with $L_1$ norm.
+The resulting algorith, known as \emph{latent-variable graphical lasso} or \emph{LV-GLASSO} \cite{Chandrasekaran:2010,Ma:2013}, produces estimator $\hat\Sigma_{d,\lambda}$:
+\begin{equation}
+\begin{split}
+-\hat\Sigma_{d,\lambda}^{-1} &= \hat S_\lambda + \hat L_d
+\\
+(\hat S_\lambda, \hat L_d) &= \arg\min\limits_{S,L} \loss{-(S+L),\Sigma_0} + \lambda \| S \|_1 \quad \mbox{such that}  \quad \rank(L) = d
+\end{split}
+\end{equation}

### 2013-10-01 sections/Acquisition.tex
updated the Theory section.
redistributed text between files, cleaned up text files.

--- 2013-07-29 sections/Acquisition.tex
@@
-\subsection{Data acquisition}
-Describe the two-photon experiment, refer to Fig.\;1
+\subsection*{Data acquisition}
+We recorded the calcium activity  of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}.
+Numerious repetitions of full-field drifting gratings (Fig.
+1A and 1B) were presented to the eye contralateral to the imaged site.
+This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
+Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.\;1C and 1D).
+The average stimulus response was subtracted to remove stimulus covariance and the sample noise covariance matrix was computed (Fig.\;1E).
+
+\input{sections/Figure1.tex}

### 2013-10-01 sections/Estimators.tex
redistributed text between files, cleaned up text files.

@@
 The problem can be made computationally tractable by replacing the $L_0$ norm with the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.
-The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Banerjee:2008,Friedman:2008}, produces the following estimator
+The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator
 \begin{equation}

### 2013-10-01 sections/Findings.tex
redistributed text between files, cleaned up text files.

--- 2013-09-30 sections/Results.tex
@@
-\subsection*{Data acquisition}
-We recorded the calcium activity  of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}.
-Numerious repetitions of full-field drifting gratings (Fig.
-1A and 1B) were presented to the eye contralateral to the imaged site.
-This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
-Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.\;1C and 1D).
-The average stimulus response was subtracted to remove stimulus covariance and the sample noise covariance matrix was computed (Fig.\;1E).
-
-\input{sections/Figure1.tex}
-
-\subsection*{The sample covariance matrix}
-We aim to estimate the true covariance matrix $\Sigma = \E{(x-\mu)(x-\mu)^\T}$ of the instantaneous activity vector $x$ of a population of $p$ neurons.
-Here $\E{\cdot}$ denotes expectation \TODO{for the true data generating process} and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.
-
-For more rigorous notation, definitions, and derivations, see Appendix.
-
-The usual estimator of $\Sigma$ is the sample covariance matrix
-\begin{equation}
-\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
-\end{equation}
-where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals; $\nu$ is the number of degrees of freedom.
-For independent observations $\nu=n-1$ because estimation of the mean $\mu$ accounts for one degree of freedom.
-When observations are correlated, as is the case with calcium signals, $\nu < n-1$ and may be estimated from the signal.
-
-The sample covariance matrix is constructed to be unbiased, such that $\E{\hat\Sigma_0} = \Sigma$.
-
-\subsection*{Evaluation of covariance matrix estimates}
-The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.
-The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.
-
-For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function:
-\begin{equation}\label{eq:loss}
-\loss{\hat\Sigma,\Sigma} = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
-\end{equation}
-This choice is motivated by mathematical convenience.
-Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$, Stein's entropy loss, and quadratic loss \cite{James:1961,Ledoit:2004,Schafer:2005,Fan:2008}.
-We assert that the main conclusions of our study are unlikely to change drastically other well behaved loss functions.
-
-The aim of our project is produce covariance matrix estimates that minimize the expected loss.  The expected loss of an estimator is known as its \emph{risk}: 
-\begin{equation}
-r = \E{\loss{\hat\Sigma, \Sigma}}
-\end{equation}
-
-In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through validation: 
-Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$. Then \emph{empirical loss} is 
-\begin{equation}
-\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
-\end{equation}
- and its expected value or \emph{empirical risk} is
-\begin{equation}\label{eq:empiricalRisk}
-\hat r = \E{\hat\ell} = \mathbb E_{\hat\Sigma} \left[ \mathbb E_{\hat\Sigma_0^\prime} \left[ {\loss{\hat\Sigma,\hat\Sigma_0^\prime}} \right] \right]
-\end{equation}
-Because the chosen loss function is linear in its second argument in the sense that
-\begin{equation}
-\loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
-\end{equation}
-the expection on the second argument in Eq.~\ref{eq:empiricalRisk} may be taken inside the loss function:
-\begin{equation}
-\hat r = \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}  = \E{\loss{\hat\Sigma,\Sigma}} = r
-\end{equation}
-Thus the empirical loss $\loss{\hat \Sigma,\hat \Sigma_0^\prime}$ serves as an unbiased estimate of risk $r$.
-
-Because the loss function is equivalent to the negative normal log likelihood, the above derivation led us to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes the normal log likelihood of the validation dataset.
-
-\subsection*{Regularization}
-Under many loss functions\footnote{
-The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}.
-However, the equality does hold for its close cousin, Stein's \emph{entropy loss},  which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$.
-This defficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually.
-The bias-variance decomposition is presented here to motivate the use of regularization.},
-the estimator's risk can be decomposed as the sum
-\begin{equation}\label{eq:bias-variance}
-r = b + \varepsilon
-\end{equation}
-of \emph{approximation error} (``bias'' or systematic error)
-\begin{equation}
-b = \loss{\bar\Sigma,\Sigma}
-\end{equation}
-and \emph{estimation error} (``variance'') 
-\begin{equation}
-\varepsilon = \E{\loss{\hat\Sigma, \bar\Sigma}}
-\end{equation}
-where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.
-
-The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thereby minimizes the approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
-
-The estimator risk can be reduced by \emph{regularization}.
-Regularization is the deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
-A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimates aiming to minimize the estimator's risk.
-Various regularization schemes focus on the dimensionality reduction part \TODO{rephrase} by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.
-Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}.
-Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.
-
-\input{sections/Estimators.tex}
-
-\subsection*{In simulation, estimator selection `recognizes' the covariance structure}
-\input{sections/Figure3.tex}
-
 \subsection*{In visual cortex, Estimator D dominates}

### 2013-10-01 sections/Theory.tex
redistributed text between files, cleaned up text files.

@@
-\subsection*{Basic concepts}
-Let $x(t) \in \mathcal X, t=1\ldots,n$ denote a sample of consecutive observations of population activity of $p$ neurons in time bins $t$.  
-The activity of each neuron is denoted by the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
-For example, in our application, the firing rate is estimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x(t)$ to take on negative and non-integer values.
+\subsection*{The sample covariance matrix}
+We aim to estimate the true covariance matrix $\Sigma = \E{(x-\mu)(x-\mu)^\T}$ of the instantaneous activity vector $x$ of a population of $p$ neurons.
+Here $\E{\cdot}$ denotes expectation \TODO{for the true data generating process} and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.
 
-We do not assume that the observations $x(t)$ are independently distributed.
-Rather, the data generating process is assumed to be \emph{ergodic}, \emph{i.e.}\;described by a \emph{true distribution} $F: \mathcal X \mapsto  \mathbb [0, 1]$ (cumulative) over long periods of time such that the \emph{empirical distribution} $\hat F_n$ from a given sample will convege to $F$ with increasing sample size $n$.
+For more rigorous notation, definitions, and derivations, see Appendix.
 
-Formally, the emprical distribution is defined as \TODO{this may be unecessary, but I leave it for now.}
+The usual estimator of $\Sigma$ is the sample covariance matrix
 \begin{equation}
-\hat F_n(x) = \frac 1 n \sum\limits_{t=1}^n \mathbf{1}(x \ge x(t))
+\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
 \end{equation}
-where $\mathbf 1(x \ge x(t))$ is the indicator function which equals 1 when all elements of $x$ are greater than the corresponding elements of $x(t)$ and 0 otherwise.
-\TODO{KJ: Are you using cumulative distributions to avoid binning?
-DY: Yes.
-Many theorems such as Glivenko-Catelli Theorem and Skorohod's Representation are proven using  cumulative representation.
-The empirical distribution is nearly always represented in its cumulative form.
-Expressing convergence is requires binning or using cumulative; the latter is simpler.}
+where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals; $\nu$ is the number of degrees of freedom.
+For independent observations $\nu=n-1$ because estimation of the mean $\mu$ accounts for one degree of freedom.
+When observations are correlated, as is the case with calcium signals, $\nu < n-1$ and may be estimated from the signal.
 
-For an ergodic process, $\max\limits_{x\in\mathcal X} \left|\hat F_n(x) - F(x)\right| \to 0$ as $n$ increases.
-\TODO{If you decide to leave this in, provide a reference to a standard book in the field.}
+The sample covariance matrix is constructed to be unbiased, such that $\E{\hat\Sigma_0} = \Sigma$.
 
-The true covariance matrix $\Sigma \in \Theta$ is defined as a function of $F$:
+\subsection*{Evaluation of covariance matrix estimates}
+The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.
+The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.
+
+For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function:
+\begin{equation}\label{eq:loss}
+\loss{\hat\Sigma,\Sigma} = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
+\end{equation}
+This choice is motivated by mathematical convenience.
+Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$, Stein's entropy loss, and quadratic loss \cite{James:1961,Ledoit:2004,Schafer:2005,Fan:2008}.
+We assert that the main conclusions of our study are unlikely to change drastically other well behaved loss functions.
+
+The aim of our project is produce covariance matrix estimates that minimize the expected loss.  The expected loss of an estimator is known as its \emph{risk}: 
 \begin{equation}
-\Sigma = \int\limits_{z\in\mathcal X} (z - \mu)(z - \mu)^\T \dif F(z)
-\quad\mbox{where}\;
-\mu = \int\limits_{z\in\mathcal X} z \dif F(z)
-\end{equation}
-The domain $\Theta$ is the set of all positive-definite $p\times p$ matrices, which is a cone in $\mathbb R^{p\times(p+1)/2}$.
-
-But the usual estimator of the true covariance matrix $\Sigma$ is  the sample covariance matrix:
-\begin{equation}
-\hat\Sigma_0 =  \frac 1 {n-c} \sum\limits_{t=1}^n (x(t) -\hat\mu) (x(t) - \hat\mu)^\T
-\quad\mbox{where}\;
-\hat\mu = \frac 1 n \sum\limits_{t=1}^n x(t)
-\end{equation}
-where $c$ is the sum of temporal correlations \TODO{there must be a better description of this} \TODO{Which?
-I think the equation is OK.
-Do you mean $c$?
-This is standard, and again, I suggest just providing a reference.}.
-For independent observations, $c=1$;  $c>1$ when nearby samples are correlated; the value of $c$ can be estimated from the data.
-\TODO{Under the Gaussian loss, $\mathcal L_g$ (\ref{eq:GaussLoss}) the scaling is irrelevant, so we will not spend any time talking about the scaling.
-But it would be important if we used $\mathcal L_e$ (\ref{eq:MSE}).}
-
-$\hat\Sigma_0$ is designed to be unbiased such that $\mathbb E\left[\hat\Sigma_0\right]=\Sigma$.
-Here and througout, $\mathbb E[\cdot]$ denotes the expected value under the  unknown true distribution $F$.
-All variables with a hat (\emph{e.g.}\;$\hat \Sigma_0,\hat \mu$) are functions of the empirical distribution $\hat F_n$, which itself is a random variable.
-
-%Nonlinear functions of an unbiased estimate can be biased: $\mathbb E\left[\hat\Sigma_0\right]=\Sigma \centernot\implies  E\left[\varphi(\hat\Sigma_0)\right]=\varphi(\Sigma)$.
-For example, if $u_{\max}(S)$ is the largest eigevalue of square matrix $S$, then $\mathbb E\left[u_{\max}(\hat\Sigma_0)\right] > u_{\max}(\Sigma)$.
-Some covariance matrix estimators are designed to correct for the eigenspectrum bias at the cost  of adding bias to covariance coefficients \cite{Ledoit:2004}.
-
-\subsection*{Loss and risk}
-The optimization of a covariance matrix estimate $\hat\Sigma$ is performed with respect to a \emph{loss function} $\mathcal L(\hat\Sigma,\Sigma)$, which expresses the discrepancy between $\hat\Sigma$ and $\Sigma$ and attains its minimum when $\hat\Sigma=\Sigma$.
-Then \emph{excess loss}  
-\begin{equation}
-\ell(\hat\Sigma,\Sigma) = \mathcal L(\hat\Sigma,\Sigma)-\mathcal L(\Sigma,\Sigma)
-\end{equation}
-assumes zero at its minimum.
-
-A particularly useful loss function is the mean squared error (MSE), which is proportional to the square of the Frobenius  norm $\|\cdot\|_F$ of the difference between the two matrices: 
-\begin{equation}\label{eq:MSE}
-\mathcal L_e(\hat\Sigma,\Sigma) =\frac 1 p \|\hat\Sigma-\Sigma\|_F^2 = \frac 1 p \Tr\left((\hat \Sigma-\Sigma)(\hat\Sigma-\Sigma)^\T\right)
-\end{equation}
-Since $\mathcal L_e(\Sigma,\Sigma) \equiv 0$, the MSE is its own excess loss: $\ell_e(\hat\Sigma,\Sigma) \equiv \mathcal L_e(\hat\Sigma,\Sigma)$.
-
-The Gaussian loss function $\mathcal L_g$  arises from the theory of multivariate normal distributions.
-When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{t=1}^n x(t) x(t)^\T$ is
-\begin{equation}
-L\left(\Sigma \mid \hat\Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
-\end{equation}
-Then $\mathcal L_g$ is constructed by rescaling $L\left(\Sigma \mid \hat\Sigma\right)$ and dropping the constant term:
-\begin{equation}\label{eq:GaussLoss}
-\mathcal L_g(\hat\Sigma,\Sigma) 
-=  -\frac 2 {pn} L\left(\hat\Sigma \mid \Sigma \right) - \frac 1 p \ln(2\pi) 
-\equiv  \frac 1 p\left(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma) \right) 
-\end{equation}
-\TODO{Note that the switch from $L\left(\Sigma \mid \hat\Sigma \right)$ to $L\left(\hat\Sigma \mid \Sigma \right)$ is on purpose, which will become clear in the next few steps.}
-The corresponding excess loss 
-\begin{equation}
-\ell_g(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) - \mathcal L_g(\Sigma,\Sigma)  
-= \frac 1 p \left(-\ln \det (\hat \Sigma^{-1} \Sigma) + \Tr(\hat \Sigma^{-1}\Sigma)\right) - 1
-\end{equation}
-is known as \emph{entropy loss} \cite{James:1961}. \TODO{The order of the operands is reversed. Check that it's okay.}
-
-Despite the fact that entropy loss is derived from normal theory, the choice of a loss function is not equivalent to assuming a specific form of $F$.
-The loss function measures the discrepancy between distribution parameters rather than the distance between the distributions themselves.
-\TODO{explain Bregman divergence?}
-
-The expected value of excess loss is the \emph{estimator risk}:
-\begin{equation}\label{eq:risk}
-r = \mathbb E\left[\ell(\hat\Sigma,\Sigma)\right]
-\end{equation}
-The estimator risk is the primary quality criterion for covariance estimation.
-Estimator $\hat\Sigma_a$ is considered more \emph{efficient} than estimator $\hat\Sigma_b$ if $\hat\Sigma_a$ has lower risk for the given sample than $\hat\Sigma_b$.
-\TODO{explain that the risk depends not only on $\hat\Sigma$ but also on the distribution of $\Sigma$ in the specific domain.}
-
-\subsection*{Cross validation}
-In practice, the true covariance matrix $\Sigma$ is not accessible and the estimator risk $r$ must be estimated from data.
-This requires a separate \emph{validation} empirical distribution $\hat F_m^\prime$ of $m$ observations sampled from $F$ independently of the \emph{training} distribution $\hat F_n$.
-Then let  $\hat \Sigma_0^\prime$ be the sample covariance obtained from $\hat F_m^\prime$.
-
-Then the \emph{empirical loss} of $\hat\Sigma$ is $\mathcal L(\hat\Sigma,\hat\Sigma_0)$ and its expectation is the \emph{empirical risk}  
-\begin{equation}
-\hat r = \mathbb E\left[\mathcal L(\hat\Sigma,\hat\Sigma_0^\prime)\right]
+r = \E{\loss{\hat\Sigma, \Sigma}}
 \end{equation}
 
- The two loss functions $\mathcal L_e(\hat\Sigma,\Sigma)$ and $\mathcal L_g(\hat\Sigma,\Sigma)$ are particularly suitable for risk estimation thanks to their linearity with respect to $\Sigma$ in the sense that
+In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through validation: 
+Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$. Then \emph{empirical loss} is 
 \begin{equation}
-\mathcal L\left(\hat\Sigma,\alpha S_1 + (1-\alpha)S_2\right) 
-\equiv 
-\alpha\mathcal L(\hat \Sigma,S_1) + (1-\alpha)\mathcal L(\hat \Sigma,S_2)
+\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
 \end{equation}
-which allows bringing the expectation inside the empirical loss function:
+ and its expected value or \emph{empirical risk} is
+ \begin{equation}\label{eq:empiricalRisk}
+ \hat r = \E{\hat\ell} = \mathbb E_{\hat\Sigma} \left[ \mathbb E_{\hat\Sigma_0^\prime} \left[ {\loss{\hat\Sigma,\hat\Sigma_0^\prime}} \right] \right]
+ \end{equation}
+ Because the chosen loss function is linear in its second argument in the sense that
  \begin{equation}
-\hat r = 
-\mathbb E\left[ \mathcal L(\hat\Sigma, \hat\Sigma_0^\prime) \right] 
-=
-\mathbb E\left[ \mathcal L\left(\hat\Sigma, \mathbb E\left[\hat\Sigma_0^\prime\right]\right) \right] 
-=
-\mathbb E\left[ \mathcal L(\hat\Sigma, \Sigma) \right] 
-= 
-r + \mathcal L(\Sigma,\Sigma)
+ \loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
  \end{equation}
-This means, that the empirical loss $\mathcal L(\hat\Sigma,\hat\Sigma_0^\prime)$ is an unbiased estimate of the estimator risk $r$ (up to a constant offset): minimization of the empirical loss implies minimization of the true estimator risk.
+ the expection on the second argument in Eq.~\ref{eq:empiricalRisk} may be taken inside the loss function:
+ \begin{equation}
+ \hat r = \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}  = \E{\loss{\hat\Sigma,\Sigma}} = r
+ \end{equation}
+ Thus the empirical loss $\loss{\hat \Sigma,\hat \Sigma_0^\prime}$ serves as an unbiased estimate of risk $r$.
 
-Recording a separate independent validation sample is usually not sensible.
-Instead, the recorded data are split into a training sample and validation sample.
-In $K$-fold \emph{cross-validation} the data are split into $K$ roughly equal-sized subsets.
-Each subset then successively serve as the testing sample while the remainder of the data serves as the training sample.
-This procedure produces $K$ estimates of the  risk, which are then averaged together to produce a better averaged estimate.
+ Because the loss function is equivalent to the negative normal log likelihood, the above derivation led us to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes the normal log likelihood of the validation dataset.
 
-\subsection*{Bias/variance decomposition}
-The estimator risk (\ref{eq:risk}) can be decomposed into \emph{approximation error} or \emph{bias}   
+ \subsection*{Regularization}
+ Under many loss functions\footnote{
+ The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}.
+However, the equality does hold for its close cousin, Stein's \emph{entropy loss},  which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$.
+This defficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually.
+The bias-variance decomposition is presented here to motivate the use of regularization.},
+ the estimator's risk can be decomposed as the sum
+ \begin{equation}\label{eq:bias-variance}
+ r = b + \varepsilon
+ \end{equation}
+ of \emph{approximation error} (``bias'' or systematic error)
  \begin{equation}
-b^2 = \ell \left( \mathbb E[\hat \Sigma],\Sigma\right)
+ b = \loss{\bar\Sigma,\Sigma}
  \end{equation}
-and \emph{estimation error} or \emph{variance}
+ and \emph{estimation error} (``variance'') 
  \begin{equation}
-\varepsilon^2 = \mathbb E \left[ \ell\left(\hat \Sigma, 
-\mathbb E[\hat \Sigma]\right) \right]
+ \varepsilon = \E{\loss{\hat\Sigma, \bar\Sigma}}
  \end{equation}
+ where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.
 
-Under the MSE loss function $\mathcal L_e$ (\ref{eq:MSE}), the decomposition is the simple sum $r = b^2 + \varepsilon^2$. 
-Under other loss functions, $r$ is an increasing function of both $b^2$ and $\varepsilon^2$ although not generally a simple sum.  
-\TODO{As far as I know, the terms `bias' and `variance' are only applicable under a quadratic loss function  like $\mathcal L_e$.
-Under other loss functions, we may need to use approximation/estimation error.}
+ The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thereby minimizes the approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
 
-The sample covariance matrix $\hat\Sigma_0$ has zero bias but high variance. Other estimators may be biased but have lower estimation error and potentially lower risk.
+ The estimator risk can be reduced by \emph{regularization}.
+Regularization is the deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
+A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimates aiming to minimize the estimator's risk.
+Various regularization schemes focus on the dimensionality reduction part \TODO{rephrase} by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.
+Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}.
+Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.

### 2013-10-01 sections/Figure3.tex
updated estimator descriptions.
put all figures and sections in place.

--- 2013-08-09 sections/Simulation.tex
@@
-To illustrate
-
-\begin{figure}
+\begin{figure}[htp]
 \centering
@@
 \caption{
-%\textbf{Figure 3.}{
-Simulation results.
-\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
+In simulation, selection amongst estimators recognizes the true covariance atructure. 
+{\bf A--D} are examples of true $100\times100$ covariance matrices of multivariate normal distributions used in simulations.
+Matrix A has no low-dimensional structure.
+Matrices B--D have low-dimensional structture corresponding to the respective graphical models in Fig.~2B--D.
+{\bf E--H} are examples of sample covariance matrices obtained from samples of size $n=1000$ from distributions with respective true covariance matrices A--D.
+{\bf I--L:} Performance evaluation of covariance estimators A, B, C, D from samples obtained from models A, B, C, D.
+Excess loss $\loss{\hat\Sigma,\Sigma}-\loss{\Sigma,\Sigma}$ is only accessible with the true covariance matrix $\Sigma$ is known and zero excess loss implies absolutely correct estimation.
+Estimators that can represent the graphical model outperform those that cannot.
+Error bars indicate the standard deviation.
+{\bf M--P:} Performance evaluation using cross-validated \emph{empirical loss} from Eq.~\ref{eq:empiricalLoss} without knowledge of true $\Sigma$.
+Cross-validation reproduces the same relationship between performances of the estimators.
 }\label{fig:03}
-%}
 \end{figure}

### 2013-10-01 sections/Figure5.tex
put all figures and sections in place.

--- 2013-09-25 sections/Figure2.tex
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figures/Figure2.pdf}
+\includegraphics[width=1.0\textwidth]{figures/Figure5.pdf}
 \caption{
-Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
-\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies.
-\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
-\textbf{C}: In the Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from sparse pairwise linear interactions between visible units.
-\textbf{D}: In the Gaussian graphical model with latent units, correlations arise  between pairs of nodes
-}\label{fig:02}
+Example of covariance structure revealed by the sparse+low-rank estimator and its relationship to the circuit arcitecture.
+}
+\label{fig:05}
 \end{figure}

### 2013-10-01 sections/Findings.tex
put all figures and sections in place.

@@
 \input{sections/Figure4.tex}
-This is the full matrix of all comparisons of the estimators.
-The plots represents the histograms of the differences of the loss function between the pair of estimators for all 31 sites.
+Figure \ref{fig:04} showthe full matrix of all comparisons of the estimators.
+The plots represents the histograms of the differences of the empirical loss between the pair of estimators for all 31 recorded sites.
 When the differences are positive, the first estimate in the title outperforms the second estimate.

### 2013-10-01 sections/Simulation.tex
minor
expanded the section on covariance estimators
included figure5.
redistributed text between files, cleaned up text files.
put all figures and sections in place.

@@
-To illustrate
+\subsection*{In simulation, estimator selection `recognizes' the covariance structure}
+\input{sections/Figure3.tex}
 
-\begin{figure}
-\centering
-\includegraphics[width=1.0\textwidth]{figures/Figure3.pdf}
-\caption{
-%\textbf{Figure 3.}{
-Simulation results.
-\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
-}\label{fig:03}
-%}
-\end{figure}
+Sample covariance matrices such as E,F,G,H are computed from samples of 1000 independent observations from multivariate normal distributions with true covariance matrices in A,B,C,D, respectively.
+However, the true covariance matrices have various low-dimensional structures.
+
+True matrix A has no low-dimensional structure.  It is computed as the sample covariance matrix of a finite sample (n=1000) of a multivariate Gaussian with independent identically distributed nodes.
+
+True matrix B is computed by factor analysis of true matrix A with six factors.
+
+True matrix C is computed by reducing to zero a large fraction the coefficients of the inverse of true matrix A.
+
+True matrix D is computed by representing the inverse of true matrix A as the sum of a sparse component and a low-low rank component.
+
+When we estimate the true covariance matrix from the samples of various sizes, we find that for true covariance matrices that have a low-dimensional structure, the covariance estimator whose target has the same type of low-dimensional structure attains lower loss.
+Plots I,J,K,L show the excess loss, which equals zero when the estimate equals truth (requires knowledge of truth).
+Plots M,N,O,P show empirical loss computed by cross-validation.
+Empirical loss does not require knowledge of truth but is also shifted so that its absolute value is not particularity meaningful but it can be used for comparison.
+The error bars in these plots indicate the standard deviation, not standard error.
+
+The loss function here is the multivariate log likelihood just as used in the previous post on real data.
+
+The fact that estimate D outperformed the other estimates can be interpreted that the low-dimensional structure of the covariance matrices obtained from dense microcircuits in the visual cortex is best represented by the sparse+low-rank inverse model.
+
+We can be then justified to examine the optimal fitted model rather than the sample covariance model in the analysis of neural correlations.
+The sparse component represents the graph of interactions in the model, which we can now relate to the circuit architecture.
+The low-rank component can be interpreted as common input into the circuit whose properties can be studied separately.
+
+In this way, the process of selecting the best of several regularized covariance estimators serves as the method for recognizing the low-dimensional structure of covariance matrices in a specific domain.
+
+Note that estimator D performs well on both purely sparse and purely low-rank models. But it does not do better.  Only when the truth is sparse+lowrank does it outperform the other estimates.

### 2013-10-01 sections/Regularization.tex
clarified "Evaluation of estimators" based on Kreso's input.

--- 2013-09-30 sections/Results.tex
@@
-\subsection*{Data acquisition}
-We recorded the calcium activity  of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}.
-Numerious repetitions of full-field drifting gratings (Fig.
-1A and 1B) were presented to the eye contralateral to the imaged site.
-This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
-Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.\;1C and 1D).
-The average stimulus response was subtracted to remove stimulus covariance and the sample noise covariance matrix was computed (Fig.\;1E).
-
-\input{sections/Figure1.tex}
-
-\subsection*{The sample covariance matrix}
-We aim to estimate the true covariance matrix $\Sigma = \E{(x-\mu)(x-\mu)^\T}$ of the instantaneous activity vector $x$ of a population of $p$ neurons.
-Here $\E{\cdot}$ denotes expectation \TODO{for the true data generating process} and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.
-
-For more rigorous notation, definitions, and derivations, see Appendix.
-
-The usual estimator of $\Sigma$ is the sample covariance matrix
-\begin{equation}
-\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
-\end{equation}
-where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals; $\nu$ is the number of degrees of freedom.
-For independent observations $\nu=n-1$ because estimation of the mean $\mu$ accounts for one degree of freedom.
-When observations are correlated, as is the case with calcium signals, $\nu < n-1$ and may be estimated from the signal.
-
-The sample covariance matrix is constructed to be unbiased, such that $\E{\hat\Sigma_0} = \Sigma$.
-
-\subsection*{Evaluation of covariance matrix estimates}
-The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.
-The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.
-
-For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function:
-\begin{equation}\label{eq:loss}
-\loss{\hat\Sigma,\Sigma} = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
-\end{equation}
-This choice is motivated by mathematical convenience.
-Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$, Stein's entropy loss, and quadratic loss \cite{James:1961,Ledoit:2004,Schafer:2005,Fan:2008}.
-We assert that the main conclusions of our study are unlikely to change drastically other well behaved loss functions.
-
-The aim of our project is produce covariance matrix estimates that minimize the expected loss.  The expected loss of an estimator is known as its \emph{risk}: 
-\begin{equation}
-r = \E{\loss{\hat\Sigma, \Sigma}}
-\end{equation}
-
-In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through validation: 
-Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$. Then \emph{empirical loss} is 
-\begin{equation}
-\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
-\end{equation}
- and its expected value or \emph{empirical risk} is
-\begin{equation}\label{eq:empiricalRisk}
-\hat r = \E{\hat\ell} = \mathbb E_{\hat\Sigma} \left[ \mathbb E_{\hat\Sigma_0^\prime} \left[ {\loss{\hat\Sigma,\hat\Sigma_0^\prime}} \right] \right]
-\end{equation}
-Because the chosen loss function is linear in its second argument in the sense that
-\begin{equation}
-\loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
-\end{equation}
-the expection on the second argument in Eq.~\ref{eq:empiricalRisk} may be taken inside the loss function:
-\begin{equation}
-\hat r = \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}  = \E{\loss{\hat\Sigma,\Sigma}} = r
-\end{equation}
-Thus the empirical loss $\loss{\hat \Sigma,\hat \Sigma_0^\prime}$ serves as an unbiased estimate of risk $r$.
-
-Because the loss function is equivalent to the negative normal log likelihood, the above derivation led us to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes the normal log likelihood of the validation dataset.
-
 \subsection*{Regularization}
@@
 \begin{equation}
-\varepsilon = \E{\loss{\hat\Sigma, \bar\Sigma}}
+   \varepsilon = \E[\hat\Sigma]{\loss{\hat\Sigma, \bar\Sigma}}
 \end{equation}
@@
 Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.
-
-\input{sections/Estimators.tex}
-
-\subsection*{In simulation, estimator selection `recognizes' the covariance structure}
-\input{sections/Figure3.tex}
-
-\subsection*{In visual cortex, Estimator D dominates}
-\input{sections/Figure4.tex}
-This is the full matrix of all comparisons of the estimators.
-The plots represents the histograms of the differences of the loss function between the pair of estimators for all 31 sites.
-When the differences are positive, the first estimate in the title outperforms the second estimate.
-
-Of particular interest is the first row, which shows that all regularized estimates  outperformed the sample covariance estimate.
-Even more informative is the last column, which shows that estimate D (sparse+lowrank) significantly outperformed all other estimates.

### 2013-10-01 sections/Theory.tex
clarified "Evaluation of estimators" based on Kreso's input.

@@
 
-\subsection*{Evaluation of covariance matrix estimates}
+\subsection*{Evaluation of covariance matrix estimators}
 The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.
@@
 Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$, Stein's entropy loss, and quadratic loss \cite{James:1961,Ledoit:2004,Schafer:2005,Fan:2008}.
-We assert that the main conclusions of our study are unlikely to change drastically other well behaved loss functions.
+We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.
 
-The aim of our project is produce covariance matrix estimates that minimize the expected loss.  The expected loss of an estimator is known as its \emph{risk}: 
+The aim of our project is to produce covariance matrix estimates that minimize the expected loss 
 \begin{equation}
@@
 \end{equation}
+which is know as the \emph{risk} of $\hat\Sigma$.
 
-In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through validation: 
-Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$. Then \emph{empirical loss} is 
+In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through \emph{validation}. 
+Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$.
+
+$\loss{\cdot,\cdot}$ is additive in its second argument such that
+ \begin{equation}\label{eq:additivity}
+ \loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
+ \end{equation}
+Then \emph{validation loss}  \TODO{check appropriate term}
 \begin{equation}
@@
 \end{equation}
- and its expected value or \emph{empirical risk} is
+is an unbiased estimate of the risk:
  \begin{equation}\label{eq:empiricalRisk}
- \hat r = \E{\hat\ell} = \mathbb E_{\hat\Sigma} \left[ \mathbb E_{\hat\Sigma_0^\prime} \left[ {\loss{\hat\Sigma,\hat\Sigma_0^\prime}} \right] \right]
+\E[(\hat\Sigma, \hat\Sigma_0^\prime)] {\hat\ell} 
+= \E[(\hat\Sigma, \hat\Sigma_0^\prime)]{\loss{\hat\Sigma,\hat\Sigma_0^\prime}}
+= \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}
+= \E{\loss{\hat\Sigma,\Sigma}} = r
  \end{equation}
- Because the chosen loss function is linear in its second argument in the sense that
- \begin{equation}
- \loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
- \end{equation}
- the expection on the second argument in Eq.~\ref{eq:empiricalRisk} may be taken inside the loss function:
- \begin{equation}
- \hat r = \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}  = \E{\loss{\hat\Sigma,\Sigma}} = r
- \end{equation}
- Thus the empirical loss $\loss{\hat \Sigma,\hat \Sigma_0^\prime}$ serves as an unbiased estimate of risk $r$.
+Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
 
- Because the loss function is equivalent to the negative normal log likelihood, the above derivation led us to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes the normal log likelihood of the validation dataset.
+Other popular loss functions such as entropy loss do not comply with Eq.~\ref{eq:additivity} and their validation loss is not an unbiased estimate of risk.
 
- \subsection*{Regularization}
- Under many loss functions\footnote{
- The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}.
-However, the equality does hold for its close cousin, Stein's \emph{entropy loss},  which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$.
-This defficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually.
-The bias-variance decomposition is presented here to motivate the use of regularization.},
- the estimator's risk can be decomposed as the sum
- \begin{equation}\label{eq:bias-variance}
- r = b + \varepsilon
- \end{equation}
- of \emph{approximation error} (``bias'' or systematic error)
- \begin{equation}
- b = \loss{\bar\Sigma,\Sigma}
- \end{equation}
- and \emph{estimation error} (``variance'') 
- \begin{equation}
- \varepsilon = \E{\loss{\hat\Sigma, \bar\Sigma}}
- \end{equation}
- where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.
-
- The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thereby minimizes the approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
-
- The estimator risk can be reduced by \emph{regularization}.
-Regularization is the deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
-A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimates aiming to minimize the estimator's risk.
-Various regularization schemes focus on the dimensionality reduction part \TODO{rephrase} by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.
-Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}.
-Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.
+Since $\loss{\cdot,\cdot}$ is equivalent to negative normal log likelihood, the above derivation leads to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes cross-validated normal log likelihood.

### 2013-10-02 sections/Regularization.tex
minor

@@
 The bias-variance decomposition is presented here to motivate the use of regularization.}, 
-the estimator's risk can be decomposed as the sum
+the risk can be decomposed as the sum
 \begin{equation}\label{eq:bias-variance}
@@
 
-The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thereby minimizes the approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
+The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thereby minimizes approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
 

### 2013-10-02 sections/Simulation.tex
minor

@@
-\subsection*{In simulation, estimator selection `recognizes' the covariance structure}
+\subsection*{Simulation: estimators perform better when their target matches the low-dimensional structure of the true covariance matrix}
 \input{sections/Figure3.tex}
+
+A common misconception about regularization is that its effect depends on accurate prior knowlege about the structure of the data.
+However, substantial improvement can be attained by shrinking the unbiased estimate toward an arbitrary target as long as the target is less variable than the unbiased estimator.
+This unintuitive phenomenon is known as \emph{Stein's paradox} \cite{Efron:1977}.
+The more accurate description of regularization is the as the optimal tradeoff between estimation and approximation error, the so-called ``bias-variance tradeoff'' as described above.
+
+Therefore, taken in isolation, the fact that a given regularized estimator improved the estimate should not be interpreted to suggest that its target estimate has the same low-dimensional structure as the data generating process.
+However, when multiple estimators are compared against each other, the ones whose target estimate can closely approximate the true value with the smallest number of parameters will be able to reduce the estimation error with mininimum increases in approximation error when compared against other estimators whose targets cannot consistenly approximate the true covariance matrix.
 

### 2013-10-03 sections/Estimators.tex
minor
finished the Simulation section.

@@
 
-\subsubsection*{Estimator A: Shrinkage toward diagonal.}
-In \emph{estimator A}, the sample covariance matrix $\hat\Sigma_0$ is shrunk toward the independent model.  The covariance matrix of a GGM in which all units are independent will be diagonal.  
+\subsubsection*{Estimator $\mathcal A$: Shrinkage toward diagonal.}
+\emph{Estimator $\mathcal A$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward an independent model of the type show in Fig.~\ref{fig:02}A.
+The covariance matrix of a GGM in which all units are independent will be diagonal.
 For example, we  could leave the sample variances on the diagonal and zero out the off-diagonal products, making the target estimate 
@@
 
-\subsubsection*{Estimate B: Shrinkage toward a factor model}
-\emph{Estimator B} shrinks the sample covariance matrix $\hat\Sigma_0$ toward a target estimated by \emph{factor analysis} with $d$ factors.
+\subsubsection*{Estimate $\mathcal B$: Shrinkage toward a factor model}
+\emph{Estimator $\mathcal B$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward a target estimated by \emph{factor analysis} with $d$ factors, which corresponds to graphical models of the kind in Fig.~\ref{fig:02}B.
 For multivariate normal distributions, factor analysis estimates a graphical model in which the observed units do not interact directly but are influenced by $d$ common inputs.
@@
 
-\subsubsection*{Estimator C: Spare inverse}
-\emph{Estimator C} produces a low-dimensional estimate by finding an approximation of $\hat\Sigma_0$ that has many zeros in its inverse.
+\subsubsection*{Estimator $\mathcal C$: Spare inverse}
+\emph{Estimator $\mathcal C$} produces a low-dimensional estimate by finding an approximation of $\hat\Sigma_0$ that has many zeros in its inverse.
 This approximation problem is known as \emph{covariance selection} \cite{Dempster:1972}.
+Covariance matrices of this family correspond to graphical models of the type depicted in Fig.~\ref{fig:02}C.
 
@@
 
-\subsubsection*{Estimator D: Sparse inverse + low-rank}
-\emph{Estimator D} produces a low-dimensional estimate by finding an apprxomation of $\hat\Sigma_0$ approximating the sample covariance matrix such that its inverse is the sum of a spase matrix and a low-rank matrix.
+\subsubsection*{Estimator $\mathcal D$: Sparse inverse + low-rank}
+\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by finding an apprxomation of $\hat\Sigma_0$ approximating the sample covariance matrix such that its inverse is the sum of a spase matrix and a low-rank matrix.
+Covariance matrices of this form correspond to Gaussian graphical models of the form depicted in Fig.~\ref{fig:02}D.
 
@@
 \end{equation}
+The optimal values of hyperparameters $\lambda$ and $d$ are found cross-validation within the training dataset.

### 2013-10-03 sections/Simulation.tex
finished the Simulation section.

--- 2013-08-09 sections/Simulation.tex
@@
-To illustrate
+\subsection*{Simulation: estimators perform better when their target matches the low-dimensional structure of the true covariance matrix}
+\input{sections/Figure3.tex}
 
-\begin{figure}
-\centering
-\includegraphics[width=1.0\textwidth]{figures/Figure3.pdf}
-\caption{
-%\textbf{Figure 3.}{
-Simulation results.
-\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies. 
-}\label{fig:03}
-%}
-\end{figure}
+Which covariance estimator is better is an empirical question.  The answer depends profoundly on the properties of the data generating system under investigation.
+
+A common misconception about regularization is that its effect depends on accurate prior knowlege about the structure of the data.
+However, substantial improvement can be attained by shrinking the unbiased estimate toward an arbitrary target as long as the target is less variable than the unbiased estimator.
+This unintuitive phenomenon is known as \emph{Stein's phenomenon} or \emph{Stein's paradox} \cite{Efron:1977} named after its discoverer Charles Stein \cite{Stein:1956}.
+The more accurate description of regularization is as of the optimal tradeoff between estimation and approximation error, the so-called ``bias-variance tradeoff'' as described in earlier sections. 
+Taken in isolation, the fact that a given regularized estimator improves the estimate should not be interpreted to suggest that its target estimate has the same low-dimensional structure as the data generating process.
+However, when multiple estimators are compared against each other, the ones whose target estimate can closely approximate the true value with the smallest number of parameters will be able to reduce the estimation error with mininimum increases in approximation error when compared against other estimators whose targets cannot consistenly approximate the true covariance matrix.
+
+To illustrate the convergence of the four regularized covariance matrix estimators,  we constructed several kinds of covariance matrices to be used as the true covariance matrices for multivariate normal distributions to be used in simulations studies.
+Four of such true matrices, sized $100\times100$ are shown in Fig.~\ref{fig:03} A--D.
+It is computed as the sample covariance matrix of a finite sample (n=1000) drawn from a multivariate normal distribution with independent units and unit variances.
+True matrix A has no low-dimensional structure: it takes $p\times(p+1)/2=5,050$ parameters to be fully defined.
+True matrix B is computed by factor analysis from matrix A with $d=6$ factors.
+As such, it has a low dimensional structure: it is fully defined with $p\times d +p = 700$ parameters.
+True matrix C is computed by reducing to zero a large fraction the coefficients of the inverse of true matrix A \TODO{using GLASSO}.
+In this instance, 64.6\% of its off-diagonal elements are zero, and the entire matrix is fully defined with 1,825 parameters.
+Finally, true matrix D is an approximation of $A$ such that its inverse is a sum of a sparse matrix (74.1\% zeros of the digonal) and a low-rank component with rank $d=3$, with the number of parameters totaling 1582.
+
+The true matrices B, C, and D correspond to Gaussian Graphical Models of the families depicted in Fig.~\ref{fig:02} B, C, and D, respectively.
+\TODO{Note that we do not have a true table that corresponds to the graphical model in Fig.~\ref{fig:02} simply because it is not of practical interest.}
+
+Panels E, F, G, and H show sample covariance matrices computed from samples of 1000 independent observations from multivariate normal distributions with true covariance matrices in A, B, C, and D, respectively.
+
+Plots I, J, K, L show the quality of covariance matrix computed by the sample covariance matrix $\hat\Sigma_0$ as well as the four regularized covariance estimators $\mathcal A$, $\mathcal B$, $\mathcal C$, and $\mathcal D$.
+Since we have access to the true covariance matrices, we use \emph{excess loss} $\loss{\hat\Sigma,\Sigma}-\loss{\Sigma,\Sigma}$ to evaluate the quality of the estimate $\hat\Sigma$.
+Excess loss approaches zero when the estimate coverges to the true value.
+The error bars indicate the standard deviation (rather than standard error) for multiple samples drawn from the same true table.
+
+All five covariance matrix estimators are \emph{consistent}: they converge to the true value as the sample size increases.
+All regularized estimators dominate the sample covariance matrix independent of the low-dimensional structure of truth.
+This is an example of Stein's phenomenon.
+The regularized estimators are not equally \emph{efficient}: for a given sample size, some covariance estimators produce consistently lower loss than others.
+Which estimator is most efficient depends on which estimator best matchest the low-dimensional structure of the true covariance matrix.
+True covariance A does not have a low-dimensional structure and all four regularized estimators ($\mathcal A$,$\mathcal B$,$\mathcal C$,$\mathcal D$) perform similarly.
+However, in cases when the true covariance matrix has a low-dimensional structure, the esimator whose target estimate can match that structure dominates the others.
+Thus for truth B, estimator $\mathcal B$ domainates; for truth C, estimator $\mathcal C$ dominates; and for truth D,  estimator $\mathcal D$ dominates.
+Note, however, that estimator $\mathcal D$'s target can accurately approximate the low dimensional structure of truths B and C and performed very closely to their matching estimators.
+\TODO{This does not mean however that we should design a ``universal'' estimator that can capture all kinds of low-dimensional structure because (a) it will fall for false patterns and (b) it will have a high number of hyperparameters, which will become intractable}.
+
+Plots M, N, O, and P reproduce the same comparison as in I, J, K, and L but use validation loss from Eq.~\ref{eq:validationLoss} computed by 10-fold cross-validation.
+This analysis is performed without access to the true covariance matrix and can thus be applied to empirical data.
+The results are qualitatively comparable to those obtained with excess loss.
+Because validation loss is computed by comparing estimates to noisy sample covariance matrices from smaller validation set, validation loss is a much noisier measurement than excess loss.
+In addition, validation loss converges to the unknown value $\loss{\Sigma,\Sigma}$ rather than zero.

### 2013-10-03 sections/Theory.tex
minor
finished the Simulation section.

@@
 Then \emph{validation loss}  \TODO{check appropriate term}
-\begin{equation}
+\begin{equation}\label{eq:validationLoss}
 \hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
@@
 
-Since $\loss{\cdot,\cdot}$ is equivalent to negative normal log likelihood, the above derivation leads to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes cross-validated normal log likelihood.
+Since $\loss{\cdot,\cdot}$ is equivalent to negative normal log likelihood, the above derivation has led to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes cross-validated normal log likelihood.

### 2013-10-03 sections/Estimators.tex
math corrections in Estimators.tex

@@
 \begin{equation}
-\hat\Sigma_{\alpha,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha 
+\hat\Sigma_{\alpha,\lambda}^\mathcal{A} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha 
 \end{equation}
@@
 \begin{equation}
-\hat\Sigma_{d,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d
+\hat\Sigma_{d,\lambda}^\mathcal{B} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d
 \end{equation}
@@
 \begin{equation}
--\hat\Sigma_\rho^{-1} =\arg\min\limits_S \loss{-S^{-1},\hat\Sigma_0}  \quad \mbox{such that} \quad  \|S_\rho\|_0 = \rho
+\hat\Sigma_\rho = \hat S^{-1}
+\quad\mbox{with}\quad
+\hat S = \argmin\limits_{\substack{S\succ 0 \\ \|S\|_0 = \rho}} \loss{S^{-1},\hat\Sigma_0}   
 \end{equation}
-where $S_\rho$ is a sparse matrix with $\rho$ non-zero coefficients. $\|S\|_0$ denotes the number of non-zero elements of matrix $S$.
+where $S\succ 0$ signifies the constraint that $S$ be positive-definite. $\|S\|_0=\rho$ signifies the constraint  that matrix $S$ have $\rho$ non-zero coefficients.
 Solving this approximation in this form is computationally challenging.
-The problem can be made computationally tractable by replacing the $L_0$ norm with the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.
+The problem can be made computationally tractable by relaxing the $L_0$ norm to the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.
 The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator
 \begin{equation}
--\hat\Sigma_\lambda^{-1} = \arg\min\limits_S \loss{-S^{-1},\hat\Sigma_0} + \lambda \|S\|_1
+\hat\Sigma_\lambda^\mathcal{C} = \hat S^{-1}
+\quad\mbox{with}\quad
+\hat S = \argmin\limits_{S\succ 0} \loss{S^{-1},\hat\Sigma_0} + \lambda \|S\|_1
 \end{equation}
@@
 
-The inverse $S=-\Sigma^{-1}$ of the covariance matrix of a GGM constitutes the graph of conditional dependencies (interactions) between the units.
+The inverse $S=\Sigma^{-1}$ of the covariance matrix of a GGM constitutes the graph of conditional dependencies (interactions) between the units.
 With the assumption of sparse organization of interactions in the network, $S$ is assumed to be sparse.
@@
 \begin{equation}
-\Sigma = -\left(S_{11} - S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
+\Sigma = \left(S_{11} - S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
 \end{equation} 
 The rank of matrix $S_{12}S_{22}^{-1}S_{12}$ equals the number of latent units.
-Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $- S_{12}S_{22}^{-1}S_{12}$ with on low-rank matrix $L_d=- S_{12}S_{22}^{-1}S_{12}$ with rank $d$.
+Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $ S_{12}S_{22}^{-1}S_{12}$ with on low-rank matrix $L_d= S_{12}S_{22}^{-1}S_{12}$ with rank $d$.
 
@@
 \begin{equation}
-\begin{split}
--\hat\Sigma_{d,\rho}^{-1} &= \hat S_\rho + \hat L_d
-\\
-(\hat S_\rho, \hat L_d) &= \arg\min\limits_{S,L} \loss{-(S+L),\Sigma_0} \quad \mbox{such that} \quad \|S\|_0=\rho \quad \mbox{and} \quad \rank(L) = d
-\end{split}
+\hat\Sigma_{d,\rho} = \left(\hat L-\hat S\right)^{-1}
+\quad\mbox{where}\quad
+(\hat S, \hat L) = \argmin\limits_{\substack{S,L\\ (L-S)\succ 0 \\ \|S\|_0=\rho \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(L-S)^{-1},\Sigma_0}
 \end{equation}
@@
 The problem can be made convex by replacing the $L_0$ norm with $L_1$ norm.
-The resulting algorith, known as \emph{latent-variable graphical lasso} or \emph{LV-GLASSO} \cite{Chandrasekaran:2010,Ma:2013}, produces estimator $\hat\Sigma_{d,\lambda}$:
+The resulting algorith, known as \emph{latent-variable graphical lasso} or \emph{LV-GLASSO} \cite{Chandrasekaran:2010,Ma:2013}, produces estimator $\hat\Sigma_{d,\lambda}^{\mathcal D}$:
 \begin{equation}
-\begin{split}
--\hat\Sigma_{d,\lambda}^{-1} &= \hat S_\lambda + \hat L_d
-\\
-(\hat S_\lambda, \hat L_d) &= \arg\min\limits_{S,L} \loss{-(S+L),\Sigma_0} + \lambda \| S \|_1 \quad \mbox{such that}  \quad \rank(L) = d
-\end{split}
+\hat\Sigma_{d,\lambda} ^{\mathcal D} = \left(\hat L-\hat S\right)^{-1}
+\quad\mbox{where}\quad
+(\hat S, \hat L) = \argmin\limits_{\substack{ S,L \\ (L-S)\succ 0 \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(L-S)^{-1},\Sigma_0} + \lambda \| S \|_1 
 \end{equation}

### 2013-10-03 plos.tex
minor
added Figure1.pdf
ongoing changes in the Results section
minor edits
worked on the description of estimators.
updated estimator descriptions.
completed estimator description.
redistributed text between files, cleaned up text files.
put all figures and sections in place.
clarified "Evaluation of estimators" based on Kreso's input.
math corrections in Estimators.tex

@@
 \DeclareMathOperator{\Tr}{tr}
+\DeclareMathOperator*{\argmin}{arg\,min}
 \newcommand{\mcond}{\,\middle\vert\,}
 \newcommand{\cond}{\,\vert\,}
+\newcommand{\loss}[1]{\mathcal L\left(#1\right)} 
 \newcommand{\T}{\intercal}
+\newcommand{\E}[2][]{\mathbb E_{#1}\left[ #2\right]}    % expected value
 \newcommand{\TODO}[1]{\emph{\small\color{blue}$\langle\langle$#1$\rangle\rangle$}}
 \newcommand*\dif{\mathop{}\,d}
+\DeclareMathOperator{\rank}{rank}
 
@@
 \\
-\bf{2} Max Planck Institute of Biological Cybernetics, T\"ubingen, Germany
+\bf{2} Department of Mathematics and Department of Biology and Biochemistry, University of Houston, Houston, TX, USA
 \\
-\bf{3} Department of Mathematics and Department of Biology and Biochemistry, University of Houston, Houston, TX, USA
+\bf{3} Max Planck Institute of Biological Cybernetics, T\"ubingen, Germany
 \\
@@
 \section*{Results}
-\input{sections/Results.tex}
+\input{sections/Acquisition.tex}
+\input{sections/Theory.tex}
+\input{sections/Regularization.tex}
+\input{sections/Estimators.tex}
+\input{sections/Simulation.tex}
+\input{sections/Findings.tex}
+\input{sections/Structure.tex}
 
 \section*{Discussion}
-%\input{sections/Discussion.tex}
+\input{sections/Discussion.tex}
 
@@
 \section*{Materials and Methods}
+\input{sections/Methods.tex}
 
@@
 \section*{Figure Legends}
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=4in]{figures/Figure1.pdf}
-\end{center}
-\caption{
-{\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\tt A.} Visual stimuli comprising full-field  
-}
-\label{Figure_label}
-\end{figure}
+\TODO{presently embedded in the text}
 

### 2013-10-03 sections/Theory.tex
minor

@@
 This choice is motivated by mathematical convenience.
-Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$, Stein's entropy loss, and quadratic loss \cite{James:1961,Ledoit:2004,Schafer:2005,Fan:2008}.
+Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, and quadratic loss \cite{James:1961,Fan:2008}.
 We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.

### 2013-10-04 sections/Figure5.tex
added contents to Fig. 1 and Fig. 5

@@
 Example of covariance structure revealed by the sparse+low-rank estimator and its relationship to the circuit arcitecture.
+{\bf A} The correlation matrix corresponding to the sample covariance matrix in Fig.~\ref{fig:01}D.
+{\bf B} The inverse of the sample covariance matrix (sample concentration matrix).
+{\bf D} shows the result of covariance matrix estimation by estimator $\mathcal D$ and {\bf E} show its inverse.
+The sparse ({\bf C}) and low-rank ({\bf F}) components of the inverse.
+{\bf G} depicts the spatial and functional organization of the connections represented by the sparse component of the estimated concentration matrix.
 }

### 2013-10-04 sections/Estimators.tex
minor

@@
 
-Just as in estimator A, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$.
+Just as in estimator $\mathcal A$, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$.
 \begin{equation}
@@
 \quad\mbox{with}\quad
-\hat S = \argmin\limits_{\substack{S\succ 0 \\ \|S\|_0 = \rho}} \loss{S^{-1},\hat\Sigma_0}   
+\hat S = \argmin\limits_{\substack{S\succ 0 \\ \|S\|_0 \le \rho}} \loss{S^{-1},\hat\Sigma_0}   
 \end{equation}
-where $S\succ 0$ signifies the constraint that $S$ be positive-definite. $\|S\|_0=\rho$ signifies the constraint  that matrix $S$ have $\rho$ non-zero coefficients.
+where $S\succ 0$ signifies the constraint that $S$ be positive-definite. $\|S\|_0\le\rho$ signifies the constraint  that matrix $S$ have at most $\rho$ non-zero coefficients.
 Solving this approximation in this form is computationally challenging.
-The problem can be made computationally tractable by relaxing the $L_0$ norm to the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.
+The problem can be made computationally tractable by relaxing \TODO{use "complex relaxation"} the $L_0$ norm to the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.
 The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator
@@
 
-\subsubsection*{Estimator $\mathcal D$: Sparse inverse + low-rank}
-\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by finding an apprxomation of $\hat\Sigma_0$ approximating the sample covariance matrix such that its inverse is the sum of a spase matrix and a low-rank matrix.
+\subsubsection*{Estimator $\mathcal D$: Sparse + low-rank inverse}
+\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by approximating $\hat\Sigma_0$ with a matrix whose inverse is the sum of a spase matrix and a low-rank matrix.
 Covariance matrices of this form correspond to Gaussian graphical models of the form depicted in Fig.~\ref{fig:02}D.
@@
 The rank of matrix $S_{12}S_{22}^{-1}S_{12}$ equals the number of latent units.
-Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $ S_{12}S_{22}^{-1}S_{12}$ with on low-rank matrix $L_d= S_{12}S_{22}^{-1}S_{12}$ with rank $d$.
+Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $ S_{12}S_{22}^{-1}S_{12}$ with one low-rank matrix $L= S_{12}S_{22}^{-1}S_{12}$.
 
@@
 \begin{equation}
-\hat\Sigma_{d,\rho} = \left(\hat L-\hat S\right)^{-1}
+\hat\Sigma_{d,\rho} = \left(\hat S-\hat L\right)^{-1}
 \quad\mbox{where}\quad
-(\hat S, \hat L) = \argmin\limits_{\substack{S,L\\ (L-S)\succ 0 \\ \|S\|_0=\rho \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(L-S)^{-1},\Sigma_0}
+(\hat S, \hat L) = \argmin\limits_{\substack{S,L\\ (S-L)\succ 0 \\ \|S\|_0\le\rho \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0}
 \end{equation}
 
-Again, in this formulation, the problem is computationally intractible.
+Again, in this formulation, the approximation problem is computationally intractible.
 The problem can be made convex by replacing the $L_0$ norm with $L_1$ norm.
@@
 \begin{equation}
-\hat\Sigma_{d,\lambda} ^{\mathcal D} = \left(\hat L-\hat S\right)^{-1}
+\hat\Sigma_{d,\lambda} ^{\mathcal D} = \left(\hat S-\hat L\right)^{-1}
 \quad\mbox{where}\quad
-(\hat S, \hat L) = \argmin\limits_{\substack{ S,L \\ (L-S)\succ 0 \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(L-S)^{-1},\Sigma_0} + \lambda \| S \|_1 
+(\hat S, \hat L) = \argmin\limits_{\substack{ S,L \\ (S-L)\succ 0 \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0} + \lambda \| S \|_1 
 \end{equation}

### 2013-10-04 plos.tex
Kreso's edits

@@
 
+\newcommand{\Kcomment}[1]{{\color{blue}{[KJ: #1]}}}
+
 \DeclareMathOperator{\Tr}{tr}
-\DeclareMathOperator*{\argmin}{arg\,min}
 \newcommand{\mcond}{\,\middle\vert\,}
@@
 \\
-\bf{2} Department of Mathematics and Department of Biology and Biochemistry, University of Houston, Houston, TX, USA
+\bf{2} Max Planck Institute of Biological Cybernetics, T\"ubingen, Germany
 \\
-\bf{3} Max Planck Institute of Biological Cybernetics, T\"ubingen, Germany
+\bf{3} Department of Mathematics and Department of Biology and Biochemistry, University of Houston, Houston, TX, USA
 \\

### 2013-10-04 sections/Estimators.tex
Kreso's edits

--- 2013-10-03 sections/Estimators.tex
@@
 Gaussian graphical models have a straightforward relationship to their covariance matrix $\Sigma$:  Zero elements in the inverse of $\Sigma$ indicate conditional independence between the corresponding pair of variables.
+\Kcomment{It may be important to note that we are not fitting Gaussian graphical models, however.}
 
@@
 \subsubsection*{Estimator $\mathcal A$: Shrinkage toward diagonal.}
-\emph{Estimator $\mathcal A$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward an independent model of the type show in Fig.~\ref{fig:02}A.
+\emph{Estimator $\mathcal A$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward that of the independent model shown in Fig.~\ref{fig:02}A.
 The covariance matrix of a GGM in which all units are independent will be diagonal.
-For example, we  could leave the sample variances on the diagonal and zero out the off-diagonal products, making the target estimate 
+For example, we  could leave the sample variances on the diagonal set the off-diagonal terms to zero. The target estimate
+is then 
 \begin{equation}\label{eq:sampleVariance}
@@
 where $\circ$ denotes the entrywise matrix product (Hadamard product). 
-Alternatively, the target could have equal variances, all equal to the average sample variance in the population $v = \frac 1 p \Tr(\hat\Sigma_0)$.
+Alternatively, the target could be a diagonal matrix with all variances equal to the average sample variance, $v = \frac 1 p \Tr(\hat\Sigma_0)$, so that
 \begin{equation}\label{eq:equalVariance}
@@
 It is not immediately clear which target is better in a particular application.
-We therefore use the linear mixture of the two targets paramaterized by the mixing coefficient $\alpha\in[0,1]$:
+We therefore use the linear mixture of the two targets paramaterized by the mixing coefficient $\alpha\in[0,1]$,
 \begin{equation}
-\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I
+\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I.
 \end{equation}
-Then the estimator $\hat\Sigma_{\alpha,\lambda}$ is the linear shrinkage of the sample covariance matrix $\hat\Sigma_0$ toward $\hat T_\alpha$ controlled by the mixing proportion $\lambda\in[0,1]$:
+The estimator $\hat\Sigma_{\alpha,\lambda}$ is the linear shrinkage of the sample covariance matrix $\hat\Sigma_0$ toward $\hat T_\alpha$ controlled by the mixing proportion $\lambda\in[0,1]$,
 \begin{equation}
-\hat\Sigma_{\alpha,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha 
+\hat\Sigma_{\alpha,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha.
 \end{equation}
-This covariance matrix estimator is implemented by R's {\tt corpcor} package \cite{Schaefer:2010} and enjoys widespread use in many applications \cite{Schafer:2005}.
+This covariance matrix estimator is implemented by R's {\tt corpcor} package \cite{Schaefer:2010} and is used widely~\cite{Schafer:2005}.
 
@@
 \TODO{more detail?} \TODO{discuss analytic optimization of shrinkage intensities.}
+
+\Kcomment{Yes -- you give no details about the cross-validation.  You should provide some details at least in 
+one case, or have a separate section that discusses how you implemented it.}
 
@@
 where $\hat L_d$ is a $p\times p$ matrix of rank $d$ and $\hat \Psi$ is a $p\times p$ diagonal matrix. The value of $\hat T_d$ is found by minimizing $\loss{\hat T_d, \hat\Sigma_0}$. 
+\Kcomment{I would give a more explicit representation of $\hat L_d$ as a product to show that the factors can be understood as common input.}
 
-Just as in estimator A, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$.
+As with estimator A, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$,
 \begin{equation}
-\hat\Sigma_{d,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d
+\hat\Sigma_{d,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d,
 \end{equation}
@@
 
-Factor analysis has a rich history of testing hypotheses about the dominance of a few common inputs \cite{findagoodone}.
+Factor analysis has long been used to test hypotheses about the dominance of a few common inputs \cite{findagoodone}.
 In neuroscience, the hypothesis that the population activity in small local circuits is largely driven by a few latent factors has been investigated \cite{Yu:2009,Ecker:2013}.
@@
 
-For Gaussian Graphical Models, zeros in the inverse covariance matrix indicate statisitcal independence between the corresponding pairs.
+For Gaussian Graphical Models, zeros in the inverse covariance matrix indicate statistical independence between the corresponding pairs.
 Thus the inverse covariance matrix $S=-\Sigma^{-1}$ is also the adjacency matrix of the GGM, indicating which pairs are conditionally dependent.
-\TODO{discuss how quickly (or slowly) this breaks down with deviations from gaussianity, e.g.~MaxEnt models, Ising models.} \TODO{Also, motivate this from the perpsective of linear regression as in \cite{Varoquaux:2012}.} The off-diagonal elements of $S$ are proportional to partial pairwise correlations between corresponding pairs of neurons.
+\TODO{discuss how quickly (or slowly) this breaks down with deviations from gaussianity, e.g.~MaxEnt models, Ising models.} \TODO{Also, motivate this from the perpsective of linear regression as in \cite{Varoquaux:2012}.}  \Kcomment{This is a bit confusing -- in the first sentence, you have statistical independence, in the second, you talk about conditional independence.
+You want to emphasize that a 0 inverse means that -- in the GGM -- the two nodes only ``talk'' indirectly through other nodes.} The off-diagonal elements of $S$ are proportional to partial pairwise correlations between corresponding pairs of neurons.
 \TODO{Expand.} Regularization of covariance matrix estimation by sparsifying the inverse of the estimate has been using in function genomics \cite{Schafer:2005,other} and fMRI functional connectivity \cite{Varoquaux:2012}.
@@
 \end{equation}
-where $\lambda \|S\|_1$ is the $L_1$ norm of the matrix. \TODO{clarify}.
+where $\lambda \|S\|_1$ is the $L_1$ norm of the matrix. \TODO{clarify}. \Kcomment{Yes - at least define the norm, if you are already taking about it. No need to expand a lot more, though.}
 
 Just as in the previous estimators, the optimal value of the regularization parameter $\lambda$ is found from training data by cross-validation.
+\Kcomment{I would say this up front, so you don't have to repeat it.
+Again, I think more details about cross-validation should be given.}
 
 \subsubsection*{Estimator $\mathcal D$: Sparse inverse + low-rank}
-\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by finding an apprxomation of $\hat\Sigma_0$ approximating the sample covariance matrix such that its inverse is the sum of a spase matrix and a low-rank matrix.
+\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by finding an approximation of $\hat\Sigma_0$ approximating the sample covariance matrix such that its inverse is the sum of a spase matrix and a low-rank matrix.
 Covariance matrices of this form correspond to Gaussian graphical models of the form depicted in Fig.~\ref{fig:02}D.
@@
 The rank of matrix $S_{12}S_{22}^{-1}S_{12}$ equals the number of latent units.
-Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $- S_{12}S_{22}^{-1}S_{12}$ with on low-rank matrix $L_d=- S_{12}S_{22}^{-1}S_{12}$ with rank $d$.
+Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $- S_{12}S_{22}^{-1}S_{12}$ with a low-rank matrix $L_d=- S_{12}S_{22}^{-1}S_{12}$ with rank $d$.
 

### 2013-10-04 sections/Regularization.tex
Kreso's edits

@@
 
-The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thereby minimizes approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
+An unbiased estimator, such as the sample covariance,  $\hat\Sigma_0,$ makes $\bar\Sigma=\Sigma$ and thereby minimizes approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
 
-The estimator risk can be reduced by \emph{regularization}.
-Regularization is the deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
-A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimates aiming to minimize the estimator's risk.  
+The estimator risk can be reduced by \emph{regularization} -- a deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
+A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimate to minimize the estimator's risk.  
 Various regularization schemes focus on the dimensionality reduction part \TODO{rephrase} by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.

### 2013-10-04 sections/Theory.tex
Kreso's edits

--- 2013-10-03 sections/Theory.tex
@@
 
-For more rigorous notation, definitions, and derivations, see Appendix.
+For detailed explanation of the notation, definitions, and derivations, see Appendix.
 
@@
 
-The sample covariance matrix is constructed to be unbiased, such that $\E{\hat\Sigma_0} = \Sigma$.
+The sample covariance matrix is constructed to be unbiased, so that $\E{\hat\Sigma_0} = \Sigma$.
 
@@
 The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.
+\Kcomment{Could we say that it quantifies the error in the estimate?}
 
@@
 \end{equation}
-which is know as the \emph{risk} of $\hat\Sigma$.
+which is known as the \emph{risk} of $\hat\Sigma$.
 
-In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through \emph{validation}. 
+In practice, the true value $\Sigma$ is not accessible and an estimators' risks must be estimated from the data.
+This may be accomplished through \emph{validation}.
+\Kcomment{If you will discuss validation later, point to it.}
 Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$.
 
-$\loss{\cdot,\cdot}$ is additive in its second argument such that
+$\loss{\cdot,\cdot}$ is additive in its second argument so that \Kcomment{You frequently use ``such that'' instead of ``so that''}
  \begin{equation}\label{eq:additivity}
@@
 Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
+\Kcomment{This is probably not a very insightful question -- but why do we assume that the sample covariance
+is what we compare $\hat\Sigma$ when we validate the results.  Our point is that this estimate is not very good.}
 

### 2013-10-05 sections/Estimators.tex

@@
 \begin{equation}
-\hat\Sigma_{\alpha,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha.
+\hat\Sigma_{\alpha,\lambda}^\mathcal{A} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha.
 \end{equation}
@@
 
-As with estimator A, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$,
+Just as with estimator $\mathcal A$, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$,
 \begin{equation}
-\hat\Sigma_{d,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d,
+\hat\Sigma_{d,\lambda}^\mathcal{B} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d,
 \end{equation}
@@
 \begin{equation}
--\hat\Sigma_\rho^{-1} =\arg\min\limits_S \loss{-S^{-1},\hat\Sigma_0}  \quad \mbox{such that} \quad  \|S_\rho\|_0 = \rho
+\hat\Sigma_\rho = \hat S^{-1}
+\quad\mbox{with}\quad
+\hat S = \argmin\limits_{\substack{S\succ 0 \\ \|S\|_0 \le \rho}} \loss{S^{-1},\hat\Sigma_0}   
 \end{equation}
-where $S_\rho$ is a sparse matrix with $\rho$ non-zero coefficients. $\|S\|_0$ denotes the number of non-zero elements of matrix $S$.
+where $S\succ 0$ signifies the constraint that $S$ be positive-definite. $\|S\|_0\le\rho$ signifies the constraint  that matrix $S$ have at most $\rho$ non-zero coefficients.
 Solving this approximation in this form is computationally challenging.
-The problem can be made computationally tractable by replacing the $L_0$ norm with the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.
+The problem can be made computationally tractable by relaxing \TODO{use "complex relaxation"} the $L_0$ norm to the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.
 The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator
 \begin{equation}
--\hat\Sigma_\lambda^{-1} = \arg\min\limits_S \loss{-S^{-1},\hat\Sigma_0} + \lambda \|S\|_1
+\hat\Sigma_\lambda^\mathcal{C} = \hat S^{-1}
+\quad\mbox{with}\quad
+\hat S = \argmin\limits_{S\succ 0} \loss{S^{-1},\hat\Sigma_0} + \lambda \|S\|_1
 \end{equation}
@@
 
-\subsubsection*{Estimator $\mathcal D$: Sparse inverse + low-rank}
-\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by finding an approximation of $\hat\Sigma_0$ approximating the sample covariance matrix such that its inverse is the sum of a spase matrix and a low-rank matrix.
+\subsubsection*{Estimator $\mathcal D$: Sparse + low-rank inverse}
+\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by approximating $\hat\Sigma_0$ with a matrix whose inverse is the sum of a spase matrix and a low-rank matrix.
 Covariance matrices of this form correspond to Gaussian graphical models of the form depicted in Fig.~\ref{fig:02}D.
 
-The inverse $S=-\Sigma^{-1}$ of the covariance matrix of a GGM constitutes the graph of conditional dependencies (interactions) between the units.
+The inverse $S=\Sigma^{-1}$ of the covariance matrix of a GGM constitutes the graph of conditional dependencies (interactions) between the units.
 With the assumption of sparse organization of interactions in the network, $S$ is assumed to be sparse.
@@
 \begin{equation}
-\Sigma = -\left(S_{11} - S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
+\Sigma = \left(S_{11} - S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
 \end{equation} 
 The rank of matrix $S_{12}S_{22}^{-1}S_{12}$ equals the number of latent units.
-Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $- S_{12}S_{22}^{-1}S_{12}$ with a low-rank matrix $L_d=- S_{12}S_{22}^{-1}S_{12}$ with rank $d$.
+Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $ S_{12}S_{22}^{-1}S_{12}$ with a low-rank matrix $L= S_{12}S_{22}^{-1}S_{12}$.
 
@@
 \begin{equation}
-\begin{split}
--\hat\Sigma_{d,\rho}^{-1} &= \hat S_\rho + \hat L_d
-\\
-(\hat S_\rho, \hat L_d) &= \arg\min\limits_{S,L} \loss{-(S+L),\Sigma_0} \quad \mbox{such that} \quad \|S\|_0=\rho \quad \mbox{and} \quad \rank(L) = d
-\end{split}
+\hat\Sigma_{d,\rho} = \left(\hat S-\hat L\right)^{-1}
+\quad\mbox{where}\quad
+(\hat S, \hat L) = \argmin\limits_{\substack{S,L\\ (S-L)\succ 0 \\ \|S\|_0\le\rho \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0}
 \end{equation}
 
-Again, in this formulation, the problem is computationally intractible.
+Again, in this formulation, the approximation problem is computationally intractible.
 The problem can be made convex by replacing the $L_0$ norm with $L_1$ norm.
-The resulting algorith, known as \emph{latent-variable graphical lasso} or \emph{LV-GLASSO} \cite{Chandrasekaran:2010,Ma:2013}, produces estimator $\hat\Sigma_{d,\lambda}$:
+The resulting algorith, known as \emph{latent-variable graphical lasso} or \emph{LV-GLASSO} \cite{Chandrasekaran:2010,Ma:2013}, produces estimator $\hat\Sigma_{d,\lambda}^{\mathcal D}$:
 \begin{equation}
-\begin{split}
--\hat\Sigma_{d,\lambda}^{-1} &= \hat S_\lambda + \hat L_d
-\\
-(\hat S_\lambda, \hat L_d) &= \arg\min\limits_{S,L} \loss{-(S+L),\Sigma_0} + \lambda \| S \|_1 \quad \mbox{such that}  \quad \rank(L) = d
-\end{split}
+\hat\Sigma_{d,\lambda} ^{\mathcal D} = \left(\hat S-\hat L\right)^{-1}
+\quad\mbox{where}\quad
+(\hat S, \hat L) = \argmin\limits_{\substack{ S,L \\ (S-L)\succ 0 \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0} + \lambda \| S \|_1 
 \end{equation}

### 2013-10-05 sections/Figure2.tex

@@
 \textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
-\textbf{C}: In the Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from sparse pairwise linear interactions between visible units.
-\textbf{D}: In the Gaussian graphical model with latent units, correlations arise  between pairs of nodes
+\textbf{C}: Graphical model with conditional dependencies between some pairs of observed neurons, no latent units. 
+\textbf{D}: Graphical model with conditional dependencies between some pairs of observed neurons and dense interactions with a few latent units.
 }\label{fig:02}

### 2013-10-05 sections/Introduction.tex
Added Author Summary

@@
 Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically.
-For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \cite{Hayes:1981}.
+For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \cite{Ledoit:2004}.
 Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.
@@
 	\item Devise more efficient estimators of neural covariance matrices in recordings from large populations of neurons.
-	\item Aid interpretation of population activity  by isolating the essential structure of the neural covariance matrix.
+	\item Aid interpretation of neural covariance matrices.
 \end{enumerate}

### 2013-10-16 sections/Simulation.tex

@@
 However, substantial improvement can be attained by shrinking the unbiased estimate toward an arbitrary target as long as the target is less variable than the unbiased estimator.
-This unintuitive phenomenon is known as \emph{Stein's phenomenon} or \emph{Stein's paradox} \cite{Efron:1977} named after its discoverer Charles Stein \cite{Stein:1956}.
+This counter--intuitive phenomenon is known as \emph{Stein's phenomenon} or \emph{Stein's paradox} \cite{Efron:1977} named after its discoverer Charles Stein \cite{Stein:1956}.
 The more accurate description of regularization is as of the optimal tradeoff between estimation and approximation error, the so-called ``bias-variance tradeoff'' as described in earlier sections. 
@@
 However, when multiple estimators are compared against each other, the ones whose target estimate can closely approximate the true value with the smallest number of parameters will be able to reduce the estimation error with mininimum increases in approximation error when compared against other estimators whose targets cannot consistenly approximate the true covariance matrix.
+\Kcomment{This seems a bit too wordy for the main part of the text.  I think the same can be said in a couple sentences.
+People will likely wonder what this has to do with the analysis you are doing, if you spend too much time on generalities.}
 
@@
 Four of such true matrices, sized $100\times100$ are shown in Fig.~\ref{fig:03} A--D.
-It is computed as the sample covariance matrix of a finite sample (n=1000) drawn from a multivariate normal distribution with independent units and unit variances.
+\Kcomment{What does the ``It'' in the following part of the text refer to?
+Do you mean ``Each''?} It is computed as the sample covariance matrix of a finite sample (n=1000) drawn from a multivariate normal distribution with independent units and unit variances.
 True matrix A has no low-dimensional structure: it takes $p\times(p+1)/2=5,050$ parameters to be fully defined.

### 2013-10-22 sections/Estimators.tex
updated figure 5.

@@
 
-Factor analysis has long been used to test hypotheses about the dominance of a few common inputs \cite{findagoodone}.
+Factor analysis has long been used in natural sciences to infer the influence of a one or several latent variables exerting linear effects on a larger population of observed variables \cite{Jvreskog:1996}.
 In neuroscience, the hypothesis that the population activity in small local circuits is largely driven by a few latent factors has been investigated \cite{Yu:2009,Ecker:2013}.
@@
 You want to emphasize that a 0 inverse means that -- in the GGM -- the two nodes only ``talk'' indirectly through other nodes.} The off-diagonal elements of $S$ are proportional to partial pairwise correlations between corresponding pairs of neurons.
-\TODO{Expand.} Regularization of covariance matrix estimation by sparsifying the inverse of the estimate has been using in function genomics \cite{Schafer:2005,other} and fMRI functional connectivity \cite{Varoquaux:2012}.
+\TODO{Expand.} Regularization of covariance matrix estimation by sparsifying the inverse of the estimate has been using in functional genomics \cite{Schafer:2005,Peng:2009} and fMRI functional connectivity \cite{Varoquaux:2012,Ryali:2012}.
 

### 2013-10-22 sections/Figure1.tex
updated figure 5.

@@
 \begin{center}
-\includegraphics[width=2.5in]{figures/Figure1.pdf}
+\includegraphics[width=0.75\textwidth]{figures/Figure1.pdf}
 \end{center}
@@
 {\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\bf A.} Visual stimuli comprising brief (500 ms) presentatios of full-field drifting gratings separated by blank screens.
-{\bf B.} Two-photon fast 3D imaging of calcium signals in an awake mouse.
-{\bf C.} Deconvolved calcium signals.
-{\bf D.} The sample covariance matrix of neuronal calcium signals in 200 ms time bins.
+{\bf B.} Two-photon fast 3D imaging of calcium signals.
+{\bf C.} The first and last 30 seconds of a 30-minute recording of population calcium signals.
+The signals are deconvovled and binned at 150 ms.
+{\bf D.} The sample noise correlation matrix of neuronal calcium signals from {\bf C} after subtracting the average stimulus response.
+{\bf E.} The hIstogram of off-diagonal coefficients from the correlation matrix in {\bf D}.
 }
-\label{Figure_label}
+\label{fig:01}
 \end{figure}

### 2013-10-22 sections/Figure5.tex
updated figure 5.

--- 2013-10-01 sections/Figure5.tex
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/Figure5.pdf}
+\includegraphics[width=0.5\textwidth]{figures/Figure5.pdf}
 \caption{
 Example of covariance structure revealed by the sparse+low-rank estimator and its relationship to the circuit arcitecture.
+{\bf A:} The correlation matrix from Fig.~\ref{fig:01}D. 
+{\bf B:} Optimal cross-validated sparse+low-rank estimate of the corrleation matrix from the same dataset. 
+{\bf C:} Scaled negative inverse of the correlation matrix in {\bf A} (partial pairwise correlations). 
+{\bf D:} Negative inverse of the sparse+low-rank estimate in {\bf B}. 
+{\bf E:} The sparse component of {\bf D} with 96.5\% off-diagonal coefficients set to zero.
+{\bf F:} The low-rank component of {\bf D}, rank=7.
+{\bf G:} The spatial arrangement of cells used in the analysis, connected by edges that correspond to the non-zero elements of the sparse component in {\bf E}.
+Green edges denote positive weights, red edges denote negative weights.
+Color-filled cells are significantly tuned to grating orientation as indicated by the color code in the legend.
 }

### 2013-10-24 sections/Figure0.tex
added Figure 6.

\begin{figure}[htp]
\begin{center}
\includegraphics[width=0.5\textwidth]{figures/Figure0.pdf}
\end{center}
\caption{
{\bf An illustration of regularized estimaton of the covariance matrix.}  \TODO{generate a smaller and simple example}
{\bf A:} The sample covariance matrix of population calcium signals of 204 cells.
{\bf B:} The sample precison matrix (inverse of the sample covariance matrix in {\bf B}.
{\bf C:} A estimate of the covariance matrix produced by bounding the $L_1$ norm of the precision matrix. 
{\bf D:} The precision matrix of the estimate in {\bf C}.
}
\label{fig:00}
\end{figure}

### 2013-10-24 sections/Figure4.tex
redid figure 4 to show only estimator D.
added Figure 6.

@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/Figure4.pdf}
-\caption{Estimator D dominates over the other estimators in estimation of noise covariances in dense recordings from primary visual cortex.
+\includegraphics[width=0.5\textwidth]{figures/Figure4.pdf}
+\caption{{\bf Estimator $\mathcal D$ (sparse+full-rank inverse covariance) dominates in dense recordings from primary visual cortex.}
+{\bf A:}  Histogram of improvements in validation loss by covariance estimator $\mathcal D$ relative to the samle covariance matrix from the 31 imaged sites used in the study.
+The vertical red line denotes the median improvment.
+{\bf B:} Improvements relative to estimator $\mathcal A$ (shrinkage toward the diagonal).
+{\bf C:} Improvements relative to estimator $\mathcal B$ (shrinkage toward the factor model).
+{\bf D:} Improvements relative to estimator $\mathcal C$ (sparse inverse covariance).
 }\label{fig:04}

### 2013-10-25 plos.tex
Revised results to be more concise. Removed details still need to be put back into the methods. Fixed my name & affiliations

--- 2013-10-03 plos.tex
@@
 
+\newcommand{\Kcomment}[1]{{\color{blue}{[KJ: #1]}}}
+\newcommand{\Acomment}[1]{{\color{red}{[AE: #1]}}}
+
 \DeclareMathOperator{\Tr}{tr}
-\DeclareMathOperator*{\argmin}{arg\,min}
 \newcommand{\mcond}{\,\middle\vert\,}
@@
 \newcommand{\loss}[1]{\mathcal L\left(#1\right)} 
-\newcommand{\T}{\intercal}
+\newcommand{\T}{{\rm T}}
 \newcommand{\E}[2][]{\mathbb E_{#1}\left[ #2\right]}    % expected value
@@
 \newcommand*\dif{\mathop{}\,d}
+\DeclareMathOperator*{\argmin}{arg\,min}
 \DeclareMathOperator{\rank}{rank}
@@
 \\
-Dimitri Yatsenko$^{1,\ast}$, 
+Dimitri Yatsenko,$^{1,\ast}$, 
 Kre\v{s}imir Josi\'{c}$^{2}$,
-Alexander Ecker$^{3,1}$,
-Emmanouil Foudarakis$^{1}$,
+Alexander S.~Ecker$^{1,3,4}$,
+Emmanouil Froudarakis$^{1}$,
 R.~James Cotton$^{1}$,
-Andreas S.~Tolias$^{1}$
+Andreas S.~Tolias$^{1,5}$
 \\
@@
 \\
-\bf{3} Max Planck Institute of Biological Cybernetics, T\"ubingen, Germany
+\bf{3}  Werner Reichardt Center for Integrative Neuroscience and Institute for Theoretical Physics, University of T\"ubingen, Germany
 \\
+\bf{4} Bernstein Center for Computational Neuroscience, T\"ubingen, Germany
+\\
+\bf{5} Department of Computational and Applied Mathematics, Rice University, Houston, TX, USA
+
 $\ast$ E-mail: yatsenko@cns.bcm.edu
@@
 \section*{Results}
-\input{sections/Acquisition.tex}
-\input{sections/Theory.tex}
-\input{sections/Regularization.tex}
-\input{sections/Estimators.tex}
-\input{sections/Simulation.tex}
-\input{sections/Findings.tex}
-\input{sections/Structure.tex}
+
+\input{sections/Results.tex}
+
+%\input{sections/Acquisition.tex}
+%\input{sections/Theory.tex}
+%\input{sections/Regularization.tex}
+%\input{sections/Estimators.tex}
+%\input{sections/Simulation.tex}
+%\input{sections/Findings.tex}
+%\input{sections/Structure.tex}
 

### 2013-10-25 sections/Results.tex
Revised results to be more concise. Removed details still need to be put back into the methods. Fixed my name & affiliations

@@
-\subsection*{Data acquisition}
+\subsection*{Covariance estimation}
+
+Our goal is to estimate the covariance matrix, $\Sigma$, of a population of $n$ neurons.
+For small populations or in the limit of infinite data this can be done in a straightforward manner by using the sample covariance estimator
+\begin{equation}
+\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T, 
+\end{equation}
+where $x(t)$ are sequential observations of population activity and $\nu$ is the number of degrees of freedom ($\nu=n-1$ if observations are independent).
+
+However, for large population sizes and limited data this approach becomes ineffective at capturing the true structure of the covariance matrix (REFS).
+The number of parameters to be estimated grows quadratically with the population size.
+Thus, to obtain a "good" estimate of the true covariance, we need to rely on regularization -- the deliberate biasing of the estimate toward a low-dimensional, less variable target estimate \cite{Bickel:2006,Ledoit:2004}.
+Various such targets exist, each expressing a different idea about what a parsimonious model would look like.
+
+We considered four different estimators, based on a few simple concepts for regularizing the covariance: independence, low-rank structure and sparse interactions.
+We optimized all four estimators with respect to Gaussian log-likelihood (see Discussion for a more in-depth justification of this loss function) and used cross-validation to determine hyper-parameters.
+
+The first and probably simplest regularized estimator is shrinkage towards a diagonal matrix
+\begin{equation}
+\hat \Sigma_{\rm diag} = (1-\lambda) \hat \Sigma_0 + \lambda D,
+\end{equation}
+where $D$ is a diagonal matrix and the mixing proportion, $\lambda$, is determined by cross-validation.
+This estimator expresses the idea that in the absence of evidence for strong correlations between cells we assume they are independent (Fig.~1A).
+This approach is sometimes also referred to as ridge regression.
+
+The second estimator we consider is shrinkage towards a low-rank matrix
+\begin{equation}
+\hat \Sigma_{\rm low-rank} = (1-\lambda) \hat \Sigma_0 + \lambda T,
+\end{equation}
+where $T = LL^\T + \Psi$ is a factor analysis model (REFS) consisting of a low-rank component $L$ with rank $d < n$ and a diagonal matrix $\Psi$.
+This estimator expresses the idea that the fluctuations in population activity are driven by a small number of unobserved sources that affect many cells (Fig.~1B).
+
+The third estimator is based on sparsifying the inverse covariance (precision) matrix.
+It expresses the idea that, similar to synaptic connections, interactions between neurons are sparse (Fig.~1C).
+The estimator produces a low-dimensional approximation to the covariance that has many zeros in its inverse.
+This approximation problem is also known as covariance selection.
+The covariance matrix estimate is given by
+\begin{equation}
+\hat\Sigma = \hat S^{-1}
+\quad\mbox{with}\quad
+\hat S = \argmin\limits_{\|S\|_0 \le \rho} \loss{S^{-1},\hat\Sigma_0}   
+\end{equation}
+where $\mathcal L$ is a loss function (see below) and $\|S\|_0\le\rho$ signifies the constraint that $S$ has at most $\rho$ non-zero coefficients.
+While solving the problem in this form is computationally challenging, relaxing the $L_0$ norm to the $L_1$ norm converts it into one of convex optimization \cite{Donoho:2000}, which is computationally much more tractable.
+The resulting algorithm, which is known as graphical lasso \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator
+\begin{equation}
+\hat\Sigma_{\rm sparse}^\mathcal{C} = \hat S^{-1}
+\quad\mbox{with}\quad
+\hat S = \argmin \loss{S^{-1},\hat\Sigma_0} + \lambda \|S\|_1,
+\end{equation}
+where $\|S\|_1$ is the $L_1$ norm of the precision matrix (see Methods for details) and $\lambda$ is again determined by cross-validation.
+
+Finally, we consider a fourth estimator, which combines common inputs with sparse interactions (Fig.~1D). For this estimator the covariance is
+\begin{equation}
+\hat\Sigma_{\rm combined} = (S + LL^\T)^{-1},
+\end{equation}
+where, as above, $S$ is sparsified using $L_1$ norm and $L$ is a low-rank matrix with the rank determined by cross-validation.
+
+\subsection*{Simulation using toy data}
+
+To verify our approach and to illustrate the performance of our four regularized estimators and the sample covariance estimator, we constructed five model populations with different underlying structures.
+Each population contained 100~neurons.
+The first four populations matched the low-dimensional structure of our four estimators: independent (Fig.~2\,A), low-rank (Fig.~2\,B), sparse inverse (Fig.~2\,C) and low-rank combined with sparse inverse (Fig.~2\,D).
+\Acomment{Integrate schematics from Fig.~1 into this figure (make it the first row) for better correspondence} In addition, we considered a fifth population that had no low-dimensional structure (Fig.~2\,E).
+
+To evaluate the performance of the different estimators, we computed the excess loss for all combinations of model populations and estimators, including the sample covariance.
+The first striking observation is that in all cases all four regularized estimators performed substantially better than the sample covariance (Fig.~2, fourth row).
+In particular, this is even true for the case where the population did not have any low-dimensional structure at all (Fig.~2\,E).
+While it may appear surprising at first sight, this counter-intuitive phenomenon is known as \emph{Stein's phenomenon} or \emph{Stein's paradox} \cite{Efron:1977}, named after its discoverer Charles Stein \cite{Stein:1956}.
+A common misconception about regularization is that its effect depends on accurate prior knowledge about the structure of the data.
+However, substantial improvement can be attained by shrinking the unbiased estimate toward an arbitrary target as long as the target is less variable than the unbiased estimator.
+The more accurate description of regularization is as of the optimal tradeoff between estimation and approximation error -- the so-called ``bias-variance tradeoff''.
+
+Thus, if taken in isolation a regularized estimator improves the estimate we should not interpret this result to suggest that the estimator's target has the same low-dimensional structure as the data-generating process.
+However, when comparing multiple estimators against each other, the one whose target estimate most closely matches the true value with the smallest number of parameters will reduce the estimation error with the least increase in approximation error.
+Indeed, in all four toy examples with a low-dimensional structure, the estimator with the matching regularization target performed best (Fig.~2\,A--D, fourth row).
+Since the estimator combining low-rank and sparse interactions combines to low-dimensional targets and includes the simpler estimators as special cases, it performed almost as well even when the low-dimensional structure did not include either a low-rank component or sparse interactions (Fig.~2\,B,\,C).
+\Acomment{May want to include a smaller sample size as well since here the more reduced estimators (B, C) may actually outperform estimator D when the population structure matches their target and there is little data (say, 1.5x number of neurons)}
+
+Since the above evaluations of excess loss require knowledge of ground truth, this analysis can be done only on toy data.
+However, an unbiased empirical estimate of the excess loss exists (see Methods, Eq.~\ref{eq:validationLoss} \Acomment{This may need a brief sentence about the intuition what's the difference, or a reference to the relevant methods section}).
+To verify that our approach should also work in a realistic situation without access to ground truth, we also computed the empirical loss.
+Indeed, without access to ground truth, this analysis revealed a pattern of results that reproduced the results obtained with access to ground truth above (Fig.~2, last row).
+Because validation loss is computed by comparing estimates to noisy sample covariance matrices from smaller validation sets, empirical loss is a much noisier measurement than excess loss.
+In addition, it does not converge to zero with increasing sample size as excess loss does.
+
+\subsection*{Covariance estimation on neural data}
+
 We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}.
-Numerious repetitions of full-field drifting gratings (Fig.
-1A and 1B) were presented to the eye contralateral to the imaged site.
-This technique allowed recording from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
-Somatic calcium signals were deconvolved using  sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.\;1C and 1D).
-The average stimulus response was subtracted to remove stimulus covariance and the sample noise covariance matrix was computed (Fig.\;1E).
+We presented numerous repetitions of full-field drifting gratings (Fig.
+1A and 1B) to the eye contralateral to the imaged site.
+This technique allowed us to record from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
+We deconvolved somatic calcium signals using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.\;1C and 1D) and subtracted the average stimulus response to remove effects of the stimulus.
+From this residual response we computed the sample noise covariance matrix (Fig.\;1E).
 
-\input{sections/Figure1.tex}
+In such highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability.
+At the same time, most correlations are relatively small (Fig.~1\,E), suggesting that a simple shrinkage towards independence may provide a sufficiently well regularized estimate.
+By comparing the performance of our differently regularized estimators we can gain insights into which of these aspects are important in our data.
 
-\subsection*{The sample covariance matrix}
-We aim to estimate the true covariance matrix $\Sigma = \E{(x-\mu)(x-\mu)^\T}$ of the instantaneous activity vector $x$ of a population of $p$ neurons.
-Here $\E{\cdot}$ denotes expectation \TODO{for the true data generating process} and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.
+As expected, all regularized estimators outperformed the sample covariance estimator substantially (Fig.~4) \Acomment{I think we should show that}.
+Say something about how shrinkage, low-rank and sparse inverse relate \Acomment{Why do we do pairwise comparisons instead of just showing median (across sites) log loss relative to the combined estimator?
+I've seen many people do this and it provides an ordering.
+I feel like we should come up with a way of ordering them somehow, even if it's under certain assumptions that may not be entirely correct.
+I'm pretty sure if we don't do it the reviewers will bring it up anyway...}.
+Finally, the combined sparse and low-rank estimator dominated all others significantly (Fig.~4), showing that both hidden units and direct interactions are important in our data.
+The improvement from the low-rank estimator to the combined one is much larger than that from the sparse inverse to the combined one, suggesting that in this dataset direct interactions contribute more strongly to the correlation structure than hidden units do.
 
-For more rigorous notation, definitions, and derivations, see Appendix.
+\subsection*{Relationship between functional covariance structure and circuit architecture}
 
-The usual estimator of $\Sigma$ is the sample covariance matrix
-\begin{equation}
-\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
-\end{equation}
-where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals; $\nu$ is the number of degrees of freedom.
-For independent observations $\nu=n-1$ because estimation of the mean $\mu$ accounts for one degree of freedom.
-When observations are correlated, as is the case with calcium signals, $\nu < n-1$ and may be estimated from the signal.
+\begin{itemize}
+\item Linear and partial correlations versus spatial separation (lateral and vertical). Discuss whether to include or remove thresholded correlations.
+\item Magnitude of common input versus spatial location in the volume (lateral and vertical separation from center).
+\item Linear and partial correlations versus orientation preference
+\item Distribution of sparsity over sites
+\item Distribution of number of hidden units over sites
+\item Distribution of eigenvalues for sample covariance versus combined estimator
+\item etc. etc.
 
-The sample covariance matrix is constructed to be unbiased, such that $\E{\hat\Sigma_0} = \Sigma$.
-
-\subsection*{Evaluation of covariance matrix estimates}
-The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.
-The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.
-
-For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function:
-\begin{equation}\label{eq:loss}
-\loss{\hat\Sigma,\Sigma} = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
-\end{equation}
-This choice is motivated by mathematical convenience.
-Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$, Stein's entropy loss, and quadratic loss \cite{James:1961,Ledoit:2004,Schafer:2005,Fan:2008}.
-We assert that the main conclusions of our study are unlikely to change drastically other well behaved loss functions.
-
-The aim of our project is produce covariance matrix estimates that minimize the expected loss.  The expected loss of an estimator is known as its \emph{risk}: 
-\begin{equation}
-r = \E{\loss{\hat\Sigma, \Sigma}}
-\end{equation}
-
-In practice, the true value $\Sigma$ is not accessible and estimators' risks must be estimated from the data.  This may be accomplished through validation: 
-Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$. Then \emph{empirical loss} is 
-\begin{equation}
-\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
-\end{equation}
- and its expected value or \emph{empirical risk} is
-\begin{equation}\label{eq:empiricalRisk}
-\hat r = \E{\hat\ell} = \mathbb E_{\hat\Sigma} \left[ \mathbb E_{\hat\Sigma_0^\prime} \left[ {\loss{\hat\Sigma,\hat\Sigma_0^\prime}} \right] \right]
-\end{equation}
-Because the chosen loss function is linear in its second argument in the sense that
-\begin{equation}
-\loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
-\end{equation}
-the expection on the second argument in Eq.~\ref{eq:empiricalRisk} may be taken inside the loss function:
-\begin{equation}
-\hat r = \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}  = \E{\loss{\hat\Sigma,\Sigma}} = r
-\end{equation}
-Thus the empirical loss $\loss{\hat \Sigma,\hat \Sigma_0^\prime}$ serves as an unbiased estimate of risk $r$.
-
-Because the loss function is equivalent to the negative normal log likelihood, the above derivation led us to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes the normal log likelihood of the validation dataset.
-
-\subsection*{Regularization}
-Under many loss functions\footnote{
-The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}.
-However, the equality does hold for its close cousin, Stein's \emph{entropy loss},  which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$.
-This defficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually.
-The bias-variance decomposition is presented here to motivate the use of regularization.},
-the estimator's risk can be decomposed as the sum
-\begin{equation}\label{eq:bias-variance}
-r = b + \varepsilon
-\end{equation}
-of \emph{approximation error} (``bias'' or systematic error)
-\begin{equation}
-b = \loss{\bar\Sigma,\Sigma}
-\end{equation}
-and \emph{estimation error} (``variance'') 
-\begin{equation}
-\varepsilon = \E{\loss{\hat\Sigma, \bar\Sigma}}
-\end{equation}
-where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.
-
-The unbiased estimator $\hat\Sigma_0$ makes $\bar\Sigma=\Sigma$ and thereby minimizes the approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
-
-The estimator risk can be reduced by \emph{regularization}.
-Regularization is the deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
-A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimates aiming to minimize the estimator's risk.
-Various regularization schemes focus on the dimensionality reduction part \TODO{rephrase} by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.
-Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}.
-Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.
-
-\input{sections/Estimators.tex}
-
-\subsection*{In simulation, estimator selection `recognizes' the covariance structure}
-\input{sections/Figure3.tex}
-
-\subsection*{In visual cortex, Estimator D dominates}
-\input{sections/Figure4.tex}
-This is the full matrix of all comparisons of the estimators.
-The plots represents the histograms of the differences of the loss function between the pair of estimators for all 31 sites.
-When the differences are positive, the first estimate in the title outperforms the second estimate.
-
-Of particular interest is the first row, which shows that all regularized estimates  outperformed the sample covariance estimate.
-Even more informative is the last column, which shows that estimate D (sparse+lowrank) significantly outperformed all other estimates.
+\end{itemize}

### 2013-10-29 sections/Figure5.tex
added Figure 6.
refined Alex's revision of the results sections -- to be continued.

@@
 {\bf A:} The correlation matrix from Fig.~\ref{fig:01}D. 
-{\bf B:} Optimal cross-validated sparse+low-rank estimate of the corrleation matrix from the same dataset. 
-{\bf C:} Scaled negative inverse of the correlation matrix in {\bf A} (partial pairwise correlations). 
-{\bf D:} Negative inverse of the sparse+low-rank estimate in {\bf B}. 
-{\bf E:} The sparse component of {\bf D} with 96.5\% off-diagonal coefficients set to zero.
+{\bf B:} Partial pairwise correlations (scaled negative inverse of the correlation matrix in {\bf A}) 
+{\bf C:} Optimal cross-validated sparse+latent estimate of the corrleation matrix from the same dataset. 
+{\bf D:} Negative inverse of the sparse+latent estimate in {\bf C}. 
+{\bf E:} The sparse component of {\bf D} with 96.5\% off-diagonal coefficients set to zero with the average of 5.1 edges per cell.
 {\bf F:} The low-rank component of {\bf D}, rank=7.

### 2013-10-29 sections/Figure6.tex
refined Alex's revision of the results sections -- to be continued.

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure6.pdf}
\caption{
Relationship between the structure of the estimate and the functional and spatial organization of the circuit.
{\bf A:} Average sample correlation, average partial sample correlation, and average partial correlation in the sparse+low-rank estimate for site in Fig.~\ref{fig:05} as a function of difference in cortical depth within the same column.
{\bf C:} Average sample correlation, average partial sample correlation, and average partial correlation in the sparse+low-rank estimate for site in Fig.~\ref{fig:05} as a function of lateral distance at the same cortical depth.
{\bf E,G:} Average sample correlation, and interaction probability obtained from the sparse+low-rank estimate as well as by thresholding the sample correlation matrix to equal sparsity as function of difference in cortical depnth and lateral distance.
{\bf I:} Average sample correlation, and interaction probability obtained from the sparse+low-rank estimate as well as by thresholding the sample correlation matrix to equal sparsity as a function of difference in preffered orientation.
}
\label{fig:06}
\end{figure}

### 2013-10-29 sections/Results.tex
refined Alex's revision of the results sections -- to be continued.

@@
 \subsection*{Covariance estimation}
+We aim to estimate the true covariance matrix 
+\begin{equation}
+\Sigma = \E{(x-\mu)(x-\mu)^\T}
+\end{equation}
+where $\E{\cdot}$ denotes expectation  and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.
 
-Our goal is to estimate the covariance matrix, $\Sigma$, of a population of $n$ neurons.
-For small populations or in the limit of infinite data this can be done in a straightforward manner by using the sample covariance estimator
+The usual estimator of the covariance matrix is the \emph{sample covariance matrix} computed from the empirical sample  $x(1),\ldots,x(n)$:
 \begin{equation}
@@
 
-However, for large population sizes and limited data this approach becomes ineffective at capturing the true structure of the covariance matrix (REFS).
-The number of parameters to be estimated grows quadratically with the population size.
-Thus, to obtain a "good" estimate of the true covariance, we need to rely on regularization -- the deliberate biasing of the estimate toward a low-dimensional, less variable target estimate \cite{Bickel:2006,Ledoit:2004}.
-Various such targets exist, each expressing a different idea about what a parsimonious model would look like.
+The sample covariance matrix is unbiased i.e.\;$\E{\hat\Sigma_0}-\Sigma=0$.
+For finite sample sizes, however, $\hat\Sigma_0$ is not as close to $\Sigma$ as a number of biased estimators that rely on \emph{regularization}.
+Regularization is the deliberate biasing of the estimate toward a low-dimensional, less variable \emph{target estimate} to strike a favorable balance between bias and variability of the estimate \cite{Bickel:2006,Ledoit:2004}.
+The estimator whose target estimate most closely matches the true low-dimensional structure of the data is likely to outperform other estimators.
 
-We considered four different estimators, based on a few simple concepts for regularizing the covariance: independence, low-rank structure and sparse interactions.
-We optimized all four estimators with respect to Gaussian log-likelihood (see Discussion for a more in-depth justification of this loss function) and used cross-validation to determine hyper-parameters.
+We considered four regularized estimators based on distinct families of low-dimensional target estimates: independent, latent factors, sparse partial correlations, and sparse partial correlations with latent factors.
+These target estimates are depicted graphically in Fig.~\ref{fig:02}.
 
-The first and probably simplest regularized estimator is shrinkage towards a diagonal matrix
+\input{sections/Figure2.tex}
+
+The relative performance of the four estimators on synthetic and empirical data was evaluated by cross-validated multivariate normal log-likelihood (see Methods for detailed justification).  
+Each estimator has one or two \emph{hyperparameters}, which are estimated within training datasets by nested cross-validation (See Methods).
+
+In the first regularized estimator, the target estimate is a diagonal matrix $\hat D$.
+The optimized estimate $\hat\Sigma_{\rm diag}$ is obtained by the linear \emph{shrinkage} of the unbiased estimate $\hat\Sigma_0$ toward $\hat D$ as regulated by scalar \emph{shrinkage intensity} $\lambda \in [0, 1]$:
 \begin{equation}
-\hat \Sigma_{\rm diag} = (1-\lambda) \hat \Sigma_0 + \lambda D,
+\hat \Sigma_{\rm diag} = (1-\lambda) \hat \Sigma_0 + \lambda \hat D
 \end{equation}
-where $D$ is a diagonal matrix and the mixing proportion, $\lambda$, is determined by cross-validation.
-This estimator expresses the idea that in the absence of evidence for strong correlations between cells we assume they are independent (Fig.~1A).
-This approach is sometimes also referred to as ridge regression.
+The diagonal target estimate expresses the idea of lack of dependence (or of linear association) between the activity of observed neurons (Fig.~\ref{fig:02}A).
+If this assumption aptly describes recorded data, then strong shrinkage toward $\hat D$ will add little bias while strongly reducing the variability of the estimate.
 
-The second estimator we consider is shrinkage towards a low-rank matrix
+In the second regularized estimator, the target estimate is the factor model $\hat F = LL^\T + \Psi$ with $d$ factors so that $L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\Psi$ contains the indvidual variances of each neuron.
+Then the estimate is 
 \begin{equation}
-\hat \Sigma_{\rm low-rank} = (1-\lambda) \hat \Sigma_0 + \lambda T,
+\hat \Sigma_{\rm factor} = (1-\lambda) \hat \Sigma_0 + \lambda \hat F,
 \end{equation}
-where $T = LL^\T + \Psi$ is a factor analysis model (REFS) consisting of a low-rank component $L$ with rank $d < n$ and a diagonal matrix $\Psi$.
-This estimator expresses the idea that the fluctuations in population activity are driven by a small number of unobserved sources that affect many cells (Fig.~1B).
+This estiamtor has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
+The target estimate $\hat F$ expresses the idea that the fluctuations in population activity must be driven by relatively small number of unobserved sources that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}B).
 

### 2013-10-29 sections/Theory.tex
refined Alex's revision of the results sections -- to be continued.

@@
-\subsection*{The sample covariance matrix}
-We aim to estimate the true covariance matrix $\Sigma = \E{(x-\mu)(x-\mu)^\T}$ of the instantaneous activity vector $x$ of a population of $p$ neurons.
-Here $\E{\cdot}$ denotes expectation \TODO{for the true data generating process} and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.
-
-For detailed explanation of the notation, definitions, and derivations, see Appendix.
-
-The usual estimator of $\Sigma$ is the sample covariance matrix
-\begin{equation}
-\hat \Sigma_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
-\end{equation}
-where $x(t),\;t=1,\ldots,n$ are sequential observations of population activity inferred from calcium signals; $\nu$ is the number of degrees of freedom.
-For independent observations $\nu=n-1$ because estimation of the mean $\mu$ accounts for one degree of freedom.
-When observations are correlated, as is the case with calcium signals, $\nu < n-1$ and may be estimated from the signal.
-
-The sample covariance matrix is constructed to be unbiased, so that $\E{\hat\Sigma_0} = \Sigma$.
-
 \subsection*{Evaluation of covariance matrix estimators}
 The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.
-The loss function quantifies the deviation of $\hat\Sigma$ from $\Sigma$ and attains its minimum  when $\hat\Sigma = \Sigma$.
-\Kcomment{Could we say that it quantifies the error in the estimate?}
+The loss function quantifies the error of the estimate, \emph{i.e}.~the deviation of $\hat\Sigma$ from $\Sigma$.
+The loss function must attain its minimum  when $\hat\Sigma = \Sigma$.
 
@@
 
-The aim of our project is to produce covariance matrix estimates that minimize the expected loss 
+The aim of our project is to identify which covariance matrix estimator minimizes the expected loss 
 \begin{equation}

### 2013-10-30 sections/Results.tex
updated the brief description of the covariance estimators in Results.

@@
 \begin{equation}
-\hat \Sigma_{\rm diag} = (1-\lambda) \hat \Sigma_0 + \lambda \hat D
+\hat \Sigma_{\sf diag} = (1-\lambda) \hat \Sigma_0 + \lambda \hat D
 \end{equation}
@@
 
-In the second regularized estimator, the target estimate is the factor model $\hat F = LL^\T + \Psi$ with $d$ factors so that $L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\Psi$ contains the indvidual variances of each neuron.
+In the second regularized estimator, the target estimate is the factor model $\hat F = \hat L \hat L^\T + \hat \Psi$ with $d$ factors so that $\hat L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\hat \Psi$ contains the independent variances of each neuron.
 Then the estimate is 
 \begin{equation}
-\hat \Sigma_{\rm factor} = (1-\lambda) \hat \Sigma_0 + \lambda \hat F,
+\hat \Sigma_{\sf factor} = (1-\lambda) \hat \Sigma_0 + \lambda \hat F,
 \end{equation}
 This estiamtor has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $\hat F$ expresses the idea that the fluctuations in population activity must be driven by relatively small number of unobserved sources that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}B).
+The target estimate $\hat F$ expresses the assumption that the fluctuations in population activity are driven by relatively small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}B).
 
-The third estimator is based on sparsifying the inverse covariance (precision) matrix.
-It expresses the idea that, similar to synaptic connections, interactions between neurons are sparse (Fig.~1C).
-The estimator produces a low-dimensional approximation to the covariance that has many zeros in its inverse.
-This approximation problem is also known as covariance selection.
-The covariance matrix estimate is given by
+The third estimator is based on the assumption that only a small fraction of pairs of neurons should have nonzero partial correlations.
+While usual correlations are calculated from the marginal distribution without conditioning on all the other neurons, the partial correlation between a pair of neurons expresses the correlation of their activity conditioned on the activity of all the other cells.
+If cells only exert linear effects on each other and all neurons are visible, partial correlations express the direct interactions between neurons.
+Therefore, this estimator is biased toward the assumption that correlations arise due to interactions between a subset of pairs of recorded neurons (Fig.~\ref{fig:02}C).
+When the partial correlation between a pair of neurons is zero, then the corresponding element of the inverse of the covariance matrix (often referred to as the \emph{precision matrix} or \emph{concentration matrix}) must be zero as well.
+Then the estimator is
 \begin{equation}
-\hat\Sigma = \hat S^{-1}
-\quad\mbox{with}\quad
-\hat S = \argmin\limits_{\|S\|_0 \le \rho} \loss{S^{-1},\hat\Sigma_0}   
+\hat\Sigma_{\sf sparse} = \hat S_\rho^{-1}
 \end{equation}
-where $\mathcal L$ is a loss function (see below) and $\|S\|_0\le\rho$ signifies the constraint that $S$ has at most $\rho$ non-zero coefficients.
-While solving the problem in this form is computationally challenging, relaxing the $L_0$ norm to the $L_1$ norm converts it into one of convex optimization \cite{Donoho:2000}, which is computationally much more tractable.
-The resulting algorithm, which is known as graphical lasso \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator
+where $\hat S\rho$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements.
+The  one hyperparameter $\rho$ expresses the sparsity (fraction of off-diagonal zeros) in $\hat\Sigma_\rho$.
+
+Finally, we consider a fourth estimator, which allows for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}D).
+This estimator has the form
 \begin{equation}
-\hat\Sigma_{\rm sparse}^\mathcal{C} = \hat S^{-1}
-\quad\mbox{with}\quad
-\hat S = \argmin \loss{S^{-1},\hat\Sigma_0} + \lambda \|S\|_1,
+\hat\Sigma_{\sf sparse+latent} = (\hat S_\rho + \hat L\hat L^\T)^{-1},
 \end{equation}
-where $\|S\|_1$ is the $L_1$ norm of the precision matrix (see Methods for details) and $\lambda$ is again determined by cross-validation.
+where, as above, $\hat S_\rho$ is a sparse matrix and $\hat L$ is a $d\times p$ matrix of factor loadings.
+The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of the sparsity of local interactions $\rho$.
 
-Finally, we consider a fourth estimator, which combines common inputs with sparse interactions (Fig.~1D). For this estimator the covariance is
-\begin{equation}
-\hat\Sigma_{\rm combined} = (S + LL^\T)^{-1},
-\end{equation}
-where, as above, $S$ is sparsified using $L_1$ norm and $L$ is a low-rank matrix with the rank determined by cross-validation.
-
-\subsection*{Simulation using toy data}
+\subsection*{Simulation}
 

### 2013-10-30 sections/Figure2.tex
Continuing to integrate Alex's changes.

@@
 Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
-\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies.
-\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
-\textbf{C}: Graphical model with conditional dependencies between some pairs of observed neurons, no latent units. 
-\textbf{D}: Graphical model with conditional dependencies between some pairs of observed neurons and dense interactions with a few latent units.
+It assumed that all interactions are linear, \emph{i.e.}\;the mean firing rate of any neuron can be computed by a linear combination of the firing rates of other neurons and latent units.
+\textbf{A}: ``Independent.'' A population with no interactions corresponds to the diagonal covariance matrix.
+\textbf{B}: ``Latent factors.'' Observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
+\textbf{C}: ``Sparse.'' Partial correlations between a subset of pairs of observed neurons, zero partial correlations between all other pairs.  
+\textbf{D}: ``Sparse+latent.''  Partial correlations between a subset of pairs of observed neurons and a few latent factors interacting with the entire population. 
 }\label{fig:02}

### 2013-10-30 sections/Results.tex
Continuing to integrate Alex's changes.

@@
 
-To verify our approach and to illustrate the performance of our four regularized estimators and the sample covariance estimator, we constructed five model populations with different underlying structures.
+To verify our approach and to illustrate the performance of the four regularized estimators, we constructed five model populations with different underlying structures.
 Each population contained 100~neurons.
-The first four populations matched the low-dimensional structure of our four estimators: independent (Fig.~2\,A), low-rank (Fig.~2\,B), sparse inverse (Fig.~2\,C) and low-rank combined with sparse inverse (Fig.~2\,D).
-\Acomment{Integrate schematics from Fig.~1 into this figure (make it the first row) for better correspondence} In addition, we considered a fifth population that had no low-dimensional structure (Fig.~2\,E).
+The first four populations matched the low-dimensional structure of our four estimators: independent (Fig.~\ref{fig:03}A), low-rank (Fig.~\ref{fig:03}B), sparse inverse (Fig.~\ref{fig:03}C) and low-rank combined with sparse inverse (~\ref{fig:03}D).
+\Acomment{Integrate schematics from Fig.~1 into this figure (make it the first row) for better correspondence} In addition, we considered a fifth population that had no low-dimensional structure (Fig.~\ref{fig:03}\,E).
+
+\input{sections/Figure3.tex}
 
@@
 We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}.
-We presented numerous repetitions of full-field drifting gratings (Fig.
-1A and 1B) to the eye contralateral to the imaged site.
+We presented numerous repetitions of full-field drifting gratings (Fig.~\ref{fig:01}A and \ref{fig:01}B) to the eye contralateral to the imaged site.
 This technique allowed us to record from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
-We deconvolved somatic calcium signals using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.\;1C and 1D) and subtracted the average stimulus response to remove effects of the stimulus.
-From this residual response we computed the sample noise covariance matrix (Fig.\;1E).
+We deconvolved somatic calcium signals using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:01}C and \ref{fig:01}D) and subtracted the average stimulus response to remove effects of the stimulus.
+From this residual response we computed the sample noise covariance matrix (Fig.~\ref{fig:01}E).
+
+\input{sections/Figure1.tex}
 
 In such highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability.
-At the same time, most correlations are relatively small (Fig.~1\,E), suggesting that a simple shrinkage towards independence may provide a sufficiently well regularized estimate.
+At the same time, most correlations are relatively small (Fig.~\ref{fig:01}E), suggesting that a simple shrinkage towards independence may provide a sufficiently well regularized estimate.
 By comparing the performance of our differently regularized estimators we can gain insights into which of these aspects are important in our data.

### 2013-10-30 sections/Figure3.tex
minor
merged the graphical model into the simulation results figure.

@@
 \caption{
-In simulation, selection amongst estimators recognizes the true covariance atructure. 
-{\bf A--D} are examples of true $100\times100$ covariance matrices of multivariate normal distributions used in simulations.
+Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
+It assumed that all interactions are linear, \emph{i.e.}\;the mean firing rate of any neuron can be computed by a linear combination of the firing rates of other neurons and latent units.
+%\textbf{A}: ``Independent.'' A population with no interactions corresponds to the diagonal covariance matrix.
+%\textbf{B}: ``Latent factors.'' Observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
+%\textbf{C}: ``Sparse.'' Partial correlations between a subset of pairs of observed neurons, zero partial correlations between all other pairs.  
+%\textbf{D}: ``Sparse+latent.''  Partial correlations between a subset of pairs of observed neurons and a few latent factors interacting with the entire population. 
+{\bf E--H} are examples of true $100\times100$ covariance matrices of multivariate normal distributions used in simulations.
 Matrix A has no low-dimensional structure.
 Matrices B--D have low-dimensional structture corresponding to the respective graphical models in Fig.~2B--D.
-{\bf E--H} are examples of sample covariance matrices obtained from samples of size $n=1000$ from distributions with respective true covariance matrices A--D.
-{\bf I--L:} Performance evaluation of covariance estimators A, B, C, D from samples obtained from models A, B, C, D.
+{\bf I--L} are examples of sample covariance matrices obtained from samples of size $n=1000$ from distributions with respective true covariance matrices A--D.
+{\bf M--P:} Performance evaluation of covariance estimators A, B, C, D from samples obtained from models A, B, C, D.
 Excess loss $\loss{\hat\Sigma,\Sigma}-\loss{\Sigma,\Sigma}$ is only accessible with the true covariance matrix $\Sigma$ is known and zero excess loss implies absolutely correct estimation.
@@
 Error bars indicate the standard deviation.
-{\bf M--P:} Performance evaluation using cross-validated \emph{empirical loss} from Eq.~\ref{eq:empiricalLoss} without knowledge of true $\Sigma$.
+{\bf Q--T:} Performance evaluation using \emph{validated loss} from Eq.~\ref{eq:validationLoss} without knowledge of true $\Sigma$.
 Cross-validation reproduces the same relationship between performances of the estimators.

### 2013-10-30 sections/Results.tex
merged the graphical model into the simulation results figure.

@@
 We considered four regularized estimators based on distinct families of low-dimensional target estimates: independent, latent factors, sparse partial correlations, and sparse partial correlations with latent factors.
-These target estimates are depicted graphically in Fig.~\ref{fig:02}.
-
-\input{sections/Figure2.tex}
+These target estimates are depicted graphically in Fig.~\ref{fig:03}A-D.
 
@@
 \end{equation}
-The diagonal target estimate expresses the idea of lack of dependence (or of linear association) between the activity of observed neurons (Fig.~\ref{fig:02}A).
+The diagonal target estimate expresses the idea of lack of dependence (or of linear association) between the activity of observed neurons (Fig.~\ref{fig:03}A).
 If this assumption aptly describes recorded data, then strong shrinkage toward $\hat D$ will add little bias while strongly reducing the variability of the estimate.
@@
 This estiamtor has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $\hat F$ expresses the assumption that the fluctuations in population activity are driven by relatively small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}B).
+The target estimate $\hat F$ expresses the assumption that the fluctuations in population activity are driven by relatively small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:03}B).
 
-The third estimator is based on the assumption that only a small fraction of pairs of neurons should have nonzero partial correlations.
-While usual correlations are calculated from the marginal distribution without conditioning on all the other neurons, the partial correlation between a pair of neurons expresses the correlation of their activity conditioned on the activity of all the other cells.
+The third estimator is based on the assumption that all correlations are the result of direct linear interactions between a fraction of observed cell pairs.
+This assumption is enforced by reducing to zero the majority of pairwise \emph{partial correlations} in the recorded population.
+While usual correlations are calculated from the marginal distribution without conditioning on all the other neurons, the partial correlation expresses the pairwise correlation conditioned on the activity of all the other cells.
 If cells only exert linear effects on each other and all neurons are visible, partial correlations express the direct interactions between neurons.
-Therefore, this estimator is biased toward the assumption that correlations arise due to interactions between a subset of pairs of recorded neurons (Fig.~\ref{fig:02}C).
+Therefore, this estimator is biased toward the assumption that correlations arise due to interactions between a subset of pairs of recorded neurons (Fig.~\ref{fig:03}C).
 When the partial correlation between a pair of neurons is zero, then the corresponding element of the inverse of the covariance matrix (often referred to as the \emph{precision matrix} or \emph{concentration matrix}) must be zero as well.
@@
 \begin{equation}
-\hat\Sigma_{\sf sparse} = \hat S_\rho^{-1}
+\hat\Sigma_{\sf sparse} = \hat S^{-1}
 \end{equation}
-where $\hat S\rho$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements.
-The  one hyperparameter $\rho$ expresses the sparsity (fraction of off-diagonal zeros) in $\hat\Sigma_\rho$.
+where $\hat S$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements. The  one hyperparameter $\rho$ expresses the sparsity (fraction of off-diagonal zeros) in $\hat\Sigma$.
 
-Finally, we consider a fourth estimator, which allows for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}D).
+Finally, we consider a fourth estimator, which allows for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:03}D).
 This estimator has the form
 \begin{equation}
-\hat\Sigma_{\sf sparse+latent} = (\hat S_\rho + \hat L\hat L^\T)^{-1},
+\hat\Sigma_{\sf sparse+latent} = (\hat S - \hat L\hat L^\T)^{-1},
 \end{equation}
-where, as above, $\hat S_\rho$ is a sparse matrix and $\hat L$ is a $d\times p$ matrix of factor loadings.
+where, as above, $\hat S$ is a sparse matrix and $\hat L$ is a $d\times p$ matrix of factor loadings.
 The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of the sparsity of local interactions $\rho$.

### 2013-10-31 sections/Figure3.tex

--- 2013-10-01 sections/Figure3.tex
@@
 \caption{
-In simulation, selection amongst estimators recognizes the true covariance atructure. 
-{\bf A--D} are examples of true $100\times100$ covariance matrices of multivariate normal distributions used in simulations.
+Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
+It assumed that all interactions are linear, \emph{i.e.}\;the mean firing rate of any neuron can be computed by a linear combination of the firing rates of other neurons and latent units.
+\textsf{A}: ``Independent.'' A population with no interactions corresponds to the diagonal covariance matrix.
+\textsf{B}: ``Latent factors.'' Observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
+\textsf{C}: ``Sparse.'' Partial correlations between a subset of pairs of observed neurons, zero partial correlations between all other pairs.  
+\textsf{D}: ``Sparse+latent.''  Partial correlations between a subset of pairs of observed neurons and a few latent factors interacting with the entire population. 
+{\sf E--H} are examples of true $100\times100$ covariance matrices of multivariate normal distributions used in simulations.
 Matrix A has no low-dimensional structure.
 Matrices B--D have low-dimensional structture corresponding to the respective graphical models in Fig.~2B--D.
-{\bf E--H} are examples of sample covariance matrices obtained from samples of size $n=1000$ from distributions with respective true covariance matrices A--D.
-{\bf I--L:} Performance evaluation of covariance estimators A, B, C, D from samples obtained from models A, B, C, D.
+{\sf I--L} are examples of sample covariance matrices obtained from samples of size $n=1000$ from distributions with respective true covariance matrices A--D.
+{\sf M--P:} Performance evaluation of covariance estimators A, B, C, D from samples obtained from models A, B, C, D.
 Excess loss $\loss{\hat\Sigma,\Sigma}-\loss{\Sigma,\Sigma}$ is only accessible with the true covariance matrix $\Sigma$ is known and zero excess loss implies absolutely correct estimation.
@@
 Error bars indicate the standard deviation.
-{\bf M--P:} Performance evaluation using cross-validated \emph{empirical loss} from Eq.~\ref{eq:empiricalLoss} without knowledge of true $\Sigma$.
+{\sf Q--T:} Performance evaluation using \emph{validated loss} from Eq.~\ref{eq:validationLoss} without knowledge of true $\Sigma$.
 Cross-validation reproduces the same relationship between performances of the estimators.

### 2013-10-31 sections/Figure4.tex

@@
 \includegraphics[width=0.5\textwidth]{figures/Figure4.pdf}
-\caption{{\bf Estimator $\mathcal D$ (sparse+full-rank inverse covariance) dominates in dense recordings from primary visual cortex.}
+\caption{{\bf Estimator $\mathcal D$ (sparse+low-rank inverse covariance) dominates in dense recordings from primary visual cortex.}
 {\bf A:}  Histogram of improvements in validation loss by covariance estimator $\mathcal D$ relative to the samle covariance matrix from the 31 imaged sites used in the study.

### 2013-10-31 sections/Results.tex
minor

@@
 This estiamtor has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $\hat F$ expresses the assumption that the fluctuations in population activity are driven by relatively small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:03}B).
+The target estimate $\hat F$ expresses the assumption that correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:03}B).
 
@@
 This estimator has the form
+$\rho$.
 \begin{equation}
@@
 where, as above, $\hat S$ is a sparse matrix and $\hat L$ is a $d\times p$ matrix of factor loadings.
-The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of the sparsity of local interactions $\rho$.
-
+The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of local interactions
 \subsection*{Simulation}
 
-To verify our approach and to illustrate the performance of the four regularized estimators, we constructed five model populations with different underlying structures.
-Each population contained 100~neurons.
-The first four populations matched the low-dimensional structure of our four estimators: independent (Fig.~\ref{fig:03}A), low-rank (Fig.~\ref{fig:03}B), sparse inverse (Fig.~\ref{fig:03}C) and low-rank combined with sparse inverse (~\ref{fig:03}D).
-\Acomment{Integrate schematics from Fig.~1 into this figure (make it the first row) for better correspondence} In addition, we considered a fifth population that had no low-dimensional structure (Fig.~\ref{fig:03}\,E).
+To verify our approach and to illustrate the performance of the four regularized estimators, we constructed five model populations with different covariance structures.
+Each population contained 50~neurons.
+The first four populations matched the low-dimensional structure of our four estimators: independent (Fig.~\ref{fig:03}A), latent factors (Fig.~\ref{fig:03}B), sparse partial correlations (Fig.~\ref{fig:03}C) and low-rank combined with sparse inverse (Fig.~\ref{fig:03}D).
+In addition, we considered a fifth population that had no low-dimensional structure (Fig.~\ref{fig:03}\,E).
 
@@
 However, when comparing multiple estimators against each other, the one whose target estimate most closely matches the true value with the smallest number of parameters will reduce the estimation error with the least increase in approximation error.
-Indeed, in all four toy examples with a low-dimensional structure, the estimator with the matching regularization target performed best (Fig.~2\,A--D, fourth row).
-Since the estimator combining low-rank and sparse interactions combines to low-dimensional targets and includes the simpler estimators as special cases, it performed almost as well even when the low-dimensional structure did not include either a low-rank component or sparse interactions (Fig.~2\,B,\,C).
-\Acomment{May want to include a smaller sample size as well since here the more reduced estimators (B, C) may actually outperform estimator D when the population structure matches their target and there is little data (say, 1.5x number of neurons)}
+Indeed, in all four toy examples with a low-dimensional structure, the estimator with the matching regularization target performed best (Fig.~\ref{fig:03}\,A--D, fourth row).
+Since the estimator combining low-rank and sparse interactions combines to low-dimensional targets and includes the simpler estimators as special cases, it performed almost as well even when the low-dimensional structure did not include either a low-rank component or sparse interactions (Fig.~\ref{fig:03}\,B,\,C).
 
-Since the above evaluations of excess loss require knowledge of ground truth, this analysis can be done only on toy data.
-However, an unbiased empirical estimate of the excess loss exists (see Methods, Eq.~\ref{eq:validationLoss} \Acomment{This may need a brief sentence about the intuition what's the difference, or a reference to the relevant methods section}).
-To verify that our approach should also work in a realistic situation without access to ground truth, we also computed the empirical loss.
-Indeed, without access to ground truth, this analysis revealed a pattern of results that reproduced the results obtained with access to ground truth above (Fig.~2, last row).
-Because validation loss is computed by comparing estimates to noisy sample covariance matrices from smaller validation sets, empirical loss is a much noisier measurement than excess loss.
-In addition, it does not converge to zero with increasing sample size as excess loss does.
+Since the above evaluations of excess loss require knowledge of ground truth, this analysis can be done only on simulated data.
+However, \emph{validation loss} can be used as an unbiased empirical estimate of loss  (see Methods, Eq.~\ref{eq:validationLoss}.
+Validation loss uses an independent validation sample to estimate the value of loss without access to ground truth.
+Indeed, this analysis revealed a pattern of results that reproduced the results obtained with access to ground truth above (Fig.~\ref{fig:03}, last row).
+Because validation loss is computed by comparing estimates to noisy sample covariance matrices from smaller validation sets, validation loss has higher variance than excess loss.
+Under our chosen loss function, validation loss does not converge to zero.
 
-\subsection*{Covariance estimation on neural data}
+\subsection*{Covariance estimation in neural data}
 
@@
 This technique allowed us to record from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
-We deconvolved somatic calcium signals using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:01}C and \ref{fig:01}D) and subtracted the average stimulus response to remove effects of the stimulus.
+We deconvolved somatic calcium signals using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:01}C and \ref{fig:01}D) and subtracted the average stimulus response to remove averaged linear effects of the stimulus.
 From this residual response we computed the sample noise covariance matrix (Fig.~\ref{fig:01}E).
@@
 
-As expected, all regularized estimators outperformed the sample covariance estimator substantially (Fig.~4) \Acomment{I think we should show that}.
+As expected, all regularized estimators outperformed the sample covariance estimator substantially (Fig.~\ref{fig:04}) \Acomment{I think we should show that}.
 Say something about how shrinkage, low-rank and sparse inverse relate \Acomment{Why do we do pairwise comparisons instead of just showing median (across sites) log loss relative to the combined estimator?
@@
 
+\input{sections/Figure4.tex}
+
 \subsection*{Relationship between functional covariance structure and circuit architecture}
+\input{sections/Figure5.tex}
+\input{sections/Figure6.tex}
 

### 2013-11-01 sections/Results.tex
changes in response to Kreso's comments this morning: wording in the Results section.

@@
 \end{equation}
-where $\E{\cdot}$ denotes expectation  and $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates discretized into bins of duration $\Delta t$ and $\mu = \E{x}$.
+where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$. The vector of mean firing rates is $\mu = \E{x}$.
 
@@
 Regularization is the deliberate biasing of the estimate toward a low-dimensional, less variable \emph{target estimate} to strike a favorable balance between bias and variability of the estimate \cite{Bickel:2006,Ledoit:2004}.
-The estimator whose target estimate most closely matches the true low-dimensional structure of the data is likely to outperform other estimators.
+
+The estimator whose target estimate most closely matches the true low-dimensional structure of the data is likely to outperform other estimators. This principle provides the logic of this study.
 
@@
 
-In the first regularized estimator, the target estimate is a diagonal matrix $\hat D$.
+In the first regularized estimator, the target estimate is the diagonal matrix $\hat D$ containing on its diagonal estimates of the variances for each neuron.
 The optimized estimate $\hat\Sigma_{\rm diag}$ is obtained by the linear \emph{shrinkage} of the unbiased estimate $\hat\Sigma_0$ toward $\hat D$ as regulated by scalar \emph{shrinkage intensity} $\lambda \in [0, 1]$:
@@
 \end{equation}
-where $\hat S$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements. The  one hyperparameter $\rho$ expresses the sparsity (fraction of off-diagonal zeros) in $\hat\Sigma$.
+where $\hat S$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements.
+The estimate has one hyperparameter to regulate the sparsity (fraction of off-diagonal zeros) in $\hat\Sigma$.
 

### 2013-11-10 sections/Figure1.tex
added Figure 6.
publication-quality figures: Simulation, Acquisition, and Domination.

--- 2013-09-25 sections/Figure1.tex
@@
 \begin{center}
-\includegraphics[width=2.5in]{figures/Figure1.pdf}
+\includegraphics{figures/Acquisition.pdf}
 \end{center}
 \caption{
-{\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\bf A.} Visual stimuli comprising brief (500 ms) presentatios of full-field drifting gratings separated by blank screens.
-{\bf B.} Two-photon fast 3D imaging of calcium signals in an awake mouse.
-{\bf C.} Deconvolved calcium signals.
-{\bf D.} The sample covariance matrix of neuronal calcium signals in 200 ms time bins.
+{\bf Acquistion of neural population activity using two-photon fluorescence imaging of calcium signal.}  {\sf A.} Visual stimuli comprising brief (500 ms) presentatios of full-field drifting gratings separated by blank screens.
+{\sf B.} Two-photon fast 3D imaging of calcium signals.
+{\sf D.} The first and last 30 seconds of a 30-minute recording of population calcium signals.
+The signals are deconvovled and binned at 150 ms.
+{\sf E.} The sample noise correlation matrix of neuronal calcium signals from {\sf D} after subtracting the average stimulus response.
+{\sf F.} The histogram of off-diagonal coefficients from the correlation matrix in {\bf D}.
 }
-\label{Figure_label}
+\label{fig:01}
 \end{figure}

### 2013-11-19 cosyne/Yatsenko-Cosyne-2014.tex
Added Cosyne abtract

\documentclass[12pt]{amsart}
\usepackage[left=.5in,top=.5in,right=.5in,bottom=.5in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{float}
\usepackage{multicol}
\usepackage{cite}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\setlength{\parskip}{3pt plus2pt}
\setlength{\parindent}{0.0cm}
\usepackage{times}
\usepackage{wrapfig}
\usepackage{hyperref}
\newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}
\usepackage[small]{caption}
\pagenumbering{gobble}
\usepackage[small,compact]{titlesec}

\begin{document}

\begin{flushleft}
{\Large
Low-dimensional structure of neural correlations in cortical microcircuits
}
\\
\vspace{6pt}
Dimitri Yatsenko, 
Kre\v{s}imir Josi\'{c},
Alexander S.~Ecker,
Emmanouil Froudarakis,
R.~James Cotton,
Andreas S.~Tolias
\end{flushleft}
\subsection*{Summary}
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons in behaving animals. It is commonly anticipated that correlations measured in such recordings will reveal the functional organization of neural circuits. However, estimation and interpretation of  large correlation matrices from finite recordings can be challenging.  Estimation can be improved by regularization: the biasing of the estimate toward a low-dimensional, less variable approximation.  The amount of improvement depends on how closely the reduced approximation captures the dominant interactions with fewest terms. Therefore, the selection of the most efficient estimator is an empirical question dependent on the system under investigation.

We compared regularized correlation matrix estimators biased toward four respective families of low-dimensional correlation structures: independent, latent factors, sparse partial correlations, and sparse partial correlations + latent factors.  The estimators' performance was evaluated on the noise correlation matrices from spatially compact groups of 91--314 neurons (average 206) in mouse visual cortex (31 sites in 24 animals) acquired by \emph{in vivo} fast 3D random-access two-photon imaging of calcium signals. Recordings lasting between 15 and 20 minutes were deconvolved and binned at 150 ms. Each estimator was optimized and evaluated by nested cross-validation.  We found that both sparse partial correlations and latent factors were required for efficient estimation.  In addition to reducing estimation error, optimized estimates provided compact representations of the correlation structure with a sparse network of interactions and several global latent inputs. The interaction connectivity depended strongly on the sign of the interaction, physical distances, differences in cortical depth, and differences in preferred orientation between pairs of cells. Such networks differed from networks constructed by considering only most significant correlations. In ongoing studies, we seek to relate such correlation structures to anatomical organization of circuits across various stimulus conditions and brain states.

\subsection*{Additional Details}
\subsubsection*{Central Questions} We pursued two related aims: (a) improved estimation of correlation matrices and (b)~discovery of low-dimensional structure of correlations in recordings of multineuronal activity.
\subsubsection*{Approach}
\begin{wrapfigure}[18]{r}{0.4\textwidth}
\centering
\vspace{-6pt}
\includegraphics{./Simulation.pdf}
\caption{Graphical structure of low-dimensional correlation matrix estimates used as target estimates in regularization. Recorded neurons are in green whereas inferred latent factors are light-colored. Edges denote linear associations (non-zero partial correlations).}
\label{fig:graphical}
\end{wrapfigure}
Both aims benefit from \emph{regularization}: the deliberate biasing (`shrinkage') of the usual sample correlation matrix toward a low-dimensional \emph{target estimate}.

We considered regularized estimates based on four respective low-dimensional structures of correlation matrices (\autoref{fig:graphical}): 
The target of the first estimator has a diagonal covariance matrix, which describes a population of independent neurons \cite{Ledoit:2004}. 
The target of the second estimator is a multifactor model \cite{Fan:2008}, which describes a population of neurons driven together by a small number of latent factors with negligible direct interactions.
The third estimator is based on the assumption that all correlations are the result of direct linear interactions between a fraction of observed cell pairs. This assumption is enforced by reducing to zero the majority of coefficients of the inverse of the correlation matrix \cite{Dempster:1972}. 
Finally, the fourth estimator allows for both common latent units and sparse local interactions between recorded neurons.  
In this estimator, the correlation matrix is modeled as the inverse of a sum of a sparse and a low-rank component \cite{Ma:2013}.

\subsubsection*{Results}
Each estimator was optimized and evaluated with respect to cross-validated multivariate Gaussian log-likelihood.  First, with synthetic data, we demonstrated that estimators whose regularization targets matched the true correlation structure outperformed other estimators (data to be presented).

We then estimated the correlation matrices of population calcium signals of large groups of cells (91--314 neurons, averarge 206 in 31 sites from 24 animals) in small volumes of visual cortex during visual stimulation. We found that the sparse+latent estimator consistently outperformed all other estimators (data to be presented).

In addition to minimizing the estimation error, the sparse+latent estimator revealed sparse networks of linear interactions such as shown in \autoref{fig:reconstruction} and several shared latent units (not shown).
Individual imaged sites varied greatly in the number of significant latent units and the sparsity of direct interations (data to be presented).
We analyzed how the distribution of these interactions compared to the distribution of the usual sample correlations as a function of the difference in preferred orientation, vertical distance, and lateral distance in the cortex \autoref{fig:summary}. The patterns of inferred interactions differed substantially from patterns of most significant correlations (data to be presented).
\begin{multicols}{2}
\begin{Figure}
  \includegraphics{./Reconstruction.pdf}
  \captionof{figure}{An example of 3D arrangement of recorded cells with color-coded orientation preference and edges indicating the network of positive and negative interactions (magenta) as revealed by the sparse+latent estimator. The latent units are not shown.}
  \label{fig:reconstruction}
\end{Figure}

\begin{Figure}
  \centering
  \includegraphics{./Summary.pdf}
  \captionof{figure}{Top row: average sample correlations across all sites. Bottom row: normalized interaction probabilities of the sparse+latent estimator with positive interactions shown in green and negative interactions show in red. Left column: as a function of difference in preferred orientation. Middle column: as a function of difference in cortical depth at small lateral distances. Right column: as a function of lateral distance at the same cortical depth.}
  \label{fig:summary}
\end{Figure}
\end{multicols}

\subsubsection*{Significance}
By an empirical study, we identied the most statistically efficient of several estimates of the correlation matrix of the activity of densly recorded cortical circuit. The resulting estimates suggest network of sparse interactions between pairs of cells and global common inputs. The physiological interpretation of these interactions remains an open question. In future studies, we aim to relate the structure of interactions to specific cell types and brain states.

\bibliographystyle{plos2009}
\bibliography{references.bib}

\end{document}

### 2013-11-19 cosyne/Yatsenko-Cosyne-2014.tex
final version of the CoSyNe abstract.

@@
 \DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
-\setlength{\parskip}{3pt plus2pt}
-\setlength{\parindent}{0.0cm}
+\setlength{\parskip}{0pt}
+\setlength{\parindent}{0.5cm}
 \usepackage{times}
@@
 Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons in behaving animals.
-It is commonly anticipated that correlations measured in such recordings will reveal the functional organization of neural circuits.
+It is commonly anticipated that correlations measured in such recordings will uncover aspects of functional organization of neural circuits.
 However, estimation and interpretation of large correlation matrices from finite recordings can be challenging.
@@
 We found that both sparse partial correlations and latent factors were required for efficient estimation.
-In addition to reducing estimation error, optimized estimates provided compact representations of the correlation structure with a sparse network of interactions and several global latent inputs.
-The interaction connectivity depended strongly on the sign of the interaction, physical distances, differences in cortical depth, and differences in preferred orientation between pairs of cells.
-Such networks differed from networks constructed by considering only most significant correlations.
-In ongoing studies, we seek to relate such correlation structures to anatomical organization of circuits across various stimulus conditions and brain states.
+In addition to reducing estimation error, optimized estimates provided compact representations of the correlation structure decomposed into a sparse network of pairwise interactions and global diffuse fluctuations.
+The density of positive interactions in these networks decreased rapidly with distance and with difference in orientation preference, whereas negative interactions were less selective.
+In ongoing studies, we seek to relate these elements of the correlation structure to circuits' anatomical organization, stimulus conditions, and brain states.
 
@@
 \subsubsection*{Central Questions} We pursued two related aims: (a) improved estimation of correlation matrices and (b)~discovery of low-dimensional structure of correlations in recordings of multineuronal activity.
+
 \subsubsection*{Approach}
-\begin{wrapfigure}[18]{r}{0.4\textwidth}
+\begin{wrapfigure}[16]{r}{0.4\textwidth}
 \centering
@@
 \includegraphics{./Simulation.pdf}
-\caption{Graphical structure of low-dimensional correlation matrix estimates used as target estimates in regularization.
+\caption{Families of low-dimensional correlation target estimates for regularization.
 Recorded neurons are in green whereas inferred latent factors are light-colored.
 Edges denote linear associations (non-zero partial correlations).}
+
 \label{fig:graphical}
@@
 
-We considered regularized estimates based on four respective low-dimensional structures of correlation matrices (\autoref{fig:graphical}): 
+We constructed for covariance estimators based on four low-dimensional target estimates of covariance matrices  (\autoref{fig:graphical}): 
 The target of the first estimator has a diagonal covariance matrix, which describes a population of independent neurons \cite{Ledoit:2004}. 
@@
 Finally, the fourth estimator allows for both common latent units and sparse local interactions between recorded neurons.  
-In this estimator, the correlation matrix is modeled as the inverse of a sum of a sparse and a low-rank component \cite{Ma:2013}.
+In this estimator, the correlation matrix is modeled as the inverse of a sum of a sparse and a low-rank component \cite{Chandrasekaran:2010}.
 
@@
 
-We then estimated the correlation matrices of population calcium signals of large groups of cells (91--314 neurons, averarge 206 in 31 sites from 24 animals) in small volumes of visual cortex during visual stimulation.
+We then estimated the correlation matrices of population calcium signals of large groups of cells (91--314 neurons, averarge 206 in 31 sites from 24 animals) in small volumes of visual cortex during visual stimulation \cite{Cotton:2013}.
 We found that the sparse+latent estimator consistently outperformed all other estimators (data to be presented).
@@
   \includegraphics{./Reconstruction.pdf}
-  \captionof{figure}{An example of 3D arrangement of recorded cells with color-coded orientation preference and edges indicating the network of positive and negative interactions (magenta) as revealed by the sparse+latent estimator.
+  \captionof{figure}{An example of 3D arrangement of recorded cells with color-coded orientation preference and edges indicating the network of positive (green) and negative interactions (magenta) as revealed by the sparse+latent estimator.
 The latent units are not shown.}
@@
 Middle column: as a function of difference in cortical depth at small lateral distances.
-Right column: as a function of lateral distance at the same cortical depth.}
+Right column: as a function of lateral distance at the same cortical depth.
+Dottted lines indicate averages across all conditions across all sites included in each analysis.}
   \label{fig:summary}
@@
 \subsubsection*{Significance}
-By an empirical study, we identied the most statistically efficient of several estimates of the correlation matrix of the activity of densly recorded cortical circuit.
-The resulting estimates suggest network of sparse interactions between pairs of cells and global common inputs.
-The physiological interpretation of these interactions remains an open question.
-In future studies, we aim to relate the structure of interactions to specific cell types and brain states.
+In this empirical study, we identified the most efficient of several estimators of correlation matrices of activity fluctuations in cortical microcircuits.
+The resulting estimates effectively decomposed the correlation structure into direct interactions between pairs of neurons and common diffuse fluctuations.
+The physiological interpretation of these components remains in question and must be addressed by additional experiements where synaptic connectivity, cell types, and local field potentials may be used to corroborate such interpretations.
 

### 2013-11-19 cosyne/Yatsenko-Cosyne-2014.tex
final version of the CoSyNe abstract

@@
 
-We then estimated the correlation matrices of population calcium signals of large groups of cells (91--314 neurons, averarge 206 in 31 sites from 24 animals) in small volumes of visual cortex during visual stimulation \cite{Cotton:2013}.
+We then estimated the correlation matrices of population calcium signals of large groups of cells (91--314 neurons, average 206 in 31 sites from 24 animals) in small volumes of visual cortex during visual stimulation \cite{Cotton:2013}.
 We found that the sparse+latent estimator consistently outperformed all other estimators (data to be presented).
@@
 In addition to minimizing the estimation error, the sparse+latent estimator revealed sparse networks of linear interactions such as shown in \autoref{fig:reconstruction} and several shared latent units (not shown).
-Individual imaged sites varied greatly in the number of significant latent units and the sparsity of direct interations (data to be presented).
+Individual imaged sites varied greatly in the number of significant latent units and the sparsity of direct interactions (data to be presented).
 We analyzed how the distribution of these interactions compared to the distribution of the usual sample correlations as a function of the difference in preferred orientation, vertical distance, and lateral distance in the cortex \autoref{fig:summary}.
@@
 The resulting estimates effectively decomposed the correlation structure into direct interactions between pairs of neurons and common diffuse fluctuations.
-The physiological interpretation of these components remains in question and must be addressed by additional experiements where synaptic connectivity, cell types, and local field potentials may be used to corroborate such interpretations.
+The physiological interpretation of these components remains in question and must be addressed by additional experiments where synaptic connectivity, cell types, and local field potentials may be used to corroborate such interpretations.
 

### 2013-11-29 plos.tex
refined Alex's revision of the results sections -- to be continued.
updated the brief description of the covariance estimators in Results.
Renamed figures to Figure0#.pdf and completed the corresponding captions.

@@
 % Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
 
@@
 \newcommand{\loss}[1]{\mathcal L\left(#1\right)} 
-\newcommand{\T}{{\rm T}}
+\newcommand{\T}{{\sf T}}
 \newcommand{\E}[2][]{\mathbb E_{#1}\left[ #2\right]}    % expected value
@@
 \begin{document}
-
 % Title must be 150 characters or less
@@
 {\Large
-Improved estimates of neural correlations suggest detailed interactions in visual cortex
+Improved estimation of neural correlations suggests detailed interactions in visual cortex
 }
@@
 
+\section*{Abstract}
 % Please keep the abstract between 250 and 300 words
-\section*{Abstract}
-\input{sections/Abstract.tex}
+Correlations between the spiking activity of pairs of neurons are among the most familiar descriptive statistics of neural activity.
+Multineuronal recordings provide richer information than the equivalent number of neuron pairs.
+For example, the full covariance  matrix reveals correlated activity across the entire population as well as partial correlations between pairs.
+Estimation of covariance matrices can be improved by regularization, \emph{i.e.}\;by imposing some kind of structure.
+Optimal regularization must be determined empirically since its effect depends on how closely the imposed structure matches the underlying regularities.
+For example, we can use low-rank parameterizations of the covariance matrix to account for common fluctuations across the recorded population.
+Conversely, if correlations are strongly influenced by a small fraction of pairwise linear associations between the observed neurons, we can impose sparsity on the inverse covariance matrix.
+
+We can also use a \lq sparse+low rank\rq inverse covariance representation to account for both common fluctuations and pairwise interactions.  
+To select the optimal structure of neural correlations in a local neural circuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance.
+We inferred instantaneous firing rates in 200 ms bins from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.
+Each covariance estimator was optimized and evaluated by cross-validation.
+As expected, covariance shrinkage reliably outperformed the sample covariance estimate.
+In turn, factor analysis-based estimates significantly outperformed covariance shrinkage.
+Yet sparse inverse covariance with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage estimators.
+The superior performance of the sparse inverse covariance estimator suggests the relative importance of detailed network interactions over common diffuse input in the circuit we studied.
 
@@
 \section*{Author Summary}
-\input{sections/Summary.tex}
+Covariance matrices of the activity of populations of neurons are useful descriptions of the functional organization of neural circuits with implications for stimulus coding and neural circuit architecture.
+Estimates of the covariance matrix can be dramatically improved by \emph{regularization}, \emph{i.e.}\;the biasing of the usual estimate toward one of several possible low-dimensional approximations, the \emph{target estimate}.
+However, the amount of improvement depends on how closely the selected approximation reflects the dominant interactions in the circuit.
+In simulation, we demonstrated that no regularized estimate was universally superior to others: each performed best when the target estimate was the closest match to the structure of the data generating process.
+We then evaluated covariance matrix estimates on the calcium activity recorded \emph{in vivo} from large populations of closely spaced neurons  in mouse primary visual cortex, after subtracting the average stimulus response.
+Using cross-validation, we found that the best estimate was produced when the covariance matrix was approximated as that of a sparse Gaussian graphical models with a small number of latent units. 
+Thanks to the paucity of their parameters, these optimally estimated covariance matrices also help interpretation of the covariance structure and allow relating the functional organization of the neural circuit to its cytoarchitecture.
 
 \section*{Introduction}
-\input{sections/Introduction.tex}
+Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Renart:2010}.
+For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been of particular interest.  
+Noise correlations can have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}, and have been interpreted to indicate detailed and specific functional organization.
+Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity and stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, cell-type specificity (?), progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
+
+However, neural correlations do not come with ready or unambiguous mechanistic interpretations.
+Theoretical and simulation studies have shown that neural correlations, rise at various temporal scales from combinations of multiple underlying mechanisms.
+These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \cite{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
+
+Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \cite{Shadlen:1998,Zohary:1994}.
+Multineuronal recordings allow estimation of covariance matrices of large populations of neurons.
+Such estimates provide more information than the equivalent number of pairwise correlations assessed in isolation.
+Indeed, the correlation matrix is greater than the sum of its parts: it can be transformed into other representations that accentuate different aspects of the correlation structure and may suggest different mechanistic interpretations.
+For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients.
+In contrast, the off-diagonal elements of the inverse of the correlation matrix constitute scaled partial correlations between neuron pairs, which reflect their specific linear dependencies, after accounting for the activity of all the other recorded cells; a strong interaction between a pair of neurons may be expressed by a single partial correlation but its effects may propagate to multiple correlations and eigenvalues.
+The inverse of the covariance matrix plays an important role in decoding schemes such as linear discriminant analysis, for example.
+The mutliple  representations of the covariance matrix with their alternative interpretations add both complications and opportunities into the search for fundamental regularities in neural population activity.
+
+With large numbers of recorded cells, the usual estimations of the covariance matrix from finite recordings become ill-conditioned.
+Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically.
+For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \cite{Ledoit:2004}.
+Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.
+
+With this study, we pursue two related goals: 
+\begin{enumerate}
+	\item Devise more efficient estimators of neural covariance matrices in recordings from large populations of neurons.
+	\item Aid interpretation of neural covariance matrices.
+\end{enumerate}
+
+To select the best covariance matrix estimate for a specific neural circuit, we evaluate the performance of four estimators using different regularization schemes, each motivated by  a different hypothesis for the origin of neural correlations.
+\emph{Regularization} is the deliberate biasing the solution toward a simplified, low-dimensional (`sparse') \emph{target estimate} \cite{Schafer:2005,Bickel:2006}.
+Regularized estimators allow reaching a favorable tradeoff between bias and variance in order to minimize the total error.
 
@@
 
-\input{sections/Results.tex}
+\subsection*{Covariance estimation}
+We aim to estimate the true covariance matrix 
+\begin{equation}\label{eq:true-covariance}
+\Sigma = \E{(x-\mu)(x-\mu)^\T}
+\end{equation}
+where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$. The vector of mean firing rates is $\mu = \E{x}$.
 
-%\input{sections/Acquisition.tex}
-%\input{sections/Theory.tex}
-%\input{sections/Regularization.tex}
-%\input{sections/Estimators.tex}
-%\input{sections/Simulation.tex}
-%\input{sections/Findings.tex}
-%\input{sections/Structure.tex}
+The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_0$ computed from the empirical sample of observations $x(1),\ldots,x(n)$ as
+\begin{equation}
+C_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T, 
+\end{equation}
+where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent).  
+The sample covariance matrix is unbiased, \emph{i.e.} $\E{C_0}-\Sigma=0$.
+For finite sample sizes, however, $C_0$ is not as close to $\Sigma$ as a number of biased estimators.
+
+Regularization is the deliberate biasing of the estimate toward a low-dimensional, less variable \emph{target estimate} to strike a favorable balance between bias and variability of the estimate \cite{Bickel:2006,Ledoit:2004}.
+
+The estimator whose target estimate most closely matches the true low-dimensional structure of the data is likely to outperform other estimators. This principle provides the logic of this study.
+
+We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (Fig.~\ref{fig:02}\;row 1).
+
+In the first regularized estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ containing on its diagonal estimates of the variances.
+The regularized estimate is obtained by linear \emph{shrinkage} of the unbiased estimate $C_0$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
+\begin{equation}
+C_{\sf diag} = (1-\lambda) C_0 + \lambda D
+\end{equation}
+The diagonal target estimate expresses the idea of a lack of dependence (or of linear association) between the activity of observed neurons (Fig.~\ref{fig:02}\;A).  
+If this assumption aptly describes recorded data, then strong shrinkage toward $D$ will add little bias while strongly reducing the variability of the estimate.
+Shrinkage allows for partial commitment to the low-dimensional representation.
+
+In the second regularized estimator $C_{\sf factor}$, the target estimate is the factor model $F =  L L^\T + \Psi$ with $d$ factors so that $L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\Psi$ contains the independent variances of each neuron.
+Then the estimate is 
+\begin{equation}
+C_{\sf factor} = (1-\lambda) C_0 + \lambda F
+\end{equation}
+This estimator has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
+The target estimate $F$ expresses the assumption that correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}\;B).
+
+The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations are the result of direct linear interactions between pairs of observed cells and that such interactions occur only between a fraction of such pairs (Fig.~\ref{fig:02}\;C).
+This assumption is enforced by reducing to zero the majority of pairwise \emph{partial correlations} in the recorded population. 
+While usual correlations are calculated from the marginal distribution without conditioning on all the other neurons, the partial correlation expresses the pairwise correlation conditioned on the activity of all the other cells.
+If cells only exert linear effects on each other and all neurons are recorded, partial correlations express the direct interactions between neurons. 
+Therefore, this estimator is biased toward the assumption that correlations arise due to interactions between a subset of pairs of recorded neurons (Fig.~\ref{fig:03}C).
+When the partial correlation between a pair of neurons is zero, then the corresponding element of the inverse of the covariance matrix (often referred to as the \emph{precision matrix} or \emph{concentration matrix}) must be zero as well.
+Then the estimator is
+\begin{equation}
+C_{\sf sparse} = S^{-1}
+\end{equation}
+where $\hat S$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements. The estimate has one hyperparameter to regulate the sparsity (fraction of off-diagonal zeros) in $S$.
+
+Finally, we consider the fourth estimator $C_{\sf sparse+lowrank}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\;D).
+This estimator has the form
+\begin{equation}
+C_{\sf sparse+latent} = (S - LL^\T)^{-1},
+\end{equation}
+where, as above, $S$ is a sparse matrix and $L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.
+
+\subsection*{Simulation}
+To verify our approach and to illustrate the performance of the four regularized estimators, we constructed five model populations with different covariance structures.
+Each population contained 50~neurons.
+The first four populations matched the low-dimensional structure of our four estimators: independent (Fig.~\ref{fig:03}A), latent factors (Fig.~\ref{fig:03}B), sparse partial correlations (Fig.~\ref{fig:03}C) and low-rank combined with sparse inverse (Fig.~\ref{fig:03}D).
+In addition, we considered a fifth population that had no low-dimensional structure (Fig.~\ref{fig:03}\,E).
+
+To evaluate the performance of the different estimators, we computed the excess loss for all combinations of model populations and estimators, including the sample covariance.
+The first striking observation is that in all cases all four regularized estimators performed substantially better than the sample covariance (Fig.~2, fourth row).
+In particular, this is even true for the case where the population did not have any low-dimensional structure at all (Fig.~2\,E).
+While it may appear surprising at first sight, this counter-intuitive phenomenon is known as \emph{Stein's phenomenon} or \emph{Stein's paradox} \cite{Efron:1977}, named after its discoverer Charles Stein \cite{Stein:1956}.
+A common misconception about regularization is that its effect depends on accurate prior knowledge about the structure of the data.
+However, substantial improvement can be attained by shrinking the unbiased estimate toward an arbitrary target as long as the target is less variable than the unbiased estimator.
+The more accurate description of regularization is as of the optimal tradeoff between estimation and approximation error -- the so-called ``bias-variance tradeoff''.
+
+Thus, if taken in isolation a regularized estimator improves the estimate we should not interpret this result to suggest that the estimator's target has the same low-dimensional structure as the data-generating process.
+However, when comparing multiple estimators against each other, the one whose target estimate most closely matches the true value with the smallest number of parameters will reduce the estimation error with the least increase in approximation error.
+Indeed, in all four toy examples with a low-dimensional structure, the estimator with the matching regularization target performed best (Fig.~\ref{fig:03}\,A--D, fourth row).
+Since the estimator combining low-rank and sparse interactions combines to low-dimensional targets and includes the simpler estimators as special cases, it performed almost as well even when the low-dimensional structure did not include either a low-rank component or sparse interactions (Fig.~\ref{fig:03}\,B,\,C).
+
+Since the above evaluations of excess loss require knowledge of ground truth, this analysis can be done only on simulated data.
+However, \emph{validation loss} can be used as an unbiased empirical estimate of loss  (see Methods, Eq.~\ref{eq:validation-loss}.
+Validation loss uses an independent validation sample to estimate the value of loss without access to ground truth.
+Indeed, this analysis revealed a pattern of results that reproduced the results obtained with access to ground truth above (Fig.~\ref{fig:03}, last row).
+Because validation loss is computed by comparing estimates to noisy sample covariance matrices from smaller validation sets, validation loss has higher variance than excess loss.
+Under our chosen loss function, validation loss does not converge to zero.
+
+\subsection*{Covariance estimation in neural data}
+
+We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}.
+We presented numerous repetitions of full-field drifting gratings (Fig.~\ref{fig:01}A and \ref{fig:01}B) to the eye contralateral to the imaged site.
+This technique allowed us to record from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
+We deconvolved somatic calcium signals using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:01}C and \ref{fig:01}D) and subtracted the average stimulus response to remove averaged linear effects of the stimulus.
+From this residual response we computed the sample noise covariance matrix (Fig.~\ref{fig:01}E).
+
+In such highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability.
+At the same time, most correlations are relatively small (Fig.~\ref{fig:01}E), suggesting that a simple shrinkage towards independence may provide a sufficiently well regularized estimate.
+By comparing the performance of our differently regularized estimators we can gain insights into which of these aspects are important in our data.
+
+As expected, all regularized estimators outperformed the sample covariance estimator substantially (Fig.~\ref{fig:04}) \Acomment{I think we should show that}.
+Say something about how shrinkage, low-rank and sparse inverse relate \Acomment{Why do we do pairwise comparisons instead of just showing median (across sites) log loss relative to the combined estimator?
+I've seen many people do this and it provides an ordering.
+I feel like we should come up with a way of ordering them somehow, even if it's under certain assumptions that may not be entirely correct.
+I'm pretty sure if we don't do it the reviewers will bring it up anyway...}.
+Finally, the combined sparse and low-rank estimator dominated all others significantly (Fig.~4), showing that both hidden units and direct interactions are important in our data.
+The improvement from the low-rank estimator to the combined one is much larger than that from the sparse inverse to the combined one, suggesting that in this dataset direct interactions contribute more strongly to the correlation structure than hidden units do.
+
+\subsection*{Relationship between functional covariance structure and circuit architecture}
+
+\begin{itemize}
+\item Linear and partial correlations versus spatial separation (lateral and vertical). Discuss whether to include or remove thresholded correlations.
+\item Magnitude of common input versus spatial location in the volume (lateral and vertical separation from center).
+\item Linear and partial correlations versus orientation preference
+\item Distribution of sparsity over sites
+\item Distribution of number of hidden units over sites
+\item Distribution of eigenvalues for sample covariance versus combined estimator
+\item etc. etc.
+
+\end{itemize}
 
 \section*{Discussion}
-\input{sections/Discussion.tex}
 
@@
 % authors may use "Analysis" 
-\section*{Materials and Methods}
-\input{sections/Methods.tex}
+\section*{Methods}
 
@@
 
-%\section*{References}
 % The bibtex filename
@@
 \section*{Figure Legends}
-\TODO{presently embedded in the text}
 
-%\section*{Tables}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
+\begin{figure}[!ht]
+    \begin{center}
+        %\includegraphics[width=8.3cm]{Figure01.pdf}
+    \end{center}
+    \caption{{\bf Illustration of regularized estimation of partial correlations.}
+        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
+        The outlined square fragment is magnified in {\bf B}.
+        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
+        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
+    }
+    \label{fig:01}
+\end{figure}
+
+\begin{figure}[!ht]
+    \begin{center}
+        %\includegraphics[width=17.35cm]{Figure02.pdf}
+    \end{center}
+    \caption{{\bf Estimators whose low-dimensional regularization targets can represent the structure of the true covariance matrix outperform other estimators.}
+        {\bf Row 1.} Graphical representations of four types of low-dimensional structures of interactions between observed neurons (green spheres) and latent units (light-shaded spheres).
+        In the \lq independent model\rq\;({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units. 
+        In \lq latent factors\rq\;({\bf B}), the correlated activity of observed cells is driven by several latent units. 
+        In \lq sparse interactions\rq\;({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
+        In \lq sparse+latent\rq\;({\bf D}), correlations arise through direct linear interactions between some pairs of observed neurons and through interactions with common latent units. 
+        {\bf Row 2.} Examples of $50\times 50$ correlation matrices corresponding to each type of low-dimensional structure. 
+        The factor model ({\bf B}) has three latent units. 
+        The partial correlation matrix of the sparse model ({\bf C}) is 73\% sparse.
+        The \lq sparse+latent\rq\;matrix has one latent unit and its direct interactions are 78\% sparse.
+        {\bf Row 3.} Examples of sample correlation matrices calculated from samples of 1000 observations taken from simulated random processes with corresponding correlation matrices from row 2.
+        {\bf Row 4.} Average normalized Gaussian log-likelihood \emph{excess loss} (\ref{eq:excess-loss}) attained by each of the five estimators as a function of sample size.
+The error bars indicate the standard error of the mean based on 30 samples.
+        {\bf Row 5.} Average normalized Gaussian log-likelihood \emph{validation loss} (\ref{eq:validation-loss}) attained by each of the five estimators.
+The values are relative to the validation loss of the estimator that matches the low-dimensional structure of the true covariance matrix.
+The error bars indicate the standard error of the mean based on 30 samples.
+    }
+    \label{fig:02}
+\end{figure}
+
+\begin{figure}[!ht]
+    \begin{center}
+        %\includegraphics[width=8.3cm]{Figure03.pdf}
+    \end{center}
+    \caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
+    Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) were presented to anesthetized mice while two-photon recordings of somatic calcium signals were collected using fast 3D random-access microscopy ({\bf B}).
+The visual stimulus included an initial period with 16 directions of motion for orientation tuning followed by a longer (15--20 min) period of stimulation with only 2 or 5 directions of motion for the computation of the noise correlation matrix.
+    {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. The inferred firing rate binned in 150 ms intervals are indicated by red ticks below each trace.
+    {\bf D.} The raster plot of the inferred firing rates, binned in 150 ms intervals, from the entire population from the first (left) and last (right) minute of the entire recording.
+The traces from {\bf C} are highlighted in red.
+    {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site.
+    {\bf F.} The noise correlation matrix of the activity of the neural population. 
+    {\bf G.} The histogram of the noise correlation coefficients with the mean indicated by the red line.
+}
+    \label{fig:03}
+\end{figure}
+
+\begin{figure}[!ht]
+    \begin{center}
+        %\includegraphics[width=8.3cm]{Figure04.pdf}
+    \end{center}
+    \caption{{\bf The sparse+lowrank estimator $C_{\sf sparse+lowrank}$ outperforms the other estimators on neural data.}
+    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_0$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+lowrank}$. 
+    The histograms are based on 31 imaged sites in 24 mice. 
+    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+lowrank}$ over the other estimators. 
+    The green arrows indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
+    }
+    \label{fig:04}
+\end{figure}
+
+\begin{figure}[!ht]
+    \begin{center}
+        %\includegraphics[width=17.35cm]{Figure05.pdf}
+    \end{center}
+    \caption{{\bf Example of low-dimensional correlation structure revealed by the sparse+low-rank estimator.}
+    {\bf A.} The regularized estimate of the correlation matrix (top-right) closely approximates the sample correlation matrix (bottom left). 
+    This close approximation is  also demonstrated by the scatter plot of the correlation coefficients produced by the two estimates ({\bf D}). 
+    However, the partial correlation matrices from the two estimate show more pronounced differences ({\bf B} and {\bf E}). 
+    {\bf C.} Furthermore, the partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.2\% off-diagonal zeros (bottom-left) and low-rank component of rank 15 (top-right).
+    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations.
+The gray interval indicates the range of correlations containing 82.2\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
+This interval contained 58.9\% of the partial correlations.
+    {\bf G.} The graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
+The line density is proportional to the magnitude of the correlation.
+    {\bf H.} A subset of neurons from the center of the cluster shown in {\bf G} showing the regularized partial correlations.
+    {\bf I.} The same subset with sample correlations thresholded to match the sparsity of the regularized interactions.
+}
+\label{fig:05}
+\end{figure}
+
+\begin{figure}[!ht]
+    \begin{center}
+        %\includegraphics[width=17.35]{Figure06.pdf}
+    \end{center}
+    \caption{{\bf Properties of sparse+low-rank regularized estimates from all imaged sites}
+    {\bf A.} The average sample correlations vs.~average partial correlations for each imaged site. In each plot, the red asterisk indicates the site shown in figures \ref{fig:03} and \ref{fig:05}.
+    {\bf B.} The average node degree for sparse partial correlations vs.~population size in each imaged site. 
+    {\bf C.} The number of inferred latent units vs.~population size in each imaged site.
+    {\bf D.} The number latent units vs.~average node degree for sparse partial correlations for each site.
+}
+\label{fig:06}
+\end{figure}
+
+\begin{figure}[!ht]
+    \begin{center}
+        %\includegraphics[width=17.35]{Figure07.pdf}
+    \end{center}
+    \caption{{\bf Dependence of correlations and partial correlations on orientation tuning differences and physical distance between cell pairs.}
+    {\bf A--C.} Average sample correlations (black) and regularized partial correlations (red) between pairs of cells in the example site shown in previous figures.
+The correlations were normalized by the respective average correlations (\ref{fig:06}\;A).
+    {\bf A.} Average correlations between pairs of neurons tuned to orientation with differences in preferred orientation in the intervals of 0--15$^\circ$, 15--45$^\circ$ and 45--90$^\circ$. 
+    {\bf B.} Average correlations between pairs of neurons located at the same depth ($\pm$25$\mu$m) separated by lateral distances in the intervals of 0--25 $\mu$m, 25--75 $\mu$m, 75--150 $\mu$m, and 150+ $\mu$m.
+    {\bf C.} Average correlations between pairs of neurons displaced laterally by less than 25 $\mu$m separated in depth by distances in the intervals of 0-25 $\mu$m, 25--60 $\mu$m, and 60+ $\mu$m.
+    {\bf D--F.} Same measurements as {\bf A--D} averaged across multiple sites.
+Only sites that had at least 20 qualifying cell pairs in each of the intervals were included in the averages.
+The error bars indicate the standard error of the mean (often too small to be seen).
+    {\bf G--I.} Normalized connectivity of positive (green) and negative (dark red) interactions from the sparse component of the regularized partial correlations in the example site show in previous figures.
+Normalized connectivity was computed as the fraction of pairs connected by interactions of corresponding signs in each interval divided by fraction of non-zero interactions across the entire site.
+The intervals are identical to those in {\bf A--C}.
+    {\bf J--L.} Same measurements as in {\bf G--I} averaged across multiple sites.
+The error bars indicate the error of the mean.
+Only sites that had at least 20 qualifying pairs in each interval were included in the averages.
+}
+\label{fig:07}
+\end{figure}
 

### 2013-11-30 plos.tex
consolidated entire writeup into one text file, removed ./sections

@@
 
-We can also use a \lq sparse+low rank\rq inverse covariance representation to account for both common fluctuations and pairwise interactions.  
+We can also use a `sparse+low' rank inverse covariance representation to account for both common fluctuations and pairwise interactions.  
 To select the optimal structure of neural correlations in a local neural circuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance.
@@
 where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent).  
-The sample covariance matrix is unbiased, \emph{i.e.} $\E{C_0}-\Sigma=0$.
+The sample covariance matrix is unbiased, \emph{i.e.} $\E{C_0}=\Sigma$.
 For finite sample sizes, however, $C_0$ is not as close to $\Sigma$ as a number of biased estimators.
@@
 % authors may use "Analysis" 
+Fundamental questions in systems neuroscience concern the relationship between the function of neural microcircuits and their cytoarchitecture: circuit topology, cell types, and patterns of synaptic connectivity.
+
+Considerable progress has been made in sensory areas where the functional characterizations of cells are defined by their reproducible responses to external stimuli (reviewed in \cite{Reid:2012}).
+
+Many key questions in neuroscience revolve around the relationship between the cytoarchitecture of neural microcircuits and the functional organization of their activity \cite{Reid:2012}.
+Investigators in this line of inquiry aim to construct networks of functional associations within groups of neurons inferred from observations of their activity under a variety of conditions.
+The inferred functional network  could then related to the synaptic connectivity patterns, cell types, cell properties, or the cells' spatial arrangment in the attempt to uncover organizational principles of neural computation.
+In addition, the network itself can serve as subtrate for subsequent graph-theoretical analysis \cite{Feldt:2011}
+
+Advances in electrophysiology and optical imaging have enabled simultaneous recordings from dozens to hundreds or thousands of cells. 
+In particular, recent advances in two-photon imaging of calcium signals have allowed in vivo recordings from nearly every neuron brain-wide in zebrafish \cite{Leung:2013,Ahrens:2013} and in a 3D volume of a few hundred microns in diameter in mouse neocortex \cite{Katona:2012,Cotton:2013}.
+Ambitious projects currently under way aim to record the spiking activity of large fractions of cells from entire circuits and systems in behaving animals \cite{Alivisatos:2012}.
+Direct observations of the population activity  of entire circuits open new possibilities for incisive statistical descriptions of the functional connectivity in the circuit.
+
+One general approach is to construct statistical models reflecting hypothesized models of neural circuit function, including temporal dynamics, nonlinear interactions between neurons, and background network activity \cite{Pillow:2008,Buonomano:2009,Yu:2009}.
+Another general approach is to search for alternative statistics that best describe the population activity without assuming a specific model.
+Such evaluations rely on constructing surrogate datasets \cite{Okun:2012} or maximum-entropy distributions that reproduce the statistic of interest \cite{Ganmor:2011,Tkacik:2012} with subsequent testing of how well such constructs can reproduce other observed properties of population activity.
+
+Such more sophisticated statistical descriptions are unlikely to entirely replace neural correlations as the initial descriptive statistic of population activity in experimental neuroscience.
+Neural correlations have several strengths that will ensure their continued prominence: (a) correlations are well established, familiar, and intuitive to most researchers in biological sciences, (b) correlations are easy to estimate as they are less susceptible to the curse of dimensionality than more sophisticated models, (c) correlations make relatively weak mechanistic assumptions (at the cost of not revealing many potentially important aspects), and (d) correlations serve as input into other models of population activity or decoding algorithms.
+
+Linear correlations between pairs of neurons are among the most common and familiar descriptions of functional associations in networks and of the collective activity of neuronal populations.
+For example, it is tempting to consider the network constructed from the highest correlations in the recorded population (e.g.~\cite{Malmersjo:2013}) for subsequent graph-theoretical analysis \cite{Feldt:2011}.
+In functional genomics, networks constructed from highest correlations are called \emph{relevance networks}.
+
+However, pairwise correlations unreliable proxies of direct functional association such as direct synaptic connectivity as they can arise due to a wide variety of physiological interactions including direct synaptic interactions, indirect chains of synaptic connections, common inputs, correlations in inputs, synchrony, fluctuations in global network states, and others \cite{Shadlen:1998,Ostojic:2009,Pernice:2011,Schneidman:2006}.
+
+Other experimental evidence suggests that detailed interactions in local microcircuits have only weak effects on overall population activity, which is then best characterized by collective features such as global population dynamics or population synchrony \cite{Okun:2012,Tkacik:2012,Tkacik:2013}.
+From this point of view, the imporant aspects of the correlation structure lie it its global aspects, such as the eigenspectrum, while the individual pairwise correlations are not particularly meaningful.
+
 \section*{Methods}
+\subsection*{Evaluation of covariance matrix estimators}
+The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.  
+The loss function quantifies the error of the estimate, \emph{i.e.}~the deviation of $\hat\Sigma$ from $\Sigma$. The loss function must attain its minimum  when $\hat\Sigma = \Sigma$.
+
+For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function:
+\begin{equation}\label{eq:loss}
+\loss{\hat\Sigma,\Sigma} = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
+\end{equation}
+This choice is motivated by mathematical convenience.
+Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, and quadratic loss \cite{James:1961,Fan:2008}.
+We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.
+
+The aim of our project is to identify which covariance matrix estimator minimizes the expected loss 
+\begin{equation}
+r = \E{\loss{\hat\Sigma, \Sigma}}
+\end{equation}
+which is known as the \emph{risk} of $\hat\Sigma$.
+
+In practice, the true value $\Sigma$ is not accessible and an estimators' risks must be estimated from the data.
+This may be accomplished through \emph{validation}.
+\Kcomment{If you will discuss validation later, point to it.}
+Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$.
+
+$\loss{\cdot,\cdot}$ is additive in its second argument so that \Kcomment{You frequently use ``such that'' instead of ``so that''}
+ \begin{equation}\label{eq:additivity}
+ \loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
+ \end{equation}
+Then \emph{validation loss}  \TODO{check appropriate term}
+\begin{equation}\label{eq:validationLoss}
+\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
+\end{equation}
+is an unbiased estimate of the risk:
+ \begin{equation}\label{eq:empiricalRisk}
+\E[(\hat\Sigma, \hat\Sigma_0^\prime)] {\hat\ell} 
+= \E[(\hat\Sigma, \hat\Sigma_0^\prime)]{\loss{\hat\Sigma,\hat\Sigma_0^\prime}}
+= \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}
+= \E{\loss{\hat\Sigma,\Sigma}} = r
+ \end{equation}
+Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
+\Kcomment{This is probably not a very insightful question -- but why do we assume that the sample covariance
+is what we compare $\hat\Sigma$ when we validate the results.  Our point is that this estimate is not very good.}
+
+Other popular loss functions such as entropy loss do not comply with Eq.~\ref{eq:additivity} and their validation loss is not an unbiased estimate of risk.
+
+Since $\loss{\cdot,\cdot}$ is equivalent to negative normal log likelihood, the above derivation has led to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes cross-validated normal log likelihood.
+
+\subsection*{Regularization}
+Under many loss functions\footnote{
+The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}. 
+However, the equality does hold for its close cousin, Stein's \emph{entropy loss}, which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$.
+This deficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually. 
+The bias-variance decomposition is presented here to motivate the use of regularization.}, 
+the risk can be decomposed as the sum
+\begin{equation}\label{eq:bias-variance}
+    r = b + \varepsilon
+\end{equation}
+of \emph{approximation error} (``bias'' or systematic error)
+\begin{equation}
+   b = \loss{\bar\Sigma,\Sigma}
+\end{equation}
+and \emph{estimation error} (``variance'') 
+\begin{equation}
+   \varepsilon = \E[\hat\Sigma]{\loss{\hat\Sigma, \bar\Sigma}}
+\end{equation}
+where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.
+
+An unbiased estimator, such as the sample covariance,  $\hat\Sigma_0,$ makes $\bar\Sigma=\Sigma$ and thereby minimizes approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
+
+The estimator risk can be reduced by \emph{regularization} -- a deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
+A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimate to minimize the estimator's risk.  
+Various regularization schemes focus on the dimensionality reduction part \TODO{rephrase} by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.
+Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}. 
+Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.
+
+\subsection*{Estimators}
+We evaluated four regularized estimators denoted here as A, B, C, and D.
+Their target estimates correspond to covariance matrices of Gaussian graphical models depicted in Figure 2 where green spheres represent the recorded neurons,  light-colored spheres represent latent units of the graphical model, and edges connecting them represent conditional dependencies or interactions.
+
+A \emph{graphical model} is a multivarate probability distribution with a specified graph of conditional dependencies between its variables \cite{Koller:2009}.
+When this distribution is a multivariate normal distribution, the model becomes a \emph{Gaussian graphical model} (GGM) or, equivalently, a \emph{Gaussian Markov Random Field}.
+Gaussian graphical models have a straightforward relationship to their covariance matrix $\Sigma$:  Zero elements in the inverse of $\Sigma$ indicate conditional independence between the corresponding pair of variables.
+\Kcomment{It may be important to note that we are not fitting Gaussian graphical models, however.}
+
+Fitting a graphical model to data involves two tasks: (1) the selection of the set of non-zero covariances or \emph{covariance selection} \cite{Dempster:1972} and (b) the fitting of the non-zero elements.
+
+\subsubsection*{Estimator $\mathcal A$: Shrinkage toward diagonal.}
+\emph{Estimator $\mathcal A$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward that of the independent model shown in Fig.~\ref{fig:02}A.
+The covariance matrix of a GGM in which all units are independent will be diagonal.
+For example, we  could leave the sample variances on the diagonal set the off-diagonal terms to zero. The target estimate
+is then 
+\begin{equation}\label{eq:sampleVariance}
+\hat T= \hat\Sigma_0\circ I
+\end{equation}
+where $\circ$ denotes the entrywise matrix product (Hadamard product). 
+Alternatively, the target could be a diagonal matrix with all variances equal to the average sample variance, $v = \frac 1 p \Tr(\hat\Sigma_0)$, so that
+\begin{equation}\label{eq:equalVariance}
+\hat T=\frac 1 p \Tr(\hat\Sigma_0) I
+\end{equation}
+The equal-variance target in Eq.~\ref{eq:equalVariance} has only one degree of freedom and thus lower estimation error but higher bias than the target with sample variances in Eq.~\ref{eq:sampleVariance}.
+It is not immediately clear which target is better in a particular application.
+We therefore use the linear mixture of the two targets paramaterized by the mixing coefficient $\alpha\in[0,1]$,
+\begin{equation}
+\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I.
+\end{equation}
+The estimator $\hat\Sigma_{\alpha,\lambda}$ is the linear shrinkage of the sample covariance matrix $\hat\Sigma_0$ toward $\hat T_\alpha$ controlled by the mixing proportion $\lambda\in[0,1]$,
+\begin{equation}
+\hat\Sigma_{\alpha,\lambda}^\mathcal{A} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha.
+\end{equation}
+This covariance matrix estimator is implemented by R's {\tt corpcor} package \cite{Schaefer:2010} and is used widely~\cite{Schafer:2005}.
+
+We estimated the optimal values of hyperparameters $ \alpha$ and $ \lambda$  by  cross-validation within the training dataset.
+\TODO{more detail?} \TODO{discuss analytic optimization of shrinkage intensities.}
+
+\Kcomment{Yes -- you give no details about the cross-validation.  You should provide some details at least in 
+one case, or have a separate section that discusses how you implemented it.}
+
+\subsubsection*{Estimate $\mathcal B$: Shrinkage toward a factor model}
+\emph{Estimator $\mathcal B$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward a target estimated by \emph{factor analysis} with $d$ factors, which corresponds to graphical models of the kind in Fig.~\ref{fig:02}B.
+For multivariate normal distributions, factor analysis estimates a graphical model in which the observed units do not interact directly but are influenced by $d$ common inputs.
+
+The target estimate is then 
+\begin{equation}
+\hat T_d = \hat L_d + \hat \Psi
+\end{equation}
+where $\hat L_d$ is a $p\times p$ matrix of rank $d$ and $\hat \Psi$ is a $p\times p$ diagonal matrix. The value of $\hat T_d$ is found by minimizing $\loss{\hat T_d, \hat\Sigma_0}$. 
+\Kcomment{I would give a more explicit representation of $\hat L_d$ as a product to show that the factors can be understood as common input.}
+
+Just as with estimator $\mathcal A$, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$,
+\begin{equation}
+\hat\Sigma_{d,\lambda}^\mathcal{B} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d,
+\end{equation}
+and the optimal values of $d$ and $\lambda$ are found by cross-validation within the training set.
+
+Factor analysis has long been used in natural sciences to infer the influence of a one or several latent variables exerting linear effects on a larger population of observed variables \cite{Jvreskog:1996}.
+In neuroscience, the hypothesis that the population activity in small local circuits is largely driven by a few latent factors has been investigated \cite{Yu:2009,Ecker:2013}.
+Estimation of covariance matrices by shrinkage toward a factor model has been used in finance for portfolio risk assessment \cite{Ledoit:2003,Fan:2008}.
+
+\subsubsection*{Estimator $\mathcal C$: Spare inverse}
+\emph{Estimator $\mathcal C$} produces a low-dimensional estimate by finding an approximation of $\hat\Sigma_0$ that has many zeros in its inverse.
+This approximation problem is known as \emph{covariance selection} \cite{Dempster:1972}.
+Covariance matrices of this family correspond to graphical models of the type depicted in Fig.~\ref{fig:02}C.
+
+For Gaussian Graphical Models, zeros in the inverse covariance matrix indicate statistical independence between the corresponding pairs.
+Thus the inverse covariance matrix $S=-\Sigma^{-1}$ is also the adjacency matrix of the GGM, indicating which pairs are conditionally dependent.
+\TODO{discuss how quickly (or slowly) this breaks down with deviations from gaussianity, e.g.~MaxEnt models, Ising models.} \TODO{Also, motivate this from the perpsective of linear regression as in \cite{Varoquaux:2012}.}  \Kcomment{This is a bit confusing -- in the first sentence, you have statistical independence, in the second, you talk about conditional independence.
+You want to emphasize that a 0 inverse means that -- in the GGM -- the two nodes only ``talk'' indirectly through other nodes.} The off-diagonal elements of $S$ are proportional to partial pairwise correlations between corresponding pairs of neurons.
+\TODO{Expand.} Regularization of covariance matrix estimation by sparsifying the inverse of the estimate has been using in functional genomics \cite{Schafer:2005,Peng:2009} and fMRI functional connectivity \cite{Varoquaux:2012,Ryali:2012}.
+
+The covariance matrix estimate is then 
+\begin{equation}
+\hat\Sigma_\rho = \hat S^{-1}
+\quad\mbox{with}\quad
+\hat S = \argmin\limits_{\substack{S\succ 0 \\ \|S\|_0 \le \rho}} \loss{S^{-1},\hat\Sigma_0}   
+\end{equation}
+where $S\succ 0$ signifies the constraint that $S$ be positive-definite. $\|S\|_0\le\rho$ signifies the constraint  that matrix $S$ have at most $\rho$ non-zero coefficients.
+Solving this approximation in this form is computationally challenging.
+The problem can be made computationally tractable by relaxing \TODO{use "complex relaxation"} the $L_0$ norm to the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.
+The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator
+\begin{equation}
+\hat\Sigma_\lambda^\mathcal{C} = \hat S^{-1}
+\quad\mbox{with}\quad
+\hat S = \argmin\limits_{S\succ 0} \loss{S^{-1},\hat\Sigma_0} + \lambda \|S\|_1
+\end{equation}
+where $\lambda \|S\|_1$ is the $L_1$ norm of the matrix. \TODO{clarify}. \Kcomment{Yes - at least define the norm, if you are already taking about it. No need to expand a lot more, though.}
+
+Just as in the previous estimators, the optimal value of the regularization parameter $\lambda$ is found from training data by cross-validation.
+\Kcomment{I would say this up front, so you don't have to repeat it.
+Again, I think more details about cross-validation should be given.}
+
+\subsubsection*{Estimator $\mathcal D$: Sparse + low-rank inverse}
+\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by approximating $\hat\Sigma_0$ with a matrix whose inverse is the sum of a spase matrix and a low-rank matrix.
+Covariance matrices of this form correspond to Gaussian graphical models of the form depicted in Fig.~\ref{fig:02}D.
+
+The inverse $S=\Sigma^{-1}$ of the covariance matrix of a GGM constitutes the graph of conditional dependencies (interactions) between the units.
+With the assumption of sparse organization of interactions in the network, $S$ is assumed to be sparse.
+When the observed units are part of a larger GGM that includes other latent units, let the negative inverse $S^*$ of the joint distribution of both observed and latent units be composed of four partitions
+\begin{equation}
+S^* = 
+\begin{pmatrix}
+S_{11} & S_{12} \\
+S_{12}^\T & S_{22} 
+\end{pmatrix}
+\end{equation}
+where partition $S_{11}$ describes the interactions between the visible units, $S_{12}$ describe the interactions between visible and latent units, and $S_{22}$ describes the interactions between the latent units.
+Then the covariance matrix of between the observed unit is 
+\begin{equation}
+\Sigma = \left(S_{11} - S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
+\end{equation} 
+The rank of matrix $S_{12}S_{22}^{-1}S_{12}$ equals the number of latent units.
+Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $ S_{12}S_{22}^{-1}S_{12}$ with a low-rank matrix $L= S_{12}S_{22}^{-1}S_{12}$.
+
+Then we obtain estimator $\hat\Sigma_{d,\rho}$, parameterized by the sparsity parameter $\rho$ and  the number of latent units $d$:
+\begin{equation}
+\hat\Sigma_{d,\rho} = \left(\hat S-\hat L\right)^{-1}
+\quad\mbox{where}\quad
+(\hat S, \hat L) = \argmin\limits_{\substack{S,L\\ (S-L)\succ 0 \\ \|S\|_0\le\rho \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0}
+\end{equation}
+
+Again, in this formulation, the approximation problem is computationally intractible.
+The problem can be made convex by replacing the $L_0$ norm with $L_1$ norm.
+The resulting algorith, known as \emph{latent-variable graphical lasso} or \emph{LV-GLASSO} \cite{Chandrasekaran:2010,Ma:2013}, produces estimator $\hat\Sigma_{d,\lambda}^{\mathcal D}$:
+\begin{equation}
+\hat\Sigma_{d,\lambda} ^{\mathcal D} = \left(\hat S-\hat L\right)^{-1}
+\quad\mbox{where}\quad
+(\hat S, \hat L) = \argmin\limits_{\substack{ S,L \\ (S-L)\succ 0 \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0} + \lambda \| S \|_1 
+\end{equation}
+The optimal values of hyperparameters $\lambda$ and $d$ are found cross-validation within the training dataset.
 
@@
 \section*{Acknowledgments}
+Genevera Allen
+
+Eftychios Pnevmatikakis
 
@@
     \begin{center}
-        %\includegraphics[width=8.3cm]{Figure01.pdf}
+        %\includegraphics[width=8.3cm]{./figures/Figure01.pdf}
     \end{center}
@@
         {\bf Row 1.} Graphical representations of four types of low-dimensional structures of interactions between observed neurons (green spheres) and latent units (light-shaded spheres).
-        In the \lq independent model\rq\;({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units. 
-        In \lq latent factors\rq\;({\bf B}), the correlated activity of observed cells is driven by several latent units. 
-        In \lq sparse interactions\rq\;({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
-        In \lq sparse+latent\rq\;({\bf D}), correlations arise through direct linear interactions between some pairs of observed neurons and through interactions with common latent units. 
+        In the `independent model' ({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units. 
+        In `latent factors' ({\bf B}), the correlated activity of observed cells is driven by several latent units. 
+        In `sparse interactions' ({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
+        In `sparse+latent' ({\bf D}), correlations arise through direct linear interactions between some pairs of observed neurons and through interactions with common latent units. 
         {\bf Row 2.} Examples of $50\times 50$ correlation matrices corresponding to each type of low-dimensional structure. 
@@
         The partial correlation matrix of the sparse model ({\bf C}) is 73\% sparse.
-        The \lq sparse+latent\rq\;matrix has one latent unit and its direct interactions are 78\% sparse.
+        The `sparse+latent' matrix has one latent unit and its direct interactions are 78\% sparse.
         {\bf Row 3.} Examples of sample correlation matrices calculated from samples of 1000 observations taken from simulated random processes with corresponding correlation matrices from row 2.
@@
     \begin{center}
-        %\includegraphics[width=8.3cm]{Figure03.pdf}
+        %\includegraphics[width=8.3cm]{figures/Figure03.pdf}
     \end{center}

### 2013-11-30 plos.tex
Updated Abstract, Authors' Summary, and Introduction.

@@
 % Please keep the abstract between 250 and 300 words
-Correlations between the spiking activity of pairs of neurons are among the most familiar descriptive statistics of neural activity.
-Multineuronal recordings provide richer information than the equivalent number of neuron pairs.
-For example, the full covariance  matrix reveals correlated activity across the entire population as well as partial correlations between pairs.
-Estimation of covariance matrices can be improved by regularization, \emph{i.e.}\;by imposing some kind of structure.
-Optimal regularization must be determined empirically since its effect depends on how closely the imposed structure matches the underlying regularities.
-For example, we can use low-rank parameterizations of the covariance matrix to account for common fluctuations across the recorded population.
-Conversely, if correlations are strongly influenced by a small fraction of pairwise linear associations between the observed neurons, we can impose sparsity on the inverse covariance matrix.
+Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons in behaving animals.
+It is anticipated that correlations measured in such recordings will uncover aspects of functional organization of neural circuits.
+However, estimation and interpretation of large correlation matrices from finite recordings can be challenging.
+Estimation can be improved by regularization: the biasing of the estimate toward a low-dimensional, less variable approximation.
+The amount of improvement depends on how closely the reduced approximation captures the dominant interactions with fewest terms.
+Therefore, the selection of the most efficient estimator is an empirical question dependent on the system under investigation.
 
-We can also use a `sparse+low' rank inverse covariance representation to account for both common fluctuations and pairwise interactions.  
-To select the optimal structure of neural correlations in a local neural circuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance.
-We inferred instantaneous firing rates in 200 ms bins from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.
-Each covariance estimator was optimized and evaluated by cross-validation.
-As expected, covariance shrinkage reliably outperformed the sample covariance estimate.
-In turn, factor analysis-based estimates significantly outperformed covariance shrinkage.
-Yet sparse inverse covariance with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage estimators.
-The superior performance of the sparse inverse covariance estimator suggests the relative importance of detailed network interactions over common diffuse input in the circuit we studied.
+We compared regularized correlation matrix estimators biased toward four respective families of low-dimensional correlation structures: independent, latent factors, sparse partial correlations, and sparse partial correlations with latent factors.
+The estimators' performance was evaluated on the noise correlation matrices from spatially compact groups of 91--314 neurons (average 206) in mouse visual cortex (31 sites in 24 animals) acquired by \emph{in vivo} fast 3D random-access two-photon imaging of calcium signals.
+Recordings lasting between 15 and 20 minutes were deconvolved and binned at 150 ms.
+Each estimator was optimized and evaluated by nested cross-validation.
+We found that both sparse partial correlations and latent factors were required for efficient estimation.
+In addition to reducing estimation error, optimized estimates provided compact representations of the correlation structure consisting of a sparse network of pairwise interactions and global diffuse fluctuations.
+The density of positive interactions in these networks decreased rapidly with distance and with difference in orientation preference, whereas negative interactions were less selective.
+The physiological interpretation of these components remains in question and must be addressed by future experiments where synaptic connectivity, cell types, and local field potentials may be used to corroborate such interpretations.
 
+\section*{Author Summary}
 % Please keep the Author Summary between 150 and 200 words
@@
 % Author Summary not valid for PLoS ONE submissions.   
-\section*{Author Summary}
-Covariance matrices of the activity of populations of neurons are useful descriptions of the functional organization of neural circuits with implications for stimulus coding and neural circuit architecture.
-Estimates of the covariance matrix can be dramatically improved by \emph{regularization}, \emph{i.e.}\;the biasing of the usual estimate toward one of several possible low-dimensional approximations, the \emph{target estimate}.
-However, the amount of improvement depends on how closely the selected approximation reflects the dominant interactions in the circuit.
-In simulation, we demonstrated that no regularized estimate was universally superior to others: each performed best when the target estimate was the closest match to the structure of the data generating process.
-We then evaluated covariance matrix estimates on the calcium activity recorded \emph{in vivo} from large populations of closely spaced neurons  in mouse primary visual cortex, after subtracting the average stimulus response.
-Using cross-validation, we found that the best estimate was produced when the covariance matrix was approximated as that of a sparse Gaussian graphical models with a small number of latent units. 
-Thanks to the paucity of their parameters, these optimally estimated covariance matrices also help interpretation of the covariance structure and allow relating the functional organization of the neural circuit to its cytoarchitecture.
-
+Correlation matrices of the activity of populations of neurons are useful descriptors of their functional organization with implications for stimulus coding and circuit architecture.
+Estimates of the correlation matrix can be dramatically improved by \emph{regularization}: the biasing of the estimate toward one of several possible low-dimensional approximations.
+Using synthetic data, we demonstrated that different estimators performed better or worse than others depending on how well their regularization targets captured the structure of the true correlation matrix.
+Then, we compared the performance of several estimators on the population calcium activity of relatively large and dense groups of neurons in mouse visual cortex under visual stimulation.
+We found that the best estimator for such neural data was composed of a sparse network of partial correlations between pairs of neurons combined with several latent units interacting with the entire population.
+We hypothesize that such succinct representations of the correlation structure may more accurately reflect the circuit's functional organization than the conventional correlations.
 \section*{Introduction}
 Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Renart:2010}.
-For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been of particular interest.  
-Noise correlations can have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}, and have been interpreted to indicate detailed and specific functional organization.
-Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity and stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, cell-type specificity (?), progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
+For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}.
+In addition, neural correlations are hypothesized to reflect aspects of functional connectivity in neural circuits.
+Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
 
-However, neural correlations do not come with ready or unambiguous mechanistic interpretations.
-Theoretical and simulation studies have shown that neural correlations, rise at various temporal scales from combinations of multiple underlying mechanisms.
-These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \cite{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
+However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.
+Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including  direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
 
-Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \cite{Shadlen:1998,Zohary:1994}.
-Multineuronal recordings allow estimation of covariance matrices of large populations of neurons.
-Such estimates provide more information than the equivalent number of pairwise correlations assessed in isolation.
-Indeed, the correlation matrix is greater than the sum of its parts: it can be transformed into other representations that accentuate different aspects of the correlation structure and may suggest different mechanistic interpretations.
-For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients.
-In contrast, the off-diagonal elements of the inverse of the correlation matrix constitute scaled partial correlations between neuron pairs, which reflect their specific linear dependencies, after accounting for the activity of all the other recorded cells; a strong interaction between a pair of neurons may be expressed by a single partial correlation but its effects may propagate to multiple correlations and eigenvalues.
-The inverse of the covariance matrix plays an important role in decoding schemes such as linear discriminant analysis, for example.
-The mutliple  representations of the covariance matrix with their alternative interpretations add both complications and opportunities into the search for fundamental regularities in neural population activity.
+A correlation matrix provides more information than the equivalent number of pairwise correlations assessed in isolation.
+Yet early studies of neural correlations were based on measurements from isolated pairs of neurons; the effects of correlations on neural coding were then extrapolated in simulations and theoretical studies \cite{Shadlen:1998,Zohary:1994}.
+Modern multineuronal recordings from increasingly large and dense subsets of neurons in intact circuits allow estimating the entire correlation matrix of large populations of highly interconnected neurons.
+Such estimates can be transformed into various representations that express different aspects of the correlation structure and may suggest different mechanistic interpretations.
+For example, factor analysis or principle component analysis may reveal common activity fluctuations across the entire population \cite{Yu:2009}.
+The correlation matrix also allows computing the \emph{partial correlations} between specific pairs of neurons, \emph{i.e.}\;correlations that remain after accounting for the linear effects of all other recorded cells.
+Under several assumptions (predominance of linear effects, sufficiently complete sampling of interacting neurons, and sufficient amount of data), partial correlations may better reflect the direct statistical association between pairs of neurons.
 
-With large numbers of recorded cells, the usual estimations of the covariance matrix from finite recordings become ill-conditioned.
-Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically.
-For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \cite{Ledoit:2004}.
-Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.
+Large sizes of recorded populations present major challenges to the estimation and interpretation of correlations.
+As the population size increases, the amount of recorded data grows only linearly whereas the number of estimated coefficients grows quadratically.
+This mismatch leads to increased opportunities for spurious correlations, overestimation of shared activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned estimates of the partial correlations \cite{Schafer:2005}.
 
-With this study, we pursue two related goals: 
-\begin{enumerate}
-	\item Devise more efficient estimators of neural covariance matrices in recordings from large populations of neurons.
-	\item Aid interpretation of neural covariance matrices.
-\end{enumerate}
+In this study, we pursued two related aims: (a)~improved estimation of neural correlation matrices and (b)~discovery of low-dimensional structure of correlations in recordings of multineuronal activity to facilitate their interpretation.
 
-To select the best covariance matrix estimate for a specific neural circuit, we evaluate the performance of four estimators using different regularization schemes, each motivated by  a different hypothesis for the origin of neural correlations.
-\emph{Regularization} is the deliberate biasing the solution toward a simplified, low-dimensional (`sparse') \emph{target estimate} \cite{Schafer:2005,Bickel:2006}.
-Regularized estimators allow reaching a favorable tradeoff between bias and variance in order to minimize the total error.
+Both aims benefit from \emph{regularization}: the deliberate biasing (\emph{`shrinkage'}) of the estimate toward one of several possible low-dimensional approximations called here \emph{target estimates} \cite{Schafer:2005,Bickel:2006}.
+Low-dimensional estimates are less susceptible to sampling noise in the empirical data and are thus less variable.
+However, they are also biased away from the true correlation matrix toward its projection onto the space of low-dimensional approximations.
+In what is called \emph{Stein's paradox}, arbitrary low-dimensional targets can be used to reduce
 

### 2013-12-01 plos.tex
Edited the introduction.

@@
 The correlation matrix also allows computing the \emph{partial correlations} between specific pairs of neurons, \emph{i.e.}\;correlations that remain after accounting for the linear effects of all other recorded cells.
-Under several assumptions (predominance of linear effects, sufficiently complete sampling of interacting neurons, and sufficient amount of data), partial correlations may better reflect the direct statistical association between pairs of neurons.
+Under several assumptions (predominance of linear effects ( sufficiently complete sampling of interacting neurons, and sufficient sample size), partial correlations may better reflect the direct statistical associations or interactions between pairs of neurons.
+In multivariate normal distributions, zero partial correlations between pairs of variables indicate conditional independence or lack of interaction between the pair and the network of non-zero interactions constitute Gaussian Graphical Models or Gauss-Markov Random Fields \cite{Koller:2009}.
+With other distributions, the correspondence between conditional dependence and partial correlations diminishes or breaks down \cite{Loh:2012}.
+Estimation of networks of partial correlations (sometimes called \emph{association networks}) have been applied to map gene interactions \cite{Schafer:2005,Peng:2009} or to establish functional connectivity between brain regions from fMRI data \cite{Varoquaux:2012,Ryali:2012}.
 
-Large sizes of recorded populations present major challenges to the estimation and interpretation of correlations.
+Multineuronal recordings also present major challenges for the estimation and interpretation of correlations.
 As the population size increases, the amount of recorded data grows only linearly whereas the number of estimated coefficients grows quadratically.
@@
 
-Both aims benefit from \emph{regularization}: the deliberate biasing (\emph{`shrinkage'}) of the estimate toward one of several possible low-dimensional approximations called here \emph{target estimates} \cite{Schafer:2005,Bickel:2006}.
-Low-dimensional estimates are less susceptible to sampling noise in the empirical data and are thus less variable.
-However, they are also biased away from the true correlation matrix toward its projection onto the space of low-dimensional approximations.
-In what is called \emph{Stein's paradox}, arbitrary low-dimensional targets can be used to reduce
+Estimation of the correlation matrix relies on estimating the means and variances of the activity of individual cells.
+To estimate the mean activity, we resort to the usual \emph{sample mean} without attempting to improve it.
+An estimate of the \emph{covariance matrix}  combined estimate of the variances and correlations constitutes an estimate covariance matrix.
+The problem can then be formulated as finding the optimal estimation of the covariance matrix given the estimate of the mean firing rates.
 
+Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward one of several possible low-dimensional approximations called here \emph{target estimates} \cite{Schafer:2005,Bickel:2006}.
+The usual estimate, the \emph{sample covariance matrix} has the advantage of being unbiased but, on average, it falls far from the true covariance matrix because of its sensitivity to sampling noise in the data.
+Low-dimensional estimates of various forms are less susceptible to sampling noise but are also liable to introduce their respective biases away from the true covariance matrix.
+Regularization works by striking a favorable balance between bias and variability.
+Regularization can produce \emph{some} improvement even with an arbitrary target estimate that has no relation to the true covariance structure in the data.
+Yet, when the target estimate is well suited for capturing the important features of the true covariance matrix with few terms, it will introduce minimal bias and outperform other estimators.
+
+To illustrate key concepts, Figure \ref{fig:01} illustrates a regularization scheme based on \emph{covariance selection} \cite{Dempster:1972}}.
+In covariance selection, the estimate is produced by fitting only an optimally selected subset of the coefficients of the \emph{precision matrix} (the inverse of the covariance matrix) while setting the remaining coefficients to zero.
+Scaling the precision matrix so that its diagonal is composed of all $-1$ produces the matrix of partial correlations.
+Thus zeros in the precision matrix produce zero partial correlations between the corresponding pairs of neurons.
+Panel \ref{fig:01}\,A depicts the sample correlation matrix of a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex.
+Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are much higher than typical firing rate correlations on shorter temporal scales.
+Panel \ref{fig:01}\,B magnifies the fragment of the sample correlation matrix outlined in \ref{fig:01}\,A whereas \ref{fig:01}\,C shows the corresponding fragment of the regularized estimate.
+The regularized estimate is produced from the same data by zeroing the optimal set of 31501 (71.2\%) of off-diagonal coefficients of the precision matrix and fitting the remaining coefficients.
+Panels \ref{fig:01}\,D and E show the same fragments of the respective estimates of the partial correlation matrix.
+The apparent similarity of the two estimates of the correlation matrix belies the utterly different partial correlation structure.
+The regularized estimate is less sensitive to sample noise in the data, is likely to be closer to truth in most cases than the unbiased estimate, which may be ascertained by cross-validation studies.
+
+It may be tempting to interpret the regularized estimates of partial correlations as a more accurate representation of the functional connectivity than the original correlation coefficients: after all, the regularized solution consistently produces better estimates and has fewer parameters which are easier to interpret.
+However, an alternative regularization scheme using 43 latent factors interacting linearly with all neurons while having about the same number of parameters, perform even better than covariance selection on this dataset.
+
+It is therefore important to compare multiple families of regularized estimators on the system under investigation.
+The family that consistently outperforms the others may be proposed as candidate low-dimensional representation of the correlation structure.
+
+In this study, we compared four families of regularized combining partial interactions and latent units: `shrinkage toward diagonal', `shrinkage toward a multifactor model', `sparse partial correlations', and 'sparse+latent partial correlations'.
+First, in simulation with known low-dimensional structure of correlations, we demonstrate that regularized estimators from matching families of low-dimensional targets generally outperform all other estimators.
+We then repeat this study on neural data from dense groups of neurons in mouse visual cortex under visual stimulation.
+We find that optimal estimation was achieved by modeling correlations with a combination of sparse partial correlations and interactions with several latent units.
+Finally, we analyze the relationship between the elements of these regularized estimates and other properties of the circuit such as physical distances between neurons and their preferred orientations.
+
+\section*{Results}
 % Results and Discussion can be combined.
-\section*{Results}
 
@@
 \end{equation}
-The diagonal target estimate expresses the idea of a lack of dependence (or of linear association) between the activity of observed neurons (Fig.~\ref{fig:02}\;A).  
+The diagonal target estimate expresses the idea of a lack of dependence (or of linear association) between the activity of observed neurons (Fig.~\ref{fig:02}\,A).  
 If this assumption aptly describes recorded data, then strong shrinkage toward $D$ will add little bias while strongly reducing the variability of the estimate.
@@
 This estimator has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $F$ expresses the assumption that correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}\;B).
+The target estimate $F$ expresses the assumption that correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}\,B).
 
-The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations are the result of direct linear interactions between pairs of observed cells and that such interactions occur only between a fraction of such pairs (Fig.~\ref{fig:02}\;C).
+The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations are the result of direct linear interactions between pairs of observed cells and that such interactions occur only between a fraction of such pairs (Fig.~\ref{fig:02}\,C).
 This assumption is enforced by reducing to zero the majority of pairwise \emph{partial correlations} in the recorded population. 
@@
 
-Finally, we consider the fourth estimator $C_{\sf sparse+lowrank}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\;D).
+Finally, we consider the fourth estimator $C_{\sf sparse+lowrank}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D).
 This estimator has the form
@@
 
-As expected, all regularized estimators outperformed the sample covariance estimator substantially (Fig.~\ref{fig:04}) \Acomment{I think we should show that}.
-Say something about how shrinkage, low-rank and sparse inverse relate \Acomment{Why do we do pairwise comparisons instead of just showing median (across sites) log loss relative to the combined estimator?
-I've seen many people do this and it provides an ordering.
-I feel like we should come up with a way of ordering them somehow, even if it's under certain assumptions that may not be entirely correct.
-I'm pretty sure if we don't do it the reviewers will bring it up anyway...}.
+As expected, all regularized estimators outperformed the sample covariance estimator substantially (Fig.~\ref{fig:04}) Say something about how shrinkage, low-rank and sparse inverse relate.
 Finally, the combined sparse and low-rank estimator dominated all others significantly (Fig.~4), showing that both hidden units and direct interactions are important in our data.
@@
 
-In practice, the true value $\Sigma$ is not accessible and an estimators' risks must be estimated from the data.
-This may be accomplished through \emph{validation}.
-\Kcomment{If you will discuss validation later, point to it.}
+In practice, the true value $\Sigma$ is not accessible and an estimators' risks must be estimated from the data.  This may be accomplished through \emph{validation}.  
 Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$.
 
-$\loss{\cdot,\cdot}$ is additive in its second argument so that \Kcomment{You frequently use ``such that'' instead of ``so that''}
+$\loss{\cdot,\cdot}$ is additive in its second argument so that 
  \begin{equation}\label{eq:additivity}
@@
  \end{equation}
-Then \emph{validation loss}  \TODO{check appropriate term}
+Then \emph{validation loss}  
 \begin{equation}\label{eq:validationLoss}
@@
 Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
-\Kcomment{This is probably not a very insightful question -- but why do we assume that the sample covariance
-is what we compare $\hat\Sigma$ when we validate the results.  Our point is that this estimate is not very good.}
 
@@
 A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimate to minimize the estimator's risk.  
-Various regularization schemes focus on the dimensionality reduction part \TODO{rephrase} by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.
+Various regularization schemes focus on the dimensionality reduction part  by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.  
 Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}. 
@@
 Gaussian graphical models have a straightforward relationship to their covariance matrix $\Sigma$:  Zero elements in the inverse of $\Sigma$ indicate conditional independence between the corresponding pair of variables.
-\Kcomment{It may be important to note that we are not fitting Gaussian graphical models, however.}
 
@@
 We estimated the optimal values of hyperparameters $ \alpha$ and $ \lambda$  by  cross-validation within the training dataset.
-\TODO{more detail?} \TODO{discuss analytic optimization of shrinkage intensities.}
-
-\Kcomment{Yes -- you give no details about the cross-validation.  You should provide some details at least in 
-one case, or have a separate section that discusses how you implemented it.}
 
@@
 where $\hat L_d$ is a $p\times p$ matrix of rank $d$ and $\hat \Psi$ is a $p\times p$ diagonal matrix. The value of $\hat T_d$ is found by minimizing $\loss{\hat T_d, \hat\Sigma_0}$.
-\Kcomment{I would give a more explicit representation of $\hat L_d$ as a product to show that the factors can be understood as common input.}
 
@@
 Thus the inverse covariance matrix $S=-\Sigma^{-1}$ is also the adjacency matrix of the GGM, indicating which pairs are conditionally dependent.
-\TODO{discuss how quickly (or slowly) this breaks down with deviations from gaussianity, e.g.~MaxEnt models, Ising models.} \TODO{Also, motivate this from the perpsective of linear regression as in \cite{Varoquaux:2012}.}  \Kcomment{This is a bit confusing -- in the first sentence, you have statistical independence, in the second, you talk about conditional independence.
-You want to emphasize that a 0 inverse means that -- in the GGM -- the two nodes only ``talk'' indirectly through other nodes.} The off-diagonal elements of $S$ are proportional to partial pairwise correlations between corresponding pairs of neurons.
-\TODO{Expand.} Regularization of covariance matrix estimation by sparsifying the inverse of the estimate has been using in functional genomics \cite{Schafer:2005,Peng:2009} and fMRI functional connectivity \cite{Varoquaux:2012,Ryali:2012}.
+The off-diagonal elements of $S$ are proportional to partial pairwise correlations between corresponding pairs of neurons.
 
@@
 \end{equation}
-where $\lambda \|S\|_1$ is the $L_1$ norm of the matrix. \TODO{clarify}. \Kcomment{Yes - at least define the norm, if you are already taking about it. No need to expand a lot more, though.}
+where $\lambda \|S\|_1$ is the $L_1$ norm of the matrix.
 
 Just as in the previous estimators, the optimal value of the regularization parameter $\lambda$ is found from training data by cross-validation.
-\Kcomment{I would say this up front, so you don't have to repeat it.
-Again, I think more details about cross-validation should be given.}
 
@@
     {\bf A--C.} Average sample correlations (black) and regularized partial correlations (red) between pairs of cells in the example site shown in previous figures.
-The correlations were normalized by the respective average correlations (\ref{fig:06}\;A).
+The correlations were normalized by the respective average correlations (\ref{fig:06}\,A).
     {\bf A.} Average correlations between pairs of neurons tuned to orientation with differences in preferred orientation in the intervals of 0--15$^\circ$, 15--45$^\circ$ and 45--90$^\circ$. 

### 2013-12-02 plos.tex
updated the Simulation section.

@@
 \newcommand{\loss}[1]{\mathcal L\left(#1\right)} 
+\newcommand{\eloss}[1]{\mathcal L_0\left(#1\right)}
 \newcommand{\T}{{\sf T}}
@@
 To estimate the mean activity, we resort to the usual \emph{sample mean} without attempting to improve it.
-An estimate of the \emph{covariance matrix}  combined estimate of the variances and correlations constitutes an estimate covariance matrix.
-The problem can then be formulated as finding the optimal estimation of the covariance matrix given the estimate of the mean firing rates.
+Estimation of a \emph{covariance matrix} is equivalent to estimating both  variances and correlations.
+Therefore, we rephrase the problem as one of optimal estimation of the covariance matrix given an estimate of the mean firing rates.
 
-Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward one of several possible low-dimensional approximations called here \emph{target estimates} \cite{Schafer:2005,Bickel:2006}.
-The usual estimate, the \emph{sample covariance matrix} has the advantage of being unbiased but, on average, it falls far from the true covariance matrix because of its sensitivity to sampling noise in the data.
-Low-dimensional estimates of various forms are less susceptible to sampling noise but are also liable to introduce their respective biases away from the true covariance matrix.
+Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward one of several possible low-dimensional approximations referred to here as \emph{target estimates} \cite{Schafer:2005,Bickel:2006}.
+The usual estimate, the \emph{sample covariance matrix} has the advantage of being unbiased but, on average, it falls far from the true covariance matrix due to its sensitivity to sampling noise in the data.
+Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from the true covariance matrix.
 Regularization works by striking a favorable balance between bias and variability.
@@
 
-To illustrate key concepts, Figure \ref{fig:01} illustrates a regularization scheme based on \emph{covariance selection} \cite{Dempster:1972}}.
-In covariance selection, the estimate is produced by fitting only an optimally selected subset of the coefficients of the \emph{precision matrix} (the inverse of the covariance matrix) while setting the remaining coefficients to zero.
-Scaling the precision matrix so that its diagonal is composed of all $-1$ produces the matrix of partial correlations.
+To illustrate key concepts, Figure \ref{fig:01} illustrates a regularization scheme based on \emph{covariance selection} \cite{Dempster:1972}.
+In covariance selection, the estimate is produced by fitting only an optimal subset of the coefficients of the \emph{precision matrix} (the inverse of the covariance matrix) while setting the remaining coefficients to zero.
+Scaling the precision matrix so that its diagonal is composed of all $-1$'s produces the matrix of partial correlations.
 Thus zeros in the precision matrix produce zero partial correlations between the corresponding pairs of neurons.
@@
 
-It may be tempting to interpret the regularized estimates of partial correlations as a more accurate representation of the functional connectivity than the original correlation coefficients: after all, the regularized solution consistently produces better estimates and has fewer parameters which are easier to interpret.
-However, an alternative regularization scheme using 43 latent factors interacting linearly with all neurons while having about the same number of parameters, perform even better than covariance selection on this dataset.
+It may be tempting to interpret the regularized sparse estimate of partial correlations as a more accurate representation of the functional connectivity than the original correlation coefficients: after all, the regularized solution consistently produces better estimates and has fewer parameters which are easier to interpret.
+However, an alternative regularization scheme using 43 latent factors interacting linearly with all neurons while having about the same number of parameters, performed even better than covariance selection on this dataset.
 
@@
 The family that consistently outperforms the others may be proposed as candidate low-dimensional representation of the correlation structure.
+The greater efficiency of one estimator over another is not a mathematical necessity but an empirical finding depending on the prevalent types of correlation structure in the investigated system.
 
@@
 We then repeat this study on neural data from dense groups of neurons in mouse visual cortex under visual stimulation.
-We find that optimal estimation was achieved by modeling correlations with a combination of sparse partial correlations and interactions with several latent units.
-Finally, we analyze the relationship between the elements of these regularized estimates and other properties of the circuit such as physical distances between neurons and their preferred orientations.
+We find that most efficient estimation was achieved by modeling correlations as a sparse network of partial correlations combined with several latent units interacting with the entire population.
+Finally, we analyze the relationship between the elements of such estimates and other properties of the circuit such as physical distances between neurons and their preferred orientations.
 
@@
 
-Regularization is the deliberate biasing of the estimate toward a low-dimensional, less variable \emph{target estimate} to strike a favorable balance between bias and variability of the estimate \cite{Bickel:2006,Ledoit:2004}.
-
-The estimator whose target estimate most closely matches the true low-dimensional structure of the data is likely to outperform other estimators. This principle provides the logic of this study.
-
 We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (Fig.~\ref{fig:02}\;row 1).
@@
 This estimator has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $F$ expresses the assumption that correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}\,B).
+The target estimate $F$ expresses the assumption that correlated fluctuations in the population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}\,B).
 
@@
 \end{equation}
-where $\hat S$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements. The estimate has one hyperparameter to regulate the sparsity (fraction of off-diagonal zeros) in $S$.
+where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements. The estimate has one hyperparameter to regulate the sparsity (fraction of off-diagonal zeros) in $S$.
 
@@
 \subsection*{Simulation}
-To verify our approach and to illustrate the performance of the four regularized estimators, we constructed five model populations with different covariance structures.
-Each population contained 50~neurons.
-The first four populations matched the low-dimensional structure of our four estimators: independent (Fig.~\ref{fig:03}A), latent factors (Fig.~\ref{fig:03}B), sparse partial correlations (Fig.~\ref{fig:03}C) and low-rank combined with sparse inverse (Fig.~\ref{fig:03}D).
-In addition, we considered a fifth population that had no low-dimensional structure (Fig.~\ref{fig:03}\,E).
+To verify our approach and to illustrate the performance of the four regularized estimators, we constructed four model populations of size $p=50$~neurons each with correlation structures of the same type as the target estimates of the first (Fig.~\ref{fig:02} Row 2).
+Panel A2 contains a diagonal correlation matrix matching the target of $C_{\sf diag}$.
+Panel B2 is a factor model with 3 factors conforming to the family of target estimates of $C_{\sf factor}$.
+Panel B3 is a correlation matrix with sparse partial correlations (76\% off diagonal zeros in the precision matrix).
+Finally, Panel B4 is a correlation matrix whose inverse is composed the sum of a sparse matrix (82\% sparse) and a low-rank component (rank=1).
+Row 3 contains sample correlation matrices calculated from samples of size $n=1000$ from a multivariate normal distributions with the respective correlation matrices from Row 2.
+
+To evaluate the performance of a covariance matrix $C$, we define a real-valued \emph{loss function} $\loss{C,\Sigma}$ such that it attains its minimum when $C=\Sigma$.
+The loss function quantifies the error of the estimate, \emph{i.e.}~the deviation of the estimate $C$ from truth $\Sigma$.
+
+For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss}\footnote{
+The loss function $\loss(C,\Sigma)$ is related to the multivariate log-likelihood function $L(\Sigma|C_0)$ as 
+\begin{equation}
+    L(\Sigma|C_0) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_0}
+\end{equation}
+}
+defined as:
+\begin{equation}\label{eq:loss}
+    \loss{C,\Sigma} = \frac 1 p\left[\ln \det C + \Tr(C^{-1}\Sigma)\right]
+\end{equation}
+
+We normalized the loss function  by $\frac 1 p$ to make its value less sensitive population size (For independent neurons, the log likelihood scales linearly with $p$).
+
+This choice is motivated by mathematical convenience.
+Other popular choices for the loss function are the Frobenius norm of the difference $C-\Sigma$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, and quadratic loss \cite{James:1961,Fan:2008}.
+We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.
+
+To make a loss function that assume zero at its minimum when $C=\Sigma$, we define \emph{excess loss} as
+\begin{equation}\label{eq:excess-loss}
+    \eloss{C,\Sigma} = \loss{C,\Sigma}-\loss{\Sigma,\Sigma}
+\end{equation}
+
+In Figure \ref{fig:02}, row 4 shows the mean excess losses and their standard errors calculated from 30 samples of sizes n=250, 500, 1000, 2000, and 4000 for each estimator and each kind of ground truth.
+For each estimator, its hyperparameters were estimated by nested cross-validation (See Methods for more details.)
+
+As expected, estimators with matching low-dimensional structures typically outperformed the other estimators.
+There were two exceptions to this observation.
+For small sample sizes, before the data were sufficient to reveal the true correlation structure and to  allow the correct model to dominate, estimates with simpler targets often outperformed the more matching estimate.
+
+With empirical data, the truth $\Sigma$ is not accessible and the relative values of loss must be estimated solely from the data.
+This may be accomplished through \emph{validation}.
+In validation, the data are split into a training subset from which the estimate is obtained and a validation subset.
+We then compute the \emph{validation loss}
+
+\begin{equation}\label{eq:validation-loss}
+    \loss{C,C^\prime_0}
+\end{equation}
+
+$\loss{\cdot,\cdot}$ is additive in its second argument so that 
+ \begin{equation}\label{eq:additivity}
+ \loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
+ \end{equation}
+Then \emph{validation loss}  
+\begin{equation}\label{eq:validationLoss}
+\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
+\end{equation}
+is an unbiased estimate of the risk:
+ \begin{equation}\label{eq:empiricalRisk}
+\E[(\hat\Sigma, \hat\Sigma_0^\prime)] {\hat\ell} 
+= \E[(\hat\Sigma, \hat\Sigma_0^\prime)]{\loss{\hat\Sigma,\hat\Sigma_0^\prime}}
+= \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}
+= \E{\loss{\hat\Sigma,\Sigma}} = r
+ \end{equation}
+Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
+
+Other popular loss functions such as entropy loss do not comply with Eq.~\ref{eq:additivity} and their validation loss is not an unbiased estimate of risk.
+
+Since $\loss{\cdot,\cdot}$ is equivalent to negative normal log likelihood, the above derivation has led to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes cross-validated normal log likelihood.
 
@@
 
-\begin{itemize}
-\item Linear and partial correlations versus spatial separation (lateral and vertical). Discuss whether to include or remove thresholded correlations.
-\item Magnitude of common input versus spatial location in the volume (lateral and vertical separation from center).
-\item Linear and partial correlations versus orientation preference
-\item Distribution of sparsity over sites
-\item Distribution of number of hidden units over sites
-\item Distribution of eigenvalues for sample covariance versus combined estimator
-\item etc. etc.
-
-\end{itemize}
-
-\section*{Discussion}
-
-% You may title this section "Methods" or "Models". 
-% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
-% authors may use "Analysis" 
-Fundamental questions in systems neuroscience concern the relationship between the function of neural microcircuits and their cytoarchitecture: circuit topology, cell types, and patterns of synaptic connectivity.
-
-Considerable progress has been made in sensory areas where the functional characterizations of cells are defined by their reproducible responses to external stimuli (reviewed in \cite{Reid:2012}).
-
-Many key questions in neuroscience revolve around the relationship between the cytoarchitecture of neural microcircuits and the functional organization of their activity \cite{Reid:2012}.
-Investigators in this line of inquiry aim to construct networks of functional associations within groups of neurons inferred from observations of their activity under a variety of conditions.
-The inferred functional network  could then related to the synaptic connectivity patterns, cell types, cell properties, or the cells' spatial arrangment in the attempt to uncover organizational principles of neural computation.
-In addition, the network itself can serve as subtrate for subsequent graph-theoretical analysis \cite{Feldt:2011}
-
-Advances in electrophysiology and optical imaging have enabled simultaneous recordings from dozens to hundreds or thousands of cells. 
-In particular, recent advances in two-photon imaging of calcium signals have allowed in vivo recordings from nearly every neuron brain-wide in zebrafish \cite{Leung:2013,Ahrens:2013} and in a 3D volume of a few hundred microns in diameter in mouse neocortex \cite{Katona:2012,Cotton:2013}.
-Ambitious projects currently under way aim to record the spiking activity of large fractions of cells from entire circuits and systems in behaving animals \cite{Alivisatos:2012}.
-Direct observations of the population activity  of entire circuits open new possibilities for incisive statistical descriptions of the functional connectivity in the circuit.
-
-One general approach is to construct statistical models reflecting hypothesized models of neural circuit function, including temporal dynamics, nonlinear interactions between neurons, and background network activity \cite{Pillow:2008,Buonomano:2009,Yu:2009}.
-Another general approach is to search for alternative statistics that best describe the population activity without assuming a specific model.
-Such evaluations rely on constructing surrogate datasets \cite{Okun:2012} or maximum-entropy distributions that reproduce the statistic of interest \cite{Ganmor:2011,Tkacik:2012} with subsequent testing of how well such constructs can reproduce other observed properties of population activity.
-
-Such more sophisticated statistical descriptions are unlikely to entirely replace neural correlations as the initial descriptive statistic of population activity in experimental neuroscience.
-Neural correlations have several strengths that will ensure their continued prominence: (a) correlations are well established, familiar, and intuitive to most researchers in biological sciences, (b) correlations are easy to estimate as they are less susceptible to the curse of dimensionality than more sophisticated models, (c) correlations make relatively weak mechanistic assumptions (at the cost of not revealing many potentially important aspects), and (d) correlations serve as input into other models of population activity or decoding algorithms.
-
-Linear correlations between pairs of neurons are among the most common and familiar descriptions of functional associations in networks and of the collective activity of neuronal populations.
-For example, it is tempting to consider the network constructed from the highest correlations in the recorded population (e.g.~\cite{Malmersjo:2013}) for subsequent graph-theoretical analysis \cite{Feldt:2011}.
-In functional genomics, networks constructed from highest correlations are called \emph{relevance networks}.
-
-However, pairwise correlations unreliable proxies of direct functional association such as direct synaptic connectivity as they can arise due to a wide variety of physiological interactions including direct synaptic interactions, indirect chains of synaptic connections, common inputs, correlations in inputs, synchrony, fluctuations in global network states, and others \cite{Shadlen:1998,Ostojic:2009,Pernice:2011,Schneidman:2006}.
-
-Other experimental evidence suggests that detailed interactions in local microcircuits have only weak effects on overall population activity, which is then best characterized by collective features such as global population dynamics or population synchrony \cite{Okun:2012,Tkacik:2012,Tkacik:2013}.
-From this point of view, the imporant aspects of the correlation structure lie it its global aspects, such as the eigenspectrum, while the individual pairwise correlations are not particularly meaningful.
-
 \section*{Methods}
-\subsection*{Evaluation of covariance matrix estimators}
-The quality of a covariance matrix estimate $\hat\Sigma$ is measured by a real-valued \emph{loss function} $\loss{\hat\Sigma,\Sigma}$.  
-The loss function quantifies the error of the estimate, \emph{i.e.}~the deviation of $\hat\Sigma$ from $\Sigma$. The loss function must attain its minimum  when $\hat\Sigma = \Sigma$.
-
-For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss} function:
-\begin{equation}\label{eq:loss}
-\loss{\hat\Sigma,\Sigma} = \frac 1 p\left[\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma)\right]
-\end{equation}
-This choice is motivated by mathematical convenience.
-Other popular choices for the loss function are the Frobenius norm of the difference $\hat\Sigma-\Sigma$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, and quadratic loss \cite{James:1961,Fan:2008}.
-We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.
-
-The aim of our project is to identify which covariance matrix estimator minimizes the expected loss 
-\begin{equation}
-r = \E{\loss{\hat\Sigma, \Sigma}}
-\end{equation}
-which is known as the \emph{risk} of $\hat\Sigma$.
-
-In practice, the true value $\Sigma$ is not accessible and an estimators' risks must be estimated from the data.  This may be accomplished through \emph{validation}.  
-Let $\hat\Sigma_0^\prime$ denote a sample covariance matrix measured from an independent sample that was not used included in the computation of $\hat\Sigma$.
-
-$\loss{\cdot,\cdot}$ is additive in its second argument so that 
- \begin{equation}\label{eq:additivity}
- \loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
- \end{equation}
-Then \emph{validation loss}  
-\begin{equation}\label{eq:validationLoss}
-\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
-\end{equation}
-is an unbiased estimate of the risk:
- \begin{equation}\label{eq:empiricalRisk}
-\E[(\hat\Sigma, \hat\Sigma_0^\prime)] {\hat\ell} 
-= \E[(\hat\Sigma, \hat\Sigma_0^\prime)]{\loss{\hat\Sigma,\hat\Sigma_0^\prime}}
-= \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}
-= \E{\loss{\hat\Sigma,\Sigma}} = r
- \end{equation}
-Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
-
-Other popular loss functions such as entropy loss do not comply with Eq.~\ref{eq:additivity} and their validation loss is not an unbiased estimate of risk.
-
-Since $\loss{\cdot,\cdot}$ is equivalent to negative normal log likelihood, the above derivation has led to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes cross-validated normal log likelihood.
+\subsection*{Estimation of parameters and hyperparameters of covariance estimators}
 
@@
 
-\subsection*{Estimators}
-We evaluated four regularized estimators denoted here as A, B, C, and D.
-Their target estimates correspond to covariance matrices of Gaussian graphical models depicted in Figure 2 where green spheres represent the recorded neurons,  light-colored spheres represent latent units of the graphical model, and edges connecting them represent conditional dependencies or interactions.
-
-A \emph{graphical model} is a multivarate probability distribution with a specified graph of conditional dependencies between its variables \cite{Koller:2009}.
-When this distribution is a multivariate normal distribution, the model becomes a \emph{Gaussian graphical model} (GGM) or, equivalently, a \emph{Gaussian Markov Random Field}.
-Gaussian graphical models have a straightforward relationship to their covariance matrix $\Sigma$:  Zero elements in the inverse of $\Sigma$ indicate conditional independence between the corresponding pair of variables.
-
-Fitting a graphical model to data involves two tasks: (1) the selection of the set of non-zero covariances or \emph{covariance selection} \cite{Dempster:1972} and (b) the fitting of the non-zero elements.
+\subsection*{}
 
@@
         {\bf Row 3.} Examples of sample correlation matrices calculated from samples of 1000 observations taken from simulated random processes with corresponding correlation matrices from row 2.
-        {\bf Row 4.} Average normalized Gaussian log-likelihood \emph{excess loss} (\ref{eq:excess-loss}) attained by each of the five estimators as a function of sample size.
+        {\bf Row 4.} Average normalized Gaussian log-likelihood \emph{excess loss} (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators as a function of sample size.
 The error bars indicate the standard error of the mean based on 30 samples.
-        {\bf Row 5.} Average normalized Gaussian log-likelihood \emph{validation loss} (\ref{eq:validation-loss}) attained by each of the five estimators.
+        {\bf Row 5.} Average normalized Gaussian log-likelihood \emph{validation loss} (Eq.~\ref{eq:validation-loss}) attained by each of the five estimators.
 The values are relative to the validation loss of the estimator that matches the low-dimensional structure of the true covariance matrix.

### 2013-12-02 plos.tex
updated the Simulation section

@@
 Panel \ref{fig:01}\,B magnifies the fragment of the sample correlation matrix outlined in \ref{fig:01}\,A whereas \ref{fig:01}\,C shows the corresponding fragment of the regularized estimate.
-The regularized estimate is produced from the same data by zeroing the optimal set of 31501 (71.2\%) of off-diagonal coefficients of the precision matrix and fitting the remaining coefficients.
+The regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) off-diagonal coefficients of the precision matrix and fitting the remaining coefficients.
 Panels \ref{fig:01}\,D and E show the same fragments of the respective estimates of the partial correlation matrix.
@@
 
-For the purposes of this study, we adopted the \emph{negative normal log-likelihood loss}\footnote{
-The loss function $\loss(C,\Sigma)$ is related to the multivariate log-likelihood function $L(\Sigma|C_0)$ as 
-\begin{equation}
+In this study, we adopted \emph{negative normal log-likelihood loss}\footnote{
+The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_0)$: 
+\begin{equation*}
     L(\Sigma|C_0) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_0}
-\end{equation}
+\end{equation*}
 }
-defined as:
+defined as
 \begin{equation}\label{eq:loss}
@@
 
-We normalized the loss function  by $\frac 1 p$ to make its value less sensitive population size (For independent neurons, the log likelihood scales linearly with $p$).
+We normalized the loss function  by $\frac 1 p$ to make its value comparable across different population sizes.
 
 This choice is motivated by mathematical convenience.
-Other popular choices for the loss function are the Frobenius norm of the difference $C-\Sigma$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, and quadratic loss \cite{James:1961,Fan:2008}.
+Other popular choices for the loss function are the Frobenius norm of the difference $\|C-\Sigma\|_F$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, and quadratic loss \cite{James:1961,Fan:2008}.
 We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.
 
-To make a loss function that assume zero at its minimum when $C=\Sigma$, we define \emph{excess loss} as
+To make the loss function to assume zero at its minimum when $C=\Sigma$, we define \emph{excess loss} as
 \begin{equation}\label{eq:excess-loss}
@@
 There were two exceptions to this observation.
-For small sample sizes, before the data were sufficient to reveal the true correlation structure and to  allow the correct model to dominate, estimates with simpler targets often outperformed the more matching estimate.
+For small sample sizes, before the data were sufficient to reveal the true correlation structure and to  allow the correct model to dominate, estimates with simpler targets often outperformed the matching estimator.
 
-With empirical data, the truth $\Sigma$ is not accessible and the relative values of loss must be estimated solely from the data.
-This may be accomplished through \emph{validation}.
-In validation, the data are split into a training subset from which the estimate is obtained and a validation subset.
-We then compute the \emph{validation loss}
+When the ground truth $\Sigma$ is not accessible, loss can be estimated solely from the data through \emph{validation}.
+In validation, an additional independent \emph{testing dataset} is used to compute an independent sample covariance estimate $C_0^\prime$.
+Then \emph{validation loss} $\loss{C,C_0^\prime}$ can be used as an unbiased estimate\footnote
+{
+    Validation loss $\loss{C,C_0^\prime}$ is an unbiased estimate of loss $\loss{C,\Sigma}$ when $\loss{\cdot,\cdot}$ is \emph{additive} in its second argument so that 
+ \begin{equation*}\label{eq:additivity}
+ \loss{C,z_1} + \loss{C,z_2} \equiv \loss{C,z_1+z_2}
+ \end{equation*}
 
-\begin{equation}\label{eq:validation-loss}
-    \loss{C,C^\prime_0}
-\end{equation}
+Then the absence of bias is shown:
+\begin{equation*}
+    \E{\loss{C,C_0^\prime}}=\loss{C,\E{C_0^\prime}}=\loss{C,\Sigma}
+\end{equation*}
+} 
+of loss $\loss{C,\Sigma}$.
 
-$\loss{\cdot,\cdot}$ is additive in its second argument so that 
- \begin{equation}\label{eq:additivity}
- \loss{\hat\Sigma,X_1} + \loss{\hat\Sigma,X_2} \equiv \loss{\hat\Sigma,X_1+X_2}
- \end{equation}
-Then \emph{validation loss}  
-\begin{equation}\label{eq:validationLoss}
-\hat \ell = \loss{\hat\Sigma,\hat\Sigma_0^\prime}
-\end{equation}
-is an unbiased estimate of the risk:
- \begin{equation}\label{eq:empiricalRisk}
-\E[(\hat\Sigma, \hat\Sigma_0^\prime)] {\hat\ell} 
-= \E[(\hat\Sigma, \hat\Sigma_0^\prime)]{\loss{\hat\Sigma,\hat\Sigma_0^\prime}}
-= \E{\loss{\hat\Sigma,\E{\hat\Sigma_0^\prime}}}
-= \E{\loss{\hat\Sigma,\Sigma}} = r
- \end{equation}
 Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
 
-Other popular loss functions such as entropy loss do not comply with Eq.~\ref{eq:additivity} and their validation loss is not an unbiased estimate of risk.
+\noindent\rule{8cm}{4pt}
 

### 2013-12-02 plos.tex
updated the Results section.

@@
 
-The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_0$ computed from the empirical sample of observations $x(1),\ldots,x(n)$ as
+The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(1),\ldots,x(n)$ as
 \begin{equation}
-C_0 = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T, 
+C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T, 
 \end{equation}
 where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent).  
-The sample covariance matrix is unbiased, \emph{i.e.} $\E{C_0}=\Sigma$.
-For finite sample sizes, however, $C_0$ is not as close to $\Sigma$ as a number of biased estimators.
+The sample covariance matrix is unbiased, \emph{i.e.} $\E{C_{\sf 0}}=\Sigma$.
+For finite sample sizes, however, $C_{\sf 0}$ is not as close to $\Sigma$ as a number of biased estimators.
 
@@
 In the first regularized estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ containing on its diagonal estimates of the variances.
-The regularized estimate is obtained by linear \emph{shrinkage} of the unbiased estimate $C_0$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
+The regularized estimate is obtained by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
 \begin{equation}
-C_{\sf diag} = (1-\lambda) C_0 + \lambda D
+C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
 \end{equation}
@@
 \begin{equation}
-C_{\sf factor} = (1-\lambda) C_0 + \lambda F
+C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda F
 \end{equation}
@@
 In this study, we adopted \emph{negative normal log-likelihood loss}\footnote{
-The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_0)$: 
+The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$: 
 \begin{equation*}
-    L(\Sigma|C_0) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_0}
+    L(\Sigma|C_{\sf 0}) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_{\sf 0}}
 \end{equation*}
@@
 
-When the ground truth $\Sigma$ is not accessible, loss can be estimated solely from the data through \emph{validation}.
-In validation, an additional independent \emph{testing dataset} is used to compute an independent sample covariance estimate $C_0^\prime$.
-Then \emph{validation loss} $\loss{C,C_0^\prime}$ can be used as an unbiased estimate\footnote
+When ground truth $\Sigma$ is not accessible, loss can be estimated solely from the data through \emph{validation}.
+In validation, an additional independent \emph{testing sample} is used to compute an independent sample covariance estimate $C_{\sf 0}^\prime$.
+Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can be used as an unbiased estimate\footnote
 {
-    Validation loss $\loss{C,C_0^\prime}$ is an unbiased estimate of loss $\loss{C,\Sigma}$ when $\loss{\cdot,\cdot}$ is \emph{additive} in its second argument so that 
+    Validation loss $\loss{C,C_{\sf 0}^\prime}$ is an unbiased estimate of loss $\loss{C,\Sigma}$ when $\loss{\cdot,\cdot}$ is \emph{additive} in its second argument so that 
  \begin{equation*}\label{eq:additivity}
@@
 
-Then the absence of bias is shown:
+Then the absence of bias can be shown:
 \begin{equation*}
-    \E{\loss{C,C_0^\prime}}=\loss{C,\E{C_0^\prime}}=\loss{C,\Sigma}
+    \E{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E{C_{\sf 0}^\prime}}=\loss{C,\Sigma}
 \end{equation*}
+
+This property does not hold for some other popular losses such as Stein's entropy loss, for example, which prevents their substitution with corresponding validation losses.
+However, various other losses do conform to this constraints and could be used in this study.
 } 
-of loss $\loss{C,\Sigma}$.
+of loss $\loss{C,\Sigma}$.  Thus, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
 
-Therefore, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
+With empirical data, acquiring a separate testing sample is not practical.
+Instead, $K$-fold cross-validation is used.
+In cross-validation, the sample is divided into $K$ subsets of approximately equal size.
+In all computations in this paper $K=10$ was used.
+Then each  subset is used as the validation sample with the other $K-1$ serving as the training dataset.
+The validation losses from each of such `folds' are averaged to produce \emph{cross-validation loss} or \emph{CV-loss} for short.
+Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of  estimator $C$ trained on the remaining $K-1$ subsets.
+Then cross-validation loss for estimator $C$ is
+\begin{equation}\label{eq:cv-loss}
+    \ell_C=\frac 1 K \sum\limits_{k=1}^K \loss{C^{\{\setminus k\}},C_{\sf 0}^{\{k\}}}
+\end{equation}
+In the present formulation, $\ell_C$ is known as the cross-validated Gaussian log-likelihood (up to a constant offset and multiplier).
 
-\noindent\rule{8cm}{4pt}
+Since we only need to compare estimators against each other, we are only interested in \emph{relative} CV-loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$:
+\begin{equation}\label{eq:rel-cv-loss}
+    \ell_{C,C_{\sf ref}} = \frac 1 K \sum\limits_{k=1}^K \left[
+        \loss{C^{\{\setminus k\}},C_{\sf 0}^{\{k\}}} -
+    \loss{C_{\sf ref}^{\{\setminus k\}},C_{\sf 0}^{\{k\}}} 
+\right]
+\end{equation}
 
-Since $\loss{\cdot,\cdot}$ is equivalent to negative normal log likelihood, the above derivation has led to the familiar criterion that the optimal covariance matrix estimator is one that consistently maximizes cross-validated normal log likelihood.
+In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (Figure \ref{fig:02}, Row 5).
+For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth: $\ell_{C,C_{\sf diag}}$, $\ell_{C,C_{\sf factor}}$, $\ell_{C,C_{\sf sparse}}$, and $\ell_{C,C_{\sf sparse+lowrank}}$.
+Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples  taken for each ground truth and for each sample size.
 
-To evaluate the performance of the different estimators, we computed the excess loss for all combinations of model populations and estimators, including the sample covariance.
-The first striking observation is that in all cases all four regularized estimators performed substantially better than the sample covariance (Fig.~2, fourth row).
-In particular, this is even true for the case where the population did not have any low-dimensional structure at all (Fig.~2\,E).
-While it may appear surprising at first sight, this counter-intuitive phenomenon is known as \emph{Stein's phenomenon} or \emph{Stein's paradox} \cite{Efron:1977}, named after its discoverer Charles Stein \cite{Stein:1956}.
-A common misconception about regularization is that its effect depends on accurate prior knowledge about the structure of the data.
-However, substantial improvement can be attained by shrinking the unbiased estimate toward an arbitrary target as long as the target is less variable than the unbiased estimator.
-The more accurate description of regularization is as of the optimal tradeoff between estimation and approximation error -- the so-called ``bias-variance tradeoff''.
-
-Thus, if taken in isolation a regularized estimator improves the estimate we should not interpret this result to suggest that the estimator's target has the same low-dimensional structure as the data-generating process.
-However, when comparing multiple estimators against each other, the one whose target estimate most closely matches the true value with the smallest number of parameters will reduce the estimation error with the least increase in approximation error.
-Indeed, in all four toy examples with a low-dimensional structure, the estimator with the matching regularization target performed best (Fig.~\ref{fig:03}\,A--D, fourth row).
-Since the estimator combining low-rank and sparse interactions combines to low-dimensional targets and includes the simpler estimators as special cases, it performed almost as well even when the low-dimensional structure did not include either a low-rank component or sparse interactions (Fig.~\ref{fig:03}\,B,\,C).
-
-Since the above evaluations of excess loss require knowledge of ground truth, this analysis can be done only on simulated data.
-However, \emph{validation loss} can be used as an unbiased empirical estimate of loss  (see Methods, Eq.~\ref{eq:validation-loss}.
-Validation loss uses an independent validation sample to estimate the value of loss without access to ground truth.
-Indeed, this analysis revealed a pattern of results that reproduced the results obtained with access to ground truth above (Fig.~\ref{fig:03}, last row).
-Because validation loss is computed by comparing estimates to noisy sample covariance matrices from smaller validation sets, validation loss has higher variance than excess loss.
-Under our chosen loss function, validation loss does not converge to zero.
+These simulation results demonstrated that, with sufficiently large sample sizes, the selection of the most efficient of several regularized estimators could be used to infer the type of low-dimensional structure present in the data, at least when one of the regularized estimators could capture such structure.
 
@@
 
-We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}.
-We presented numerous repetitions of full-field drifting gratings (Fig.~\ref{fig:01}A and \ref{fig:01}B) to the eye contralateral to the imaged site.
-This technique allowed us to record from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4.
-We deconvolved somatic calcium signals using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:01}C and \ref{fig:01}D) and subtracted the average stimulus response to remove averaged linear effects of the stimulus.
-From this residual response we computed the sample noise covariance matrix (Fig.~\ref{fig:01}E).
+We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005,Katona:2012,Cotton:2013}}.
+We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (Fig.~\ref{fig:03}\,A) on one side of the visual field of anesthetized mice.
+Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (Fig.~\ref{fig:03}\,B and E) in the visual cortex on the contralateral side to the stimulus.
+With acousto-optic modulator (AOD) 3D steering of the laser, the located cells were imaged at high sampling rates with concurrent motion detection.
+This technique allowed fast sampling (100-150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:03}\,C).
+After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:03}\,C and D).
+Only cells that produced detectable calcium activity were included in the analysis.
+The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (Fig.~\ref{fig:03}\,F and G).
+The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.
+Besides the high-repetition stimulus protocol for noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cells (Fig.~\ref{fig:03}\,E).
+Overall, 31 imaged sites from 24 animals were included in the study.
 
-In such highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability.
-At the same time, most correlations are relatively small (Fig.~\ref{fig:01}E), suggesting that a simple shrinkage towards independence may provide a sufficiently well regularized estimate.
-By comparing the performance of our differently regularized estimators we can gain insights into which of these aspects are important in our data.
+In such highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+lowrank}$) would be more efficient than others.
+At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), to the advantage of estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$).
+Finally, there is the possibility that the correlations structure is best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$).
 
-As expected, all regularized estimators outperformed the sample covariance estimator substantially (Fig.~\ref{fig:04}) Say something about how shrinkage, low-rank and sparse inverse relate.
-Finally, the combined sparse and low-rank estimator dominated all others significantly (Fig.~4), showing that both hidden units and direct interactions are important in our data.
-The improvement from the low-rank estimator to the combined one is much larger than that from the sparse inverse to the combined one, suggesting that in this dataset direct interactions contribute more strongly to the correlation structure than hidden units do.
+We found that estimator $C_{\sf sparse+lowrank}$ was more efficient than the other estimators.
+The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+lowrank}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).
+
+\hfill\rule{8cm}{4pt}
 
@@
 Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.
-
-\subsection*{}
 
@@
     \caption{{\bf The sparse+lowrank estimator $C_{\sf sparse+lowrank}$ outperforms the other estimators on neural data.}
-    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_0$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+lowrank}$. 
+    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+lowrank}$. 
     The histograms are based on 31 imaged sites in 24 mice. 

### 2013-12-03 plos.tex
Completed the  Results section.

@@
 With acousto-optic modulator (AOD) 3D steering of the laser, the located cells were imaged at high sampling rates with concurrent motion detection.
-This technique allowed fast sampling (100-150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:03}\,C).
+This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:03}\,C).
 After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:03}\,C and D).
@@
 
-In such highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+lowrank}$) would be more efficient than others.
+In these highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+lowrank}$) would be more efficient than others.
 At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), to the advantage of estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$).
@@
 
-We found that estimator $C_{\sf sparse+lowrank}$ was more efficient than the other estimators.
+We found that estimator $C_{\sf sparse+lowrank}$ was more efficient than the other estimators (Fig.~\ref{fig:04}).
 The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+lowrank}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).
 
-\hfill\rule{8cm}{4pt}
+\subsection*{Relationship of correlation structure to circuit architecture}
+Having demonstrated that $C_{\sf sparse+lowrank}$ dominated the other estimators, we examined the elements of the low-dimensional structure revealed by this estimator for individual imaged sites.
+By its design, the sparse+lowrank estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
+If this estimate dominates all other estimators evaluated thus far, it seems reasonable to hypothesize that these interaction may reflect underlying physiological interactions more closely than the usual correlations.
+In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
 
-\subsection*{Relationship between functional covariance structure and circuit architecture}
+We examined the structure of the sparse+lowrank estimate in individual sites such as the example site first depicted in Fig.\;\ref{fig:03}.
+In this site, as in others, the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (Fig.~\ref{fig:05}\,A and D).
+However, the corresponding partial correlation matrices differed dramatically (Fig.~\ref{fig:05}\,B and E).
+The partial correlation was decomposed into the sparse and low-rank components (Fig.\,\ref{fig:05}\,C).
+Although correlations were mostly positive, the sparse partial correlations (or `interactions'), although mostly positive, had a much larger proportion of negative values.
+The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G).
+The low-rank component was of rank 17.
+
+In previous publications that looked at the structure of correlations, the sparse network was obtained by thresholding the correlations coefficients  above some level deemed significant and examining the network of correlations above that threshold \cite{Golshani:2009,Malmersjo:2013}.
+There was fairly little overlap between the network of interactions revealed by this thresholding method and those revealed by the sparse+lowrank estimator (Fig.\;\ref{fig:05}\,F).
+When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (Fig.\;\ref{fig:05}\,F, H, and I).
+In particular, many low sample correlations translated into negative interactions in the regularized estimate.
+Indeed, the absence of a correlation between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient.
+Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.
+
+The average partial correlations  revealed by the regularized estimator in all 31 sites were about 5 times lower than the sample correlations and  less variable across sites (Fig.\;\ref{fig:06}\,A).
+The average node degree of the sparse component of the partial correlations and the number of  inferred latent units varied widely between sites but generally increased with the recorded population size (Fig.\;\ref{fig:06}\,B and C).
+However, there was an inverse relationship between the number of latent units and the average node degree (Fig.\;\ref{fig:06}\,D).
+Several sites, even with relatively large population sizes, had fairly few pairwise interactions and were dominated by latent units.
+These differences have multiple explanations, including differences in brain states and recording quality (neuropil contamination, motion).
+
+We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells  to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}).
+The average partial correlations fell more rapidly with with difference in preferred orientation (Fig.\;\ref{fig:07}\,A and D) and lateral displacements at equal depths ($\pm 25\mu$m) (Fig.\;\ref{fig:07}\,B and E), and differences in depth at small ($\pm 25\mu$m) lateral displacements (Fig.\;\ref{fig:07}\,B and E).
+
+The connectivity of partial correlations in the regularized estimate had different organization for positive and negative interactions.
+Positive interactions fell rapidly as a function of difference in preferred orientation (Fig.\;\ref{fig:07}\,G and J), lateral displacements (Fig.\;\ref{fig:07}\,H and K), and displacement in depth (Fig.\;\ref{fig:07}\,I and L).
+Negative interactions were much less selective (Fig.\;\ref{fig:07}\,G--L).
+
+\section*{Discussion}
+TODO
 
 \section*{Methods}
-\subsection*{Estimation of parameters and hyperparameters of covariance estimators}
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis" 
+\subsection*{Data Acquisition}
+Fast 3D two-photon imaging of calcium signals \emph{in vivo} was performed as described in \cite{Cotton:2013}
+\subsection*{Covariance Estimation}
+TODO
+\subsection*{Simulation}
+TODO
+\subsection*{Cross-validation}}
+TODO
 
-\subsection*{Regularization}
-Under many loss functions\footnote{
-The strict equality in Eq.~\ref{eq:bias-variance} does not hold under the loss function in Eq.~\ref{eq:loss}. 
-However, the equality does hold for its close cousin, Stein's \emph{entropy loss}, which only differs by the order of its arguments and a constant offset: $\mathcal L_s(\hat\Sigma,\Sigma) \equiv \loss{\Sigma,\hat\Sigma} - \loss{\Sigma,\Sigma}$.
-This deficiency presents no difficulty because we minimize the risk directly, without assessing the two error components individually. 
-The bias-variance decomposition is presented here to motivate the use of regularization.}, 
-the risk can be decomposed as the sum
-\begin{equation}\label{eq:bias-variance}
-    r = b + \varepsilon
-\end{equation}
-of \emph{approximation error} (``bias'' or systematic error)
-\begin{equation}
-   b = \loss{\bar\Sigma,\Sigma}
-\end{equation}
-and \emph{estimation error} (``variance'') 
-\begin{equation}
-   \varepsilon = \E[\hat\Sigma]{\loss{\hat\Sigma, \bar\Sigma}}
-\end{equation}
-where $\bar\Sigma = \E{\hat\Sigma}$ is the expected value of the estimate.
-
-An unbiased estimator, such as the sample covariance,  $\hat\Sigma_0,$ makes $\bar\Sigma=\Sigma$ and thereby minimizes approximation error, but may be excessively susceptible to sample noise and result in high estimation error.
-
-The estimator risk can be reduced by \emph{regularization} -- a deliberate biasing (\emph{``shrinkage''}) of the estimate toward a low-dimensional, less variable \emph{target estimate} \cite{Bickel:2006,Ledoit:2004}.
-A regularized estimator solves the bias-variance tradeoff to produce a biased but less variable estimate to minimize the estimator's risk.  
-Various regularization schemes focus on the dimensionality reduction part  by selecting the optimal target estimate from a family of estimates with reduced dimensionality \cite{findit}.  
-Other estimators only shrink the sample covariance matrix toward a single target estimator \cite{Schafer:2005}. 
-Yet other regularizers effectively combine shrinkage and dimensionality reduction \cite{findit}.
-
-\subsubsection*{Estimator $\mathcal A$: Shrinkage toward diagonal.}
-\emph{Estimator $\mathcal A$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward that of the independent model shown in Fig.~\ref{fig:02}A.
-The covariance matrix of a GGM in which all units are independent will be diagonal.
-For example, we  could leave the sample variances on the diagonal set the off-diagonal terms to zero. The target estimate
-is then 
-\begin{equation}\label{eq:sampleVariance}
-\hat T= \hat\Sigma_0\circ I
-\end{equation}
-where $\circ$ denotes the entrywise matrix product (Hadamard product). 
-Alternatively, the target could be a diagonal matrix with all variances equal to the average sample variance, $v = \frac 1 p \Tr(\hat\Sigma_0)$, so that
-\begin{equation}\label{eq:equalVariance}
-\hat T=\frac 1 p \Tr(\hat\Sigma_0) I
-\end{equation}
-The equal-variance target in Eq.~\ref{eq:equalVariance} has only one degree of freedom and thus lower estimation error but higher bias than the target with sample variances in Eq.~\ref{eq:sampleVariance}.
-It is not immediately clear which target is better in a particular application.
-We therefore use the linear mixture of the two targets paramaterized by the mixing coefficient $\alpha\in[0,1]$,
-\begin{equation}
-\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I.
-\end{equation}
-The estimator $\hat\Sigma_{\alpha,\lambda}$ is the linear shrinkage of the sample covariance matrix $\hat\Sigma_0$ toward $\hat T_\alpha$ controlled by the mixing proportion $\lambda\in[0,1]$,
-\begin{equation}
-\hat\Sigma_{\alpha,\lambda}^\mathcal{A} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha.
-\end{equation}
-This covariance matrix estimator is implemented by R's {\tt corpcor} package \cite{Schaefer:2010} and is used widely~\cite{Schafer:2005}.
-
-We estimated the optimal values of hyperparameters $ \alpha$ and $ \lambda$  by  cross-validation within the training dataset.
-
-\subsubsection*{Estimate $\mathcal B$: Shrinkage toward a factor model}
-\emph{Estimator $\mathcal B$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward a target estimated by \emph{factor analysis} with $d$ factors, which corresponds to graphical models of the kind in Fig.~\ref{fig:02}B.
-For multivariate normal distributions, factor analysis estimates a graphical model in which the observed units do not interact directly but are influenced by $d$ common inputs.
-
-The target estimate is then 
-\begin{equation}
-\hat T_d = \hat L_d + \hat \Psi
-\end{equation}
-where $\hat L_d$ is a $p\times p$ matrix of rank $d$ and $\hat \Psi$ is a $p\times p$ diagonal matrix. The value of $\hat T_d$ is found by minimizing $\loss{\hat T_d, \hat\Sigma_0}$.
-
-Just as with estimator $\mathcal A$, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$,
-\begin{equation}
-\hat\Sigma_{d,\lambda}^\mathcal{B} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d,
-\end{equation}
-and the optimal values of $d$ and $\lambda$ are found by cross-validation within the training set.
-
-Factor analysis has long been used in natural sciences to infer the influence of a one or several latent variables exerting linear effects on a larger population of observed variables \cite{Jvreskog:1996}.
-In neuroscience, the hypothesis that the population activity in small local circuits is largely driven by a few latent factors has been investigated \cite{Yu:2009,Ecker:2013}.
-Estimation of covariance matrices by shrinkage toward a factor model has been used in finance for portfolio risk assessment \cite{Ledoit:2003,Fan:2008}.
-
-\subsubsection*{Estimator $\mathcal C$: Spare inverse}
-\emph{Estimator $\mathcal C$} produces a low-dimensional estimate by finding an approximation of $\hat\Sigma_0$ that has many zeros in its inverse.
-This approximation problem is known as \emph{covariance selection} \cite{Dempster:1972}.
-Covariance matrices of this family correspond to graphical models of the type depicted in Fig.~\ref{fig:02}C.
-
-For Gaussian Graphical Models, zeros in the inverse covariance matrix indicate statistical independence between the corresponding pairs.
-Thus the inverse covariance matrix $S=-\Sigma^{-1}$ is also the adjacency matrix of the GGM, indicating which pairs are conditionally dependent.
-The off-diagonal elements of $S$ are proportional to partial pairwise correlations between corresponding pairs of neurons.
-
-The covariance matrix estimate is then 
-\begin{equation}
-\hat\Sigma_\rho = \hat S^{-1}
-\quad\mbox{with}\quad
-\hat S = \argmin\limits_{\substack{S\succ 0 \\ \|S\|_0 \le \rho}} \loss{S^{-1},\hat\Sigma_0}   
-\end{equation}
-where $S\succ 0$ signifies the constraint that $S$ be positive-definite. $\|S\|_0\le\rho$ signifies the constraint  that matrix $S$ have at most $\rho$ non-zero coefficients.
-Solving this approximation in this form is computationally challenging.
-The problem can be made computationally tractable by relaxing \TODO{use "complex relaxation"} the $L_0$ norm to the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.
-The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator
-\begin{equation}
-\hat\Sigma_\lambda^\mathcal{C} = \hat S^{-1}
-\quad\mbox{with}\quad
-\hat S = \argmin\limits_{S\succ 0} \loss{S^{-1},\hat\Sigma_0} + \lambda \|S\|_1
-\end{equation}
-where $\lambda \|S\|_1$ is the $L_1$ norm of the matrix.
-
-Just as in the previous estimators, the optimal value of the regularization parameter $\lambda$ is found from training data by cross-validation.
-
-\subsubsection*{Estimator $\mathcal D$: Sparse + low-rank inverse}
-\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by approximating $\hat\Sigma_0$ with a matrix whose inverse is the sum of a spase matrix and a low-rank matrix.
-Covariance matrices of this form correspond to Gaussian graphical models of the form depicted in Fig.~\ref{fig:02}D.
-
-The inverse $S=\Sigma^{-1}$ of the covariance matrix of a GGM constitutes the graph of conditional dependencies (interactions) between the units.
-With the assumption of sparse organization of interactions in the network, $S$ is assumed to be sparse.
-When the observed units are part of a larger GGM that includes other latent units, let the negative inverse $S^*$ of the joint distribution of both observed and latent units be composed of four partitions
-\begin{equation}
-S^* = 
-\begin{pmatrix}
-S_{11} & S_{12} \\
-S_{12}^\T & S_{22} 
-\end{pmatrix}
-\end{equation}
-where partition $S_{11}$ describes the interactions between the visible units, $S_{12}$ describe the interactions between visible and latent units, and $S_{22}$ describes the interactions between the latent units.
-Then the covariance matrix of between the observed unit is 
-\begin{equation}
-\Sigma = \left(S_{11} - S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
-\end{equation} 
-The rank of matrix $S_{12}S_{22}^{-1}S_{12}$ equals the number of latent units.
-Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $ S_{12}S_{22}^{-1}S_{12}$ with a low-rank matrix $L= S_{12}S_{22}^{-1}S_{12}$.
-
-Then we obtain estimator $\hat\Sigma_{d,\rho}$, parameterized by the sparsity parameter $\rho$ and  the number of latent units $d$:
-\begin{equation}
-\hat\Sigma_{d,\rho} = \left(\hat S-\hat L\right)^{-1}
-\quad\mbox{where}\quad
-(\hat S, \hat L) = \argmin\limits_{\substack{S,L\\ (S-L)\succ 0 \\ \|S\|_0\le\rho \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0}
-\end{equation}
-
-Again, in this formulation, the approximation problem is computationally intractible.
-The problem can be made convex by replacing the $L_0$ norm with $L_1$ norm.
-The resulting algorith, known as \emph{latent-variable graphical lasso} or \emph{LV-GLASSO} \cite{Chandrasekaran:2010,Ma:2013}, produces estimator $\hat\Sigma_{d,\lambda}^{\mathcal D}$:
-\begin{equation}
-\hat\Sigma_{d,\lambda} ^{\mathcal D} = \left(\hat S-\hat L\right)^{-1}
-\quad\mbox{where}\quad
-(\hat S, \hat L) = \argmin\limits_{\substack{ S,L \\ (S-L)\succ 0 \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0} + \lambda \| S \|_1 
-\end{equation}
-The optimal values of hyperparameters $\lambda$ and $d$ are found cross-validation within the training dataset.
-
-% Do NOT remove this, even if you are not including acknowledgments
+\subsection*{Evaluation}
 \section*{Acknowledgments}

### 2013-12-03 plos.tex
minor corrections

@@
 
-We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005,Katona:2012,Cotton:2013}}.
+We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005,Katona:2012,Cotton:2013}.
 We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (Fig.~\ref{fig:03}\,A) on one side of the visual field of anesthetized mice.
@@
 TODO
-\subsection*{Cross-validation}}
+\subsection*{Cross-validation}
 TODO
@@
         {\bf Row 3.} Examples of sample correlation matrices calculated from samples of 1000 observations taken from simulated random processes with corresponding correlation matrices from row 2.
-        {\bf Row 4.} Average normalized Gaussian log-likelihood \emph{excess loss} (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators as a function of sample size.
+        {\bf Row 4.} Mean \emph{excess loss} (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators as a function of sample size.
 The error bars indicate the standard error of the mean based on 30 samples.
-        {\bf Row 5.} Average normalized Gaussian log-likelihood \emph{validation loss} (Eq.~\ref{eq:validation-loss}) attained by each of the five estimators.
+        {\bf Row 5.} Mean \emph{cross-validation loss} (Eq.~\ref{eq:rel-cv-loss}) of covariance estimators with respect to the matching estimator .
 The values are relative to the validation loss of the estimator that matches the low-dimensional structure of the true covariance matrix.

### 2013-12-06 plos.tex
minor
incorporated some of Andreas' comments.

@@
 Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons in behaving animals.
-It is anticipated that correlations measured in such recordings will uncover aspects of functional organization of neural circuits.
-However, estimation and interpretation of large correlation matrices from finite recordings can be challenging.
-Estimation can be improved by regularization: the biasing of the estimate toward a low-dimensional, less variable approximation.
-The amount of improvement depends on how closely the reduced approximation captures the dominant interactions with fewest terms.
-Therefore, the selection of the most efficient estimator is an empirical question dependent on the system under investigation.
+It is anticipated that correlations measured in such recordings will uncover important aspects of the functional organization of neural circuits.
+However, estimation and interpretation of large correlation matrices from finite recordings are challenging.
+Estimation can be improved by regularization: the biasing of the estimate towards a low-dimensional, less variable approximation.
+The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies.
+Therefore, the selection of the most efficient estimator is an empirical question that depends on the system under investigation.
 
-We compared regularized correlation matrix estimators biased toward four respective families of low-dimensional correlation structures: independent, latent factors, sparse partial correlations, and sparse partial correlations with latent factors.
-The estimators' performance was evaluated on the noise correlation matrices from spatially compact groups of 91--314 neurons (average 206) in mouse visual cortex (31 sites in 24 animals) acquired by \emph{in vivo} fast 3D random-access two-photon imaging of calcium signals.
-Recordings lasting between 15 and 20 minutes were deconvolved and binned at 150 ms.
-Each estimator was optimized and evaluated by nested cross-validation.
-We found that both sparse partial correlations and latent factors were required for efficient estimation.
-In addition to reducing estimation error, optimized estimates provided compact representations of the correlation structure consisting of a sparse network of pairwise interactions and global diffuse fluctuations.
-The density of positive interactions in these networks decreased rapidly with distance and with difference in orientation preference, whereas negative interactions were less selective.
-The physiological interpretation of these components remains in question and must be addressed by future experiments where synaptic connectivity, cell types, and local field potentials may be used to corroborate such interpretations.
+In this study, we sought to identify the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.
+Using fast 3D random-access laser scanning microscopy of calcium signals we recorded the activity of nearly every neuron in volumes of cortex about 100 $\mu$m in diameter (150--350 cells).
+We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with several shared latent factors to capture unobserved inputs and network fluctuations.
+To test this hypothesis, we demonstrated that, in cross-validation tests, a covariance matrix estimator based on this correlation structure outperformed estimators regularized toward several other plausible low-dimensional approximations.
+We analyzed how the inferred network of interactions was related to the physical distances between cells and their orientation preferences.
+Consistent with known synaptic connectivity, the density of positive interactions decreased rapidly with distance and with differences in preferred orientation, whereas negative interactions were less selective.
+To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and local field potentials.
 
@@
 \begin{equation}
-C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T, 
+C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
 \end{equation}
@@
 Finally, we consider the fourth estimator $C_{\sf sparse+lowrank}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D).
-This estimator has the form
+This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}
 \begin{equation}
-C_{\sf sparse+latent} = (S - LL^\T)^{-1},
+C_{\sf sparse+latent} = (S - LL^\T)^{-1}
 \end{equation}
@@
 \section*{Discussion}
-TODO
+\subsection*{Big neural data}
+\subsection*{Meaning of interactions}
+\subsection*{Fitting probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
+\subsection*{Future directions}
 
@@
 TODO
+\subsection*{Cross-validation}
+TODO
 \subsection*{Simulation}
 TODO
-\subsection*{Cross-validation}
-TODO
 
-\subsection*{Evaluation}
 \section*{Acknowledgments}

### 2013-12-06 plos.tex
Updated Abstract and Author summary, Figure 7.

@@
 % Please keep the abstract between 250 and 300 words
-Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons in behaving animals.
-It is anticipated that correlations measured in such recordings will uncover important aspects of the functional organization of neural circuits.
+Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.
+Correlations measured in such recordings are anticipated to uncover important aspects of the functional organization of neural circuits.
 However, estimation and interpretation of large correlation matrices from finite recordings are challenging.
-Estimation can be improved by regularization: the biasing of the estimate towards a low-dimensional, less variable approximation.
-The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies.
-Therefore, the selection of the most efficient estimator is an empirical question that depends on the system under investigation.
+Estimation can be improved by regularization: biasing of the estimate towards a low-dimensional, less variable approximation.
+The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.
+Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies present in the system.
 
-In this study, we sought to identify the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.
-Using fast 3D random-access laser scanning microscopy of calcium signals we recorded the activity of nearly every neuron in volumes of cortex about 100 $\mu$m in diameter (150--350 cells).
-We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with several shared latent factors to capture unobserved inputs and network fluctuations.
-To test this hypothesis, we demonstrated that, in cross-validation tests, a covariance matrix estimator based on this correlation structure outperformed estimators regularized toward several other plausible low-dimensional approximations.
-We analyzed how the inferred network of interactions was related to the physical distances between cells and their orientation preferences.
-Consistent with known synaptic connectivity, the density of positive interactions decreased rapidly with distance and with differences in preferred orientation, whereas negative interactions were less selective.
-To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and local field potentials.
+In this study, we sought the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.
+Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 100 $\mu$m in diameter (150--350 cells) in mouse visual cortex.
+We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.
+Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward other plausible low-dimensional approximations.
+Furthermore, we found consistent relationships between the organization of sparse interactions inferred by the estimator and the physical distances and preferred orientation differences of pairs of cells.
+Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with differences in preferred orientation, whereas negative interactions were less selective.
+To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and brain states.
 
@@
 % Author Summary not valid for PLoS ONE submissions.   
-Correlation matrices of the activity of populations of neurons are useful descriptors of their functional organization with implications for stimulus coding and circuit architecture.
-Estimates of the correlation matrix can be dramatically improved by \emph{regularization}: the biasing of the estimate toward one of several possible low-dimensional approximations.
-Using synthetic data, we demonstrated that different estimators performed better or worse than others depending on how well their regularization targets captured the structure of the true correlation matrix.
-Then, we compared the performance of several estimators on the population calcium activity of relatively large and dense groups of neurons in mouse visual cortex under visual stimulation.
-We found that the best estimator for such neural data was composed of a sparse network of partial correlations between pairs of neurons combined with several latent units interacting with the entire population.
-We hypothesize that such succinct representations of the correlation structure may more accurately reflect the circuit's functional organization than the conventional correlations.
+Correlations of the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.
+Estimation of correlation matrices can be improved by imposing some kind of structure on the estimate with greatest improvement attained when the imposed structure closely matches the real dependencies in the data.
+Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the cross-validated performance of correlation matrix estimators that imposed different kinds of structure.
+The correlation structure of the estimator that proved most efficient comprised a sparse network of partial correlations between pairs of neurons combined with several common latent factors, with both components required for efficient estimation.
+Because of its superior benefit for estimation, we proposed that this inferred structure could prove more relevant for the description of functional connectivity than the usual correlation matrix in densely sampled neural recordings.
+As a first application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.
+
 \section*{Introduction}
-Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Renart:2010}.
+Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.
 For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}.
-In addition, neural correlations are hypothesized to reflect aspects of functional connectivity in neural circuits.
+In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.
 Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
@@
 
+By their construction, the correlation between a pair of cells expresses the degree to which the activity of one cell predicts the activity of the other cell.
+With multiple cells If interactions between variables have a strong linear components, then correlations also statistical dependence.
+
 A correlation matrix provides more information than the equivalent number of pairwise correlations assessed in isolation.
@@
 Such estimates can be transformed into various representations that express different aspects of the correlation structure and may suggest different mechanistic interpretations.
-For example, factor analysis or principle component analysis may reveal common activity fluctuations across the entire population \cite{Yu:2009}.
+For example, factor analysis or principal component analysis may reveal common activity fluctuations across the entire population \cite{Yu:2009}.
 The correlation matrix also allows computing the \emph{partial correlations} between specific pairs of neurons, \emph{i.e.}\;correlations that remain after accounting for the linear effects of all other recorded cells.

### 2013-12-11 plos.tex
updated introduction

@@
 Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.
-Correlations measured in such recordings are anticipated to uncover important aspects of the functional organization of neural circuits.
+Correlations measured in such recordings are anticipated to reveal important aspects of the functional organization of neural circuits.
 However, estimation and interpretation of large correlation matrices from finite recordings are challenging.
-Estimation can be improved by regularization: biasing of the estimate towards a low-dimensional, less variable approximation.
+Estimation can be improved by regularization: biasing of the estimate toward a low-dimensional, less variable approximation.
 The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.
@@
 Furthermore, we found consistent relationships between the organization of sparse interactions inferred by the estimator and the physical distances and preferred orientation differences of pairs of cells.
-Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with differences in preferred orientation, whereas negative interactions were less selective.
+Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with difference in preferred orientation whereas negative interactions were less selective.
 To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and brain states.
@@
 
-By their construction, the correlation between a pair of cells expresses the degree to which the activity of one cell predicts the activity of the other cell.
-With multiple cells If interactions between variables have a strong linear components, then correlations also statistical dependence.
+Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
+In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998,Zohary:1994}.
+In contrast, modern multineuronal recordings from increasingly large and dense populations of highly interconnected neurons allow estimation of various low-dimensional representations of the entire correlation matrix, which can enable inquiry into the nature of underlying physiological mechanisms and computational principles.
 
-A correlation matrix provides more information than the equivalent number of pairwise correlations assessed in isolation.
-Yet early studies of neural correlations were based on measurements from isolated pairs of neurons; the effects of correlations on neural coding were then extrapolated in simulations and theoretical studies \cite{Shadlen:1998,Zohary:1994}.
-Modern multineuronal recordings from increasingly large and dense subsets of neurons in intact circuits allow estimating the entire correlation matrix of large populations of highly interconnected neurons.
-Such estimates can be transformed into various representations that express different aspects of the correlation structure and may suggest different mechanistic interpretations.
+In this study, we focused on low-dimensional representations composed of two types of elements: the \emph{low-rank correlation structure} and \emph{partial correlations} between pairs of neurons.
+The former is most suitable for expressing common circuit fluctuations whereas the latter are most suitable for expressing detailed interactions between specific neurons.
+Low-rank neural correlation structure extracted by principal component analysis or factor analysis has been analyzed for temporal dynamics in in several studies \cite{Nicolelis:1995,Yu:2006,Yu:2009}.
+
+Partial correlations between pairs of neurons are linear correlations remaining after subtracting the linear effect of all the other neurons.
+In systems where effects exerted by variables on one another have strong linear components, partial correlations describe conditional dependency: zero  partial correlation between two variables implies conditional independence.
+Conditional independence between two variables, if all relevant variables are observed, implies a lack of direct interaction between the underlying processes.
+Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been applied to infer interactions between components of complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
+In multivariate normal distributions\footnote{or any other \emph{elliptical} distribution of the form $g(x-\mu)^\T\Sigma^{-1}(x-\mu))$, where $g$ is an arbitrary function}, are completely defined by linear effects.
+A multivariate normal distribution defined through a sparse network of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
+With other distributions where nonlinear effects are more significant, the correspondence between conditional dependence and partial correlations diminishes or breaks down \cite{Loh:2012}.
+
 For example, factor analysis or principal component analysis may reveal common activity fluctuations across the entire population \cite{Yu:2009}.
 The correlation matrix also allows computing the \emph{partial correlations} between specific pairs of neurons, \emph{i.e.}\;correlations that remain after accounting for the linear effects of all other recorded cells.
-Under several assumptions (predominance of linear effects ( sufficiently complete sampling of interacting neurons, and sufficient sample size), partial correlations may better reflect the direct statistical associations or interactions between pairs of neurons.
+Under several assumptions (predominance of linear effects (sufficiently complete sampling of interacting neurons, and sufficient sample size), partial correlations may better reflect the direct statistical associations or interactions between pairs of neurons.
 In multivariate normal distributions, zero partial correlations between pairs of variables indicate conditional independence or lack of interaction between the pair and the network of non-zero interactions constitute Gaussian Graphical Models or Gauss-Markov Random Fields \cite{Koller:2009}.

### 2013-12-12 plos.tex
updtated introduction

@@
 In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.
-Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
+Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
 
@@
 
-In this study, we focused on low-dimensional representations composed of two types of elements: the \emph{low-rank correlation structure} and \emph{partial correlations} between pairs of neurons.
-The former is most suitable for expressing common circuit fluctuations whereas the latter are most suitable for expressing detailed interactions between specific neurons.
-Low-rank neural correlation structure extracted by principal component analysis or factor analysis has been analyzed for temporal dynamics in in several studies \cite{Nicolelis:1995,Yu:2006,Yu:2009}.
+For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation.
+Low-rank approximations are particularly suitable for capturing shared fluctuations across the recorded population.
+Extracted through principal component or factor analysis, low-rank components of neural correlations have been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2006,Yu:2009}.
 
-Partial correlations between pairs of neurons are linear correlations remaining after subtracting the linear effect of all the other neurons.
-In systems where effects exerted by variables on one another have strong linear components, partial correlations describe conditional dependency: zero  partial correlation between two variables implies conditional independence.
-Conditional independence between two variables, if all relevant variables are observed, implies a lack of direct interaction between the underlying processes.
-Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been applied to infer interactions between components of complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
-In multivariate normal distributions\footnote{or any other \emph{elliptical} distribution of the form $g(x-\mu)^\T\Sigma^{-1}(x-\mu))$, where $g$ is an arbitrary function}, are completely defined by linear effects.
-A multivariate normal distribution defined through a sparse network of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
-With other distributions where nonlinear effects are more significant, the correspondence between conditional dependence and partial correlations diminishes or breaks down \cite{Loh:2012}.
-
-For example, factor analysis or principal component analysis may reveal common activity fluctuations across the entire population \cite{Yu:2009}.
-The correlation matrix also allows computing the \emph{partial correlations} between specific pairs of neurons, \emph{i.e.}\;correlations that remain after accounting for the linear effects of all other recorded cells.
-Under several assumptions (predominance of linear effects (sufficiently complete sampling of interacting neurons, and sufficient sample size), partial correlations may better reflect the direct statistical associations or interactions between pairs of neurons.
-In multivariate normal distributions, zero partial correlations between pairs of variables indicate conditional independence or lack of interaction between the pair and the network of non-zero interactions constitute Gaussian Graphical Models or Gauss-Markov Random Fields \cite{Koller:2009}.
-With other distributions, the correspondence between conditional dependence and partial correlations diminishes or breaks down \cite{Loh:2012}.
-Estimation of networks of partial correlations (sometimes called \emph{association networks}) have been applied to map gene interactions \cite{Schafer:2005,Peng:2009} or to establish functional connectivity between brain regions from fMRI data \cite{Varoquaux:2012,Ryali:2012}.
-
-Multineuronal recordings also present major challenges for the estimation and interpretation of correlations.
-As the population size increases, the amount of recorded data grows only linearly whereas the number of estimated coefficients grows quadratically.
-This mismatch leads to increased opportunities for spurious correlations, overestimation of shared activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned estimates of the partial correlations \cite{Schafer:2005}.
+Alternatively, the correlation matrix can be formulated through the partial correlations between pairs of neurons.
+The partial correlation between a pair of neurons is the linear correlation remaining after accounting for correlations to all the other neurons.
+Partial correlations carry particular significance when conditional dependencies between all the variables are thought to be dominated by linear effects.
+In such systems, zero  partial correlation between two variables implies their conditional independence.
+Conditional independence between a pair of variables suggests a lack of direct interaction between the underlying processes.
+Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been used to infer interactions between components in complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
+In \emph{elliptical distributions} (of the form $g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies.
+A multivariate normal distribution specified through its graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
+With increasing departure from linearity, the correspondence between conditional dependence and partial correlations diminishes or completely breaks down \cite{Loh:2012}.
+For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations.
 
 In this study, we pursued two related aims: (a)~improved estimation of neural correlation matrices and (b)~discovery of low-dimensional structure of correlations in recordings of multineuronal activity to facilitate their interpretation.
+
+The true correlation matrix is the normalized version of the true covariance matrix defined as
+\begin{equation}\label{eq:true-covariance}
+    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
+\end{equation}
+where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates.
+The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
+\begin{equation}
+    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
+\end{equation}
+where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent).
+In this study, we keep the sample mean $\bar x$ as the estimate of the true mean $\mu$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.
+
+Estimation of the covariance matrix from a large population presents a number of numerical challenges.
+The amount of recorded data grows only linearly with the population size whereas the number of estimated coefficients increases quadratically.
+This mismatch leads to increased opportunities for spurious correlations, to overestimation of shared activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and to poorly conditioned estimates of the partial correlations \cite{Schafer:2005}.
+
+Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation referred to here as the \emph{target estimate} \cite{Schafer:2005,Bickel:2006}.
+The usual estimate $C_{\sf 0}$ has the advantage of being unbiased, \emph{i.e.}\;$\E{C_0}=\Sigma$ but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in the data.
+Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from truth $\Sigma$.
+Regularization works by striking a favorable balance between bias and variability.
+Regularization can produce \emph{some} improvement even with an arbitrary target estimate that has no relation to the true dependencies in the data.
+Yet, when the target estimate is well suited for capturing the important features of the true covariance matrix with few terms, it will introduce minimal bias and outperform other estimators.
+
+Figure \ref{fig:01} illustrates the challenges of estimating the correlation matrix from a finite sample.
+Panel \ref{fig:01}\,A depicts the sample correlation matrix $R_{\sf 0}$ of a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex.
+$R_{\sf 0}$ is computed by normalizing the covariance matrix $C_{\sf 0}$ by the sample variances:
+\begin{equation}
+    R_{\sf 0} = \left(I\circ C_{\sf 0}\right)^{-\frac 1 2} C_{\sf 0} \left(I\circ C_{\sf 0}\right)^{-\frac 1 2}
+\end{equation}
+Here $\circ$ denotes entrywise matrix product (Hadamard product). Thus $I\circ C_{\sf 0}$ represents the diagonal matrix with the diagonal elements from $C_{\sf 0}$, \emph{i.e.} the sample variances. 
+Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are much higher than typical firing rate correlations on shorter temporal scales.
+To show greater detail, panel \ref{fig:01}\,B magnifies the fragment of the sample correlation matrix that is outlined in \ref{fig:01}\,A.
+The matrix of partial correlations $P$ is obtained by scaling the inverse of the correlation matrix:
+\begin{equation}
+    P = -D^{-1/2} R^{-1} D^{-1/2}
+\end{equation}
+where $D=I\circ R^{-1}$ is the diagonal matrix with the diagonal elements of the inverse of the correlation matrix $R$.
+Thus $P$ contains $-1$s on its diagonal and the off-diagonal elements contain the pairwise partial correlations.
+Figure \ref{fig:01}\,C shows the magnified fragment of the partial correlation matrix $P$.
+
+Panels \ref{fig:01}\,D and E depict a different estimate obtained from the same data by \emph{covariance selection} \cite{Dempster:1972}.
+In covariance selection, only a subset of partial correlations is fitted while the rest are fixed at zero.
+The resulting estimate closely reproduces the sample correlation matrix (Fig.~\ref{fig:01}\,D), but its partial correlation structure appears substantially different(Fig.~\ref{fig:01}\,E).
+In this particular example  31,501 of the 44,253 (71.2\%) total partial correlations are set to zero.
+This alternative estimate is attractive for a few reasons.
+First, it has fewer free parameters and is therefore less susceptible to sampling noise.
+Additionally, assuming that partial correlations do indeed reflect elements of the functional connectivity, the second estimate also appears more physiologically plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.
+
+It is therefore important to compare multiple families of regularized estimators on the system under investigation.
+The family that consistently outperforms the others may be proposed as candidate low-dimensional representation of the correlation structure.
+The greater efficiency of one estimator over another is not a mathematical necessity but an empirical finding depending on the prevalent types of correlation structure in the investigated system.
 
@@
 Therefore, we rephrase the problem as one of optimal estimation of the covariance matrix given an estimate of the mean firing rates.
-
-Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward one of several possible low-dimensional approximations referred to here as \emph{target estimates} \cite{Schafer:2005,Bickel:2006}.
-The usual estimate, the \emph{sample covariance matrix} has the advantage of being unbiased but, on average, it falls far from the true covariance matrix due to its sensitivity to sampling noise in the data.
-Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from the true covariance matrix.
-Regularization works by striking a favorable balance between bias and variability.
-Regularization can produce \emph{some} improvement even with an arbitrary target estimate that has no relation to the true covariance structure in the data.
-Yet, when the target estimate is well suited for capturing the important features of the true covariance matrix with few terms, it will introduce minimal bias and outperform other estimators.
-
-To illustrate key concepts, Figure \ref{fig:01} illustrates a regularization scheme based on \emph{covariance selection} \cite{Dempster:1972}.
-In covariance selection, the estimate is produced by fitting only an optimal subset of the coefficients of the \emph{precision matrix} (the inverse of the covariance matrix) while setting the remaining coefficients to zero.
-Scaling the precision matrix so that its diagonal is composed of all $-1$'s produces the matrix of partial correlations.
-Thus zeros in the precision matrix produce zero partial correlations between the corresponding pairs of neurons.
-Panel \ref{fig:01}\,A depicts the sample correlation matrix of a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex.
-Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are much higher than typical firing rate correlations on shorter temporal scales.
-Panel \ref{fig:01}\,B magnifies the fragment of the sample correlation matrix outlined in \ref{fig:01}\,A whereas \ref{fig:01}\,C shows the corresponding fragment of the regularized estimate.
-The regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) off-diagonal coefficients of the precision matrix and fitting the remaining coefficients.
-Panels \ref{fig:01}\,D and E show the same fragments of the respective estimates of the partial correlation matrix.
-The apparent similarity of the two estimates of the correlation matrix belies the utterly different partial correlation structure.
-The regularized estimate is less sensitive to sample noise in the data, is likely to be closer to truth in most cases than the unbiased estimate, which may be ascertained by cross-validation studies.
-
-It may be tempting to interpret the regularized sparse estimate of partial correlations as a more accurate representation of the functional connectivity than the original correlation coefficients: after all, the regularized solution consistently produces better estimates and has fewer parameters which are easier to interpret.
-However, an alternative regularization scheme using 43 latent factors interacting linearly with all neurons while having about the same number of parameters, performed even better than covariance selection on this dataset.
-
-It is therefore important to compare multiple families of regularized estimators on the system under investigation.
-The family that consistently outperforms the others may be proposed as candidate low-dimensional representation of the correlation structure.
-The greater efficiency of one estimator over another is not a mathematical necessity but an empirical finding depending on the prevalent types of correlation structure in the investigated system.
 
@@
 \subsection*{Covariance estimation}
-We aim to estimate the true covariance matrix 
-\begin{equation}\label{eq:true-covariance}
-\Sigma = \E{(x-\mu)(x-\mu)^\T}
-\end{equation}
-where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$. The vector of mean firing rates is $\mu = \E{x}$.
-
-The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(1),\ldots,x(n)$ as
-\begin{equation}
-C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\mu)(x(t)-\mu)^\T 
-\end{equation}
-where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent).  
-The sample covariance matrix is unbiased, \emph{i.e.} $\E{C_{\sf 0}}=\Sigma$.
-For finite sample sizes, however, $C_{\sf 0}$ is not as close to $\Sigma$ as a number of biased estimators.
-
 We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (Fig.~\ref{fig:02}\;row 1).

### 2013-12-15 plos.tex
minor

@@
 \section*{Introduction}
+\subsection*{Neural correlations}
 Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.
@@
 
+\subsection*{Additional information in neural correlation matrices}
 Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
@@
 Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been used to infer interactions between components in complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
-In \emph{elliptical distributions} (of the form $g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies.
+In \emph{elliptical distributions} (those of the form $\frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is a normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies.
 A multivariate normal distribution specified through its graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
@@
 
+\subsection*{Estimation of neural correlation matrices}
 In this study, we pursued two related aims: (a)~improved estimation of neural correlation matrices and (b)~discovery of low-dimensional structure of correlations in recordings of multineuronal activity to facilitate their interpretation.
@@
 
+Given a covariance matrix estimate $C$, the correlation matrix $R$ and partial correlation matrix $P$ are computed as
+\begin{equation}
+    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
+\end{equation}
+\begin{equation}
+    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
+\end{equation}
+Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix.
+$R$ contains 1s on its diagonal whereas $P$ contains $-1$s on the diagonal.
+The off-diagonal elements of $P$ contain the partial correlations between corresponding pairs of neurons.
+
 Estimation of the covariance matrix from a large population presents a number of numerical challenges.
@@
 
-Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation referred to here as the \emph{target estimate} \cite{Schafer:2005,Bickel:2006}.
-The usual estimate $C_{\sf 0}$ has the advantage of being unbiased, \emph{i.e.}\;$\E{C_0}=\Sigma$ but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in the data.
+Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.
+The usual estimate $C_{\sf 0}$ has the advantage of being unbiased ($\E{C_0}=\Sigma$) but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in the data.
 Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from truth $\Sigma$.
@@
 
-Figure \ref{fig:01} illustrates the challenges of estimating the correlation matrix from a finite sample.
-Panel \ref{fig:01}\,A depicts the sample correlation matrix $R_{\sf 0}$ of a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex.
-$R_{\sf 0}$ is computed by normalizing the covariance matrix $C_{\sf 0}$ by the sample variances:
-\begin{equation}
-    R_{\sf 0} = \left(I\circ C_{\sf 0}\right)^{-\frac 1 2} C_{\sf 0} \left(I\circ C_{\sf 0}\right)^{-\frac 1 2}
-\end{equation}
-Here $\circ$ denotes entrywise matrix product (Hadamard product). Thus $I\circ C_{\sf 0}$ represents the diagonal matrix with the diagonal elements from $C_{\sf 0}$, \emph{i.e.} the sample variances. 
-Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are much higher than typical firing rate correlations on shorter temporal scales.
-To show greater detail, panel \ref{fig:01}\,B magnifies the fragment of the sample correlation matrix that is outlined in \ref{fig:01}\,A.
-The matrix of partial correlations $P$ is obtained by scaling the inverse of the correlation matrix:
-\begin{equation}
-    P = -D^{-1/2} R^{-1} D^{-1/2}
-\end{equation}
-where $D=I\circ R^{-1}$ is the diagonal matrix with the diagonal elements of the inverse of the correlation matrix $R$.
-Thus $P$ contains $-1$s on its diagonal and the off-diagonal elements contain the pairwise partial correlations.
-Figure \ref{fig:01}\,C shows the magnified fragment of the partial correlation matrix $P$.
+Figure \ref{fig:01} illustrates the challenges of estimating the correlation matrix from a finite sample and one example of regularization.
+Panel A depicts the correlation matrix computed from the sample covariance matrix $C_{\sf 0}$ for a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex.
+To show detail, panel B magnifies the fragment outlined in panel A.
+Panel C shows the same fragment from the partial correlation matrix also computed from $C_{\sf 0}$.
+Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
 
-Panels \ref{fig:01}\,D and E depict a different estimate obtained from the same data by \emph{covariance selection} \cite{Dempster:1972}.
+Panels \ref{fig:01}\,D and E depict the correlation matrix and the partial correlation matrix computed from a different estimate, which we will refer to here as $C_{\sf sparse}$, which is computed through \emph{covariance selection} \cite{Dempster:1972}.
 In covariance selection, only a subset of partial correlations is fitted while the rest are fixed at zero.
@@
 The greater efficiency of one estimator over another is not a mathematical necessity but an empirical finding depending on the prevalent types of correlation structure in the investigated system.
-
-Estimation of the correlation matrix relies on estimating the means and variances of the activity of individual cells.
-To estimate the mean activity, we resort to the usual \emph{sample mean} without attempting to improve it.
-Estimation of a \emph{covariance matrix} is equivalent to estimating both  variances and correlations.
-Therefore, we rephrase the problem as one of optimal estimation of the covariance matrix given an estimate of the mean firing rates.
 
@@
 \section*{Discussion}
-\subsection*{Big neural data}
+\subsection*{Functional connectivity from big neural data}
+Breakthroughs in neuroscience
+
 \subsection*{Meaning of interactions}

### 2013-12-16 plos.tex
outline of discussion

@@
 
-It is therefore important to compare multiple families of regularized estimators on the system under investigation.
-The family that consistently outperforms the others may be proposed as candidate low-dimensional representation of the correlation structure.
-The greater efficiency of one estimator over another is not a mathematical necessity but an empirical finding depending on the prevalent types of correlation structure in the investigated system.
+\subsection*{Model selection}
+Covariance selection is but one of numerous possible low-dimensional approximations of the covariance matrix.
+Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix.
+Models that accurately reflect the dominant functional dependencies in the circuit, its so-called ``functional connectivity'', will tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators.
+An estimator's superior efficiency can be ascertained by cross-validation.
+Because different types of neural circuits are governed by different types of dependencies, different probabilistic models and covariance matrix estimators may dominate in each case.
+The estimator that is shown to dominate all others for a specific type of neural circuit may be proposed as the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
+Its structure may then be analyzed and related to the circuits anatomical organization.
 
-In this study, we compared four families of regularized combining partial interactions and latent units: `shrinkage toward diagonal', `shrinkage toward a multifactor model', `sparse partial correlations', and 'sparse+latent partial correlations'.
-First, in simulation with known low-dimensional structure of correlations, we demonstrate that regularized estimators from matching families of low-dimensional targets generally outperform all other estimators.
-We then repeat this study on neural data from dense groups of neurons in mouse visual cortex under visual stimulation.
-We find that most efficient estimation was achieved by modeling correlations as a sparse network of partial correlations combined with several latent units interacting with the entire population.
-Finally, we analyze the relationship between the elements of such estimates and other properties of the circuit such as physical distances between neurons and their preferred orientations.
+\subsection*{Summary of findings}
+In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and 'sparse partial correlations with latent units' or $C_{\sf sparse+lowrank}$.
+First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators.
+We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex.
+We found that $C_{\sf sparse+lowrank}$ consistently dominated the other estimators.
+Estimate produced by $C_{\sf sparse+lowrank}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.
 
@@
 \subsection*{Functional connectivity from big neural data}
-Breakthroughs in neuroscience
+Transformational breakthroughs in neuroscience are anticipated to emerge from technological developments that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far.
 

### 2013-12-16 plos.tex
Rev23.  Incorporated some of Alex' comments.

@@
 The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.
-Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies present in the system.
+Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies governing the system.
 
@@
 Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward other plausible low-dimensional approximations.
-Furthermore, we found consistent relationships between the organization of sparse interactions inferred by the estimator and the physical distances and preferred orientation differences of pairs of cells.
+Furthermore, we found consistent relationships between the organization of sparse interactions inferred by this estimator and the physical distances and preferred orientation differences of pairs of cells.
 Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with difference in preferred orientation whereas negative interactions were less selective.
@@
 Correlations of the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.
-Estimation of correlation matrices can be improved by imposing some kind of structure on the estimate with greatest improvement attained when the imposed structure closely matches the real dependencies in the data.
+Estimation of correlation matrices can be improved by imposing some kind of structure on the estimate with greatest improvement attained when the imposed structure efficiently captures real dependencies in the data.
 Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the cross-validated performance of correlation matrix estimators that imposed different kinds of structure.
@@
 Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.
-For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}.
+For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Wilke:2002,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}.
 In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.
-Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
+Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
 
@@
 Low-rank approximations are particularly suitable for capturing shared fluctuations across the recorded population.
-Extracted through principal component or factor analysis, low-rank components of neural correlations have been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2006,Yu:2009}.
+Extracted through principal component or factor analysis, low-rank components of neural correlations have been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.
 
@@
 Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been used to infer interactions between components in complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
-In \emph{elliptical distributions} (those of the form $\frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is a normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies.
+In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies.
 A multivariate normal distribution specified through its graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
 With increasing departure from linearity, the correspondence between conditional dependence and partial correlations diminishes or completely breaks down \cite{Loh:2012}.
-For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations.
+For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006,Tkacik:2006}.
 
@@
 
-Given a covariance matrix estimate $C$, the correlation matrix $R$ and partial correlation matrix $P$ are computed as
+Given a covariance matrix estimate $C$ (not necessarily sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
 \begin{equation}
@@
 \end{equation}
+Similarly, the matrix of partial correlations $P$ is computed as the negative normalized version of the \emph{precision matrix} (inverse covariance matrix) $C^{-1}$:
 \begin{equation}
@@
 Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix.
-$R$ contains 1s on its diagonal whereas $P$ contains $-1$s on the diagonal.
-The off-diagonal elements of $P$ contain the partial correlations between corresponding pairs of neurons.
 
@@
 Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.
-The usual estimate $C_{\sf 0}$ has the advantage of being unbiased ($\E{C_0}=\Sigma$) but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in the data.
+The usual estimate $C_{\sf 0}$ has the advantage of being unbiased ($\E{C_0}=\Sigma$) but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in training sample.
 Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from truth $\Sigma$.
@@
 
-Figure \ref{fig:01} illustrates the challenges of estimating the correlation matrix from a finite sample and one example of regularization.
-Panel A depicts the correlation matrix computed from the sample covariance matrix $C_{\sf 0}$ for a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex.
-To show detail, panel B magnifies the fragment outlined in panel A.
-Panel C shows the same fragment from the partial correlation matrix also computed from $C_{\sf 0}$.
+To illustrate challenges of estimating the correlation matrix from a finite sample, we briefly consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
+In covariance selection, the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero.
+We estimated the covariance matrix of a period of a somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (Fig.\;\ref{fig:01}\,A, B) and covariance selection (Fig.\;\ref{fig:01}\,C).
 Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
-
-Panels \ref{fig:01}\,D and E depict the correlation matrix and the partial correlation matrix computed from a different estimate, which we will refer to here as $C_{\sf sparse}$, which is computed through \emph{covariance selection} \cite{Dempster:1972}.
-In covariance selection, only a subset of partial correlations is fitted while the rest are fixed at zero.
-The resulting estimate closely reproduces the sample correlation matrix (Fig.~\ref{fig:01}\,D), but its partial correlation structure appears substantially different(Fig.~\ref{fig:01}\,E).
-In this particular example  31,501 of the 44,253 (71.2\%) total partial correlations are set to zero.
-This alternative estimate is attractive for a few reasons.
-First, it has fewer free parameters and is therefore less susceptible to sampling noise.
-Additionally, assuming that partial correlations do indeed reflect elements of the functional connectivity, the second estimate also appears more physiologically plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.
+The apparent similarity of the two estimates belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): the regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remainder.
+Estimation by covariance selection is attractive for several reasons: First, it has fewer free parameters and is therefore less susceptible to sampling noise, yielding estimates that are, on average, closer to truth $\Sigma$.
+Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.
 
 \subsection*{Model selection}
-Covariance selection is but one of numerous possible low-dimensional approximations of the covariance matrix.
+However, covariance selection is but one of numerous possible low-dimensional approximations of the covariance matrix.
 Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix.
-Models that accurately reflect the dominant functional dependencies in the circuit, its so-called ``functional connectivity'', will tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators.
+Models that accurately reflect the dominant functional dependencies in the circuit, its so-called `functional connectivity', will tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} than the other estimators or to \emph{dominate} the other estimators.
 An estimator's superior efficiency can be ascertained by cross-validation.
@@
 \subsection*{Summary of findings}
-In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and 'sparse partial correlations with latent units' or $C_{\sf sparse+lowrank}$.
+In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and `sparse partial correlations with latent units' or $C_{\sf sparse+latent}$.
 First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators.
 We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex.
-We found that $C_{\sf sparse+lowrank}$ consistently dominated the other estimators.
-Estimate produced by $C_{\sf sparse+lowrank}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.
+We found that $C_{\sf sparse+latent}$ consistently dominated the other estimators.
+Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.
 
@@
 
-Finally, we consider the fourth estimator $C_{\sf sparse+lowrank}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D).
+Finally, we consider the fourth estimator $C_{\sf sparse+latent}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D).
 This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}
@@
 In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (Figure \ref{fig:02}, Row 5).
-For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth: $\ell_{C,C_{\sf diag}}$, $\ell_{C,C_{\sf factor}}$, $\ell_{C,C_{\sf sparse}}$, and $\ell_{C,C_{\sf sparse+lowrank}}$.
+For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth: $\ell_{C,C_{\sf diag}}$, $\ell_{C,C_{\sf factor}}$, $\ell_{C,C_{\sf sparse}}$, and $\ell_{C,C_{\sf sparse+latent}}$.
 Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples  taken for each ground truth and for each sample size.
@@
 
-In these highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+lowrank}$) would be more efficient than others.
+In these highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would be more efficient than others.
 At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), to the advantage of estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$).
@@
 
-We found that estimator $C_{\sf sparse+lowrank}$ was more efficient than the other estimators (Fig.~\ref{fig:04}).
-The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+lowrank}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).
+We found that estimator $C_{\sf sparse+latent}$ was more efficient than the other estimators (Fig.~\ref{fig:04}).
+The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).
 
 \subsection*{Relationship of correlation structure to circuit architecture}
-Having demonstrated that $C_{\sf sparse+lowrank}$ dominated the other estimators, we examined the elements of the low-dimensional structure revealed by this estimator for individual imaged sites.
-By its design, the sparse+lowrank estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
+Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the elements of the low-dimensional structure revealed by this estimator for individual imaged sites.
+By its design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
 If this estimate dominates all other estimators evaluated thus far, it seems reasonable to hypothesize that these interaction may reflect underlying physiological interactions more closely than the usual correlations.
@@
 
-We examined the structure of the sparse+lowrank estimate in individual sites such as the example site first depicted in Fig.\;\ref{fig:03}.
+We examined the structure of the sparse+latent estimate in individual sites such as the example site first depicted in Fig.\;\ref{fig:03}.
 In this site, as in others, the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (Fig.~\ref{fig:05}\,A and D).
@@
 
-In previous publications that looked at the structure of correlations, the sparse network was obtained by thresholding the correlations coefficients  above some level deemed significant and examining the network of correlations above that threshold \cite{Golshani:2009,Malmersjo:2013}.
-There was fairly little overlap between the network of interactions revealed by this thresholding method and those revealed by the sparse+lowrank estimator (Fig.\;\ref{fig:05}\,F).
+In previous studies of the structure of correlations, a sparse network was produced by thresholding the correlations coefficients  above some level deemed significant and examining the network of correlations above that threshold \cite{Golshani:2009,Malmersjo:2013}.
+There was fairly little overlap between the network of interactions revealed by this thresholding method and those revealed by the sparse+latent estimator (Fig.\;\ref{fig:05}\,F).
 When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (Fig.\;\ref{fig:05}\,F, H, and I).
@@
     \end{center}
-    \caption{{\bf The sparse+lowrank estimator $C_{\sf sparse+lowrank}$ outperforms the other estimators on neural data.}
-    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+lowrank}$. 
+    \caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
+    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
     The histograms are based on 31 imaged sites in 24 mice. 
-    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+lowrank}$ over the other estimators. 
+    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
     The green arrows indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}

### 2013-12-16 plos.tex
unfinished edits of the Discussion section.

@@
 However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.
-Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including  direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
+Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including  direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b,Moore:1970,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
 
@@
 
-Given a covariance matrix estimate $C$ (not necessarily sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
+Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
 \begin{equation}
@@
 Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix.
+Clearly, off-diagonal zeros in the precision matrix correspond to zero partial correlations.
 
@@
 The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations are the result of direct linear interactions between pairs of observed cells and that such interactions occur only between a fraction of such pairs (Fig.~\ref{fig:02}\,C).
-This assumption is enforced by reducing to zero the majority of pairwise \emph{partial correlations} in the recorded population.
-While usual correlations are calculated from the marginal distribution without conditioning on all the other neurons, the partial correlation expresses the pairwise correlation conditioned on the activity of all the other cells.
-If cells only exert linear effects on each other and all neurons are recorded, partial correlations express the direct interactions between neurons.
-Therefore, this estimator is biased toward the assumption that correlations arise due to interactions between a subset of pairs of recorded neurons (Fig.~\ref{fig:03}C).
-When the partial correlation between a pair of neurons is zero, then the corresponding element of the inverse of the covariance matrix (often referred to as the \emph{precision matrix} or \emph{concentration matrix}) must be zero as well.
+This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.
+Therefore, $C_{\sf sparse}$ is biased toward the assumption that correlations arise due to linear interactions between pairs of recorded neurons (Fig.~\ref{fig:03}C).
 Then the estimator is
@@
 Finally, we consider the fourth estimator $C_{\sf sparse+latent}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D).
-This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}
+This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
 \begin{equation}
@@
 In this study, we adopted \emph{negative normal log-likelihood loss}\footnote{
-The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$: 
+The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$ as
 \begin{equation*}
@@
 \subsection*{Functional connectivity from big neural data}
-Transformational breakthroughs in neuroscience are anticipated to emerge from technological developments that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far.
+Transformational breakthroughs in neuroscience are anticipated to follow technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far.
+It far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.
+Most common analyses of the functional connectivity rely on analyzing pairwise correlation coefficients, t-tests, ANCOVA, principal component analysis of correlated activity, and cross-correlations.
+Traditionally, functional connectivity between regions has been defined as the level of correlation or synchronization between regions or cells.
+Without targeted stimulation and intracellular signals, synaptic connectivity could be inferred with some degree of confidence from the presence of sharp peaks or troughs in the cross-correlograms at offsets corresponding to typical synaptic transmission delays.
+However, even such effects can be explained by multiple effects besides monosynaptic connections while some strong synaptic connections do not necessarily produce a clear spike in the cross-correlogram.
+The situations is a bit more complicated in recording modalities, such as calcium imaging, in which only firing rates at lower temporal resolutions can be inferred and cross-correlograms with millisecond precision cannot be obtained.
+Therefore, some researchers have resorted to using simple correlations as proxies of functional connectivity.
+Yet, correlations are the end product of multiple cell interactions and, which do not always correspond to
 
+\subsection*{Comparison with other uses of neural correlations}
 \subsection*{Meaning of interactions}

### 2013-12-17 plos.tex
minor

@@
 The situations is a bit more complicated in recording modalities, such as calcium imaging, in which only firing rates at lower temporal resolutions can be inferred and cross-correlograms with millisecond precision cannot be obtained.
-Therefore, some researchers have resorted to using simple correlations as proxies of functional connectivity.
+Several studies have used sample correlations to infer graphs of functional connectivities.
 Yet, correlations are the end product of multiple cell interactions and, which do not always correspond to
+
+Increasingly, functional connectivity is inferred by fitting probabilistic models such as Generalized Linear Models \cite{Pillow:2008} or Ising Models \cite{Schneidman:2006,Tkacik:2006,Tang:2008,Shlens:2009}.
+In probabilistic models, functional dependencies correspond to terms in the model.
+If two neurons do not participate in single term in the model, they are conditionally independent: the activity of one of the two neurons can be predicted from other neurons just as well with or without the second neuron.
 

### 2013-12-17 plos.tex
further progress on Discussion.

--- 2013-12-16 plos.tex
@@
 For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Wilke:2002,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}.
-In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.
-Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
+In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.
+Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Arieli:1995,Chiu:2002,Kenet:2003,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
 
@@
 Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
-The apparent similarity of the two estimates belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): the regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remainder.
+The apparent similarity of the two estimates belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): the regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
 Estimation by covariance selection is attractive for several reasons: First, it has fewer free parameters and is therefore less susceptible to sampling noise, yielding estimates that are, on average, closer to truth $\Sigma$.
@@
 \section*{Discussion}
+\subsection*{Neural correlations as indicators of functional connectivity}
+Neural correlations have a rich history in neuroscience as indicators of \emph{functional connectivity}.
+When measured on temporal scales of a few milliseconds, peaks in the cross-correlations between pairs of neurons at latencies typical of synaptic transmission are often interpreted as evidence of direct synaptic connections between pairs of cells \cite{Moore:1970,Alonso:1998,Denman:2013}.
+Without targeted stimulations of specific cells or measurements of intracellular signals, such analysis remains the best test for probable monosynaptic connections in vivo.
+At longer temporal scales, neural correlations reflect activations involving large ensembles of neurons.
+Such correlations may have emergent structure with little  correspondence to synaptic connectivity, which  may better reflect computations performed by the circuit.
+For this reason, firing rate correlations at scales of tens or hundreds milliseconds have been of interest for their implications on coding under the assumption of rate codes and linear downstream decoder \cite{Averbeck:2006}.
+The terms \emph{functional connectivity} is still often applied to slow correlations, similarly to its use in PET, fMRI, EEG, and MEG brain imaging studies, as an expression of task-related, stimulus-evoked, or spontaneous synchronization or correlation between brain regions, circuits, or cells without implying a mechanistic interpretation.
+
+Calcium imaging allows simultaneous measurements of the activity of large populations of cells, even of entire nervous systems \cite{Leung:2013,Ahrens:2013}, with cell-wise resolution, but its temporal bandwidth is limited to the scale of tens or hundreds milliseconds by the kinetics of calcium-sensitive dyes.
+This limits their use to the analysis of slower correlations.
+
+\subsection*{Correlation `structure'}
+The term `structure' implies a reduced, simplified representation expressing the essential organization.
+The structure of slow correlations is often expressed through low-rank approximations extracted by principal component analysis or factor analysis.
+These components show synchronous multineuronal activation patterns without revealing graphs of interactions between neurons.
+
+Another reduced representation of correlation matrices is obtained by sparsifying its inverse, resulting in a graph of partial correlations.
+Under the assumption of a multivariate normal distribution or, more generally, an elliptical distribution, sparse partial correlations suggest a sparse matrix of dependencies, a probabilistic graphical model.
+This property has led some investigators to suggest that, in brain imaging studies, such approximation must serve as better proxies of direct interactions between brain regions than correlations \cite{Varoquaux:2012,Ryali:2012}.
+
+\subsection*{Interpretation of functional connectivity from population calcium signals}
+Nonetheless, some investigators have examined graphs of the most significant correlations in population calcium activity as proxy of connectivity \cite{Malmersjo:2013}.
+
+While synaptic connectivity constitutes a directed weighted graph, most meaningful representation of functional organization at longer temporal scales are less certain.
+A common approach has been to examine the low-rank structure of the joint population activity extracted through principal component analysis, factor analysis, independent component analysis, linear embedding, and others.
+Functional connectivity is then represented as the eigenvectors, principle components, or factor loadings corresponding to the extracted components of the joint activity.
+
+If only most significant components are retained, these two representations turn into, respectively, low-rank approximations and sparse inverse approximations.
+Low-rank representation suggest interpretation as common input onto large subsets of neurons or highly synchronous activations in the entire recorded population.
+
+We sought to uncover the optimal low-dimensional approximation of the covariance matrix for efficient estimation of the true covariance matrices in cortical microcircuits on the temporal scale of 100--200 ms corresponding to the temporal resolution of spike rates from somatic calcium signals \cite{Cotton:2013}.
+We considered sparse-inverse approximations, low-rank approximations, and mixtures of low-rank components with sparse partial correlations.
+We used cross-validation to determine which of such representations was closest to the true correlation matrix.
+As we illustrated by simulation, none of the estimators was intrinsically more efficient than the others and their relative performance was dependent on the structure of dependencies in the empirical data.
+We found that sparse inverse approximation performed better than low-rank approximations (Fig.\;S1\;E) and approximations by combinations of sparse and low-rank components consistently outperformed the other estimators.
+This finding is specific to the circuit we examined and may not generalize to other circuits or to less dense sampling of cells in electrophysiological recordings.
+
+\subsection*{Probabilistic models and their interpretation}
+On the millisecond scale, most common analyses of the functional connectivity have relied on cross-correlation analysis in attempts to detect signatures of direct synaptic connectivity.
+
+Increasingly, the functional organization of neural circuits is inferred by fitting probabilistic models such as various forms of multivariate Gaussian models, linear-nonlinear models, Generalized Linear Models \cite{Pillow:2008}, and maximum entropy models \cite{Schneidman:2006,Tkacik:2006,Tang:2008,Shlens:2009}.
+In probabilistic models, functional dependencies correspond to multiplicative terms involving pairs or larger groups of neurons.
+If two neurons do not both participate in any of the multiplicative terms in the model, they are conditionally independent: the activity of one can be predicted from other neurons just equally well with or without considering the second neuron, suggesting a lack of direct interaction between them.
+Additionally, various probabilistic models allow inferring latent variables to explain joint patterns of in large groups of cells.
+The uncovered networks of dependencies from such statistical models may arise from mechanistic interactions such as direct monosynaptic connections or from complex emergent processes involving large ensembles of cells.
+Because statistical models of the joint activity of populations of neurons use the activity of the entire population to infer functional dependencies, they to produce more accurate and relevant representations of the functional organization of the data than available from equivalent numbers of paired recordings.
+
+However, no empirical evidence yet exists to demonstrate that statistical models of population activity reveal networks of dependences more informative than the conventional correlations with respect to the structural architecture of specific neural circuits.
+
+\subsection*{Model selection to uncover structure best supported by data}
+Model selection from various several probabilistic models
+
 \subsection*{Functional connectivity from big neural data}
-Transformational breakthroughs in neuroscience are anticipated to follow technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far.
-It far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.
-Most common analyses of the functional connectivity rely on analyzing pairwise correlation coefficients, t-tests, ANCOVA, principal component analysis of correlated activity, and cross-correlations.
-Traditionally, functional connectivity between regions has been defined as the level of correlation or synchronization between regions or cells.
-Without targeted stimulation and intracellular signals, synaptic connectivity could be inferred with some degree of confidence from the presence of sharp peaks or troughs in the cross-correlograms at offsets corresponding to typical synaptic transmission delays.
-However, even such effects can be explained by multiple effects besides monosynaptic connections while some strong synaptic connections do not necessarily produce a clear spike in the cross-correlogram.
-The situations is a bit more complicated in recording modalities, such as calcium imaging, in which only firing rates at lower temporal resolutions can be inferred and cross-correlograms with millisecond precision cannot be obtained.
-Therefore, some researchers have resorted to using simple correlations as proxies of functional connectivity.
-Yet, correlations are the end product of multiple cell interactions and, which do not always correspond to
+Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}.
+It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.
+
+Such probabilistic models of population activity feature large numbers of parameters, growing quadratically or faster with population size.
+Estimation of large number of parameters is sensitive to sampling noise.
+To solve this problem, the recording durations must increase at least proportionally to the number of parameters or the model must be constrained by regularization.
 
@@
 \subsection*{Meaning of interactions}
-\subsection*{Fitting probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
+\subsection*{Fitting and evaluation of probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
 \subsection*{Future directions}

### 2013-12-18 plos.tex
edited discussion

@@
 \subsection*{Neural correlations as indicators of functional connectivity}
-Neural correlations have a rich history in neuroscience as indicators of \emph{functional connectivity}.
-When measured on temporal scales of a few milliseconds, peaks in the cross-correlations between pairs of neurons at latencies typical of synaptic transmission are often interpreted as evidence of direct synaptic connections between pairs of cells \cite{Moore:1970,Alonso:1998,Denman:2013}.
-Without targeted stimulations of specific cells or measurements of intracellular signals, such analysis remains the best test for probable monosynaptic connections in vivo.
+Neural correlations boast a rich history in systems neuroscience as indicators of \emph{functional connectivity}.
+When measured on temporal scales of a milliseconds, peaks in the cross-correlogram between pairs of neurons at latencies typical of synaptic transmission may be interpreted as evidence of monosynaptic connection between pairs of cells \cite{Moore:1970,Alonso:1998,Denman:2013}.
+Without targeted stimulation of specific cells or measurement of intracellular signals, such analysis remains the best test for probable monosynaptic connections in vivo.
 At longer temporal scales, neural correlations reflect activations involving large ensembles of neurons.
@@
 \subsection*{Correlation `structure'}
-The term `structure' implies a reduced, simplified representation expressing the essential organization.
-The structure of slow correlations is often expressed through low-rank approximations extracted by principal component analysis or factor analysis.
+The term `structure' refers to a succinct, simplified representation that captures the essential features.
+Many questions in systems neuroscience require deducing the structure of the neural correlation structure in order to relate it to the spatial organization of the circuit, track its changes due to learning and development, or to describe effects of brain states, etc.
+Perhaps the least sophisticated approach for extracting the structure of a correlation matrix is to examine the network of its most significant or largest coefficients.
+For example, \cite{Malmersjo:2013} analyzed the graph theoretical properties of networks of thresholded correlations in development.
+Alternatively, the correlation structure is often expressed through low-rank approximations extracted by principal component analysis or factor analysis.
 These components show synchronous multineuronal activation patterns without revealing graphs of interactions between neurons.
+Another reduced representation of correlation matrices is obtained by sparsifying its inverse, resulting in a graph of partial correlations.
+Under the assumption of a multivariate normal distribution or, more generally, an elliptical distribution, sparse partial correlations suggest a sparse network of dependencies, \emph{i.e.}\;a probabilistic graphical model.
+This property has led some investigators to propose that such approximations  must  serve as closer estimates of direct interactions between brain regions in imaging studies than correlations \cite{Varoquaux:2012,Ryali:2012}.
+However, since deviations from elliptical distribution disrupt the correspondence between partial correlations and linear dependencies \cite{Loh:2012}, it is not clear whether that would indeed be the case.
+Partial correlations are meaningful when most interacting variables are observed.
+Perhaps for this reason, until now, parptial correlations had not been considered in analysis of multineuronal recordings.
 
-Another reduced representation of correlation matrices is obtained by sparsifying its inverse, resulting in a graph of partial correlations.
-Under the assumption of a multivariate normal distribution or, more generally, an elliptical distribution, sparse partial correlations suggest a sparse matrix of dependencies, a probabilistic graphical model.
-This property has led some investigators to suggest that, in brain imaging studies, such approximation must serve as better proxies of direct interactions between brain regions than correlations \cite{Varoquaux:2012,Ryali:2012}.
+We found that, for slow correlations (150 ms bins) in cortical microcircuits, networks of partial correlations were more statistically efficient than low-rank approximations of correlation matrices for cortical micr(Fig.\;S1 E), suggesting that they better reflected the correlation structure.
+Furthermore, estimation was made even more efficient when low-rank components were allowed to be combined with partial correlations.
+The correlation structure depended more strongly on distance between cells and differences in their
 
-\subsection*{Interpretation of functional connectivity from population calcium signals}
-Nonetheless, some investigators have examined graphs of the most significant correlations in population calcium activity as proxy of connectivity \cite{Malmersjo:2013}.
-
+\subsection*{Inference of functional connectivity with probabilistic models}
 While synaptic connectivity constitutes a directed weighted graph, most meaningful representation of functional organization at longer temporal scales are less certain.
@@
 Functional connectivity is then represented as the eigenvectors, principle components, or factor loadings corresponding to the extracted components of the joint activity.
-
-If only most significant components are retained, these two representations turn into, respectively, low-rank approximations and sparse inverse approximations.
-Low-rank representation suggest interpretation as common input onto large subsets of neurons or highly synchronous activations in the entire recorded population.
 

### 2013-12-19 plos.tex
nearly completed Discussion.

--- 2013-12-17 plos.tex
@@
 \usepackage{color} 
-
+\usepackage[CaptionAfterwards]{fltpage}
+\usepackage{floatrow}
 % Use doublespacing - comment out for single spacing
@@
 where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates.
+In the case of noise correlations, $\mu$ depends on the stimulus condition.
 The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
@@
 
+\subsection*{Illustration of estimation of a neural correlation matrix}
 To illustrate challenges of estimating the correlation matrix from a finite sample, we briefly consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
@@
 Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
-The apparent similarity of the two estimates belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): the regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remainder.
+The close similarity of the two estimates (Fig.\;\ref{fig:01}\,B and C) belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): the regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
 Estimation by covariance selection is attractive for several reasons: First, it has fewer free parameters and is therefore less susceptible to sampling noise, yielding estimates that are, on average, closer to truth $\Sigma$.
@@
 
+\begin{figure}[!ht]
+    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
+        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
+        The outlined square fragment is magnified in {\bf B}.
+        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
+        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
+    }
+    \label{fig:01}}
+    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
+\end{figure}
+
 \subsection*{Model selection}
-However, covariance selection is but one of numerous possible low-dimensional approximations of the covariance matrix.
+Covariance selection is but one of many conceivable low-dimensional approximations of the covariance matrix.
 Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix.
@@
 The estimator that is shown to dominate all others for a specific type of neural circuit may be proposed as the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
-Its structure may then be analyzed and related to the circuits anatomical organization.
+Its structure may then be analyzed and related to the circuit's anatomical organization.
 
@@
 We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (Fig.~\ref{fig:02}\;row 1).
+
+\begin{FPfigure}
+    \begin{center}
+        \includegraphics{./figures/Figure02.pdf}
+    \end{center}
+    \caption{{\bf Estimators whose low-dimensional regularization targets can represent the structure of the true covariance matrix outperform other estimators.}
+        {\bf Row 1.} Graphical representations of four types of low-dimensional structures of interactions between observed neurons (green spheres) and latent units (light-shaded spheres).
+        In the `independent model' ({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units. 
+        In `latent factors' ({\bf B}), the correlated activity of observed cells is driven by several latent units. 
+        In `sparse interactions' ({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
+        In `sparse+latent' ({\bf D}), correlations arise through direct linear interactions between some pairs of observed neurons and through interactions with common latent units. 
+        {\bf Row 2.} Examples of $50\times 50$ correlation matrices corresponding to each type of low-dimensional structure. 
+        The factor model ({\bf B}) has three latent units. 
+        The partial correlation matrix of the sparse model ({\bf C}) is 73\% sparse.
+        The `sparse+latent' matrix has one latent unit and its direct interactions are 78\% sparse.
+        {\bf Row 3.} Examples of sample correlation matrices calculated from samples of 1000 observations taken from simulated random processes with corresponding correlation matrices from row 2.
+        {\bf Row 4.} Mean \emph{excess loss} (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators as a function of sample size.
+The error bars indicate the standard error of the mean based on 30 samples.
+        {\bf Row 5.} Mean \emph{cross-validation loss} (Eq.~\ref{eq:rel-cv-loss}) of covariance estimators with respect to the matching estimator .
+The values are relative to the validation loss of the estimator that matches the low-dimensional structure of the true covariance matrix.
+The error bars indicate the standard error of the mean based on 30 samples.
+    }
+    \label{fig:02}
+\end{FPfigure}
 
@@
 This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.
-Therefore, $C_{\sf sparse}$ is biased toward the assumption that correlations arise due to linear interactions between pairs of recorded neurons (Fig.~\ref{fig:03}C).
+Therefore, $C_{\sf sparse}$ is biased toward the assumption that correlations arise due to linear interactions between pairs of recorded neurons (Fig.~\ref{fig:02}\,C).
 Then the estimator is
@@
 
+\begin{figure}
+    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+    {\caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
+    Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) were presented to anesthetized mice while two-photon recordings of somatic calcium signals were collected using fast 3D random-access microscopy ({\bf B}).
+The visual stimulus included an initial period with 16 directions of motion for orientation tuning followed by a longer (15--20 min) period of stimulation with only 2 or 5 directions of motion for the computation of the noise correlation matrix.
+    {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. The inferred firing rate binned in 150 ms intervals are indicated by red ticks below each trace.
+    {\bf D.} The raster plot of the inferred firing rates, binned in 150 ms intervals, from the entire population from the first (left) and last (right) minute of the entire recording.
+The traces from {\bf C} are highlighted in red.
+    {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site.
+    {\bf F.} The noise correlation matrix of the activity of the neural population. 
+    {\bf G.} Histogram of noise correlation coefficients. The red line indicates the mean.
+} \label{fig:03}}
+    {\includegraphics[width=8.3cm]{figures/Figure03.pdf}}
+\end{figure}
+
 In these highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would be more efficient than others.
@@
 
-\subsection*{Relationship of correlation structure to circuit architecture}
+\begin{figure}[!ht]
+    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+    {\caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
+    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
+    The histograms are based on 31 imaged sites in 24 mice. 
+    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
+    The green arrows indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
+    } \label{fig:04}}
+    {\includegraphics{./figures/Figure04.pdf}}
+\end{figure}
+
+\subsection*{Correlation structure and circuit architecture}
 Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the elements of the low-dimensional structure revealed by this estimator for individual imaged sites.
@@
 The low-rank component was of rank 17.
+
+\begin{FPfigure}
+    \begin{center}
+        \includegraphics[width=17.35cm]{./figures/Figure05.pdf}
+    \end{center}
+    \caption{{\bf Example of low-dimensional correlation structure revealed by the sparse+low-rank estimator.}
+    {\bf A.} The regularized estimate of the correlation matrix (top-right) closely approximates the sample correlation matrix (bottom left). 
+    This close approximation is  also demonstrated by the scatter plot of the correlation coefficients produced by the two estimates ({\bf D}). 
+    However, the partial correlation matrices from the two estimate show more pronounced differences ({\bf B} and {\bf E}). 
+    {\bf C.} Furthermore, the partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.2\% off-diagonal zeros (bottom-left) and low-rank component of rank 15 (top-right).
+    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations.
+The gray interval indicates the range of correlations containing 82.2\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
+This interval contained 58.9\% of the partial correlations.
+    {\bf G.} The graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
+The line density is proportional to the magnitude of the correlation.
+    {\bf H.} A subset of neurons from the center of the cluster shown in {\bf G} showing the regularized partial correlations.
+    {\bf I.} The same subset with sample correlations thresholded to match the sparsity of the regularized interactions.
+}
+\label{fig:05}
+\end{FPfigure}
 
@@
 
+\begin{figure}[!ht]
+    \begin{center}
+        \includegraphics{./figures/Figure06.pdf}
+    \end{center}
+    \caption{{\bf Properties of sparse+low-rank regularized estimates from all imaged sites}
+    {\bf A.} The average sample correlations vs.~average partial correlations for each imaged site. In each plot, the red asterisk indicates the site shown in figures \ref{fig:03} and \ref{fig:05}.
+    {\bf B.} The average node degree for sparse partial correlations vs.~population size in each imaged site. 
+    {\bf C.} The number of inferred latent units vs.~population size in each imaged site.
+    {\bf D.} The number latent units vs.~average node degree for sparse partial correlations for each site.
+}
+\label{fig:06}
+\end{figure}
+
 We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells  to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}).
-The average partial correlations fell more rapidly with with difference in preferred orientation (Fig.\;\ref{fig:07}\,A and D) and lateral displacements at equal depths ($\pm 25\mu$m) (Fig.\;\ref{fig:07}\,B and E), and differences in depth at small ($\pm 25\mu$m) lateral displacements (Fig.\;\ref{fig:07}\,B and E).
+The average partial correlations fell more rapidly with with difference in preferred orientation (Fig.\;\ref{fig:07}\,A and D) and lateral displacements at equal depths ($\pm 25\;\mu$m) (Fig.\;\ref{fig:07}\,B and E), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (Fig.\;\ref{fig:07}\,B and E).
+
+\begin{figure}[!ht]
+    \begin{center}
+        \includegraphics{./figures/Figure07.pdf}
+    \end{center}
+    \caption{{\bf Dependence of correlations and partial correlations on orientation tuning differences and physical distance between cell pairs.}
+    {\bf A--C.} Average partial correlations (red) estimated by $C_{\sf sparse+latent}$ and average sample correlations (black) averaged across multiple imaged sites.
+In each site, the correlations were normalized by the respective average correlation shown in Fig.\;\ref{fig:06}\,A.
+The number $n$ of sites that qualified to be included in the analysis is indicated.
+Sites were included if they had at least 20 pairs of neurons in each of the intervals.
+The error bars indicate the standard error of the mean for based on $n$.
+    {\bf A.} Average correlations between pairs of neurons tuned to orientation with differences in preferred orientation in the intervals of 0--15$^\circ$, 15--45$^\circ$ and 45--90$^\circ$. 
+    {\bf B.} Average correlations between pairs of neurons located at the same depth ($\pm$25$\mu$m) separated by lateral distances in the intervals of 0--25 $\mu$m, 25--75 $\mu$m, 75--150 $\mu$m, and 150+ $\mu$m.
+    {\bf C.} Average correlations between pairs of neurons displaced laterally by less than 25 $\mu$m separated in depth by distances in the intervals of 0-25 $\mu$m, 25--60 $\mu$m, and 60+ $\mu$m.
+    {\bf D--F.} Normalized connectivity of positive (green) and negative (dark red) interactions from the sparse component obtained from $C_{\sf sparse+latent}$.
+Normalized connectivity was computed as the fraction of pairs connected by interactions of corresponding signs in each interval divided by the fraction of non-zero interactions across the entire site.
+Site were included in the analysis only if they had at least 20 cell pairs in each interval.
+The error bars indicate the standard error of the mean based on the number $n$ of sites included in the analysis.
+}
+\label{fig:07}
+\end{figure}
 
@@
 \section*{Discussion}
-\subsection*{Functional connectivity from big neural data}
-Transformational breakthroughs in neuroscience are anticipated to follow technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far.
-It far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.
-Most common analyses of the functional connectivity rely on analyzing pairwise correlation coefficients, t-tests, ANCOVA, principal component analysis of correlated activity, and cross-correlations.
-Traditionally, functional connectivity between regions has been defined as the level of correlation or synchronization between regions or cells.
-Without targeted stimulation and intracellular signals, synaptic connectivity could be inferred with some degree of confidence from the presence of sharp peaks or troughs in the cross-correlograms at offsets corresponding to typical synaptic transmission delays.
-However, even such effects can be explained by multiple effects besides monosynaptic connections while some strong synaptic connections do not necessarily produce a clear spike in the cross-correlogram.
-The situations is a bit more complicated in recording modalities, such as calcium imaging, in which only firing rates at lower temporal resolutions can be inferred and cross-correlograms with millisecond precision cannot be obtained.
-Several studies have used sample correlations to infer graphs of functional connectivities.
-Yet, correlations are the end product of multiple cell interactions and, which do not always correspond to
+\subsection*{Functional and anatomical connectivity}
+Systems neuroscience is the quest to discover principles whereby anatomical organization of neural circuits gives rise to their function.
+This line of inquiry requires descriptions of the activity of neural circuits that have the best chance of correspondence to the underlying anatomical organization.
 
-Increasingly, functional connectivity is inferred by fitting probabilistic models such as Generalized Linear Models \cite{Pillow:2008} or Ising Models \cite{Schneidman:2006,Tkacik:2006,Tang:2008,Shlens:2009}.
-In probabilistic models, functional dependencies correspond to terms in the model.
-If two neurons do not participate in single term in the model, they are conditionally independent: the activity of one of the two neurons can be predicted from other neurons just as well with or without the second neuron.
+The term \emph{functional architecture} most commonly refers to stimulus response properties of neurons \cite{Reid:2012}.
+Over the past 50 years, tremendous progress has been made in relating the functional architecture to the anatomical, synaptic wiring of neuronal populations and individual neurons.
+This line of inquiry is based on the view of the brain as a complex filter of external stimuli.
 
-\subsection*{Comparison with other uses of neural correlations}
-\subsection*{Meaning of interactions}
-\subsection*{Fitting probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
+Although there these concepts overlap, \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or variability in stimulus responses,  not directly driven by external stimuli, and reflecting internal, recurrent connectivity in neural circuits.
+This line of inquiry ascribes greater significance to the brain's intrinsic activity (often dismissed as `noise') \cite{Yuste:2005}.
+Relatively little progress has been made in understanding the role of functional connectivity and its relationship to the anatomical connectivity.
+
+Neural correlations boast a rich history as indicators of functional connectivity.
+When measured on temporal scales of milliseconds, sharp peaks or troughs in the cross-correlogram between pairs of neurons at latencies typical of synaptic transmission provide evidence of monosynaptic connection between pairs of cells \cite{Moore:1970,Alonso:1998,Denman:2013} (although other processes can also generate sharp peaks in the cross-correlogram while many synaptic connections may not produce such signatures).
+Without targeted stimulation of specific cells or measurement of intracellular signals, such analysis remains the best method for detecting probable monosynaptic connections in vivo.
+When measured at longer temporal scales, neural correlations reflect activations involving large ensembles of neurons and may reflect direct causal anatomical interactions, indirect causal interactions, or complex network-wide emergent properties with little correspondence to anatomical connectivity.
+These slower correlations correspond more closely to the computation performed by the circuit than to its anatomical organization.
+For this reason, firing rate correlations at scales of tens to hundreds milliseconds have been studied for their effect on coding under the assumption of rate codes and linear downstream decoder \cite{Averbeck:2006} with little interest in anatomy.
+Calcium imaging allows simultaneous measurements of the activity of large populations of cells \emph{in vivo} \cite{Katona:2012,Cotton:2013} and even of entire nervous systems \cite{Leung:2013,Ahrens:2013}, but its temporal bandwidth is limited to the scale of tens or hundreds milliseconds by the kinetics of calcium-sensitive dyes, limiting its use to slower temporal scales at which correspondence between functional and anatomical connectivity is less direct.
+
+Our study is part of the effort to find descriptions of functional connectivity with the greatest chance of correspondence to anatomical, causal organization of the circuit.
+
+\subsection*{Extraction of the correlation structure}
+The word `structure' implies a succinct, simplified representation that captures only the essential, significant features.
+Several forms of correlations structure can be extracted from the same data and used to describe functional connectivity in different ways.
+
+Perhaps the least sophisticated yet common approach for extracting the structure of correlations is to threshold above a selected level of significance.
+Part of the appeal of thresholded correlations is their representation in the form of graphs that can be related to graphs of synaptic connectivity and made amenable to network analysis \cite{Bullmore:2012}.
+For example, \cite{Malmersjo:2013} analyzed the graph theoretical properties of networks of thresholded correlations in development.
+
+Alternatively, the correlation structure may be extracted in the form of low-rank approximations computed by principal component analysis, factor analysis, or activation maps conditioned on the activity of a single neuron.
+Low-rank components show synchronous multineuronal activation patterns without revealing graphs of interactions between neurons.
+They are often interpreted as emergent patterns of large circuits or brain state fluctuations \cite{Kenet:2003}.
+
+Another reduced representation of correlation matrices is obtained by sparsifying its inverse, resulting in a graph of partial correlations.
+Under the assumption of a multivariate normal distribution or, more generally, an elliptical distribution, sparse partial correlations suggest a sparse network of dependencies, \emph{i.e.}\;a probabilistic graphical model.
+This property has led some investigators to propose that sparse inverse approximation of correlations may serve as closer estimates of direct anatomical interactions between brain regions in imaging studies than correlations \cite{Varoquaux:2012,Ryali:2012}.
+However, since deviations from elliptical distribution disrupt the correspondence between partial correlations and linear dependencies \cite{Loh:2012}, it is not clear whether that would indeed be the case.
+Partial correlations are meaningful when most interacting variables are observed.
+Perhaps for this reason, until now, partial correlations had not been considered in analysis of multineuronal recordings.
+
+In this study, we sought to uncover the optimal low-dimensional approximation of the covariance matrix for efficient estimation of the true covariance matrices in cortical microcircuits on the temporal scale of 100--200 ms corresponding to the temporal resolution of spike rates from somatic calcium signals \cite{Cotton:2013}.
+We found that networks of partial correlations were more statistically efficient than low-rank approximations of correlation matrices (Fig.\;S1 E).
+Furthermore, estimation became still more efficient when low-rank components were allowed to be combined with sparse partial correlations.
+We propose that correlation structures that are supported by the data  are more likely to reflect anatomical, causal interactions.
+The network of sparse partial correlations suggests interactions between specific pairs whereas the low-rank represents emergent activity patterns involving collective activations of larger circuits or global fluctuations.
+The revealed network of partial interactions was substantially different from thresholded correlations (Fig.~\ref{fig:05}\,C, F, H, I).
+In particular, it shows that many low correlations correspond to strong negative interactions (Fig.\;\ref{fig:05}\,F).
+Intuitively, this suggests that many low interactions are as surprising or significant considering the common correlations with other neurons.
+The analysis of sample correlations does not detect surprisingly low correlations.
+
+\subsection*{Functional connectivity from probabilistic models}
+Increasingly, functional connectivity is inferred by fitting probabilistic models such as various forms of multivariate Gaussian models, linear-nonlinear models, Generalized Linear Models \cite{Pillow:2008}, and maximum entropy models \cite{Schneidman:2006,Tkacik:2006,Tang:2008,Shlens:2009}.
+In probabilistic models, functional dependencies are defined by multiplicative terms comprising the model.
+If the model can be decomposed as a product of terms so that a pair of variables can be segregated into separate terms, the two variables are conditionally independent: the activity of one neuron can be predicted from other neurons just as well with or without considering the second neuron, suggesting a lack of a direct causal interaction between them.
+Probabilistic models allow inferring latent variables to account for the activity of unobserved circuitry.
+
+\subsection*{Model selection}
+Various estimation methods reveal radically different structures of functional connectivity?
+Which of the structures should be preferred?
+How does one determine which functional connectivity structure is best supported by the empirical data?
+Machine learning literature offers a rich toolbox of model selection criteria.
+Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion}, and \emph{Bayes (or Schwartz) Information Criterion} are computationally efficient but rely on assumptions on generally unsupportable assumptions about the data generating process.
+The most assumption-free yet more computationally intensive are procedures based on cross-validation \cite{Arlot:2010} and, when models have hyperparameters, nested cross-validation to selected the optimal values of the hyperparameters.
+It is commonly assumed that probabilistic models that demonstrate the highest average cross-validated log likelihood yield a better representation of the functional connectivity.
+For example, \cite{Ganmor:2011} show the fact that their proposed Reliable Network Model (a maximum-entropy model on the multivariate binary domain constrained by a sparse network of correlations of various orders) yielded higher cross-validated log-likelihoods than the full-pairwise Ising model.
+The authors use this fact to support the claim that the former model revealed a more anatomically relevant structure of functional connectivity than the pairwise model.
+
+Under the loss function (Eq.~\ref{eq:loss}) employed in our study, the covariance matrix estimation methods are equivalent to fitting multivariate Gaussian models with various structure of dependencies between visible and inferred latent units and selecting between them on the basis of cross-validated log-likelihood.
+However, the ranking of models was preserved under other loss functions (Fig.~\ref{fig:S02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.
+
+\subsection*{Big neural data}
+Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}.
+It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.
+
+Such probabilistic models of population activity feature large numbers of parameters, growing quadratically or faster with population size.
+Estimation of large number of parameters is sensitive to sampling noise.
+To solve this problem, the recording durations must increase at least proportionally to the number of parameters or the model must be constrained by regularization.
+
 \subsection*{Future directions}
@@
 
-\section*{Figure Legends}
-
-\begin{figure}[!ht]
-    \begin{center}
-        %\includegraphics[width=8.3cm]{./figures/Figure01.pdf}
-    \end{center}
-    \caption{{\bf Illustration of regularized estimation of partial correlations.}
-        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
-        The outlined square fragment is magnified in {\bf B}.
-        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
-        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
-    }
-    \label{fig:01}
-\end{figure}
-
-\begin{figure}[!ht]
-    \begin{center}
-        %\includegraphics[width=17.35cm]{Figure02.pdf}
-    \end{center}
-    \caption{{\bf Estimators whose low-dimensional regularization targets can represent the structure of the true covariance matrix outperform other estimators.}
-        {\bf Row 1.} Graphical representations of four types of low-dimensional structures of interactions between observed neurons (green spheres) and latent units (light-shaded spheres).
-        In the `independent model' ({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units. 
-        In `latent factors' ({\bf B}), the correlated activity of observed cells is driven by several latent units. 
-        In `sparse interactions' ({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
-        In `sparse+latent' ({\bf D}), correlations arise through direct linear interactions between some pairs of observed neurons and through interactions with common latent units. 
-        {\bf Row 2.} Examples of $50\times 50$ correlation matrices corresponding to each type of low-dimensional structure. 
-        The factor model ({\bf B}) has three latent units. 
-        The partial correlation matrix of the sparse model ({\bf C}) is 73\% sparse.
-        The `sparse+latent' matrix has one latent unit and its direct interactions are 78\% sparse.
-        {\bf Row 3.} Examples of sample correlation matrices calculated from samples of 1000 observations taken from simulated random processes with corresponding correlation matrices from row 2.
-        {\bf Row 4.} Mean \emph{excess loss} (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators as a function of sample size.
-The error bars indicate the standard error of the mean based on 30 samples.
-        {\bf Row 5.} Mean \emph{cross-validation loss} (Eq.~\ref{eq:rel-cv-loss}) of covariance estimators with respect to the matching estimator .
-The values are relative to the validation loss of the estimator that matches the low-dimensional structure of the true covariance matrix.
-The error bars indicate the standard error of the mean based on 30 samples.
-    }
-    \label{fig:02}
-\end{figure}
-
-\begin{figure}[!ht]
-    \begin{center}
-        %\includegraphics[width=8.3cm]{figures/Figure03.pdf}
-    \end{center}
-    \caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
-    Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) were presented to anesthetized mice while two-photon recordings of somatic calcium signals were collected using fast 3D random-access microscopy ({\bf B}).
-The visual stimulus included an initial period with 16 directions of motion for orientation tuning followed by a longer (15--20 min) period of stimulation with only 2 or 5 directions of motion for the computation of the noise correlation matrix.
-    {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. The inferred firing rate binned in 150 ms intervals are indicated by red ticks below each trace.
-    {\bf D.} The raster plot of the inferred firing rates, binned in 150 ms intervals, from the entire population from the first (left) and last (right) minute of the entire recording.
-The traces from {\bf C} are highlighted in red.
-    {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site.
-    {\bf F.} The noise correlation matrix of the activity of the neural population. 
-    {\bf G.} The histogram of the noise correlation coefficients with the mean indicated by the red line.
-}
-    \label{fig:03}
-\end{figure}
-
-\begin{figure}[!ht]
-    \begin{center}
-        %\includegraphics[width=8.3cm]{Figure04.pdf}
-    \end{center}
-    \caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
-    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
-    The histograms are based on 31 imaged sites in 24 mice. 
-    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
-    The green arrows indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
-    }
-    \label{fig:04}
-\end{figure}
-
-\begin{figure}[!ht]
-    \begin{center}
-        %\includegraphics[width=17.35cm]{Figure05.pdf}
-    \end{center}
-    \caption{{\bf Example of low-dimensional correlation structure revealed by the sparse+low-rank estimator.}
-    {\bf A.} The regularized estimate of the correlation matrix (top-right) closely approximates the sample correlation matrix (bottom left). 
-    This close approximation is  also demonstrated by the scatter plot of the correlation coefficients produced by the two estimates ({\bf D}). 
-    However, the partial correlation matrices from the two estimate show more pronounced differences ({\bf B} and {\bf E}). 
-    {\bf C.} Furthermore, the partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.2\% off-diagonal zeros (bottom-left) and low-rank component of rank 15 (top-right).
-    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations.
-The gray interval indicates the range of correlations containing 82.2\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
-This interval contained 58.9\% of the partial correlations.
-    {\bf G.} The graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
-The line density is proportional to the magnitude of the correlation.
-    {\bf H.} A subset of neurons from the center of the cluster shown in {\bf G} showing the regularized partial correlations.
-    {\bf I.} The same subset with sample correlations thresholded to match the sparsity of the regularized interactions.
-}
-\label{fig:05}
-\end{figure}
-
-\begin{figure}[!ht]
-    \begin{center}
-        %\includegraphics[width=17.35]{Figure06.pdf}
-    \end{center}
-    \caption{{\bf Properties of sparse+low-rank regularized estimates from all imaged sites}
-    {\bf A.} The average sample correlations vs.~average partial correlations for each imaged site. In each plot, the red asterisk indicates the site shown in figures \ref{fig:03} and \ref{fig:05}.
-    {\bf B.} The average node degree for sparse partial correlations vs.~population size in each imaged site. 
-    {\bf C.} The number of inferred latent units vs.~population size in each imaged site.
-    {\bf D.} The number latent units vs.~average node degree for sparse partial correlations for each site.
-}
-\label{fig:06}
-\end{figure}
-
-\begin{figure}[!ht]
-    \begin{center}
-        %\includegraphics[width=17.35]{Figure07.pdf}
-    \end{center}
-    \caption{{\bf Dependence of correlations and partial correlations on orientation tuning differences and physical distance between cell pairs.}
-    {\bf A--C.} Average sample correlations (black) and regularized partial correlations (red) between pairs of cells in the example site shown in previous figures.
-The correlations were normalized by the respective average correlations (\ref{fig:06}\,A).
-    {\bf A.} Average correlations between pairs of neurons tuned to orientation with differences in preferred orientation in the intervals of 0--15$^\circ$, 15--45$^\circ$ and 45--90$^\circ$. 
-    {\bf B.} Average correlations between pairs of neurons located at the same depth ($\pm$25$\mu$m) separated by lateral distances in the intervals of 0--25 $\mu$m, 25--75 $\mu$m, 75--150 $\mu$m, and 150+ $\mu$m.
-    {\bf C.} Average correlations between pairs of neurons displaced laterally by less than 25 $\mu$m separated in depth by distances in the intervals of 0-25 $\mu$m, 25--60 $\mu$m, and 60+ $\mu$m.
-    {\bf D--F.} Same measurements as {\bf A--D} averaged across multiple sites.
-Only sites that had at least 20 qualifying cell pairs in each of the intervals were included in the averages.
-The error bars indicate the standard error of the mean (often too small to be seen).
-    {\bf G--I.} Normalized connectivity of positive (green) and negative (dark red) interactions from the sparse component of the regularized partial correlations in the example site show in previous figures.
-Normalized connectivity was computed as the fraction of pairs connected by interactions of corresponding signs in each interval divided by fraction of non-zero interactions across the entire site.
-The intervals are identical to those in {\bf A--C}.
-    {\bf J--L.} Same measurements as in {\bf G--I} averaged across multiple sites.
-The error bars indicate the error of the mean.
-Only sites that had at least 20 qualifying pairs in each interval were included in the averages.
-}
-\label{fig:07}
-\end{figure}
+%\section*{Figure Legends}
+\section*{Supplementary Figures}
 

### 2013-12-19 plos.tex
Kreso's comments

--- 2013-12-17 plos.tex
@@
 The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.
-Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies governing the system.
+Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies present in the system.
+\Kcomment{Simplify previous sentence.}
 
@@
 We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.
-Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward other plausible low-dimensional approximations.
-Furthermore, we found consistent relationships between the organization of sparse interactions inferred by this estimator and the physical distances and preferred orientation differences of pairs of cells.
+Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward \Kcomment{Will the phrase ``regularized toward'' be understandable in an abstract?
+Can we say something more explicit -- something like ``estimators based on different assumptions about network structure''?} other plausible low-dimensional approximations.
+Furthermore, we found consistent relationships between the organization of sparse interactions inferred by the estimator and the physical distances and preferred orientation differences of pairs of cells.
 Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with difference in preferred orientation whereas negative interactions were less selective.
@@
 Correlations of the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.
-Estimation of correlation matrices can be improved by imposing some kind of structure on the estimate with greatest improvement attained when the imposed structure efficiently captures real dependencies in the data.
+Estimation of correlation matrices can be improved by imposing some kind of structure on the estimate with greatest improvement attained when the imposed structure closely matches the real dependencies in the data.
 Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the cross-validated performance of correlation matrix estimators that imposed different kinds of structure.
@@
 Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.
-For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Wilke:2002,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}.
-In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.
-Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Arieli:1995,Chiu:2002,Kenet:2003,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
+For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}.
+In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.
+Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
 
 However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.
-Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including  direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b,Moore:1970,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
+Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including  direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
 
@@
 Low-rank approximations are particularly suitable for capturing shared fluctuations across the recorded population.
-Extracted through principal component or factor analysis, low-rank components of neural correlations have been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.
+Extracted through principal component or factor analysis, low-rank components of neural correlations have been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2006,Yu:2009}.
 
-Alternatively, the correlation matrix can be formulated through the partial correlations between pairs of neurons.
+Alternatively, the correlation matrix can be formulated through the \emph{partial correlations} between pairs of neurons.
 The partial correlation between a pair of neurons is the linear correlation remaining after accounting for correlations to all the other neurons.
@@
 Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been used to infer interactions between components in complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
-In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies.
+In \emph{elliptical distributions} (those of the form $\frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is a normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies.
 A multivariate normal distribution specified through its graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
 With increasing departure from linearity, the correspondence between conditional dependence and partial correlations diminishes or completely breaks down \cite{Loh:2012}.
-For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006,Tkacik:2006}.
+For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations.
+\Kcomment{Part of this - especially
+a discussion of where the approach is applicable  and when it breaks down - should be expanded a bit in the discussion, and maybe 
+should be mentioned only there.}
 
 \subsection*{Estimation of neural correlation matrices}
+
+\Kcomment{Is this still the Introduction?}
+
 In this study, we pursued two related aims: (a)~improved estimation of neural correlation matrices and (b)~discovery of low-dimensional structure of correlations in recordings of multineuronal activity to facilitate their interpretation.
@@
 
-Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
+Given a covariance matrix estimate $C$, the correlation matrix $R$ and partial correlation matrix $P$ are computed as
 \begin{equation}
@@
 \end{equation}
-Similarly, the matrix of partial correlations $P$ is computed as the negative normalized version of the \emph{precision matrix} (inverse covariance matrix) $C^{-1}$:
 \begin{equation}
@@
 Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix.
-Clearly, off-diagonal zeros in the precision matrix correspond to zero partial correlations.
+$R$ contains 1s on its diagonal whereas $P$ contains $-1$s on the diagonal.
+The off-diagonal elements of $P$ contain the partial correlations between corresponding pairs of neurons.
 
@@
 Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.
-The usual estimate $C_{\sf 0}$ has the advantage of being unbiased ($\E{C_0}=\Sigma$) but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in training sample.
+The usual estimate $C_{\sf 0}$ has the advantage of being unbiased ($\E{C_0}=\Sigma$) but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in the data.
 Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from truth $\Sigma$.
@@
 Yet, when the target estimate is well suited for capturing the important features of the true covariance matrix with few terms, it will introduce minimal bias and outperform other estimators.
+\Kcomment{$\rightarrow$ ``Yet, when the target estimate captures important features of the true covariance matrix with few terms, it introduces minimal bias and outperform other estimators.''  I think we need a reference for this statement.
+Also, here
+minimal is used imprecisely, and another term may be better.}
 
-To illustrate challenges of estimating the correlation matrix from a finite sample, we briefly consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
-In covariance selection, the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero.
-We estimated the covariance matrix of a period of a somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (Fig.\;\ref{fig:01}\,A, B) and covariance selection (Fig.\;\ref{fig:01}\,C).
+Figure \ref{fig:01} illustrates the challenges of estimating the correlation matrix from a finite sample and one example of regularization.
+\Kcomment{``and one example of regularization'' seems like an add-on.}  Panel A depicts the correlation matrix computed from the sample covariance matrix $C_{\sf 0}$ for a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex.
+To show detail, panel B magnifies the fragment outlined in panel A.
+Panel C shows the same fragment from the partial correlation matrix also computed from $C_{\sf 0}$.
 Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
-The apparent similarity of the two estimates belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): the regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
-Estimation by covariance selection is attractive for several reasons: First, it has fewer free parameters and is therefore less susceptible to sampling noise, yielding estimates that are, on average, closer to truth $\Sigma$.
-Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.
+
+Panels \ref{fig:01}\,D and E depict the correlation matrix and the partial correlation matrix computed from a different estimate, which we will refer to here as $C_{\sf sparse}$, which is computed through \emph{covariance selection} \cite{Dempster:1972}.
+In covariance selection, only a subset of partial correlations is fitted while the rest are fixed at zero.
+The resulting estimate closely reproduces the sample correlation matrix (Fig.~\ref{fig:01}\,D), but its partial correlation structure appears substantially different(Fig.~\ref{fig:01}\,E).
+In this particular example  31,501 of the 44,253 (71.2\%) total partial correlations are set to zero.
+This alternative estimate is attractive for a few reasons.
+First, it has fewer free parameters and is therefore less susceptible to sampling noise.
+Additionally, assuming that partial correlations do indeed reflect elements of the functional connectivity, the second estimate also appears more physiologically plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.
 
 \subsection*{Model selection}
-However, covariance selection is but one of numerous possible low-dimensional approximations of the covariance matrix.
+Covariance selection is but one of numerous possible low-dimensional approximations of the covariance matrix.
 Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix.
-Models that accurately reflect the dominant functional dependencies in the circuit, its so-called `functional connectivity', will tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} than the other estimators or to \emph{dominate} the other estimators.
+Models that accurately reflect the dominant functional dependencies in the circuit, its so-called ``functional connectivity'', will tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators.
 An estimator's superior efficiency can be ascertained by cross-validation.
@@
 Its structure may then be analyzed and related to the circuits anatomical organization.
+\Kcomment{Need a couple refs to standard books here.}
 
 \subsection*{Summary of findings}
-In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and `sparse partial correlations with latent units' or $C_{\sf sparse+latent}$.
+In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and 'sparse partial correlations with latent units' or $C_{\sf sparse+lowrank}$.
 First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators.
 We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex.
-We found that $C_{\sf sparse+latent}$ consistently dominated the other estimators.
-Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.
+We found that $C_{\sf sparse+lowrank}$ consistently dominated the other estimators.
+Estimate produced by $C_{\sf sparse+lowrank}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.
 
@@
 Shrinkage allows for partial commitment to the low-dimensional representation.
+\Kcomment{This was already stated in introduction.}
 
@@
 The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations are the result of direct linear interactions between pairs of observed cells and that such interactions occur only between a fraction of such pairs (Fig.~\ref{fig:02}\,C).
-This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.
-Therefore, $C_{\sf sparse}$ is biased toward the assumption that correlations arise due to linear interactions between pairs of recorded neurons (Fig.~\ref{fig:03}C).
+This assumption is enforced by reducing to zero the majority of pairwise \emph{partial correlations} in the recorded population.
+While usual correlations are calculated from the marginal distribution without conditioning on all the other neurons, the partial correlation expresses the pairwise correlation conditioned on the activity of all the other cells.
+If cells only exert linear effects on each other and all neurons are recorded, partial correlations express the direct interactions between neurons.
+Therefore, this estimator is biased toward the assumption \Kcomment{What does it mean to be ``biased towards an assumption''?} that correlations arise due to interactions between a subset of pairs of recorded neurons (Fig.~\ref{fig:03}C).
+When the partial correlation between a pair of neurons is zero, then the corresponding element of the inverse of the covariance matrix (often referred to as the \emph{precision matrix} or \emph{concentration matrix}) must be zero as well.
 Then the estimator is
@@
 
-Finally, we consider the fourth estimator $C_{\sf sparse+latent}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D).
-This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
+Finally, we consider the fourth estimator $C_{\sf sparse+lowrank}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D).
+This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}
 \begin{equation}
@@
 In this study, we adopted \emph{negative normal log-likelihood loss}\footnote{
-The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$ as
+The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$: 
 \begin{equation*}
@@
 Other popular choices for the loss function are the Frobenius norm of the difference $\|C-\Sigma\|_F$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, and quadratic loss \cite{James:1961,Fan:2008}.
-We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.
+\Kcomment{Is this list necessary here?} We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.
 
@@
 In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (Figure \ref{fig:02}, Row 5).
-For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth: $\ell_{C,C_{\sf diag}}$, $\ell_{C,C_{\sf factor}}$, $\ell_{C,C_{\sf sparse}}$, and $\ell_{C,C_{\sf sparse+latent}}$.
+For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth: $\ell_{C,C_{\sf diag}}$, $\ell_{C,C_{\sf factor}}$, $\ell_{C,C_{\sf sparse}}$, and $\ell_{C,C_{\sf sparse+lowrank}}$.
 Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples  taken for each ground truth and for each sample size.
@@
 
-In these highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would be more efficient than others.
+In these highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+lowrank}$) would be more efficient than others.
 At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), to the advantage of estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$).
@@
 
-We found that estimator $C_{\sf sparse+latent}$ was more efficient than the other estimators (Fig.~\ref{fig:04}).
-The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).
+We found that estimator $C_{\sf sparse+lowrank}$ was more efficient than the other estimators (Fig.~\ref{fig:04}).
+The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+lowrank}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).
 
 \subsection*{Relationship of correlation structure to circuit architecture}
-Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the elements of the low-dimensional structure revealed by this estimator for individual imaged sites.
-By its design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
+Having demonstrated that $C_{\sf sparse+lowrank}$ dominated the other estimators, we examined the elements of the low-dimensional structure revealed by this estimator for individual imaged sites.
+By its design, the sparse+lowrank estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
 If this estimate dominates all other estimators evaluated thus far, it seems reasonable to hypothesize that these interaction may reflect underlying physiological interactions more closely than the usual correlations.
@@
 
-We examined the structure of the sparse+latent estimate in individual sites such as the example site first depicted in Fig.\;\ref{fig:03}.
+We examined the structure of the sparse+lowrank estimate in individual sites such as the example site first depicted in Fig.\;\ref{fig:03}.
 In this site, as in others, the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (Fig.~\ref{fig:05}\,A and D).
@@
 
-In previous studies of the structure of correlations, a sparse network was produced by thresholding the correlations coefficients  above some level deemed significant and examining the network of correlations above that threshold \cite{Golshani:2009,Malmersjo:2013}.
-There was fairly little overlap between the network of interactions revealed by this thresholding method and those revealed by the sparse+latent estimator (Fig.\;\ref{fig:05}\,F).
+In previous publications that looked at the structure of correlations, the sparse network was obtained by thresholding the correlations coefficients  above some level deemed significant and examining the network of correlations above that threshold \cite{Golshani:2009,Malmersjo:2013}.
+There was fairly little overlap between the network of interactions revealed by this thresholding method and those revealed by the sparse+lowrank estimator (Fig.\;\ref{fig:05}\,F).
 When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (Fig.\;\ref{fig:05}\,F, H, and I).
@@
 \section*{Discussion}
-\subsection*{Neural correlations as indicators of functional connectivity}
-Neural correlations have a rich history in neuroscience as indicators of \emph{functional connectivity}.
-When measured on temporal scales of a few milliseconds, peaks in the cross-correlations between pairs of neurons at latencies typical of synaptic transmission are often interpreted as evidence of direct synaptic connections between pairs of cells \cite{Moore:1970,Alonso:1998,Denman:2013}.
-Without targeted stimulations of specific cells or measurements of intracellular signals, such analysis remains the best test for probable monosynaptic connections in vivo.
-At longer temporal scales, neural correlations reflect activations involving large ensembles of neurons.
-Such correlations may have emergent structure with little  correspondence to synaptic connectivity, which  may better reflect computations performed by the circuit.
-For this reason, firing rate correlations at scales of tens or hundreds milliseconds have been of interest for their implications on coding under the assumption of rate codes and linear downstream decoder \cite{Averbeck:2006}.
-The terms \emph{functional connectivity} is still often applied to slow correlations, similarly to its use in PET, fMRI, EEG, and MEG brain imaging studies, as an expression of task-related, stimulus-evoked, or spontaneous synchronization or correlation between brain regions, circuits, or cells without implying a mechanistic interpretation.
+\subsection*{Functional connectivity from big neural data}
+Transformational breakthroughs in neuroscience are anticipated to emerge from technological developments that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far.
 
-Calcium imaging allows simultaneous measurements of the activity of large populations of cells, even of entire nervous systems \cite{Leung:2013,Ahrens:2013}, with cell-wise resolution, but its temporal bandwidth is limited to the scale of tens or hundreds milliseconds by the kinetics of calcium-sensitive dyes.
-This limits their use to the analysis of slower correlations.
-
-\subsection*{Correlation `structure'}
-The term `structure' implies a reduced, simplified representation expressing the essential organization.
-The structure of slow correlations is often expressed through low-rank approximations extracted by principal component analysis or factor analysis.
-These components show synchronous multineuronal activation patterns without revealing graphs of interactions between neurons.
-
-Another reduced representation of correlation matrices is obtained by sparsifying its inverse, resulting in a graph of partial correlations.
-Under the assumption of a multivariate normal distribution or, more generally, an elliptical distribution, sparse partial correlations suggest a sparse matrix of dependencies, a probabilistic graphical model.
-This property has led some investigators to suggest that, in brain imaging studies, such approximation must serve as better proxies of direct interactions between brain regions than correlations \cite{Varoquaux:2012,Ryali:2012}.
-
-\subsection*{Interpretation of functional connectivity from population calcium signals}
-Nonetheless, some investigators have examined graphs of the most significant correlations in population calcium activity as proxy of connectivity \cite{Malmersjo:2013}.
-
-While synaptic connectivity constitutes a directed weighted graph, most meaningful representation of functional organization at longer temporal scales are less certain.
-A common approach has been to examine the low-rank structure of the joint population activity extracted through principal component analysis, factor analysis, independent component analysis, linear embedding, and others.
-Functional connectivity is then represented as the eigenvectors, principle components, or factor loadings corresponding to the extracted components of the joint activity.
-
-If only most significant components are retained, these two representations turn into, respectively, low-rank approximations and sparse inverse approximations.
-Low-rank representation suggest interpretation as common input onto large subsets of neurons or highly synchronous activations in the entire recorded population.
-
-We sought to uncover the optimal low-dimensional approximation of the covariance matrix for efficient estimation of the true covariance matrices in cortical microcircuits on the temporal scale of 100--200 ms corresponding to the temporal resolution of spike rates from somatic calcium signals \cite{Cotton:2013}.
-We considered sparse-inverse approximations, low-rank approximations, and mixtures of low-rank components with sparse partial correlations.
-We used cross-validation to determine which of such representations was closest to the true correlation matrix.
-As we illustrated by simulation, none of the estimators was intrinsically more efficient than the others and their relative performance was dependent on the structure of dependencies in the empirical data.
-We found that sparse inverse approximation performed better than low-rank approximations (Fig.\;S1\;E) and approximations by combinations of sparse and low-rank components consistently outperformed the other estimators.
-This finding is specific to the circuit we examined and may not generalize to other circuits or to less dense sampling of cells in electrophysiological recordings.
-
-\subsection*{Probabilistic models and their interpretation}
-On the millisecond scale, most common analyses of the functional connectivity have relied on cross-correlation analysis in attempts to detect signatures of direct synaptic connectivity.
-
-Increasingly, the functional organization of neural circuits is inferred by fitting probabilistic models such as various forms of multivariate Gaussian models, linear-nonlinear models, Generalized Linear Models \cite{Pillow:2008}, and maximum entropy models \cite{Schneidman:2006,Tkacik:2006,Tang:2008,Shlens:2009}.
-In probabilistic models, functional dependencies correspond to multiplicative terms involving pairs or larger groups of neurons.
-If two neurons do not both participate in any of the multiplicative terms in the model, they are conditionally independent: the activity of one can be predicted from other neurons just equally well with or without considering the second neuron, suggesting a lack of direct interaction between them.
-Additionally, various probabilistic models allow inferring latent variables to explain joint patterns of in large groups of cells.
-The uncovered networks of dependencies from such statistical models may arise from mechanistic interactions such as direct monosynaptic connections or from complex emergent processes involving large ensembles of cells.
-Because statistical models of the joint activity of populations of neurons use the activity of the entire population to infer functional dependencies, they to produce more accurate and relevant representations of the functional organization of the data than available from equivalent numbers of paired recordings.
-
-However, no empirical evidence yet exists to demonstrate that statistical models of population activity reveal networks of dependences more informative than the conventional correlations with respect to the structural architecture of specific neural circuits.
-
-\subsection*{Model selection to uncover structure best supported by data}
-Model selection from various several probabilistic models
-
-\subsection*{Functional connectivity from big neural data}
-Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}.
-It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.
-
-Such probabilistic models of population activity feature large numbers of parameters, growing quadratically or faster with population size.
-Estimation of large number of parameters is sensitive to sampling noise.
-To solve this problem, the recording durations must increase at least proportionally to the number of parameters or the model must be constrained by regularization.
-
-\subsection*{Comparison with other uses of neural correlations}
 \subsection*{Meaning of interactions}
-\subsection*{Fitting and evaluation of probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
+\subsection*{Fitting probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
 \subsection*{Future directions}
@@
     \end{center}
-    \caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
-    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
+    \caption{{\bf The sparse+lowrank estimator $C_{\sf sparse+lowrank}$ outperforms the other estimators on neural data.}
+    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+lowrank}$. 
     The histograms are based on 31 imaged sites in 24 mice. 
-    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
+    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+lowrank}$ over the other estimators. 
     The green arrows indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}

### 2013-12-19 plos.tex
complete but untidy Discussion

--- 2013-12-19 plos.tex
@@
 \subsection*{Model selection}
-Various estimation methods reveal radically different structures of functional connectivity?
-Which of the structures should be preferred?
+Various estimation methods infer very different structures of functional connectivity from the same data.
+Which structure should be preferred?
 How does one determine which functional connectivity structure is best supported by the empirical data?
-Machine learning literature offers a rich toolbox of model selection criteria.
-Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion}, and \emph{Bayes (or Schwartz) Information Criterion} are computationally efficient but rely on assumptions on generally unsupportable assumptions about the data generating process.
-The most assumption-free yet more computationally intensive are procedures based on cross-validation \cite{Arlot:2010} and, when models have hyperparameters, nested cross-validation to selected the optimal values of the hyperparameters.
+The field of machine learning offers a rich toolbox of model selection criteria.
+Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion} (AIC), and \emph{Bayes (or Schwartz) Information Criterion} (BIC) are computationally efficient but rely on generally unsupportable assumptions about the data generating process.
+The most assumption-free yet more computationally intensive are procedures based on cross-validation \cite{Arlot:2010} and, when models have hyperparameters, nested cross-validation to select their values.
 It is commonly assumed that probabilistic models that demonstrate the highest average cross-validated log likelihood yield a better representation of the functional connectivity.
-For example, \cite{Ganmor:2011} show the fact that their proposed Reliable Network Model (a maximum-entropy model on the multivariate binary domain constrained by a sparse network of correlations of various orders) yielded higher cross-validated log-likelihoods than the full-pairwise Ising model.
-The authors use this fact to support the claim that the former model revealed a more anatomically relevant structure of functional connectivity than the pairwise model.
+For example, \cite{Ganmor:2011} demonstrated that their proposed Reliable Network Model (a maximum-entropy model on the multivariate binary domain constrained by a sparse heuristically selected network of correlations of several orders) yielded higher cross-validated log-likelihoods than the full-pairwise Ising model.
+The authors use this fact to support the claim that the former model revealed a more anatomically relevant structure of functional connectivity than the pairwise Ising model.
 
-Under the loss function (Eq.~\ref{eq:loss}) employed in our study, the covariance matrix estimation methods are equivalent to fitting multivariate Gaussian models with various structure of dependencies between visible and inferred latent units and selecting between them on the basis of cross-validated log-likelihood.
-However, the ranking of models was preserved under other loss functions (Fig.~\ref{fig:S02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.
+In simulation, we showed that, more constrained, simpler models often outperformed complex models in cross-validation studies even when their structure was differed from the true structure (Fig.~\ref{fig:02}\;row 5), only because these more constrained models were less sensitive to sampling noise.
+Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process.
+For example, a properly sparsified pairwise Ising model could outperform the Reliable Network model.
+Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.
+
+In our study we selected between families of models based on their cross-validation performance.
+The winning model family was more flexible than the others and therefore more susceptible to noise.The fact that it still performed better indicated that the additional features were required.
+We conclude that under these circumstances, the selected model can be asserted to better represent the true structure of the data generating process, at least in comparison to the other three models.
+
+In our study, we focused on estimating a specific parameter of the distribution (its covariance matrix) rather than fitting a specific parametric distribution.
+Of course, under the loss function (Eq.~\ref{eq:loss}) we selected, the covariance matrix estimation methods are equivalent to fitting a multivariate Gaussian model and selecting the best structure on the basis of cross-validated log-likelihood.
+However, the ranking of models was preserved under other loss functions (Fig.~\ref{fig:Supp02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.
+We defend the practice of evaluating and optimizing probabilistic models with respect to their parameters that are more relevant for functional connectivity rather than maximizing their cross-validated log likelihoods, which are sensitive to all parameters.
+For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix.
+We maintained that model B must be more relevant for studies of functional connectivity because correlations do reflect aspects of functional connectivity.
+
+The correlation structures we considered also corresponded to structures of conditional dependencies in multivariate Gaussian models or, more generally, elliptical distributions.
+For example, zero partial correlation indicates conditional independence between a pair of variables under an elliptical distribution, but this correspondence weakens or breaks down with deviation from elliptical distributions \cite{Loh:2012}.
+Other probabilistic models, when optimally sparsified for best cross-validated estimation of the covariance matrix, could suggest very different forms of low-dimensional representations (structures) of the covariance matrix.
+For example, spatiotemporal Ising models, marginalized to reproduce the correlation matrix at the same temporal scale, would suggest an alternative correlation structure from the same data.
+At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
+If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+lowrank}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity.
+Various forms of GLMs could also be evaluated with in the same way: sparsified for best estimation of the covariance matrix (not log-likelihood) and compared to the estimators studied in this study.
+Again, if this new covariance estimator outperforms, $C_{\sf sparse+lowrank}$, we will be compelled to adapt it for estimating the functional connectivity.
 
@@
 It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.
+Large datasets aggravate the numerical problems and the model selection caveats described in this paper.
+The number of parameters in models of population activity grows quadratically.
+Models of joint population activity feature large numbers of parameters, that growing quadratically or faster with population size.
+Without ways to restrict such models, they quickly become ill conditioned (e.g.
+\cite{Roudi:2009}.
+Statistical models of population activity will need to have both correct mathematical forms of interactions and optimal ways of restricting them so that the total number of estimated parameters remains commensurate with the amount of measurements.
 
-Such probabilistic models of population activity feature large numbers of parameters, growing quadratically or faster with population size.
-Estimation of large number of parameters is sensitive to sampling noise.
-To solve this problem, the recording durations must increase at least proportionally to the number of parameters or the model must be constrained by regularization.
+Our study employs regularization, which restricts the models based on the data itself.
+Other models may use restrictions that are based on various heuristics in addition to regularization.
+These methods have numerous caveats when it comes to inferring the optimal structure of functional connectivity from such methods.
+Improved estimation of covariances and the general framework of producing such estimators for specific neural populations can be applied to various circuits and serve as a statistical description of their functional organization.
 
-\subsection*{Future directions}
+\subsection*{Correspondence to anatomical organization}
+If the ultimate goal of this line of inquiry to find descriptions of functional connectivity that are most informative about the anatomical and physiological organization of the circuit, then the models must be directly validated according to their correspondence to the anatomical organization.
+Little empirical evidence presently exists to show which models of connectivity, when fitted to the recordings of activity, reveal functional connectivity with best correspondence to anatomical connectivity.
+
+In this study we find that the structure of partial correlation of the optimal estimator had stronger dependencies on the separation in depth, lateral position, and orientation preferences than sample correlations (Fig.~\ref{fig:07}\;A--C).
+The inferred positive and negative partial correlations too had distinct relationships to structure  (Fig.~\ref{fig:07}\;D--F).
+Also the average partial correlations were more consistent between imaged sites than average sample correlations (Fig.~\ref{fig:06}\;A).
+Although these effects could conceivably results from trivial numerical artifacts, they are consistent with  improved representation of functional connectivity by the proposed estimator.
+
+Additional studies with additional markers of circuit architecture such as cell types and synaptic connectivity could shed additional light on the ability of models to reveal meaningful functional connectivity.
+For example, labeled interneurons could be studied for their distinct connectivity patterns contrasted with pyramidal cells.
+
+\subsection*{Conclusions and future work}
+We proposed an alternative statistically verified method for estimating the functional connectivity in neural circuits at temporal scales attainable by calcium imaging.
 

### 2013-12-20 plos.tex
rev27 (rev26 after responding to Kreso's changes/comments)

@@
 % Please keep the abstract between 250 and 300 words
-Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.
-Correlations measured in such recordings are anticipated to reveal important aspects of the functional organization of neural circuits.
-However, estimation and interpretation of large correlation matrices from finite recordings are challenging.
-Estimation can be improved by regularization: biasing of the estimate toward a low-dimensional, less variable approximation.
-The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.
-Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies governing the system.
+Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.
+Correlations in neural activity measured in such recordings can reveal important aspects of the functional organization of neural circuits.
+However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.
+Estimation can be improved by regularization: biasing of the usual unbiased estimate toward its low-dimensional approximation.
+The amount of improvement is greatest when the approximation chosen as the target of regularization parsimoniously captures the dominant dependencies in the data.
+Therefore, the construction of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.
 
-In this study, we sought the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.
+In this study, we sought most statistically efficient estimators of neural correlations in recordings from large, dense groups of cortical neurons.
 Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 100 $\mu$m in diameter (150--350 cells) in mouse visual cortex.
-We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.
-Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward other plausible low-dimensional approximations.
-Furthermore, we found consistent relationships between the organization of sparse interactions inferred by this estimator and the physical distances and preferred orientation differences of pairs of cells.
-Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with difference in preferred orientation whereas negative interactions were less selective.
-To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and brain states.
-
+We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.
+Indeed, in cross-validation tests, the covariance matrix estimator based on such approximations outperformed other regularized estimators.
+Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.
+The inferred function connectivity obtained using these methods can provide hypotheses about synaptic architecture that can guide future experimental work.
 \section*{Author Summary}
@@
 % Author Summary not valid for PLoS ONE submissions.   
-Correlations of the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.
-Estimation of correlation matrices can be improved by imposing some kind of structure on the estimate with greatest improvement attained when the imposed structure efficiently captures real dependencies in the data.
-Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the cross-validated performance of correlation matrix estimators that imposed different kinds of structure.
-The correlation structure of the estimator that proved most efficient comprised a sparse network of partial correlations between pairs of neurons combined with several common latent factors, with both components required for efficient estimation.
-Because of its superior benefit for estimation, we proposed that this inferred structure could prove more relevant for the description of functional connectivity than the usual correlation matrix in densely sampled neural recordings.
-As a first application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.
+Correlations in the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.
+Estimation of correlation matrices can be improved by imposing some structure on the estimate.
+Greatest improvements are attained when the imposed structure efficiently captures real dependencies in the data.
+Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the performance of correlation matrix estimators that imposed different kinds of structure.
+The correlation structure of the estimator that led to greatest improvement comprised a sparse network of partial correlations between pairs of neurons combined with partial correlations with several common latent factors.
+Both components were necessary for efficient estimation.
+Since it leads to better estimates, we proposed that this sparse+latent structure provides a better picture of the  functional connectivity than that obtained from the usual sample correlation matrix in densely sampled neural recordings.
+As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.
 
@@
 \subsection*{Neural correlations}
-Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.
-For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Wilke:2002,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}.
+Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.
+For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}.
 In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.
-Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Arieli:1995,Chiu:2002,Kenet:2003,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
+This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Arieli:1995,Chiu:2002,Kenet:2003,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
 
@@
 For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation.
-Low-rank approximations are particularly suitable for capturing shared fluctuations across the recorded population.
-Extracted through principal component or factor analysis, low-rank components of neural correlations have been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.
+Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.
+Low-rank components can be  extracted using principal component or factor analysis.
+The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.
 
@@
 
-\begin{figure}[!ht]
-    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
     {\caption{{\bf Illustration of regularized estimation of partial correlations.}
@@
 
-\begin{figure}
-    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
     {\caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
@@
 
-\begin{figure}[!ht]
-    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
     {\caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
@@
 
-\subsection*{Conclusions and future work}
-We proposed an alternative statistically verified method for estimating the functional connectivity in neural circuits at temporal scales attainable by calcium imaging.
-
 \section*{Methods}

### 2013-12-21 plos.tex
Rev28 -- Kreso's edits of Abstract--Results.

@@
 The amount of improvement is greatest when the approximation chosen as the target of regularization parsimoniously captures the dominant dependencies in the data.
-Therefore, the construction of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.
+Therefore, the selection of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.
 
@@
 Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.
-The inferred function connectivity obtained using these methods can provide hypotheses about synaptic architecture that can guide future experimental work.
+The inferred function connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.
 \section*{Author Summary}
@@
 
-Alternatively, the correlation matrix can be formulated through the partial correlations between pairs of neurons.
-The partial correlation between a pair of neurons is the linear correlation remaining after accounting for correlations to all the other neurons.
-Partial correlations carry particular significance when conditional dependencies between all the variables are thought to be dominated by linear effects.
-In such systems, zero  partial correlation between two variables implies their conditional independence.
+Alternatively, the correlation matrix can be expressed through the partial pairwise correlations.
+The partial pairwise correlation is the linear correlation remaining after accounting for correlations with all the other neurons.
+Partial correlations carry particular significance when the relationship between the variables is approximately linear .
+In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent.
+
 Conditional independence between a pair of variables suggests a lack of direct interaction between the underlying processes.
-Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been used to infer interactions between components in complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
-In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies.
-A multivariate normal distribution specified through its graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
-With increasing departure from linearity, the correspondence between conditional dependence and partial correlations diminishes or completely breaks down \cite{Loh:2012}.
+Therefore, networks of partial correlations (sometimes called \emph{association networks}) can be used to infer interactions between the components of a complex system.
+This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005,Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
+In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies.
+A multivariate normal distribution specified by a weighted graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
+With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}.
 For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006,Tkacik:2006}.
@@
 \subsection*{Estimation of neural correlation matrices}
-In this study, we pursued two related aims: (a)~improved estimation of neural correlation matrices and (b)~discovery of low-dimensional structure of correlations in recordings of multineuronal activity to facilitate their interpretation.
+Here, we pursued two related aims: (a)~the improved estimation of neural correlation matrices and (b)~discovery of the low-dimensional structure of correlations in recordings of multineuronal activity.
+An accurate characterization of the patterns of correlations helps in interpreting them, and to better relate them to the architecture and dynamics of the underlying neuronal networks.
 
@@
 where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent).
-In this study, we keep the sample mean $\bar x$ as the estimate of the true mean $\mu$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.
+In this study, we estimate the true mean, $\mu$,  using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.
 
@@
 
-Estimation of the covariance matrix from a large population presents a number of numerical challenges.
-The amount of recorded data grows only linearly with the population size whereas the number of estimated coefficients increases quadratically.
-This mismatch leads to increased opportunities for spurious correlations, to overestimation of shared activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and to poorly conditioned estimates of the partial correlations \cite{Schafer:2005}.
+Estimation of large covariance matrices presents a number of numerical challenges.
+The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
+This mismatch leads to an increase in spurious correlations, overestimation of shared activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned estimates of the partial correlations \cite{Schafer:2005}.
 
 Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.
-The usual estimate $C_{\sf 0}$ has the advantage of being unbiased ($\E{C_0}=\Sigma$) but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in training sample.
-Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from truth $\Sigma$.
-Regularization works by striking a favorable balance between bias and variability.
-Regularization can produce \emph{some} improvement even with an arbitrary target estimate that has no relation to the true dependencies in the data.
-Yet, when the target estimate is well suited for capturing the important features of the true covariance matrix with few terms, it will introduce minimal bias and outperform other estimators.
+The sample estimate $C_{\sf 0}$ is unbiased ($\E{C_0}=\Sigma$) but, on average, falls far from $\Sigma$ due to its sensitivity to sampling noise.
+Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.
+Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate  Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
 
 \subsection*{Illustration of estimation of a neural correlation matrix}
-To illustrate challenges of estimating the correlation matrix from a finite sample, we briefly consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
-In covariance selection, the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero.
-We estimated the covariance matrix of a period of a somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (Fig.\;\ref{fig:01}\,A, B) and covariance selection (Fig.\;\ref{fig:01}\,C).
+To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
+Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero.
+We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (Fig.\;\ref{fig:01}\,A, B) and covariance selection (Fig.\;\ref{fig:01}\,C).
 Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
-The close similarity of the two estimates (Fig.\;\ref{fig:01}\,B and C) belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): the regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
-Estimation by covariance selection is attractive for several reasons: First, it has fewer free parameters and is therefore less susceptible to sampling noise, yielding estimates that are, on average, closer to truth $\Sigma$.
+The close similarity of the two estimates of the correlation matrix (Fig.\;\ref{fig:01}\,B and C) belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
+Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth.
 Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.
@@
 Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix.
-Models that accurately reflect the dominant functional dependencies in the circuit, its so-called `functional connectivity', will tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} than the other estimators or to \emph{dominate} the other estimators.
+Models that accurately reflect the dominant regularities in the activity of the circuit, its so-called `functional connectivity', tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators.
 An estimator's superior efficiency can be ascertained by cross-validation.
-Because different types of neural circuits are governed by different types of dependencies, different probabilistic models and covariance matrix estimators may dominate in each case.
-The estimator that is shown to dominate all others for a specific type of neural circuit may be proposed as the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
+Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case.
+The estimator that is shown to dominate all others for a specific type of neural circuit may provide the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
 Its structure may then be analyzed and related to the circuit's anatomical organization.
@@
         In `sparse interactions' ({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
-        In `sparse+latent' ({\bf D}), correlations arise through direct linear interactions between some pairs of observed neurons and through interactions with common latent units. 
+        In `sparse+latent' ({\bf D}), correlations arise through direct linear interactions in a sparse subset of observed neurons and through interactions with common latent units. 
         {\bf Row 2.} Examples of $50\times 50$ correlation matrices corresponding to each type of low-dimensional structure. 
@@
 The error bars indicate the standard error of the mean based on 30 samples.
-        {\bf Row 5.} Mean \emph{cross-validation loss} (Eq.~\ref{eq:rel-cv-loss}) of covariance estimators with respect to the matching estimator .
+        {\bf Row 5.} Mean \emph{cross-validation loss} (Eq.~\ref{eq:rel-cv-loss}) of covariance estimators with respect to the matching estimator.
 The values are relative to the validation loss of the estimator that matches the low-dimensional structure of the true covariance matrix.
-The error bars indicate the standard error of the mean based on 30 samples.
+Error bars indicate the standard error of the mean based on 30 samples.
     }
@@
 
-In the first regularized estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ containing on its diagonal estimates of the variances.
-The regularized estimate is obtained by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
+For the first regularized estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ whose diagonal contains  variance estimates.
+Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
 \begin{equation}
@@
 \end{equation}
-The diagonal target estimate expresses the idea of a lack of dependence (or of linear association) between the activity of observed neurons (Fig.~\ref{fig:02}\,A).  
-If this assumption aptly describes recorded data, then strong shrinkage toward $D$ will add little bias while strongly reducing the variability of the estimate.
+The diagonal target has structure corresponding to the absence of linear associations between the activity of observed neurons (Fig.~\ref{fig:02}\,A).
+If this structure accurately describes recorded data, then strong shrinkage toward $D$ adds little bias while strongly reducing the variability of the estimate.
 Shrinkage allows for partial commitment to the low-dimensional representation.
 
-In the second regularized estimator $C_{\sf factor}$, the target estimate is the factor model $F =  L L^\T + \Psi$ with $d$ factors so that $L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\Psi$ contains the independent variances of each neuron.
-Then the estimate is 
+The target of the second regularized estimator $C_{\sf factor}$, is the factor model $F =  L L^\T + \Psi$ with $d$ factors.
+Therefore $L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\Psi$ contains the independent variances of each neuron.
+The estimator,
 \begin{equation}
-C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda F
+C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda F,
 \end{equation}
-This estimator has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $F$ expresses the assumption that correlated fluctuations in the population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}\,B).
+has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
+The target estimate $F$ has the structure that arises when correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}\,B).
 
-The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations are the result of direct linear interactions between pairs of observed cells and that such interactions occur only between a fraction of such pairs (Fig.~\ref{fig:02}\,C).
+The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (Fig.~\ref{fig:02}\,C).
 This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.
-Therefore, $C_{\sf sparse}$ is biased toward the assumption that correlations arise due to linear interactions between pairs of recorded neurons (Fig.~\ref{fig:02}\,C).
-Then the estimator is
+Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (Fig.~\ref{fig:02}\,C).
+The estimator takes the form
 \begin{equation}
-C_{\sf sparse} = S^{-1}
+C_{\sf sparse} = S^{-1},
 \end{equation}
-where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal elements. The estimate has one hyperparameter to regulate the sparsity (fraction of off-diagonal zeros) in $S$.
+where $S$ is a sparse matrix with a large fraction of zeros off the diagonal. The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.
 
-Finally, we consider the fourth estimator $C_{\sf sparse+latent}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D).
-This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
+The fourth estimator $C_{\sf sparse+latent}$, combines the assumptions of the previous two.
+The underlying structure arises from sparse interactions between the recorded neurons and latent factors driving their activity (Fig.~\ref{fig:02}\,D).
+The estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
 \begin{equation}
@@
 \subsection*{Simulation}
-To verify our approach and to illustrate the performance of the four regularized estimators, we constructed four model populations of size $p=50$~neurons each with correlation structures of the same type as the target estimates of the first (Fig.~\ref{fig:02} Row 2).
-Panel A2 contains a diagonal correlation matrix matching the target of $C_{\sf diag}$.
-Panel B2 is a factor model with 3 factors conforming to the family of target estimates of $C_{\sf factor}$.
-Panel B3 is a correlation matrix with sparse partial correlations (76\% off diagonal zeros in the precision matrix).
-Finally, Panel B4 is a correlation matrix whose inverse is composed the sum of a sparse matrix (82\% sparse) and a low-rank component (rank=1).
-Row 3 contains sample correlation matrices calculated from samples of size $n=1000$ from a multivariate normal distributions with the respective correlation matrices from Row 2.
+To illustrate the performance of the four regularized estimators, we constructed four model populations of $p=50$~neurons each with each matching the correlation structures of a different target estimate (Fig.~\ref{fig:02} Row 2).
+Sample correlation matrices calculated from samples of size $n=1000$ contain substantial sampling noise (Fig.~\ref{fig:02) Row 3) that makes the correlation structure less obvious.
 
-To evaluate the performance of a covariance matrix $C$, we define a real-valued \emph{loss function} $\loss{C,\Sigma}$ such that it attains its minimum when $C=\Sigma$.
+To evaluate the performance of a covariance matrix estimator $C$, we define a real-valued \emph{loss function} $\loss{C,\Sigma}$ that attains a minimum when $C=\Sigma$.
 The loss function quantifies the error of the estimate, \emph{i.e.}~the deviation of the estimate $C$ from truth $\Sigma$.
@@
 
-This choice is motivated by mathematical convenience.
-Other popular choices for the loss function are the Frobenius norm of the difference $\|C-\Sigma\|_F$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, and quadratic loss \cite{James:1961,Fan:2008}.
-We expect that the main conclusions of our study will not change qualitatively under other well behaved loss functions.
+The choice of loss function is motivated by mathematical convenience.
+We expect that the main conclusions of our study will not change qualitatively with other well behaved loss functions, such as the Frobenius norm of the difference $\|C-\Sigma\|_F$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, or quadratic loss \cite{James:1961,Fan:2008}.
 
-To make the loss function to assume zero at its minimum when $C=\Sigma$, we define \emph{excess loss} as
+We define the \emph{excess loss} as
 \begin{equation}\label{eq:excess-loss}
-    \eloss{C,\Sigma} = \loss{C,\Sigma}-\loss{\Sigma,\Sigma}
+    \eloss{C,\Sigma} = \loss{C,\Sigma}-\loss{\Sigma,\Sigma},
 \end{equation}
+which assumes zero at its minimum.
 
-In Figure \ref{fig:02}, row 4 shows the mean excess losses and their standard errors calculated from 30 samples of sizes n=250, 500, 1000, 2000, and 4000 for each estimator and each kind of ground truth.
-For each estimator, its hyperparameters were estimated by nested cross-validation (See Methods for more details.)
+In Figure \ref{fig:02}, row 4 shows the mean excess losses and their standard errors calculated from 30 samples of sizes n=250, 500, 1000, 2000, and 4000 for each estimator and each ground truth.
+For each estimator, hyperparameters were estimated by nested cross-validation (See Methods for more details.)
 
-As expected, estimators with matching low-dimensional structures typically outperformed the other estimators.
-There were two exceptions to this observation.
-For small sample sizes, before the data were sufficient to reveal the true correlation structure and to  allow the correct model to dominate, estimates with simpler targets often outperformed the matching estimator.
+As expected, estimators with whose structure matched that of the true model typically outperformed the other estimators.
+There were two exceptions: First, for small sample sizes, there was insufficient data to reveal the true correlation structure.
+In this case estimates with simpler targets often outperformed the estimator based on the correct model.
+Second, the sparse+latent estimator often performed equally well to the sparse estimator even when the data did not have any latent units.
+In these cases, it inferred the right number of latent units: zero.
+In larger models, the falsely identified latent factors are more common.
 
-When ground truth $\Sigma$ is not accessible, loss can be estimated solely from the data through \emph{validation}.
-In validation, an additional independent \emph{testing sample} is used to compute an independent sample covariance estimate $C_{\sf 0}^\prime$.
+When the ground truth, $\Sigma,$ is not accessible, loss can be estimated solely from the data through \emph{validation}.
+In validation, an additional independent \emph{testing sample} is used to compute a sample covariance estimate $C_{\sf 0}^\prime$.
 Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can be used as an unbiased estimate\footnote
@@
 
-This property does not hold for some other popular losses such as Stein's entropy loss, for example, which prevents their substitution with corresponding validation losses.
-However, various other losses do conform to this constraints and could be used in this study.
+This property does not hold for some other popular loss functions such as Stein's entropy loss, for example, which prevents their substitution with corresponding validation losses.
+However, various other loss functions do have this property and could have been used.
 } 
@@
 With empirical data, acquiring a separate testing sample is not practical.
-Instead, $K$-fold cross-validation is used.
-In cross-validation, the sample is divided into $K$ subsets of approximately equal size.
-In all computations in this paper $K=10$ was used.
-Then each  subset is used as the validation sample with the other $K-1$ serving as the training dataset.
+Instead, $K$-fold cross-validation is used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).
+Then each  subset is used as the validation sample with the other $K-1$ serving as  a training dataset.
 The validation losses from each of such `folds' are averaged to produce \emph{cross-validation loss} or \emph{CV-loss} for short.
@@
 
-Since we only need to compare estimators against each other, we are only interested in \emph{relative} CV-loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$:
+Since we only need to compare estimators to each other, we are only interested in \emph{relative} CV-loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$:
 \begin{equation}\label{eq:rel-cv-loss}
@@
 
-These simulation results demonstrated that, with sufficiently large sample sizes, the selection of the most efficient of several regularized estimators could be used to infer the type of low-dimensional structure present in the data, at least when one of the regularized estimators could capture such structure.
+These simulations demonstrate that with sufficiently large sample sizes the most efficient among several regularized estimators could be used to infer the most likely type of low-dimensional structure present in the data.
 
 \subsection*{Covariance estimation in neural data}
-
-We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005,Katona:2012,Cotton:2013}.
+We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005,Katona:2012,Cotton:2013}.
 We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (Fig.~\ref{fig:03}\,A) on one side of the visual field of anesthetized mice.
@@
 The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.
-Besides the high-repetition stimulus protocol for noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cells (Fig.~\ref{fig:03}\,E).
+Besides the high-repetition stimulus protocol used to estimate noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cell (Fig.~\ref{fig:03}\,E).
 Overall, 31 imaged sites from 24 animals were included in the study.
@@
 
-In these highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would be more efficient than others.
-At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), to the advantage of estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$).
-Finally, there is the possibility that the correlations structure is best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$).
+In these highly localized populations both direct interactions between cells and common diffuse inputs likely contribute to overall population co-variability.
+We therefore hypothesized that regularized estimates capturing a sparse partial correlation structure, and/or interactions with latent units would be the most efficient.
+At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), giving a potential advantage to estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$).
+Finally, correlation structure could possibly be best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$).
+Thus it was a priori unclear which estimator would perform best.
 
-We found that estimator $C_{\sf sparse+latent}$ was more efficient than the other estimators (Fig.~\ref{fig:04}).
+We found that estimator $C_{\sf sparse+latent}$ was the most efficient (Fig.~\ref{fig:04}).
 The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).
@@
 \subsection*{Correlation structure and circuit architecture}
-Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the elements of the low-dimensional structure revealed by this estimator for individual imaged sites.
-By its design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
-If this estimate dominates all other estimators evaluated thus far, it seems reasonable to hypothesize that these interaction may reflect underlying physiological interactions more closely than the usual correlations.
+Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites.
+By design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
+If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations.
 In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
 
-We examined the structure of the sparse+latent estimate in individual sites such as the example site first depicted in Fig.\;\ref{fig:03}.
-In this site, as in others, the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (Fig.~\ref{fig:05}\,A and D).
+We examined the structure of the sparse+latent estimate at individual sites (an example site is depicted in Fig.\;\ref{fig:03}).
+At these sites the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (Fig.~\ref{fig:05}\,A and D).
 However, the corresponding partial correlation matrices differed dramatically (Fig.~\ref{fig:05}\,B and E).
-The partial correlation was decomposed into the sparse and low-rank components (Fig.\,\ref{fig:05}\,C).
-Although correlations were mostly positive, the sparse partial correlations (or `interactions'), although mostly positive, had a much larger proportion of negative values.
+The partial correlation was decomposed into sparse and low-rank components (Fig.\,\ref{fig:05}\,C).
+Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations.
 The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G).
-The low-rank component was of rank 17.
+The low-rank component had rank 17.
 
@@
 
-In previous studies of the structure of correlations, a sparse network was produced by thresholding the correlations coefficients  above some level deemed significant and examining the network of correlations above that threshold \cite{Golshani:2009,Malmersjo:2013}.
-There was fairly little overlap between the network of interactions revealed by this thresholding method and those revealed by the sparse+latent estimator (Fig.\;\ref{fig:05}\,F).
+In previous studies of the structure of correlations, a sparse network was produced by thresholding the correlations coefficients at a level deemed significant.
+A network was formed by connecting neurons whose total correlations exceed this threshold \cite{Golshani:2009,Malmersjo:2013}.
+There was fairly little overlap between the network of interactions revealed by such thresholding and those revealed by the sparse+latent estimator (Fig.\;\ref{fig:05}\,F).
 When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (Fig.\;\ref{fig:05}\,F, H, and I).
 In particular, many low sample correlations translated into negative interactions in the regularized estimate.
-Indeed, the absence of a correlation between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient.
+Indeed, the absence of correlations between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient.
 Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.
 
-The average partial correlations  revealed by the regularized estimator in all 31 sites were about 5 times lower than the sample correlations and  less variable across sites (Fig.\;\ref{fig:06}\,A).
-The average node degree of the sparse component of the partial correlations and the number of  inferred latent units varied widely between sites but generally increased with the recorded population size (Fig.\;\ref{fig:06}\,B and C).
+The average partial correlations  revealed by the sparse+latent estimator at all 31 sites were about 5 times lower than the sample correlations and  less variable across sites (Fig.\;\ref{fig:06}\,A).
+The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (Fig.\;\ref{fig:06}\,B and C).
 However, there was an inverse relationship between the number of latent units and the average node degree (Fig.\;\ref{fig:06}\,D).
@@
 We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}).
-The average partial correlations fell more rapidly with with difference in preferred orientation (Fig.\;\ref{fig:07}\,A and D) and lateral displacements at equal depths ($\pm 25\;\mu$m) (Fig.\;\ref{fig:07}\,B and E), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (Fig.\;\ref{fig:07}\,B and E).
+The average partial correlations fell more rapidly with difference in preferred orientation (Fig.\;\ref{fig:07}\,A and D) and lateral displacements at equal depths ($\pm 25\;\mu$m) (Fig.\;\ref{fig:07}\,B and E), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (Fig.\;\ref{fig:07}\,B and E).
 
@@
 
-The connectivity of partial correlations in the regularized estimate had different organization for positive and negative interactions.
+The connectivity expressed by the partial correlations in the regularized estimate had different organization for positive and negative interactions.
 Positive interactions fell rapidly as a function of difference in preferred orientation (Fig.\;\ref{fig:07}\,G and J), lateral displacements (Fig.\;\ref{fig:07}\,H and K), and displacement in depth (Fig.\;\ref{fig:07}\,I and L).

### 2013-12-22 plos.tex

@@
 \newcommand{\cond}{\,\vert\,}
+\newcommand{\figref}[2]{Fig.\;\ref{fig:#1}\,#2}
 \newcommand{\loss}[1]{\mathcal L\left(#1\right)} 
@@
 However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.
-Estimation can be improved by regularization: biasing of the usual unbiased estimate toward its low-dimensional approximation.
+Estimation can be improved by regularization: biasing of the usual unbiased estimate toward a low-dimensional approximation[1].
 The amount of improvement is greatest when the approximation chosen as the target of regularization parsimoniously captures the dominant dependencies in the data.
@@
 In this study, we sought most statistically efficient estimators of neural correlations in recordings from large, dense groups of cortical neurons.
-Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 100 $\mu$m in diameter (150--350 cells) in mouse visual cortex.
+Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.
 We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.
@@
 Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.
-The inferred function connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.
+The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.[2]
 \section*{Author Summary}
@@
 \section*{Introduction}
-\subsection*{Neural correlations}
+\paragraph{Neural correlations}
 Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.
 For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}.
-In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.
-This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Arieli:1995,Chiu:2002,Kenet:2003,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
+In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits[3].
+This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Arieli:1995,Chiu:2002,Kenet:2003,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b[4]}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
 
@@
 
-\subsection*{Additional information in neural correlation matrices}
-Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
+\paragraph{Additional information in neural correlation matrices}
+Yet, a complete [5]correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
 In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998,Zohary:1994}.
@@
 For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation.
-Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.
-Low-rank components can be extracted using principal component or factor analysis.
-The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.
+Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.[6] Low-rank components can be extracted using principal component or factor analysis.
+The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009[7]}.
+[8]
 
 Alternatively, the correlation matrix can be expressed through the partial pairwise correlations.
-The partial pairwise correlation is the linear correlation remaining after accounting for correlations with all the other neurons.
-Partial correlations carry particular significance when the relationship between the variables is approximately linear .
-In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent.
+The partial pairwise correlation is the linear correlation remaining[9] after accounting for correlations with all the other neurons.
+Partial correlations carry particular significance when the relationship between the variables is approximately linear.
+In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent[10].
 
@@
 This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005,Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
-In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies.
-A multivariate normal distribution specified by a weighted graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
+In \emph{elliptical distributions} [11](those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies.
+A multivariate normal distribution specified by a weighted graph[12] of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
 With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}.
-For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006,Tkacik:2006}.
+For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006,Tkacik:2006}.[13]
 
-\subsection*{Estimation of neural correlation matrices}
+\paragraph{Estimation of neural correlation matrices}
 Here, we pursued two related aims: (a)~the improved estimation of neural correlation matrices and (b)~discovery of the low-dimensional structure of correlations in recordings of multineuronal activity.
@@
 In the case of noise correlations, $\mu$ depends on the stimulus condition.
-The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
+The usual estimator [14]of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
 \begin{equation}
@@
 Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.
-Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate  Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
+Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate  [15]Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
 
-\subsection*{Illustration of estimation of a neural correlation matrix}
+\paragraph{Illustration of estimation of a neural correlation matrix}[16]
 To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
 Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero.
-We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (Fig.\;\ref{fig:01}\,A, B) and covariance selection (Fig.\;\ref{fig:01}\,C).
+We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}).
 Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
-The close similarity of the two estimates of the correlation matrix (Fig.\;\ref{fig:01}\,B and C) belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
+The close similarity of the two estimates of the correlation matrix (\figref{ 01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
 Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth.
@@
 
-\subsection*{Model selection}
+\paragraph{Model selection}
 Covariance selection is but one of many conceivable low-dimensional approximations of the covariance matrix.
@@
 Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case.
-The estimator that is shown to dominate all others for a specific type of neural circuit may provide the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
+The estimator that is shown to dominate all others for a specific type of neural circuit may provide [17]the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
 Its structure may then be analyzed and related to the circuit's anatomical organization.
 
-\subsection*{Summary of findings}
+\paragraph{Summary of findings}
 In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and `sparse partial correlations with latent units' or $C_{\sf sparse+latent}$.
@@
 Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.
+[18]
 
@@
 
-\subsection*{Covariance estimation}
-We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (Fig.~\ref{fig:02}\;row 1).
+\paragraph{Covariance estimation}
+We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' [19](\figref{02}{row 1}).
 
@@
 Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
-\begin{equation}
+\begin{equation}\label{eq:c-diag}
 C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
 \end{equation}
-The diagonal target has structure corresponding to the absence of linear associations between the activity of observed neurons (Fig.~\ref{fig:02}\,A).
+The diagonal target has structure corresponding to the absence of linear associations between the activity of observed neurons (\figref{02}{A}).
 If this structure accurately describes recorded data, then strong shrinkage toward $D$ adds little bias while strongly reducing the variability of the estimate.
@@
 The estimator,
-\begin{equation}
+\begin{equation}\label{eq:c-factor}
 C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda F,
@@
 has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $F$ has the structure that arises when correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:02}\,B).
+The target estimate $F$ has the structure that arises when correlated fluctuations in population activity are driven by a small number [20]of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
 
-The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (Fig.~\ref{fig:02}\,C).
-This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.
-Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (Fig.~\ref{fig:02}\,C).
+The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (\figref{ 02}{C}).
+This assumption is enforced by setting to zero an optimal[21] subset of the off-diagonal elements of the precision matrix.
+Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{ 02}{C}).
 The estimator takes the form
-\begin{equation}
+\begin{equation}\label{eq:c-sparse}
 C_{\sf sparse} = S^{-1},
@@
 The fourth estimator $C_{\sf sparse+latent}$, combines the assumptions of the previous two.
-The underlying structure arises from sparse interactions between the recorded neurons and latent factors driving their activity (Fig.~\ref{fig:02}\,D).
+The underlying structure arises from sparse interactions between the recorded neurons and latent factors driving their activity (\figref{02}{D}).
 The estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
-\begin{equation}
+\begin{equation}{eq:c-sl}
 C_{\sf sparse+latent} = (S - LL^\T)^{-1}
@@
 
-\subsection*{Simulation}
-To illustrate the performance of the four regularized estimators, we constructed four model populations of $p=50$~neurons each with each matching the correlation structures of a different target estimate (Fig.~\ref{fig:02} Row 2).
-Sample correlation matrices calculated from samples of size $n=1000$ contain substantial sampling noise (Fig.~\ref{fig:02) Row 3) that makes the correlation structure less obvious.
+\paragraph{Simulation}
+To illustrate the performance of the four regularized estimators, we constructed four model populations of $p=50$~neurons each with each matching the correlation structures of a different target estimate (\figref{02}{Row 2}).
+Sample correlation matrices calculated from samples of size $n=1000$ contain substantial sampling noise (\figref{02}{Row 3}) that makes the correlation structure less obvious.
 
@@
 In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (Figure \ref{fig:02}, Row 5).
-For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth: $\ell_{C,C_{\sf diag}}$, $\ell_{C,C_{\sf factor}}$, $\ell_{C,C_{\sf sparse}}$, and $\ell_{C,C_{\sf sparse+latent}}$.
-Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.
+For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth.
+Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.[22]
 
@@
 
-\subsection*{Covariance estimation in neural data}
+\paragraph{Covariance estimation in neural data}
 We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005,Katona:2012,Cotton:2013}.
@@
 Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (Fig.~\ref{fig:03}\,B and E) in the visual cortex on the contralateral side to the stimulus.
-With acousto-optic modulator (AOD) 3D steering of the laser, the located cells were imaged at high sampling rates with concurrent motion detection.
-This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:03}\,C).
-After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:03}\,C and D).
-Only cells that produced detectable calcium activity were included in the analysis.
-The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (Fig.~\ref{fig:03}\,F and G).
+Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Ecker:2013}.
+[23]This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
+After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}).
+Only cells that produced detectable calcium activity were included in the analysis[24] (See Methods).[25] The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}).
 The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.
-Besides the high-repetition stimulus protocol used to estimate noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cell (Fig.~\ref{fig:03}\,E).
+Besides the high-repetition stimulus protocol used to estimate noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cell (\figref{03}{E}).
 Overall, 31 imaged sites from 24 animals were included in the study.
@@
 We therefore hypothesized that regularized estimates capturing a sparse partial correlation structure, and/or interactions with latent units would be the most efficient.
-At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), giving a potential advantage to estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$).
+At the same time, average correlations were relatively small (\figref{03}{G}), giving a potential advantage to estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$).
 Finally, correlation structure could possibly be best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$).
@@
 
-\subsection*{Correlation structure and circuit architecture}
+\paragraph{Correlation structure and circuit architecture}
 Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites.
-By design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
-If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations.
-In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
+By design, the sparse+latent estimator finds the optimal[26] balance between a sparse network of partial correlations and shared common latent units.
+If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations[27].
+[28]In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
 
@@
 Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations.
-The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G).
-The low-rank component had rank 17.
+[29]The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G).
+The low-rank component had rank 17.[30]
 
@@
 Indeed, the absence of correlations between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient.
-Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.
+[31]Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.
 
@@
 We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}).
-The average partial correlations fell more rapidly with difference in preferred orientation (Fig.\;\ref{fig:07}\,A and D) and lateral displacements at equal depths ($\pm 25\;\mu$m) (Fig.\;\ref{fig:07}\,B and E), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (Fig.\;\ref{fig:07}\,B and E).
+The average partial correlations fell more rapidly with difference in preferred orientation (Fig.\;\ref{fig:07}\,A and D) and lateral displacements at equal depths ($\pm 25\;\mu$m) (Fig.\;\ref{fig:07}\,B and E), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (Fig.\;\ref{fig:07}\,B and E).[32]
 
@@
 \section*{Discussion}
-\subsection*{Functional and anatomical connectivity}
+\paragraph{Functional and anatomical connectivity}
 Systems neuroscience is the quest to discover principles whereby anatomical organization of neural circuits gives rise to their function.
-This line of inquiry requires descriptions of the activity of neural circuits that have the best chance of correspondence to the underlying anatomical organization.
+This line of inquiry requires descriptions of the activity of neural circuits that have the best chance [33]of correspondence to the underlying anatomical organization.
 
@@
 
-Although there these concepts overlap, \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or variability in stimulus responses,  not directly driven by external stimuli, and reflecting internal, recurrent connectivity in neural circuits.
-This line of inquiry ascribes greater significance to the brain's intrinsic activity (often dismissed as `noise') \cite{Yuste:2005}.
-Relatively little progress has been made in understanding the role of functional connectivity and its relationship to the anatomical connectivity.
+Although there these concepts [34]overlap[35], \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or variability in stimulus responses,  not directly driven by external stimuli, and reflecting internal, recurrent connectivity in neural circuits.
+This line of inquiry ascribes greater significance to the brain's intrinsic activity (often dismissed as `noise') \cite{Yuste:2005}[36].
+However, relatively little progress has been made in understanding the role of functional connectivity and its relationship to the anatomical connectivity.
 
@@
 When measured on temporal scales of milliseconds, sharp peaks or troughs in the cross-correlogram between pairs of neurons at latencies typical of synaptic transmission provide evidence of monosynaptic connection between pairs of cells \cite{Moore:1970,Alonso:1998,Denman:2013} (although other processes can also generate sharp peaks in the cross-correlogram while many synaptic connections may not produce such signatures).
-Without targeted stimulation of specific cells or measurement of intracellular signals, such analysis remains the best method for detecting probable monosynaptic connections in vivo.
-When measured at longer temporal scales, neural correlations reflect activations involving large ensembles of neurons and may reflect direct causal anatomical interactions, indirect causal interactions, or complex network-wide emergent properties with little correspondence to anatomical connectivity.
-These slower correlations correspond more closely to the computation performed by the circuit than to its anatomical organization.
+Without targeted stimulation of specific cells or measurement of intracellular signals[37], such analysis remains the best method for detecting probable monosynaptic connections in vivo.
+When measured at longer temporal scales, neural correlations reflect activations involving large ensembles of neurons and may reflect direct causal anatomical interactions, indirect causal interactions, or complex network-wide emergent properties with little correspondence to local ???
+anatomical connectivity.
+These slower correlations correspond more closely to the computation [38]performed by the circuit than to its anatomical organization.
 For this reason, firing rate correlations at scales of tens to hundreds milliseconds have been studied for their effect on coding under the assumption of rate codes and linear downstream decoder \cite{Averbeck:2006} with little interest in anatomy.
@@
 
-\subsection*{Extraction of the correlation structure}
+\paragraph{Extraction of the correlation structure}
 The word `structure' implies a succinct, simplified representation that captures only the essential, significant features.
@@
 Intuitively, this suggests that many low interactions are as surprising or significant considering the common correlations with other neurons.
-The analysis of sample correlations does not detect surprisingly low correlations.
+The analysis of sample correlations does not ascribe any significance to correlation coefficients that are surprisingly low, based on their common correlations with the other cells.[39]
 
-\subsection*{Functional connectivity from probabilistic models}
+\paragraph{Functional connectivity from probabilistic models}
 Increasingly, functional connectivity is inferred by fitting probabilistic models such as various forms of multivariate Gaussian models, linear-nonlinear models, Generalized Linear Models \cite{Pillow:2008}, and maximum entropy models \cite{Schneidman:2006,Tkacik:2006,Tang:2008,Shlens:2009}.
@@
 Probabilistic models allow inferring latent variables to account for the activity of unobserved circuitry.
+[40]
 
-\subsection*{Model selection}
+\paragraph{Model selection}
 Various estimation methods infer very different structures of functional connectivity from the same data.
@@
 
-In simulation, we showed that, more constrained, simpler models often outperformed complex models in cross-validation studies even when their structure was differed from the true structure (Fig.~\ref{fig:02}\;row 5), only because these more constrained models were less sensitive to sampling noise.
-Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process.
+However, in simulation, we showed that, more constrained, simpler models often outperformed complex models in cross-validation studies even when their structure was differed from the true structure (Fig.~\ref{fig:02}\;row 5), only because these more constrained models were less sensitive to sampling noise.
+[41]Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process.
 For example, a properly sparsified pairwise Ising model could outperform the Reliable Network model.
-Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.
+Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.[42]
 
 In our study we selected between families of models based on their cross-validation performance.
-The winning model family was more flexible than the others and therefore more susceptible to noise.The fact that it still performed better indicated that the additional features were required.
+The winning model family was more flexible than the others and therefore more susceptible to noise.[43] The fact that it still performed better indicated that the additional features were required.
 We conclude that under these circumstances, the selected model can be asserted to better represent the true structure of the data generating process, at least in comparison to the other three models.
@@
 We defend the practice of evaluating and optimizing probabilistic models with respect to their parameters that are more relevant for functional connectivity rather than maximizing their cross-validated log likelihoods, which are sensitive to all parameters.
-For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix.
+[44]For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix.
 We maintained that model B must be more relevant for studies of functional connectivity because correlations do reflect aspects of functional connectivity.
+[45]
 
@@
 For example, spatiotemporal Ising models, marginalized to reproduce the correlation matrix at the same temporal scale, would suggest an alternative correlation structure from the same data.
-At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
+[46]At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
 If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+lowrank}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity.
@@
 
-\subsection*{Big neural data}
+\paragraph{Big neural data}
 Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}.
@@
 
-\subsection*{Correspondence to anatomical organization}
+\paragraph{Correspondence to anatomical organization}
 If the ultimate goal of this line of inquiry to find descriptions of functional connectivity that are most informative about the anatomical and physiological organization of the circuit, then the models must be directly validated according to their correspondence to the anatomical organization.
@@
 In this study we find that the structure of partial correlation of the optimal estimator had stronger dependencies on the separation in depth, lateral position, and orientation preferences than sample correlations (Fig.~\ref{fig:07}\;A--C).
-The inferred positive and negative partial correlations too had distinct relationships to structure  (Fig.~\ref{fig:07}\;D--F).
+The inferred positive and negative partial correlations too had different structures COMPARED TO NOISE CORRELATIONS ??
+[47](Fig.~\ref{fig:07}\;D--F).
 Also the average partial correlations were more consistent between imaged sites than average sample correlations (Fig.~\ref{fig:06}\;A).
@@
 For example, labeled interneurons could be studied for their distinct connectivity patterns contrasted with pyramidal cells.
+I THIKN WE NEED ONE LAST SENTNECE HERE TO SAY THAT FUNCTIONAL CONNECTIVITY IF IT RELATED TO ANALTOMICAL IS IMPORTANT SINCE IT CAN TELL US HOW CIRCUITS INTERACT TO COMPUTE MENTION EXAMPLE OF STATES, SLEEP HERE ETC.
+I.E.
+GROUPS CELLS MAY BE CONNECTED BUT ONLY INTERACTING IN STATE A AND NOT B.
 
@@
 % authors may use "Analysis" 
-\subsection*{Data Acquisition}
+\paragraph{Data Acquisition}
 Fast 3D two-photon imaging of calcium signals \emph{in vivo} was performed as described in \cite{Cotton:2013}
-\subsection*{Covariance Estimation}
-TODO
-\subsection*{Cross-validation}
-TODO
-\subsection*{Simulation}
+
+\paragraph{Cross-validation}
+We used conventional 10-fold cross validation to measure cross-validation loss (Eq.~\ref{eq:cv-loss}).
+Briefly, each recording was split into 30 blocks of equal duration.
+The 30 blocks were then grouped randomly into 10 datasets with 3 blocks in each.
+This procedure ensured sufficient independence between the 10 datasets while still ensuring that each dataset included data from different parts of the recording.
+Then, each dataset was used as the testing dataset with the rest of the data used for estimating the covariance matrix.
+
+Each regularized estimator had one or two hyperparameters.
+We estimated the optimal values of the hyperparameters independently in each of the 10 folds by \emph{nested cross-validation}: The 10-fold cross-validation procedure was repeated internally within each training dataset for given hyperparameter values.
+The search for optimal hyperparameter values consistent of two phases: random search to find a good starting point and pattern search to find the global minimum.
+
+\paragraph{Covariance Estimation}
+Each regularized estimator starts with the sample covariance matrix $C_{\sf 0}$ computed from the training dataset and calculats their low-dimensional approximation called the target estimate.
+In the last two estimators ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$), the target estimate is sufficiently flexible and is used directly.
+The first two estimators ($C_{\sf diag}$ and $C_{\sf factor}$) are computed as linear mixture (shrinkage) of $C_{\sf 0}$ with the target estimate.
+
+The diagonal matrix $V_{\sf 0}=C_{\sf 0}\circ I$ contains the sample variances.
+The mean sample variance is computed as the normalized matrix trace of the sample covariance matrix:  $v_{\sf 0} = \frac 1 p \Tr(C_{\sf 0})$.
+
+In estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag}), the diagonal matrix of regularized variances $V$ is estimated by  linearly shrinking of $V_{\sf 0}$ toward the mean variance with shrinkage intensity $\alpha$:
+\begin{equation}
+V = V_0 + \alpha \left(v_{\sf 0} I - V_{\sf 0}\right)
+\end{equation}
+
+Then the overall estimate is computed by linearly shrinking the sample covariance matrix toward $V$ with shrinkage intensity $\lambda$:
+\begin{equation}
+C_{\sf diag} = C_{\sf 0} + \lambda \left[V - C_{\sf 0}\right] 
+\end{equation}
+The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean square error whereas we optimized them with respect to the loss function in Eq.~\ref{eq:loss}.
+
+Estimator $C_{\sf factor}$ uses the factor analysis approximation of $C_{\sf 0}$:
+
+\paragraph{Simulation}
 TODO
@@
 \end{document}
+
+[1]Here we need to define what we mean by low-dimensional ..
+I found this confusing ..
+for example complete independent population i.e.
+zero non-diagonal in covariance matrix would be low-dimensional according to your definition but in fact in neural space it will be a max-entropy gauss distribution for example with flat eigenspectrum ..
+most people would not think as this as low-dimensional ..
+[2]We need stronger finishing sentence !!!! more punchy and exciting 
+[3]Reference http://www.med.upenn.edu/mulab/ .. I think he is one of the main old figures in this area of correlations and architecture .. 
+[4]Reference ecker et al 2010 where we also show so called limited range correlations 
+[5]A bit abrupt transition here .. 
+[6]We need to be careful here since dense connectivity can give rise to low dim and high eigenvalues .. so we may have to rephrase this a bit . 
+[7]Just reference a bunch of papers that use PCA, FA etc to analyze correlation structure .. 
+[8]I would remove this sentence or expand it.  It is not informative.
+[9]Is there a better way to say this?  Perhaps Òpartial correlation is a measure of association between two neurons after the effects of all other neurons are taken into account.
+[10]Is it clear that the condition is on the state of the rest of the network.
+[11]Does this really give us a lot more generality than the multivariate normal?  IÕve never seen it used in practice.  We never use it here, so it may be best to skip it.
+[12]WouldnÕt we need the means as well as the interactions to define the model?
+
+[13]Why ?? this is not quite clear to me .. 
+[14]Usual where?  I would be more specific.  ÒThe estimator of the covariance most commonly used in neuroscience applicationsÉÓ 
+[15]Give reference about Stein paradox 
+[16]I am again a bit worried about the flow.
+The previous paragraph describes shrinkage.
+However, this is not illustrated here  Perhaps just saying in a brief sentence how this estimate is related would be sufficient.
+[17]This is a fairly weak statement.  Any way to make it stronger or more convincing, while still being correct?
+[18]This sentence does not seem grammatically correct ???
+[19]These names donÕt agree with what we had above.  We should probably try to use consistent terminology.
+[20]There is an implicit assumption here that $p$ is small.
+I donÕt think this is spelled out clearly.
+This may be confusing, unless the assumption is clearly stated.
+Same comment for Òaffect many cellsÓ.
+This is another implicit assumption about the factor model.
+These hold in our examples, but we donÕt really make clear what we mean by p small and Òaffect many cellsÓ.
+[21]This is imprecise.  Please explain if possible.
+[22]I think this can appear in the caption, since it is more of a detail about how the figures were generated.
+[23]You may want to state that this (i.e. motion) is crucial for analysis of correlations ..
+[24]You should probably state how many cells were excluded.  This should give an idea of what fraction of cells you captured.
+[25]Also what war exclusion criterion .. if in methods then state so and say (see methods)
+[26]What does optimal mean here? Optimal in the sense that it describes the data best?  
+[27]There are many mentiones of Òusual correlationsÓ .
+This is unclear.
+I would stick with something more descriptive like Òsample correlation estimateÓ.
+It is also important to note that it is the estimate of correlation structure that you are talkinga bout here.
+[28]I agree .. use sample correlation 
+[29]Say some numbers here . 
+[30]ShouldnÕt you give average numbers here across sites ??? or does this come later??
+[31]This sentence is confusing .. rephrase 
+[32]Sentence is confusing .. reword 
+[33]This is a bit confusing here .. not clear what you mean exactly ??
+[34]Which concepts ???
+[35]What about starting with ÒIn contrast, or on the other hand i.e. something to counter functional architecture ..Ó
+[36]???
+[37]????
+[38]isnÕt everything to some extent computation ??? why bring computation here ???
+[39]you mean after thresholding ???
+[40]this seems out of place ..
+we need to discuss GLMs etc and how they are different from what we do here and put everything within the same context i.e.
+statistical graphical models that can make predictions ..
+[41]this was true when we had low sample numbers right?
+But the question is as we are trying to learn larger structures i.e.
+more neurons then wont we suffer from the sample size i.e.
+we will like be in a regime where sample size is low ??
+[42]this is confusing ??? I am not sure I understand the point here ???
+[43]why is this the case ???
+isnÕt this the point of cross validation ???
+in other words the best model wins in having predictive power in cross validation scheme even if this does not mean that it necessarily reflects the true physical model that generated the data.
+I think we hope that it does but ultimately one has to do more experiments to proof that ..
+
+However, I guess what you are saying is that because the most complex model did better then the other simpler models are further away from the true physical generating model ???
+is this a true mathematical statementÉ it makes sense intuitively to me but how true is it ..
+[44]This is not clear what it means here .. sounds confusing to me 
+[45]This is also confusing É 
+[46]How do we know that for this data ????
+[47]I think ending the sentence with structure is the wrong word here

### 2013-12-23 plos.tex
rev29. Integrated Kreso's comments. Added Supp Figure 1.

@@
 However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.
-Estimation can be improved by regularization: biasing of the usual unbiased estimate toward a low-dimensional approximation[1].
+Estimation can be improved by regularization: biasing of the usual unbiased estimate toward a low-dimensional approximation.
 The amount of improvement is greatest when the approximation chosen as the target of regularization parsimoniously captures the dominant dependencies in the data.
@@
 Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.
-The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.[2]
+The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.
 \section*{Author Summary}
@@
 \paragraph{Neural correlations}
-Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.
-For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}.
-In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits[3].
-This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Arieli:1995,Chiu:2002,Kenet:2003,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b[4]}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.
+Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.
+For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
+In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.
+This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.
 
 However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.
-Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b,Moore:1970,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}.
+Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.
 
 \paragraph{Additional information in neural correlation matrices}
-Yet, a complete [5]correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
-In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998,Zohary:1994}.
+Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
+In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998, Zohary:1994}.
 In contrast, modern multineuronal recordings from increasingly large and dense populations of highly interconnected neurons allow estimation of various low-dimensional representations of the entire correlation matrix, which can enable inquiry into the nature of underlying physiological mechanisms and computational principles.
@@
 For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation.
-Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.[6] Low-rank components can be extracted using principal component or factor analysis.
-The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009[7]}.
-[8]
+Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.
+Low-rank components can be extracted using principal component or factor analysis.
+The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.
 
 Alternatively, the correlation matrix can be expressed through the partial pairwise correlations.
-The partial pairwise correlation is the linear correlation remaining[9] after accounting for correlations with all the other neurons.
+The partial pairwise correlation is the linear correlation remaining after accounting for correlations with all the other neurons.
 Partial correlations carry particular significance when the relationship between the variables is approximately linear.
-In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent[10].
+In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent.
 
@@
 This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005,Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
-In \emph{elliptical distributions} [11](those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies.
-A multivariate normal distribution specified by a weighted graph[12] of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
+In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies.
+A multivariate normal distribution specified by a weighted graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
 With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}.
-For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006,Tkacik:2006}.[13]
+For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006, Tkacik:2006}.
 
@@
 In the case of noise correlations, $\mu$ depends on the stimulus condition.
-The usual estimator [14]of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
+The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
 \begin{equation}
@@
 Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.
-Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate  [15]Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
+Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate.
+Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
 
-\paragraph{Illustration of estimation of a neural correlation matrix}[16]
+\paragraph{Illustration of estimation of a neural correlation matrix}
 To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
@@
 Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case.
-The estimator that is shown to dominate all others for a specific type of neural circuit may provide [17]the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
+The estimator that is shown to dominate all others for a specific type of neural circuit may provide the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
 Its structure may then be analyzed and related to the circuit's anatomical organization.
@@
 Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.
-[18]
 
@@
 \paragraph{Covariance estimation}
-We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' [19](\figref{02}{row 1}).
+We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (\figref{02}{row 1}).
 
@@
 has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $F$ has the structure that arises when correlated fluctuations in population activity are driven by a small number [20]of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
+The target estimate $F$ has the structure that arises when correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
 
 The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (\figref{ 02}{C}).
-This assumption is enforced by setting to zero an optimal[21] subset of the off-diagonal elements of the precision matrix.
+This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.
 Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{ 02}{C}).
@@
 For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth.
-Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.[22]
+Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.
 
@@
 Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Ecker:2013}.
-[23]This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
+This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
 After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}).
-Only cells that produced detectable calcium activity were included in the analysis[24] (See Methods).[25] The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}).
+Only cells that produced detectable calcium activity were included in the analysis (See Methods).
+The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}).
 The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.
@@
 Finally, correlation structure could possibly be best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$).
-Thus it was a priori unclear which estimator would perform best.
+Thus it was \emph{a priori} unclear which estimator would perform best.
 
@@
 Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites.
-By design, the sparse+latent estimator finds the optimal[26] balance between a sparse network of partial correlations and shared common latent units.
-If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations[27].
-[28]In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
+By design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
+If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations.
+In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
 
@@
 Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations.
-[29]The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G).
-The low-rank component had rank 17.[30]
+The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G).
+The low-rank component had rank 17.
 
@@
 A network was formed by connecting neurons whose total correlations exceed this threshold \cite{Golshani:2009,Malmersjo:2013}.
-There was fairly little overlap between the network of interactions revealed by such thresholding and those revealed by the sparse+latent estimator (Fig.\;\ref{fig:05}\,F).
-When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (Fig.\;\ref{fig:05}\,F, H, and I).
+There was fairly little overlap between the network of interactions revealed by such thresholding and those revealed by the sparse+latent estimator (\figref{05}{F}).
+When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (\figref{05}{F, H, and I}).
 In particular, many low sample correlations translated into negative interactions in the regularized estimate.
 Indeed, the absence of correlations between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient.
-[31]Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.
+Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.
 
-The average partial correlations  revealed by the sparse+latent estimator at all 31 sites were about 5 times lower than the sample correlations and  less variable across sites (Fig.\;\ref{fig:06}\,A).
-The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (Fig.\;\ref{fig:06}\,B and C).
-However, there was an inverse relationship between the number of latent units and the average node degree (Fig.\;\ref{fig:06}\,D).
+The average partial correlations revealed by the sparse+latent estimator at all 31 sites were about 5 times lower than the sample correlations and  less variable across sites (\figref{06}{A}).
+The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (\figref{06}{B and C}).
+However, there was an inverse relationship between the number of latent units and the average node degree (\figref{fig:06}{D}).
 Several sites, even with relatively large population sizes, had fairly few pairwise interactions and were dominated by latent units.
@@
 We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}).
-The average partial correlations fell more rapidly with difference in preferred orientation (Fig.\;\ref{fig:07}\,A and D) and lateral displacements at equal depths ($\pm 25\;\mu$m) (Fig.\;\ref{fig:07}\,B and E), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (Fig.\;\ref{fig:07}\,B and E).[32]
+The average partial correlations fell more rapidly with difference in preferred orientation (\figref{fig:07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).
 
@@
 The connectivity expressed by the partial correlations in the regularized estimate had different organization for positive and negative interactions.
-Positive interactions fell rapidly as a function of difference in preferred orientation (Fig.\;\ref{fig:07}\,G and J), lateral displacements (Fig.\;\ref{fig:07}\,H and K), and displacement in depth (Fig.\;\ref{fig:07}\,I and L).
-Negative interactions were much less selective (Fig.\;\ref{fig:07}\,G--L).
+Positive interactions fell rapidly as a function of difference in preferred orientation (\figref{07}{D}), lateral displacements (\figref{07}{E}), and displacement in depth (\figref{07}{F}).
+Negative interactions were much less selective (\figref{07}{D--F}).
 
@@
 Systems neuroscience is the quest to discover principles whereby anatomical organization of neural circuits gives rise to their function.
-This line of inquiry requires descriptions of the activity of neural circuits that have the best chance [33]of correspondence to the underlying anatomical organization.
+This line of inquiry requires expressive descriptions of the observed activity of neural circuits.
 
@@
 
-Although there these concepts [34]overlap[35], \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or variability in stimulus responses,  not directly driven by external stimuli, and reflecting internal, recurrent connectivity in neural circuits.
-This line of inquiry ascribes greater significance to the brain's intrinsic activity (often dismissed as `noise') \cite{Yuste:2005}[36].
+Although there these concepts overlap, \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or variability in stimulus responses,  not directly driven by external stimuli, and reflecting internal, recurrent connectivity in neural circuits.
+This line of inquiry ascribes greater significance to the brain's intrinsic activity (often dismissed as `noise') \cite{Yuste:2005}.
 However, relatively little progress has been made in understanding the role of functional connectivity and its relationship to the anatomical connectivity.
@@
 When measured on temporal scales of milliseconds, sharp peaks or troughs in the cross-correlogram between pairs of neurons at latencies typical of synaptic transmission provide evidence of monosynaptic connection between pairs of cells \cite{Moore:1970,Alonso:1998,Denman:2013} (although other processes can also generate sharp peaks in the cross-correlogram while many synaptic connections may not produce such signatures).
-Without targeted stimulation of specific cells or measurement of intracellular signals[37], such analysis remains the best method for detecting probable monosynaptic connections in vivo.
-When measured at longer temporal scales, neural correlations reflect activations involving large ensembles of neurons and may reflect direct causal anatomical interactions, indirect causal interactions, or complex network-wide emergent properties with little correspondence to local ???
-anatomical connectivity.
-These slower correlations correspond more closely to the computation [38]performed by the circuit than to its anatomical organization.
+Without targeted stimulation of specific cells or measurements of subthreshold intracellular events, such analysis remains the best method for detecting probable monosynaptic connections in vivo.
+When measured at longer temporal scales, neural correlations reflect activations involving large ensembles of neurons and may reflect direct causal anatomical interactions, indirect causal interactions, or complex network-wide emergent features of activity with little correspondence to the local anatomical connectivity.
+These slower correlations correspond more closely to the computation performed by the circuit than to its anatomical organization.
 For this reason, firing rate correlations at scales of tens to hundreds milliseconds have been studied for their effect on coding under the assumption of rate codes and linear downstream decoder \cite{Averbeck:2006} with little interest in anatomy.
@@
 This property has led some investigators to propose that sparse inverse approximation of correlations may serve as closer estimates of direct anatomical interactions between brain regions in imaging studies than correlations \cite{Varoquaux:2012,Ryali:2012}.
-However, since deviations from elliptical distribution disrupt the correspondence between partial correlations and linear dependencies \cite{Loh:2012}, it is not clear whether that would indeed be the case.
+However, since deviations from elliptical distribution disrupt the correspondence between partial correlations and linear dependencies \cite{Loh:2012}, it is not clear whether this is indeed  the case.
 Partial correlations are meaningful when most interacting variables are observed.
@@
 The network of sparse partial correlations suggests interactions between specific pairs whereas the low-rank represents emergent activity patterns involving collective activations of larger circuits or global fluctuations.
-The revealed network of partial interactions was substantially different from thresholded correlations (Fig.~\ref{fig:05}\,C, F, H, I).
-In particular, it shows that many low correlations correspond to strong negative interactions (Fig.\;\ref{fig:05}\,F).
+The revealed network of partial interactions was substantially different from that revealed by thresholding correlations (\figref{05}{C, F, H, I}).
+In particular, it shows that many low correlations correspond to strong negative interactions (\figref{05}{F}).
 Intuitively, this suggests that many low interactions are as surprising or significant considering the common correlations with other neurons.
-The analysis of sample correlations does not ascribe any significance to correlation coefficients that are surprisingly low, based on their common correlations with the other cells.[39]
+The analysis of sample correlations does not ascribe significance to correlation coefficients that are surprisingly low, based on their common correlations to other cells.
 
 \paragraph{Functional connectivity from probabilistic models}
-Increasingly, functional connectivity is inferred by fitting probabilistic models such as various forms of multivariate Gaussian models, linear-nonlinear models, Generalized Linear Models \cite{Pillow:2008}, and maximum entropy models \cite{Schneidman:2006,Tkacik:2006,Tang:2008,Shlens:2009}.
+Increasingly, functional connectivity is inferred by fitting probabilistic models such as various forms of multivariate Gaussian models, linear-nonlinear models, Generalized Linear Models \cite{Pillow:2008}, and maximum entropy models \cite{Schneidman:2006, Tkacik:2006, Tang:2008, Shlens:2009}.
 In probabilistic models, functional dependencies are defined by multiplicative terms comprising the model.
@@
 Probabilistic models allow inferring latent variables to account for the activity of unobserved circuitry.
-[40]
 
@@
 Various estimation methods infer very different structures of functional connectivity from the same data.
-Which structure should be preferred?
+What kind of structure expresses most relevant information?
 How does one determine which functional connectivity structure is best supported by the empirical data?
 The field of machine learning offers a rich toolbox of model selection criteria.
-Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion} (AIC), and \emph{Bayes (or Schwartz) Information Criterion} (BIC) are computationally efficient but rely on generally unsupportable assumptions about the data generating process.
+Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion}, and \emph{Bayes (or Schwartz) Information Criterion} (BIC) are computationally efficient but rely on generally unsupportable assumptions about the data generating process.
 The most assumption-free yet more computationally intensive are procedures based on cross-validation \cite{Arlot:2010} and, when models have hyperparameters, nested cross-validation to select their values.
@@
 
-However, in simulation, we showed that, more constrained, simpler models often outperformed complex models in cross-validation studies even when their structure was differed from the true structure (Fig.~\ref{fig:02}\;row 5), only because these more constrained models were less sensitive to sampling noise.
-[41]Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process.
+However, in simulation, we showed that, more constrained, simpler models often outperformed complex models in cross-validation studies even when their structure was differed from the true structure (\figref{02}{row 5}), only because these more constrained models were less sensitive to sampling noise.
+Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process.
 For example, a properly sparsified pairwise Ising model could outperform the Reliable Network model.
-Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.[42]
+Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.
 
 In our study we selected between families of models based on their cross-validation performance.
-The winning model family was more flexible than the others and therefore more susceptible to noise.[43] The fact that it still performed better indicated that the additional features were required.
+The winning model family was more flexible than the others and therefore more susceptible to noise.
+The fact that it still performed better indicated that the additional features were required.
 We conclude that under these circumstances, the selected model can be asserted to better represent the true structure of the data generating process, at least in comparison to the other three models.
@@
 In our study, we focused on estimating a specific parameter of the distribution (its covariance matrix) rather than fitting a specific parametric distribution.
-Of course, under the loss function (Eq.~\ref{eq:loss}) we selected, the covariance matrix estimation methods are equivalent to fitting a multivariate Gaussian model and selecting the best structure on the basis of cross-validated log-likelihood.
-However, the ranking of models was preserved under other loss functions (Fig.~\ref{fig:Supp02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.
+Under the loss function (Eq.~\ref{eq:loss}) we selected, the covariance matrix estimation methods are equivalent to fitting a multivariate Gaussian model and selecting the best structure on the basis of cross-validated log-likelihood.
+However, the ranking of models was preserved under other loss functions (Fig.~\ref{supp:02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.
 We defend the practice of evaluating and optimizing probabilistic models with respect to their parameters that are more relevant for functional connectivity rather than maximizing their cross-validated log likelihoods, which are sensitive to all parameters.
-[44]For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix.
+For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix.
 We maintained that model B must be more relevant for studies of functional connectivity because correlations do reflect aspects of functional connectivity.
-[45]
 
@@
 For example, spatiotemporal Ising models, marginalized to reproduce the correlation matrix at the same temporal scale, would suggest an alternative correlation structure from the same data.
-[46]At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
+At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
 If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+lowrank}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity.
@@
 
-In this study we find that the structure of partial correlation of the optimal estimator had stronger dependencies on the separation in depth, lateral position, and orientation preferences than sample correlations (Fig.~\ref{fig:07}\;A--C).
+In this study we find that the structure of partial correlation of the optimal estimator had stronger dependencies on the separation in depth, lateral position, and orientation preferences than sample correlations (\figref{07}{A--C}).
 The inferred positive and negative partial correlations too had different structures COMPARED TO NOISE CORRELATIONS ??
-[47](Fig.~\ref{fig:07}\;D--F).
-Also the average partial correlations were more consistent between imaged sites than average sample correlations (Fig.~\ref{fig:06}\;A).
+(\figref{07}{D--F}).
+Also the average partial correlations were more consistent between imaged sites than average sample correlations (\figref{06}{A}).
 Although these effects could conceivably results from trivial numerical artifacts, they are consistent with improved representation of functional connectivity by the proposed estimator.
@@
 
-Estimator $C_{\sf factor}$ uses the factor analysis approximation of $C_{\sf 0}$:
+Estimator $C_{\sf factor}$ (Eq.~\ref{eq:c-factor}) is the linear shrinkage the factor analysis approximation of $C_{\sf 0}$:
 
@@
 %\section*{Figure Legends}
+\newpage
 \section*{Supplementary Figures}
 
+\setcounter{figure}{0}
+
+\begin{figure}[!ht]
+    \begin{center}
+        \includegraphics{./figures/Figure-Supp01.pdf}
+    \end{center}
+    \caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators.}
+}
+\label{supp:01}
+\end{figure}
+
 \end{document}
-
-[1]Here we need to define what we mean by low-dimensional ..
-I found this confusing ..
-for example complete independent population i.e.
-zero non-diagonal in covariance matrix would be low-dimensional according to your definition but in fact in neural space it will be a max-entropy gauss distribution for example with flat eigenspectrum ..
-most people would not think as this as low-dimensional ..
-[2]We need stronger finishing sentence !!!! more punchy and exciting 
-[3]Reference http://www.med.upenn.edu/mulab/ .. I think he is one of the main old figures in this area of correlations and architecture .. 
-[4]Reference ecker et al 2010 where we also show so called limited range correlations 
-[5]A bit abrupt transition here .. 
-[6]We need to be careful here since dense connectivity can give rise to low dim and high eigenvalues .. so we may have to rephrase this a bit . 
-[7]Just reference a bunch of papers that use PCA, FA etc to analyze correlation structure .. 
-[8]I would remove this sentence or expand it.  It is not informative.
-[9]Is there a better way to say this?  Perhaps Òpartial correlation is a measure of association between two neurons after the effects of all other neurons are taken into account.
-[10]Is it clear that the condition is on the state of the rest of the network.
-[11]Does this really give us a lot more generality than the multivariate normal?  IÕve never seen it used in practice.  We never use it here, so it may be best to skip it.
-[12]WouldnÕt we need the means as well as the interactions to define the model?
-
-[13]Why ?? this is not quite clear to me .. 
-[14]Usual where?  I would be more specific.  ÒThe estimator of the covariance most commonly used in neuroscience applicationsÉÓ 
-[15]Give reference about Stein paradox 
-[16]I am again a bit worried about the flow.
-The previous paragraph describes shrinkage.
-However, this is not illustrated here  Perhaps just saying in a brief sentence how this estimate is related would be sufficient.
-[17]This is a fairly weak statement.  Any way to make it stronger or more convincing, while still being correct?
-[18]This sentence does not seem grammatically correct ???
-[19]These names donÕt agree with what we had above.  We should probably try to use consistent terminology.
-[20]There is an implicit assumption here that $p$ is small.
-I donÕt think this is spelled out clearly.
-This may be confusing, unless the assumption is clearly stated.
-Same comment for Òaffect many cellsÓ.
-This is another implicit assumption about the factor model.
-These hold in our examples, but we donÕt really make clear what we mean by p small and Òaffect many cellsÓ.
-[21]This is imprecise.  Please explain if possible.
-[22]I think this can appear in the caption, since it is more of a detail about how the figures were generated.
-[23]You may want to state that this (i.e. motion) is crucial for analysis of correlations ..
-[24]You should probably state how many cells were excluded.  This should give an idea of what fraction of cells you captured.
-[25]Also what war exclusion criterion .. if in methods then state so and say (see methods)
-[26]What does optimal mean here? Optimal in the sense that it describes the data best?  
-[27]There are many mentiones of Òusual correlationsÓ .
-This is unclear.
-I would stick with something more descriptive like Òsample correlation estimateÓ.
-It is also important to note that it is the estimate of correlation structure that you are talkinga bout here.
-[28]I agree .. use sample correlation 
-[29]Say some numbers here . 
-[30]ShouldnÕt you give average numbers here across sites ??? or does this come later??
-[31]This sentence is confusing .. rephrase 
-[32]Sentence is confusing .. reword 
-[33]This is a bit confusing here .. not clear what you mean exactly ??
-[34]Which concepts ???
-[35]What about starting with ÒIn contrast, or on the other hand i.e. something to counter functional architecture ..Ó
-[36]???
-[37]????
-[38]isnÕt everything to some extent computation ??? why bring computation here ???
-[39]you mean after thresholding ???
-[40]this seems out of place ..
-we need to discuss GLMs etc and how they are different from what we do here and put everything within the same context i.e.
-statistical graphical models that can make predictions ..
-[41]this was true when we had low sample numbers right?
-But the question is as we are trying to learn larger structures i.e.
-more neurons then wont we suffer from the sample size i.e.
-we will like be in a regime where sample size is low ??
-[42]this is confusing ??? I am not sure I understand the point here ???
-[43]why is this the case ???
-isnÕt this the point of cross validation ???
-in other words the best model wins in having predictive power in cross validation scheme even if this does not mean that it necessarily reflects the true physical model that generated the data.
-I think we hope that it does but ultimately one has to do more experiments to proof that ..
-
-However, I guess what you are saying is that because the most complex model did better then the other simpler models are further away from the true physical generating model ???
-is this a true mathematical statementÉ it makes sense intuitively to me but how true is it ..
-[44]This is not clear what it means here .. sounds confusing to me 
-[45]This is also confusing É 
-[46]How do we know that for this data ????
-[47]I think ending the sentence with structure is the wrong word here

### 2013-12-23 plos.tex
worked on the Methods section.

@@
 
-The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (\figref{ 02}{C}).
+The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (\figref{02}{C}).
 This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.
-Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{ 02}{C}).
+Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{02}{C}).
 The estimator takes the form
@@
 The estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
-\begin{equation}{eq:c-sl}
+\begin{equation}\label{eq:c-sl}
 C_{\sf sparse+latent} = (S - LL^\T)^{-1}
@@
 \paragraph{Simulation}
-To illustrate the performance of the four regularized estimators, we constructed four model populations of $p=50$~neurons each with each matching the correlation structures of a different target estimate (\figref{02}{Row 2}).
-Sample correlation matrices calculated from samples of size $n=1000$ contain substantial sampling noise (\figref{02}{Row 3}) that makes the correlation structure less obvious.
+To illustrate the performance of the four regularized estimators, we constructed four model populations of $p=50$~neurons each with each matching the correlation structures of a different target estimate (\figref{02}{\,Row 2}).
+Sample correlation matrices calculated from samples of size $n=1000$ contain substantial sampling noise (\figref{02}{\,Row 3}) that makes the correlation structure less obvious.
 
@@
 
-In this study, we adopted \emph{negative normal log-likelihood loss}\footnote{
-The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$ as
-\begin{equation*}
-    L(\Sigma|C_{\sf 0}) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_{\sf 0}}
-\end{equation*}
-}
-defined as
+In this study, we adopted \emph{negative normal log-likelihood loss} defined as
 \begin{equation}\label{eq:loss}
     \loss{C,\Sigma} = \frac 1 p\left[\ln \det C + \Tr(C^{-1}\Sigma)\right]
+\end{equation}
+The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$ as
+\begin{equation}
+    L(\Sigma|C_{\sf 0}) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_{\sf 0}}
 \end{equation}
@@
 \paragraph{Covariance estimation in neural data}
-We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005,Katona:2012,Cotton:2013}.
-We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (Fig.~\ref{fig:03}\,A) on one side of the visual field of anesthetized mice.
-Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (Fig.~\ref{fig:03}\,B and E) in the visual cortex on the contralateral side to the stimulus.
+We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005, Katona:2012, Cotton:2013}.
+We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (\figref{03}{A}) on one side of the visual field of anesthetized mice.
+Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E} in the visual cortex on the contralateral side to the stimulus.
 Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Ecker:2013}.
@@
 In this study, we sought to uncover the optimal low-dimensional approximation of the covariance matrix for efficient estimation of the true covariance matrices in cortical microcircuits on the temporal scale of 100--200 ms corresponding to the temporal resolution of spike rates from somatic calcium signals \cite{Cotton:2013}.
-We found that networks of partial correlations were more statistically efficient than low-rank approximations of correlation matrices (Fig.\;S1 E).
+We found that networks of partial correlations were more statistically efficient than low-rank approximations of correlation matrices (Fig.\;S\ref{supp:01}\;E).
 Furthermore, estimation became still more efficient when low-rank components were allowed to be combined with sparse partial correlations.
@@
 Under the loss function (Eq.~\ref{eq:loss}) we selected, the covariance matrix estimation methods are equivalent to fitting a multivariate Gaussian model and selecting the best structure on the basis of cross-validated log-likelihood.
-However, the ranking of models was preserved under other loss functions (Fig.~\ref{supp:02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.
+However, the ranking of models was preserved under other loss functions (Fig.~S\ref{fig:supp02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.
 We defend the practice of evaluating and optimizing probabilistic models with respect to their parameters that are more relevant for functional connectivity rather than maximizing their cross-validated log likelihoods, which are sensitive to all parameters.
@@
 
-The diagonal matrix $V_{\sf 0}=C_{\sf 0}\circ I$ contains the sample variances.
+The diagonal matrix $D_{\sf 0}=C_{\sf 0}\circ I$ contains the sample variances.
 The mean sample variance is computed as the normalized matrix trace of the sample covariance matrix:  $v_{\sf 0} = \frac 1 p \Tr(C_{\sf 0})$.
 
-In estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag}), the diagonal matrix of regularized variances $V$ is estimated by  linearly shrinking of $V_{\sf 0}$ toward the mean variance with shrinkage intensity $\alpha$:
+In estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag}), the diagonal matrix of regularized variances $D$ is estimated by  linearly shrinking of $D_{\sf 0}$ toward the mean variance with shrinkage intensity $\alpha$:
 \begin{equation}
-V = V_0 + \alpha \left(v_{\sf 0} I - V_{\sf 0}\right)
+D = (1-\alpha)D_0 + \alpha v_{\sf 0}I 
 \end{equation} 
-
-Then the overall estimate is computed by linearly shrinking the sample covariance matrix toward $V$ with shrinkage intensity $\lambda$:
-\begin{equation}
-C_{\sf diag} = C_{\sf 0} + \lambda \left[V - C_{\sf 0}\right] 
-\end{equation}
+Then the overall estimate is computed by linearly shrinking the sample covariance matrix toward $V$ with shrinkage intensity $\lambda$ (Eq.~\ref{eq:c-diag}).
 The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean square error whereas we optimized them with respect to the loss function in Eq.~\ref{eq:loss}.
 
-Estimator $C_{\sf factor}$ (Eq.~\ref{eq:c-factor}) is the linear shrinkage the factor analysis approximation of $C_{\sf 0}$:
+Estimator $C_{\sf factor}$ is the linear shrinkage the factor analysis approximation of $C_{\sf 0}$ (Eq.~\ref{eq:c-factor}).
+The target $F$ by minimizing the loss $\loss{F,C_{\sf}}$, which is equivalent to maximizing the Gaussian log-likelihood with respect to $C_{\sf 0}$.
+To solve this minimization problem, we used the expectation-maximization (EM) algorithm.
+
+Estimator $C_{\sf sparse}$ (Eq.~\ref{eq:c-sparse}) finds the sparse precision matrix $S$ that minimizes the loss $\loss{S^{-1},C_{\sf 0}}$.
+It is computed in two steps.
+First, the zero structure of the precision matrix is obtained by minimizing the $L_1$-loss:
+\begin{equation}
+S = \argmin\limits_{S^\prime \succ 0} \loss{{S^\prime }^{-1},C_{\sf 0}} + \lambda \|S^\prime\|_1
+\end{equation}
+where $S \succ 0$ is the constraint that $S$ be a positive definite matrix and $\|S\|_1$ is the element-wise $L_1$ norm of the matrix $S$.
+This formulation is known as \emph{graphical lasso} \cite{Friedman:2008}.
+To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm developed by \cite{Ma:2013}.
+Then, the $L_1$ penalty was lifted and only the non-zero elements of $S$ where fitted to minimize the loss function $\loss{S^{-1},C_{\sf 0}}$.
+
+Estimator $C_{\sf sparse+lowrank}$ (Eq.~\ref{eq:c-sl}) finds a precision matrix $S^*$ of dimensions $(p+d)\times(p+d)$, where $d$ is the number of latent units
+\begin{equation}
+S^*=
+\begin{pmatrix}
+S_{11} & S_{12} \\
+S_{12}^\T & S_{22}
+\end{pmatrix}
+\end{equation}
+Then the covariance matrix of the observed population is 
+\begin{equation}
+C = \left(S_{11}-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
+\end{equation}
+We require that $S_{11}$ be sparse. Rather than estimate $S_{12}$ and $S_{22}$ individually, we estimate the $p\times d$ matrix $L$ such that $S_{12}S_{22}^{-1}S_{12}=LL^\T$.
 
@@
 \section*{Supplementary Figures}
-
 \setcounter{figure}{0}
-
 \begin{figure}[!ht]
@@
 \end{figure}
-
 \end{document}

### 2013-12-23 plos.tex
Rev30 -- nearly completed Methods. About to review with Kreso.

@@
 As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.
-
 \section*{Introduction}
@@
 Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
-\begin{equation}
+\begin{equation}\label{eq:precision}
     R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
@@
 Similarly, the matrix of partial correlations $P$ is computed as the negative normalized version of the \emph{precision matrix} (inverse covariance matrix) $C^{-1}$:
-\begin{equation}
+\begin{equation}\label{eq:partial}
     P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
@@
 Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.
-Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate.
-Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
+Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate  Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
 
@@
 \end{figure}
-
 \paragraph{Model selection}
@@
 Its structure may then be analyzed and related to the circuit's anatomical organization.
-
 \paragraph{Summary of findings}
@@
 % Results and Discussion can be combined.
-
 \paragraph{Covariance estimation}
@@
 
-The target of the second regularized estimator $C_{\sf factor}$, is the factor model $F =  L L^\T + \Psi$ with $d$ factors.
-Therefore $L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\Psi$ contains the independent variances of each neuron.
+The target of the second regularized estimator $C_{\sf factor}$, is the factor model $LL^\T + \Psi$, where $L$ is the $p\times d$ matrix of \emph{factor loadings}, where $d$ is the number of factors, and $\Psi$ is the diagonal matrix of independent variances.
 The estimator,
 \begin{equation}\label{eq:c-factor}
-C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda F,
+C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (LL^\T + \Psi),
 \end{equation}
 has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $F$ has the structure that arises when correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
+The target estimate $LL^\T + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
 
@@
 Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E} in the visual cortex on the contralateral side to the stimulus.
-Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Ecker:2013}.
+Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}.
 This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
@@
 
-Although there these concepts overlap, \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or variability in stimulus responses,  not directly driven by external stimuli, and reflecting internal, recurrent connectivity in neural circuits.
+In contrast to functional architecture, \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or trial-to-trial variations,  not directly driven by external stimuli.
+Functional connectivity is thought to reflect internal, recurrent connectivity in neural circuits.
 This line of inquiry ascribes greater significance to the brain's intrinsic activity (often dismissed as `noise') \cite{Yuste:2005}.
@@
 \paragraph{Cross-validation}
-We used conventional 10-fold cross validation to measure cross-validation loss (Eq.~\ref{eq:cv-loss}).
+To compare the performance of the estimator against each other, we used conventional 10-fold cross validation to measure cross-validation loss (Eq.~\ref{eq:cv-loss}).
 Briefly, each recording was split into 30 blocks of equal duration.
@@
 
-Each regularized estimator had one or two hyperparameters.
-We estimated the optimal values of the hyperparameters independently in each of the 10 folds by \emph{nested cross-validation}: The 10-fold cross-validation procedure was repeated internally within each training dataset for given hyperparameter values.
-The search for optimal hyperparameter values consistent of two phases: random search to find a good starting point and pattern search to find the global minimum.
+Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}.
+The outer loop evaluated the performance of the estimators with optimal values of the hyperparameters.
+The optimization of the hyperparameters was performed within the inner loop in two phases: random search to find a good starting point and pattern search to find the global minimum.
+The inner cross-validation loop subdivided the training dataset from the outer loop to perform 10-fold cross-validation in order to evaluate each choice of the hyperparameter values.
+Thus the size of the training dataset within the inner loop comprised 81% of the entire recording.
 
+When cross-validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.
+This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{02}{\;rows 3 and 4}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:05}--\ref{fig:07}).
 \paragraph{Covariance Estimation}
-Each regularized estimator starts with the sample covariance matrix $C_{\sf 0}$ computed from the training dataset and calculats their low-dimensional approximation called the target estimate.
-In the last two estimators ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$), the target estimate is sufficiently flexible and is used directly.
-The first two estimators ($C_{\sf diag}$ and $C_{\sf factor}$) are computed as linear mixture (shrinkage) of $C_{\sf 0}$ with the target estimate.
+Within the inner loop of cross-validation, covariance matrix estimation was performed with fixed hyperparameter values provided by the search algorithm.
+The computation of each regularized estimator only required the sample covariance matrix $C_{\sf 0}$ of the training dataset.
 
-The diagonal matrix $D_{\sf 0}=C_{\sf 0}\circ I$ contains the sample variances.
-The mean sample variance is computed as the normalized matrix trace of the sample covariance matrix:  $v_{\sf 0} = \frac 1 p \Tr(C_{\sf 0})$.
+Estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag})  used two hyperparameters: the covariance shrinkage intensity $\lambda \in [0,1]$ and variance shrinkage intensity $\alpha \in [0,1]$.
+The variances (the diagonal of $C_{\sf 0}$) were shrunk toward (linearly mixed with) their mean value:
+\begin{equation}
+D = (1-\alpha)C_{\sf 0}\circ I + \alpha \frac 1 p \Tr(C_{\sf 0}) I
+\end{equation}
+Then the diagonal matrix $D$ was used as the target of covariance shrinkage (Eq.~\ref{eq:c-diag}) to produce the final regularized estimate.
 
-In estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag}), the diagonal matrix of regularized variances $D$ is estimated by  linearly shrinking of $D_{\sf 0}$ toward the mean variance with shrinkage intensity $\alpha$:
+The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean squared error whereas we optimized them with respect to the loss function in Eq.~\ref{eq:loss}.
+
+Estimator $C_{\sf factor}$ used two hyperparameters: the number of latent factors $d$ and the shrinkage intensity $\lambda \in [0, 1]$.  
+The $p\times d$ factor loading matrix $L$ and individual variances $\Psi$ were computed by solving the minimization problem
 \begin{equation}
-D = (1-\alpha)D_0 + \alpha v_{\sf 0}I 
+(L,\Psi) = \argmin\limits_{\tilde L,\tilde\Psi} \loss{\tilde L \tilde L^\T + \tilde\Psi,C_{\sf 0}},
 \end{equation}
-Then the overall estimate is computed by linearly shrinking the sample covariance matrix toward $V$ with shrinkage intensity $\lambda$ (Eq.~\ref{eq:c-diag}).
-The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean square error whereas we optimized them with respect to the loss function in Eq.~\ref{eq:loss}.
+which we solved by an expectation-maximization (EM) algorithm.
+Under our chosen loss function (Eq.~\ref{eq:loss}), this is equivalent to maximum likelihood estimation of $L$ and $\Psi$ under the multivariate Gaussian distribution.
+The final regularized estimate is obtained by linear shrinkage toward the factor model (Eq.~\ref{eq:c-factor}).
 
-Estimator $C_{\sf factor}$ is the linear shrinkage the factor analysis approximation of $C_{\sf 0}$ (Eq.~\ref{eq:c-factor}).
-The target $F$ by minimizing the loss $\loss{F,C_{\sf}}$, which is equivalent to maximizing the Gaussian log-likelihood with respect to $C_{\sf 0}$.
-To solve this minimization problem, we used the expectation-maximization (EM) algorithm.
+Estimator $C_{\sf sparse}$  has one hyperparameter $\lambda$, which regulates the sparsity of its precision matrix $S$.
+The precision matrix is computed in two steps: First, the zero structure $Z$ is determined by minimizing the $L_1$-penalized loss.
+\begin{equation}
+Z = \argmin\limits_{\tilde Z \succ 0} \loss{{\tilde Z}^{-1},C_{\sf 0}} + \lambda \|\tilde Z \|_1
+\end{equation}
+where $\tilde Z\succ 0$ denotes the constraint that $\tilde Z$ be a positive definite matrix and $\|\tilde Z\|_1$ is the element-wise $L_1$ norm of the matrix $\tilde Z$.
+This problem formulation is known as \emph{graphical lasso} \cite{Friedman:2008}.
+To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm developed by \cite{Ma:2013}.
+Then, after the zero structure was determined, the remaining coefficients were fitted without penalty:
+\begin{equation}
+S = \argmin\limits_{\tilde S \in Z^\sharp} \loss{\tilde S,C_{\sf 0}},
+\end{equation}
+where $Z^\sharp$ denotes the set of positive-definite matrices with zeros in all entries where entries of $Z$ equal zero.
+This step was also solved by ADMM.
+Then the final estimate is the inverse of $S$ (Eq.~\ref{eq:c-sparse}).
+Unlike $C_{\sf diag}$ and $C_{\sf factor}$, this estimator does not include linear shrinkage: the selection of the sparsity level provides sufficient flexibility to fine tune the regularization level.
 
-Estimator $C_{\sf sparse}$ (Eq.~\ref{eq:c-sparse}) finds the sparse precision matrix $S$ that minimizes the loss $\loss{S^{-1},C_{\sf 0}}$.
-It is computed in two steps.
-First, the zero structure of the precision matrix is obtained by minimizing the $L_1$-loss:
+Estimator $C_{\sf sparse+lowrank}$ has two hyperparameters: the number of latent units $d$ and the sparsity level $\lambda$.
+It estimates the larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.
 \begin{equation}
-S = \argmin\limits_{S^\prime \succ 0} \loss{{S^\prime }^{-1},C_{\sf 0}} + \lambda \|S^\prime\|_1
+S^\ast=
+\begin{pmatrix}
+S & S_{12} \\
+S_{12}^\T & S_{22}
+\end{pmatrix},
 \end{equation}
-where $S \succ 0$ is the constraint that $S$ be a positive definite matrix and $\|S\|_1$ is the element-wise $L_1$ norm of the matrix $S$.
-This formulation is known as \emph{graphical lasso} \cite{Friedman:2008}.
-To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm developed by \cite{Ma:2013}.
-Then, the $L_1$ penalty was lifted and only the non-zero elements of $S$ where fitted to minimize the loss function $\loss{S^{-1},C_{\sf 0}}$.
-
-Estimator $C_{\sf sparse+lowrank}$ (Eq.~\ref{eq:c-sl}) finds a precision matrix $S^*$ of dimensions $(p+d)\times(p+d)$, where $d$ is the number of latent units
-\begin{equation}
-S^*=
-\begin{pmatrix}
-S_{11} & S_{12} \\
-S_{12}^\T & S_{22}
-\end{pmatrix}
-\end{equation}
+where the $p\times p$ partition $S$ corresponds to the visible units and expresses their partial correlation structure, and $S_{12}$ and $S_{22}$ are of size $p\times d$ and $d\times d$, respectively.
 Then the covariance matrix of the observed population is 
 \begin{equation}
-C = \left(S_{11}-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
+C_{\sf sparse+lowrank} = (S^\ast)^{-1} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
 \end{equation}
-We require that $S_{11}$ be sparse. Rather than estimate $S_{12}$ and $S_{22}$ individually, we estimate the $p\times d$ matrix $L$ such that $S_{12}S_{22}^{-1}S_{12}=LL^\T$.
+The matrix $ S_{12}S_{22}^{-1}S_{12}^\T$ has rank $d$.
+Rather than searching for the optimal sparse structure of $S_{12}$ and $S_{22}$, an ill-posed problem, we estimated these components together as a single matrix $LL^\T$ of rank $d$, where $L$ is of size $p\times d$.
+The estimate is found in two steps. First, we use the ADMM algorithm to find the zero structure of $S$ by minimizing the $L_1$-penalized loss \cite{Chandrasekaran:2010,Ma:2013}:
+\begin{equation}
+(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L\tilde L^\T} + \|\tilde Z\|_1
+\end{equation}
+Then we find the sparse and low-rank components of the precision matrix by minimizing the loss 
+\begin{equation}
+(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L \tilde L^\T}
+\end{equation}
 
+The MATLAB code for these computations is made publically available at {\tt http://github.com/atlab/cov-est}.
+
+For plots in \figref{fig:05}{B and E}, \figref{fig:06}{A}, and \figref{fig:07}{A--C}, the regularized partial correlations were obtained by normalizing the precision matrix (Eq.~\ref{eq:partial}).
+The sparse and low-rank components in \figref{fig;05}{C} are the additive components of these normalized partial correlations.
+
+The sparse partial correlations 
 \paragraph{Simulation}
-TODO
 
 \section*{Acknowledgments}
-Genevera Allen
+We Genevera Allen
 

### 2013-12-23 plos.tex
rev31 - update of Discussion after conversation with Kreso.

@@
 Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.
-Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate  Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
+Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate.
+Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
 
@@
 We examined the structure of the sparse+latent estimate at individual sites (an example site is depicted in Fig.\;\ref{fig:03}).
-At these sites the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (Fig.~\ref{fig:05}\,A and D).
-However, the corresponding partial correlation matrices differed dramatically (Fig.~\ref{fig:05}\,B and E).
-The partial correlation was decomposed into sparse and low-rank components (Fig.\,\ref{fig:05}\,C).
+At these sites the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (\figref{05}{A and D}).
+However, the corresponding partial correlation matrices differed dramatically (\figref{05}{B and E}).
+The partial correlation was decomposed into sparse and low-rank components (\figref{ 05}{C}).
 Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations.
-The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G).
+The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}).
 The low-rank component had rank 17.
@@
 The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (\figref{06}{B and C}).
-However, there was an inverse relationship between the number of latent units and the average node degree (\figref{fig:06}{D}).
+However, there was an inverse relationship between the number of latent units and the average node degree (\figref{06}{D}).
 Several sites, even with relatively large population sizes, had fairly few pairwise interactions and were dominated by latent units.
@@
 Functional connectivity is thought to reflect internal, recurrent connectivity in neural circuits.
-This line of inquiry ascribes greater significance to the brain's intrinsic activity (often dismissed as `noise') \cite{Yuste:2005}.
-However, relatively little progress has been made in understanding the role of functional connectivity and its relationship to the anatomical connectivity.
+However, less experimental evidence has been garnered to for links between functional and anatomical connectivity.
 
 Neural correlations boast a rich history as indicators of functional connectivity.
-When measured on temporal scales of milliseconds, sharp peaks or troughs in the cross-correlogram between pairs of neurons at latencies typical of synaptic transmission provide evidence of monosynaptic connection between pairs of cells \cite{Moore:1970,Alonso:1998,Denman:2013} (although other processes can also generate sharp peaks in the cross-correlogram while many synaptic connections may not produce such signatures).
+When measured on temporal scales of milliseconds, sharp peaks or troughs in the cross-correlogram between pairs of neurons at latencies typical of synaptic transmission provide evidence of monosynaptic connection between pairs of cells \cite{Moore:1970, Alonso:1998, Denman:2013} (although other processes can also generate sharp peaks in the cross-correlogram while many monosynaptic connections may not produce such signatures).
 Without targeted stimulation of specific cells or measurements of subthreshold intracellular events, such analysis remains the best method for detecting probable monosynaptic connections in vivo.
+
 When measured at longer temporal scales, neural correlations reflect activations involving large ensembles of neurons and may reflect direct causal anatomical interactions, indirect causal interactions, or complex network-wide emergent features of activity with little correspondence to the local anatomical connectivity.
-These slower correlations correspond more closely to the computation performed by the circuit than to its anatomical organization.
-For this reason, firing rate correlations at scales of tens to hundreds milliseconds have been studied for their effect on coding under the assumption of rate codes and linear downstream decoder \cite{Averbeck:2006} with little interest in anatomy.
-Calcium imaging allows simultaneous measurements of the activity of large populations of cells \emph{in vivo} \cite{Katona:2012,Cotton:2013} and even of entire nervous systems \cite{Leung:2013,Ahrens:2013}, but its temporal bandwidth is limited to the scale of tens or hundreds milliseconds by the kinetics of calcium-sensitive dyes, limiting its use to slower temporal scales at which correspondence between functional and anatomical connectivity is less direct.
-
-Our study is part of the effort to find descriptions of functional connectivity with the greatest chance of correspondence to anatomical, causal organization of the circuit.
+Calcium imaging allows simultaneous measurements of the activity of large populations of cells \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even of entire nervous systems \cite{Leung:2013, Ahrens:2013}, but its temporal bandwidth is limited to the scale of tens or hundreds milliseconds by the kinetics of calcium-sensitive dyes, limiting its use to slower temporal scales at which correspondence between functional and anatomical connectivity is less direct.
 
@@
 
-The MATLAB code for these computations is made publically available at {\tt http://github.com/atlab/cov-est}.
+For plots in \figref{05}{B and E}, \figref{06}{A}, and \figref{07}{A--C}, the regularized partial correlations were obtained by normalizing the precision matrix (Eq.~\ref{eq:partial}).
+The sparse and low-rank components in \figref{05}{C} are the additive components of these normalized partial correlations.
 
-For plots in \figref{fig:05}{B and E}, \figref{fig:06}{A}, and \figref{fig:07}{A--C}, the regularized partial correlations were obtained by normalizing the precision matrix (Eq.~\ref{eq:partial}).
-The sparse and low-rank components in \figref{fig;05}{C} are the additive components of these normalized partial correlations.
+The sparse partial correlations computed for the display of connectivity in \figref{05}{F, G, H}, was computed by normalizing only the sparse component of the precision matrix still using Eq.~\ref{eq:partial}.
 
-The sparse partial correlations 
+The MATLAB code for these computations is available at {\tt http://github.com/atlab/cov-est}.
+
 \paragraph{Simulation}

### 2013-12-24 plos.tex
minor edits of Methods.

@@
 However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.
-Estimation can be improved by regularization: biasing of the usual unbiased estimate toward a low-dimensional approximation.
+Estimation can be improved by regularization: biasing of the usual unbiased estimate toward a low-dimensional approximation[1].
 The amount of improvement is greatest when the approximation chosen as the target of regularization parsimoniously captures the dominant dependencies in the data.
@@
 Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.
-The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.
+The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.[2]
 \section*{Author Summary}
@@
 For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
-In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.
-This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.
+In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits[3].
+This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ko:2011, Smith:2013b[4]}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.
 
@@
 \paragraph{Additional information in neural correlation matrices}
-Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
+Yet, a complete [5]correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
 In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998, Zohary:1994}.
@@
 For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation.
-Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.
-Low-rank components can be extracted using principal component or factor analysis.
-The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.
+Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.[6] Low-rank components can be extracted using principal component or factor analysis.
+The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009[7]}.
+[8]
 
 Alternatively, the correlation matrix can be expressed through the partial pairwise correlations.
-The partial pairwise correlation is the linear correlation remaining after accounting for correlations with all the other neurons.
+The partial pairwise correlation is the linear correlation remaining[9] after accounting for correlations with all the other neurons.
 Partial correlations carry particular significance when the relationship between the variables is approximately linear.
-In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent.
+In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent[10].
 
@@
 This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005,Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
-In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies.
-A multivariate normal distribution specified by a weighted graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
+In \emph{elliptical distributions} [11](those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies.
+A multivariate normal distribution specified by a weighted graph[12] of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
 With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}.
-For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006, Tkacik:2006}.
+For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006, Tkacik:2006}.[13]
 
@@
 Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.
-Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate.
-Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
+Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate  [15]Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
 
-\paragraph{Illustration of estimation of a neural correlation matrix}
+\paragraph{Illustration of estimation of a neural correlation matrix}[16]
 To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
@@
 Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
-The close similarity of the two estimates of the correlation matrix (\figref{ 01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
+The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
 Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth.
@@
 Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case.
-The estimator that is shown to dominate all others for a specific type of neural circuit may provide the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
+The estimator that is shown to dominate all others for a specific type of neural circuit may provide [17]the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
 Its structure may then be analyzed and related to the circuit's anatomical organization.
@@
 \paragraph{Covariance estimation}
-We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (\figref{02}{row 1}).
+We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' [19](\figref{02}{row 1}).
 
@@
 has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $LL^\T + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
+The target estimate $LL^\T + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number [20]of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
 
 The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (\figref{02}{C}).
-This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.
+This assumption is enforced by setting to zero an optimal[21] subset of the off-diagonal elements of the precision matrix.
 Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{02}{C}).
@@
 For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth.
-Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.
+Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.[22]
 
@@
 Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}.
-This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
+[23]This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
 After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}).
-Only cells that produced detectable calcium activity were included in the analysis (See Methods).
-The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}).
+Only cells that produced detectable calcium activity were included in the analysis[24] (See Methods).[25] The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}).
 The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.
@@
 Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites.
-By design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
-If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations.
-In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
+By design, the sparse+latent estimator finds the optimal[26] balance between a sparse network of partial correlations and shared common latent units.
+If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations[27].
+[28]In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
 
@@
 However, the corresponding partial correlation matrices differed dramatically (\figref{05}{B and E}).
-The partial correlation was decomposed into sparse and low-rank components (\figref{ 05}{C}).
+The partial correlation was decomposed into sparse and low-rank components (\figref{05}{C}).
 Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations.
-The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}).
-The low-rank component had rank 17.
+[29]The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}).
+The low-rank component had rank 17.[30]
 
@@
 We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}).
-The average partial correlations fell more rapidly with difference in preferred orientation (\figref{fig:07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).
+The average partial correlations fell more rapidly with difference in preferred orientation (\figref{07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).[32]
 
@@
 Systems neuroscience is the quest to discover principles whereby anatomical organization of neural circuits gives rise to their function.
-This line of inquiry requires expressive descriptions of the observed activity of neural circuits.
+This line of inquiry requires expressive descriptions of the observed activity of neural circuits.[33]
 
@@
 
-\paragraph{Functional connectivity from probabilistic models}
+\paragraph{Functional connectivity from probabilistic models}[34]
 Increasingly, functional connectivity is inferred by fitting probabilistic models such as various forms of multivariate Gaussian models, linear-nonlinear models, Generalized Linear Models \cite{Pillow:2008}, and maximum entropy models \cite{Schneidman:2006, Tkacik:2006, Tang:2008, Shlens:2009}.
@@
 Probabilistic models allow inferring latent variables to account for the activity of unobserved circuitry.
+[35]
 
@@
 The field of machine learning offers a rich toolbox of model selection criteria.
-Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion}, and \emph{Bayes (or Schwartz) Information Criterion} (BIC) are computationally efficient but rely on generally unsupportable assumptions about the data generating process.
+Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion[36]}, and \emph{Bayes (or Schwartz) Information Criterion} (BIC) are computationally efficient but rely on generally unsupportable assumptions[37] about the data generating process.
 The most assumption-free yet more computationally intensive are procedures based on cross-validation \cite{Arlot:2010} and, when models have hyperparameters, nested cross-validation to select their values.
@@
 However, in simulation, we showed that, more constrained, simpler models often outperformed complex models in cross-validation studies even when their structure was differed from the true structure (\figref{02}{row 5}), only because these more constrained models were less sensitive to sampling noise.
-Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process.
+[38]Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process.
 For example, a properly sparsified pairwise Ising model could outperform the Reliable Network model.
-Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.
+Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.[39]
 
 In our study we selected between families of models based on their cross-validation performance.
-The winning model family was more flexible than the others and therefore more susceptible to noise.
-The fact that it still performed better indicated that the additional features were required.
+The winning model family was more flexible than the others and therefore more susceptible to noise.[40] The fact that it still performed better indicated that the additional features were required.
 We conclude that under these circumstances, the selected model can be asserted to better represent the true structure of the data generating process, at least in comparison to the other three models.
@@
 We defend the practice of evaluating and optimizing probabilistic models with respect to their parameters that are more relevant for functional connectivity rather than maximizing their cross-validated log likelihoods, which are sensitive to all parameters.
-For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix.
+[41]For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix.
 We maintained that model B must be more relevant for studies of functional connectivity because correlations do reflect aspects of functional connectivity.
+[42]
 
@@
 For example, spatiotemporal Ising models, marginalized to reproduce the correlation matrix at the same temporal scale, would suggest an alternative correlation structure from the same data.
-At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
+[43]At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
 If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+lowrank}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity.
@@
 The inferred positive and negative partial correlations too had different structures COMPARED TO NOISE CORRELATIONS ??
-(\figref{07}{D--F}).
+[44](\figref{07}{D--F}).
 Also the average partial correlations were more consistent between imaged sites than average sample correlations (\figref{06}{A}).
@@
 
-For plots in \figref{05}{B and E}, \figref{06}{A}, and \figref{07}{A--C}, the regularized partial correlations were obtained by normalizing the precision matrix (Eq.~\ref{eq:partial}).
-The sparse and low-rank components in \figref{05}{C} are the additive components of these normalized partial correlations.
-
-The sparse partial correlations computed for the display of connectivity in \figref{05}{F, G, H}, was computed by normalizing only the sparse component of the precision matrix still using Eq.~\ref{eq:partial}.
+The partial correlation matrix computed from the overall estimate, computed from $C_{\sf sparse+lowrank}$ according to Eq.~\ref{eq:partial}, includes the effects of interactions between the visible and latent units.
+This estimate of the partial correlations was used in analyses where sparsity was not required (\figref{05}{B and E}, \figref{06}{A}, and \figref{07}{A--C}).
+For analyses that required a sparse network (\figref{05}{F, G, H}, \figref{06}{B and D}, \figref{07}{D--F}) was computed using the extended precision matrix $S^\ast$:
+\begin{equation}
+P_{\sf sparse} = (S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
+\end{equation}
 
@@
 \end{document}
+
+[1]Here we need to define what we mean by low-dimensional ..
+I found this confusing ..
+for example complete independent population i.e.
+zero non-diagonal in covariance matrix would be low-dimensional according to your definition but in fact in neural space it will be a max-entropy gauss distribution for example with flat eigenspectrum ..
+most people would not think as this as low-dimensional ..
+[2]We need stronger finishing sentence !!!! more punchy and exciting 
+[3]Reference http://www.med.upenn.edu/mulab/ .. I think he is one of the main old figures in this area of correlations and architecture .. 
+[4]Reference ecker et al 2010 where we also show so called limited range correlations 
+[5]A bit abrupt transition here .. 
+[6]We need to be careful here since dense connectivity can give rise to low dim and high eigenvalues .. so we may have to rephrase this a bit . 
+[7]Just reference a bunch of papers that use PCA, FA etc to analyze correlation structure .. 
+[8]I would remove this sentence or expand it.  It is not informative.
+[9]Is there a better way to say this?  Perhaps Òpartial correlation is a measure of association between two neurons after the effects of all other neurons are taken into account.
+[10]Is it clear that the condition is on the state of the rest of the network.
+[11]Does this really give us a lot more generality than the multivariate normal?  IÕve never seen it used in practice.  We never use it here, so it may be best to skip it.
+[12]WouldnÕt we need the means as well as the interactions to define the model?
+
+[13]Why ?? this is not quite clear to me .. 
+[14]Usual where?  I would be more specific.  ÒThe estimator of the covariance most commonly used in neuroscience applicationsÉÓ 
+[15]Give reference about Stein paradox 
+[16]I am again a bit worried about the flow.
+The previous paragraph describes shrinkage.
+However, this is not illustrated here  Perhaps just saying in a brief sentence how this estimate is related would be sufficient.
+[17]This is a fairly weak statement.  Any way to make it stronger or more convincing, while still being correct?
+[18]This sentence does not seem grammatically correct ???
+[19]These names donÕt agree with what we had above.  We should probably try to use consistent terminology.
+[20]There is an implicit assumption here that $p$ is small.
+I donÕt think this is spelled out clearly.
+This may be confusing, unless the assumption is clearly stated.
+Same comment for Òaffect many cellsÓ.
+This is another implicit assumption about the factor model.
+These hold in our examples, but we donÕt really make clear what we mean by p small and Òaffect many cellsÓ.
+[21]This is imprecise.  Please explain if possible.
+[22]I think this can appear in the caption, since it is more of a detail about how the figures were generated.
+[23]You may want to state that this (i.e. motion) is crucial for analysis of correlations ..
+[24]You should probably state how many cells were excluded.  This should give an idea of what fraction of cells you captured.
+[25]Also what war exclusion criterion .. if in methods then state so and say (see methods)
+[26]What does optimal mean here? Optimal in the sense that it describes the data best?  
+[27]There are many mentiones of Òusual correlationsÓ .
+This is unclear.
+I would stick with something more descriptive like Òsample correlation estimateÓ.
+It is also important to note that it is the estimate of correlation structure that you are talkinga bout here.
+[28]I agree .. use sample correlation 
+[29]Say some numbers here . 
+[30]ShouldnÕt you give average numbers here across sites ??? or does this come later??
+[31]This sentence is confusing .. rephrase 
+[32]Sentence is confusing .. reword 
+[33]This intro is a bit too general.
+I donÕt think you need to start with Òsystems neuroscienceÓ.
+Perhaps just sturt the next paragraph with: ÒThe relation between anatomical organization and the patterns of neural activity is a fundamental question in neuroscience.Ó
+[34]I am not sure if this stands on its own.  It is a general statement about probabilistic models.  What does it say about our approach?  
+[35]this seems out of place ..
+we need to discuss GLMs etc and how they are different from what we do here and put everything within the same context i.e.
+statistical graphical models that can make predictions ..
+[36]These predate machine learning.
+[37]Why does BMC make assumption on the underlying process?  I agre with BIC and AIC, but BMC is pretty general.
+[38]this was true when we had low sample numbers right?
+But the question is as we are trying to learn larger structures i.e.
+more neurons then wont we suffer from the sample size i.e.
+we will like be in a regime where sample size is low ??
+[39]this is confusing ??? I am not sure I understand the point here ???
+[40]why is this the case ???
+isnÕt this the point of cross validation ???
+in other words the best model wins in having predictive power in cross validation scheme even if this does not mean that it necessarily reflects the true physical model that generated the data.
+I think we hope that it does but ultimately one has to do more experiments to proof that ..
+
+However, I guess what you are saying is that because the most complex model did better then the other simpler models are further away from the true physical generating model ???
+is this a true mathematical statementÉ it makes sense intuitively to me but how true is it ..
+[41]This is not clear what it means here .. sounds confusing to me 
+[42]This is also confusing É 
+[43]How do we know that for this data ????
+[44]I think ending the sentence with structure is the wrong word here

### 2013-12-26 plos.tex
completed Methods -- still needs some refining.

--- 2013-12-23 plos.tex
@@
 However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.
-Estimation can be improved by regularization: biasing of the usual unbiased estimate toward a low-dimensional approximation.
-The amount of improvement is greatest when the approximation chosen as the target of regularization parsimoniously captures the dominant dependencies in the data.
+Estimation can be improved by regularization: imposing some kind of structure on the estimate.
+The amount of improvement depends on how well the imposed structure accommodates the dominant dependencies in the data.
 Therefore, the selection of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.
@@
 However, the corresponding partial correlation matrices differed dramatically (\figref{05}{B and E}).
-The partial correlation was decomposed into sparse and low-rank components (\figref{ 05}{C}).
+The partial correlation was decomposed into sparse and low-rank components (\figref{05}{C}).
 Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations.
@@
 We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}).
-The average partial correlations fell more rapidly with difference in preferred orientation (\figref{fig:07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).
+The average partial correlations fell more rapidly with difference in preferred orientation (\figref{07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).
 
@@
 Without ways to restrict such models, they quickly become ill conditioned (e.g.
-\cite{Roudi:2009}.
+\cite{Roudi:2009}).
 Statistical models of population activity will need to have both correct mathematical forms of interactions and optimal ways of restricting them so that the total number of estimated parameters remains commensurate with the amount of measurements.
@@
 % authors may use "Analysis" 
-\paragraph{Data Acquisition}
-Fast 3D two-photon imaging of calcium signals \emph{in vivo} was performed as described in \cite{Cotton:2013}
+\paragraph{Surgery and two-photon Imaging}
+All procedures were conducted in accordance with the ethical guidelines of the National Institutes of Health and were approved by the Baylor College of Medicine IACUC.
+The surgical procedures and data acquisition were performed as descrbed in \cite{Cotton:2013}.
+Briefly, C57BL/6J mice (aged p40--60) were used.
+Anesthesia was initiated with isoflurane (3%) and the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
+A craniotomy was performed over the right primary visual area.
+Membrane-permeant calcium indicator Oregon Green 488 BAPTA-1 AM (OGB-1, Invitrogen) was loaded by bolus injection.
+The craniotomy was sealed using a glass coverslip secured with dental cement.
 
+Calcium imaging began 1 hour after dye injection.
+All imaging was performed using the 3D-RAMP two-photon microscope described in \cite{Cotton:2013}.
+First, A 3D stack was acquired and cells manually segmented.
+To collect calcium signals, the system repeated hopped between the selected neurons.
+\paragraph{Visual stimulus}
+Full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s.
+Two sets of stimuli were presented for each imaging site: the first to map directional tuning and the second to estimate noise correlations.
+Directional tuning was map using a pseudo-random sequence of drifting gratings at sixteen equally spaced directional of motion changing at 2 Hz for 3 min without blanks.
+The data for covariance estimation were collected during presentations of full-field drifting gratings with the same parameters as those used in directional tuning except only two directions (in 9 dataset) or five directions (in 22 datasets) were used and the presentations lasted 1 second and, separated by 1-second blanks.
+Each stimulus condition was presented at least 180 times.
+\paragraph{Data processing}
+All data were processed in MATLAB using the DataJoint data processing chain toolbox ({\tt https://github.com/datajoint/}) developed in our lab.
+
+The collected fluorescent traces were deconvolved to reconstruct the firing rates for each neuron.
+First, the first principal component was subtracted from the traces, which reduced common mode noise related to small movement artifacts and cardiovascular artifacts.
+The resulting traces were low-pass filtered below 0.1 Hz and downsampled to 20 Hz.
+Firing rates were estimated using a fast non-negative deconvolution algorithm \cite{Vogelstein:2010}.
+
+Orientation tuning was computed by fitting the mean firing rates in response to gratings of directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[(\cos(\phi-\theta)-1)/w\right] + c\exp\left[(-\cos(\phi-\theta)-1)/w\right]$ where $b\ge c$ are amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
+The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times.
+The $p$-value of the fit was computed as the fraction of permuted datasets for which the $R^2$ value of the tuning function fit exceeded that of the real data.
+Cells were considered tuned for $p$-values not exceeding $0.05$.
+
+For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 12 and 20 mins.
+Only traces whose average deconvolved signal was greater than 1\% of the median of the recorded population were included in the analysis.
 \paragraph{Cross-validation}
@@
 
-For plots in \figref{05}{B and E}, \figref{06}{A}, and \figref{07}{A--C}, the regularized partial correlations were obtained by normalizing the precision matrix (Eq.~\ref{eq:partial}).
-The sparse and low-rank components in \figref{05}{C} are the additive components of these normalized partial correlations.
-
-The sparse partial correlations computed for the display of connectivity in \figref{05}{F, G, H}, was computed by normalizing only the sparse component of the precision matrix still using Eq.~\ref{eq:partial}.
+The partial correlation matrix computed from the overall estimate, computed from $C_{\sf sparse+lowrank}$ according to Eq.~\ref{eq:partial}, includes the effects of interactions between the visible and latent units.
+This estimate of the partial correlations was used in analyses where sparsity was not required (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).
+For analyses that required a sparse network (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}) was computed using the extended precision matrix $S^\ast$:
+\begin{equation}
+P_{\sf sparse} = (S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
+\end{equation}
 
@@
 \paragraph{Simulation}
+For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units.
+The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{02}{\,row 2}.
+Samples were then generated by multivariate normal distributions with the respective true covariance matrices and estimated by each of the estimators.
 
 \section*{Acknowledgments}
-We Genevera Allen
-
-Eftychios Pnevmatikakis
+We thank Genevera Allen for a helpful discussion.  We thank Eftychios Pnevmatikakis for helpful suggestions and feedback.
 
@@
 \newpage
-\section*{Supplementary Figures}
+\section*{Supporting Information}
 \setcounter{figure}{0}
+
 \begin{figure}[!ht]
@@
 \end{figure}
+
 \end{document}

### 2013-12-27 plos.tex
added Supp Fig 02.

@@
 Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.
-The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.
+The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.[1]
 \section*{Author Summary}
@@
 For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
-In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.
-This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.
+In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits[2].
+This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ko:2011, Smith:2013b[3]}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.
 
@@
 \paragraph{Additional information in neural correlation matrices}
-Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
+Yet, a complete [4]correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
 In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998, Zohary:1994}.
@@
 For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation.
-Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.
-Low-rank components can be extracted using principal component or factor analysis.
-The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.
+Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.[5] Low-rank components can be extracted using principal component or factor analysis.
+The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009[6]}.
+[7]
 
 Alternatively, the correlation matrix can be expressed through the partial pairwise correlations.
-The partial pairwise correlation is the linear correlation remaining after accounting for correlations with all the other neurons.
+The partial pairwise correlation is the linear correlation remaining[8] after accounting for correlations with all the other neurons.
 Partial correlations carry particular significance when the relationship between the variables is approximately linear.
-In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent.
+In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent[9].
 
@@
 This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005,Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
-In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies.
-A multivariate normal distribution specified by a weighted graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
+In \emph{elliptical distributions} [10](those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies.
+A multivariate normal distribution specified by a weighted graph[11] of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
 With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}.
-For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006, Tkacik:2006}.
+For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006, Tkacik:2006}.[12]
 
@@
 In the case of noise correlations, $\mu$ depends on the stimulus condition.
-The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
+The usual estimator [13]of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
 \begin{equation}
@@
 Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate.
-Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
+[14]Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
 
-\paragraph{Illustration of estimation of a neural correlation matrix}
+\paragraph{Illustration of estimation of a neural correlation matrix}[15]
 To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
@@
 Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case.
-The estimator that is shown to dominate all others for a specific type of neural circuit may provide the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
+The estimator that is shown to dominate all others for a specific type of neural circuit may provide [16]the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
 Its structure may then be analyzed and related to the circuit's anatomical organization.
@@
 Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.
+[17]
 
@@
 \paragraph{Covariance estimation}
-We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (\figref{02}{row 1}).
+We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' [18](\figref{02}{row 1}).
 
@@
 has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $LL^\T + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
+The target estimate $LL^\T + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number [19]of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
 
 The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (\figref{02}{C}).
-This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.
+This assumption is enforced by setting to zero an optimal[20] subset of the off-diagonal elements of the precision matrix.
 Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{02}{C}).
@@
 \begin{equation*}
-    \E{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E{C_{\sf 0}^\prime}}=\loss{C,\Sigma}
+    \E[C]{\E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}}=\E[C]{\loss{C,\E[C_{\sf 0}]{C_{\sf 0}^\prime}}}=\E[C]{\loss{C,\Sigma}}
 \end{equation*}
 
-This property does not hold for some other popular loss functions such as Stein's entropy loss, for example, which prevents their substitution with corresponding validation losses.
-However, various other loss functions do have this property and could have been used.
+The last result expression is known as the \emph{estimator risk} which serv
+
+The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses, for example, which prevents their substitution with corresponding validation losses.
 } 
@@
 For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth.
-Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.
+Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.[21]
 
@@
 Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}.
-This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
+[22]This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
 After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}).
-Only cells that produced detectable calcium activity were included in the analysis (See Methods).
-The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}).
+Only cells that produced detectable calcium activity were included in the analysis[23] (See Methods).[24] The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}).
 The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.
@@
 Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites.
-By design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
-If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations.
-In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
+By design, the sparse+latent estimator finds the optimal[25] balance between a sparse network of partial correlations and shared common latent units.
+If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations[26].
+[27]In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
 
@@
 Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations.
-The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}).
-The low-rank component had rank 17.
+[28]The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}).
+The low-rank component had rank 17.[29]
 
@@
 Indeed, the absence of correlations between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient.
-Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.
+[30]Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.
 
@@
 We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}).
-The average partial correlations fell more rapidly with difference in preferred orientation (\figref{07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).
+The average partial correlations fell more rapidly with difference in preferred orientation (\figref{07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).[31]
 
@@
 Systems neuroscience is the quest to discover principles whereby anatomical organization of neural circuits gives rise to their function.
-This line of inquiry requires expressive descriptions of the observed activity of neural circuits.
+This line of inquiry requires expressive descriptions of the observed activity of neural circuits.[32]
 
@@
 
-\paragraph{Functional connectivity from probabilistic models}
+\paragraph{Functional connectivity from probabilistic models}[33]
 Increasingly, functional connectivity is inferred by fitting probabilistic models such as various forms of multivariate Gaussian models, linear-nonlinear models, Generalized Linear Models \cite{Pillow:2008}, and maximum entropy models \cite{Schneidman:2006, Tkacik:2006, Tang:2008, Shlens:2009}.
@@
 Probabilistic models allow inferring latent variables to account for the activity of unobserved circuitry.
+[34]
 
@@
 The field of machine learning offers a rich toolbox of model selection criteria.
-Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion}, and \emph{Bayes (or Schwartz) Information Criterion} (BIC) are computationally efficient but rely on generally unsupportable assumptions about the data generating process.
+Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion[35]}, and \emph{Bayes (or Schwartz) Information Criterion} (BIC) are computationally efficient but rely on generally unsupportable assumptions[36] about the data generating process.
 The most assumption-free yet more computationally intensive are procedures based on cross-validation \cite{Arlot:2010} and, when models have hyperparameters, nested cross-validation to select their values.
@@
 However, in simulation, we showed that, more constrained, simpler models often outperformed complex models in cross-validation studies even when their structure was differed from the true structure (\figref{02}{row 5}), only because these more constrained models were less sensitive to sampling noise.
-Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process.
+[37]Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process.
 For example, a properly sparsified pairwise Ising model could outperform the Reliable Network model.
-Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.
+Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.[38]
 
 In our study we selected between families of models based on their cross-validation performance.
-The winning model family was more flexible than the others and therefore more susceptible to noise.
-The fact that it still performed better indicated that the additional features were required.
+The winning model family was more flexible than the others and therefore more susceptible to noise.[39] The fact that it still performed better indicated that the additional features were required.
 We conclude that under these circumstances, the selected model can be asserted to better represent the true structure of the data generating process, at least in comparison to the other three models.
@@
 We defend the practice of evaluating and optimizing probabilistic models with respect to their parameters that are more relevant for functional connectivity rather than maximizing their cross-validated log likelihoods, which are sensitive to all parameters.
-For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix.
+[40]For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix.
 We maintained that model B must be more relevant for studies of functional connectivity because correlations do reflect aspects of functional connectivity.
+[41]
 
@@
 For example, spatiotemporal Ising models, marginalized to reproduce the correlation matrix at the same temporal scale, would suggest an alternative correlation structure from the same data.
-At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
+[42]At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
 If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+lowrank}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity.
@@
 The inferred positive and negative partial correlations too had different structures COMPARED TO NOISE CORRELATIONS ??
-(\figref{07}{D--F}).
+[43](\figref{07}{D--F}).
 Also the average partial correlations were more consistent between imaged sites than average sample correlations (\figref{06}{A}).
@@
 
-Orientation tuning was computed by fitting the mean firing rates in response to gratings of directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[(\cos(\phi-\theta)-1)/w\right] + c\exp\left[(-\cos(\phi-\theta)-1)/w\right]$ where $b\ge c$ are amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
+Orientation tuning was computed by fitting the mean firing rates in response to gratings of directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
 The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times.
@@
 \end{center}
-\caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators.}
+\caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators with respect to multivariate normal cross-validation loss.}
 }
@@
 
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics{./figures/Figure-Supp02.pdf}
+\end{center}
+\caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators with respect to quadratic cross-validation loss.}
+}
+\label{supp:02}
+\end{figure}
+
 \end{document}
+
+[1]We need stronger finishing sentence !!!! more punchy and exciting 
+[2]Reference http://www.med.upenn.edu/mulab/ .. I think he is one of the main old figures in this area of correlations and architecture .. 
+[3]Reference ecker et al 2010 where we also show so called limited range correlations 
+[4]A bit abrupt transition here .. 
+[5]We need to be careful here since dense connectivity can give rise to low dim and high eigenvalues .. so we may have to rephrase this a bit . 
+[6]Just reference a bunch of papers that use PCA, FA etc to analyze correlation structure .. 
+[7]I would remove this sentence or expand it.  It is not informative.
+[8]Is there a better way to say this?  Perhaps Òpartial correlation is a measure of association between two neurons after the effects of all other neurons are taken into account.
+[9]Is it clear that the condition is on the state of the rest of the network.
+[10]Does this really give us a lot more generality than the multivariate normal?  IÕve never seen it used in practice.  We never use it here, so it may be best to skip it.
+[11]WouldnÕt we need the means as well as the interactions to define the model?
+
+[12]Why ?? this is not quite clear to me .. 
+[13]Usual where?  I would be more specific.  ÒThe estimator of the covariance most commonly used in neuroscience applicationsÉÓ 
+[14]Give reference about Stein paradox 
+[15]I am again a bit worried about the flow.
+The previous paragraph describes shrinkage.
+However, this is not illustrated here  Perhaps just saying in a brief sentence how this estimate is related would be sufficient.
+[16]This is a fairly weak statement.  Any way to make it stronger or more convincing, while still being correct?
+[17]This sentence does not seem grammatically correct ???
+[18]These names donÕt agree with what we had above.  We should probably try to use consistent terminology.
+[19]There is an implicit assumption here that $p$ is small.
+I donÕt think this is spelled out clearly.
+This may be confusing, unless the assumption is clearly stated.
+Same comment for Òaffect many cellsÓ.
+This is another implicit assumption about the factor model.
+These hold in our examples, but we donÕt really make clear what we mean by p small and Òaffect many cellsÓ.
+[20]This is imprecise.  Please explain if possible.
+[21]I think this can appear in the caption, since it is more of a detail about how the figures were generated.
+[22]You may want to state that this (i.e. motion) is crucial for analysis of correlations ..
+[23]You should probably state how many cells were excluded.  This should give an idea of what fraction of cells you captured.
+[24]Also what war exclusion criterion .. if in methods then state so and say (see methods)
+[25]What does optimal mean here? Optimal in the sense that it describes the data best?  
+[26]There are many mentiones of Òusual correlationsÓ .
+This is unclear.
+I would stick with something more descriptive like Òsample correlation estimateÓ.
+It is also important to note that it is the estimate of correlation structure that you are talkinga bout here.
+[27]I agree .. use sample correlation 
+[28]Say some numbers here . 
+[29]ShouldnÕt you give average numbers here across sites ??? or does this come later??
+[30]This sentence is confusing .. rephrase 
+[31]Sentence is confusing .. reword 
+[32]This intro is a bit too general.
+I donÕt think you need to start with Òsystems neuroscienceÓ.
+Perhaps just sturt the next paragraph with: ÒThe relation between anatomical organization and the patterns of neural activity is a fundamental question in neuroscience.Ó
+[33]I am not sure if this stands on its own.  It is a general statement about probabilistic models.  What does it say about our approach?  
+[34]this seems out of place ..
+we need to discuss GLMs etc and how they are different from what we do here and put everything within the same context i.e.
+statistical graphical models that can make predictions ..
+[35]These predate machine learning.
+[36]Why does BMC make assumption on the underlying process?  I agre with BIC and AIC, but BMC is pretty general.
+[37]this was true when we had low sample numbers right?
+But the question is as we are trying to learn larger structures i.e.
+more neurons then wont we suffer from the sample size i.e.
+we will like be in a regime where sample size is low ??
+[38]this is confusing ??? I am not sure I understand the point here ???
+[39]why is this the case ???
+isnÕt this the point of cross validation ???
+in other words the best model wins in having predictive power in cross validation scheme even if this does not mean that it necessarily reflects the true physical model that generated the data.
+I think we hope that it does but ultimately one has to do more experiments to proof that ..
+
+However, I guess what you are saying is that because the most complex model did better then the other simpler models are further away from the true physical generating model ???
+is this a true mathematical statementÉ it makes sense intuitively to me but how true is it ..
+[40]This is not clear what it means here .. sounds confusing to me 
+[41]This is also confusing É 
+[42]How do we know that for this data ????
+[43]I think ending the sentence with structure is the wrong word here

### 2013-12-27 plos.tex
minor

--- 2013-12-26 plos.tex
@@
 Under the loss function (Eq.~\ref{eq:loss}) we selected, the covariance matrix estimation methods are equivalent to fitting a multivariate Gaussian model and selecting the best structure on the basis of cross-validated log-likelihood.
-However, the ranking of models was preserved under other loss functions (Fig.~S\ref{fig:supp02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.
+However, the ranking of models was preserved under other loss functions (Fig.~S\ref{supp:02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.
 We defend the practice of evaluating and optimizing probabilistic models with respect to their parameters that are more relevant for functional connectivity rather than maximizing their cross-validated log likelihoods, which are sensitive to all parameters.
@@
 
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics{./figures/Figure-Supp02.pdf}
+\end{center}
+\caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators with respect to quadratic cross-validation loss.}
+}
+\label{supp:02}
+\end{figure}
+
 \end{document}

### 2013-12-27 plos.tex
rev34. Made Supp Fig. 2 smaller by keeping only comparisons to C_{sparse+latent}

@@
 
-In this study, we sought most statistically efficient estimators of neural correlations in recordings from large, dense groups of cortical neurons.
+In this study, we sought most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.
 Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.
@@
 For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
-In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.
-This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.
+In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits and reflect interactions between neurons \cite{Gerstein:1964}.
+This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.
 
@@
 \paragraph{Additional information in neural correlation matrices}
-Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
+Despite these ambiguities, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
 In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998, Zohary:1994}.
@@
 For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation.
-Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.
+Such approximations are particularly suitable for capturing shared fluctuations across the recorded population, which may arise in local circuits or be externally driven.
 Low-rank components can be extracted using principal component or factor analysis.
@@
 This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005,Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
-In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies.
-A multivariate normal distribution specified by a weighted graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
+In multivariate normal distributions (or, more generally, \emph{elliptical distributions}) interactions are completely defined by linear effects and, therefore, partial correlations reflect conditional dependencies.
+A multivariate normal distribution specified by a weighted graph of non-zero partial correlations (along with the means and variances of individual neurons) is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
 With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}.
-For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006, Tkacik:2006}.
+For example, the coupling terms in pairwise Ising models do not directly correspond o the partial correlations \cite{Schneidman:2006, Tkacik:2006}.
 
@@
 In the case of noise correlations, $\mu$ depends on the stimulus condition.
-The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
+Usually, neural covariance matrices have been estimated as the \emph{sample covariance matrix} $C_{\sf 0}$ calculated from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
 \begin{equation}
@@
 
-Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.
-The sample estimate $C_{\sf 0}$ is unbiased ($\E{C_0}=\Sigma$) but, on average, falls far from $\Sigma$ due to its sensitivity to sampling noise.
+Estimation can be improved through \emph{regularization}: deliberately imposing some kind of structure on the estimate.
+The sample estimate $C_{\sf 0}$, which is usually used to estimated covariance matrices in neuroscience, is unbiased ($\E{C_0}=\Sigma$).
+However, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise.
 Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.
-Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate.
+Regularization strikes a favorable balance between bias and variability
+
+ A regularized estimate is produced by biasing (`shrinking’) the estimate from the unbiased estimate toward a selected low-dimensional approximation (`target’) of the unbiased estimate \cite{Schafer:2005, Bickel:2006}.
+The selection of the optimal target estimate and the optimal amount of shrinkage can be optimized based on the training data by analytical methods or by cross-validation.
+However, some regularization schemes only perform target selection (\emph{dimensionality reduction}) and replace the estimate without shrinkages whereas others only perform shrinkage given a single target estimate (\emph{shrinkage estimators}).
+
+Regularization can produce \emph{some} improvement even with an arbitrary target estimate (See Discussion).
 Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
@@
 Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case.
-The estimator that is shown to dominate all others for a specific type of neural circuit may provide the best descriptor of the low-dimensional representation of the correlation structure of the system of interest.
-Its structure may then be analyzed and related to the circuit's anatomical organization.
+The estimator that is shown to dominate all others for a specific type of neural circuit serves as an effective descriptor the circuit’s correlation structure.
+This structure may then be analyzed with respect to the circuit's anatomical organization.
 \paragraph{Summary of findings}
@@
 Only cells that produced detectable calcium activity were included in the analysis (See Methods).
-The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}).
+The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{03}{F and G}).
 The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.
@@
 Briefly, C57BL/6J mice (aged p40--60) were used.
-Anesthesia was initiated with isoflurane (3%) and the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
+Anesthesia was initiated with isoflurane (3\%) and the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
 A craniotomy was performed over the right primary visual area.
@@
 \paragraph{Data processing}
-All data were processed in MATLAB using the DataJoint data processing chain toolbox ({\tt https://github.com/datajoint/}) developed in our lab.
+All data were processed in MATLAB using the DataJoint data processing chain toolbox ({\tt datajoint.github.com}) first developed in our lab.
 
@@
 The inner cross-validation loop subdivided the training dataset from the outer loop to perform 10-fold cross-validation in order to evaluate each choice of the hyperparameter values.
-Thus the size of the training dataset within the inner loop comprised 81% of the entire recording.
+Thus the size of the training dataset within the inner loop comprised 81\% of the entire recording.
 

### 2013-12-30 plos.tex
minor

@@
 {\Large
-Improved estimation of neural correlations suggests detailed interactions in visual cortex
+Inferred partial neural correlations suggest detailed interactions in cortical circuits
 }
@@
 Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.
-The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.
+Because of its superior estimation performance, the inferred partial correlation structure obtained using these methods may provide a better description of the functional connectivity in dense circuits than the usual sample correlations.
 \section*{Author Summary}
@@
 Low-rank components can be extracted using principal component or factor analysis.
-The resulting structure of neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.
 
 Alternatively, the correlation matrix can be expressed through the partial pairwise correlations.
-The partial pairwise correlation is the linear correlation remaining after accounting for correlations with all the other neurons.
+The partial correlation between a pair of neurons is a measure of their linear association after the linear effects of all other neurons have been taken into account.
 Partial correlations carry particular significance when the relationship between the variables is approximately linear.
-In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent.
+In this case, small pairwise partial correlations indicate that the two variables are nearly independent when conditioned on the activity of the other neurons.
 
@@
 \paragraph{Illustration of estimation of a neural correlation matrix}
-To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
+To illustrate the challenges and techniques of estimation of the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
 Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero.
@@
 \end{figure}
-\paragraph{Model selection}
-Covariance selection is but one of many conceivable low-dimensional approximations of the covariance matrix.
-Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix.
-Models that accurately reflect the dominant regularities in the activity of the circuit, its so-called `functional connectivity', tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators.
-An estimator's superior efficiency can be ascertained by cross-validation.
-Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case.
-The estimator that is shown to dominate all others for a specific type of neural circuit serves as an effective descriptor the circuit’s correlation structure.
-This structure may then be analyzed with respect to the circuit's anatomical organization.
-\paragraph{Summary of findings}
-In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and `sparse partial correlations with latent units' or $C_{\sf sparse+latent}$.
-First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators.
-We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex.
-We found that $C_{\sf sparse+latent}$ consistently dominated the other estimators.
-Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.
+\paragraph{Our contribution}
+Various other regularization schemes also produce improved estimates.
+In this study, we compared the sample covariance matrix and four regularized estimators based on distinct correlation structures as follows:
+\begin{description}
+\item[$C_{\sf 0}$] -- sample covariance matrix, the usual estimator used in previous studies.
+\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix.
+\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis with an optimal number of factors.
+\item[$C_{\sf sparse}$] -- sparse inverse \emph{covariance selection} \cite{Dempster:1972, Friedman:2008}, without shrinkage.   
+\item[$C_{\sf sparse+latent}$] -- inverse matrix as the sum of a sparse  and a low-rank component \cite{Chandrasekaran:2010, Ma:2013}.
+\end{description}
+
+First, we demonstrated that, in simulations with known correlation structures, regularized estimators with matching structure were generally more efficient than the other estimators.
+
+We also performed a cross-validated evaluation to establish the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex.
+We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.
+This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units.
+Since it led to better estimates, we proposed that this structure provided a more physiologically relevant representation of the circuit’s functional connectivity than that obtained from the sample correlation matrix.
+We showed that the partial correlation structure had stronger dependence on physical distances and differences in orientation preference than sample correlations.
+In Discussion, we outline possibly approaches to corroborate the relationship between the structure of the functional connectivity revealed by this and similar method to the anatomical organization of the circuit.
 
@@
 \paragraph{Covariance estimation}
-We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (\figref{02}{row 1}).
+We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
+In probabilistic systems with linear dependencies, these correspond to graphical modes with distinct types of structure  (\figref{02}{row 1}).
 
@@
 
-The target of the second regularized estimator $C_{\sf factor}$, is the factor model $LL^\T + \Psi$, where $L$ is the $p\times d$ matrix of \emph{factor loadings}, where $d$ is the number of factors, and $\Psi$ is the diagonal matrix of independent variances.
+The target of the second regularized estimator $C_{\sf factor}$, is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix of rank $d$, and $\Psi$ is the diagonal matrix of independent variances.
 The estimator,
 \begin{equation}\label{eq:c-factor}
-C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (LL^\T + \Psi),
+C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (L + \Psi),
 \end{equation}
 has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $LL^\T + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
+The target estimate $L + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
 
@@
 
-The fourth estimator $C_{\sf sparse+latent}$, combines the assumptions of the previous two.
-The underlying structure arises from sparse interactions between the recorded neurons and latent factors driving their activity (\figref{02}{D}).
+The fourth estimator, $C_{\sf sparse+latent}$, combines the assumptions of the previous two.
+The underlying structure arises from sparse interactions between the observed neurons and several latent units (\figref{02}{D}).
 The estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
 \begin{equation}\label{eq:c-sl}
-C_{\sf sparse+latent} = (S - LL^\T)^{-1}
+C_{\sf sparse+latent} = (S - L)^{-1}
 \end{equation}
-where, as above, $S$ is a sparse matrix and $L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.
+where, as above, $S$ is a sparse matrix and $L$ is a $p\times p$ matrix of rank $d$. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.
 
@@
 
-When the ground truth, $\Sigma,$ is not accessible, loss can be estimated solely from the data through \emph{validation}.
+When the ground truth, $\Sigma$, is not accessible, loss can be estimated solely from the data through \emph{validation}.
 In validation, an additional independent \emph{testing sample} is used to compute a sample covariance estimate $C_{\sf 0}^\prime$.
-Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can be used as an unbiased estimate\footnote
-{
-    Validation loss $\loss{C,C_{\sf 0}^\prime}$ is an unbiased estimate of loss $\loss{C,\Sigma}$ when $\loss{\cdot,\cdot}$ is \emph{additive} in its second argument so that 
+Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can be used as an estimate of loss $\loss{C,\Sigma}$.
+Thus, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
+
+The negative log-likelihood loss (Eq.~{\ref{eq:loss}) is particularly convenient because it is additive in its second argument in the sense that 
  \begin{equation*}\label{eq:additivity}
@@
 
-Then the absence of bias can be shown:
+Then the cross-validation loss is in fact an unbiased estimate of the regular loss.
 \begin{equation*}
@@
 
-The last result expression is known as the \emph{estimator risk} which serv
-
-The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses, for example, which prevents their substitution with corresponding validation losses.
-} 
-of loss $\loss{C,\Sigma}$.  Thus, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
+The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses.
+With non-additive loss functions, cross-validation loss is a biased estimate of loss, but it is still generally appropriate to be used in lieu of loss.
 
@@
 At the same time, average correlations were relatively small (\figref{03}{G}), giving a potential advantage to estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$).
-Finally, correlation structure could possibly be best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$).
+Finally, correlation structure could possibly be best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf sparse+latent}$).
 Thus it was \emph{a priori} unclear which estimator would perform best.
@@
     All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
-    The green arrows indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
+    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
+{\bf A} $C_{\sf sparse+latent}$ outperforms $C_{\sf 0}$: median improvement 0.16 nats/neuron, $p=3.9\times 10^{-5}$.
+    {\bf B} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median improvement 0.0032 nats/neuron, $p=2.2\times 10^{-3}$.
+    {\bf C} $C_{\sf sparse+latent}$ outperforms $C_{\sf factor}$: median improvement 0.16 nats/neuron, $p=6.9\times 10^{-5}$.
+    {\bf D} $C_{\sf sparse+latent}$ outperforms $C_{\sf sparse}$: median improvement 0.0016 nats/neuron, $p=5.5\times 10^{-6}$.
 } \label{fig:04}}
@@
 Under the assumption of a multivariate normal distribution or, more generally, an elliptical distribution, sparse partial correlations suggest a sparse network of dependencies, \emph{i.e.}\;a probabilistic graphical model.
-This property has led some investigators to propose that sparse inverse approximation of correlations may serve as closer estimates of direct anatomical interactions between brain regions in imaging studies than correlations \cite{Varoquaux:2012,Ryali:2012}.
+This property has led some investigators to propose that sparse inverse approximation of correlations may serve as closer estimates of direct anatomical interactions between brain regions in imaging studies than correlations \cite{Varoquaux:2012, Ryali:2012}.
 However, since deviations from elliptical distribution disrupt the correspondence between partial correlations and linear dependencies \cite{Loh:2012}, it is not clear whether this is indeed  the case.
@@
 At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
-If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+lowrank}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity.
+If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+latent}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity.
 Various forms of GLMs could also be evaluated with in the same way: sparsified for best estimation of the covariance matrix (not log-likelihood) and compared to the estimators studied in this study.
-Again, if this new covariance estimator outperforms, $C_{\sf sparse+lowrank}$, we will be compelled to adapt it for estimating the functional connectivity.
+Again, if this new covariance estimator outperforms, $C_{\sf sparse+latent}$, we will be compelled to adapt it for estimating the functional connectivity.
 
@@
 \begin{equation}
-(L,\Psi) = \argmin\limits_{\tilde L,\tilde\Psi} \loss{\tilde L \tilde L^\T + \tilde\Psi,C_{\sf 0}},
+(L,\Psi) = \argmin\limits_{\tilde L,\tilde\Psi} \loss{\tilde L + \tilde\Psi,C_{\sf 0}},
 \end{equation}
@@
 
-Estimator $C_{\sf sparse+lowrank}$ has two hyperparameters: the number of latent units $d$ and the sparsity level $\lambda$.
+Estimator $C_{\sf sparse+latent}$ has two hyperparameters: the number of latent units $d$ and the sparsity level $\lambda$.
 It estimates the larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.
@@
 \begin{equation}
-C_{\sf sparse+lowrank} = (S^\ast)^{-1} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
+C_{\sf sparse+latent} = (S^\ast)^{-1} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
 \end{equation}
 The matrix $ S_{12}S_{22}^{-1}S_{12}^\T$ has rank $d$.
-Rather than searching for the optimal sparse structure of $S_{12}$ and $S_{22}$, an ill-posed problem, we estimated these components together as a single matrix $LL^\T$ of rank $d$, where $L$ is of size $p\times d$.
+Rather than searching for the optimal sparse structure of $S_{12}$ and $S_{22}$, an ill-posed problem, we estimated these components together as one positive-definite $p\times p$ matrix $L$ of rank $d$.
+
 The estimate is found in two steps. First, we use the ADMM algorithm to find the zero structure of $S$ by minimizing the $L_1$-penalized loss \cite{Chandrasekaran:2010,Ma:2013}:
 \begin{equation}
-(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L\tilde L^\T} + \|\tilde Z\|_1
+(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L} + \|\tilde Z\|_1
 \end{equation}
@@
 \begin{equation}
-(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L \tilde L^\T}
+(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L }
 \end{equation}
 
-The partial correlation matrix computed from the overall estimate, computed from $C_{\sf sparse+lowrank}$ according to Eq.~\ref{eq:partial}, includes the effects of interactions between the visible and latent units.
+The partial correlation matrix computed from the overall estimate, computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}, includes the effects of interactions between the visible and latent units.
 This estimate of the partial correlations was used in analyses where sparsity was not required (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).
@@
 \end{center}
-\caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators with respect to quadratic cross-validation loss.}
+\caption{{\bf The sparse+latent estimator outperforms the other estimators with respect to quadratic cross-validation loss $\loss{C,C_{\sf 0}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf 0}^\prime-I)^2$.}
 }

### 2013-12-30 plos.tex
new version of the discussion nearly completed.

@@
 \DeclareMathOperator{\Tr}{tr}
+\newcommand{\sq}[1]{\lq#1\rq}
 \newcommand{\mcond}{\,\middle\vert\,}
@@
 Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.
-Because of its superior estimation performance, the inferred partial correlation structure obtained using these methods may provide a better description of the functional connectivity in dense circuits than the usual sample correlations.
+Because of its superior estimation performance, the partial correlation structure by these methods will likely provide a better description of the functional connectivity in dense circuits than the usual sample correlations.
 \section*{Author Summary}
@@
 Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.
-For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
+For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
 In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits and reflect interactions between neurons \cite{Gerstein:1964}.
@@
 Such approximations are particularly suitable for capturing shared fluctuations across the recorded population, which may arise in local circuits or be externally driven.
-Low-rank components can be extracted using principal component or factor analysis.
+Low-rank components can be extracted using principal component or factor analysis \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.
 
@@
 Partial correlations carry particular significance when the relationship between the variables is approximately linear.
-In this case, small pairwise partial correlations indicate that the two variables are nearly independent when conditioned on the activity of the other neurons.
+In this case, a low pairwise partial correlation indicates that the two variables are nearly independent when conditioned on the activity of the other neurons.
 
@@
 Therefore, networks of partial correlations (sometimes called \emph{association networks}) can be used to infer interactions between the components of a complex system.
-This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005,Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012,Ryali:2012}.
+This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005, Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012, Ryali:2012}.
 In multivariate normal distributions (or, more generally, \emph{elliptical distributions}) interactions are completely defined by linear effects and, therefore, partial correlations reflect conditional dependencies.
@@
 With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}.
-For example, the coupling terms in pairwise Ising models do not directly correspond o the partial correlations \cite{Schneidman:2006, Tkacik:2006}.
+For example, the coupling terms in pairwise Ising models fitted to the same sample correlation matrix will produce a different network of dependencies than that inferred by the partial correlations \cite{Schneidman:2006, Tkacik:2006}.
 
@@
 \end{equation}
-Similarly, the matrix of partial correlations $P$ is computed as the negative normalized version of the \emph{precision matrix} (inverse covariance matrix) $C^{-1}$:
+Similarly, the matrix of partial correlations $P$ is computed by normalizing the the negative \emph{precision matrix} $C^{-1}$ (inverse of the covariance matrix):
 \begin{equation}\label{eq:partial}
@@
 Estimation can be improved through \emph{regularization}: deliberately imposing some kind of structure on the estimate.
-The sample estimate $C_{\sf 0}$, which is usually used to estimated covariance matrices in neuroscience, is unbiased ($\E{C_0}=\Sigma$).
+The sample estimate $C_{\sf 0}$, which is usually used to estimate covariance matrices in neuroscience, is unbiased ($\E{C_0}=\Sigma$).
 However, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise.
@@
 
- A regularized estimate is produced by biasing (`shrinking’) the estimate from the unbiased estimate toward a selected low-dimensional approximation (`target’) of the unbiased estimate \cite{Schafer:2005, Bickel:2006}.
-The selection of the optimal target estimate and the optimal amount of shrinkage can be optimized based on the training data by analytical methods or by cross-validation.
-However, some regularization schemes only perform target selection (\emph{dimensionality reduction}) and replace the estimate without shrinkages whereas others only perform shrinkage given a single target estimate (\emph{shrinkage estimators}).
+ A regularized estimate is produced by \sq{shrinking} the unbiased estimate toward a selected low-dimensional approximation (\sq{target estimate}) of the unbiased estimate \cite{Schafer:2005, Bickel:2006}.
+The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data analytically or by cross-validation.
+Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely, without shrinkage.
+Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).
 
@@
 
+Although regularized covariance estimates have become established in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization schemes for neural correlations.
+Since the performance of regularization schemes depends on the dominant regularities in the data, their optimal choice depends on the type of neural recording, the temporal scale, and brain states.
+For example, dense recordings from local circuits with high temporal resolution in desynchronized brain states may warrant greater emphasis on partial correlations, reflecting direct interactions between neurons.
+Recordings from distributed groups of cells, synchronous circuit activity, or recordings with low temporal resolution are likely to reveal common correlated activity across many cells with correlation matrices best modeled by low-rank components.
+Therefore, the optimal choice of the regularization scheme, ascertained from the data by cross-validation, constitutes a statistical description of a neural circuit that informs about the dominant regularities in the data.
 \paragraph{Illustration of estimation of a neural correlation matrix}
@@
 
-\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
     {\caption{{\bf Illustration of regularized estimation of partial correlations.}
@@
 \item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis with an optimal number of factors.
-\item[$C_{\sf sparse}$] -- sparse inverse \emph{covariance selection} \cite{Dempster:1972, Friedman:2008}, without shrinkage.   
-\item[$C_{\sf sparse+latent}$] -- inverse matrix as the sum of a sparse  and a low-rank component \cite{Chandrasekaran:2010, Ma:2013}.
+\item[$C_{\sf sparse}$] -- sparsified precision matrix,  \emph{covariance selection} \cite{Dempster:1972, Friedman:2008}, without shrinkage.   
+\item[$C_{\sf sparse+latent}$] -- precision matrix composed as the sum of a sparse and a low-rank component \cite{Chandrasekaran:2010, Ma:2013}.
 \end{description}
@@
         {\bf Row 1.} Graphical representations of four types of low-dimensional structures of interactions between observed neurons (green spheres) and latent units (light-shaded spheres).
-        In the `independent model' ({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units. 
-        In `latent factors' ({\bf B}), the correlated activity of observed cells is driven by several latent units. 
-        In `sparse interactions' ({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
-        In `sparse+latent' ({\bf D}), correlations arise through direct linear interactions in a sparse subset of observed neurons and through interactions with common latent units. 
+        In the \sq{independent model} ({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units. 
+        In \sq{latent factors} ({\bf B}), the correlated activity of observed cells is driven by several latent units. 
+        In \sq{sparse interactions} ({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
+        In \sq{sparse+latent} ({\bf D}), correlations arise through direct linear interactions in a sparse subset of observed neurons and through interactions with common latent units. 
         {\bf Row 2.} Examples of $50\times 50$ correlation matrices corresponding to each type of low-dimensional structure. 
@@
         The partial correlation matrix of the sparse model ({\bf C}) is 73\% sparse.
-        The `sparse+latent' matrix has one latent unit and its direct interactions are 78\% sparse.
+        The \sq{sparse+latent} matrix has one latent unit and its direct interactions are 78\% sparse.
         {\bf Row 3.} Examples of sample correlation matrices calculated from samples of 1000 observations taken from simulated random processes with corresponding correlation matrices from row 2.
@@
 Then each subset is used as the validation sample with the other $K-1$ serving as a training dataset.
-The validation losses from each of such `folds' are averaged to produce \emph{cross-validation loss} or \emph{CV-loss} for short.
+The validation losses from each of such \sq{folds} are averaged to produce \emph{cross-validation loss} or \emph{CV-loss} for short.
 Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ trained on the remaining $K-1$ subsets.
@@
 For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth.
-Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.
 
@@
 
-\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
     {\caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
@@
 The partial correlation was decomposed into sparse and low-rank components (\figref{05}{C}).
-Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations.
+Although correlations were mostly positive, the sparse partial correlations (or \sq{interactions}), had a much larger fraction of negative values than sample correlations.
 The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}).
@@
 \section*{Discussion}
-\paragraph{Functional and anatomical connectivity}
-Systems neuroscience is the quest to discover principles whereby anatomical organization of neural circuits gives rise to their function.
-This line of inquiry requires expressive descriptions of the observed activity of neural circuits.
+\paragraph{Functional connectivity}
+How the anatomical organization of neural circuits gives rise to their function is a foremost question in neuroscience.
+Effective phenomenological descriptions of neural activity, once discovered, enable investigation of underlying physiological mechanisms.
+For example, the discoveries of receptive fields and tuning properties of individual cells, over the past 60 years, have led to fruitful hypotheses about the underlying patterns of synaptic connectivity, principles of neural coding, and to the discovery of other organizational principles such as various cortical feature maps \cite{Reid:2012}.
 
-The term \emph{functional architecture} most commonly refers to stimulus response properties of neurons \cite{Reid:2012}.
-Over the past 50 years, tremendous progress has been made in relating the functional architecture to the anatomical, synaptic wiring of neuronal populations and individual neurons.
-This line of inquiry is based on the view of the brain as a complex filter of external stimuli.
+The term \emph{functional connectivity} encompasses various phenomenological descriptions of patterns of activity in populations of neurons.
+It is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons that is not accounted by the independent response properties of the individual cells.
+Neuroscientists have hypothesized that spontaneous activity and trial-to-trial variations in population responses are not random noise but endogenously generated activity with fundamental roles in stimulus coding, sensory inference, attention, learning, and cognition and that further progress depends on understanding of patterns of activity in populations of cells.
+This line of inquiry may require characterizing changes in functional connectivity due to development, learning, brain states, and behavioral tasks and finding its relationships to the circuits’ anatomical organization.
+\paragraph{Temporal scales of functional connectivity}
+Analyses of functional connectivity depend profoundly on their temporal resolution.
+At millisecond-scale resolution, cross-correlograms of the spiking activity of pairs of cells manifest statistical signatures of monosynaptic connections \cite{Gerstein:1964, Moore:1970, Gerstein:1989, Alonso:1998, Denman:2013}.
+Other methods require temporal resolution on the order of membrane time constants (milliseconds to tens of milliseconds): \emph{e.g.}~statistical models that mimic membrane integration (\cite{Pillow:2008}) and information-theoretic and statistical physics models that bin spike trains finely to ensure a binary representation \cite{Schneidman:2006, Tkacik:2006, Tkacik:2013}.
+At slower scales of tens to hundreds milliseconds, temporal sequencing of spikes can no longer resolve directions of monosynaptic interactions and any expression of functional connectivity summarizes chains of synaptic transmissions with more complex, emergent properties of network computation: the analysis shifts from to detecting co-active cell assemblies and the low-rank structure of correlations \cite{Ch:2010, Lopes:2011, Lopes:2013}.
 
-In contrast to functional architecture, \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or trial-to-trial variations,  not directly driven by external stimuli.
-Functional connectivity is thought to reflect internal, recurrent connectivity in neural circuits.
-However, less experimental evidence has been garnered to for links between functional and anatomical connectivity.
+The temporal resolution of calcium signals is limited by calcium dye kinetics.
+Fast imaging techniques combined with advanced spike inference algorithms have been demonstrated to provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010, Katona:2012, Cotton:2013}.
+However, high temporal precision comes at the cost of the accuracy of the inferred spike rates: better accuracy can be achieved when calcium signals are analyzed on the temporal scales of tens of milliseconds \cite{Cotton:2013}.
+Therefore, functional connectivity studies from calcium signals have focused on the slower temporal scales (tens or hundreds of milliseconds).
+In this study, too, we chose 20 Hz for the temporal precision of spike inference and bin widths of 150 ms for computing correlations.
 
-Neural correlations boast a rich history as indicators of functional connectivity.
-When measured on temporal scales of milliseconds, sharp peaks or troughs in the cross-correlogram between pairs of neurons at latencies typical of synaptic transmission provide evidence of monosynaptic connection between pairs of cells \cite{Moore:1970, Alonso:1998, Denman:2013} (although other processes can also generate sharp peaks in the cross-correlogram while many monosynaptic connections may not produce such signatures).
-Without targeted stimulation of specific cells or measurements of subthreshold intracellular events, such analysis remains the best method for detecting probable monosynaptic connections in vivo.
+\paragraph{Estimation of functional connectivity depends on recorded population size and density}
+Analyses of functional connectivity depend profoundly on the sizes and densities the recorded population.
+Pairwise recordings may contain statistical signatures of monosynaptic connections but conditioning on the activity of all the other cells may help disambiguate the network of interactions in the circuit.
+This distinction has been described as \emph{effective connectivity}: the best guess at the most direct effects exerted by neurons on each other \cite{Gerstein:1989}.
+For example, the interaction terms in spatiotemporal generalized linear models \cite{Pillow:2008} or Ising-like maximum entropy models \cite{Schneidman:2006, Tkacik:2006, Ganmor:2011} express the conditional dependencies between neurons, which only have weak relationship with the pairwise statistics to which they are fitted.
+If neurons can be assumed to exert nearly linear effects on each other, then conditional dependencies between neurons are expressed by the partial correlations.
+For effectual computation of conditional dependencies, the recording must include a sufficient fraction of the neurons in the embedding circuit, or even the complete population.
+Otherwise, interactions with the unobserved portions of the circuit will introduce false conditional dependencies between the observed neurons.
+For this reason, the most successful applications of statistical models of population activity thus far have been in in vitro preparations of the retina or cell cultures where high-quality recordings from the complete populations have been possible.
 
-When measured at longer temporal scales, neural correlations reflect activations involving large ensembles of neurons and may reflect direct causal anatomical interactions, indirect causal interactions, or complex network-wide emergent features of activity with little correspondence to the local anatomical connectivity.
-Calcium imaging allows simultaneous measurements of the activity of large populations of cells \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even of entire nervous systems \cite{Leung:2013, Ahrens:2013}, but its temporal bandwidth is limited to the scale of tens or hundreds milliseconds by the kinetics of calcium-sensitive dyes, limiting its use to slower temporal scales at which correspondence between functional and anatomical connectivity is less direct.
+Multielectrode recordings implanted in the cortex record from relatively small groups of cells separated by relatively long distances.
+Perhaps for this reason, the structure of partial correlations has not been, to our knowledge, been used to describe the functional connectivity in multielectrode recordings.
+In contrast, two-photon calcium imaging allows simultaneous measurements of the activity of nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
+Therefore, two-photon imaging opens opportunities for computing the structure of conditional dependencies between neurons with arguably closer correspondence to the underlying physiological interactions.
 
-\paragraph{Extraction of the correlation structure}
-The word `structure' implies a succinct, simplified representation that captures only the essential, significant features.
-Several forms of correlations structure can be extracted from the same data and used to describe functional connectivity in different ways.
+\paragraph{Finding improved representation of the correlation structure}
+We were initially motivated by the somewhat simplistic hypothesis that partial correlations in local cortical microcircuits should provide more meaningful representations of the functional connectivity than the conventional correlations.
+Since neurons are known to form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we imposed the constraint that most of the partial correlations must be zero.
+To account for interactions with the rest of the brain, we applied the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables\cite{Chandrasekaran:2010, Ma:2013}.
+A priori, the validity of this assumption was debatable:  On the temporal scales of 150 ms, the pairwise effects between neurons could fade in prominence compared to the co-activation of larger cell assemblies and global circuit fluctuations, which would be poorly fitted by sparse partial correlations.
+In addition, strong nonlinearities in interactions between neurons could disrupt the relationship between conditional independence and absence of partial correlation.
 
-Perhaps the least sophisticated yet common approach for extracting the structure of correlations is to threshold above a selected level of significance.
-Part of the appeal of thresholded correlations is their representation in the form of graphs that can be related to graphs of synaptic connectivity and made amenable to network analysis \cite{Bullmore:2012}.
-For example, \cite{Malmersjo:2013} analyzed the graph theoretical properties of networks of thresholded correlations in development.
+To provide evidence for the validity of this hypothesis, we compared the effect of imposing this structure on the estimation of the noise covariance matrices relative to estimations based on other assumptions.
+The finding that the sparse+latent structure of correlations led to better estimation than the other estimators lends a degree of credence to its structural representation of the circuit's functional connectivity.
+Of course, this finding could be overturned or augmented by finding an even more efficient covariance estimator for these types of recordings and we welcome the challenge.
 
-Alternatively, the correlation structure may be extracted in the form of low-rank approximations computed by principal component analysis, factor analysis, or activation maps conditioned on the activity of a single neuron.
-Low-rank components show synchronous multineuronal activation patterns without revealing graphs of interactions between neurons.
-They are often interpreted as emergent patterns of large circuits or brain state fluctuations \cite{Kenet:2003}.
-
-Another reduced representation of correlation matrices is obtained by sparsifying its inverse, resulting in a graph of partial correlations.
-Under the assumption of a multivariate normal distribution or, more generally, an elliptical distribution, sparse partial correlations suggest a sparse network of dependencies, \emph{i.e.}\;a probabilistic graphical model.
-This property has led some investigators to propose that sparse inverse approximation of correlations may serve as closer estimates of direct anatomical interactions between brain regions in imaging studies than correlations \cite{Varoquaux:2012, Ryali:2012}.
-However, since deviations from elliptical distribution disrupt the correspondence between partial correlations and linear dependencies \cite{Loh:2012}, it is not clear whether this is indeed  the case.
-Partial correlations are meaningful when most interacting variables are observed.
-Perhaps for this reason, until now, partial correlations had not been considered in analysis of multineuronal recordings.
-
-In this study, we sought to uncover the optimal low-dimensional approximation of the covariance matrix for efficient estimation of the true covariance matrices in cortical microcircuits on the temporal scale of 100--200 ms corresponding to the temporal resolution of spike rates from somatic calcium signals \cite{Cotton:2013}.
-We found that networks of partial correlations were more statistically efficient than low-rank approximations of correlation matrices (Fig.\;S\ref{supp:01}\;E).
-Furthermore, estimation became still more efficient when low-rank components were allowed to be combined with sparse partial correlations.
-We propose that correlation structures that are supported by the data  are more likely to reflect anatomical, causal interactions.
-The network of sparse partial correlations suggests interactions between specific pairs whereas the low-rank represents emergent activity patterns involving collective activations of larger circuits or global fluctuations.
-The revealed network of partial interactions was substantially different from that revealed by thresholding correlations (\figref{05}{C, F, H, I}).
-In particular, it shows that many low correlations correspond to strong negative interactions (\figref{05}{F}).
-Intuitively, this suggests that many low interactions are as surprising or significant considering the common correlations with other neurons.
-The analysis of sample correlations does not ascribe significance to correlation coefficients that are surprisingly low, based on their common correlations to other cells.
-
-\paragraph{Functional connectivity from probabilistic models}
-Increasingly, functional connectivity is inferred by fitting probabilistic models such as various forms of multivariate Gaussian models, linear-nonlinear models, Generalized Linear Models \cite{Pillow:2008}, and maximum entropy models \cite{Schneidman:2006, Tkacik:2006, Tang:2008, Shlens:2009}.
-In probabilistic models, functional dependencies are defined by multiplicative terms comprising the model.
-If the model can be decomposed as a product of terms so that a pair of variables can be segregated into separate terms, the two variables are conditionally independent: the activity of one neuron can be predicted from other neurons just as well with or without considering the second neuron, suggesting a lack of a direct causal interaction between them.
-Probabilistic models allow inferring latent variables to account for the activity of unobserved circuitry.
-
-\paragraph{Model selection}
-Various estimation methods infer very different structures of functional connectivity from the same data.
-What kind of structure expresses most relevant information?
-How does one determine which functional connectivity structure is best supported by the empirical data?
-The field of machine learning offers a rich toolbox of model selection criteria.
-Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion}, and \emph{Bayes (or Schwartz) Information Criterion} (BIC) are computationally efficient but rely on generally unsupportable assumptions about the data generating process.
-The most assumption-free yet more computationally intensive are procedures based on cross-validation \cite{Arlot:2010} and, when models have hyperparameters, nested cross-validation to select their values.
-It is commonly assumed that probabilistic models that demonstrate the highest average cross-validated log likelihood yield a better representation of the functional connectivity.
-For example, \cite{Ganmor:2011} demonstrated that their proposed Reliable Network Model (a maximum-entropy model on the multivariate binary domain constrained by a sparse heuristically selected network of correlations of several orders) yielded higher cross-validated log-likelihoods than the full-pairwise Ising model.
-The authors use this fact to support the claim that the former model revealed a more anatomically relevant structure of functional connectivity than the pairwise Ising model.
-
-However, in simulation, we showed that, more constrained, simpler models often outperformed complex models in cross-validation studies even when their structure was differed from the true structure (\figref{02}{row 5}), only because these more constrained models were less sensitive to sampling noise.
-Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process.
-For example, a properly sparsified pairwise Ising model could outperform the Reliable Network model.
-Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.
-
-In our study we selected between families of models based on their cross-validation performance.
-The winning model family was more flexible than the others and therefore more susceptible to noise.
-The fact that it still performed better indicated that the additional features were required.
-We conclude that under these circumstances, the selected model can be asserted to better represent the true structure of the data generating process, at least in comparison to the other three models.
-
-In our study, we focused on estimating a specific parameter of the distribution (its covariance matrix) rather than fitting a specific parametric distribution.
-Under the loss function (Eq.~\ref{eq:loss}) we selected, the covariance matrix estimation methods are equivalent to fitting a multivariate Gaussian model and selecting the best structure on the basis of cross-validated log-likelihood.
-However, the ranking of models was preserved under other loss functions (Fig.~S\ref{supp:02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.
-We defend the practice of evaluating and optimizing probabilistic models with respect to their parameters that are more relevant for functional connectivity rather than maximizing their cross-validated log likelihoods, which are sensitive to all parameters.
-For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix.
-We maintained that model B must be more relevant for studies of functional connectivity because correlations do reflect aspects of functional connectivity.
-
-The correlation structures we considered also corresponded to structures of conditional dependencies in multivariate Gaussian models or, more generally, elliptical distributions.
-For example, zero partial correlation indicates conditional independence between a pair of variables under an elliptical distribution, but this correspondence weakens or breaks down with deviation from elliptical distributions \cite{Loh:2012}.
-Other probabilistic models, when optimally sparsified for best cross-validated estimation of the covariance matrix, could suggest very different forms of low-dimensional representations (structures) of the covariance matrix.
-For example, spatiotemporal Ising models, marginalized to reproduce the correlation matrix at the same temporal scale, would suggest an alternative correlation structure from the same data.
-At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices.
-If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+latent}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity.
-Various forms of GLMs could also be evaluated with in the same way: sparsified for best estimation of the covariance matrix (not log-likelihood) and compared to the estimators studied in this study.
-Again, if this new covariance estimator outperforms, $C_{\sf sparse+latent}$, we will be compelled to adapt it for estimating the functional connectivity.
-
-\paragraph{Big neural data}
-Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}.
-It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.
-Large datasets aggravate the numerical problems and the model selection caveats described in this paper.
-The number of parameters in models of population activity grows quadratically.
-Models of joint population activity feature large numbers of parameters, that growing quadratically or faster with population size.
-Without ways to restrict such models, they quickly become ill conditioned (e.g.
-\cite{Roudi:2009}).
-Statistical models of population activity will need to have both correct mathematical forms of interactions and optimal ways of restricting them so that the total number of estimated parameters remains commensurate with the amount of measurements.
-
-Our study employs regularization, which restricts the models based on the data itself.
-Other models may use restrictions that are based on various heuristics in addition to regularization.
-These methods have numerous caveats when it comes to inferring the optimal structure of functional connectivity from such methods.
-Improved estimation of covariances and the general framework of producing such estimators for specific neural populations can be applied to various circuits and serve as a statistical description of their functional organization.
+In our study, the covariance matrix estimators were evaluated with respect to the cross-validated normal log likelihood.
+This does not limit the applicability of its conclusion normal distributions.
+Indeed the major findings in this paper could be reproduced with respect to other loss functions (compare Fig.\ref{fig:04} and Fig.~S\ref{supp:02}).
+Other probabilistic models, fitted to the same data, could be used to produce their respective estimates of the covariance matrix.
+If another model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of functional connectivity.
+We advocate comparing models by how well they estimate the covariance matrix due to their well established relevance to connectivity and neural coding.
 
@@
 GROUPS CELLS MAY BE CONNECTED BUT ONLY INTERACTING IN STATE A AND NOT B.
+
+\paragraph{Challenges of big neural data}
+The extraction and interpretation of functional connectivity depended on the density and size of the recorded neuronal population.
+
+In this study, we sought to uncover the optimal low-dimensional approximation of the covariance matrix for efficient estimation of the true covariance matrices in cortical microcircuits on the temporal scale of 100--200 ms corresponding to the temporal resolution of spike rates from somatic calcium signals \cite{Cotton:2013}.
+We found that networks of partial correlations were more statistically efficient than low-rank approximations of correlation matrices (Fig.\;S\ref{supp:01}\;E).
+Furthermore, estimation became still more efficient when low-rank components were allowed to be combined with sparse partial correlations.
+We propose that correlation structures that are supported by the data  are more likely to reflect anatomical, causal interactions.
+The network of sparse partial correlations suggests interactions between specific pairs whereas the low-rank represents emergent activity patterns involving collective activations of larger circuits or global fluctuations.
+The revealed network of partial interactions was substantially different from that revealed by thresholding correlations (\figref{05}{C, F, H, I}).
+In particular, it shows that many low correlations correspond to strong negative interactions (\figref{05}{F}).
+Intuitively, this suggests that many low interactions are as surprising or significant considering the common correlations with other neurons.
+The analysis of sample correlations does not ascribe significance to correlation coefficients that are surprisingly low, based on their common correlations to other cells.
+
+Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}.
+It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.
+Large datasets aggravate the numerical problems and the model selection caveats described in this paper.
+The number of parameters in models of population activity grows quadratically.
+Models of joint population activity feature large numbers of parameters, that growing quadratically or faster with population size.
+Without ways to restrict such models, they quickly become ill conditioned (e.g.
+\cite{Roudi:2009}).
+Statistical models of population activity will need to have both correct mathematical forms of interactions and optimal ways of restricting them so that the total number of estimated parameters remains commensurate with the amount of measurements.
+
+Our study employs regularization, which restricts the models based on the data itself.
+Other models may use restrictions that are based on various heuristics in addition to regularization.
+These methods have numerous caveats when it comes to inferring the optimal structure of functional connectivity from such methods.
+Improved estimation of covariances and the general framework of producing such estimators for specific neural populations can be applied to various circuits and serve as a statistical description of their functional organization.
 
@@
 
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics{./figures/Figure-Supp02.pdf}
-\end{center}
-\caption{{\bf The sparse+latent estimator outperforms the other estimators with respect to quadratic cross-validation loss $\loss{C,C_{\sf 0}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf 0}^\prime-I)^2$.}
+\begin{figure}
+\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+{\caption{{\bf The sparse+latent estimator outperforms the other estimators with respect to quadratic cross-validation loss $\loss{C,C_{\sf 0}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf 0}^\prime-I)^2$.}
 }
-\label{supp:02}
+\label{supp:02}}
+{\includegraphics{./figures/Figure-Supp02.pdf}}
 \end{figure}

### 2014-01-03 plos.tex
updated introduction

@@
 {\Large
-Inferred partial neural correlations suggest detailed interactions in cortical circuits
+Partial neural correlations suggest detailed interactions in cortical circuits
 }
@@
 However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.
-Estimation can be improved by regularization: imposing some kind of structure on the estimate.
-The amount of improvement depends on how well the imposed structure accommodates the dominant dependencies in the data.
-Therefore, the selection of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.
+Estimation can be improved by regularization, \emph{i.e.}\;by imposing a structure on the estimate.
+The amount of improvement depends on how closely the assumed structure represents dependencies in the data.
+Therefore, the selection of the most efficient correlation matrix estimator for a given neural circuit must be determined empirically.
+Furthermore, the identity and structure of the most efficient estimator informs about the types of dominant dependencies governing the system.
 
-In this study, we sought most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.
+We sought the most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.
 Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.
-We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.
-Indeed, in cross-validation tests, the covariance matrix estimator based on such approximations outperformed other regularized estimators.
-Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.
-Because of its superior estimation performance, the partial correlation structure by these methods will likely provide a better description of the functional connectivity in dense circuits than the usual sample correlations.
+We hypothesized that in these dense recordings, the correlation matrix should be most efficiently represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.
+Indeed, in cross-validation tests, the covariance matrix estimator with such structure consistently outperformed other regularized estimators.
+The connectivity patterns inferred from the sparse component of the estimate manifested a strong relationship to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distance and difference in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
+Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than sample correlations.
+
 \section*{Author Summary}
@@
 % Author Summary not valid for PLoS ONE submissions.   
-Correlations in the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.
-Estimation of correlation matrices can be improved by imposing some structure on the estimate.
-Greatest improvements are attained when the imposed structure efficiently captures real dependencies in the data.
-Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the performance of correlation matrix estimators that imposed different kinds of structure.
-The correlation structure of the estimator that led to greatest improvement comprised a sparse network of partial correlations between pairs of neurons combined with partial correlations with several common latent factors.
-Both components were necessary for efficient estimation.
-Since it leads to better estimates, we proposed that this sparse+latent structure provides a better picture of the functional connectivity than that obtained from the usual sample correlation matrix in densely sampled neural recordings.
-As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.
+Correlations in the activity of neural populations have proven useful to describe their functional connectivity.
+Correlations may arise from various underlying processes that reflect differently on the structure of the correlation matrix:  Common network fluctuations or external inputs are well approximated by low-rank representations whereas linear effects between specific pairs of neurons are well approximated by fitting their pairwise partial correlations.
+What kinds of correlation effects dominate the population activity in cortical microcircuits?
+To answer this question, we imposed various correlation structures on empirical estimates of neural correlation matrices and evaluated which of them produced greatest estimation improvement.
+In fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the greatest improvement was produced by decomposing the correlation matrix as a sparse network of partial correlations and a low-rank component.
+Since it leads to the best estimates, we propose that this structure provides a better picture of the functional connectivity than that obtained from the usual correlations.
+As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and to differences in their preferred orientations and found basic resemblance to results from studies of synaptic connectivity.
 \section*{Introduction}
@@
 Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.
-For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
+For example, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
 In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits and reflect interactions between neurons \cite{Gerstein:1964}.
@@
 However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.
-Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967b, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.
+Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.
 
-\paragraph{Additional information in neural correlation matrices}
-Despite these ambiguities, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations.
-In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998, Zohary:1994}.
-In contrast, modern multineuronal recordings from increasingly large and dense populations of highly interconnected neurons allow estimation of various low-dimensional representations of the entire correlation matrix, which can enable inquiry into the nature of underlying physiological mechanisms and computational principles.
+\paragraph{Structure  of correlation matrices}
+Modern multineuronal recordings allow measuring entire correlation matrices from increasingly large and dense populations of highly interconnected neurons.
+Multivariate analysis techniques based on large correlation matrices provide information that could not be extracted from an equivalent number of isolated measurements of pairwise correlations.
+For example, when the correlation matrix is approximated in a form with a reduced number of variables, this \emph{reduced representation} can help isolate various aspects of the functional connectivity in neural circuits and possibly point to underlying physiological mechanisms and computational principles.
 
-For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation.
-Such approximations are particularly suitable for capturing shared fluctuations across the recorded population, which may arise in local circuits or be externally driven.
-Low-rank components can be extracted using principal component or factor analysis \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.
+One reduced representation of the correlation matrix is its low-rank approximation.
+Low-rank approximations are particularly suitable for capturing common fluctuations or common external inputs into the circuit and can be extract using principal component analysis, factor analysis, or  \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.
 
-Alternatively, the correlation matrix can be expressed through the partial pairwise correlations.
-The partial correlation between a pair of neurons is a measure of their linear association after the linear effects of all other neurons have been taken into account.
-Partial correlations carry particular significance when the relationship between the variables is approximately linear.
-In this case, a low pairwise partial correlation indicates that the two variables are nearly independent when conditioned on the activity of the other neurons.
+Another reduced representation of the correlation matrix can be obtained by \emph{covariance selection}, \emph{i.e.}\;by selecting a subset of the pairwise \emph{partial correlation} that are sufficient to reproduce the entire correlation matrix \cite{Dempster:1972, Friedman:2008}.
+The partial correlation between a pair of neurons is a measure of their linear association after the effects of the rest of the network have been taken into account \cite{Anderson:2003}.
+Partial correlations are particularly meaningful when relationships between variables are mostly linear and large fraction of the variables are observed.
+In this case, partial correlations approximate conditional dependency.
+A lack of conditional dependency between two neurons suggests a lack of direct interaction between them.
+Networks of partial correlations (sometimes called \emph{association networks}) have been used to infer gene interaction networks \cite{Schafer:2005, Peng:2009} and interactions between brain regions in brain imaging studies \cite{Varoquaux:2012, Ryali:2012}.
 
-Conditional independence between a pair of variables suggests a lack of direct interaction between the underlying processes.
-Therefore, networks of partial correlations (sometimes called \emph{association networks}) can be used to infer interactions between the components of a complex system.
-This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005, Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012, Ryali:2012}.
-In multivariate normal distributions (or, more generally, \emph{elliptical distributions}) interactions are completely defined by linear effects and, therefore, partial correlations reflect conditional dependencies.
-A multivariate normal distribution specified by a weighted graph of non-zero partial correlations (along with the means and variances of individual neurons) is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}.
-With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}.
-For example, the coupling terms in pairwise Ising models fitted to the same sample correlation matrix will produce a different network of dependencies than that inferred by the partial correlations \cite{Schneidman:2006, Tkacik:2006}.
+\paragraph{Improved estimation performance by regularization}
+Imposing some structure on an estimate of the covariance matrix can improve its estimation performance.
 
-\paragraph{Estimation of neural correlation matrices}
-Here, we pursued two related aims: (a)~the improved estimation of neural correlation matrices and (b)~discovery of the low-dimensional structure of correlations in recordings of multineuronal activity.
+Estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
+This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned estimation of the partial correlations \cite{Schafer:2005}.
+In neuroscience studies, the true correlation matrix is usually estimated by the \emph{sample correlation matrix} and these two concepts are rarely distinguished.
+The sample correlation matrix is unbiased but, because it has many free parameters, it is highly sensitive to sampling noise.
+As a result, on average, the sample correlation matrix is far from the true correlation matrix.
+In contrast, reduced estimates are typically less susceptible to sampling noise although can introduce their respective biases.
+
+In statistics, the technique of deliberately imposing some structure on an estimate in order to improve its performance is known as \emph{regularization} \cite{Schafer:2005, Bickel:2006}.
+To \sq{impose structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation known as the shrinkage target.
+The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data either analytically or by cross-validation.
+Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely, without shrinkage.
+Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).
+Yet others can do both target selection and shrinkage toward the target.
+
+Although regularized covariance estimates have become commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.
+\paragraph{Illustration of estimation of a neural correlation matrix}
+To illustrate the challenges and techniques of estimation of the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972, Friedman:2008}.
+Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix (inverse of the covariance matrix) while setting the rest to zero.
+Zeros in the off-diagonal elements of the precision matrix indicate zero partial correlations between the corresponding pair of neurons.
+Therefore, covariance selection amounts to finding an approximation of the covariance matrix so that a large fraction of neuron pairs are linearly independent of each other when conditioned on the activity of all the other observed cells.
+
+We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}).
+Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
+The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
+
+Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth.
+Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.
+
+\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
+        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
+        The outlined square fragment is magnified in {\bf B}.
+        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
+        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
+    }
+    \label{fig:01}}
+    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
+\end{figure}
+\paragraph{Conceptual questions and basic approach}
+What structures of correlation matrices are most suitable for the description of the multineuronal activity in specific neural circuits and brain states?
+For example, are correlations in visual cortex under during visual stimulation best explained by common fluctuations or by local interactions within the recorded microcircuit?
+
+We pursued two related aims: (a)~the improved estimation of neural correlation matrices and (b)~discovery of the low-dimensional structure of correlations in recordings of multineuronal activity.
 An accurate characterization of the patterns of correlations helps in interpreting them, and to better relate them to the architecture and dynamics of the underlying neuronal networks.
+
+ In this study, we compared the sample covariance matrix and four regularized estimators based on distinct correlation structures:
+\begin{description}
+\item[$C_{\sf 0}$] -- sample covariance matrix, the usual estimator used in previous studies.
+\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix.
+\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis with an optimal number of factors.
+\item[$C_{\sf sparse}$] -- sparsified precision matrix,  \emph{covariance selection} \cite{Dempster:1972, Friedman:2008}, without shrinkage.   
+\item[$C_{\sf sparse+latent}$] -- precision matrix composed as the sum of a sparse and a low-rank component \cite{Chandrasekaran:2010, Ma:2013}.
+\end{description}
+
+First, we demonstrated that, in simulations with known correlation structures, regularized estimators with matching structure were generally more efficient than the other estimators.
+
+We also performed a cross-validated evaluation to establish the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex.
+We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.
+This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units.
+Since it led to better estimates, we propose that this structure provides a more physiologically relevant representation of the circuit’s functional connectivity than that obtained from the sample correlation matrix.
+We showed that the partial correlation structure had stronger dependence on physical distances and differences in orientation preference than sample correlations.
+In Discussion, we outline possibly approaches to corroborate the relationship between the structure of the functional connectivity revealed by this and similar method to the anatomical organization of the circuit.
 
@@
 
-Estimation of large covariance matrices presents a number of numerical challenges.
-The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
-This mismatch leads to an increase in spurious correlations, overestimation of shared activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned estimates of the partial correlations \cite{Schafer:2005}.
-
-Estimation can be improved through \emph{regularization}: deliberately imposing some kind of structure on the estimate.
-The sample estimate $C_{\sf 0}$, which is usually used to estimate covariance matrices in neuroscience, is unbiased ($\E{C_0}=\Sigma$).
-However, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise.
-Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.
-Regularization strikes a favorable balance between bias and variability
-
- A regularized estimate is produced by \sq{shrinking} the unbiased estimate toward a selected low-dimensional approximation (\sq{target estimate}) of the unbiased estimate \cite{Schafer:2005, Bickel:2006}.
-The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data analytically or by cross-validation.
-Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely, without shrinkage.
-Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).
-
-Regularization can produce \emph{some} improvement even with an arbitrary target estimate (See Discussion).
-Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.
-
-Although regularized covariance estimates have become established in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization schemes for neural correlations.
-Since the performance of regularization schemes depends on the dominant regularities in the data, their optimal choice depends on the type of neural recording, the temporal scale, and brain states.
-For example, dense recordings from local circuits with high temporal resolution in desynchronized brain states may warrant greater emphasis on partial correlations, reflecting direct interactions between neurons.
-Recordings from distributed groups of cells, synchronous circuit activity, or recordings with low temporal resolution are likely to reveal common correlated activity across many cells with correlation matrices best modeled by low-rank components.
-Therefore, the optimal choice of the regularization scheme, ascertained from the data by cross-validation, constitutes a statistical description of a neural circuit that informs about the dominant regularities in the data.
-\paragraph{Illustration of estimation of a neural correlation matrix}
-To illustrate the challenges and techniques of estimation of the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}.
-Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero.
-We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}).
-Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
-The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
-Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth.
-Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.
-
-\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
-        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
-        The outlined square fragment is magnified in {\bf B}.
-        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
-        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
-    }
-    \label{fig:01}}
-    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
-\end{figure}
-\paragraph{Our contribution}
-Various other regularization schemes also produce improved estimates.
-In this study, we compared the sample covariance matrix and four regularized estimators based on distinct correlation structures as follows:
-\begin{description}
-\item[$C_{\sf 0}$] -- sample covariance matrix, the usual estimator used in previous studies.
-\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix.
-\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis with an optimal number of factors.
-\item[$C_{\sf sparse}$] -- sparsified precision matrix,  \emph{covariance selection} \cite{Dempster:1972, Friedman:2008}, without shrinkage.   
-\item[$C_{\sf sparse+latent}$] -- precision matrix composed as the sum of a sparse and a low-rank component \cite{Chandrasekaran:2010, Ma:2013}.
-\end{description}
-
-First, we demonstrated that, in simulations with known correlation structures, regularized estimators with matching structure were generally more efficient than the other estimators.
-
-We also performed a cross-validated evaluation to establish the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex.
-We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.
-This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units.
-Since it led to better estimates, we proposed that this structure provided a more physiologically relevant representation of the circuit’s functional connectivity than that obtained from the sample correlation matrix.
-We showed that the partial correlation structure had stronger dependence on physical distances and differences in orientation preference than sample correlations.
-In Discussion, we outline possibly approaches to corroborate the relationship between the structure of the functional connectivity revealed by this and similar method to the anatomical organization of the circuit.
+Various other regularization schemes also produce improved estimates
 
@@
 
-The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (\figref{02}{C}).
-This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.
+The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between pairs of observed cells (\figref{02}{C}).
+This assumption is enforced by fitting only a subset of the off-diagonal elements of the precision matrix while setting the rest to zero.
 Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{02}{C}).
@@
 
-Then the cross-validation loss is in fact an unbiased estimate of the regular loss.
+Then the cross-validation loss is in fact an unbiased estimate of loss.
 \begin{equation*}
-    \E[C]{\E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}}=\E[C]{\loss{C,\E[C_{\sf 0}]{C_{\sf 0}^\prime}}}=\E[C]{\loss{C,\Sigma}}
+    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}
 \end{equation*}
@@
 
-\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
     {\caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
@@
     The arrow heads indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
-{\bf A} $C_{\sf sparse+latent}$ outperforms $C_{\sf 0}$: median improvement 0.16 nats/neuron, $p=3.9\times 10^{-5}$.
-    {\bf B} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median improvement 0.0032 nats/neuron, $p=2.2\times 10^{-3}$.
-    {\bf C} $C_{\sf sparse+latent}$ outperforms $C_{\sf factor}$: median improvement 0.16 nats/neuron, $p=6.9\times 10^{-5}$.
-    {\bf D} $C_{\sf sparse+latent}$ outperforms $C_{\sf sparse}$: median improvement 0.0016 nats/neuron, $p=5.5\times 10^{-6}$.
+\\{\bf A.} $C_{\sf sparse+latent}$ outperforms $C_{\sf 0}$: median improvement 0.16 nats/neuron, $p=3.9\times 10^{-5}$.
+\\{\bf B.} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median improvement 0.0032 nats/neuron, $p=2.2\times 10^{-3}$.
+\\{\bf C.} $C_{\sf sparse+latent}$ outperforms $C_{\sf factor}$: median improvement 0.16 nats/neuron, $p=6.9\times 10^{-5}$.
+\\{\bf D.} $C_{\sf sparse+latent}$ outperforms $C_{\sf sparse}$: median improvement 0.0016 nats/neuron, $p=5.5\times 10^{-6}$.
 } \label{fig:04}}
@@
 Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites.
-By design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units.
-If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations.
+By design, the sparse+latent estimator finds the mixture of sparse partial correlations between visible neurons and common fluctuations described as latent units.
+If this estimate dominates, we can hypothesize that it better describes underlying physiological interactions than other estimators.
 In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
@@
     \caption{{\bf Properties of sparse+low-rank regularized estimates from all imaged sites}
+\\
     {\bf A.} The average sample correlations vs.~average partial correlations for each imaged site. In each plot, the red asterisk indicates the site shown in figures \ref{fig:03} and \ref{fig:05}.
@@
 Effective phenomenological descriptions of neural activity, once discovered, enable investigation of underlying physiological mechanisms.
-For example, the discoveries of receptive fields and tuning properties of individual cells, over the past 60 years, have led to fruitful hypotheses about the underlying patterns of synaptic connectivity, principles of neural coding, and to the discovery of other organizational principles such as various cortical feature maps \cite{Reid:2012}.
+For example, the discoveries of receptive fields and tuning properties of single neurons, over the past 60 years, have led to fruitful hypotheses about the underlying patterns of synaptic connectivity, principles of neural coding, and to the discovery of other organizational principles such as various cortical feature maps \cite{Reid:2012}.
 
-The term \emph{functional connectivity} encompasses various phenomenological descriptions of patterns of activity in populations of neurons.
-It is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons that is not accounted by the independent response properties of the individual cells.
-Neuroscientists have hypothesized that spontaneous activity and trial-to-trial variations in population responses are not random noise but endogenously generated activity with fundamental roles in stimulus coding, sensory inference, attention, learning, and cognition and that further progress depends on understanding of patterns of activity in populations of cells.
-This line of inquiry may require characterizing changes in functional connectivity due to development, learning, brain states, and behavioral tasks and finding its relationships to the circuits’ anatomical organization.
+Similarly, the term \sq{functional connectivity} denotes a statistical summary of observed \emph{multineuronal} patterns of activity thought to reflect internal, recurrent connectivity in neural circuits.
+In statistical terms, functional connectivity is an empirical estimate of a feature of the joint probability distribution of the neurons’ activity.
+Although functional connectivity may not provide a detailed and unambiguous interpretation of underlying mechanisms, its utility in experimental neuroscience is that of a descriptive statistic to quantify the general organization of synchronous activations of neuronal populations.
+Functional connectivity is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons.
+To remove the confounding effects of stimuli, the average stimulus responses are usually subtracted or otherwise accounted for.
+
+Pairwise correlations (including cross-correlograms and joint peristimulus histograms) between the spiking activities of neurons boast a rich history as a measure of functional connectivity.
+Interpreted as either the statistical signatures of synaptic connectivity \cite{Gerstein:1964, Perkel:1969, Denman:2013} or of Hebbian neuronal assemblies \cite{Gerstein:1989, Lopes:2013}, pairwise correlations have been instrumental in discovering various principles of the functional organization of neural circuits, as only partially enumerated in the Introduction.
+Several other measures of functions connectivity have emerged including probabilistic models
+
+Parsimony is a desired attribute of a measure of functional connectivity: insignificant or contingent detail.
+
 \paragraph{Temporal scales of functional connectivity}
 Analyses of functional connectivity depend profoundly on their temporal resolution.
-At millisecond-scale resolution, cross-correlograms of the spiking activity of pairs of cells manifest statistical signatures of monosynaptic connections \cite{Gerstein:1964, Moore:1970, Gerstein:1989, Alonso:1998, Denman:2013}.
+At millisecond-scale resolution, cross-correlograms of the spiking activity of pairs of cells manifest statistical signatures of monosynaptic connections \cite{Gerstein:1964,  Perkel:1969, Moore:1970, Gerstein:1989, Alonso:1998, Denman:2013}.
 Other methods require temporal resolution on the order of membrane time constants (milliseconds to tens of milliseconds): \emph{e.g.}~statistical models that mimic membrane integration (\cite{Pillow:2008}) and information-theoretic and statistical physics models that bin spike trains finely to ensure a binary representation \cite{Schneidman:2006, Tkacik:2006, Tkacik:2013}.
@@
 
-Multielectrode recordings implanted in the cortex record from relatively small groups of cells separated by relatively long distances.
-Perhaps for this reason, the structure of partial correlations has not been, to our knowledge, been used to describe the functional connectivity in multielectrode recordings.
-In contrast, two-photon calcium imaging allows simultaneous measurements of the activity of nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
+When inserted in cortical tissue, electrode arrays record from a small fraction of cells in a given volume.
+Perhaps due to such sparse sampling of neurons, the structure of partial correlations has not, to our knowledge, been used to describe the functional connectivity in multielectrode recordings.
+Two-photon calcium imaging allows simultaneous measurements of the activity of nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
 Therefore, two-photon imaging opens opportunities for computing the structure of conditional dependencies between neurons with arguably closer correspondence to the underlying physiological interactions.
@@
 \paragraph{Finding improved representation of the correlation structure}
-We were initially motivated by the somewhat simplistic hypothesis that partial correlations in local cortical microcircuits should provide more meaningful representations of the functional connectivity than the conventional correlations.
-Since neurons are known to form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we imposed the constraint that most of the partial correlations must be zero.
-To account for interactions with the rest of the brain, we applied the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables\cite{Chandrasekaran:2010, Ma:2013}.
+We began this study from the somewhat simplistic hypothesis that the pairwise partial correlations in local cortical microcircuits should correspond more closely to the synaptic or effective connectivity than the sample correlations.
+Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we required that the partial correlations be sparse.
+To account for interactions with the rest of the brain, we applied the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables \cite{Chandrasekaran:2010, Ma:2013}.
 A priori, the validity of this assumption was debatable:  On the temporal scales of 150 ms, the pairwise effects between neurons could fade in prominence compared to the co-activation of larger cell assemblies and global circuit fluctuations, which would be poorly fitted by sparse partial correlations.
@@
 
-To provide evidence for the validity of this hypothesis, we compared the effect of imposing this structure on the estimation of the noise covariance matrices relative to estimations based on other assumptions.
-The finding that the sparse+latent structure of correlations led to better estimation than the other estimators lends a degree of credence to its structural representation of the circuit's functional connectivity.
+To verify our hypothesis, we compared the effect of imposing this structure on the estimation of the noise covariance matrices relative to estimations based on other assumptions.
+The finding that the sparse+latent structure of correlations led to better estimation than the other estimators lends a degree of credence to its structural representation of the circuit’s functional connectivity.
 Of course, this finding could be overturned or augmented by finding an even more efficient covariance estimator for these types of recordings and we welcome the challenge.
@@
 In our study, the covariance matrix estimators were evaluated with respect to the cross-validated normal log likelihood.
-This does not limit the applicability of its conclusion normal distributions.
-Indeed the major findings in this paper could be reproduced with respect to other loss functions (compare Fig.\ref{fig:04} and Fig.~S\ref{supp:02}).
-Other probabilistic models, fitted to the same data, could be used to produce their respective estimates of the covariance matrix.
-If another model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of functional connectivity.
-We advocate comparing models by how well they estimate the covariance matrix due to their well established relevance to connectivity and neural coding.
+This does not limit the applicability of its conclusion to normal distributions.
+Indeed the major findings in this paper could be reproduced with respect to other loss functions (compare Fig.~\ref{fig:04} and Fig.~S\ref{supp:02}).
+Other probabilistic models, fitted to the same data, would also serve as estimators of the covariance matrix.
+If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of functional connectivity.
 
@@
 These methods have numerous caveats when it comes to inferring the optimal structure of functional connectivity from such methods.
-Improved estimation of covariances and the general framework of producing such estimators for specific neural populations can be applied to various circuits and serve as a statistical description of their functional organization.
-
+Improved estimation of covariances and the general framework of producing such estimators for specific neural populations can be applied to various circuits and serve as a statistical description of their functional organization
 \section*{Methods}
@@
 The collected fluorescent traces were deconvolved to reconstruct the firing rates for each neuron.
-First, the first principal component was subtracted from the traces, which reduced common mode noise related to small movement artifacts and cardiovascular artifacts.
+First, the first principal component was subtracted from the traces, which reduced common mode noise related to small movement and cardiovascular artifacts.
 The resulting traces were low-pass filtered below 0.1 Hz and downsampled to 20 Hz.
@@
 
-\begin{figure}
+\begin{figure}[!ht]
 \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]

### 2014-01-04 plos.tex
Changes in Introduction, Discussion, Results

@@
 Indeed, in cross-validation tests, the covariance matrix estimator with such structure consistently outperformed other regularized estimators.
-The connectivity patterns inferred from the sparse component of the estimate manifested a strong relationship to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distance and difference in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
+The connectivity patterns inferred from the sparse component of the estimate manifested a strong relationship to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distance and the difference in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
 Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than sample correlations.
@@
 % Author Summary not valid for PLoS ONE submissions.   
-Correlations in the activity of neural populations have proven useful to describe their functional connectivity.
+Correlations in the activity of neural populations have proven useful as descriptions of the functional connectivity in the embedding circuit.
 Correlations may arise from various underlying processes that reflect differently on the structure of the correlation matrix:  Common network fluctuations or external inputs are well approximated by low-rank representations whereas linear effects between specific pairs of neurons are well approximated by fitting their pairwise partial correlations.
@@
 In fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the greatest improvement was produced by decomposing the correlation matrix as a sparse network of partial correlations and a low-rank component.
-Since it leads to the best estimates, we propose that this structure provides a better picture of the functional connectivity than that obtained from the usual correlations.
+Since it leads to the best estimates, we propose that this structure provides a better picture of the functional connectivity than that obtained from sample correlations.
 As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and to differences in their preferred orientations and found basic resemblance to results from studies of synaptic connectivity.
@@
 For example, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
-In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits and reflect interactions between neurons \cite{Gerstein:1964}.
-This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.
+In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the functional connectivity in neural circuits \cite{Gerstein:1964}.
+This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.
 
@@
 Multivariate analysis techniques based on large correlation matrices provide information that could not be extracted from an equivalent number of isolated measurements of pairwise correlations.
-For example, when the correlation matrix is approximated in a form with a reduced number of variables, this \emph{reduced representation} can help isolate various aspects of the functional connectivity in neural circuits and possibly point to underlying physiological mechanisms and computational principles.
+For example, when the correlation matrix is approximated in a form with a reduced number of variables, this \emph{reduced representation} can isolate an important aspect of the functional connectivity in neural circuits and possibly point to underlying physiological mechanisms and computational principles.
 
 One reduced representation of the correlation matrix is its low-rank approximation.
-Low-rank approximations are particularly suitable for capturing common fluctuations or common external inputs into the circuit and can be extract using principal component analysis, factor analysis, or  \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.
+Low-rank approximations are particularly suitable for capturing common fluctuations and common external inputs and can be extracted using principal component analysis, independent component analysis, factor analysis, or  \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.
 
-Another reduced representation of the correlation matrix can be obtained by \emph{covariance selection}, \emph{i.e.}\;by selecting a subset of the pairwise \emph{partial correlation} that are sufficient to reproduce the entire correlation matrix \cite{Dempster:1972, Friedman:2008}.
+Another reduced representation of the correlation matrix can be obtained by \emph{covariance selection}, \emph{i.e.}\;by selecting a subset of the pairwise \emph{partial correlations} that are sufficient to reproduce the entire correlation matrix \cite{Dempster:1972, Friedman:2008}.
 The partial correlation between a pair of neurons is a measure of their linear association after the effects of the rest of the network have been taken into account \cite{Anderson:2003}.
-Partial correlations are particularly meaningful when relationships between variables are mostly linear and large fraction of the variables are observed.
-In this case, partial correlations approximate conditional dependency.
+Partial correlations are particularly meaningful when all effects between the variables are nearly linear and a large fraction of the variables are observed; in this case, partial correlations approximate conditional dependency.
 A lack of conditional dependency between two neurons suggests a lack of direct interaction between them.
@@
 \paragraph{Improved estimation performance by regularization}
-Imposing some structure on an estimate of the covariance matrix can improve its estimation performance.
+Imposing a structure on an estimate of the covariance matrix can improve its estimation performance.
 
@@
 
-In statistics, the technique of deliberately imposing some structure on an estimate in order to improve its performance is known as \emph{regularization} \cite{Schafer:2005, Bickel:2006}.
+In statistics, the technique of deliberately imposing a structure on an estimate in order to improve its performance is known as \emph{regularization} \cite{Schafer:2005, Bickel:2006}.
 To \sq{impose structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation known as the shrinkage target.
 The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data either analytically or by cross-validation.
-Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely, without shrinkage.
+Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely.
 Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).
@@
 
-Although regularized covariance estimates have become commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.
+Although regularized covariance matrix estimates have become commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.
 \paragraph{Illustration of estimation of a neural correlation matrix}
@@
 Zeros in the off-diagonal elements of the precision matrix indicate zero partial correlations between the corresponding pair of neurons.
-Therefore, covariance selection amounts to finding an approximation of the covariance matrix so that a large fraction of neuron pairs are linearly independent of each other when conditioned on the activity of all the other observed cells.
+Therefore, covariance selection amounts to finding an approximation of the covariance matrix that maximizes the number of neuron pairs that are linearly independent of each other when conditioned on the activity of all the other observed cells.
 
@@
 \paragraph{Conceptual questions and basic approach}
-What structures of correlation matrices are most suitable for the description of the multineuronal activity in specific neural circuits and brain states?
-For example, are correlations in visual cortex under during visual stimulation best explained by common fluctuations or by local interactions within the recorded microcircuit?
+What structures of correlation matrices best describe the multineuronal activity in specific circuits and in specific brain states?
+More specifically, are correlations in the visual cortex during visual stimulation best explained by common fluctuations or by local interactions within the recorded microcircuit?
 
-We pursued two related aims: (a)~the improved estimation of neural correlation matrices and (b)~discovery of the low-dimensional structure of correlations in recordings of multineuronal activity.
-An accurate characterization of the patterns of correlations helps in interpreting them, and to better relate them to the architecture and dynamics of the underlying neuronal networks.
-
- In this study, we compared the sample covariance matrix and four regularized estimators based on distinct correlation structures:
+To address these questions, we evaluated five covariance matrix estimators based on different assumptions about the underlying correlation structure:
 \begin{description}
-\item[$C_{\sf 0}$] -- sample covariance matrix, the usual estimator used in previous studies.
+\item[$C_{\sf 0}$] -- sample covariance matrix, the unbiased, unregularized estimator.
 \item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix. 
-\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis with an optimal number of factors.
-\item[$C_{\sf sparse}$] -- sparsified precision matrix,  \emph{covariance selection} \cite{Dempster:1972, Friedman:2008}, without shrinkage.   
-\item[$C_{\sf sparse+latent}$] -- precision matrix composed as the sum of a sparse and a low-rank component \cite{Chandrasekaran:2010, Ma:2013}.
+\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis.
+\item[$C_{\sf sparse}$] --  sparse partial correlations, \emph{i.e.}\;covariance selection.
+\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons combined with partial correlations with several inferred latent units.
 \end{description}
 
-First, we demonstrated that, in simulations with known correlation structures, regularized estimators with matching structure were generally more efficient than the other estimators.
+First, we demonstrated that, in simulations with known true correlation structures, estimators that imposed a matching structure were generally more efficient than the other estimators.
+The estimators’ relative performance was evaluated by cross-validation.
+The logic of this study relies on the converse of this relationship: the covariance matrix estimator that is found to be most efficient on empirical data is proposed to provide a better picture of the correlation structure than other estimators.
+Caveats for this kind of inference are reviewed in the Discussion.
 
-We also performed a cross-validated evaluation to establish the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex.
+Therefore, we performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with fast 3D random-access two-photon imaging of calcium signals.
 We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.
 This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units.
-Since it led to better estimates, we propose that this structure provides a more physiologically relevant representation of the circuit’s functional connectivity than that obtained from the sample correlation matrix.
-We showed that the partial correlation structure had stronger dependence on physical distances and differences in orientation preference than sample correlations.
-In Discussion, we outline possibly approaches to corroborate the relationship between the structure of the functional connectivity revealed by this and similar method to the anatomical organization of the circuit.
+We examined the structure of the sparse partial correlations inferred by the estimate with respect to the neurons’ geometrical positions and preferred orientations and found that stronger dependencies than those of sample correlations.
+In Discussion, we outline possible approaches to further corroborate the relationship between the structure of the functional connectivity revealed by such methods to the anatomical organization of the circuit.
 
-The true correlation matrix is the normalized version of the true covariance matrix defined as
+\section*{Results}
+% Results and Discussion can be combined.
+\paragraph{Covariance estimation}
+The icovariance matrix defined as
 \begin{equation}\label{eq:true-covariance}
@@
 
-\section*{Results}
-% Results and Discussion can be combined.
-\paragraph{Covariance estimation}
 We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
@@
 
-For the first regularized estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ whose diagonal contains  variance estimates.
+For estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ whose diagonal contains  variance estimates.
 Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
@@
 \end{equation}
-The diagonal target has structure corresponding to the absence of linear associations between the activity of observed neurons (\figref{02}{A}).
-If this structure accurately describes recorded data, then strong shrinkage toward $D$ adds little bias while strongly reducing the variability of the estimate.
-Shrinkage allows for partial commitment to the low-dimensional representation.
+The structure of target $D$ favors the absence of linear associations between the activity of observed neurons (\figref{02}{A}).
+If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ should outperform the other estimators.
 
-The target of the second regularized estimator $C_{\sf factor}$, is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix of rank $d$, and $\Psi$ is the diagonal matrix of independent variances.
+The target of estimator $C_{\sf factor}$ is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix of rank $d$, and $\Psi$ is the diagonal matrix of independent variances.
 The estimator,
@@
 has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $L + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
+The target estimate $L + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
 
-The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between pairs of observed cells (\figref{02}{C}).
+Estimator $C_{\sf sparse}$ is based on the assumption that correlations result from direct linear associations between pairs of observed neurons and that only a fraction of neuron pairs interact (\figref{02}{C}).
 This assumption is enforced by fitting only a subset of the off-diagonal elements of the precision matrix while setting the rest to zero.
-Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{02}{C}).
 The estimator takes the form
@@
 \end{equation}
-where $S$ is a sparse matrix with a large fraction of zeros off the diagonal. The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.
+where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal.  The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.
 
-The fourth estimator, $C_{\sf sparse+latent}$, combines the assumptions of the previous two.
-The underlying structure arises from sparse interactions between the observed neurons and several latent units (\figref{02}{D}).
+Estimator $C_{\sf sparse+latent}$ allow the assumptions of both $C_{\sf factor}$ and $C_{\sf sparse}$.
+Its correlation structure can be explained by linear effects exerted between some pairs of observed neurons and several latent units (\figref{02}{D}).
 The estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
@@
 \end{equation}
-where, as above, $S$ is a sparse matrix and $L$ is a $p\times p$ matrix of rank $d$. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.
+where, as above, $S$ is a sparse matrix and $L$ is a $p\times p$ matrix of rank $d$.
+The estimator has two hyperparameters: one to regulate the number of latent units, the other to regulate the sparsity of matrix $S$.
 
@@
 
-To evaluate the performance of a covariance matrix estimator $C$, we define a real-valued \emph{loss function} $\loss{C,\Sigma}$ that attains a minimum when $C=\Sigma$.
-The loss function quantifies the error of the estimate, \emph{i.e.}~the deviation of the estimate $C$ from truth $\Sigma$.
+An evaluation of the performance of a covariance matrix estimator $C$ requires a real-valued \emph{loss function} $\loss{C,\Sigma}$.
+The loss loss function quantifies the deviation of the estimate $C$ from truth $\Sigma$ and attains its minimum when $C=\Sigma$.
 
-In this study, we adopted \emph{negative normal log-likelihood loss} defined as
+In this study, we adopted the \emph{normal loss} defined as
 \begin{equation}\label{eq:loss}
@@
 \end{equation}
-The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelihood function $L(\Sigma|C_{\sf 0})$ as
+
+The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelih\emph{normal log-likelihood loss} defined as ood function $L(\Sigma|C_{\sf 0})$ as
 \begin{equation}
@@
 \end{equation}
+Although the normal log-likelihood loss and the normal log-likelihood appear similar in form, they differ conceptually: The log-likelihood is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the distance from the estimate $C$ to the truth $\Sigma$.
 
@@
 
-In Figure \ref{fig:02}, row 4 shows the mean excess losses and their standard errors calculated from 30 samples of sizes n=250, 500, 1000, 2000, and 4000 for each estimator and each ground truth.
-For each estimator, hyperparameters were estimated by nested cross-validation (See Methods for more details.)
-
-As expected, estimators with whose structure matched that of the true model typically outperformed the other estimators.
-There were two exceptions: First, for small sample sizes, there was insufficient data to reveal the true correlation structure.
-In this case estimates with simpler targets often outperformed the estimator based on the correct model.
-Second, the sparse+latent estimator often performed equally well to the sparse estimator even when the data did not have any latent units.
-In these cases, it inferred the right number of latent units: zero.
-In larger models, the falsely identified latent factors are more common.
+We computed the excess losses for 30 samples for each sample size $n=250$, 500, 10000, 2000, and 4000 for each of the five estimators and for each ground truth (\figref{02}{\,row 5}).
+For each estimator, hyperparameters were estimated by nested cross-validation (See Methods.).
+As expected, estimators whose structure matched that of the true model typically outperformed the other estimators.
+There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with the simpler structure, $C_{\sf diag}$ and $C_{\sf factor}$ outperformed the estimators with matching structure (\figref{02}{\,columns C and D, rows 5 and 6}).
+Second, $C_{\sf sparse+latent}$ often performed equally well to the sparse estimator even when the ground truth did not have any latent units.
+In these cases, it inferred the right number of latent units, zero, and estimated optimally.
 
@@
 \begin{equation*}
-    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}
+    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}^\prime]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}
 \end{equation*}
@@
 The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses.
-With non-additive loss functions, cross-validation loss is a biased estimate of loss, but it is still generally appropriate to be used in lieu of loss.
+With non-additive loss functions, validation loss is a biased estimate of loss, but it is still generally appropriate to be used in lieu of loss.
 
@@
 Instead, $K$-fold cross-validation is used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).
-Then each subset is used as the validation sample with the other $K-1$ serving as a training dataset.
-The validation losses from each of such \sq{folds} are averaged to produce \emph{cross-validation loss} or \emph{CV-loss} for short.
+Then each subset is used as the validation sample with the other $K-1$ serving as the training sample.
+The validation losses from each of such \sq{folds} are averaged to produce the \emph{cross-validation loss} or \emph{CV-loss} for short.
 Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ trained on the remaining $K-1$ subsets.
@@
 \end{equation}
-In the present formulation, $\ell_C$ is known as the cross-validated Gaussian log-likelihood (up to a constant offset and multiplier).
+In this formulation, $\ell_C$ can be recognized as the cross-validated Gaussian log-likelihood (up to a constant offset and multiplier).
+However, the present framework is more general, allowing for other loss functions to be used.
 
-Since we only need to compare estimators to each other, we are only interested in \emph{relative} CV-loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$:
+Since we only need to compare estimators to each other, we are only interested in the \emph{relative} CV-loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$:
 \begin{equation}\label{eq:rel-cv-loss}
-    \ell_{C,C_{\sf ref}} = \frac 1 K \sum\limits_{k=1}^K \left[
-        \loss{C^{\{\setminus k\}},C_{\sf 0}^{\{k\}}} -
-    \loss{C_{\sf ref}^{\{\setminus k\}},C_{\sf 0}^{\{k\}}} 
-\right]
+    \Delta\ell_{C,C_{\sf ref}} = \ell_{C}-\ell_{C_{\sf ref}}
 \end{equation}
 
-In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (Figure \ref{fig:02}, Row 5).
+In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (\figref{02}{Row 6}).
 For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth.
 
-These simulations demonstrate that with sufficiently large sample sizes the most efficient among several regularized estimators could be used to infer the most likely type of low-dimensional structure present in the data.
+These simulations demonstrated that, with sufficiently large sample sizes, the most efficient among several regularized estimators could be used to infer the most likely type of correlation structure present in the data.
 
@@
     All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
-    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
+    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}.
 \\{\bf A.} $C_{\sf sparse+latent}$ outperforms $C_{\sf 0}$: median improvement 0.16 nats/neuron, $p=3.9\times 10^{-5}$.
@@
 By design, the sparse+latent estimator finds the mixture of sparse partial correlations between visible neurons and common fluctuations described as latent units.
-If this estimate dominates, we can hypothesize that it better describes underlying physiological interactions than other estimators.
+If this estimate dominates, we can hypothesize that it better describes underlying physiological interactions than other estimators, including the sample covariance.
 In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
@@
 
-\paragraph{Correspondence to anatomical organization}
-If the ultimate goal of this line of inquiry to find descriptions of functional connectivity that are most informative about the anatomical and physiological organization of the circuit, then the models must be directly validated according to their correspondence to the anatomical organization.
-Little empirical evidence presently exists to show which models of connectivity, when fitted to the recordings of activity, reveal functional connectivity with best correspondence to anatomical connectivity.
-
-In this study we find that the structure of partial correlation of the optimal estimator had stronger dependencies on the separation in depth, lateral position, and orientation preferences than sample correlations (\figref{07}{A--C}).
-The inferred positive and negative partial correlations too had different structures COMPARED TO NOISE CORRELATIONS ??
-(\figref{07}{D--F}).
-Also the average partial correlations were more consistent between imaged sites than average sample correlations (\figref{06}{A}).
-Although these effects could conceivably results from trivial numerical artifacts, they are consistent with improved representation of functional connectivity by the proposed estimator.
-
-Additional studies with additional markers of circuit architecture such as cell types and synaptic connectivity could shed additional light on the ability of models to reveal meaningful functional connectivity.
-For example, labeled interneurons could be studied for their distinct connectivity patterns contrasted with pyramidal cells.
-I THIKN WE NEED ONE LAST SENTNECE HERE TO SAY THAT FUNCTIONAL CONNECTIVITY IF IT RELATED TO ANALTOMICAL IS IMPORTANT SINCE IT CAN TELL US HOW CIRCUITS INTERACT TO COMPUTE MENTION EXAMPLE OF STATES, SLEEP HERE ETC.
-I.E.
-GROUPS CELLS MAY BE CONNECTED BUT ONLY INTERACTING IN STATE A AND NOT B.
-
-\paragraph{Challenges of big neural data}
-The extraction and interpretation of functional connectivity depended on the density and size of the recorded neuronal population.
-
-In this study, we sought to uncover the optimal low-dimensional approximation of the covariance matrix for efficient estimation of the true covariance matrices in cortical microcircuits on the temporal scale of 100--200 ms corresponding to the temporal resolution of spike rates from somatic calcium signals \cite{Cotton:2013}.
-We found that networks of partial correlations were more statistically efficient than low-rank approximations of correlation matrices (Fig.\;S\ref{supp:01}\;E).
-Furthermore, estimation became still more efficient when low-rank components were allowed to be combined with sparse partial correlations.
-We propose that correlation structures that are supported by the data  are more likely to reflect anatomical, causal interactions.
-The network of sparse partial correlations suggests interactions between specific pairs whereas the low-rank represents emergent activity patterns involving collective activations of larger circuits or global fluctuations.
-The revealed network of partial interactions was substantially different from that revealed by thresholding correlations (\figref{05}{C, F, H, I}).
-In particular, it shows that many low correlations correspond to strong negative interactions (\figref{05}{F}).
-Intuitively, this suggests that many low interactions are as surprising or significant considering the common correlations with other neurons.
-The analysis of sample correlations does not ascribe significance to correlation coefficients that are surprisingly low, based on their common correlations to other cells.
-
-Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}.
-It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.
-Large datasets aggravate the numerical problems and the model selection caveats described in this paper.
-The number of parameters in models of population activity grows quadratically.
-Models of joint population activity feature large numbers of parameters, that growing quadratically or faster with population size.
-Without ways to restrict such models, they quickly become ill conditioned (e.g.
-\cite{Roudi:2009}).
-Statistical models of population activity will need to have both correct mathematical forms of interactions and optimal ways of restricting them so that the total number of estimated parameters remains commensurate with the amount of measurements.
-
-Our study employs regularization, which restricts the models based on the data itself.
-Other models may use restrictions that are based on various heuristics in addition to regularization.
-These methods have numerous caveats when it comes to inferring the optimal structure of functional connectivity from such methods.
-Improved estimation of covariances and the general framework of producing such estimators for specific neural populations can be applied to various circuits and serve as a statistical description of their functional organization
 \section*{Methods}

### 2014-01-06 plos.tex
complete Discussion

@@
 In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the functional connectivity in neural circuits \cite{Gerstein:1964}.
-This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.
+This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.
 
@@
 
-\paragraph{Improved estimation performance by regularization}
+\paragraph{Improving estimation through regularization}
 Imposing a structure on an estimate of the covariance matrix can improve its estimation performance.
@@
 Estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
-This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned estimation of the partial correlations \cite{Schafer:2005}.
-In neuroscience studies, the true correlation matrix is usually estimated by the \emph{sample correlation matrix} and these two concepts are rarely distinguished.
+This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
+
+In neuroscience studies, the true correlation matrix is usually estimated by the \emph{sample correlation matrix}, which is rarely distinguished from the true correlation matrix.
 The sample correlation matrix is unbiased but, because it has many free parameters, it is highly sensitive to sampling noise.
@@
 The logic of this study relies on the converse of this relationship: the covariance matrix estimator that is found to be most efficient on empirical data is proposed to provide a better picture of the correlation structure than other estimators.
-Caveats for this kind of inference are reviewed in the Discussion.
+Caveats for this type of inference are reviewed in the Discussion.
 
@@
 In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.
+In the case of noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.
 
@@
 
-Various other regularization schemes also produce improved estimates
-
 We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
-In probabilistic systems with linear dependencies, these correspond to graphical modes with distinct types of structure  (\figref{02}{row 1}).
+In probabilistic models with exclusively linear dependencies, their targets estimates can be described by distinct families of graphical models (\figref{02}{row 1}).
 
@@
 
-The loss function $\loss{C,\Sigma}$ is related to the multivariate log-likelih\emph{normal log-likelihood loss} defined as ood function $L(\Sigma|C_{\sf 0})$ as
+This loss function relates to the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$ through the following expression:
 \begin{equation}
@@
 \end{equation}
-Although the normal log-likelihood loss and the normal log-likelihood appear similar in form, they differ conceptually: The log-likelihood is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the distance from the estimate $C$ to the truth $\Sigma$.
+Although the two functions appear similar in form, they differ conceptually: The log-likelihood is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the distance from an arbitrary estimate $C$ to the given truth $\Sigma$.
 
-We normalized the loss function by $\frac 1 p$ to make its value comparable across different population sizes.
+The normalization by $\frac 1 p$ is not standard but serves to make the values of the loss functions comparable across different population sizes.
 
@@
 We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (\figref{03}{A}) on one side of the visual field of anesthetized mice.
-Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E} in the visual cortex on the contralateral side to the stimulus.
+Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E}) in the visual cortex on the contralateral side to the stimulus.
 Using acousto-optic deflectors (AOD) to steer the laser in 3D, we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}.
@@
 \paragraph{Functional connectivity}
-How the anatomical organization of neural circuits gives rise to their function is a foremost question in neuroscience.
-Effective phenomenological descriptions of neural activity, once discovered, enable investigation of underlying physiological mechanisms.
-For example, the discoveries of receptive fields and tuning properties of single neurons, over the past 60 years, have led to fruitful hypotheses about the underlying patterns of synaptic connectivity, principles of neural coding, and to the discovery of other organizational principles such as various cortical feature maps \cite{Reid:2012}.
+Effective phenomenological or statistical descriptions of neural activity can serve as a starting point for investigation of underlying physiological mechanisms.
+For example, the discoveries of receptive fields and tuning properties of single neurons, over the past 60 years, have led to fruitful hypotheses about the underlying patterns of synaptic connectivity and principles of neural coding, and to the discovery of other organizational principles such as various cortical feature maps \cite{Reid:2012}.
+The search for similarly effective descriptions of observed \emph{multineuronal} activity patterns that are not reducible to the response properties of the individual cells has produced various statistical expressions collectively known as \sq{functional connectivity}.
+Functional connectivity is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons and is thought to reflect internal local recurrent connectivity, shared interactions with other regions, or endogenous network activity.
+To remove the confounding effects of stimuli, the average stimulus responses are usually subtracted or otherwise accounted for.
+Although various measures of functional connectivity are only phenomenological descriptions of population activity, they provide experimental neuroscientists with tools to relate population activity to the anatomical and functional architecture of neural circuits and to generate hypotheses about underlying mechanisms.
 
-Similarly, the term \sq{functional connectivity} denotes a statistical summary of observed \emph{multineuronal} patterns of activity thought to reflect internal, recurrent connectivity in neural circuits.
-In statistical terms, functional connectivity is an empirical estimate of a feature of the joint probability distribution of the neurons’ activity.
-Although functional connectivity may not provide a detailed and unambiguous interpretation of underlying mechanisms, its utility in experimental neuroscience is that of a descriptive statistic to quantify the general organization of synchronous activations of neuronal populations.
-Functional connectivity is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons.
-To remove the confounding effects of stimuli, the average stimulus responses are usually subtracted or otherwise accounted for.
+\paragraph{Functional connectivity as a network of pairwise interactions}
+The intended purpose of many studies of functional connectivity is to estimate the circuit’s anatomical connectivity solely from the observed multineuronal spiking activity.
+For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections or shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
+Such signatures are still ambiguous as they can arise from diverse network effects other than direct synaptic connections \cite{Aertsen:1989}.
+With simultaneous recordings from a larger subset of neurons in the circuit, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.
+Since direct causal interactions between neurons produce statistical dependency even after conditioning on the state of the rest of the network, conditionally independent neurons can be inferred to lack a direct causal influence.
+Such conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity.
+For example, generalized linear models (GLMs) have been constructed to include biophysically plausible synaptic integration and membrane kinetics, individual neurons’ stimulus drive, and a network of synaptic couplings \cite{Pillow:2008}.
+Another family of models are maximum entropy models that are constrained by the observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2009, Shlens:2009} .
+Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with fewest other assumptions about the distribution: the investigator need only choose which statistics to match.
+Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
+Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \emph{i.e.}\;conditional dependencies between pairs of neurons.
 
-Pairwise correlations (including cross-correlograms and joint peristimulus histograms) between the spiking activities of neurons boast a rich history as a measure of functional connectivity.
-Interpreted as either the statistical signatures of synaptic connectivity \cite{Gerstein:1964, Perkel:1969, Denman:2013} or of Hebbian neuronal assemblies \cite{Gerstein:1989, Lopes:2013}, pairwise correlations have been instrumental in discovering various principles of the functional organization of neural circuits, as only partially enumerated in the Introduction.
-Several other measures of functions connectivity have emerged including probabilistic models
+The inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the interaction between a pair of neurons on the activity of the other neurons under a given probabilistic model, the recording must include all the other neurons with which the pair can potentially interact.
+Otherwise, interactions with the unobserved portions of the circuit will introduce false conditional dependencies between the observed neurons.
+For this reason, the most successful applications of statistical models of population activity thus far have been in in vitro preparations of the retina or cell cultures where high-quality recordings from the complete populations have been possible \cite{Pillow:2008}.
+When inserted in cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the conditional dependencies between pairs of cells.
+Perhaps for this reason, until our study, partial correlations have not been used to describe the functional connectivity between cortical neurons.
+Despite this limitation, many studies have fitted probabilistic models to multielectrode recordings and analyzed the structure of the resulting conditional dependencies \cite{Tang:2008, Yu:2008}.
 
-Parsimony is a desired attribute of a measure of functional connectivity: insignificant or contingent detail.
+Because the validity of the inference of interactions from probabilistic models depends on both the correctness of their mathematical form in the model and the completeness of the recorded population, it is far from clear which approach provides best correspondence to the anatomical connectivity in the circuit, with no experimental evidence yet available to address this question.
+The result is often treated as a graphical representation of the functional connectivity without a statement about its anatomical implications.
+Whatever these implications may be, investigators have applied graph-theoretical analyses to the inferred networks to make statements about the design principles and efficiency of circuit architectures \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
 
-\paragraph{Temporal scales of functional connectivity}
-Analyses of functional connectivity depend profoundly on their temporal resolution.
-At millisecond-scale resolution, cross-correlograms of the spiking activity of pairs of cells manifest statistical signatures of monosynaptic connections \cite{Gerstein:1964,  Perkel:1969, Moore:1970, Gerstein:1989, Alonso:1998, Denman:2013}.
-Other methods require temporal resolution on the order of membrane time constants (milliseconds to tens of milliseconds): \emph{e.g.}~statistical models that mimic membrane integration (\cite{Pillow:2008}) and information-theoretic and statistical physics models that bin spike trains finely to ensure a binary representation \cite{Schneidman:2006, Tkacik:2006, Tkacik:2013}.
-At slower scales of tens to hundreds milliseconds, temporal sequencing of spikes can no longer resolve directions of monosynaptic interactions and any expression of functional connectivity summarizes chains of synaptic transmissions with more complex, emergent properties of network computation: the analysis shifts from to detecting co-active cell assemblies and the low-rank structure of correlations \cite{Ch:2010, Lopes:2011, Lopes:2013}.
-
+Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity despite some of the disadvantages.
 The temporal resolution of calcium signals is limited by calcium dye kinetics.
 Fast imaging techniques combined with advanced spike inference algorithms have been demonstrated to provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010, Katona:2012, Cotton:2013}.
-However, high temporal precision comes at the cost of the accuracy of the inferred spike rates: better accuracy can be achieved when calcium signals are analyzed on the temporal scales of tens of milliseconds \cite{Cotton:2013}.
-Therefore, functional connectivity studies from calcium signals have focused on the slower temporal scales (tens or hundreds of milliseconds).
-In this study, too, we chose 20 Hz for the temporal precision of spike inference and bin widths of 150 ms for computing correlations.
+However, such high temporal precision comes at the cost of the accuracy of the inferred spike rates: better accuracy can be achieved when calcium signals are analyzed on the temporal scales of tens of milliseconds \cite{Cotton:2013}.
+For what calcium imaging lack in temporal resolution, it make up in spatial resolution and the ability to characterize the spatial arrangements of cells and identify their types.
+Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
+Such completeness may open new opportunities for inferring networks of conditional dependencies between neurons with greater validity than with electrophysiological recordings.
 
-\paragraph{Estimation of functional connectivity depends on recorded population size and density}
-Analyses of functional connectivity depend profoundly on the sizes and densities the recorded population.
-Pairwise recordings may contain statistical signatures of monosynaptic connections but conditioning on the activity of all the other cells may help disambiguate the network of interactions in the circuit.
-This distinction has been described as \emph{effective connectivity}: the best guess at the most direct effects exerted by neurons on each other \cite{Gerstein:1989}.
-For example, the interaction terms in spatiotemporal generalized linear models \cite{Pillow:2008} or Ising-like maximum entropy models \cite{Schneidman:2006, Tkacik:2006, Ganmor:2011} express the conditional dependencies between neurons, which only have weak relationship with the pairwise statistics to which they are fitted.
-If neurons can be assumed to exert nearly linear effects on each other, then conditional dependencies between neurons are expressed by the partial correlations.
-For effectual computation of conditional dependencies, the recording must include a sufficient fraction of the neurons in the embedding circuit, or even the complete population.
-Otherwise, interactions with the unobserved portions of the circuit will introduce false conditional dependencies between the observed neurons.
-For this reason, the most successful applications of statistical models of population activity thus far have been in in vitro preparations of the retina or cell cultures where high-quality recordings from the complete populations have been possible.
+The low temporal resolution of calcium signals limits the use of functional connectivity methods that rely on millisecond-scale binning of signals (cross-correlograms, some GLMs, and maximum entropy models on the binary domain).
+Most studies of functional connectivity have relied on instantaneous sample correlations \cite{Greenberg:2008, Golshani:2009, Hofer:2011, Malmersjo:2013} .
+Although some investigators have interpreted such correlations as indicators of (chemical or electrical) synaptic connectivity, most used them as more general indicators of functional connectivity without a statement about underlying mechanisms.
 
-When inserted in cortical tissue, electrode arrays record from a small fraction of cells in a given volume.
-Perhaps due to such sparse sampling of neurons, the structure of partial correlations has not, to our knowledge, been used to describe the functional connectivity in multielectrode recordings.
-Two-photon calcium imaging allows simultaneous measurements of the activity of nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
-Therefore, two-photon imaging opens opportunities for computing the structure of conditional dependencies between neurons with arguably closer correspondence to the underlying physiological interactions.
+In our study, we sought to express pairwise connectivity networks among neurons in local cortical microcircuits.
+We hypothesized that partial correlations would correspond more closely to underlying mechanisms than sample correlations.
+Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we a priori favored solutions in with sparse partial correlations.
+The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution.
+The interactions (conditional dependencies) between pairs of neurons inferred by this distributed are expressed by the precision matrix (inverse of the covariance) or, equivalently, by the matrix of the pairwise partial correlations.
+Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than sample correlations.
 
-\paragraph{Finding improved representation of the correlation structure}
-We began this study from the somewhat simplistic hypothesis that the pairwise partial correlations in local cortical microcircuits should correspond more closely to the synaptic or effective connectivity than the sample correlations.
-Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we required that the partial correlations be sparse.
-To account for interactions with the rest of the brain, we applied the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables \cite{Chandrasekaran:2010, Ma:2013}.
-A priori, the validity of this assumption was debatable:  On the temporal scales of 150 ms, the pairwise effects between neurons could fade in prominence compared to the co-activation of larger cell assemblies and global circuit fluctuations, which would be poorly fitted by sparse partial correlations.
-In addition, strong nonlinearities in interactions between neurons could disrupt the relationship between conditional independence and absence of partial correlation.
+\paragraph{Functional connectivity as coactivations}
+Another approach to describe the functional connectivity of a circuit is to isolate patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}.
+Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}.
+Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \emph{i.e.}\;groups of tightly interconnected groups of cells involved in a common computation.
+Coactivation patterns could also be explained as shared input from unobserved parts of the circuit.
+Coactivations could also be explained by global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.
 
-To verify our hypothesis, we compared the effect of imposing this structure on the estimation of the noise covariance matrices relative to estimations based on other assumptions.
-The finding that the sparse+latent structure of correlations led to better estimation than the other estimators lends a degree of credence to its structural representation of the circuit’s functional connectivity.
-Of course, this finding could be overturned or augmented by finding an even more efficient covariance estimator for these types of recordings and we welcome the challenge.
+Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity, but analysis of coactivations method shifts the focus from detailed interactions to the collective behavior.
 
+In our study, the analysis of functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.
+
+\paragraph{Combining pairwise interactions and coactivations}
+In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables \cite{Koster:2013},  high-order correlations \cite{Ganmor:2011}, or global network fluctuations \cite{Tkacik:2013}.
+
+In our study, we combined pairwise interactions with collective coactivations by applying the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables \cite{Chandrasekaran:2010, Ma:2013}.
+The resulting estimator, $C_{\sf sparse+latent}$, effectively decomposed the functional connectivity into a sparse network of pairwise interactions and coactivation mode vectors.
+\paragraph{Addressing ill-posedness}
+Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.
+The problem becomes increasingly worse as the size of the recorded population of neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
+As the recorded neuronal population sizes increased gradually, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption are observed\cite{Tkacik:2013}.
+However, with ambitious projects, such as the BRAIN initiative currently underway \cite{Alivisatos:2013} aiming to record from significantly larger populations, the problem must be addressed by regularizing the solution.
+Regularization limits the solution to a smaller subspace to counteract the effect of the sampling noise in the empirical data.
+However, limiting the solution to an inappropriate subspace does not allow significant improvement.
+
+Several strategies have been developed to limit the model space in order to improve the quality of the estimate.
+For example, \cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in retina.
+Generalized linear models typically employ $L_1$ penalty terms to constrain the solution space and to effectively reduce the dimensionality of the solution \cite{Pillow:2008}.
+
+In our study, regularization was also accomplished by dimensionality reduction (feature selection) schemes to produce sparse, constrained solutions.
+
+Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in big neural data will entail the search for optimal regularizations that may involve combinations of heuristic rules and numerical techniques specifically designed for each type of neural circuit.
+
+\paragraph{Model selection}
 In our study, the covariance matrix estimators were evaluated with respect to the cross-validated normal log likelihood.
-This does not limit the applicability of its conclusion to normal distributions.
+However, this does not limit the applicability of its conclusions to normal distributions.
 Indeed the major findings in this paper could be reproduced with respect to other loss functions (compare Fig.~\ref{fig:04} and Fig.~S\ref{supp:02}).
 Other probabilistic models, fitted to the same data, would also serve as estimators of the covariance matrix.
-If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of functional connectivity.
+If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.
+
+The results of model selection must be interpreted with caution.
+As we demonstrated by simulation, a simple model can produce a better estimate even if it does not correctly represent the real dependencies in the data than a more complex model with correct representation of dependencies.
+Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the inference that it reveals a better representation of dependencies in the data.
+This caveat is related to the phenomenon known as \emph{Stein’s Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement.
+
+\paragraph{Physiological interpretation and future directions}
+Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, alludes synaptic interactions, it is unlikely that the anatomical and functional networks are identical.
+The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.
+The inferred interaction network was substantially different from that inferred from by considering the network of most significant sample correlations (\figref{05}{F, H, and I}).
+In particular, sample correlations do not draw attention to the absence of correlation between a pair of neurons that should be strongly correlated based on their common correlations with other neurons.
+The $C_{\sf sparse+latent}$ estimator identified many such pairs and inferred negative interactions between them (\figref{05}{F}).
+The resulting network had a larger fraction of negative interactions than the fraction of negative correlations in the sample correlation matrix.
+The negative interactions had a distinct spatial organization from   Reproducing our study with labeled interneurons of various types could indicate whether the negative interactions
+
+The estimator effectively decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.
+Since the distinction led to improved estimation, the two types of components could be further investigated.
+For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected on the common components/latent units but much less on the pairwise interactions.
+Previous studied found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.
+When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but does not affect the pairwise interactions as strongly.
+The correlation structure also changes as a function of the stimulus condition \cite{Cotton:2013} on time scales that are much faster than changes in the anatomical connectivity.
+Descriptions of specific changes in the functional connectivity due to stimulus properties could shed light on how circuits orchestrate their activity to represent diverse stimuli.
+
+The malleability of the joint structure of activity in populations of neurons makes functional connectivity statistics indispensible wholly apart from the ability to infer the anatomical connectivity.
+Effective statistical descriptions of the functional connectivity will need to be developed to infer the most significant features of the functional organization in large ensembles of neurons from limited data.
+These approaches will need to rely on techniques exemplified by our study: selection among multiple mathematical representations of dependencies, regularization, and model selection.
 

### 2014-01-16 plos.tex
removed illustration in Introduction

@@
 \newcommand{\E}[2][]{\mathbb E_{#1}\left[ #2\right]}    % expected value
-\newcommand{\TODO}[1]{\emph{\small\color{blue}$\langle\langle$#1$\rangle\rangle$}}
-\newcommand*\dif{\mathop{}\,d}
+
 \DeclareMathOperator*{\argmin}{arg\,min}
@@
 Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.
-Correlations in neural activity measured in such recordings can reveal important aspects of the functional organization of neural circuits.
-However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.
+Correlations in neural activity measured in such recordings can reveal important aspects of the organization of neural circuits.
+However, estimating and interpreting large correlation matrices is challenging.
 Estimation can be improved by regularization, \emph{i.e.}\;by imposing a structure on the estimate.
@@
 Therefore, the selection of the most efficient correlation matrix estimator for a given neural circuit must be determined empirically.
-Furthermore, the identity and structure of the most efficient estimator informs about the types of dominant dependencies governing the system.
+Importantly, the identity and structure of the most efficient estimator informs about the types of dominant dependencies governing the system.
 
-We sought the most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.
+We sought statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.
 Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.
-We hypothesized that in these dense recordings, the correlation matrix should be most efficiently represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.
-Indeed, in cross-validation tests, the covariance matrix estimator with such structure consistently outperformed other regularized estimators.
-The connectivity patterns inferred from the sparse component of the estimate manifested a strong relationship to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distance and the difference in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
-Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than sample correlations.
+We hypothesized that in these dense recordings, the correlation matrix should be best represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.
+Indeed, in cross-validation tests, the covariance matrix estimator with such a structure consistently outperformed other regularized estimators.
+The connectivity patterns inferred from the sparse component of the estimate were strongly related to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with  differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
+Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.
 
@@
 % Author Summary not valid for PLoS ONE submissions.
-Correlations in the activity of neural populations have proven useful as descriptions of the functional connectivity in the embedding circuit.
-Correlations may arise from various underlying processes that reflect differently on the structure of the correlation matrix:  Common network fluctuations or external inputs are well approximated by low-rank representations whereas linear effects between specific pairs of neurons are well approximated by fitting their pairwise partial correlations.
-What kinds of correlation effects dominate the population activity in cortical microcircuits?
-To answer this question, we imposed various correlation structures on empirical estimates of neural correlation matrices and evaluated which of them produced greatest estimation improvement.
-In fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the greatest improvement was produced by decomposing the correlation matrix as a sparse network of partial correlations and a low-rank component.
-Since it leads to the best estimates, we propose that this structure provides a better picture of the functional connectivity than that obtained from sample correlations.
-As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and to differences in their preferred orientations and found basic resemblance to results from studies of synaptic connectivity.
+
+It is now possible to record the spiking activity of hundreds of neurons at the same time.
+A meaningful statistical description of the collective activity of these neural populations -- their \sq{functional connectivity} -- is a forefront challenge in neuroscience.
+We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.
+Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.
+In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (\sq{interactions}) combined with a low-rank component.
+The inferred interactions were both positive (\sq{excitatory}) and negative (\sq{inhibitory}) and reflected the spatial organization and orientation preferences of the interacting cells.
+We propose that the composition of the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.
 \section*{Introduction}
-\paragraph{Neural correlations}
-Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.
-For example, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
-In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the functional connectivity in neural circuits \cite{Gerstein:1964}.
-This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.
+\emph{Functional connectivity} is a statistical description of \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells.
+Functional connectivity is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons or in the form of joint coactivation patterns.
+Functional connectivity is thought to reflect local synaptic connections, shared inputs from other regions, and endogenous network activity.
+To remove the confounding effects of the stimulus, the average stimulus responses are subtracted or otherwise accounted for.
+Although functional connectivity is only a phenomenological description, it can serve as a tool to generate hypotheses about how population activity relates to the anatomical architecture of neural circuits and informs about how information is encoded and processed at the population level.
 
-However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.
-Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.
+Pearson correlations between the spiking activity of pairs of neurons, are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.
+In particular, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
+In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the anatomical connectivity \cite{Gerstein:1964}.
+This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.
 
-\paragraph{Structure  of correlation matrices}
-Modern multineuronal recordings allow measuring entire correlation matrices from increasingly large and dense populations of highly interconnected neurons.
-Multivariate analysis techniques based on large correlation matrices provide information that could not be extracted from an equivalent number of isolated measurements of pairwise correlations.
-For example, when the correlation matrix is approximated in a form with a reduced number of variables, this \emph{reduced representation} can isolate an important aspect of the functional connectivity in neural circuits and possibly point to underlying physiological mechanisms and computational principles.
+Neural correlations do not come with ready or unambiguous mechanistic interpretations.
+They can arise from monosynaptic or polysynaptic interactions, common or correlated inputs, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.
+But multineuronal recordings do provide more information than available from an equivalent number of pairs of cells considered separately.
+For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients.
+In contrast, the off-diagonal elements of the precision matrix (inverse of the covariance) constitute scaled partial correlations between neuron pairs.
+The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.
+For this reason, partial correlations have been proposed to more closely reflect the direct effects between components of complex systems and have been used to describe gene interactions \cite{Schafer:2005, Peng:2009} and the interactions between brain regions \cite{Varoquaux:2012, Ryali:2012}.
 
-One reduced representation of the correlation matrix is its low-rank approximation.
-Low-rank approximations are particularly suitable for capturing common fluctuations and common external inputs and can be extracted using principal component analysis, independent component analysis, factor analysis, or  \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.
-
-Another reduced representation of the correlation matrix can be obtained by \emph{covariance selection}, \emph{i.e.}\;by selecting a subset of the pairwise \emph{partial correlations} that are sufficient to reproduce the entire correlation matrix \cite{Dempster:1972, Friedman:2008}.
-The partial correlation between a pair of neurons is a measure of their linear association after the effects of the rest of the network have been taken into account \cite{Anderson:2003}.
-Partial correlations are particularly meaningful when all effects between the variables are nearly linear and a large fraction of the variables are observed; in this case, partial correlations approximate conditional dependency.
-A lack of conditional dependency between two neurons suggests a lack of direct interaction between them.
-Networks of partial correlations (sometimes called \emph{association networks}) have been used to infer gene interaction networks \cite{Schafer:2005, Peng:2009} and interactions between brain regions in brain imaging studies \cite{Varoquaux:2012, Ryali:2012}.
-
-\paragraph{Improving estimation through regularization}
-Imposing a structure on an estimate of the covariance matrix can improve its estimation performance.
-
-Estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
+However, estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
 This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
-
-In neuroscience studies, the true correlation matrix is usually estimated by the \emph{sample correlation matrix}, which is rarely distinguished from the true correlation matrix.
-The sample correlation matrix is unbiased but, because it has many free parameters, it is highly sensitive to sampling noise.
-As a result, on average, the sample correlation matrix is far from the true correlation matrix.
-In contrast, reduced estimates are typically less susceptible to sampling noise although can introduce their respective biases.
+The \emph{sample correlation matrix}, which is an unbiased estimate, is often used to estimate the true correlation matrix (ref).
+However, it has many free parameters, and is thus highly sensitive to sampling noise.
+As a result, on average, the sample correlation matrix is farther from the true correlation matrix than some structured estimates.
 
@@
 Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).
-Yet others can do both target selection and shrinkage toward the target.
+Yet others do both target selection and shrinkage.
 
-Although regularized covariance matrix estimates have become commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.
-\paragraph{Illustration of estimation of a neural correlation matrix}
-To illustrate the challenges and techniques of estimation of the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972, Friedman:2008}.
-Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix (inverse of the covariance matrix) while setting the rest to zero.
-Zeros in the off-diagonal elements of the precision matrix indicate zero partial correlations between the corresponding pair of neurons.
-Therefore, covariance selection amounts to finding an approximation of the covariance matrix that maximizes the number of neuron pairs that are linearly independent of each other when conditioned on the activity of all the other observed cells.
+Although regularized covariance matrix estimation is commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.
+All prior empirical findings from neural correlations have relied on sample correlations.
+Improved estimation is not the only or even the main benefit of regularization.
+When we select the best performing estimator among many regularized estimators, we learn something about the system itself:  The structure imposed by the best estimator represents the underlying dependencies in the system more closely than the structures imposed by the other estimators.
+We can then examine this structure as a more effective description of the functional connectivity in the circuit than the sample correlation matrix.
 
-We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}).
-Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.
-The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.
-
-Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth.
-Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.
-
-\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
-        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
-        The outlined square fragment is magnified in {\bf B}.
-        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
-        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
-    }
-    \label{fig:01}}
-    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
-\end{figure}
-\paragraph{Conceptual questions and basic approach}
+%\paragraph{Conceptual questions and basic approach}
 What structures of correlation matrices best describe the multineuronal activity in specific circuits and in specific brain states?
@@
 
-To address these questions, we evaluated five covariance matrix estimators based on different assumptions about the underlying correlation structure:
+To address these questions, in addition to the sample covariance matrix, we evaluated four regularized covariance matrix estimators that imposed different kinds of structure on the estimate.
+The estimators are designated as follows:
 \begin{description}
@@
 
-First, we demonstrated that, in simulations with known true correlation structures, estimators that imposed a matching structure were generally more efficient than the other estimators.
-The estimators’ relative performance was evaluated by cross-validation.
-The logic of this study relies on the converse of this relationship: the covariance matrix estimator that is found to be most efficient on empirical data is proposed to provide a better picture of the correlation structure than other estimators.
-Caveats for this type of inference are reviewed in the Discussion.
+First, we demonstrated that, in simulations with known correlation structures, estimators that imposed a matching structure were generally more efficient than other estimators.
+We therefore reasoned as follows: The most efficient covariance matrix estimator is likely to provide a better picture of the correlation structure than less efficient estimators (See Discussion).
 
-Therefore, we performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with fast 3D random-access two-photon imaging of calcium signals.
+We therefore performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals.
 We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.
-This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units.
-We examined the structure of the sparse partial correlations inferred by the estimate with respect to the neurons’ geometrical positions and preferred orientations and found that stronger dependencies than those of sample correlations.
-In Discussion, we outline possible approaches to further corroborate the relationship between the structure of the functional connectivity revealed by such methods to the anatomical organization of the circuit.
+This estimator revealed a sparse network of partial correlations between the observed neurons and inferred several latent units with partial correlations with the observed cells.
+We analyzed these networks of partial correlations and found the following: Whereas noise correlations were predominantly positive the inferred partial correlations had a large fraction of negative values possibly reflecting inhibitory circuitry.
+Moreover, we found that the inferred positive partial correlations exhibited stronger dependencies with respect to physical distance and differences in preferred orientations than noise correlations.
+In contrast, the inferred negative partial correlations were less selective.
 
@@
 \paragraph{Covariance estimation}
-The icovariance matrix defined as
+The covariance matrix defined as
 \begin{equation}\label{eq:true-covariance}
@@
 where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates.
-In the case of noise correlations, $\mu$ depends on the stimulus condition.
 Usually, neural covariance matrices have been estimated as the \emph{sample covariance matrix} $C_{\sf 0}$ calculated from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
@@
 
-Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
+Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0})$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
 \begin{equation}\label{eq:precision}
-    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
+    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2},
 \end{equation}
-Similarly, the matrix of partial correlations $P$ is computed by normalizing the the negative \emph{precision matrix} $C^{-1}$ (inverse of the covariance matrix):
+where $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. 
+Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$ (inverse of the covariance matrix):
 \begin{equation}\label{eq:partial}
@@
 \end{equation}
-Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix.
-Clearly, off-diagonal zeros in the precision matrix correspond to zero partial correlations.
 
 We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
-In probabilistic models with exclusively linear dependencies, their targets estimates can be described by distinct families of graphical models (\figref{02}{row 1}).
+In probabilistic models with exclusively linear dependencies, their targets estimates can be described by distinct families of graphical models (\figref{02}{\,Row}{row 1}).
 
@@
     \end{center}
-    \caption{{\bf Estimators whose low-dimensional regularization targets can represent the structure of the true covariance matrix outperform other estimators.}
-        {\bf Row 1.} Graphical representations of four types of low-dimensional structures of interactions between observed neurons (green spheres) and latent units (light-shaded spheres).
-        In the \sq{independent model} ({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units. 
+    \caption{{\bf Estimators whose regularization targets represent the structure of the true covariance matrix outperform other estimators.}
+        {\bf Row 1.} Graphical representations of the regularization targets in the four regularized covariance matrix estimators used in this study,   structures of interactions between observed neurons (green spheres) and latent units (light-shaded spheres).
+        In $C_{\sf diag}$, the regularization target is  \sq{diagonal} ({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units.
         In \sq{latent factors} ({\bf B}), the correlated activity of observed cells is driven by several latent units. 
@@
 
-For estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ whose diagonal contains  variance estimates.
+For estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ whose diagonal elements contain the variance estimates.
 Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
@@
 \end{equation}
-The structure of target $D$ favors the absence of linear associations between the activity of observed neurons (\figref{02}{A}).
+The structure of target $D$ favors the absence of linear associations between the activity of observed neurons (\figref{02}{A Row 1}).
 If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ should outperform the other estimators.
@@
 
-Estimator $C_{\sf sparse}$ is based on the assumption that correlations result from direct linear associations between pairs of observed neurons and that only a fraction of neuron pairs interact (\figref{02}{C}).
+Estimator $C_{\sf sparse}$ favors correlations that arise from direct linear effects between pairs of observed neurons when only a fraction of neuron pairs interact (\figref{02}{C}).
 This assumption is enforced by fitting only a subset of the off-diagonal elements of the precision matrix while setting the rest to zero.
@@
 
-Estimator $C_{\sf sparse+latent}$ allow the assumptions of both $C_{\sf factor}$ and $C_{\sf sparse}$.
-Its correlation structure can be explained by linear effects exerted between some pairs of observed neurons and several latent units (\figref{02}{D}).
-The estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
+Estimator $C_{\sf sparse+latent}$ allows the assumptions of both $C_{\sf factor}$ and $C_{\sf sparse}$.
+Its correlation structure can be explained by linear effects between some pairs of observed neurons and by several latent units (\figref{02}{D}).
+The estimator has the form \cite{Chandrasekaran:2010,Ma:2013} (See Methods):
 \begin{equation}\label{eq:c-sl}
@@
 \end{equation}
-where, as above, $S$ is a sparse matrix and $L$ is a $p\times p$ matrix of rank $d$.
-The estimator has two hyperparameters: one to regulate the number of latent units, the other to regulate the sparsity of matrix $S$.
+where, as above, $S$ is a sparse matrix and $L$ is a $p\times p$ matrix of low rank $d$.
+The estimator has two hyperparameters: one to regulate the rank $d$ and the other to regulate the sparsity of matrix $S$.
 
@@
 An evaluation of the performance of a covariance matrix estimator $C$ requires a real-valued \emph{loss function} $\loss{C,\Sigma}$.
-The loss loss function quantifies the deviation of the estimate $C$ from truth $\Sigma$ and attains its minimum when $C=\Sigma$.
+The loss function quantifies the deviation of the estimate $C$ from truth $\Sigma$ and attains its minimum when $C=\Sigma$.
 
@@
 \end{equation}
-Although the two functions appear similar in form, they differ conceptually: The log-likelihood is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the distance from an arbitrary estimate $C$ to the given truth $\Sigma$.
+Although the two functions appear similar in form, they differ conceptually: The log-likelihood is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the distance from an arbitrary estimate $C$ to the truth $\Sigma$.
 
-The normalization by $\frac 1 p$ is not standard but serves to make the values of the loss functions comparable across different population sizes.
+The normalization by $p$ makes the values of the loss function comparable across different population sizes.  Therefore, the loss function is measured in nats per neuron per observation (or frame).
 
 The choice of loss function is motivated by mathematical convenience.
-We expect that the main conclusions of our study will not change qualitatively with other well behaved loss functions, such as the Frobenius norm of the difference $\|C-\Sigma\|_F$ \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, or quadratic loss \cite{James:1961,Fan:2008}.
+We expect that the main conclusions of our study will not change qualitatively with other well behaved loss functions, such as the Frobenius norm of the difference \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, or quadratic loss \cite{James:1961,Fan:2008}.
 
@@
 
-The negative log-likelihood loss (Eq.~{\ref{eq:loss}) is particularly convenient because it is additive in its second argument in the sense that 
+The normal loss (Eq.~\ref{eq:loss}) is particularly convenient because it is additive in its second argument in the sense that 
  \begin{equation*}\label{eq:additivity}
@@
 
-Then the cross-validation loss is in fact an unbiased estimate of loss.
+Then the cross-validation loss is an unbiased estimate of loss:
 \begin{equation*}
@@
 
-In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (\figref{02}{Row 6}).
+In simulation, CV-losses accurately reproduced the differences between the estimators' excess losses, although with greater variability (\figref{02}{Row 6}).
 For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth.
@@
 We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (\figref{03}{A}) on one side of the visual field of anesthetized mice.
-Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E}) in the visual cortex on the contralateral side to the stimulus.
+Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E} in the visual cortex on the contralateral side to the stimulus.
 Using acousto-optic deflectors (AOD) to steer the laser in 3D, we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}.
@@
 We found that estimator $C_{\sf sparse+latent}$ was the most efficient (Fig.~\ref{fig:04}).
-The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).
+The median relative CV loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).
 
@@
 If this estimate dominates, we can hypothesize that it better describes underlying physiological interactions than other estimators, including the sample covariance.
-In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.
+In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggests common fluctuations such as those caused by common inputs or other collective synchronized activity.
 
@@
     This close approximation is  also demonstrated by the scatter plot of the correlation coefficients produced by the two estimates ({\bf D}). 
-    However, the partial correlation matrices from the two estimate show more pronounced differences ({\bf B} and {\bf E}). 
+    However, the partial correlation matrices from the two estimates show more pronounced differences ({\bf B} and {\bf E}). 
     {\bf C.} Furthermore, the partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.2\% off-diagonal zeros (bottom-left) and low-rank component of rank 15 (top-right).
@@
 \section*{Discussion}
-\paragraph{Functional connectivity}
-Effective phenomenological or statistical descriptions of neural activity can serve as a starting point for investigation of underlying physiological mechanisms.
-For example, the discoveries of receptive fields and tuning properties of single neurons, over the past 60 years, have led to fruitful hypotheses about the underlying patterns of synaptic connectivity and principles of neural coding, and to the discovery of other organizational principles such as various cortical feature maps \cite{Reid:2012}.
-The search for similarly effective descriptions of observed \emph{multineuronal} activity patterns that are not reducible to the response properties of the individual cells has produced various statistical expressions collectively known as \sq{functional connectivity}.
-Functional connectivity is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons and is thought to reflect internal local recurrent connectivity, shared interactions with other regions, or endogenous network activity.
-To remove the confounding effects of stimuli, the average stimulus responses are usually subtracted or otherwise accounted for.
-Although various measures of functional connectivity are only phenomenological descriptions of population activity, they provide experimental neuroscientists with tools to relate population activity to the anatomical and functional architecture of neural circuits and to generate hypotheses about underlying mechanisms.
 
 \paragraph{Functional connectivity as a network of pairwise interactions}
-The intended purpose of many studies of functional connectivity is to estimate the circuit’s anatomical connectivity solely from the observed multineuronal spiking activity.
+Functional connectivity is often expressed as a network of pairwise interactions.
+The intended purpose of many studies of functional connectivity is to estimate the circuitÕs anatomical connectivity solely from the observed multineuronal spiking activity.
 For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections or shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
-Such signatures are still ambiguous as they can arise from diverse network effects other than direct synaptic connections \cite{Aertsen:1989}.
+Such signatures are still ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.
 With simultaneous recordings from a larger subset of neurons in the circuit, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.
@@
 Such conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity.
-For example, generalized linear models (GLMs) have been constructed to include biophysically plausible synaptic integration and membrane kinetics, individual neurons’ stimulus drive, and a network of synaptic couplings \cite{Pillow:2008}.
-Another family of models are maximum entropy models that are constrained by the observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2009, Shlens:2009} .
+For example, generalized linear models (GLMs) can include biophysically plausible synaptic integration and membrane kinetics, individual neuronsÕ stimulus drive, and a network of synaptic couplings \cite{Pillow:2008}.
+Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .
 Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with fewest other assumptions about the distribution: the investigator need only choose which statistics to match.
@@
 
-The inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the interaction between a pair of neurons on the activity of the other neurons under a given probabilistic model, the recording must include all the other neurons with which the pair can potentially interact.
-Otherwise, interactions with the unobserved portions of the circuit will introduce false conditional dependencies between the observed neurons.
-For this reason, the most successful applications of statistical models of population activity thus far have been in in vitro preparations of the retina or cell cultures where high-quality recordings from the complete populations have been possible \cite{Pillow:2008}.
-When inserted in cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the conditional dependencies between pairs of cells.
-Perhaps for this reason, until our study, partial correlations have not been used to describe the functional connectivity between cortical neurons.
+Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the interaction between a pair of neurons on the activity of the other neurons under a given probabilistic model, the recording must include all the other neurons with which the pair can potentially interact.
+Otherwise, interactions with unobserved portions of the circuit can introduce false conditional dependencies between observed neurons.
+For this reason, the most successful applications of statistical models of population activity have been in \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations have been possible \cite{Pillow:2008}.
+In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the conditional dependencies between pairs of cells.
+Perhaps for this reason, partial correlations have not been previously used to describe the functional connectivity between cortical neurons.
 Despite this limitation, many studies have fitted probabilistic models to multielectrode recordings and analyzed the structure of the resulting conditional dependencies \cite{Tang:2008, Yu:2008}.
@@
 The result is often treated as a graphical representation of the functional connectivity without a statement about its anatomical implications.
-Whatever these implications may be, investigators have applied graph-theoretical analyses to the inferred networks to make statements about the design principles and efficiency of circuit architectures \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
+Regardless, investigators have applied graph-theoretical analyses to the inferred networks to make statements about the design principles and efficiency of circuit architectures \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
 
 Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity despite some of the disadvantages.
-The temporal resolution of calcium signals is limited by calcium dye kinetics.
-Fast imaging techniques combined with advanced spike inference algorithms have been demonstrated to provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010, Katona:2012, Cotton:2013}.
+While, the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with advanced spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}.
 However, such high temporal precision comes at the cost of the accuracy of the inferred spike rates: better accuracy can be achieved when calcium signals are analyzed on the temporal scales of tens of milliseconds \cite{Cotton:2013}.
-For what calcium imaging lack in temporal resolution, it make up in spatial resolution and the ability to characterize the spatial arrangements of cells and identify their types.
+The major advantage of imaging is its spatial resolution and the ability to characterize the spatial arrangement and type of recorded cells.
 Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
-Such completeness may open new opportunities for inferring networks of conditional dependencies between neurons with greater validity than with electrophysiological recordings.
+These techniques may open new opportunities for inferring networks of conditional dependencies between neurons with greater validity than with electrophysiological recordings.
 
@@
 
-In our study, we sought to express pairwise connectivity networks among neurons in local cortical microcircuits.
+In our study, we sought to infer pairwise functional connectivity networks among neurons in cortical microcircuits.
 We hypothesized that partial correlations would correspond more closely to underlying mechanisms than sample correlations.
-Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we a priori favored solutions in with sparse partial correlations.
+Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we a priori favored solutions with sparse partial correlations.
 The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution.
 The interactions (conditional dependencies) between pairs of neurons inferred by this distributed are expressed by the precision matrix (inverse of the covariance) or, equivalently, by the matrix of the pairwise partial correlations.
-Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than sample correlations.
+Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
 
@@
 Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the inference that it reveals a better representation of dependencies in the data.
-This caveat is related to the phenomenon known as \emph{Stein’s Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement.
+This caveat is related to the phenomenon known as \emph{SteinÕs Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement.
 
@@
 The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.
-The inferred interaction network was substantially different from that inferred from by considering the network of most significant sample correlations (\figref{05}{F, H, and I}).
+The inferred interaction network was substantially different from that inferred from by considering the network of most significant sample correlations \figref{05}{F, H, and I}.
 In particular, sample correlations do not draw attention to the absence of correlation between a pair of neurons that should be strongly correlated based on their common correlations with other neurons.
-The $C_{\sf sparse+latent}$ estimator identified many such pairs and inferred negative interactions between them (\figref{05}{F}).
+The $C_{\sf sparse+latent}$ estimator identified many such pairs and inferred negative interactions between them \figref{05}{F}.
 The resulting network had a larger fraction of negative interactions than the fraction of negative correlations in the sample correlation matrix.
@@
 For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected on the common components/latent units but much less on the pairwise interactions.
-Previous studied found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.
+Previous studies found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.
 When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but does not affect the pairwise interactions as strongly.
@@
 These approaches will need to rely on techniques exemplified by our study: selection among multiple mathematical representations of dependencies, regularization, and model selection.
+
+We propose that the structure of the most efficient among many estimators provides a better picture of the functional connectivity than that obtained from sample correlations.
 
@@
 where $\tilde Z\succ 0$ denotes the constraint that $\tilde Z$ be a positive definite matrix and $\|\tilde Z\|_1$ is the element-wise $L_1$ norm of the matrix $\tilde Z$.
-This problem formulation is known as \emph{graphical lasso} \cite{Friedman:2008}.
+This problem formulation is known as \emph{graphical lasso} \cite{Meinshausen:2006, Friedman:2008}.
 To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm developed by \cite{Ma:2013}.

### 2014-01-20 plos.tex
rev37. Responded to all comments.

@@
 \newcommand{\E}[2][]{\mathbb E_{#1}\left[ #2\right]}    % expected value
+\newcommand{\ie}{\emph{i.e.}\;}
+\newcommand{\eg}{\emph{e.g.}\;}
 
@@
 However, estimating and interpreting large correlation matrices is challenging.
-Estimation can be improved by regularization, \emph{i.e.}\;by imposing a structure on the estimate.
+Estimation can be improved by regularization, \ie by imposing a structure on the estimate.
 The amount of improvement depends on how closely the assumed structure represents dependencies in the data.
@@
 Pearson correlations between the spiking activity of pairs of neurons, are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.
-In particular, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
+In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
 In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the anatomical connectivity \cite{Gerstein:1964}.
-This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.
+This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.
 
@@
 For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients.
-In contrast, the off-diagonal elements of the precision matrix (inverse of the covariance) constitute scaled partial correlations between neuron pairs.
+In contrast, the off-diagonal elements of the precision matrix (inverse of the covariance matrix) constitute scaled partial correlations between neuron pairs.
 The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.
@@
 However, estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
-This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
-The \emph{sample correlation matrix}, which is an unbiased estimate, is often used to estimate the true correlation matrix (ref).
-However, it has many free parameters, and is thus highly sensitive to sampling noise.
+This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
+The \emph{sample correlation matrix}, which is an unbiased estimate, is often used to estimate the true correlation matrix.
+It has many free parameters making it highly sensitive to sampling noise.
 As a result, on average, the sample correlation matrix is farther from the true correlation matrix than some structured estimates.
@@
 \item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis.
-\item[$C_{\sf sparse}$] --  sparse partial correlations, \emph{i.e.}\;covariance selection.
+\item[$C_{\sf sparse}$] --  sparse partial correlations, \ie covariance selection.
 \item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons combined with partial correlations with several inferred latent units.
@@
 \paragraph{Covariance estimation}
-The covariance matrix defined as
+The covariance matrix is defined as
 \begin{equation}\label{eq:true-covariance}
@@
 \end{equation}
-where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates.
-Usually, neural covariance matrices have been estimated as the \emph{sample covariance matrix} $C_{\sf 0}$ calculated from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
+where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.
+The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf 0}$ from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
 \begin{equation}
@@
 \end{equation}
-where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent).
+where $\nu$ is the number of degrees of freedom in one neuron’s activity within the sample ($\nu=n-1$ if observations are independent).
 In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.
@@
 
-Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0})$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
+Given an arbitrary covariance matrix estimate $C$, the corresponding correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
 \begin{equation}\label{eq:precision}
@@
 where $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. 
-Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$ (inverse of the covariance matrix):
+Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$:
 \begin{equation}\label{eq:partial}
@@
 
-We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
-In probabilistic models with exclusively linear dependencies, their targets estimates can be described by distinct families of graphical models (\figref{02}{\,Row}{row 1}).
+We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
+In probabilistic models with exclusively linear dependencies, the targets estimates of these for estimators can correspond to distinct families of graphical models (\figref{02}{\,Row}{row 1}).
 
@@
     \caption{{\bf Estimators whose regularization targets represent the structure of the true covariance matrix outperform other estimators.}
-        {\bf Row 1.} Graphical representations of the regularization targets in the four regularized covariance matrix estimators used in this study,   structures of interactions between observed neurons (green spheres) and latent units (light-shaded spheres).
-        In $C_{\sf diag}$, the regularization target is  \sq{diagonal} ({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units.
-        In \sq{latent factors} ({\bf B}), the correlated activity of observed cells is driven by several latent units. 
-        In \sq{sparse interactions} ({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
-        In \sq{sparse+latent} ({\bf D}), correlations arise through direct linear interactions in a sparse subset of observed neurons and through interactions with common latent units. 
-        {\bf Row 2.} Examples of $50\times 50$ correlation matrices corresponding to each type of low-dimensional structure. 
-        The factor model ({\bf B}) has three latent units. 
-        The partial correlation matrix of the sparse model ({\bf C}) is 73\% sparse.
-        The \sq{sparse+latent} matrix has one latent unit and its direct interactions are 78\% sparse.
-        {\bf Row 3.} Examples of sample correlation matrices calculated from samples of 1000 observations taken from simulated random processes with corresponding correlation matrices from row 2.
-        {\bf Row 4.} Mean \emph{excess loss} (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators as a function of sample size.
+        {\bf Row 1.} Graphical representations of the target estimates used by the four respective regularized covariance matrix estimators.
+Recorded neurons are denotes as green spheres and latent units as light-shaded spheres.
+The edges between them denote non-zero partial correlations.
+        In estimator $C_{\sf diag}$, the regularization target is a diagonal matrix.
+This target corresponds to a model wherein observed neurons exert no linear effects on one another ({\bf Row 1, A}).
+        In estimator $C_{\sf factor}$, the regularization target is a factor model (low-rank matrix plus a diagonal matrix).
+This target describes a model wherein the activity of observed neurons is linearly driven by several latent units ({\bf  Row 1, B}).
+        In estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix.
+This approximation describes a model wherein correlations arise due to linear effects exerted by the recorded cells onto one another other ({\bf  Row 1, C}).
+        In estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of a matrix constructed as the sum of a sparse and a low-rank component.
+This approximation describes a model wherein correlations arise due to linear affects exerted by the recorded cells onto one another and due to linear effects from several latent units ({\bf  Row 1, D}).
+\\
+{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: the diagonal correlation matrix ({\bf  A}), a factor model with three latent units ({\bf B}),  a correlation matrix with 73\%  off-diagonal zeros ({\bf  C}), and a correlation matrix whose inverse is the some a rank-1 (\ie one latent unit) component and a sparse component with 78\%[1] off-diagonal zeros.
+\\
+{\bf Row 3:} Sample correlation matrices calculated from samples of 1000 observations drawn from simulated random processes with respective correlation matrices from row 2.
+The structure of the correlation matrix is much more difficult to discern by eye.
+\\
+{\bf Row 4:} Regularized correlation matrix estimates computed from the same samples as in Row 3.
+For each kind of ground truth, the estimator with the matching structure was used and the regularization parameters optimized by cross-validation using the data in the same sample.
+The regularized estimates are closer to truth than sample correlation matrices.
+\\
+{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators for  various sample sizes.
 The error bars indicate the standard error of the mean based on 30 samples.
-        {\bf Row 5.} Mean \emph{cross-validation loss} (Eq.~\ref{eq:rel-cv-loss}) of covariance estimators with respect to the matching estimator.
-The values are relative to the validation loss of the estimator that matches the low-dimensional structure of the true covariance matrix.
+Estimators with the matching structure converged to zero faster than the other estimators when the sample size was large enough to allow the more complex true structures outperform simpler structures.
+\\
+{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) attained  by each of the five estimators with respect to the matching estimator.
 Error bars indicate the standard error of the mean based on 30 samples.
@@
 
-For estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ whose diagonal elements contain the variance estimates.
+The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons’ variances.
 Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
@@
 \end{equation}
-The structure of target $D$ favors the absence of linear associations between the activity of observed neurons (\figref{02}{A Row 1}).
-If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ should outperform the other estimators.
+The structure of imposed by $C_{\sf diag}$ favors populations with no linear associations between the neurons (\figref{02}{Row 1, A}).
+If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ is expected to get closer to truth than other estimators.
 
-The target of estimator $C_{\sf factor}$ is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix of rank $d$, and $\Psi$ is the diagonal matrix of independent variances.
+The target of estimator $C_{\sf factor}$ is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix with low rank and $\Psi$ is the diagonal matrix of independent variances.
 The estimator,
@@
 \end{equation}
-has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$.
-The target estimate $L + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).
+has two hyperparameters: the number of factors in $L$ and the shrinkage intensity $\lambda$.
+The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{Row 1, B}).
 
-Estimator $C_{\sf sparse}$ favors correlations that arise from direct linear effects between pairs of observed neurons when only a fraction of neuron pairs interact (\figref{02}{C}).
-This assumption is enforced by fitting only a subset of the off-diagonal elements of the precision matrix while setting the rest to zero.
-The estimator takes the form
+Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by a matrix whose inverse matrix is sparse: 
 \begin{equation}\label{eq:c-sparse}
@@
 \end{equation}
-where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal.  The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.
+where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal.
+The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$.
+The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{02}{Row 1, C}).
+We will sometimes refer to sparse partial interactions between neurons as simply \sq{interactions}.
 
-Estimator $C_{\sf sparse+latent}$ allows the assumptions of both $C_{\sf factor}$ and $C_{\sf sparse}$.
-Its correlation structure can be explained by linear effects between some pairs of observed neurons and by several latent units (\figref{02}{D}).
-The estimator has the form \cite{Chandrasekaran:2010,Ma:2013} (See Methods):
+Estimator $C_{\sf sparse+latent}$ is produced by approximating the sample covariance matrix by a matrix whose inverse is the sum of a sparse component and a low-rank component: 
 \begin{equation}\label{eq:c-sl}
-C_{\sf sparse+latent} = (S - L)^{-1}
+C_{\sf sparse+latent} = (S - L)^{-1},
 \end{equation}
-where, as above, $S$ is a sparse matrix and $L$ is a $p\times p$ matrix of low rank $d$.
-The estimator has two hyperparameters: one to regulate the rank $d$ and the other to regulate the sparsity of matrix $S$.
+where, as above, $S$ is a sparse matrix and $L$ is a low rank matrix.
+The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$.
+The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (\figref{02}{Row 1, D}) \cite{Chandrasekaran:2010,Ma:2013}.
 
 \paragraph{Simulation}
-To illustrate the performance of the four regularized estimators, we constructed four model populations of $p=50$~neurons each with each matching the correlation structures of a different target estimate (\figref{02}{\,Row 2}).
-Sample correlation matrices calculated from samples of size $n=1000$ contain substantial sampling noise (\figref{02}{\,Row 3}) that makes the correlation structure less obvious.
+To demonstrate how the selection between regularized estimators allows discerning the structure of correlation, we constructed four $50\times 50$ covariance matrices with the structures imposed by the four respective regularized estimators (\figref{02}{\,Row 2, A--D}).
+See the Methods section for details of how these matrices were generated.
+We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes.
+Even with relatively large sample sizes (\eg $n=1000$), the sample correlation matrices were contaminated by sampling noise and their underlying structure was difficult to discern (\figref{02}{\,Row 3}).
 
-An evaluation of the performance of a covariance matrix estimator $C$ requires a real-valued \emph{loss function} $\loss{C,\Sigma}$.
-The loss function quantifies the deviation of the estimate $C$ from truth $\Sigma$ and attains its minimum when $C=\Sigma$.
+An evaluation of a covariance matrix estimators requires a \emph{loss function} $\loss{C,\Sigma}$ to quantify the discrepancy between the estimate, $C$, and the truth, $\Sigma$.
+The loss function must attain its minimum when $C=\Sigma$.
 
-In this study, we adopted the \emph{normal loss} defined as
+In this study, we adopted the \emph{normal loss} function defined as
 \begin{equation}\label{eq:loss}
@@
 
-This loss function relates to the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$ through the following expression:
+This loss function derives from the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$ through the following expression:
 \begin{equation}
-    L(\Sigma|C_{\sf 0}) = -\frac p 2 \ln(2\pi) -\frac p 2 \loss{\Sigma,C_{\sf 0}}
+     \loss{\Sigma,C_{\sf 0}} = -\ln(2\pi) - \frac 2 p L(\Sigma|C_{\sf 0})
 \end{equation}
-Although the two functions appear similar in form, they differ conceptually: The log-likelihood is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the distance from an arbitrary estimate $C$ to the truth $\Sigma$.
+The normalization by $p$ makes the values of the loss function comparable across different population sizes.
 
-The normalization by $p$ makes the values of the loss function comparable across different population sizes.  Therefore, the loss function is measured in nats per neuron per observation (or frame).
+Although the two functions appear similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf 0})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.
 
-The choice of loss function is motivated by mathematical convenience.
-We expect that the main conclusions of our study will not change qualitatively with other well behaved loss functions, such as the Frobenius norm of the difference \cite{Ledoit:2004,Schafer:2005}, Stein's entropy loss, or quadratic loss \cite{James:1961,Fan:2008}.
+The choice of normal loss function is motivated by mathematical convenience.
+We expect that the main conclusions of our study will not change qualitatively with other well behaved loss functions, such as Stein's entropy loss or quadratic loss \cite{James:1961, Fan:2008, Ledoit:2004, Schafer:2005}.
 
-We define the \emph{excess loss} as
+The \emph{excess loss} function is produced by shifting the loss function to zero at its minimum:
 \begin{equation}\label{eq:excess-loss}
@@
 \end{equation}
-which assumes zero at its minimum.
 
-We computed the excess losses for 30 samples for each sample size $n=250$, 500, 10000, 2000, and 4000 for each of the five estimators and for each ground truth (\figref{02}{\,row 5}).
-For each estimator, hyperparameters were estimated by nested cross-validation (See Methods.).
-As expected, estimators whose structure matched that of the true model typically outperformed the other estimators.
-There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with the simpler structure, $C_{\sf diag}$ and $C_{\sf factor}$ outperformed the estimators with matching structure (\figref{02}{\,columns C and D, rows 5 and 6}).
-Second, $C_{\sf sparse+latent}$ often performed equally well to the sparse estimator even when the ground truth did not have any latent units.
-In these cases, it inferred the right number of latent units, zero, and estimated optimally.
+We drew 30 independent samples from each model of size $n=250$, 500, 1000, 2000, and 4000 and measured the excess loss produced by each of the five estimators.
+The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.
+All the regularized estimators produced better estimates (lower excess loss) than the sample covariance matrix.
+However, estimators whose structure matched the true model typically outperformed the other estimators (\figref{02}{\,row 5}).
+There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$ often outperformed the estimators with matching structure (\figref{02}{\,Row 5, C and D}).
+Second, $C_{\sf sparse+latent}$ often performed equally well to $C_{\sf sparse}$ on the model without latent units.
+This can be explained by the fact that $C_{\sf sparse+latent}$ correctly inferred the right number of latent units, zero.
 
-When the ground truth, $\Sigma$, is not accessible, loss can be estimated solely from the data through \emph{validation}.
-In validation, an additional independent \emph{testing sample} is used to compute a sample covariance estimate $C_{\sf 0}^\prime$.
-Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can be used as an estimate of loss $\loss{C,\Sigma}$.
-Thus, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.
+Even though, with increasing sample sizes, all estimators converged to the truth (zero excess loss), the distinction between the estimators did not decrease.
+
+In empirical data the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly.
+Instead, the loss function can be estimated solely from the data through \emph{validation}.
+In validation, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf 0}^\prime$ to validate the estimate $C$ computed from the main, \emph{training sample}.
+Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can substitute the loss function $\loss{C,\Sigma}$.
 
@@
 
-Then the cross-validation loss is an unbiased estimate of loss:
+In this case, the cross-validation loss is an unbiased estimate of the loss function:
 \begin{equation*}
@@
 
-The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses.
-With non-additive loss functions, validation loss is a biased estimate of loss, but it is still generally appropriate to be used in lieu of loss.
+The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses; their validation losses are biased estimates of the loss function.
 
-With empirical data, acquiring a separate testing sample is not practical.
-Instead, $K$-fold cross-validation is used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).
+Acquiring a separate testing sample is not practical.
+Instead, $K$-fold cross-validation can be used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).
 Then each subset is used as the validation sample with the other $K-1$ serving as the training sample.
@@
 
-Since we only need to compare estimators to each other, we are only interested in the \emph{relative} CV-loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$:
+Since we only need to compare estimators to each other, we are only interested in the \emph{relative} CV loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$:
 \begin{equation}\label{eq:rel-cv-loss}
@@
 
-In simulation, CV-losses accurately reproduced the differences between the estimators' excess losses, although with greater variability (\figref{02}{Row 6}).
-For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth.
+To examine how well the relative CV-losses reproduced the relationships between the excess losses, we evaluated the relative CV-loss with respect to the estimator that matched the structure of the data-generating model.
+Indeed, CV-losses accurately reproduced the relationships between the excess losses of the estimators, with somewhat greater variability (\figref{02}{Row 6}).
 
-These simulations demonstrated that, with sufficiently large sample sizes, the most efficient among several regularized estimators could be used to infer the most likely type of correlation structure present in the data.
+This simulations study demonstrated that, with sufficiently large sample sizes, imposing the correct type of structure on the estimate, led to greater estimation improvement than imposing incorrect structures
 
 \paragraph{Covariance estimation in neural data}
-We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005, Katona:2012, Cotton:2013}.
-We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (\figref{03}{A}) on one side of the visual field of anesthetized mice.
-Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E} in the visual cortex on the contralateral side to the stimulus.
-Using acousto-optic deflectors (AOD) to steer the laser in 3D, we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}.
+We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{03}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
 This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
-After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}).
-Only cells that produced detectable calcium activity were included in the analysis (See Methods).
-The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{03}{F and G}).
-The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.
-Besides the high-repetition stimulus protocol used to estimate noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cell (\figref{03}{E}).
-Overall, 31 imaged sites from 24 animals were included in the study.
+The firing rates of neurons were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} to produce (\figref{03}{C and D}).
+Only cells that produced detectable calcium activity were included in the analysis (For details, see Methods).
+First, full-field gratings moving in 16 directions were presented to determine the orientation tuning of the recorded cells (\figref{03}{E}).
+To estimate the noise correlation matrix, we presented 300[2] repetitions of full-field drifting gratings with two directions of motion in some experiments or 150[3] repetitions with five directions in other experiments.
+Each grating lasted 1 second and was followed by a 1-second blank.
+The average stimulus response was subtracted from each trial; the traces were then binned into 150 ms intervals for the estimation of the covariance matrix.
+The sample correlation coefficients were largely positive and low (0.01--0.03 in most images sites) (\figref{03}{F and G}).
 
@@
     {\caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
-    Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) were presented to anesthetized mice while two-photon recordings of somatic calcium signals were collected using fast 3D random-access microscopy ({\bf B}).
-The visual stimulus included an initial period with 16 directions of motion for orientation tuning followed by a longer (15--20 min) period of stimulation with only 2 or 5 directions of motion for the computation of the noise correlation matrix.
+    Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) presented during two-photon recordings of somatic calcium signals using fast 3D random-access microscopy ({\bf B}).
+    {\bf C--G.} Calcium activity data from an example site.
     {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. The inferred firing rate binned in 150 ms intervals are indicated by red ticks below each trace.
@@
 
-In these highly localized populations both direct interactions between cells and common diffuse inputs likely contribute to overall population co-variability.
-We therefore hypothesized that regularized estimates capturing a sparse partial correlation structure, and/or interactions with latent units would be the most efficient.
-At the same time, average correlations were relatively small (\figref{03}{G}), giving a potential advantage to estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$).
-Finally, correlation structure could possibly be best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf sparse+latent}$).
-Thus it was \emph{a priori} unclear which estimator would perform best.
+In these dense local populations direct interactions between cells are likely to contribute substantially to the patterns of population.
+We therefore hypothesized that covariance matrix estimates that explicitly model the partial correlation between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$ would have a performance advantage over those that do not.
+However, the observed neurons must also be strongly influenced by global activity fluctuations and by unobserved common inputs, giving an advantage to estimators that explicitly model common fluctuations of the entire population $C_{\sf factor}$ and $C_{\sf sparse+latent}$.
+If both types of effects are significant, then $C_{\sf sparse+latent}$ should outperform the other estimators.
 
-We found that estimator $C_{\sf sparse+latent}$ was the most efficient (Fig.~\ref{fig:04}).
-The median relative CV loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).
-
+To test this hypothesis, we estimated the relative CV loss (Eq.~\ref{eq:rel-cv-loss}) of estimators  $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$.
+The hyperparameters of each estimator were optimized by nested cross-validation (See Methods for details).
+Indeed, the median relative CV loss was significantly above zero (Fig.~\ref{fig:04}), indicating that $C_{\sf sparse+latent}$ produced estimates that were consistently closest to the true covariance matrix.
 \begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
@@
 
+Furthermore, $C_{\sf sparse}$ outperformed all other estimators except $C_{\sf sparse+latent}$ (Fig.~S\ref{supp:01}) suggesting the greater importance of modeling interactions between pairs of neurons than the common fluctuations of activity across the entire population.
+
+Evaluation with respect to the quadratic loss function,
+\begin{equation}\label{eq:quadratic}
+\loss{C,C_{\sf 0}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf 0}^\prime-I)^2,
+\end{equation}
+produced the same relationship between the estimators as the normal loss (Fig.~S\ref{supp:02}) suggesting that the results of this study do not depend on the assumption of gaussianity and reflect real regularities in the data.
+
 \paragraph{Correlation structure and circuit architecture}
-Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites.
-By design, the sparse+latent estimator finds the mixture of sparse partial correlations between visible neurons and common fluctuations described as latent units.
-If this estimate dominates, we can hypothesize that it better describes underlying physiological interactions than other estimators, including the sample covariance.
-In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggests common fluctuations such as those caused by common inputs or other collective synchronized activity.
+Because of its superior estimation performance, $C_{\sf sparse+latent}$ likely provides a more physiologically relevant representation of the functional connectivity than the other estimators.
+The estimator segregates two sources of correlations: linear dependencies between pairs of neurons (sparse component) suggestive of direct interactions and common fluctuations across the entire population (low-rank) component suggestive of common inputs, cell assembly activation, network fluctuations or other forms of synchronized activations.
 
-We examined the structure of the sparse+latent estimate at individual sites (an example site is depicted in Fig.\;\ref{fig:03}).
-At these sites the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (\figref{05}{A and D}).
-However, the corresponding partial correlation matrices differed dramatically (\figref{05}{B and E}).
-The partial correlation was decomposed into sparse and low-rank components (\figref{05}{C}).
-Although correlations were mostly positive, the sparse partial correlations (or \sq{interactions}), had a much larger fraction of negative values than sample correlations.
-The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}).
+We examined the structure of the $C_{\sf sparse+latent}$ estimate in each imaged site.
+The regularized estimate of the correlation matrix appeared very similar to the sample correlation matrix (\figref{05}{A and D}) but the corresponding partial correlation matrices differed substantially.
+The partial correlation matrix (\figref{05}{B and E})  and a low-rank component (\figref{05}{C}).
+While sample correlations were mostly positive, the sparse partial correlations had a high fraction of negative interactions.
+In the example site (Fig.~\ref{fig:03} and Fig.~\ref{fig:05})),  the sparse component had 82.1\% [4]sparsity (or, conversely, 17.9\% [5]connectivity) with average node degree of 52.5[6] (\figref{05}{G}).
 The low-rank component had rank 17.
@@
     However, the partial correlation matrices from the two estimates show more pronounced differences ({\bf B} and {\bf E}). 
-    {\bf C.} Furthermore, the partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.2\% off-diagonal zeros (bottom-left) and low-rank component of rank 15 (top-right).
+    {\bf C.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.1\%[7] off-diagonal zeros (bottom-left) and low-rank component of rank 17[8] (top-right).
     {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations.
-The gray interval indicates the range of correlations containing 82.2\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
-This interval contained 58.9\% of the partial correlations.
-    {\bf G.} The graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
+The gray interval indicates the range of correlations containing 82.1\%[9] of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
+This interval contained 58.9\%[10] of the partial correlations.
+    {\bf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
 The line density is proportional to the magnitude of the correlation.
@@
 
-In previous studies of the structure of correlations, a sparse network was produced by thresholding the correlations coefficients at a level deemed significant.
-A network was formed by connecting neurons whose total correlations exceed this threshold \cite{Golshani:2009,Malmersjo:2013}.
-There was fairly little overlap between the network of interactions revealed by such thresholding and those revealed by the sparse+latent estimator (\figref{05}{F}).
-When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (\figref{05}{F, H, and I}).
-In particular, many low sample correlations translated into negative interactions in the regularized estimate.
-Indeed, the absence of correlations between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient.
-Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.
+We compared the graphs of partial correlations revealed by $C_{\sf sparse+latent}$ to the graph obtained by thresholding the sample correlations.
+Thresholded correlations have been used in several studies as to isolate the likely pairwise interactions with subsequent graph theoretical analysis \cite{ Malmersjo:2013}.
+We set the threshold to match the connectivity of the $C_{\sf sparse+latent}$ graph.
+The networks revealed by the two methods differed substantially.
+In the example site with 17.9\%[11] connectivity, only 42\%[12] of cell pairs connected in the $C_{\sf sparse+latent}$ graph were also connected in the thresholded correlation graph (\figref{05}{F, H, and I}).
+In particular, the thresholded sample correlations were predominantly positive whereas 38\%[13] of the interactions in $C_{sparse+latent}$ were negative.
+Most of the negative interactions corresponded to low sample correlations (\figref{05}{F}).
+Thus the $C_{\sf sparse+latent}$ estimates reveal negative interactions that account for low correlations where high correlations are predicted from all the other interactions.
 
-The average partial correlations revealed by the sparse+latent estimator at all 31 sites were about 5 times lower than the sample correlations and  less variable across sites (\figref{06}{A}).
+The average partial correlations revealed by $C_{\sf sparse+latent}$ estimates in all 31 sites were about 5 times lower than the sample correlations and less variable across sites (\figref{06}{A}): the coefficient of variation of the average sample correlations across all sites was 0.41 whereas that of the average partial correlations was only 0.35.
 The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (\figref{06}{B and C}).
@@
 
-We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}).
-The average partial correlations fell more rapidly with difference in preferred orientation (\figref{07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).
+We also examined the relationship of the correlation structure to the differences in orientation preference and to the geometric distances between the cell pairs (Fig.\;\ref{fig:07}).
+The partial correlations of the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation (\figref{07}{A}).
+Similarly, the regularized partial correlations decreased more rapidly than sample correlations with geometric distance laterally  (\figref{07}{B}) and in depth (\figref{07}{B}).
 
@@
 
-The connectivity expressed by the partial correlations in the regularized estimate had different organization for positive and negative interactions.
-Positive interactions fell rapidly as a function of difference in preferred orientation (\figref{07}{D}), lateral displacements (\figref{07}{E}), and displacement in depth (\figref{07}{F}).
-Negative interactions were much less selective (\figref{07}{D--F}).
+Furthermore, the positive and negative interactions in the $C_{\sf sparse+latent}$ estimator were organized differently: The positive interactions fell rapidly with differences in the preferred orientation (\figref{07}{D}), lateral displacements (\figref{07}{E}), and displacement in depth (\figref{07}{F}) while the negative interactions were much less selective (\figref{07}{D--F}).
 
 \section*{Discussion}
-
 \paragraph{Functional connectivity as a network of pairwise interactions}
-Functional connectivity is often expressed as a network of pairwise interactions.
-The intended purpose of many studies of functional connectivity is to estimate the circuitÕs anatomical connectivity solely from the observed multineuronal spiking activity.
+Functional connectivity is often expressed as a graph of pairwise interactions.
+The intended purpose of many studies of functional connectivity is to estimate the circuit’s anatomical connectivity solely from the observed multineuronal spiking activity.
 For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections or shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
@@
 With simultaneous recordings from a larger subset of neurons in the circuit, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.
-Since direct causal interactions between neurons produce statistical dependency even after conditioning on the state of the rest of the network, conditionally independent neurons can be inferred to lack a direct causal influence.
+Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the rest of the network.
+Therefore, conditionally independent neurons can be inferred to lack a direct causal influence.
 Such conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity.
-For example, generalized linear models (GLMs) can include biophysically plausible synaptic integration and membrane kinetics, individual neuronsÕ stimulus drive, and a network of synaptic couplings \cite{Pillow:2008}.
+For example, generalized linear models (GLMs) can include biophysically plausible synaptic integration and membrane kinetics, individual neurons’ stimulus drive, and a network of synaptic couplings \cite{Pillow:2008}.
 Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .
-Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with fewest other assumptions about the distribution: the investigator need only choose which statistics to match.
+Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with minimum other assumptions about the distribution: the investigator need only choose which statistics to match.
 Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
-Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \emph{i.e.}\;conditional dependencies between pairs of neurons.
+Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of neurons.
 
@@
 
-Because the validity of the inference of interactions from probabilistic models depends on both the correctness of their mathematical form in the model and the completeness of the recorded population, it is far from clear which approach provides best correspondence to the anatomical connectivity in the circuit, with no experimental evidence yet available to address this question.
-The result is often treated as a graphical representation of the functional connectivity without a statement about its anatomical implications.
-Regardless, investigators have applied graph-theoretical analyses to the inferred networks to make statements about the design principles and efficiency of circuit architectures \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
+Correct inference of causal interactions between neurons depends on correct mathematical representations of the interactions in the statistical model and the completeness of the recorded population.
+It is not yet clear which approach provides best correspondence to the anatomical connectivity in the circuit, with little experimental evidence yet available to address this question.
+The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without a statement about its anatomical implications.
+Topological properties of such graphs have been interpreted to indicate principles of circuit organization such as small-world organization \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
+However, the topological properties of graphs depend critically on the method of their inference (Zalesky:2012) and, until the physiological meaning of the functional connectivity is confirmed, such analyses may be questionable.
 
-Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity despite some of the disadvantages.
-While, the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with advanced spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}.
+Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.
+While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with advanced spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}.
 However, such high temporal precision comes at the cost of the accuracy of the inferred spike rates: better accuracy can be achieved when calcium signals are analyzed on the temporal scales of tens of milliseconds \cite{Cotton:2013}.
-The major advantage of imaging is its spatial resolution and the ability to characterize the spatial arrangement and type of recorded cells.
+The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.
 Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
-These techniques may open new opportunities for inferring networks of conditional dependencies between neurons with greater validity than with electrophysiological recordings.
+These techniques may open new opportunities for more incisive of functional connectivity than with electrophysiological recordings.
 
@@
 The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution.
-The interactions (conditional dependencies) between pairs of neurons inferred by this distributed are expressed by the precision matrix (inverse of the covariance) or, equivalently, by the matrix of the pairwise partial correlations.
+The interactions (conditional dependencies) between pairs of neurons inferred by this distributed are expressed by the precision matrix or, equivalently, by the matrix of the pairwise partial correlations.
 Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
@@
 Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}.
-Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \emph{i.e.}\;groups of tightly interconnected groups of cells involved in a common computation.
+Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \ie groups of tightly interconnected groups of cells involved in a common computation.
 Coactivation patterns could also be explained as shared input from unobserved parts of the circuit.
@@
 Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the inference that it reveals a better representation of dependencies in the data.
-This caveat is related to the phenomenon known as \emph{SteinÕs Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement.
+This caveat is related to the phenomenon known as \emph{Stein’s Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
 
 \paragraph{Physiological interpretation and future directions}
-Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, alludes synaptic interactions, it is unlikely that the anatomical and functional networks are identical.
+Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.
 The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.
 The inferred interaction network was substantially different from that inferred from by considering the network of most significant sample correlations \figref{05}{F, H, and I}.
-In particular, sample correlations do not draw attention to the absence of correlation between a pair of neurons that should be strongly correlated based on their common correlations with other neurons.
-The $C_{\sf sparse+latent}$ estimator identified many such pairs and inferred negative interactions between them \figref{05}{F}.
-The resulting network had a larger fraction of negative interactions than the fraction of negative correlations in the sample correlation matrix.
-The negative interactions had a distinct spatial organization from   Reproducing our study with labeled interneurons of various types could indicate whether the negative interactions
+The $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that are not present in the sample correlation matrix (\figref{05}{F}).
+These negative interactions are produced as explanations for low sample correlation values where high values are expected based on common correlations with other neurons.
+The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.
+The resulting network had a larger fraction of negative interactions than in the sample correlation matrix.
+The negative interactions had a distinct spatial and functional organization from the positive interactions.
+To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.
 
@@
 
-The malleability of the joint structure of activity in populations of neurons makes functional connectivity statistics indispensible wholly apart from the ability to infer the anatomical connectivity.
-Effective statistical descriptions of the functional connectivity will need to be developed to infer the most significant features of the functional organization in large ensembles of neurons from limited data.
-These approaches will need to rely on techniques exemplified by our study: selection among multiple mathematical representations of dependencies, regularization, and model selection.
-
-We propose that the structure of the most efficient among many estimators provides a better picture of the functional connectivity than that obtained from sample correlations.
+We propose that the structure of the most efficient among many estimators provides a better picture of the functional connectivity than that obtained from unbiased sample correlations.
 
@@
 
-For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 12 and 20 mins.
-Only traces whose average deconvolved signal was greater than 1\% of the median of the recorded population were included in the analysis.
+For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 12 and 20 minutes[14].
+Only spikes trains whose variance was greater than 1\%[15] of the median across all the traces in the recording were included in the analysis.
 \paragraph{Cross-validation}
@@
 \begin{equation}
-(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L} + \|\tilde Z\|_1
+(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L, C_{\sf 0}} + \|\tilde Z\|_1
 \end{equation}
@@
 \begin{equation}
-(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L }
+(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L, C_{\sf 0} }
 \end{equation}
@@
 \setcounter{figure}{0}
+\renewcommand{\figurename}{Figure S}
 
@@
 \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-{\caption{{\bf The sparse+latent estimator outperforms the other estimators with respect to quadratic cross-validation loss $\loss{C,C_{\sf 0}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf 0}^\prime-I)^2$.}
+{\caption{{\bf The sparse+latent estimator outperforms the other estimators under the quadratic loss (Eq.~\ref{eq:quadratic}) just as it did under the normal loss (Eq.~\ref{eq:loss}) used in Fig.~\ref{fig:04}}.
 }
@@
 \end{document}
+
+[1]verify
+[2]verify
+[3]verify
+[4]Verify
+[5]Verify
+[6]Verify
+[7]Verify
+[8]Verify
+[9]Verify
+[10]Verify
+[11]Verify
+[12]Verify
+[13]Verify
+[14]Verify
+[15]Verify

### 2014-01-20 plos.tex
rev39 -- verified all numbers

@@
 \\
-{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: the diagonal correlation matrix ({\bf  A}), a factor model with three latent units ({\bf B}),  a correlation matrix with 73\%  off-diagonal zeros ({\bf  C}), and a correlation matrix whose inverse is the some a rank-1 (\ie one latent unit) component and a sparse component with 78\%[1] off-diagonal zeros.
+{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: the diagonal correlation matrix ({\bf  A}), a factor model with two latent units ({\bf B}),  a correlation matrix with 85\%  off-diagonal zeros ({\bf  C}), and a correlation matrix whose inverse is the some a rank-1 (\ie one latent unit) component and a sparse component with 89\% off-diagonal zeros.
 \\
@@
 \paragraph{Simulation}
-To demonstrate how the selection between regularized estimators allows discerning the structure of correlation, we constructed four $50\times 50$ covariance matrices with the structures imposed by the four respective regularized estimators (\figref{02}{\,Row 2, A--D}).
+To demonstrate how the selection between regularized estimators allows discerning the structure of correlations, we constructed four $50\times 50$ covariance matrices with the structures imposed by the four respective regularized estimators (\figref{02}{\,Row 2, A--D}).
 See the Methods section for details of how these matrices were generated.
@@
 
-This loss function derives from the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$ through the following expression:
+This loss function derives from the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$ as:
 \begin{equation}
-     \loss{\Sigma,C_{\sf 0}} = -\ln(2\pi) - \frac 2 p L(\Sigma|C_{\sf 0})
+     \loss{X,Y} \equiv -\ln(2\pi) - \frac 2 p L(X|Y)
 \end{equation}
@@
 First, full-field gratings moving in 16 directions were presented to determine the orientation tuning of the recorded cells (\figref{03}{E}).
-To estimate the noise correlation matrix, we presented 300[2] repetitions of full-field drifting gratings with two directions of motion in some experiments or 150[3] repetitions with five directions in other experiments.
+To estimate the noise correlation matrix, we presented between 100 and 300 repetitions of full-field drifting gratings at each direction of motion.
 Each grating lasted 1 second and was followed by a 1-second blank.
@@
     {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
-    The histograms are based on 31 imaged sites in 24 mice. 
+    The histograms are based on 31 imaged sites in 15 mice. 
     All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
@@
 While sample correlations were mostly positive, the sparse partial correlations had a high fraction of negative interactions.
-In the example site (Fig.~\ref{fig:03} and Fig.~\ref{fig:05})),  the sparse component had 82.1\% [4]sparsity (or, conversely, 17.9\% [5]connectivity) with average node degree of 52.5[6] (\figref{05}{G}).
-The low-rank component had rank 17.
+In the example site (Fig.~\ref{fig:03} and Fig.~\ref{fig:05})),  the sparse component had 82.2\% sparsity (or, conversely, 17.8\% connectivity) with average node degree of 52.8 (\figref{05}{G}).
+The estimate also inferred 15 latent units contributing to the low-rank component of the partial correlation matrix.
 
@@
     However, the partial correlation matrices from the two estimates show more pronounced differences ({\bf B} and {\bf E}). 
-    {\bf C.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.1\%[7] off-diagonal zeros (bottom-left) and low-rank component of rank 17[8] (top-right).
+    {\bf C.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.2\% off-diagonal zeros (bottom-left) and low-rank component of rank 15 (top-right).
     {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations.
-The gray interval indicates the range of correlations containing 82.1\%[9] of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
-This interval contained 58.9\%[10] of the partial correlations.
+The gray interval indicates the range of correlations containing 82.2\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
+This interval contained 58.3\% of the partial correlations.
     {\bf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
@@
 The networks revealed by the two methods differed substantially.
-In the example site with 17.9\%[11] connectivity, only 42\%[12] of cell pairs connected in the $C_{\sf sparse+latent}$ graph were also connected in the thresholded correlation graph (\figref{05}{F, H, and I}).
-In particular, the thresholded sample correlations were predominantly positive whereas 38\%[13] of the interactions in $C_{sparse+latent}$ were negative.
+In the example site with 17.8\% connectivity, only 41.7\% of cell pairs connected in the $C_{\sf sparse+latent}$ graph were also connected in the thresholded correlation graph (\figref{05}{F, H, and I}).
+In particular, the thresholded sample correlations were predominantly positive whereas 43.3\% of the interactions in $C_{\sf sparse+latent}$ were negative.
 Most of the negative interactions corresponded to low sample correlations (\figref{05}{F}).
@@
 The collected fluorescent traces were deconvolved to reconstruct the firing rates for each neuron.
-First, the first principal component was subtracted from the traces, which reduced common mode noise related to small movement and cardiovascular artifacts.
+First, the first principal component was subtracted from the traces, which reduced common mode noise related to small movement and cardiovascular artifacts \cite{Cotton:2013}.
 The resulting traces were low-pass filtered below 0.1 Hz and downsampled to 20 Hz.
@@
 
-For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 12 and 20 minutes[14].
-Only spikes trains whose variance was greater than 1\%[15] of the median across all the traces in the recording were included in the analysis.
+For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 14 and 27 minutes (mean 22 minutes).
+Only spikes trains whose variance was greater than 1\% of the median of the variances of all the traces in each recording were included in the analysis.
 \paragraph{Cross-validation}
@@
 \end{document}
-
-[1]verify
-[2]verify
-[3]verify
-[4]Verify
-[5]Verify
-[6]Verify
-[7]Verify
-[8]Verify
-[9]Verify
-[10]Verify
-[11]Verify
-[12]Verify
-[13]Verify
-[14]Verify
-[15]Verify

### 2014-01-21 plos.tex
changed title

@@
 {\Large
-Partial neural correlations suggest detailed interactions in cortical circuits
+Improved estimation and interpretation of correlations in neural circuits
 }
@@
 % Please keep the abstract between 250 and 300 words
-Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.
+Ambitious projects aim to record the activity of ever larger and denser neuronal populations \emph{in vivo}.
 Correlations in neural activity measured in such recordings can reveal important aspects of the organization of neural circuits.
@@
 We hypothesized that in these dense recordings, the correlation matrix should be best represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.
-Indeed, in cross-validation tests, the covariance matrix estimator with such a structure consistently outperformed other regularized estimators.
+Indeed, in cross-validation tests, the covariance matrix estimator with this structure consistently outperformed other regularized estimators.
 The connectivity patterns inferred from the sparse component of the estimate were strongly related to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with  differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
@@
 To remove the confounding effects of the stimulus, the average stimulus responses are subtracted or otherwise accounted for.
-Although functional connectivity is only a phenomenological description, it can serve as a tool to generate hypotheses about how population activity relates to the anatomical architecture of neural circuits and informs about how information is encoded and processed at the population level.
+Functional connectivity is only a phenomenological description with no simple correspondence to anatomical connectivity, it can be used to generate or test hypotheses about the anatomical architecture of the neural circuit and how information is encoded and processed at the population level.
 
-Pearson correlations between the spiking activity of pairs of neurons, are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.
+Pearson correlations between the spiking activity of pairs of neurons are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.
 In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
@@
 \item[$C_{\sf sparse}$] --  sparse partial correlations, \ie covariance selection.
-\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons combined with partial correlations with several inferred latent units.
+\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons combined with partial correlations to several inferred latent units.
 \end{description}
@@
 \end{equation}
-where $\nu$ is the number of degrees of freedom in one neuron’s activity within the sample ($\nu=n-1$ if observations are independent).
+where $\nu$ is the number of degrees of freedom in one neuron's activity within the sample ($\nu=n-1$ if observations are independent).
 In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.
@@
 
-Although the two functions appear similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf 0})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.
+Although the two functions appear similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf 0})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.
 
@@
 
-In empirical data the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly.
+In empirical data, the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly.
 Instead, the loss function can be estimated solely from the data through \emph{validation}.
@@
     \end{center}
-    \caption{{\bf Dependence of correlations and partial correlations on orientation tuning differences and physical distance between cell pairs.}
+    \caption{{\bf Dependence of sample correlations and regularized partial correlations on orientation tuning and physical distance.}
     {\bf A--C.} Average partial correlations (red) estimated by $C_{\sf sparse+latent}$ and average sample correlations (black) averaged across multiple imaged sites.
@@
 Directional tuning was map using a pseudo-random sequence of drifting gratings at sixteen equally spaced directional of motion changing at 2 Hz for 3 min without blanks.
-The data for covariance estimation were collected during presentations of full-field drifting gratings with the same parameters as those used in directional tuning except only two directions (in 9 dataset) or five directions (in 22 datasets) were used and the presentations lasted 1 second and, separated by 1-second blanks.
+The data for covariance estimation were collected during presentations of full-field drifting gratings with the same parameters as those used in directional tuning except only two directions (in 9 datasets) or five directions (in 22 datasets) were used and the presentations lasted 1 second and, separated by 1-second blanks.
 Each stimulus condition was presented at least 180 times.
@@
 \begin{equation}
-(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L, C_{\sf 0}} + \|\tilde Z\|_1
+(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L, C_{\sf 0}} + \lambda\|\tilde Z\|_1
 \end{equation}
@@
 
-The partial correlation matrix computed from the overall estimate, computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}, includes the effects of interactions between the visible and latent units.
-This estimate of the partial correlations was used in analyses where sparsity was not required (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).
-For analyses that required a sparse network (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}) was computed using the extended precision matrix $S^\ast$:
+The partial correlation matrix with both sparse and low-rank components is computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}; it includes the effects of interactions between the visible and latent units.
+This estimate of the partial correlations was used in analyses of average partial correlations (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).
+The sparse component was used separately in analyses of connectivity graphs (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component:
 \begin{equation}

### 2014-01-22 plos.tex
corrections

@@
 \emph{Functional connectivity} is a statistical description of \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells.
-Functional connectivity is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons or in the form of joint coactivation patterns.
+Functional connectivity is expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons, or in the form of multineuronal coactivation patterns, or by fitting a probabilistic model of the population activity.
 Functional connectivity is thought to reflect local synaptic connections, shared inputs from other regions, and endogenous network activity.
-To remove the confounding effects of the stimulus, the average stimulus responses are subtracted or otherwise accounted for.
-Functional connectivity is only a phenomenological description with no simple correspondence to anatomical connectivity, it can be used to generate or test hypotheses about the anatomical architecture of the neural circuit and how information is encoded and processed at the population level.
+Although functional connectivity is only a phenomenological description with no simple mechanistic interpretation, it can be used to generate or test hypotheses about the anatomical architecture of the neural circuit and the processing of information at the population level.
 
@@
 In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
-In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the anatomical connectivity \cite{Gerstein:1964}.
-This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.
+In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of the synaptic connectivity \cite{Gerstein:1964}.
+The interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.
 
@@
 For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients.
-In contrast, the off-diagonal elements of the precision matrix (inverse of the covariance matrix) constitute scaled partial correlations between neuron pairs.
+On the other hand, a correlation matrix can be specified through the  partial correlations between pairs of the recorded neurons.
 The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.
-For this reason, partial correlations have been proposed to more closely reflect the direct effects between components of complex systems and have been used to describe gene interactions \cite{Schafer:2005, Peng:2009} and the interactions between brain regions \cite{Varoquaux:2012, Ryali:2012}.
+For this reason, partial correlations have been proposed to more closely reflect the direct causal effects between components of complex systems than correlations; partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
 
-However, estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
+However, estimation of covariance matrices from large populations presents a number of numerical challenges.
+The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
 This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
-The \emph{sample correlation matrix}, which is an unbiased estimate, is often used to estimate the true correlation matrix.
-It has many free parameters making it highly sensitive to sampling noise.
+All prior empirical findings from neural correlations have used \emph{sample correlations}.
+The \emph{sample correlation matrix} is an unbiased estimate but its many free parameters make it sensitive to the sampling noise.
 As a result, on average, the sample correlation matrix is farther from the true correlation matrix than some structured estimates.
 
-In statistics, the technique of deliberately imposing a structure on an estimate in order to improve its performance is known as \emph{regularization} \cite{Schafer:2005, Bickel:2006}.
-To \sq{impose structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation known as the shrinkage target.
-The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data either analytically or by cross-validation.
-Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely.
-Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).
-Yet others do both target selection and shrinkage.
+Estimation can be improved through \emph{regularization},  the technique of deliberately imposing a structure on an estimate in order to reduce its estimation error \cite{Schafer:2005, Bickel:2006}.
+To \sq{impose a structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation with fewer free parameters, which is often referred to as the \emph{target estimate}.
+The selection of the optimal target estimate and the optimal amount of shrinkage is deduced from the data sample either analytically \cite{Ledoit:2003, Ledoit:2004, Schafer:2005}  or by cross-validation \cite{Friedman:1989}.
+An estimator that produces estimates that are, on average, closer to the truth for a given sample size is said to be more \emph{efficient} than other estimators.
 
 Although regularized covariance matrix estimation is commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.
-All prior empirical findings from neural correlations have relied on sample correlations.
-Improved estimation is not the only or even the main benefit of regularization.
-When we select the best performing estimator among many regularized estimators, we learn something about the system itself:  The structure imposed by the best estimator represents the underlying dependencies in the system more closely than the structures imposed by the other estimators.
-We can then examine this structure as a more effective description of the functional connectivity in the circuit than the sample correlation matrix.
+
+Reduced estimation error is not the only benefit of regularization.
+By finding the most efficient estimator for a specific system among many regularized estimators, we learn something about the system itself: Since imposing this structure leads best estimation, the elements of the structure may provide a better representation of the underlying dependencies in the data than previous analyses of sample correlations.
 
@@
 
-To address these questions, in addition to the sample covariance matrix, we evaluated four regularized covariance matrix estimators that imposed different kinds of structure on the estimate.
-The estimators are designated as follows:
+To address these questions, we evaluated four regularized covariance matrix estimators that imposed different structures on the estimate. The estimators are designated as follows:
 \begin{description}
 \item[$C_{\sf 0}$] -- sample covariance matrix, the unbiased, unregularized estimator.
-\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix. 
-\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis.
-\item[$C_{\sf sparse}$] --  sparse partial correlations, \ie covariance selection.
-\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons combined with partial correlations to several inferred latent units.
+\item[$C_{\sf diag}$] -- linear shrinkage of correlations toward zero.
+\item[$C_{\sf factor}$] -- linear shrinkage of the covariances toward a low-rank matrix, representing inputs from unobserved factors.
+\item[$C_{\sf sparse}$] -- sparse partial correlations, \ie a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
+\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons \emph{and} linear interactions with a number of latent units.
 \end{description}
 
-First, we demonstrated that, in simulations with known correlation structures, estimators that imposed a matching structure were generally more efficient than other estimators.
-We therefore reasoned as follows: The most efficient covariance matrix estimator is likely to provide a better picture of the correlation structure than less efficient estimators (See Discussion).
+First, we used simulated data to demonstrate that the selection of the optimal estimator indeed pointed to the true structure of the dependencies in the data.
 
@@
 We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.
-This estimator revealed a sparse network of partial correlations between the observed neurons and inferred several latent units with partial correlations with the observed cells.
-We analyzed these networks of partial correlations and found the following: Whereas noise correlations were predominantly positive the inferred partial correlations had a large fraction of negative values possibly reflecting inhibitory circuitry.
-Moreover, we found that the inferred positive partial correlations exhibited stronger dependencies with respect to physical distance and differences in preferred orientations than noise correlations.
+This estimator revealed a sparse network of partial correlations between the observed neurons and inferred several latent units exerting linear effects on the observed neurons.
+We analyzed these networks of partial correlations and found the following: Whereas noise correlations were predominantly positive, the inferred partial correlations had a large fraction of negative values possibly reflecting inhibitory circuitry.
+Moreover, we found that the inferred positive partial correlations exhibited stronger dependencies with respect to physical distances and differences in preferred orientations than noise correlations.
 In contrast, the inferred negative partial correlations were less selective.
@@
 where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.
-The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf 0}$ from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
+The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf 0}$ from the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
 \begin{equation}
@@
 where $\nu$ is the number of degrees of freedom in one neuron's activity within the sample ($\nu=n-1$ if observations are independent).
-In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.
-In the case of noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.
+For noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.
+In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than the sample correlation matrix, $C_{\sf 0}$.
 
@@
 We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
-In probabilistic models with exclusively linear dependencies, the targets estimates of these for estimators can correspond to distinct families of graphical models (\figref{02}{\,Row}{row 1}).
+In probabilistic models with exclusively linear dependencies, the targets estimates of these for estimators correspond to distinct families of graphical models (\figref{02}{\,Row 1}}).
 
@@
     \end{center}
-    \caption{{\bf Estimators whose regularization targets represent the structure of the true covariance matrix outperform other estimators.}
-        {\bf Row 1.} Graphical representations of the target estimates used by the four respective regularized covariance matrix estimators.
+    \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
+        {\bf Row 1.} Graphical representations of the target estimates used of the four respective regularized covariance matrix estimators.
 Recorded neurons are denotes as green spheres and latent units as light-shaded spheres.
 The edges between them denote non-zero partial correlations.
-        In estimator $C_{\sf diag}$, the regularization target is a diagonal matrix.
-This target corresponds to a model wherein observed neurons exert no linear effects on one another ({\bf Row 1, A}).
-        In estimator $C_{\sf factor}$, the regularization target is a factor model (low-rank matrix plus a diagonal matrix).
-This target describes a model wherein the activity of observed neurons is linearly driven by several latent units ({\bf  Row 1, B}).
+        {\bf Row 1, A}.  In estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear associations. 
+        {\bf  Row 1, B.} In estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), favoring systems in which correlations arise due to common input from latent units.
+        {\bf  Row 1, C}.
 In estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix.
-This approximation describes a model wherein correlations arise due to linear effects exerted by the recorded cells onto one another other ({\bf  Row 1, C}).
-        In estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of a matrix constructed as the sum of a sparse and a low-rank component.
-This approximation describes a model wherein correlations arise due to linear affects exerted by the recorded cells onto one another and due to linear effects from several latent units ({\bf  Row 1, D}).
+This approximation describes systems in which correlations arise from linear associations between the observed units.
+        {\bf  Row 1, D}.
+In estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix.
+This approximation describes a model wherein correlations arise due to linear associations between the recorded cells \emph{and} with several latent units.
 \\
-{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: the diagonal correlation matrix ({\bf  A}), a factor model with two latent units ({\bf B}),  a correlation matrix with 85\%  off-diagonal zeros ({\bf  C}), and a correlation matrix whose inverse is the some a rank-1 (\ie one latent unit) component and a sparse component with 89\% off-diagonal zeros.
+{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros.
 \\
-{\bf Row 3:} Sample correlation matrices calculated from samples of 1000 observations drawn from simulated random processes with respective correlation matrices from row 2.
-The structure of the correlation matrix is much more difficult to discern by eye.
+{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices from row 2.
+The structure of the sample correlation matrix is difficult to discern by eye.
 \\
-{\bf Row 4:} Regularized correlation matrix estimates computed from the same samples as in Row 3.
-For each kind of ground truth, the estimator with the matching structure was used and the regularization parameters optimized by cross-validation using the data in the same sample.
-The regularized estimates are closer to truth than sample correlation matrices.
+{\bf Row 4:} Correlation matrix estimates computed by estimators with matching structure from the same data as the sample correlation matrices in Row 3.
+The regularized estimates are closer to the truth than the sample correlation matrices.
 \\
-{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators for  various sample sizes.
+{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) attained by the five estimators as a function of sample size.
 The error bars indicate the standard error of the mean based on 30 samples.
-Estimators with the matching structure converged to zero faster than the other estimators when the sample size was large enough to allow the more complex true structures outperform simpler structures.
+Estimators with the matching structure converged to zero faster than the other estimators.
+Some exceptions occur when the sample is not large enough to reveal the true structure; in these cases estimators with simpler structures can be more efficient.
 \\
-{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) attained  by each of the five estimators with respect to the matching estimator.
+{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) attained the five estimators with respect to the matching estimator.
 Error bars indicate the standard error of the mean based on 30 samples.
@@
 
-The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons’ variances.
+The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons' variances.
 Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
@@
 The structure of imposed by $C_{\sf diag}$ favors populations with no linear associations between the neurons (\figref{02}{Row 1, A}).
-If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ is expected to get closer to truth than other estimators.
+If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.
 
@@
 \end{equation}
-has two hyperparameters: the number of factors in $L$ and the shrinkage intensity $\lambda$.
+has two hyperparameters: the rank of $L$ (\ie the number of latent factors) and the shrinkage intensity $\lambda$.
 The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{Row 1, B}).
@@
 \end{equation}
-where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal.
+where $S$ is a sparse matrix, \ie one in which a large fraction of off-diagonal elements are set to zero.
 The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$.

### 2014-01-22 plos.tex
fixed a buncha things in the manuscript

@@
 {\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices from row 2.
-The structure of the sample correlation matrix is difficult to discern by eye.
+The structure of the sample correlation matrix is difficult to discern.
 \\
@@
 We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes.
-Even with relatively large sample sizes (\eg $n=1000$), the sample correlation matrices were contaminated by sampling noise and their underlying structure was difficult to discern (\figref{02}{\,Row 3}).
+Even with relatively large sample sizes (\eg $n=1000$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (\figref{02}{\,Row 3}).
 
-An evaluation of a covariance matrix estimators requires a \emph{loss function} $\loss{C,\Sigma}$ to quantify the discrepancy between the estimate, $C$, and the truth, $\Sigma$.
+The evaluation of a covariance matrix estimator $C$ is performed with respect to a \emph{loss function} $\loss{C,\Sigma}$ to quantify the discrepancy between the estimate and the truth, $\Sigma$.
 The loss function must attain its minimum when $C=\Sigma$.
@@
 
-We drew 30 independent samples from each model of size $n=250$, 500, 1000, 2000, and 4000 and measured the excess loss produced by each of the five estimators.
+We drew 30 independent samples of size $n=250$, 500, 1000, 2000, and 4000 from model and computed the excess loss resulting from the five estimators.
 The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.
@@
 However, estimators whose structure matched the true model typically outperformed the other estimators (\figref{02}{\,row 5}).
-There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$ often outperformed the estimators with matching structure (\figref{02}{\,Row 5, C and D}).
-Second, $C_{\sf sparse+latent}$ often performed equally well to $C_{\sf sparse}$ on the model without latent units.
-This can be explained by the fact that $C_{\sf sparse+latent}$ correctly inferred the right number of latent units, zero.
+There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$, often outperformed the estimators with the matching structure (\figref{02}{\,Row 5, C and D}).
+Second, $C_{\sf sparse+latent}$ often performed equally well to $C_{\sf sparse}$ on the data from models without latent units because it reliably inferred the right number of latent units, zero.
 
@@
 Instead, the loss function can be estimated solely from the data through \emph{validation}.
-In validation, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf 0}^\prime$ to validate the estimate $C$ computed from the main, \emph{training sample}.
+In validation, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf 0}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.
 Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can substitute the loss function $\loss{C,\Sigma}$.
@@
 
-In this case, the cross-validation loss is an unbiased estimate of the loss function:
+In this case, the validation loss is an unbiased estimate of the loss function:
 \begin{equation*}
@@
 Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ trained on the remaining $K-1$ subsets.
-Then cross-validation loss for estimator $C$ is
+Then the cross-validation loss for estimator $C$ is
 \begin{equation}\label{eq:cv-loss}
@@
 \end{equation}
+The relative cross-valiation loss was used for the main findings in this study.
 
-To examine how well the relative CV-losses reproduced the relationships between the excess losses, we evaluated the relative CV-loss with respect to the estimator that matched the structure of the data-generating model.
+To examine how well the relative CV-losses reproduced the relationships between the excess losses, we evaluated the relative CV-loss with respect to the estimator that matched the true structure of the model model.
 Indeed, CV-losses accurately reproduced the relationships between the excess losses of the estimators, with somewhat greater variability (\figref{02}{Row 6}).
 
-This simulations study demonstrated that, with sufficiently large sample sizes, imposing the correct type of structure on the estimate, led to greater estimation improvement than imposing incorrect structures
+This simulation study demonstrated that, with sufficiently large sample sizes, imposing the correct type of structure on the estimate led to greater estimation improvement than imposing other structures.
 
 \paragraph{Covariance estimation in neural data}
-We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{03}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
+We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{03}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
 This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
-The firing rates of neurons were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} to produce (\figref{03}{C and D}).
+The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} to produce (\figref{03}{C and D}).
 Only cells that produced detectable calcium activity were included in the analysis (For details, see Methods).
-First, full-field gratings moving in 16 directions were presented to determine the orientation tuning of the recorded cells (\figref{03}{E}).
-To estimate the noise correlation matrix, we presented between 100 and 300 repetitions of full-field drifting gratings at each direction of motion.
-Each grating lasted 1 second and was followed by a 1-second blank.
-The average stimulus response was subtracted from each trial; the traces were then binned into 150 ms intervals for the estimation of the covariance matrix.
+First, 30 repetitions of full-field drifting gratings of 16 direction were presented in random order for 500 ms without blanks.
+This stimulus was used to compute the orientation tuning of the recorded cells (\figref{03}{E}).
+To estimate the noise correlation matrix, we presented only 2 or 5 distinct directions with between 100 and 300 repetitions of each direction.
+In this stimulus, each grating lasted 1 second and was followed by a 1-second blank.
+The average stimulus response was subtracted from each trial.
+The traces were then binned into 150 ms intervals for the estimation of the covariance matrix.
 The sample correlation coefficients were largely positive and low (0.01--0.03 in most images sites) (\figref{03}{F and G}).
@@
     {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site.
+The cells' colors indicate their orienation preferences.
+The gray cells were not significantly tuned to orientation.
     {\bf F.} The noise correlation matrix of the activity of the neural population. 
@@
 
-In these dense local populations direct interactions between cells are likely to contribute substantially to the patterns of population.
-We therefore hypothesized that covariance matrix estimates that explicitly model the partial correlation between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$ would have a performance advantage over those that do not.
-However, the observed neurons must also be strongly influenced by global activity fluctuations and by unobserved common inputs, giving an advantage to estimators that explicitly model common fluctuations of the entire population $C_{\sf factor}$ and $C_{\sf sparse+latent}$.
+In these dense populations, direct interactions between cells are likely to influence the patterns of population activity.
+We therefore hypothesized that covariance matrix estimates that explicitly model the partial correlation between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would have a performance advantage.
+However, the observed neurons must also be strongly influenced by global activity fluctuations and by unobserved common inputs, to the advantage of estimators that explicitly model common fluctuations of the entire population: $C_{\sf factor}$ and $C_{\sf sparse+latent}$.
 If both types of effects are significant, then $C_{\sf sparse+latent}$ should outperform the other estimators.
 
-To test this hypothesis, we estimated the relative CV loss (Eq.~\ref{eq:rel-cv-loss}) of estimators  $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$.
+To test this hypothesis, we computed the relative CV losses (Eq.~\ref{eq:rel-cv-loss}) of estimators  $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in 31 imaged sites in 15 mice.
 The hyperparameters of each estimator were optimized by nested cross-validation (See Methods for details).
@@
 \begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-    {\caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
-    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
+    {\caption{{\bf Estimator $C_{\sf sparse+latent}$ is more efficient than the other estimators on neural data.}
+    {\bf A--D.} Histograms of the relative cross-validation loss of estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$.
     The histograms are based on 31 imaged sites in 15 mice. 
@@
 
-Evaluation with respect to the quadratic loss function,
+The same evaluation based on the quadratic loss function,
 \begin{equation}\label{eq:quadratic}
@@
 \end{equation}
-produced the same relationship between the estimators as the normal loss (Fig.~S\ref{supp:02}) suggesting that the results of this study do not depend on the assumption of gaussianity and reflect real regularities in the data.
+instead of the normal loss function reproduced the same relationship between the estimators (Fig.~S\ref{supp:02}).
+This suggests that the results of this study are robust to the choice of the loss function and do not depend on the assumption of gaussianity.
 
@@
 Because of its superior estimation performance, $C_{\sf sparse+latent}$ likely provides a more physiologically relevant representation of the functional connectivity than the other estimators.
-The estimator segregates two sources of correlations: linear dependencies between pairs of neurons (sparse component) suggestive of direct interactions and common fluctuations across the entire population (low-rank) component suggestive of common inputs, cell assembly activation, network fluctuations or other forms of synchronized activations.
+The estimator segregates two sources of correlations: linear dependencies between pairs of neurons (sparse component) suggestive of direct interactions and common fluctuations across the entire population (low-rank component) component suggestive of common inputs, cell assembly activation, network fluctuations or other forms of synchronized activations.
 
-We examined the structure of the $C_{\sf sparse+latent}$ estimate in each imaged site.
-The regularized estimate of the correlation matrix appeared very similar to the sample correlation matrix (\figref{05}{A and D}) but the corresponding partial correlation matrices differed substantially.
-The partial correlation matrix (\figref{05}{B and E})  and a low-rank component (\figref{05}{C}).
+We examined the composition of the $C_{\sf sparse+latent}$ estimates in each imaged site.
+The regularized estimates appeared similar to the sample correlation matrix (\figref{05}{A and D}) but the corresponding partial correlation matrices differed substantially (\figref{05}{B and E}).
+The partial correlation matrix of the regularized estimate was decomposed into a sparse component and a low-rank component (\figref{05}{C}).
 While sample correlations were mostly positive, the sparse partial correlations had a high fraction of negative interactions.
-In the example site (Fig.~\ref{fig:03} and Fig.~\ref{fig:05})),  the sparse component had 82.2\% sparsity (or, conversely, 17.8\% connectivity) with average node degree of 52.8 (\figref{05}{G}).
+In the example site (Fig.~\ref{fig:03} and Fig.~\ref{fig:05}),  the sparse component had 82.2\% sparsity (or, conversely, 17.8\% connectivity) with average node degree of 52.8 (\figref{05}{G}).
 The estimate also inferred 15 latent units contributing to the low-rank component of the partial correlation matrix.
@@
     \end{center}
-    \caption{{\bf Example of low-dimensional correlation structure revealed by the sparse+low-rank estimator.}
-    {\bf A.} The regularized estimate of the correlation matrix (top-right) closely approximates the sample correlation matrix (bottom left). 
-    This close approximation is  also demonstrated by the scatter plot of the correlation coefficients produced by the two estimates ({\bf D}). 
-    However, the partial correlation matrices from the two estimates show more pronounced differences ({\bf B} and {\bf E}). 
+    \caption{{\bf Example of the structure revealed by $C_{\sf sparse+latent}$ the sparse+low-rank estimator.}
+    {\bf A and D.} The regularized estimate $C_{\sf sparse+latent}$ closely approximates the sample correlation matrix $C_{\sf 0}$. 
+    {\bf B and E.} However, the partial correlation matrices from the two estimates differ substantially.
     {\bf C.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.2\% off-diagonal zeros (bottom-left) and low-rank component of rank 15 (top-right).
-    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations.
-The gray interval indicates the range of correlations containing 82.2\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
-This interval contained 58.3\% of the partial correlations.
-    {\bf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
+    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations: the gray interval indicates the range of correlations containing 82.2\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
+The significant correlations were outside this interval.
+Yet 58.3\% of the interactions inferred by $C_{\sf sparse+latent}$  linked pairs of neurons whose correlation was below the threshold.
+Only the remaning 41.7\% of the interactions overlapped with sample correlations above the theshold.
+    {ebf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
 The line density is proportional to the magnitude of the correlation.
@@
 We compared the graphs of partial correlations revealed by $C_{\sf sparse+latent}$ to the graph obtained by thresholding the sample correlations.
-Thresholded correlations have been used in several studies as to isolate the likely pairwise interactions with subsequent graph theoretical analysis \cite{ Malmersjo:2013}.
+Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013}.
 We set the threshold to match the connectivity of the $C_{\sf sparse+latent}$ graph.
 The networks revealed by the two methods differed substantially.
-In the example site with 17.8\% connectivity, only 41.7\% of cell pairs connected in the $C_{\sf sparse+latent}$ graph were also connected in the thresholded correlation graph (\figref{05}{F, H, and I}).
+In the example site with 17.8\% connectivity, only 41.7\% of cell pairs connected in the $C_{\sf sparse+latent}$ graph were also connected in the graph of thresholded sample correlations (\figref{05}{F, H, and I}).
 In particular, the thresholded sample correlations were predominantly positive whereas 43.3\% of the interactions in $C_{\sf sparse+latent}$ were negative.
@@
 Several sites, even with relatively large population sizes, had fairly few pairwise interactions and were dominated by latent units.
-These differences have multiple explanations, including differences in brain states and recording quality (neuropil contamination, motion).
+These differences may be explained by differences in brain states, recording quality (neuropil contamination, motion).
 
@@
 We also examined the relationship of the correlation structure to the differences in orientation preference and to the geometric distances between the cell pairs (Fig.\;\ref{fig:07}).
-The partial correlations of the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation (\figref{07}{A}).
-Similarly, the regularized partial correlations decreased more rapidly than sample correlations with geometric distance laterally  (\figref{07}{B}) and in depth (\figref{07}{B}).
+The partial correlations of the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation (\figref{07}{A}) than sample correlations.
+They also decreased more rapidly with geometric distance, both laterally  (\figref{07}{B}) and in depth (\figref{07}{B}).
 
@@
 
-Furthermore, the positive and negative interactions in the $C_{\sf sparse+latent}$ estimator were organized differently: The positive interactions fell rapidly with differences in the preferred orientation (\figref{07}{D}), lateral displacements (\figref{07}{E}), and displacement in depth (\figref{07}{F}) while the negative interactions were much less selective (\figref{07}{D--F}).
+Furthermore, the positive and negative interactions in the $C_{\sf sparse+latent}$ estimator were organized differently: The probability of positive interactions fell rapidly with differences in the preferred orientation (\figref{07}{D}), lateral displacements (\figref{07}{E}), and displacement in depth (\figref{07}{F}) while the negative interactions were much less selective (\figref{07}{D--F}).
 
@@
 Functional connectivity is often expressed as a graph of pairwise interactions.
-The intended purpose of many studies of functional connectivity is to estimate the circuit’s anatomical connectivity solely from the observed multineuronal spiking activity.
+The intended purpose of many studies of functional connectivity is to estimate the circuit's anatomical connectivity solely from the observed multineuronal spiking activity.
 For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections or shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
@@
 Such conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity.
-For example, generalized linear models (GLMs) can include biophysically plausible synaptic integration and membrane kinetics, individual neurons’ stimulus drive, and a network of synaptic couplings \cite{Pillow:2008}.
+For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, individual neurons' stimulus drive, and a network of synaptic couplings \cite{Pillow:2008}.
 Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .
-Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with minimum other assumptions about the distribution: the investigator need only choose which statistics to match.
+Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with minimal other assumptions about the distribution: the investigator need only choose which statistics to match.
 Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
@@
 The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without a statement about its anatomical implications.
-Topological properties of such graphs have been interpreted to indicate principles of circuit organization such as small-world organization \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
-However, the topological properties of graphs depend critically on the method of their inference (Zalesky:2012) and, until the physiological meaning of the functional connectivity is confirmed, such analyses may be questionable.
+Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
+However, the topological properties of the functional connectivity graphs vary with the method of their inference (Zalesky:2012); until the physiological meaning of the functional connectivity is established, the validity of such analyses remains in question.
 
@@
 Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
-These techniques may open new opportunities for more incisive of functional connectivity than with electrophysiological recordings.
+These techniques may open new opportunities for more incisive measurements of functional connectivity than with electrophysiological recordings.
 
 The low temporal resolution of calcium signals limits the use of functional connectivity methods that rely on millisecond-scale binning of signals (cross-correlograms, some GLMs, and maximum entropy models on the binary domain).
-Most studies of functional connectivity have relied on instantaneous sample correlations \cite{Greenberg:2008, Golshani:2009, Hofer:2011, Malmersjo:2013} .
+Hence, most studies of functional connectivity have relied on instantaneous sample correlations \cite{Greenberg:2008, Golshani:2009, Hofer:2011, Malmersjo:2013} .
 Although some investigators have interpreted such correlations as indicators of (chemical or electrical) synaptic connectivity, most used them as more general indicators of functional connectivity without a statement about underlying mechanisms.
@@
 In our study, we sought to infer pairwise functional connectivity networks among neurons in cortical microcircuits.
-We hypothesized that partial correlations would correspond more closely to underlying mechanisms than sample correlations.
+We hypothesized that partial correlations would correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.
 Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we a priori favored solutions with sparse partial correlations.
 The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution.
-The interactions (conditional dependencies) between pairs of neurons inferred by this distributed are expressed by the precision matrix or, equivalently, by the matrix of the pairwise partial correlations.
 Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
@@
 The problem becomes increasingly worse as the size of the recorded population of neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
-As the recorded neuronal population sizes increased gradually, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption are observed\cite{Tkacik:2013}.
-However, with ambitious projects, such as the BRAIN initiative currently underway \cite{Alivisatos:2013} aiming to record from significantly larger populations, the problem must be addressed by regularizing the solution.
+As the recorded neuronal population sizes increased gradually, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.
+However, with ambitious projects such as the BRAIN initiative currently underway \cite{Alivisatos:2013}, aiming to record from significantly larger populations, the problem must be addressed by regularizing the solution.
 Regularization limits the solution to a smaller subspace to counteract the effect of the sampling noise in the empirical data.
-However, limiting the solution to an inappropriate subspace does not allow significant improvement.
+However, limiting the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.
 
@@
 The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.
-The inferred interaction network was substantially different from that inferred from by considering the network of most significant sample correlations \figref{05}{F, H, and I}.
+The inferred graph of interactions was substantially different from the graph of the most significant sample correlations \figref{05}{F, H, and I}.
 The $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that are not present in the sample correlation matrix (\figref{05}{F}).
@@
 The estimator effectively decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.
-Since the distinction led to improved estimation, the two types of components could be further investigated.
+Since the distinction led to improved estimation, the two types of components could be further analyzed and related to physiological processes.
 For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected on the common components/latent units but much less on the pairwise interactions.
 Previous studies found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.
-When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but does not affect the pairwise interactions as strongly.
+When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but has little effect on the pairwise interactions.
 The correlation structure also changes as a function of the stimulus condition \cite{Cotton:2013} on time scales that are much faster than changes in the anatomical connectivity.
@@
 % authors may use "Analysis" 
-\paragraph{Surgery and two-photon Imaging}
+\paragraph{Surgery and two-photon imaging}
 All procedures were conducted in accordance with the ethical guidelines of the National Institutes of Health and were approved by the Baylor College of Medicine IACUC.
@@
 Briefly, C57BL/6J mice (aged p40--60) were used.
-Anesthesia was initiated with isoflurane (3\%) and the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
-A craniotomy was performed over the right primary visual area.
+Anesthesia was initiated with isoflurane (3\%) and maintained with the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
+A craniotomy was performed over the right primary visual cortex.
 Membrane-permeant calcium indicator Oregon Green 488 BAPTA-1 AM (OGB-1, Invitrogen) was loaded by bolus injection.
@@
 Calcium imaging began 1 hour after dye injection.
-All imaging was performed using the 3D-RAMP two-photon microscope described in \cite{Cotton:2013}.
-First, A 3D stack was acquired and cells manually segmented.
-To collect calcium signals, the system repeated hopped between the selected neurons.
+All imaging was performed using the 3D-RAMP two-photon microscope as described in \cite{Cotton:2013}.
+First, a 3D stack was acquired and cells manually segmented.
+Then calcium signal were collected by sampling in the center of each cell at rates of 100 Hz or higher, depending on the number of cells.
+
 \paragraph{Visual stimulus}
@@
 The collected fluorescent traces were deconvolved to reconstruct the firing rates for each neuron.
-First, the first principal component was subtracted from the traces, which reduced common mode noise related to small movement and cardiovascular artifacts \cite{Cotton:2013}.
+First, the first principal component was subtracted from the traces, which reduced common mode noise related to small cardiovascular movements \cite{Cotton:2013}.
 The resulting traces were low-pass filtered below 0.1 Hz and downsampled to 20 Hz.
@@
 
-Orientation tuning was computed by fitting the mean firing rates in response to gratings of directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
+Orientation tuning was computed by fitting the mean firing rates in response to gratings of directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
 The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times.
-The $p$-value of the fit was computed as the fraction of permuted datasets for which the $R^2$ value of the tuning function fit exceeded that of the real data.
-Cells were considered tuned for $p$-values not exceeding $0.05$.
+The $p$-value of the fit was computed as the fraction of the permuted datasets for which the $R^2$ value of the tuning function fit exceeded that of the real data.
+Cells were considered tuned for $p$-values below 0.05.
 
 For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 14 and 27 minutes (mean 22 minutes).
-Only spikes trains whose variance was greater than 1\% of the median of the variances of all the traces in each recording were included in the analysis.
+Spike trains that did not have substantial spiking activity (those whose variance was less than 2\% of the $90^{th}$ percentile in each site), presumably from glial cells or inactive neurons, were exculded from the analysis.
+
 \paragraph{Cross-validation}
-To compare the performance of the estimator against each other, we used conventional 10-fold cross validation to measure cross-validation loss (Eq.~\ref{eq:cv-loss}).
+To compare the performance of the estimator against each other, we used conventional 10-fold cross validation.
 Briefly, each recording was split into 30 blocks of equal duration.
 The 30 blocks were then grouped randomly into 10 datasets with 3 blocks in each.
-This procedure ensured sufficient independence between the 10 datasets while still ensuring that each dataset included data from different parts of the recording.
+This splitting of data was chosen to achieve sufficient independence between the 10 datasets for cross-validaton purposes.
+Splitting the data into smaller blocks would increase their dependence due to temporal effects.
 Then, each dataset was used as the testing dataset with the rest of the data used for estimating the covariance matrix.
@@
 When cross-validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.
-This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{02}{\;rows 3 and 4}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:05}--\ref{fig:07}).
+This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{02}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:05}--\ref{fig:07}).
 \paragraph{Covariance Estimation}
 Within the inner loop of cross-validation, covariance matrix estimation was performed with fixed hyperparameter values provided by the search algorithm.
-The computation of each regularized estimator only required the sample covariance matrix $C_{\sf 0}$ of the training dataset.
+The computation of regularized estimators only required the sample covariance matrix $C_{\sf 0}$ of the training dataset.
 
@@
 
-The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean squared error whereas we optimized them with respect to the loss function in Eq.~\ref{eq:loss}.
+The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean squared error whereas we optimized them with respect to the normal loss function (Eq.~\ref{eq:loss}).
 
 Estimator $C_{\sf factor}$ used two hyperparameters: the number of latent factors $d$ and the shrinkage intensity $\lambda \in [0, 1]$.  
-The $p\times d$ factor loading matrix $L$ and individual variances $\Psi$ were computed by solving the minimization problem
+The factor matrix $L$ of rank $d$ and the diagonal matrix of individual variances $\Psi$ were computed by solving the minimization problem
 \begin{equation}
@@
 
-Estimator $C_{\sf sparse}$  has one hyperparameter $\lambda$, which regulates the sparsity of its precision matrix $S$.
-The precision matrix is computed in two steps: First, the zero structure $Z$ is determined by minimizing the $L_1$-penalized loss.
+Estimator $C_{\sf sparse}$  has one hyperparameter $\lambda$ to regulate the sparsity of its precision matrix $S$.
+The precision matrix is computed in two steps: First, the zero structure is determined by minimizing the $L_1$-penalized loss:
 \begin{equation}
@@
 This problem formulation is known as \emph{graphical lasso} \cite{Meinshausen:2006, Friedman:2008}.
-To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm developed by \cite{Ma:2013}.
+To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm \cite{Ma:2013}.
 Then, after the zero structure was determined, the remaining coefficients were fitted without penalty:
@@
 Estimator $C_{\sf sparse+latent}$ has two hyperparameters: the number of latent units $d$ and the sparsity level $\lambda$.
-It estimates the larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.
+$C_{\sf sparse+latent}$ estimates a larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.
 \begin{equation}
@@
 \begin{equation}
-C_{\sf sparse+latent} = (S^\ast)^{-1} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
+C_{\sf sparse+latent} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
 \end{equation}
-The matrix $ S_{12}S_{22}^{-1}S_{12}^\T$ has rank $d$.
-Rather than searching for the optimal sparse structure of $S_{12}$ and $S_{22}$, an ill-posed problem, we estimated these components together as one positive-definite $p\times p$ matrix $L$ of rank $d$.
+The $p\times p$  matrix $L=S_{12}S_{22}^{-1}S_{12}^\T$ has rank $d$.
+Rather than searching for the optimal sparse structure of $S_{12}$ and $S_{22}$, an ill-posed problem, we estimated these components together as the low-rank matrix $L$.
 
@@
 \end{equation}
-Then we find the sparse and low-rank components of the precision matrix by minimizing the loss 
+Then we find the sparse and low-rank components of the precision matrix by solving (also using ADMM) 
 \begin{equation}
@@
 This estimate of the partial correlations was used in analyses of average partial correlations (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).
-The sparse component was used separately in analyses of connectivity graphs (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component:
+The sparse component $S$ was used separately in analyses of the connectivity graphs (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component alone without the low-rank component:
 \begin{equation}
@@
 For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units.
-The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{02}{\,row 2}.
-Samples were then generated by multivariate normal distributions with the respective true covariance matrices and estimated by each of the estimators.
+The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{02}{\,Row 2}.
+Samples were then generated by multivariate normal distributions with the respective true covariance matrices to be estimated by each of the estimators.
 

### 2014-01-24 plos.tex
Rev 41. Incorporated Andreas' comments.

--- 2014-01-22 plos.tex
@@
 Correlations in neural activity measured in such recordings can reveal important aspects of the organization of neural circuits.
-However, estimating and interpreting large correlation matrices is challenging.
+However, estimating and interpreting large correlation matrices is statistically challenging.
 Estimation can be improved by regularization, \ie by imposing a structure on the estimate.
@@
 We propose that the composition of the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.
+
 \section*{Introduction}
-\emph{Functional connectivity} is a statistical description of \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells.
-Functional connectivity is expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons, or in the form of multineuronal coactivation patterns, or by fitting a probabilistic model of the population activity.
+\emph{Functional connectivity} is a statistical description of observed \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells.
 Functional connectivity is thought to reflect local synaptic connections, shared inputs from other regions, and endogenous network activity.
@@
 For this reason, partial correlations have been proposed to more closely reflect the direct causal effects between components of complex systems than correlations; partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
+These opportunities have not yet been explored in neurophysiological studies: most studies have only consider the distribution of pairwise correlations \cite{Zohary:1994, Bair:2001, Smith:2008, Ecker:2010}.
 
@@
 This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
-All prior empirical findings from neural correlations have used \emph{sample correlations}.
-The \emph{sample correlation matrix} is an unbiased estimate but its many free parameters make it sensitive to the sampling noise.
+The \emph{sample correlation matrix} is an unbiased estimate of the true correlations but its many free parameters make it sensitive to the sampling noise.
 As a result, on average, the sample correlation matrix is farther from the true correlation matrix than some structured estimates.
@@
 
-Reduced estimation error is not the only benefit of regularization.
-By finding the most efficient estimator for a specific system among many regularized estimators, we learn something about the system itself: Since imposing this structure leads best estimation, the elements of the structure may provide a better representation of the underlying dependencies in the data than previous analyses of sample correlations.
+Improved estimation of the correlation matrix is beneficial in itself.
+For example, improved estimates can be used to optimize  decoding of the population activity.
+But reduced estimation error is not the only benefit of regularization.
+The amount of improvement of estimation is greatest when the structure imposed by regularization limits the solution space but still represents the dominant dependencies in the data.
+Therefore, by finding the most efficient among many regularized estimators, we learn something about the system itself: the imposed structure of the most efficient estimator is a parsimonious representation of the regularities in the data.
+The elements of this structure likely provide a better representation of the underlying dependencies in the data than previous analyses of sample correlations.
+
+The necessity and benefits of regularization increase with the size of the recorded population.
+With the advent of the age of big neural data, the search for optimal regularization schemes will become increasingly relevant for all models of functional connectivity.
+Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in big neural data will entail the search for optimal regularization schemes that may involve combinations of heuristic rules and numerical techniques specially designed for each neural circuit.
 
@@
 \item[$C_{\sf diag}$] -- linear shrinkage of correlations toward zero.
-\item[$C_{\sf factor}$] -- linear shrinkage of the covariances toward a low-rank matrix, representing inputs from unobserved factors.
+\item[$C_{\sf factor}$] -- linear shrinkage of the covariances toward a low-rank matrix, representing inputs from unobserved factors (latent units).
 \item[$C_{\sf sparse}$] -- sparse partial correlations, \ie a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
@@
 
-We therefore performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals.
-We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.
+We then performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals.
+In our data, the sample correlation coefficients were largely positive and low.
+We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.
 This estimator revealed a sparse network of partial correlations between the observed neurons and inferred several latent units exerting linear effects on the observed neurons.
@@
 \end{equation}
-where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.
+where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.
 The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf 0}$ from the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
@@
 We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
-In probabilistic models with exclusively linear dependencies, the targets estimates of these for estimators correspond to distinct families of graphical models (\figref{02}{\,Row 1}}).
+In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{02}{\,Row 1}}).
 
@@
 \end{equation}
-The structure of imposed by $C_{\sf diag}$ favors populations with no linear associations between the neurons (\figref{02}{Row 1, A}).
-If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.
+The structure imposed by $C_{\sf diag}$ favors populations with no linear associations between the neurons (\figref{02}{Row 1, A}).
+If sample correlations arise from spurious effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.
 
@@
     \end{center}
-    \caption{{\bf Example of low-dimensional correlation structure revealed by the sparse+low-rank estimator.}
+    \caption{{\bf Example of low-dimensional correlation structure revealed by the sparse+latent estimator.}
     {\bf A.} The regularized estimate of the correlation matrix (top-right) closely approximates the sample correlation matrix (bottom left). 
@@
     \end{center}
-    \caption{{\bf Properties of sparse+low-rank regularized estimates from all imaged sites}
+    \caption{{\bf Properties of sparse+latent estimates from all imaged sites}
 \\
@@
 In each site, the correlations were normalized by the respective average correlation shown in Fig.\;\ref{fig:06}\,A.
-The number $n$ of sites that qualified to be included in the analysis is indicated.
+The number of sites, $n$, that qualified to be included in the analysis is indicated in each panel.
 Sites were included if they had at least 20 pairs of neurons in each of the intervals.
-The error bars indicate the standard error of the mean for based on $n$.
+The error bars indicate the standard error of the mean based on on the number of sites $n$.
     {\bf A.} Average correlations between pairs of neurons tuned to orientation with differences in preferred orientation in the intervals of 0--15$^\circ$, 15--45$^\circ$ and 45--90$^\circ$. 
@@
 
-Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in big neural data will entail the search for optimal regularizations that may involve combinations of heuristic rules and numerical techniques specifically designed for each type of neural circuit.
-
 \paragraph{Model selection}
@@
 Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the inference that it reveals a better representation of dependencies in the data.
-This caveat is related to the phenomenon known as \emph{Stein’s Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
+This caveat is related to the phenomenon known as \emph{Stein's Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
 
@@
 The resulting network had a larger fraction of negative interactions than in the sample correlation matrix.
-The negative interactions had a distinct spatial and functional organization from the positive interactions.
+The negative interactions had a distinct spatial and functional organization from the positive interactions (\figref{fig:07}).
 To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.
@@
 
-We propose that the structure of the most efficient among many estimators provides a better picture of the functional connectivity than that obtained from unbiased sample correlations.
+Here we showed that among several models we tested, we found that for cortical microcircuits, a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix.
+This finding valuable in itself: it can be used to improve studies aiming to understand the nature of population coding.
+Moreover, this estimation approach also provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.
+For example, we find a specific structure of negative interactions which may indicate inhibitory circuitry.
+Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2012}  with our methods can be used to relate functional connectivity to the circuit architecture.
+Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation: this opens a window to describe what kinds of interactions are activated by stimulation conditions.
+Extending the estimation framework to temporal domain is important for future work.
+
+\begin{enumerate}
+    \item Better estimator of correlations -- better studies of coding, etc.
+    \item The interaction model may give more relevant - validation with new experiments 
+    \item Hopefully go beyond static connectivity -- but real-time, emergent, connectivity 
+    \item The ideal  experiments  would have causal manipulations 
+\end{enumerate}
 

### 2014-01-24 plos.tex

--- 2014-01-22 plos.tex
@@
     \end{center}
-    \caption{{\bf Example of the structure revealed by $C_{\sf sparse+latent}$ the sparse+low-rank estimator.}
+    \caption{{\bf Example of the structure revealed by $C_{\sf sparse+latent}$ the sparse+latent estimator.}
     {\bf A and D.} The regularized estimate $C_{\sf sparse+latent}$ closely approximates the sample correlation matrix $C_{\sf 0}$. 
@@
 Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the inference that it reveals a better representation of dependencies in the data.
-This caveat is related to the phenomenon known as \emph{Stein’s Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
+This caveat is related to the phenomenon known as \emph{Stein's Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
 
 \paragraph{Physiological interpretation and future directions}
+Although the graph of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.
+The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.
+The inferred graph of interactions differed substantially from the graph of most significant sample correlations \figref{05}{F, H, and I}.
+The $C_{\sf sparse+latent}$ estimator revealed a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}).
+These negative interactions are often produced as explanations for low sample correlation values where high values were expected based on common correlations with other neurons.
+The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.
+The resulting network had a larger fraction of negative interactions than in the sample correlation matrix.
+The negative interactions had a distinct spatial and functional organization from the positive interactions (\figref{fig:07}).
+To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.
+
+The estimator effectively decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.
+Since the distinction led to improved estimation, the two types of components could be further investigated.
+For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected on the common components/latent units but much less on the pairwise interactions.
+Previous studies found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.
+When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but does not affect the pairwise interactions as strongly.
+The correlation structure also changes as a function of the stimulus condition \cite{Cotton:2013} on time scales that are much faster than changes in the anatomical connectivity.
+Descriptions of specific changes in the functional connectivity due to stimulus properties could shed light on how circuits orchestrate their activity to represent diverse stimuli.
+
+Here we showed that among several models we tested, we found that for cortical microcircuits, a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix.
+This finding valuable in itself: it can be used to improve studies aiming to understand the nature of population coding.
+Moreover, this estimation approach also provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.
+For example, we find a specific structure of negative interactions which may indicate inhibitory circuitry.
+Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2012}  with our methods can be used to relate functional connectivity to the circuit architecture.
+Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation: this opens a window to describe what kinds of interactions are activated by stimulation conditions.
+Extending the estimation framework to temporal domain is important for future work.
+
 Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.
@@
 
-We propose that the structure of the most efficient among many estimators provides a better picture of the functional connectivity than that obtained from unbiased sample correlations.
+\begin{enumerate}
+    \item Better estimator of correlations -- better studies of coding, etc.
+    \item The interaction model may give more relevant - validation with new experiments 
+    \item Hopefully go beyond static connectivity -- but real-time, emergent, connectivity 
+    \item The ideal  experiments  would have causal manipulations 
+\end{enumerate}
 

### 2014-01-24 plos.tex
Rev 41.

@@
 In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of the synaptic connectivity \cite{Gerstein:1964}.
-The interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.
+The interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.
 
@@
 We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
-In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{02}{\,Row 1}}).
+In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{02}{\,Row 1}).
 
@@
 For example, we find a specific structure of negative interactions which may indicate inhibitory circuitry.
-Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2012}  with our methods can be used to relate functional connectivity to the circuit architecture.
+Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2011,Ko:2013}  with our methods can be used to relate functional connectivity to the circuit architecture.
 Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation: this opens a window to describe what kinds of interactions are activated by stimulation conditions.

### 2014-01-24 plos.tex
Kreso's edits of rev 41

@@
 Ambitious projects aim to record the activity of ever larger and denser neuronal populations \emph{in vivo}.
-Correlations in neural activity measured in such recordings can reveal important aspects of the organization of neural circuits.
+Correlations in neural activity measured in such recordings can reveal important aspects of  neural circuit organization.
 However, estimating and interpreting large correlation matrices is statistically challenging.
@@
 Indeed, in cross-validation tests, the covariance matrix estimator with this structure consistently outperformed other regularized estimators.
-The connectivity patterns inferred from the sparse component of the estimate were strongly related to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with  differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
-Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.
+The connectivity patterns inferred from the sparse component of the estimate were strongly related to the geometrical arrangement and orientation tuning properties of the recorded cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with  differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
+Because of its superior performance, this ``sparse + latent'' estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.
 
@@
 It is now possible to record the spiking activity of hundreds of neurons at the same time.
-A meaningful statistical description of the collective activity of these neural populations -- their \sq{functional connectivity} -- is a forefront challenge in neuroscience.
+A meaningful statistical description of the collective activity of these neural populations which reveals their \sq{functional connectivity} is a forefront challenge in neuroscience.
 We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.
@@
 The inferred interactions were both positive (\sq{excitatory}) and negative (\sq{inhibitory}) and reflected the spatial organization and orientation preferences of the interacting cells.
-We propose that the composition of the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.
+We propose that  the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.
 
@@
 \emph{Functional connectivity} is a statistical description of observed \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells.
-Functional connectivity is thought to reflect local synaptic connections, shared inputs from other regions, and endogenous network activity.
-Although functional connectivity is only a phenomenological description with no simple mechanistic interpretation, it can be used to generate or test hypotheses about the anatomical architecture of the neural circuit and the processing of information at the population level.
+Functional connectivity reflects local synaptic connections, shared inputs from other regions, and endogenous network activity.
+Although functional connectivity is a phenomenological description with no simple mechanistic interpretation, it can be used to generate hypotheses about the anatomical architecture of the neural circuit and test hypotheses about the processing of information at the population level.
 
@@
 In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
-In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of the synaptic connectivity \cite{Gerstein:1964}.
-The interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.
+In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of synaptic connectivity \cite{Gerstein:1964}.
+Interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010}.
 
 Neural correlations do not come with ready or unambiguous mechanistic interpretations.
-They can arise from monosynaptic or polysynaptic interactions, common or correlated inputs, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.
-But multineuronal recordings do provide more information than available from an equivalent number of pairs of cells considered separately.
+They can arise from monosynaptic or polysynaptic interactions, common or correlated inputs, oscillations, top-down modulation, and background network fluctuations, or a combination of these and other mechanisms \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.
+But multineuronal recordings do provide more information than an equivalent number of separately recorded pairs of cells.
 For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients.
-On the other hand, a correlation matrix can be specified through the  partial correlations between pairs of the recorded neurons.
-The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.
-For this reason, partial correlations have been proposed to more closely reflect the direct causal effects between components of complex systems than correlations; partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
-These opportunities have not yet been explored in neurophysiological studies: most studies have only consider the distribution of pairwise correlations \cite{Zohary:1994, Bair:2001, Smith:2008, Ecker:2010}.
+On the other hand, a correlation matrix can be specified using the emph{partial correlations} between pairs of the recorded neurons.
+The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all other recorded cells.
+For this reason, partial correlations more closely reflect the direct causal effects between components of complex systems than correlations; partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
+These opportunities have not yet been explored in neurophysiological studies, as most studies have only consider the distribution of pairwise correlations \cite{Zohary:1994, Bair:2001, Smith:2008, Ecker:2010}.
 
@@
 This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
-The \emph{sample correlation matrix} is an unbiased estimate of the true correlations but its many free parameters make it sensitive to the sampling noise.
+The \emph{sample correlation matrix} is an unbiased estimate of the true correlations but its many free parameters make it sensitive to sampling noise.
 As a result, on average, the sample correlation matrix is farther from the true correlation matrix than some structured estimates.
@@
 To \sq{impose a structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation  with fewer free parameters, which is often referred to as the \emph{target estimate}.
-The selection of the optimal target estimate and the optimal amount of shrinkage is deduced from the data sample either analytically \cite{Ledoit:2003, Ledoit:2004, Schafer:2005}  or by cross-validation \cite{Friedman:1989}.
+The optimal target estimate and the optimal amount of shrinkage can be obtained from the data sample either analytically \cite{Ledoit:2003, Ledoit:2004, Schafer:2005},  or by cross-validation \cite{Friedman:1989}.
 An estimator that produces estimates that are, on average, closer to the truth for a given sample size is said to be more \emph{efficient} than other estimators.
 
-Although regularized covariance matrix estimation is commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.
+Although regularized covariance matrix estimation is commonplace in finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identify optimal regularization of neural correlation matrices.
 
@@
 But reduced estimation error is not the only benefit of regularization.
-The amount of improvement of estimation is greatest when the structure imposed by regularization limits the solution space but still represents the dominant dependencies in the data.
-Therefore, by finding the most efficient among many regularized estimators, we learn something about the system itself: the imposed structure of the most efficient estimator is a parsimonious representation of the regularities in the data.
-The elements of this structure likely provide a better representation of the underlying dependencies in the data than previous analyses of sample correlations.
+The amount of improvement of estimation is greatest when the structure imposed by regularization limits the solution space, but still represents the dominant dependencies in the data.
+Therefore, by finding the most efficient among many regularized estimators, we learn something about the system itself: the structure of the most efficient estimator is a parsimonious representation of the regularities in the data.
 
 The necessity and benefits of regularization increase with the size of the recorded population.
-With the advent of the age of big neural data, the search for optimal regularization schemes will become increasingly relevant for all models of functional connectivity.
-Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in big neural data will entail the search for optimal regularization schemes that may involve combinations of heuristic rules and numerical techniques specially designed for each neural circuit.
+With the advent of  big neural data, the search for optimal regularization schemes will become increasingly relevant for all models of functional connectivity.
+Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in large scale neural data will entail the search for optimal regularization schemes that may involve combinations of heuristic rules and numerical techniques specially designed for each neural circuit.
 
@@
     \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
-        {\bf Row 1.} Graphical representations of the target estimates used of the four respective regularized covariance matrix estimators.
-Recorded neurons are denotes as green spheres and latent units as light-shaded spheres.
-The edges between them denote non-zero partial correlations.
-        {\bf Row 1, A}.  In estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear associations. 
-        {\bf  Row 1, B.} In estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), favoring systems in which correlations arise due to common input from latent units.
+        {\bf Row 1.} Graphical representations of the target estimates of the four respective regularized covariance matrix estimators.
+Recorded neurons are represented by green spheres and latent units by light-shaded spheres.
+Edges represent non-zero partial correlations.
+        {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear associations. 
+        {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units.
         {\bf  Row 1, C}.
-In estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix.
-This approximation describes systems in which correlations arise from linear associations between the observed units.
+For estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix.
+This approximation describes systems in which correlations arise from a sparse set of  linear associations between the observed units.
         {\bf  Row 1, D}.
-In estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix.
-This approximation describes a model wherein correlations arise due to linear associations between the recorded cells \emph{and} with several latent units.
+For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix.
+This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} with several latent units.
 \\
@@
 \\
-{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) attained by the five estimators as a function of sample size.
+{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) for the five estimators as a function of sample size.
 The error bars indicate the standard error of the mean based on 30 samples.
-Estimators with the matching structure converged to zero faster than the other estimators.
-Some exceptions occur when the sample is not large enough to reveal the true structure; in these cases estimators with simpler structures can be more efficient.
+Estimators with structure that matches the model used to generate the data converged to zero faster than the other estimators.
+Some exceptions occur when the sample is too small to reveal the structure in the data.
+In these cases estimators with simpler structures can be more efficient.
 \\
-{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) attained the five estimators with respect to the matching estimator.
+{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) for the five estimators with respect to the matching estimator.
 Error bars indicate the standard error of the mean based on 30 samples.
@@
 \end{equation}
-The structure imposed by $C_{\sf diag}$ favors populations with no linear associations between the neurons (\figref{02}{Row 1, A}).
+The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (\figref{02}{Row 1, A}).
 If sample correlations arise from spurious effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.
@@
 
-Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by a matrix whose inverse matrix is sparse: 
+Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix whose inverse is sparse: 
 \begin{equation}\label{eq:c-sparse}
-C_{\sf sparse} = S^{-1},
+C_{\sf sparse} = S^{-1}.
 \end{equation}
-where $S$ is a sparse matrix, \ie one in which a large fraction of off-diagonal elements are set to zero.
+Here $S$ is a sparse matrix, \ie one in which a large fraction of off-diagonal elements are set to zero.
 The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$.
 The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{02}{Row 1, C}).
-We will sometimes refer to sparse partial interactions between neurons as simply \sq{interactions}.
+We will sometimes refer to sparse partial correlations between neurons simply  as \sq{interactions}.
 
-Estimator $C_{\sf sparse+latent}$ is produced by approximating the sample covariance matrix by a matrix whose inverse is the sum of a sparse component and a low-rank component: 
+Estimator $C_{\sf sparse+latent}$ is obtained by approximating the sample covariance matrix by a matrix whose inverse is the sum of a sparse component and a low-rank component: 
 \begin{equation}\label{eq:c-sl}
@@
 \paragraph{Simulation}
-To demonstrate how the selection between regularized estimators allows discerning the structure of correlations, we constructed four $50\times 50$ covariance matrices with the structures imposed by the four respective regularized estimators (\figref{02}{\,Row 2, A--D}).
-See the Methods section for details of how these matrices were generated.
+We next demonstrat how the most efficient among different regularized estimators can reveal the structure of correlations.
+We constructed four $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (\figref{02}{\,Row 2, A--D}, see Methods for details about how these matrices were generated).
 We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes.
@@
 
-The evaluation of a covariance matrix estimator $C$ is performed with respect to a \emph{loss function} $\loss{C,\Sigma}$ to quantify the discrepancy between the estimate and the truth, $\Sigma$.
+The evaluation of a covariance matrix estimator, $C,$ is performed with respect to a \emph{loss function} $\loss{C,\Sigma}$ to quantify the discrepancy between the estimate and the truth, $\Sigma$.
 The loss function must attain its minimum when $C=\Sigma$.
@@
 
-This loss function derives from the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$ as:
+This loss function is related to the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$,
 \begin{equation}
@@
 \end{equation}
-The normalization by $p$ makes the values of the loss function comparable across different population sizes.
+Normalization by $p$ makes the values of the loss function comparable across different population sizes.
 
-Although the two functions appear similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf 0})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.
+Although the two functions are similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf 0})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.
 
@@
 \begin{equation}\label{eq:excess-loss}
-    \eloss{C,\Sigma} = \loss{C,\Sigma}-\loss{\Sigma,\Sigma},
+    \eloss{C,\Sigma} = \loss{C,\Sigma}-\loss{\Sigma,\Sigma}
 \end{equation}
 
-We drew 30 independent samples of size $n=250$, 500, 1000, 2000, and 4000 from model and computed the excess loss resulting from the five estimators.
+We drew 30 independent samples of size $n=250$, 500, 1000, 2000, and 4000 from model and computed the excess loss for each of the five estimators.
 The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.
@@
 There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$, often outperformed the estimators with the matching structure (\figref{02}{\,Row 5, C and D}).
-Second, $C_{\sf sparse+latent}$ often performed equally well to $C_{\sf sparse}$ on the data from models without latent units because it reliably inferred the right number of latent units, zero.
+Second, $C_{\sf sparse+latent}$ often performed as well as $C_{\sf sparse}$ on data from models without latent units because.
+However, in theses cases it reliably inferred the correct number of latent units which was zero.
 
-Even though, with increasing sample sizes, all estimators converged to the truth (zero excess loss), the distinction between the estimators did not decrease.
+With increasing sample sizes, all estimators converged to the truth (zero excess loss).  However, 
+the discriminability between their performances increased with sample size (data not shown).
 
-In empirical data, the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly.
-Instead, the loss function can be estimated solely from the data through \emph{validation}.
-In validation, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf 0}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.
-Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can substitute the loss function $\loss{C,\Sigma}$.
+With empirical data, the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly.
+Instead, the loss function must be estimated from the sample through \emph{validation}.
+In a validation procedure, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf 0}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.
+The estimate $C_{\sf 0}^\prime$ is used to
+compute the \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$, which approximates the loss  $\loss{C,\Sigma}$.
 
-The normal loss (Eq.~\ref{eq:loss}) is particularly convenient because it is additive in its second argument in the sense that 
+The normal loss (Eq.~\ref{eq:loss}) is particularly convenient because it is additive in its second argument, \ie
  \begin{equation*}\label{eq:additivity}
@@
  \end{equation*}
-
-In this case, the validation loss is an unbiased estimate of the loss function:
+With this property of additivity, the validation loss is an unbiased estimate of the loss function:
 \begin{equation*}
-    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}^\prime]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}
+    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}^\prime]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}.
 \end{equation*}
@@
 Then each subset is used as the validation sample with the other $K-1$ serving as the training sample.
-The validation losses from each of such \sq{folds} are averaged to produce the \emph{cross-validation loss} or \emph{CV-loss} for short.
-Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ trained on the remaining $K-1$ subsets.
+The validation losses from all such \sq{folds} are averaged to produce the \emph{cross-validation loss} or \emph{CV-loss} for short.
+Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ when the remaining $K-1$ subsets are treated as the training sample.
 Then the cross-validation loss for estimator $C$ is
@@
 In this formulation, $\ell_C$ can be recognized as the cross-validated Gaussian log-likelihood (up to a constant offset and multiplier).
-However, the present framework is more general, allowing for other loss functions to be used.
+However, the present framework is more general, as other loss functions can be used.
 
-Since we only need to compare estimators to each other, we are only interested in the \emph{relative} CV loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$:
+Since we only need to compare estimators to each other, we are  interested in the \emph{relative} CV loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$,
 \begin{equation}\label{eq:rel-cv-loss}
@@
 \end{equation}
-The relative cross-valiation loss was used for the main findings in this study.
+The relative cross-validation loss was used for the main findings in this study.
 
-To examine how well the relative CV-losses reproduced the relationships between the excess losses, we evaluated the relative CV-loss with respect to the estimator that matched the true structure of the model model.
+To examine how well the relative CV-losses reproduced the relationships between the excess losses, we evaluated the relative CV-loss with respect to the estimator that matched the true structure of the model.
 Indeed, CV-losses accurately reproduced the relationships between the excess losses of the estimators, with somewhat greater variability (\figref{02}{Row 6}).
 
-This simulation study demonstrated that, with sufficiently large sample sizes, imposing the correct type of structure on the estimate led to greater estimation improvement than imposing other structures.
+This simulation study demonstrated that, with sufficiently large sample sizes, imposing the correct type of structure on the estimate leads to greater estimation improvement than imposing other structures.
 
@@
 Only cells that produced detectable calcium activity were included in the analysis (For details, see Methods).
-First, 30 repetitions of full-field drifting gratings of 16 direction were presented in random order for 500 ms without blanks.
+First, 30 repetitions of full-field drifting gratings of 16 directions were presented in random order.
+Each grating was presented for 500 ms, without intervening blanks.
 This stimulus was used to compute the orientation tuning of the recorded cells (\figref{03}{E}).
-To estimate the noise correlation matrix, we presented only 2 or 5 distinct directions with between 100 and 300 repetitions of each direction.
+To estimate the noise correlation matrix, we presented only 2 distinct directions in some experiments or 5 directions in others with between 100 and 300 repetitions of each direction.
 In this stimulus, each grating lasted 1 second and was followed by a 1-second blank.
@@
     {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site.
-The cells' colors indicate their orienation preferences.
-The gray cells were not significantly tuned to orientation.
-    {\bf F.} The noise correlation matrix of the activity of the neural population. 
+The cells' colors indicate their orientation preferences.
+The gray cells were not significantly tuned.
+    {\bf F.} The sample noise correlation matrix of the activity of the neural population. 
     {\bf G.} Histogram of noise correlation coefficients. The red line indicates the mean.
@@
     {\bf A--D.} Histograms of the relative cross-validation loss of estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$.
-    The histograms are based on 31 imaged sites in 15 mice. 
+    The histograms were obtained from 31 imaged sites in 15 mice. 
     All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
@@
 Because of its superior estimation performance, $C_{\sf sparse+latent}$ likely provides a more physiologically relevant representation of the functional connectivity than the other estimators.
-The estimator segregates two sources of correlations: linear dependencies between pairs of neurons (sparse component) suggestive of direct interactions and common fluctuations across the entire population (low-rank component) component suggestive of common inputs, cell assembly activation, network fluctuations or other forms of synchronized activations.
+The estimator segregates two sources of correlations:  The first source was the network of linear dependencies between pairs of neurons.
+The second source was the set of latent units reflecting common fluctuations across the entire population.
+These could be caused by  common inputs, cell assembly activation, network fluctuations or other forms of synchronized activations.
 
-We examined the composition of the $C_{\sf sparse+latent}$ estimates in each imaged site.
-The regularized estimates appeared similar to the sample correlation matrix (\figref{05}{A and D}) but the corresponding partial correlation matrices differed substantially (\figref{05}{B and E}).
+We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site.
+Although the regularized estimates were similar to the sample correlation matrix (\figref{05}{A and D}), the corresponding partial correlation matrices differed substantially (\figref{05}{B and E}).
 The partial correlation matrix of the regularized estimate was decomposed into a sparse component and a low-rank component (\figref{05}{C}).
-While sample correlations were mostly positive, the sparse partial correlations had a high fraction of negative interactions.
+While sample correlations were mostly positive, the sparse partial correlations contained a high fraction of negative interactions.
 In the example site (Fig.~\ref{fig:03} and Fig.~\ref{fig:05}),  the sparse component had 82.2\% sparsity (or, conversely, 17.8\% connectivity) with average node degree of 52.8 (\figref{05}{G}).
@@
 The networks revealed by the two methods differed substantially.
-In the example site with 17.8\% connectivity, only 41.7\% of cell pairs connected in the $C_{\sf sparse+latent}$ graph were also connected in the graph of thresholded sample correlations (\figref{05}{F, H, and I}).
-In particular, the thresholded sample correlations were predominantly positive whereas 43.3\% of the interactions in $C_{\sf sparse+latent}$ were negative.
-Most of the negative interactions corresponded to low sample correlations (\figref{05}{F}).
-Thus the $C_{\sf sparse+latent}$ estimates reveal negative interactions that account for low correlations where high correlations are predicted from all the other interactions.
+In the example site with 17.8\% connectivity inferred by $C_{\sf sparse+latent}$, only 41.7\% of connections coincided with above-threshold sample correlations (\figref{05}{F, H, and I}).
+In particular, the thresholded sample correlations were predominantly positive whereas 43.3\% of the interactions obtained using $C_{\sf sparse+latent}$ were negative.
+Most of the negative interactions corresponded to low sample correlations (\figref{05}{F}) where high correlations were expected given the rest of the correlation matrix.
 
-The average partial correlations revealed by $C_{\sf sparse+latent}$ estimates in all 31 sites were about 5 times lower than the sample correlations and less variable across sites (\figref{06}{A}): the coefficient of variation of the average sample correlations across all sites was 0.41 whereas that of the average partial correlations was only 0.35.
+The average partial correlations revealed by $C_{\sf sparse+latent}$ estimates at all 31 sites were about 5 times lower and less variable across sites than sample correlations (\figref{06}{A}): the coefficient of variation of the average sample correlations across all sites was 0.41 whereas that of the average partial correlations was only 0.35.
 The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (\figref{06}{B and C}).
@@
 Several sites, even with relatively large population sizes, had fairly few pairwise interactions and were dominated by latent units.
-These differences may be explained by differences in brain states, recording quality (neuropil contamination, motion).
+These differences may be explained by differences in brain states and recording quality.
+For example neuropil contamination, motion and other factors may have degraded the quality of the data.
 
@@
     {\bf A.} The average sample correlations vs.~average partial correlations for each imaged site. In each plot, the red asterisk indicates the site shown in figures \ref{fig:03} and \ref{fig:05}.
-    {\bf B.} The average node degree for sparse partial correlations vs.~population size in each imaged site. 
+    {\bf B.} The average node degree for sparse partial correlations vs.~population size at each imaged site. 
     {\bf C.} The number of inferred latent units vs.~population size in each imaged site.
@@
 
-We also examined the relationship of the correlation structure to the differences in orientation preference and to the geometric distances between the cell pairs (Fig.\;\ref{fig:07}).
+We also examined the relationship of the correlation structure to the differences in orientation preference and to the physical distances between the cell pairs (Fig.\;\ref{fig:07}).
 The partial correlations of the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation (\figref{07}{A}) than sample correlations.
-They also decreased more rapidly with geometric distance, both laterally  (\figref{07}{B}) and in depth (\figref{07}{B}).
+They also decreased more rapidly with physical distance, both laterally  (\figref{07}{B}) and in depth (\figref{07}{B}).
 
@@
     {\bf A--C.} Average partial correlations (red) estimated by $C_{\sf sparse+latent}$ and average sample correlations (black) averaged across multiple imaged sites.
-In each site, the correlations were normalized by the respective average correlation shown in Fig.\;\ref{fig:06}\,A.
+At each site the correlations were normalized by the respective average correlation shown in Fig.\;\ref{fig:06}\,A.
 The number of sites, $n$, that qualified to be included in the analysis is indicated in each panel.
@@
 Normalized connectivity was computed as the fraction of pairs connected by interactions of corresponding signs in each interval divided by the fraction of non-zero interactions across the entire site.
-Site were included in the analysis only if they had at least 20 cell pairs in each interval.
-The error bars indicate the standard error of the mean based on the number $n$ of sites included in the analysis.
 }
@@
 \section*{Acknowledgments}
-We thank Genevera Allen for a helpful discussion.  We thank Eftychios Pnevmatikakis for helpful suggestions and feedback.
+We thank Genevera Allen for a helpful discussion, and Eftychios Pnevmatikakis for helpful suggestions and feedback.
 

### 2014-01-24 plos.tex
minor

@@
 We sought statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.
-Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.
+Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.
 We hypothesized that in these dense recordings, the correlation matrix should be best represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.
 Indeed, in cross-validation tests, the covariance matrix estimator with this structure consistently outperformed other regularized estimators.
-The connectivity patterns inferred from the sparse component of the estimate were strongly related to the geometrical arrangement and orientation tuning properties of the recorded cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with  differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
+The sparse component of the estimate expressed a graph of interactions which were organized with respect to the physical distances and orientation tuning properties of cells: The density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
 Because of its superior performance, this ``sparse + latent'' estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.
@@
 We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.
-This estimator revealed a sparse network of partial correlations between the observed neurons and inferred several latent units exerting linear effects on the observed neurons.
-We analyzed these networks of partial correlations and found the following: Whereas noise correlations were predominantly positive, the inferred partial correlations had a large fraction of negative values possibly reflecting inhibitory circuitry.
-Moreover, we found that the inferred positive partial correlations exhibited stronger dependencies with respect to physical distances and differences in preferred orientations than noise correlations.
-In contrast, the inferred negative partial correlations were less selective.
+This estimator revealed a sparse network of partial correlations, which we call \sq{interactions} here, between the observed neurons and inferred several latent units exerting linear effects on the observed neurons.
+We analyzed these networks of partial correlations and found the following: Whereas noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.
+Moreover, we found that these interactions exhibited a stronger relationship on the physical distances and differences in preferred orientations than noise correlations.
+In contrast, the inferred negative interactions were less selective.
 

### 2014-01-26 plos.tex
Kreso's changes of the Discussion.

--- 2014-01-24 plos.tex
@@
 \paragraph{Functional connectivity as a network of pairwise interactions}
-Functional connectivity is often expressed as a graph of pairwise interactions.
-The intended purpose of many studies of functional connectivity is to estimate the circuit's anatomical connectivity solely from the observed multineuronal spiking activity.
-For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections or shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
-Such signatures are still ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.
-With simultaneous recordings from a larger subset of neurons in the circuit, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.
-Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the rest of the network.
+Functional connectivity is often represented as a graph of pairwise interactions.
+The goal of many studies of functional connectivity was to estimate  anatomical connectivity from  observed multineuronal spiking activity.
+For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
+Such signatures are ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.
+With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.
+Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network.
 Therefore, conditionally independent neurons can be inferred to lack a direct causal influence.
-Such conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity.
-For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, individual neurons' stimulus drive, and a network of synaptic couplings \cite{Pillow:2008}.
+
+Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity.
+For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.
 Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .
-Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with minimal other assumptions about the distribution: the investigator need only choose which statistics to match.
+%Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with minimal other assumptions about the distribution: the investigator need only choose which statistics to match.
 Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
-Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of neurons.
+Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.
 
-Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the interaction between a pair of neurons on the activity of the other neurons under a given probabilistic model, the recording must include all the other neurons with which the pair can potentially interact.
-Otherwise, interactions with unobserved portions of the circuit can introduce false conditional dependencies between observed neurons.
-For this reason, the most successful applications of statistical models of population activity have been in \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations have been possible \cite{Pillow:2008}.
-In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the conditional dependencies between pairs of cells.
-Perhaps for this reason, partial correlations have not been previously used to describe the functional connectivity between cortical neurons.
-Despite this limitation, many studies have fitted probabilistic models to multielectrode recordings and analyzed the structure of the resulting conditional dependencies \cite{Tang:2008, Yu:2008}.
+%Correct inference of causal interactions between neurons depends on correct mathematical representations of the interactions in the statistical model and the completeness of the recorded population.
+It is not yet clear which approach provides the best correspondence with anatomical connectivity.
+Little experimental evidence is available to answer this question.
+The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without examining their relation to anatomical implications.
+Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
+However, the topological properties of functional connectivity graphs vary with the method of their inference (Zalesky:2012).
+Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.
 
-Correct inference of causal interactions between neurons depends on correct mathematical representations of the interactions in the statistical model and the completeness of the recorded population.
-It is not yet clear which approach provides best correspondence to the anatomical connectivity in the circuit, with little experimental evidence yet available to address this question.
-The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without a statement about its anatomical implications.
-Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
-However, the topological properties of the functional connectivity graphs vary with the method of their inference (Zalesky:2012); until the physiological meaning of the functional connectivity is established, the validity of such analyses remains in question.
+Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the pairwise interactions on the activity of the other neurons, we must record from all neurons with which the pair interacts.
+Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons.
+For this reason, the most statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}.
+In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the pairwise conditional dependencies.
+This may be the reason that partial correlations have not previously been used to describe  functional in cortical populations.
+% Despite this limitation, many studies have fitted probabilistic models to multielectrode recordings and analyzed the structure of the resulting conditional dependencies \cite{Tang:2008, Yu:2008}.  
+\Kcomment{I thought this was a good way to end, and the last sentence went in a different direction.}
 
 Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.
-While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with advanced spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}.
-However, such high temporal precision comes at the cost of the accuracy of the inferred spike rates: better accuracy can be achieved when calcium signals are analyzed on the temporal scales of tens of milliseconds \cite{Cotton:2013}.
+While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}.
+However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.
+Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.
 The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.
-Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
-These techniques may open new opportunities for more incisive measurements of functional connectivity than with electrophysiological recordings.
+Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
+These techniques may provide more incisive measurements of functional connectivity than classical electrophysiological recordings.
 
-The low temporal resolution of calcium signals limits the use of functional connectivity methods that rely on millisecond-scale binning of signals (cross-correlograms, some GLMs, and maximum entropy models on the binary domain).
+The low temporal resolution of calcium signals limits the use of functional connectivity methods that rely on millisecond-scale binning of signals (cross-correlograms, some GLMs, and binary maximum entropy models).
 Hence, most studies of functional connectivity have relied on instantaneous sample correlations \cite{Greenberg:2008, Golshani:2009, Hofer:2011, Malmersjo:2013} .
-Although some investigators have interpreted such correlations as indicators of (chemical or electrical) synaptic connectivity, most used them as more general indicators of functional connectivity without a statement about underlying mechanisms.
+Although some investigators have interpreted such correlations as indicators of (chemical or electrical) synaptic connectivity, most used them as more general indicators of functional connectivity without relating them to underlying mechanisms.
 
-In our study, we sought to infer pairwise functional connectivity networks among neurons in cortical microcircuits.
-We hypothesized that partial correlations would correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.
-Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we a priori favored solutions with sparse partial correlations.
+In our study, we sought to infer pairwise functional connectivity networks  in cortical microcircuits.
+We hypothesized that partial correlations correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.
+Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we \emph{a priori} favored solutions with sparse partial correlations.
 The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution.
 Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
+\Kcomment{I am not sure about the previous argument.}
 
 \paragraph{Functional connectivity as coactivations}
-Another approach to describe the functional connectivity of a circuit is to isolate patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}.
+Another approach to describing the functional connectivity of a circuit is to isolate patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}.
 Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}.
 Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \ie groups of tightly interconnected groups of cells involved in a common computation.
-Coactivation patterns could also be explained as shared input from unobserved parts of the circuit.
-Coactivations could also be explained by global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.
+Coactivation patterns could also result from shared input from unobserved parts of the circuit, or global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.
 
-Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity, but analysis of coactivations method shifts the focus from detailed interactions to the collective behavior.
-
+Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity, but an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.
 In our study, the analysis of functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.
@@
 The resulting estimator, $C_{\sf sparse+latent}$, effectively decomposed the functional connectivity into a sparse network of pairwise interactions and coactivation mode vectors.
+
 \paragraph{Addressing ill-posedness}
 Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.
-The problem becomes increasingly worse as the size of the recorded population of neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
-As the recorded neuronal population sizes increased gradually, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.
-However, with ambitious projects such as the BRAIN initiative currently underway \cite{Alivisatos:2013}, aiming to record from significantly larger populations, the problem must be addressed by regularizing the solution.
+The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
+As  techniques improved to allow for recording from larger neuronal populations, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.
+However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
+Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators.
 Regularization limits the solution to a smaller subspace to counteract the effect of the sampling noise in the empirical data.
-However, limiting the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.
+\Kcomment{This is not really true -- it doesn't limit the solution to a smaller subspace.
+It just shrinks it toward it.} However, limiting the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.
 
@@
 \paragraph{Model selection}
-In our study, the covariance matrix estimators were evaluated with respect to the cross-validated normal log likelihood.
+We evaluated the covariance matrix estimators using the cross-validated normal log likelihood.
 However, this does not limit the applicability of its conclusions to normal distributions.
-Indeed the major findings in this paper could be reproduced with respect to other loss functions (compare Fig.~\ref{fig:04} and Fig.~S\ref{supp:02}).
+Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:04} and Fig.~S\ref{supp:02}).
 Other probabilistic models, fitted to the same data, would also serve as estimators of the covariance matrix.
-If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.
+\Kcomment{A probabilistic model is not an estimator.} If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.
 
 The results of model selection must be interpreted with caution.
-As we demonstrated by simulation, a simple model can produce a better estimate even if it does not correctly represent the real dependencies in the data than a more complex model with correct representation of dependencies.
-Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the inference that it reveals a better representation of dependencies in the data.
-This caveat is related to the phenomenon known as \emph{Stein's Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
+As we demonstrated by simulation, a wrong simple model can produce a better estimate than a more complex model with correct representation of dependencies.
+Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the conclusion that it reveals a better representation of dependencies in the data.
+This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
 
@@
 Although the graph of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.
-The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.
+The improved estimation performance gives the model additional credence over other estimates of the functional connectivity.
 The inferred graph of interactions differed substantially from the graph of most significant sample correlations \figref{05}{F, H, and I}.
 The $C_{\sf sparse+latent}$ estimator revealed a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}).
-These negative interactions are often produced as explanations for low sample correlation values where high values were expected based on common correlations with other neurons.
+%These negative interactions are often produced as explanations for low sample correlation values where high values were expected based on common correlations with other neurons.
 The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.
@@
 The negative interactions had a distinct spatial and functional organization from the positive interactions (\figref{fig:07}).
-To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.
+To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether negative interactions preferentially involve inhibitory cells.
 
-The estimator effectively decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.
+The  $C_{\sf sparse+latent}$ estimator decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.
 Since the distinction led to improved estimation, the two types of components could be further investigated.
-For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected on the common components/latent units but much less on the pairwise interactions.
-Previous studies found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.
+For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected in the common components/latent units but much less in pairwise interactions.
+Previous studies found that correlations measured under anesthesia retained little in common with correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.
 When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but does not affect the pairwise interactions as strongly.
@@
 
-Here we showed that among several models we tested, we found that for cortical microcircuits, a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix.
-This finding valuable in itself: it can be used to improve studies aiming to understand the nature of population coding.
+Here we showed that among several models we tested, for cortical microcircuits, a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix.
+This finding is valuable in itself: it can be used to improve studies aiming to understand the nature of population coding.
 Moreover, this estimation approach also provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.
@@
 Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2011,Ko:2013}  with our methods can be used to relate functional connectivity to the circuit architecture.
-Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation: this opens a window to describe what kinds of interactions are activated by stimulation conditions.
+Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation.
+This opens a window to describe what kinds of interactions are activated by stimulation conditions.
 Extending the estimation framework to temporal domain is important for future work.
 
-Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.
+%Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.
 The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.
@@
 To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.
+\Kcomment{The paragraph here (erased) was repeated from above.}
 
@@
 Descriptions of specific changes in the functional connectivity due to stimulus properties could shed light on how circuits orchestrate their activity to represent diverse stimuli.
+\Kcomment{This is also repeated.}
 

### 2014-01-26 plos.tex
rev43: completed Discussion, added two more panels to Figure 6.

@@
 The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf 0}$ from the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
-\begin{equation}
+\begin{equation}\label{eq:sample}
     C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
@@
 Yet 58.3\% of the interactions inferred by $C_{\sf sparse+latent}$  linked pairs of neurons whose correlation was below the threshold.
-Only the remaning 41.7\% of the interactions overlapped with sample correlations above the theshold.
-    {ebf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
+Only the remaining 41.7\% of the interactions overlapped with sample correlations above the threshold.
+    {\bf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
 The line density is proportional to the magnitude of the correlation.
@@
 
-We compared the graphs of partial correlations revealed by $C_{\sf sparse+latent}$ to the graph obtained by thresholding the sample correlations.
+We compared the graphs of interactions inferred by $C_{\sf sparse+latent}$ to the graph obtained by thresholding the sample correlations.
 Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013}.
@@
 
-The average partial correlations revealed by $C_{\sf sparse+latent}$ estimates at all 31 sites were about 5 times lower and less variable across sites than sample correlations (\figref{06}{A}): the coefficient of variation of the average sample correlations across all sites was 0.41 whereas that of the average partial correlations was only 0.35.
-The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (\figref{06}{B and C}).
-However, there was an inverse relationship between the number of latent units and the average node degree (\figref{06}{D}).
+The average partial correlations revealed by $C_{\sf sparse+latent}$ estimates at all 31 sites were about 5 times lower and less variable across sites than sample correlations (\figref{06}{D}): the coefficient of variation of the average sample correlations across all sites was 0.41 whereas that of the average partial correlations was only 0.35.
+The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (\figref{06}{A and B}).
+However, there was an inverse relationship between the number of latent units and the average node degree (\figref{06}{C}).
 Several sites, even with relatively large population sizes, had fairly few pairwise interactions and were dominated by latent units.
@@
 
-\begin{figure}[!ht]
-    \begin{center}
-        \includegraphics{./figures/Figure06.pdf}
-    \end{center}
-    \caption{{\bf Properties of sparse+latent estimates from all imaged sites}
+\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
+    {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites} Each point represents an imaged site with its color indicating the population size as shown in panels A and B.
+The example site from Figures \ref{fig:03} and \ref{fig:05} is circled in blue.
 \\
-    {\bf A.} The average sample correlations vs.~average partial correlations for each imaged site. In each plot, the red asterisk indicates the site shown in figures \ref{fig:03} and \ref{fig:05}.
+    {\bf A.} The number of inferred latent units vs.~population size in each imaged site.
     {\bf B.} The average node degree for sparse partial correlations vs.~population size at each imaged site.
-    {\bf C.} The number of inferred latent units vs.~population size in each imaged site.
-    {\bf D.} The number latent units vs.~average node degree for sparse partial correlations for each site.
+    {\bf C.} The average node degree in the sparse component in each site vs.~the number of latent units described by the low-rank component of the estimate.
+    {\bf D.} The average sample correlations  vs.~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
+    {\bf E.} Percentage overlap between the highest sample correlations and the  interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie 1-sparsity) of its sparse component.
+    {\bf F.} The percentage of negative interactions vs.~connectivity.
 }
-\label{fig:06}
+\label{fig:06}}
+{\includegraphics{./figures/Figure06.pdf}}
 \end{figure}
@@
     {\bf A--C.} Average partial correlations (red) estimated by $C_{\sf sparse+latent}$ and average sample correlations (black) averaged across multiple imaged sites.
-At each site the correlations were normalized by the respective average correlation shown in Fig.\;\ref{fig:06}\,A.
+At each site the correlations were normalized by the respective average correlation shown in \figref{06}{D}.
 The number of sites, $n$, that qualified to be included in the analysis is indicated in each panel.
@@
 Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .
-%Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with minimal other assumptions about the distribution: the investigator need only choose which statistics to match.
 Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
@@
 
-%Correct inference of causal interactions between neurons depends on correct mathematical representations of the interactions in the statistical model and the completeness of the recorded population.
 It is not yet clear which approach provides the best correspondence with anatomical connectivity.
@@
 Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
-However, the topological properties of functional connectivity graphs vary with the method of their inference (Zalesky:2012).
+However, the topological properties of functional connectivity graphs vary with the method of their inference \cite{Zalesky:2012}.
 Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.
@@
 This may be the reason that partial correlations have not previously been used to describe  functional in cortical populations.
-% Despite this limitation, many studies have fitted probabilistic models to multielectrode recordings and analyzed the structure of the resulting conditional dependencies \cite{Tang:2008, Yu:2008}.  
-\Kcomment{I thought this was a good way to end, and the last sentence went in a different direction.}
 
@@
 Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
-\Kcomment{I am not sure about the previous argument.}
+%\Kcomment{I am not sure about the previous argument.}
 
@@
 Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators.
-Regularization limits the solution to a smaller subspace to counteract the effect of the sampling noise in the empirical data.
-\Kcomment{This is not really true -- it doesn't limit the solution to a smaller subspace.
-It just shrinks it toward it.} However, limiting the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.
+Regularization biases the solution toward a smaller subspace to counteract the effect of the sampling noise in the empirical data.
+However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.
 
@@
 Other probabilistic models, fitted to the same data, would also serve as estimators of the covariance matrix.
-\Kcomment{A probabilistic model is not an estimator.} If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.
+If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.
 
@@
 \paragraph{Physiological interpretation and future directions}
-Although the graph of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.
-The improved estimation performance gives the model additional credence over other estimates of the functional connectivity.
-The inferred graph of interactions differed substantially from the graph of most significant sample correlations \figref{05}{F, H, and I}.
-The $C_{\sf sparse+latent}$ estimator revealed a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}).
-%These negative interactions are often produced as explanations for low sample correlation values where high values were expected based on common correlations with other neurons.
-The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.
- The resulting network had a larger fraction of negative interactions than in the sample correlation matrix.
-The negative interactions had a distinct spatial and functional organization from the positive interactions (\figref{fig:07}).
-To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether negative interactions preferentially involve inhibitory cells.
 
-The  $C_{\sf sparse+latent}$ estimator decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.
-Since the distinction led to improved estimation, the two types of components could be further investigated.
-For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected in the common components/latent units but much less in pairwise interactions.
-Previous studies found that correlations measured under anesthesia retained little in common with correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.
-When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but does not affect the pairwise interactions as strongly.
-The correlation structure also changes as a function of the stimulus condition \cite{Cotton:2013} on time scales that are much faster than changes in the anatomical connectivity.
-Descriptions of specific changes in the functional connectivity due to stimulus properties could shed light on how circuits orchestrate their activity to represent diverse stimuli.
+Here we showed that among several models a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix for cortical microcircuits.
+This finding is valuable in itself: improved estimates of the noise covariance matrix for large data sets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Ecker:2011}
 
-Here we showed that among several models we tested, for cortical microcircuits, a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix.
-This finding is valuable in itself: it can be used to improve studies aiming to understand the nature of population coding.
-Moreover, this estimation approach also provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.
-For example, we find a specific structure of negative interactions which may indicate inhibitory circuitry.
-Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2011,Ko:2013}  with our methods can be used to relate functional connectivity to the circuit architecture.
-Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation.
-This opens a window to describe what kinds of interactions are activated by stimulation conditions.
-Extending the estimation framework to temporal domain is important for future work.
-
-%Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.
-The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.
-The inferred graph of interactions was substantially different from the graph of the most significant sample correlations \figref{05}{F, H, and I}.
-The $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that are not present in the sample correlation matrix (\figref{05}{F}).
-These negative interactions are produced as explanations for low sample correlation values where high values are expected based on common correlations with other neurons.
-The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.
-The resulting network had a larger fraction of negative interactions than in the sample correlation matrix.
-The negative interactions had a distinct spatial and functional organization from the positive interactions.
-To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.
-\Kcomment{The paragraph here (erased) was repeated from above.}
-
-The estimator effectively decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.
-Since the distinction led to improved estimation, the two types of components could be further analyzed and related to physiological processes.
-For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected on the common components/latent units but much less on the pairwise interactions.
-Previous studies found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.
-When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but has little effect on the pairwise interactions.
-The correlation structure also changes as a function of the stimulus condition \cite{Cotton:2013} on time scales that are much faster than changes in the anatomical connectivity.
-Descriptions of specific changes in the functional connectivity due to stimulus properties could shed light on how circuits orchestrate their activity to represent diverse stimuli.
-\Kcomment{This is also repeated.}
-
-\begin{enumerate}
-    \item Better estimator of correlations -- better studies of coding, etc.
-    \item The interaction model may give more relevant - validation with new experiments 
-    \item Hopefully go beyond static connectivity -- but real-time, emergent, connectivity 
-    \item The ideal  experiments  would have causal manipulations 
-\end{enumerate}
+Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.
+Importantly, the inferred functional interactions were substantially different from those measured by considering the network of the most significant sample correlations \figref{05}{F, H, and I}.
+For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}) that may reflect inhibitory circuitry.
+Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity it may accurately reflect average connectivity profiles.
+Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \ref{Hofer:2012, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
+Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation.
+For example, these latent units may be modulated under different brain states (e.g.
+slow-wave sleep, attention) and stimulus conditions (e.g.
+certain types of stimuli may engage feedback connections)
 

### 2014-01-27 plos.tex
minor corrections

@@
 Indeed, in cross-validation tests, the covariance matrix estimator with this structure consistently outperformed other regularized estimators.
-The sparse component of the estimate expressed a graph of interactions which were organized with respect to the physical distances and orientation tuning properties of cells: The density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
+The sparse component of the estimate expressed a graph of interactions organized with respect to the physical distances and orientation tuning properties of cells: The density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
 Because of its superior performance, this ``sparse + latent'' estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.
@@
 It is now possible to record the spiking activity of hundreds of neurons at the same time.
-A meaningful statistical description of the collective activity of these neural populations which reveals their \sq{functional connectivity} is a forefront challenge in neuroscience.
+A meaningful statistical description of the collective activity of these neural populations -- their \sq{functional connectivity} -- is a forefront challenge in neuroscience.
 We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.
@@
 But multineuronal recordings do provide more information than an equivalent number of separately recorded pairs of cells.
-For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients.
-On the other hand, a correlation matrix can be specified using the emph{partial correlations} between pairs of the recorded neurons.
-The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all other recorded cells.
-For this reason, partial correlations more closely reflect the direct causal effects between components of complex systems than correlations; partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
-These opportunities have not yet been explored in neurophysiological studies, as most studies have only consider the distribution of pairwise correlations \cite{Zohary:1994, Bair:2001, Smith:2008, Ecker:2010}.
+For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by only a few eigenvectors that affect all correlation coefficients.
+On the other hand, a correlation matrix can be specified using the \emph{partial correlations} between pairs of the recorded neurons.
+The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.
+As such, partial correlations more closely reflect the direct causal effects between components of complex systems than correlations; partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
+These opportunities have not yet been explored in neurophysiological studies, as most studies have only considered the distributions of pairwise correlations \cite{Zohary:1994, Bair:2001, Smith:2008, Ecker:2010}.
 
-However, estimation of covariance matrices from large populations presents a number of numerical challenges.
+However, estimation of correlation matrices from large populations presents a number of numerical challenges.
 The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
-This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
+This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of the largest eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
 The \emph{sample correlation matrix} is an unbiased estimate of the true correlations but its many free parameters make it sensitive to sampling noise.
@@
 Improved estimation of the correlation matrix is beneficial in itself.
-For example, improved estimates can be used to optimize  decoding of the population activity.
+For example, improved estimates can be used to optimize  decoding of the population activity \cite{Friedman:1989}.
 But reduced estimation error is not the only benefit of regularization.
@@
 The necessity and benefits of regularization increase with the size of the recorded population.
-With the advent of  big neural data, the search for optimal regularization schemes will become increasingly relevant for all models of functional connectivity.
+With the advent of  big neural data, the search for optimal regularization schemes will become increasingly relevant in any model of population activity.
 Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in large scale neural data will entail the search for optimal regularization schemes that may involve combinations of heuristic rules and numerical techniques specially designed for each neural circuit.
@@
 Here we showed that among several models a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix for cortical microcircuits.
-This finding is valuable in itself: improved estimates of the noise covariance matrix for large data sets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Ecker:2011}
+This finding is valuable in itself: improved estimates of the noise covariance matrix for large datasets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Ecker:2011}
 
@@
 Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity it may accurately reflect average connectivity profiles.
-Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \ref{Hofer:2012, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
+Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \cite{Hofer:2012, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
 Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation.

### 2014-01-27 plos.tex
split analysis of structure in Results into two sections:

@@
 The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$.
+The problem of finding the optimal set of non-zero elements of the precision matrix is known as \emph{covariance selection} \cite{Dempster:1972}.
 The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{02}{Row 1, C}).
@@
 \paragraph{Simulation}
-We next demonstrat how the most efficient among different regularized estimators can reveal the structure of correlations.
+We next demonstrate how the most efficient among different regularized estimators can reveal the structure of correlations.
 We constructed four $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (\figref{02}{\,Row 2, A--D}, see Methods for details about how these matrices were generated).
@@
 
-\paragraph{Covariance estimation in neural data}
+\paragraph{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
 We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{03}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
@@
 The traces were then binned into 150 ms intervals for the estimation of the covariance matrix.
-The sample correlation coefficients were largely positive and low (0.01--0.03 in most images sites) (\figref{03}{F and G}).
+The sample correlation coefficients were largely positive and low (average across sites of 0.02 and all under 0.04) (\figref{03}{G} and \figref{06}{D}).
 
@@
 In these dense populations, direct interactions between cells are likely to influence the patterns of population activity.
-We therefore hypothesized that covariance matrix estimates that explicitly model the partial correlation between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would have a performance advantage.
+We therefore hypothesized that covariance matrix estimates that explicitly model the partial correlations between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would have a performance advantage.
 However, the observed neurons must also be strongly influenced by global activity fluctuations and by unobserved common inputs, to the advantage of estimators that explicitly model common fluctuations of the entire population: $C_{\sf factor}$ and $C_{\sf sparse+latent}$.
@@
 
-Furthermore, $C_{\sf sparse}$ outperformed all other estimators except $C_{\sf sparse+latent}$ (Fig.~S\ref{supp:01}) suggesting the greater importance of modeling interactions between pairs of neurons than the common fluctuations of activity across the entire population.
+Furthermore, $C_{\sf sparse}$ outperformed all the other estimators besides $C_{\sf sparse+latent}$ (Fig.~S\ref{supp:01}) suggesting the greater importance of modeling interactions between pairs of neurons than the common fluctuations of activity across the entire population.
 
@@
 
-\paragraph{Correlation structure and circuit architecture}
-Because of its superior estimation performance, $C_{\sf sparse+latent}$ likely provides a more physiologically relevant representation of the functional connectivity than the other estimators.
-The estimator segregates two sources of correlations:  The first source was the network of linear dependencies between pairs of neurons.
-The second source was the set of latent units reflecting common fluctuations across the entire population.
-These could be caused by  common inputs, cell assembly activation, network fluctuations or other forms of synchronized activations.
+\paragraph{Structure of $C_{\sf sparse+latent}$ estimates}
+Estimator $C_{\sf sparse+latent}$ separates two sources of correlations: a network of linear interactions between pairs of neurons and a set of latent units reflecting common fluctuations across the entire population.
+These could have multiple physiological causes such as common inputs, cell assembly activation, network fluctuations, or other forms of synchronized activations.
 
-We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site.
+We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:05} and Fig.~\ref{fig:06}).
 Although the regularized estimates were similar to the sample correlation matrix (\figref{05}{A and D}), the corresponding partial correlation matrices differed substantially (\figref{05}{B and E}).
-The partial correlation matrix of the regularized estimate was decomposed into a sparse component and a low-rank component (\figref{05}{C}).
-While sample correlations were mostly positive, the sparse partial correlations contained a high fraction of negative interactions.
-In the example site (Fig.~\ref{fig:03} and Fig.~\ref{fig:05}),  the sparse component had 82.2\% sparsity (or, conversely, 17.8\% connectivity) with average node degree of 52.8 (\figref{05}{G}).
-The estimate also inferred 15 latent units contributing to the low-rank component of the partial correlation matrix.
+The partial correlation matrix of the regularized estimate was the sum of a sparse component and a low-rank component (\figref{05}{C}).
+
+In the example site (Fig.~\ref{fig:05}), the sparse component had 82.2\% sparsity (or conversely, 17.8\% connectivity: $connectivity=1-sparsity$) with average node degree of 52.9 (\figref{05}{G}).
+The averge node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$.
+The low-rank component had rank 15, expressing 15 inferred latent units.
+The number of latent units and the average node degrees genearlly increased with population size, although these values varied widely between the datasets \figref{06}{A and B}.
+There was an inverse relationship between the number of latent units and the average node degree (\figref{06}{C}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
+This variability may be explained by differences in brain states and recording quality and needs further investigation.
+
+The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 31 sites were about 5 times lower than the average sample correlations (\figref{06}{D}).
+This signifies that correlations between neurons build up from large numbers of smaller partial correlations or interactions with other neurons.
+Furthermore, the avearge partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.41 whereas that of the average partial correlations was only 0.35.
+
+While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (43.3\%) of negative (\sq{inhibitory}) values (\figref{05}{F}).
+This fraction was fairly consistent across multiple sites but the fraction of negative interactions increased with the inferred connectivity \figref{06}{F}.
+This increase signifies that negative interactions are inferred only after a sufficient density of positive interactions have been uncovered.
 
@@
 In the example site with 17.8\% connectivity inferred by $C_{\sf sparse+latent}$, only 41.7\% of connections coincided with above-threshold sample correlations (\figref{05}{F, H, and I}).
-In particular, the thresholded sample correlations were predominantly positive whereas 43.3\% of the interactions obtained using $C_{\sf sparse+latent}$ were negative.
-Most of the negative interactions corresponded to low sample correlations (\figref{05}{F}) where high correlations were expected given the rest of the correlation matrix.
-
-The average partial correlations revealed by $C_{\sf sparse+latent}$ estimates at all 31 sites were about 5 times lower and less variable across sites than sample correlations (\figref{06}{D}): the coefficient of variation of the average sample correlations across all sites was 0.41 whereas that of the average partial correlations was only 0.35.
-The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (\figref{06}{A and B}).
-However, there was an inverse relationship between the number of latent units and the average node degree (\figref{06}{C}).
-Several sites, even with relatively large population sizes, had fairly few pairwise interactions and were dominated by latent units.
-These differences may be explained by differences in brain states and recording quality.
-For example neuropil contamination, motion and other factors may have degraded the quality of the data.
+In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{05}{F}) where high correlations were expected given the rest of the correlation matrix.
+This low overlap between the two types of graphs was consistent across across the datasets even though  their connectivity varied \figref{06}{E}.
 
@@
 
+\paragraph{Relationship of $C_{\sf sparse+latent}$ to orientation tuning and physical distances}
+
 We also examined the relationship of the correlation structure to the differences in orientation preference and to the physical distances between the cell pairs (Fig.\;\ref{fig:07}).
-The partial correlations of the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation (\figref{07}{A}) than sample correlations.
+The partial correlations of the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation (\figref{07}{A}) than the sample correlations.
 They also decreased more rapidly with physical distance, both laterally  (\figref{07}{B}) and in depth (\figref{07}{B}).
@@
     {\bf B.} Average correlations between pairs of neurons located at the same depth ($\pm$25$\mu$m) separated by lateral distances in the intervals of 0--25 $\mu$m, 25--75 $\mu$m, 75--150 $\mu$m, and 150+ $\mu$m.
-    {\bf C.} Average correlations between pairs of neurons displaced laterally by less than 25 $\mu$m separated in depth by distances in the intervals of 0-25 $\mu$m, 25--60 $\mu$m, and 60+ $\mu$m.
+    {\bf C.} Average correlations between pairs of neurons displaced laterally by less than 25 $\mu$m separated in depth by distances in the intervals of 0--25 $\mu$m, 25--60 $\mu$m, and 60+ $\mu$m.
     {\bf D--F.} Normalized connectivity of positive (green) and negative (dark red) interactions from the sparse component obtained from $C_{\sf sparse+latent}$.
@@
 
-Furthermore, the positive and negative interactions in the $C_{\sf sparse+latent}$ estimator were organized differently: The probability of positive interactions fell rapidly with differences in the preferred orientation (\figref{07}{D}), lateral displacements (\figref{07}{E}), and displacement in depth (\figref{07}{F}) while the negative interactions were much less selective (\figref{07}{D--F}).
+The positive and negative interactions in the $C_{\sf sparse+latent}$ estimator were organized differently: The density of the positive interactions fell rapidly with differences in the preferred orientation (\figref{07}{D}), lateral displacements (\figref{07}{E}), and displacement in depth (\figref{07}{F}) whereas the negative interactions were much less selective (\figref{07}{D--F}).
 

### 2014-01-27 plos.tex
Kreso's edits

--- 2014-01-27 plos.tex
@@
 With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.
-Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network.
+Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network \Kcomment{and the input?}.
 Therefore, conditionally independent neurons can be inferred to lack a direct causal influence.
@@
 For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.
-Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .
-Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
+Maximum entropy models are similarly constrained by \Kcomment{stimulus drive and} observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.
+\Kcomment{Assuming that the population response follows a} multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
 Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.
@@
 Little experimental evidence is available to answer this question.
-The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without examining their relation to anatomical implications.
+The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to \Kcomment{anatomy}.
 Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
-However, the topological properties of functional connectivity graphs vary with the method of their inference \cite{Zalesky:2012}.
+However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}.
 Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.
@@
 Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons.
-For this reason, the most statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}.
-In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the pairwise conditional dependencies.
-This may be the reason that partial correlations have not previously been used to describe  functional in cortical populations.
+For this reason, statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}.
+In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies.
+This may be the reason that partial correlations have not previously been used to describe  functional connectivity in cortical populations.
 
@@
 The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.
-Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
+Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
 These techniques may provide more incisive measurements of functional connectivity than classical electrophysiological recordings.
@@
 
-Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity, but an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.
-In our study, the analysis of functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.
+Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity.
+However, an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.
+In our study, the functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.
 
@@
 Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.
-The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
+\Kcomment{References for the previous statement?} The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
 As  techniques improved to allow for recording from larger neuronal populations, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.
-However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
+\Kcomment{The stationarity assumption seems a bit orthogonal to the rest discussed here.}  However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
 Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators.
-Regularization biases the solution toward a smaller subspace to counteract the effect of the sampling noise in the empirical data.
+Regularization biases the solution toward a small subspace to counteract the effect of  sampling noise in the empirical data.
 However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.
@@
 The results of model selection must be interpreted with caution.
-As we demonstrated by simulation, a wrong simple model can produce a better estimate than a more complex model with correct representation of dependencies.
+As we demonstrated by simulation, with limited data an incorrect, but simpler (lower dimension) model can produce a better estimate than a more complex model with correct representation of dependencies.
 Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the conclusion that it reveals a better representation of dependencies in the data.
-This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
+This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
 
@@
 Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity it may accurately reflect average connectivity profiles.
-Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \cite{Hofer:2012, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
+Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \cite{Hofer:2012, Ko:2013} \Kcomment{ref} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
 Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation.

### 2014-01-27 plos.tex
minor

@@
 \paragraph{Structure of $C_{\sf sparse+latent}$ estimates}
-Estimator $C_{\sf sparse+latent}$ separates two sources of correlations: a network of linear interactions between pairs of neurons and a set of latent units reflecting common fluctuations across the entire population.
+The estimator $C_{\sf sparse+latent}$ separates two sources of correlations: a network of linear interactions between pairs of neurons and a set of latent units reflecting common fluctuations across the entire population.
 These could have multiple physiological causes such as common inputs, cell assembly activation, network fluctuations, or other forms of synchronized activations.

### 2014-01-27 plos.tex
minor

--- 2014-01-27 plos.tex
@@
 \\
-Dimitri Yatsenko,$^{1,\ast}$, 
+Dimitri Yatsenko,$^{1}$, 
 Kre\v{s}imir Josi\'{c}$^{2}$,
@@
 R.~James Cotton$^{1}$,
-Andreas S.~Tolias$^{1,5}$
+Andreas S.~Tolias$^{1,5,\ast}$
 \\
@@
 
-$\ast$ E-mail: yatsenko@cns.bcm.edu
+$\ast$ E-mail: atolias@cns.bcm.edu
 \end{flushleft}
@@
 In the example site (Fig.~\ref{fig:05}), the sparse component had 82.2\% sparsity (or conversely, 17.8\% connectivity: $connectivity=1-sparsity$) with average node degree of 52.9 (\figref{05}{G}).
-The averge node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$.
+The averge node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$, where $p$ is the number of neurons.
 The low-rank component had rank 15, expressing 15 inferred latent units.
-The number of latent units and the average node degrees genearlly increased with population size, although these values varied widely between the datasets \figref{06}{A and B}.
+The number of latent units and the average node degrees generally increased with population size, although these values varied widely between the datasets (\figref{06}{A and B}).
 There was an inverse relationship between the number of latent units and the average node degree (\figref{06}{C}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
@@
 The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 31 sites were about 5 times lower than the average sample correlations (\figref{06}{D}).
-This signifies that correlations between neurons build up from large numbers of smaller partial correlations or interactions with other neurons.
-Furthermore, the avearge partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.41 whereas that of the average partial correlations was only 0.35.
+This suggests that correlations between neurons build up from large numbers of small interactions with other neurons.
+Furthermore, the avearge partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.41 whereas that of the average partial correlations was 0.35.
 
-While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (43.3\%) of negative (\sq{inhibitory}) values (\figref{05}{F}).
-This fraction was fairly consistent across multiple sites but the fraction of negative interactions increased with the inferred connectivity \figref{06}{F}.
-This increase signifies that negative interactions are inferred only after a sufficient density of positive interactions have been uncovered.
+While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (43.3\%) of negative values (\figref{05}{F}).
+This fraction was consistent across multiple sites but the fraction of negative interactions increased with the inferred connectivity \figref{06}{F}.
+This increase suggests that negative interactions are inferred only after a sufficient density of positive interactions have been uncovered.
 
@@
 
-We compared the graphs of interactions inferred by $C_{\sf sparse+latent}$ to the graph obtained by thresholding the sample correlations.
 Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013}.
-We set the threshold to match the connectivity of the $C_{\sf sparse+latent}$ graph.
+We therefore compared the interactions inferred by $C_{\sf sparse+latent}$ to those obtained by thresholding the sample correlations with the threshold set to match the connectivity of the $C_{\sf sparse+latent}$ graph.
 The networks revealed by the two methods differed substantially.
@@
 In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{05}{F}) where high correlations were expected given the rest of the correlation matrix.
-This low overlap between the two types of graphs was consistent across across the datasets even though  their connectivity varied \figref{06}{E}.
+This low overlap between the two types of graphs was consistent across the datasets \figref{06}{E}.
+t
 
@@
 With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.
-Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network \Kcomment{and the input?}.
+Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network and external input.
 Therefore, conditionally independent neurons can be inferred to lack a direct causal influence.
@@
 For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.
-Maximum entropy models are similarly constrained by \Kcomment{stimulus drive and} observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.
-\Kcomment{Assuming that the population response follows a} multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
+Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .
+Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
 Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.
@@
 Little experimental evidence is available to answer this question.
-The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to \Kcomment{anatomy}.
+The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without examining their relation to anatomical implications.
 Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
-However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}.
+However, the topological properties of functional connectivity graphs vary with the method of their inference \cite{Zalesky:2012}.
 Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.
@@
 Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons.
-For this reason, statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}.
-In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies.
-This may be the reason that partial correlations have not previously been used to describe  functional connectivity in cortical populations.
+For this reason, the most statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}.
+In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the pairwise conditional dependencies.
+This may be the reason that partial correlations have not previously been used to describe  functional in cortical populations.
 
@@
 The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.
-Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
+Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
 These techniques may provide more incisive measurements of functional connectivity than classical electrophysiological recordings.
@@
 
-Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity.
-However, an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.
-In our study, the functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.
+Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity, but an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.
+In our study, the analysis of functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.
 
@@
 Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.
-\Kcomment{References for the previous statement?} The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
+The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
 As  techniques improved to allow for recording from larger neuronal populations, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.
-\Kcomment{The stationarity assumption seems a bit orthogonal to the rest discussed here.}  However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
+However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
 Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators.
-Regularization biases the solution toward a small subspace to counteract the effect of  sampling noise in the empirical data.
+Regularization biases the solution toward a smaller subspace to counteract the effect of the sampling noise in the empirical data.
 However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.
@@
 The results of model selection must be interpreted with caution.
-As we demonstrated by simulation, with limited data an incorrect, but simpler (lower dimension) model can produce a better estimate than a more complex model with correct representation of dependencies.
+As we demonstrated by simulation, a wrong simple model can produce a better estimate than a more complex model with correct representation of dependencies.
 Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the conclusion that it reveals a better representation of dependencies in the data.
-This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
+This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
 
@@
 Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity it may accurately reflect average connectivity profiles.
-Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \cite{Hofer:2012, Ko:2013} \Kcomment{ref} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
+Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \cite{Hofer:2011, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
 Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation.
@@
 slow-wave sleep, attention) and stimulus conditions (e.g.
-certain types of stimuli may engage feedback connections)
+certain types of stimuli may engage feedback connections).
 
@@
 \section*{Acknowledgments}
-We thank Genevera Allen for a helpful discussion, and Eftychios Pnevmatikakis for helpful suggestions and feedback.
-
+We thank Genevera Allen for a helpful discussion, and Eftychios Pnevmatikakis for helpful suggestions and feedback on the manuscript.
+This work was supported by grants NEI R01-EY018847, NEI P30-EY002520-33, NEI T32-EY07001-37, NIDA R01DA028525, the NIH-Pioneer award DP1-OD008301 to AST, the McKnight Scholar Award to AST; the Arnold & Beckman Foundation Young Investigator Award to AST; and NSF grants DMS-0817649, DMS-1122094 to KJ.
 % The bibtex filename

### 2014-01-27 plos.tex

@@
     {\bf D.} The average sample correlations  vs.~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
-    {\bf E.} Percentage overlap between the highest sample correlations and the  interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie 1-sparsity) of its sparse component.
+    {\bf E.} Percentage overlap between the highest sample correlations and the  interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie $1-sparsity$) of its sparse component.
     {\bf F.} The percentage of negative interactions vs.~connectivity.
@@
 For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.
-Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .
-Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
+Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.
+Assuming that the population response follows a multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
 Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.
@@
 Little experimental evidence is available to answer this question.
-The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without examining their relation to anatomical implications.
+The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to anatomy.
 Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
-However, the topological properties of functional connectivity graphs vary with the method of their inference \cite{Zalesky:2012}.
+However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}.
 Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.
@@
 Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators.
-Regularization biases the solution toward a smaller subspace to counteract the effect of the sampling noise in the empirical data.
+Regularization biases the solution toward a small subspace to counteract the effect of  sampling noise in the empirical data.
 However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.
@@
 \begin{equation}
-P_{\sf sparse} = (S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
+P_{\sf sparse} = -(S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
 \end{equation}

### 2014-01-27 plos.tex
minor

@@
 We thank Genevera Allen for a helpful discussion, and Eftychios Pnevmatikakis for helpful suggestions and feedback on the manuscript.
-This work was supported by grants NEI R01-EY018847, NEI P30-EY002520-33, NEI T32-EY07001-37, NIDA R01DA028525, the NIH-Pioneer award DP1-OD008301 to AST, the McKnight Scholar Award to AST; the Arnold & Beckman Foundation Young Investigator Award to AST; and NSF grants DMS-0817649, DMS-1122094 to KJ.
+This work was supported by grants NEI R01-EY018847, NEI P30-EY002520-33, NEI T32-EY07001-37, NIDA R01DA028525, the NIH-Pioneer award DP1-OD008301 to AST, the McKnight Scholar Award to AST; the Arnold \& Beckman Foundation Young Investigator Award to AST; and NSF grants DMS-0817649, DMS-1122094 to KJ.
 % The bibtex filename

### 2014-01-28 plos.tex
corrected the captions in the supplementary figures.

@@
 Indeed, in cross-validation tests, the covariance matrix estimator with this structure consistently outperformed other regularized estimators.
-The sparse component of the estimate expressed a graph of interactions organized with respect to the physical distances and orientation tuning properties of cells: The density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.
+The sparse component of the estimate expressed a graph of interactions organized with respect to the physical distances and orientation tuning properties of cells: The density of positive \sq{excitatory} interactions decreased rapidly with geometric distances and with differences in orientation preference whereas the negative \sq{inhibitory} interactions were less selective.
 Because of its superior performance, this ``sparse + latent'' estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.
@@
 Neural correlations do not come with ready or unambiguous mechanistic interpretations.
-They can arise from monosynaptic or polysynaptic interactions, common or correlated inputs, oscillations, top-down modulation, and background network fluctuations, or a combination of these and other mechanisms \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.
+They can arise from monosynaptic or polysynaptic interactions, common or correlated inputs, oscillations, top-down modulation, and background network fluctuations, and other mechanisms \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.
 But multineuronal recordings do provide more information than an equivalent number of separately recorded pairs of cells.
@@
 The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.
-As such, partial correlations more closely reflect the direct causal effects between components of complex systems than correlations; partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
+When a large fraction of the interacting variables are observed, partial correlations more closely reflect the direct causal effects between components of complex systems than correlations.
+Partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
 These opportunities have not yet been explored in neurophysiological studies, as most studies have only considered the distributions of pairwise correlations \cite{Zohary:1994, Bair:2001, Smith:2008, Ecker:2010}.
@@
 Estimation can be improved through \emph{regularization},  the technique of deliberately imposing a structure on an estimate in order to reduce its estimation error \cite{Schafer:2005, Bickel:2006}.
-To \sq{impose a structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation  with fewer free parameters, which is often referred to as the \emph{target estimate}.
-The optimal target estimate and the optimal amount of shrinkage can be obtained from the data sample either analytically \cite{Ledoit:2003, Ledoit:2004, Schafer:2005},  or by cross-validation \cite{Friedman:1989}.
+To \sq{impose a structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation  with fewer free parameters, the \emph{target estimate}.
+The optimal target estimate and the optimal amount of shrinkage can be obtained from the data sample either analytically \cite{Ledoit:2003, Ledoit:2004, Schafer:2005}  or by cross-validation \cite{Friedman:1989}.
 An estimator that produces estimates that are, on average, closer to the truth for a given sample size is said to be more \emph{efficient} than other estimators.
@@
 The necessity and benefits of regularization increase with the size of the recorded population.
-With the advent of  big neural data, the search for optimal regularization schemes will become increasingly relevant in any model of population activity.
+With the advent of  big neural data \cite{Alivasatos:2013}, the search for optimal regularization schemes will become increasingly relevant in any model of population activity.
 Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in large scale neural data will entail the search for optimal regularization schemes that may involve combinations of heuristic rules and numerical techniques specially designed for each neural circuit.
@@
 \item[$C_{\sf 0}$] -- sample covariance matrix, the unbiased, unregularized estimator.
-\item[$C_{\sf diag}$] -- linear shrinkage of correlations toward zero.
+\item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \ie toward a diagonal covariance matrix.
 \item[$C_{\sf factor}$] -- linear shrinkage of the covariances toward a low-rank matrix, representing inputs from unobserved factors (latent units).
@@
 We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.
-This estimator revealed a sparse network of partial correlations, which we call \sq{interactions} here, between the observed neurons and inferred several latent units exerting linear effects on the observed neurons.
+This estimator revealed a sparse network of partial correlations, which we call \sq{interactions} here, between the observed neurons; it also inferred several latent units exerting linear effects on the observed neurons.
 We analyzed these networks of partial correlations and found the following: Whereas noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.
-Moreover, we found that these interactions exhibited a stronger relationship on the physical distances and differences in preferred orientations than noise correlations.
+Moreover, we found that these interactions exhibited a stronger relationship to the physical distances and to the differences in preferred orientations than noise correlations.
 In contrast, the inferred negative interactions were less selective.
@@
 For noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.
-In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than the sample correlation matrix, $C_{\sf 0}$.
+In this study, we estimate the true mean, $\mu$, using the sample mean, $\bar x$, but seek a better estimate of the covariance matrix than the sample correlation matrix, $C_{\sf 0}$.
 
@@
 \end{equation}
-where $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. 
+where $\circ$ denotes the entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. 
 Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$:
@@
 Recorded neurons are represented by green spheres and latent units by light-shaded spheres.
-Edges represent non-zero partial correlations.
-        {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear associations. 
+Edges represent non-zero partial correlations, \ie \sq{interactions}.
+        {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear dependencies. 
         {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units.
@@
 \\
-{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros.
+{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros.
 \\
-{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices from row 2.
+{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices shown in Row 2.
 The structure of the sample correlation matrix is difficult to discern.
@@
 The error bars indicate the standard error of the mean based on 30 samples.
-Estimators with structure that matches the model used to generate the data converged to zero faster than the other estimators.
+Estimators with structure that matches the true model converged to zero faster than the other estimators.
 Some exceptions occur when the sample is too small to reveal the structure in the data.
-In these cases estimators with simpler structures can be more efficient.
+In these cases estimators with simpler structures can be more efficient than estimators with true structure.
 \\
@@
 has two hyperparameters: the rank of $L$ (\ie the number of latent factors) and the shrinkage intensity $\lambda$.
-The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{Row 1, B}).
+The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (\figref{02}{Row 1, B}).
 
-Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix whose inverse is sparse: 
+Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by the inverse of a sparse matrix $S$: 
 \begin{equation}\label{eq:c-sparse}
@@
 The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{02}{Row 1, C}).
-We will sometimes refer to sparse partial correlations between neurons simply  as \sq{interactions}.
 
@@
 
+We refer to the sparse partial correlations in estimators $C_{\sf sparse}$ and $C_{\sf sparse+latent}$ as as \sq{interactions}.
+
 \paragraph{Simulation}
@@
 
-The evaluation of a covariance matrix estimator, $C,$ is performed with respect to a \emph{loss function} $\loss{C,\Sigma}$ to quantify the discrepancy between the estimate and the truth, $\Sigma$.
+The evaluation of any covariance matrix estimator, $C$, is performed with respect to a \emph{loss function} $\loss{C,\Sigma}$ to quantify its discrepancy from truth, $\Sigma$.
 The loss function must attain its minimum when $C=\Sigma$.
@@
 
-The choice of normal loss function is motivated by mathematical convenience.
+The choice of the normal loss function is motivated by mathematical convenience.
 We expect that the main conclusions of our study will not change qualitatively with other well behaved loss functions, such as Stein's entropy loss or quadratic loss \cite{James:1961, Fan:2008, Ledoit:2004, Schafer:2005}.
@@
 
-We drew 30 independent samples of size $n=250$, 500, 1000, 2000, and 4000 from model and computed the excess loss for each of the five estimators.
+We drew 30 independent samples of sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the excess loss for each of the five estimators.
 The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.
@@
 There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$, often outperformed the estimators with the matching structure (\figref{02}{\,Row 5, C and D}).
-Second, $C_{\sf sparse+latent}$ often performed as well as $C_{\sf sparse}$ on data from models without latent units because.
-However, in theses cases it reliably inferred the correct number of latent units which was zero.
+Second, $C_{\sf sparse+latent}$ often performed as well as $C_{\sf sparse}$ on data from models without latent units because it reliably inferred the correct number of latent units: zero.
 
@@
 
-\begin{figure}[!ht]
+\begin{FPfigure}
 \begin{center}
@@
 \caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators with respect to multivariate normal cross-validation loss.}
+\\{\bf A--D.} These panels are identical to those in \figref{04}{A--D}.
+\\{\bf E.} $C_{\sf sparse}$ outperforms $C_{\sf 0}$: median improvement 0.15 nats/neuron, $p=3.9\times 10^{-5}$.
+\\{\bf F.} $C_{\sf sparse}$ performs similarly to $C_{\sf diag}$: median improvement $1.0\times10^{-3}$ nats/neuron, $p=0.13$.
+\\{\bf G.} $C_{\sf sparse}$ outperforms $C_{\sf factor}$: median improvement $0.15$ nats/neuron, $p=1.2\times10^{-4}$.
+\\{\bf H.} $C_{\sf factor}$ outperforms $C_{\sf 0}$: median improvement $0.029$ nats/neuron, $p=2.7\times10^{-4}$.
+\\{\bf I.} $C_{\sf factor}$ performs worse than $C_{\sf diag}$: median improvement $-0.16$ nats/neuron, $p=1.6\times10^{-4}$.
+\\{\bf J.} $C_{\sf diag}$ outperforms $C_{\sf 0}$: median improvement $0.14$ nats/neuron, $p=2.7\times10^{-5}$.
 }
 \label{supp:01}
-\end{figure}
+\end{FPfigure}
 
@@
 \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-{\caption{{\bf The sparse+latent estimator outperforms the other estimators under the quadratic loss (Eq.~\ref{eq:quadratic}) just as it did under the normal loss (Eq.~\ref{eq:loss}) used in Fig.~\ref{fig:04}}.
+{\caption{{\bf The sparse+latent estimator outperforms the other estimators under the quadratic loss (Eq.~\ref{eq:quadratic}) just as it did under the normal loss (Eq.~\ref{eq:loss}) used in Fig.~\ref{fig:04}.}
+    Histograms of the relative cross-validation loss (Eq.~\ref{eq:rel-cv-loss}) of estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ using the quadratic loss function.
+    The histograms were obtained from 31 imaged sites in 15 mice. 
+    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
+    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}.
+    The $p$-values are for the two-sided Wilcoxon's signed rank test.
+    \\{\bf A.} $C_{\sf sparse+latent}$ outperforms $C_{\sf 0}$: median $\Delta\ell_{C_{\sf 0},C_{\sf sparse+latent}}$ = 0.35, $p=1.7\times 10^{-6}$.
+    \\{\bf B.} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median $\Delta\ell_{C_{\sf diag},C_{\sf sparse+latent}}$ = $2.5\times 10^{-3}$, $p=1.0\times 10^{-4}$.
+    \\{\bf C.} $C_{\sf sparse+latent}$ outperforms $C_{\sf factor}$: median $\Delta\ell_{C_{\sf factor},C_{\sf sparse+latent}}$ = 0.27, $p=1.2\times 10^{-6}$.
+    \\{\bf D.} $C_{\sf sparse+latent}$ outperforms $C_{\sf sparse}$: median $\Delta\ell_{C_{\sf sparse},C_{\sf sparse+latent}}$ = $3.7\times 10^{-4}$, $p=2.2\times 10^{-3}$.
 }

### 2014-01-29 plos.tex
minor

--- 2014-01-27 plos.tex
@@
 This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
-The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} to produce (\figref{03}{C and D}).
+The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}).
 Only cells that produced detectable calcium activity were included in the analysis (For details, see Methods).
@@
 The traces were then binned into 150 ms intervals for the estimation of the covariance matrix.
-The sample correlation coefficients were largely positive and low (average across sites of 0.02 and all under 0.04) (\figref{03}{G} and \figref{06}{D}).
+The sample correlation coefficients were largely positive and low (average across sites of 0.02 and under 0.04 in any site) (\figref{03}{G} and \figref{06}{D}).
 
@@
     {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site.
+Only spiking cells are shown.
 The cells' colors indicate their orientation preferences.
@@
 In these dense populations, direct interactions between cells are likely to influence the patterns of population activity.
-We therefore hypothesized that covariance matrix estimates that explicitly model the partial correlations between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would have a performance advantage.
+We therefore hypothesized that covariance matrix estimators that explicitly modeled the partial correlations between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would have a performance advantage.
 However, the observed neurons must also be strongly influenced by global activity fluctuations and by unobserved common inputs, to the advantage of estimators that explicitly model common fluctuations of the entire population: $C_{\sf factor}$ and $C_{\sf sparse+latent}$.

### 2014-01-29 plos.tex
Rev44: submitted version

@@
 The estimator $C_{\sf sparse+latent}$ separates two sources of correlations: a network of linear interactions between pairs of neurons and a set of latent units reflecting common fluctuations across the entire population.
-These could have multiple physiological causes such as common inputs, cell assembly activation, network fluctuations, or other forms of synchronized activations.
 
@@
 The averge node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$, where $p$ is the number of neurons.
-The low-rank component had rank 15, expressing 15 inferred latent units.
+The low-rank component had rank 15, denoting 15 inferred latent units.
 The number of latent units and the average node degrees generally increased with population size, although these values varied widely between the datasets (\figref{06}{A and B}).
@@
 
-While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (43.3\%) of negative values (\figref{05}{F}).
+While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (43.3\% in the example site) of negative values (\figref{05}{F}).
 This fraction was consistent across multiple sites but the fraction of negative interactions increased with the inferred connectivity \figref{06}{F}.
-This increase suggests that negative interactions are inferred only after a sufficient density of positive interactions have been uncovered.
+This increase suggests that negative interactions are inferred only after a sufficient density of positive interactions has been uncovered.
 
@@
 In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{05}{F}) where high correlations were expected given the rest of the correlation matrix.
-This low overlap between the two types of graphs was consistent across the datasets \figref{06}{E}.
-t
+This low overlap between the two types of graphs was consistent across the datasets (\figref{06}{E}).
 
 \begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-    {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites} Each point represents an imaged site with its color indicating the population size as shown in panels A and B.
-The example site from Figures \ref{fig:03} and \ref{fig:05} is circled in blue.
+    {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites.} 
+    Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:03} and \ref{fig:05} is circled in blue. 
 \\
-    {\bf A.} The number of inferred latent units vs.~population size in each imaged site.
-    {\bf B.} The average node degree for sparse partial correlations vs.~population size at each imaged site.
+    {\bf A.} The number of inferred latent units vs.~population size.
+    {\bf B.} The average node degree for sparse partial correlations vs.~population size.
     {\bf C.} The average node degree in the sparse component in each site vs.~the number of latent units described by the low-rank component of the estimate.
     {\bf D.} The average sample correlations  vs.~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
-    {\bf E.} Percentage overlap between the highest sample correlations and the  interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie $1-sparsity$) of its sparse component.
-    {\bf F.} The percentage of negative interactions vs.~connectivity.
+    {\bf E.} Percent overlap between the graph of highest sample correlations and the graph of interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie $1-sparsity$) of its sparse component.
+    {\bf F.} The percentage of negative interactions vs.~connectivity in the $C_{\sf sparse+latent}$ estimates.
 }
@@
 For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.
-Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.
+Maximum entropy models constrained by observed pairwise correlations also model pairwise coupling between cells \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.
 Assuming that the population response follows a multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
@@
 However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}.
-Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.
+Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question and we did not attempt them here.
 
@@
 In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies.
-This may be the reason that partial correlations have not previously been used to describe  functional connectivity in cortical populations.
+This may be the reason that partial correlations have not previously been used to describe the functional connectivity in cortical populations until this study.
 
@@
 Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we \emph{a priori} favored solutions with sparse partial correlations.
-The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution.
+The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution, where the precision matrix specifies the interactions between cell pairs.
 Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
@@
 The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
-As  techniques improved to allow for recording from larger neuronal populations, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.
-However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
-Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators.
+As techniques have improved to allow recording from larger neuronal populations, experimental neuroscientists have addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.
+However, ambitious projects, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
+Simply increasing recording duration will not be sufficient, and the problem must be addressed by using regularized estimators.
 Regularization biases the solution toward a small subspace to counteract the effect of  sampling noise in the empirical data.
-However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.
+However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hinders interpretation.
 
 Several strategies have been developed to limit the model space in order to improve the quality of the estimate.
-For example, \cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in retina.
+For example, Ganmor et al.
+\cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in retina.
 Generalized linear models typically employ $L_1$ penalty terms to constrain the solution space and to effectively reduce the dimensionality of the solution \cite{Pillow:2008}.
@@
 The results of model selection must be interpreted with caution.
-As we demonstrated by simulation, with limited data an incorrect, but simpler (lower dimension) model can produce a better estimate than a more complex model with correct representation of dependencies.
-Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the conclusion that it reveals a better representation of dependencies in the data.
-This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.
+As we demonstrated by simulation, with limited data, an incorrect but more constrained model can produce a better estimate than a more complex model with correct representation of dependencies.
+Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not necessarily support the conclusion that it reveals a better representation of dependencies in the data.
+This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary low-dimensional target is guaranteed to produce \emph{some} improvement over the best estimate with many degrees of freedom.
 
@@
 Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.
-Importantly, the inferred functional interactions were substantially different from those measured by considering the network of the most significant sample correlations \figref{05}{F, H, and I}.
-For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}) that may reflect inhibitory circuitry.
-Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity it may accurately reflect average connectivity profiles.
-Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \cite{Hofer:2011, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
+Importantly, the inferred functional interactions were substantially different from the network of the most significant sample correlations \figref{05}{F, H, and I}.
+For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}) and may reflect inhibitory circuitry.
+Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity, it may accurately reflect average connectivity profiles.
+Future experiments utilizing molecular markers for different cell types and follow-up multiple whole-cell \emph{in vitro} experiments \cite{Hofer:2011, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
 Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation.
@@
 All imaging was performed using the 3D-RAMP two-photon microscope as described in \cite{Cotton:2013}.
-First, a 3D stack was acquired and cells manually segmented.
+First, a 3D stack was acquired and cells were manually segmented.
 Then calcium signal were collected by sampling in the center of each cell at rates of 100 Hz or higher, depending on the number of cells.
@@
 \paragraph{Visual stimulus}
-Full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s.
+The visual stimulus consisted of full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s.
 Two sets of stimuli were presented for each imaging site: the first to map directional tuning and the second to estimate noise correlations.
-Directional tuning was map using a pseudo-random sequence of drifting gratings at sixteen equally spaced directional of motion changing at 2 Hz for 3 min without blanks.
-The data for covariance estimation were collected during presentations of full-field drifting gratings with the same parameters as those used in directional tuning except only two directions (in 9 datasets) or five directions (in 22 datasets) were used and the presentations lasted 1 second and, separated by 1-second blanks.
-Each stimulus condition was presented at least 180 times.
+Directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen equally spaced directions of motion, 500 ms per direction, for 3 min without blanks.
+The data for covariance estimation were collected during presentations of full-field drifting gratings with the same parameters as those used in directional tuning except only two directions (in 9 datasets) or five directions (in 22 datasets) were used and the presentations lasted 1 second and were separated by 1-second blanks.
+Each stimulus condition was presented between 100 and 300 times.
 \paragraph{Data processing}
@@
 
-Orientation tuning was computed by fitting the mean firing rates in response to gratings of directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
+Orientation tuning was computed by fitting the mean firing rates in response to gratings drifting in directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
 The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times.
 The $p$-value of the fit was computed as the fraction of the permuted datasets for which the $R^2$ value of the tuning function fit exceeded that of the real data.
-Cells were considered tuned for $p$-values below 0.05.
+Cells were considered tuned with $p$-values below 0.05.
 
 For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 14 and 27 minutes (mean 22 minutes).
-Spike trains that did not have substantial spiking activity (those whose variance was less than 2\% of the $90^{th}$ percentile in each site), presumably from glial cells or inactive neurons, were exculded from the analysis.
+Cells that did not have substantial spiking activity (those whose variance was less than 2\% of the $90^{th}$ percentile in each site) were exculded from the analysis.
 
 \paragraph{Cross-validation}
-To compare the performance of the estimator against each other, we used conventional 10-fold cross validation.
+To compare the performance of the estimators against each other, we used conventional 10-fold cross validation.
 Briefly, each recording was split into 30 blocks of equal duration.
 The 30 blocks were then grouped randomly into 10 datasets with 3 blocks in each.
-This splitting of data was chosen to achieve sufficient independence between the 10 datasets for cross-validaton purposes.
-Splitting the data into smaller blocks would increase their dependence due to temporal effects.
+This splitting of the data was chosen to achieve a high degree of independence between the 10 datasets for cross-validaton purposes.
+Splitting the data into smaller blocks could increase dependencies between them due to temporal correlations thereby violating the assumptions of cross-validation.
 Then, each dataset was used as the testing dataset with the rest of the data used for estimating the covariance matrix.
 
-Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}.
-The outer loop evaluated the performance of the estimators with optimal values of the hyperparameters.
-The optimization of the hyperparameters was performed within the inner loop in two phases: random search to find a good starting point and pattern search to find the global minimum.
+Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}:  The outer loop evaluated the performance of the estimators with the optimal hyperparameter values estimated within the inner loop.
+Hyperparameters were optimized by a two-phase search algorithm: random search to find a good starting point followed by pattern search to find the global minimum.
 The inner cross-validation loop subdivided the training dataset from the outer loop to perform 10-fold cross-validation in order to evaluate each choice of the hyperparameter values.
@@
 
-When cross-validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.
+When the cross-validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.
 This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{02}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:05}--\ref{fig:07}).
-\paragraph{Covariance Estimation}
+\paragraph{Covariance estimation}
 Within the inner loop of cross-validation, covariance matrix estimation was performed with fixed hyperparameter values provided by the search algorithm.
@@
 Then the diagonal matrix $D$ was used as the target of covariance shrinkage (Eq.~\ref{eq:c-diag}) to produce the final regularized estimate.
-
 The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean squared error whereas we optimized them with respect to the normal loss function (Eq.~\ref{eq:loss}).
@@
 Estimator $C_{\sf factor}$ used two hyperparameters: the number of latent factors $d$ and the shrinkage intensity $\lambda \in [0, 1]$.  
-The factor matrix $L$ of rank $d$ and the diagonal matrix of individual variances $\Psi$ were computed by solving the minimization problem
+The matrix $L$ of rank $d$ and the diagonal matrix of individual variances $\Psi$ were computed by solving the minimization problem
 \begin{equation}
@@
 \end{equation}
-where $Z^\sharp$ denotes the set of positive-definite matrices with zeros in all entries where entries of $Z$ equal zero.
+where $Z^\sharp$ denotes the set of positive-definite matrices with the same zero structure as $Z$.
 This step was also solved by ADMM.
@@
 The partial correlation matrix with both sparse and low-rank components is computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}; it includes the effects of interactions between the visible and latent units.
-This estimate of the partial correlations was used in analyses of average partial correlations (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).
-The sparse component $S$ was used separately in analyses of the connectivity graphs (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component alone without the low-rank component:
+This estimate of the partial correlations was used in analyses of average partial correlations (\figref{05}{B and E} and \figref{06}{D}, and \figref{07}{A--C}).
+The sparse component $S$ was used separately in analyses of the connectivity graphs (\figref{05}{F, G, H} and \figref{06}{B, C, E, and F} and \figref{07}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component alone without the low-rank component:
 \begin{equation}
@@
 For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units.
-The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{02}{\,Row 2}.
-Samples were then generated by multivariate normal distributions with the respective true covariance matrices to be estimated by each of the estimators.
+The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{02}{\,Row 2}).
+Samples were then drawn from multivariate normal distributions with the respective true covariance matrices to be estimated by each of the estimators.
 

### 2014-01-29 plos.tex
modified based on Editor's request: Ethics statement

--- 2014-01-29 plos.tex
@@
 The necessity and benefits of regularization increase with the size of the recorded population.
-With the advent of  big neural data \cite{Alivasatos:2013}, the search for optimal regularization schemes will become increasingly relevant in any model of population activity.
+With the advent of  big neural data \cite{Alivisatos:2013}, the search for optimal regularization schemes will become increasingly relevant in any model of population activity.
 Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in large scale neural data will entail the search for optimal regularization schemes that may involve combinations of heuristic rules and numerical techniques specially designed for each neural circuit.
@@
 This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
-The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}).
+The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} to produce (\figref{03}{C and D}).
 Only cells that produced detectable calcium activity were included in the analysis (For details, see Methods).
@@
 The traces were then binned into 150 ms intervals for the estimation of the covariance matrix.
-The sample correlation coefficients were largely positive and low (average across sites of 0.02 and under 0.04 in any site) (\figref{03}{G} and \figref{06}{D}).
+The sample correlation coefficients were largely positive and low (average across sites of 0.02 and all under 0.04) (\figref{03}{G} and \figref{06}{D}).
 
@@
     {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site.
-Only spiking cells are shown.
 The cells' colors indicate their orientation preferences.
@@
 In these dense populations, direct interactions between cells are likely to influence the patterns of population activity.
-We therefore hypothesized that covariance matrix estimators that explicitly modeled the partial correlations between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would have a performance advantage.
+We therefore hypothesized that covariance matrix estimates that explicitly model the partial correlations between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would have a performance advantage.
 However, the observed neurons must also be strongly influenced by global activity fluctuations and by unobserved common inputs, to the advantage of estimators that explicitly model common fluctuations of the entire population: $C_{\sf factor}$ and $C_{\sf sparse+latent}$.
@@
 % authors may use "Analysis" 
+\paragraph{Ethics statement}
+All procedures were conducted in accordance with the ethical guidelines of the National Institutes of Health and were approved by the Baylor College of Medicine IACUC.
+
 \paragraph{Surgery and two-photon imaging}
-All procedures were conducted in accordance with the ethical guidelines of the National Institutes of Health and were approved by the Baylor College of Medicine IACUC.
 The surgical procedures and data acquisition were performed as descrbed in \cite{Cotton:2013}.

### 2014-04-01 plos.tex
fixed typos but not made any other changes suggested by reviewers.

@@
 %\doublespacing
+
+\usepackage{soul}
 
@@
 
-The target of estimator $C_{\sf factor}$ is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix with low rank and $\Psi$ is the diagonal matrix of independent variances.
+The target of estimator $C_{\sf factor}$ is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive \hl{semidefinite} matrix with low rank and $\Psi$ is the diagonal matrix of independent variances.
 The estimator,
@@
 
-We refer to the sparse partial correlations in estimators $C_{\sf sparse}$ and $C_{\sf sparse+latent}$ as as \sq{interactions}.
+We refer to the sparse partial correlations in estimators $C_{\sf sparse}$ and $C_{\sf sparse+latent}$ as \sq{interactions}.
 
@@
 The traces were then binned into 150 ms intervals for the estimation of the covariance matrix.
-The sample correlation coefficients were largely positive and low (average across sites of 0.02 and under 0.04 in any site) (\figref{03}{G} and \figref{06}{D}).
+The sample correlation coefficients were largely positive and low (average across sites of 0.02 and \hl{under 0.04 in all sites}) (\figref{03}{G} and \figref{06}{D}).
 
@@
 In the example site (Fig.~\ref{fig:05}), the sparse component had 82.2\% sparsity (or conversely, 17.8\% connectivity: $connectivity=1-sparsity$) with average node degree of 52.9 (\figref{05}{G}).
-The averge node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$, where $p$ is the number of neurons.
+The average node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$, where $p$ is the number of neurons.
 The low-rank component had rank 15, denoting 15 inferred latent units.
@@
 This suggests that correlations between neurons build up from large numbers of small interactions with other neurons.
-Furthermore, the avearge partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.41 whereas that of the average partial correlations was 0.35.
+Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.41 whereas that of the average partial correlations was 0.35.
 
@@
 
-Estimator $C_{\sf sparse+latent}$ has two hyperparameters: the number of latent units $d$ and the sparsity level $\lambda$.
+Estimator $C_{\sf sparse+latent}$ has two hyperparameters: the number of latent units $d$ and \hl{hyperparameter $\lambda$ regulating the sparsity level}.
 $C_{\sf sparse+latent}$ estimates a larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.

### 2014-05-03 plos.tex
Response to reviewers.

@@
 Functional connectivity reflects local synaptic connections, shared inputs from other regions, and endogenous network activity.
-Although functional connectivity is a phenomenological description with no simple mechanistic interpretation, it can be used to generate hypotheses about the anatomical architecture of the neural circuit and test hypotheses about the processing of information at the population level.
+Although functional connectivity is a phenomenological description without a strict mechanistic interpretation, it can be used to generate hypotheses about the anatomical architecture of the neural circuit and test hypotheses about the processing of information at the population level.
 
@@
 \begin{description}
-\item[$C_{\sf 0}$] -- sample covariance matrix, the unbiased, unregularized estimator.
+\item[$C_{\sf sample}$] -- sample covariance matrix, the unbiased, unregularized estimator.
 \item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \ie toward a diagonal covariance matrix.
@@
 where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.
-The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf 0}$ from the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
+The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf sample}$ from the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
 \begin{equation}\label{eq:sample}
-    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
+    C_{\sf sample} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
 \end{equation}
@@
 For noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.
-In this study, we estimate the true mean, $\mu$, using the sample mean, $\bar x$, but seek a better estimate of the covariance matrix than the sample correlation matrix, $C_{\sf 0}$.
+In this study, we estimate the true mean, $\mu$, using the sample mean, $\bar x$, but seek a better estimate of the covariance matrix than the sample correlation matrix, $C_{\sf sample}$.
 
@@
 
+To see that this is the case, partition the covariance matrix 
+\begin{equation}
+C = \begin{pmatrix}
+C_{11} & C_{12} \\
+C_{12}^\T & C_{22}
+\end{pmatrix}
+\end{equation}
+so that $C_{11}$ is the $2\times 2$ covariance matrix of two cells of interest. Such partitioning can be done for any pair of cells by changing their order. 
+Then the covariance matrix of the conditional distribution of these two cells is \cite{Anderson:2003}:
+\begin{equation}
+\hat C_{11} = C_{11}-C_{12}C^{-1}C_{12}^\T 
+\end{equation}
+The inverse of the conditional covariance matrix can be found in the corresponding partition of the precision matrix $C^{-1}$:
+\begin{equation}
+C^{-1}=\begin{pmatrix}
+\hat C_{11}^{-1} &   -\hat C_{11}^{-1}C_{12}C_{22}^{-1} \\
+-C_{22}^{-1}C_{12}^\T\hat C_{11}^{-1}  &  C_{22}^{-1}+C_{22}^{-1}C_{12}^\T\hat C_{11}^{-1}C_{12}C_{22}^{-1}
+\end{pmatrix}
+\end{equation}
+The inverse of a $2\times 2$ correlation matrix reverses the sign of its off-diagonal elements:
+\begin{equation}
+\begin{pmatrix}
+1 & \rho \\
+\rho & 1
+\end{pmatrix}^{-1}
+\propto 
+\begin{pmatrix}
+-1 & \rho \\
+\rho & -1 
+\end{pmatrix}
+\end{equation}
+Therefore, scaling $C^{-1}$ to have 1s on the diagonal yields the partial correlation in each of its $2\times 2$ partitions, resulting in a matrix of partial pairwise correlations.
+
 We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
-In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{02}{\,Row 1}).
+In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{1}{\,Row 1}).
 
@@
     }
-    \label{fig:02}
+    \label{fig:1}
 \end{FPfigure}
@@
 The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons' variances.
-Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
+Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf sample}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
 \begin{equation}\label{eq:c-diag}
-C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
+C_{\sf diag} = (1-\lambda) C_{\sf sample} + \lambda D
 \end{equation}
-The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (\figref{02}{Row 1, A}).
+The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (\figref{1}{Row 1, A}).
 If sample correlations arise from spurious effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.
 
-The target of estimator $C_{\sf factor}$ is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive \hl{semidefinite} matrix with low rank and $\Psi$ is the diagonal matrix of independent variances.
+The target of estimator $C_{\sf factor}$ is the factor model $L + D$, where $L$ is the $p\times p$ positive \hl{semidefinite} matrix with low rank and $D$ is the diagonal matrix of independent variances.
 The estimator,
 \begin{equation}\label{eq:c-factor}
-C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (L + \Psi),
+C_{\sf factor} = (1-\lambda) C_{\sf sample} + \lambda (L + D),
 \end{equation}
 has two hyperparameters: the rank of $L$ (\ie the number of latent factors) and the shrinkage intensity $\lambda$.
-The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (\figref{02}{Row 1, B}).
+The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (\figref{1}{Row 1, B}).
 
@@
 The problem of finding the optimal set of non-zero elements of the precision matrix is known as \emph{covariance selection} \cite{Dempster:1972}.
-The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{02}{Row 1, C}).
+The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{1}{Row 1, C}).
 
@@
 The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$.
-The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (\figref{02}{Row 1, D}) \cite{Chandrasekaran:2010,Ma:2013}.
+The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (\figref{1}{Row 1, D}) \cite{Chandrasekaran:2010,Ma:2013}.
 
@@
 We next demonstrate how the most efficient among different regularized estimators can reveal the structure of correlations.
-We constructed four $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (\figref{02}{\,Row 2, A--D}, see Methods for details about how these matrices were generated).
+We constructed four $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (\figref{1}{\,Row 2, A--D}, see Methods for details about how these matrices were generated).
 We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes.
-Even with relatively large sample sizes (\eg $n=1000$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (\figref{02}{\,Row 3}).
+Even with relatively large sample sizes (\eg $n=1000$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (\figref{1}{\,Row 3}).
 
@@
 
-This loss function is related to the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$,
+This loss function is related to the multivariate normal log-likelihood function $L(\Sigma|C_{\sf sample})$,
 \begin{equation}
@@
 
-Although the two functions are similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf 0})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.
+Although the two functions are similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf sample})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf sample}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.
 
@@
 All the regularized estimators produced better estimates (lower excess loss) than the sample covariance matrix.
-However, estimators whose structure matched the true model typically outperformed the other estimators (\figref{02}{\,row 5}).
-There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$, often outperformed the estimators with the matching structure (\figref{02}{\,Row 5, C and D}).
+However, estimators whose structure matched the true model typically outperformed the other estimators (\figref{1}{\,row 5}).
+There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$, often outperformed the estimators with the matching structure (\figref{1}{\,Row 5, C and D}).
 Second, $C_{\sf sparse+latent}$ often performed as well as $C_{\sf sparse}$ on data from models without latent units because it reliably inferred the correct number of latent units: zero.
@@
 Instead, the loss function must be estimated from the sample through \emph{validation}.
-In a validation procedure, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf 0}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.
-The estimate $C_{\sf 0}^\prime$ is used to
-compute the \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$, which approximates the loss  $\loss{C,\Sigma}$.
+In a validation procedure, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf sample}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.
+The estimate $C_{\sf sample}^\prime$ is used to
+compute the \emph{validation loss} $\loss{C,C_{\sf sample}^\prime}$, which approximates the loss  $\loss{C,\Sigma}$.
 
@@
 \begin{equation*}
-    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}^\prime]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}.
+    \E[C_{\sf sample}^\prime]{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E[C_{\sf sample}^\prime]{C_{\sf sample}^\prime}}=\loss{C,\Sigma}.
 \end{equation*}
@@
 Then each subset is used as the validation sample with the other $K-1$ serving as the training sample.
-The validation losses from all such \sq{folds} are averaged to produce the \emph{cross-validation loss} or \emph{CV-loss} for short.
-Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ when the remaining $K-1$ subsets are treated as the training sample.
-Then the cross-validation loss for estimator $C$ is
-\begin{equation}\label{eq:cv-loss}
-    \ell_C=\frac 1 K \sum\limits_{k=1}^K \loss{C^{\{\setminus k\}},C_{\sf 0}^{\{k\}}}
-\end{equation}
-In this formulation, $\ell_C$ can be recognized as the cross-validated Gaussian log-likelihood (up to a constant offset and multiplier).
-However, the present framework is more general, as other loss functions can be used.
+The validation losses from all such \sq{folds} are averaged to produce the \emph{cross-validation loss}.
 
-Since we only need to compare estimators to each other, we are  interested in the \emph{relative} CV loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$,
-\begin{equation}\label{eq:rel-cv-loss}
-    \Delta\ell_{C,C_{\sf ref}} = \ell_{C}-\ell_{C_{\sf ref}}
-\end{equation}
-The relative cross-validation loss was used for the main findings in this study.
-
-To examine how well the relative CV-losses reproduced the relationships between the excess losses, we evaluated the relative CV-loss with respect to the estimator that matched the true structure of the model.
-Indeed, CV-losses accurately reproduced the relationships between the excess losses of the estimators, with somewhat greater variability (\figref{02}{Row 6}).
+Cross-validation loss accurately reproduced the relationships between the excess losses of the estimators, with somewhat greater variability (\figref{1}{Row 6}).
 
@@
 \paragraph{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
-We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{03}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
-This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).
-The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}).
+We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{2}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
+This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{2}{C}).
+The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{2}{C and D}).
 Only cells that produced detectable calcium activity were included in the analysis (For details, see Methods).
@@
 Each grating was presented for 500 ms, without intervening blanks.
-This stimulus was used to compute the orientation tuning of the recorded cells (\figref{03}{E}).
+This stimulus was used to compute the orientation tuning of the recorded cells (\figref{2}{E}).
 To estimate the noise correlation matrix, we presented only 2 distinct directions in some experiments or 5 directions in others with between 100 and 300 repetitions of each direction.
@@
 The traces were then binned into 150 ms intervals for the estimation of the covariance matrix.
-The sample correlation coefficients were largely positive and low (average across sites of 0.02 and \hl{under 0.04 in all sites}) (\figref{03}{G} and \figref{06}{D}).
+The sample correlation coefficients were largely positive and low (average across sites of 0.02 and \hl{less than 0.04 in any site}) (\figref{2}{G} and \figref{5}{D}).
 
@@
     {\bf G.} Histogram of noise correlation coefficients. The red line indicates the mean.
-} \label{fig:03}}
+} \label{fig:2}}
     {\includegraphics[width=8.3cm]{figures/Figure03.pdf}}
@@
 
-To test this hypothesis, we computed the relative CV losses (Eq.~\ref{eq:rel-cv-loss}) of estimators  $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in 31 imaged sites in 15 mice.
+To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
 The hyperparameters of each estimator were optimized by nested cross-validation (See Methods for details).
-Indeed, the median relative CV loss was significantly above zero (Fig.~\ref{fig:04}), indicating that $C_{\sf sparse+latent}$ produced estimates that were consistently closest to the true covariance matrix.
-\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-    {\caption{{\bf Estimator $C_{\sf sparse+latent}$ is more efficient than the other estimators on neural data.}
-    {\bf A--D.} Histograms of the relative cross-validation loss of estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$.
-    The histograms were obtained from 31 imaged sites in 15 mice. 
-    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
-    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}.
-\\{\bf A.} $C_{\sf sparse+latent}$ outperforms $C_{\sf 0}$: median improvement 0.16 nats/neuron, $p=3.9\times 10^{-5}$.
-\\{\bf B.} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median improvement 0.0032 nats/neuron, $p=2.2\times 10^{-3}$.
-\\{\bf C.} $C_{\sf sparse+latent}$ outperforms $C_{\sf factor}$: median improvement 0.16 nats/neuron, $p=6.9\times 10^{-5}$.
-\\{\bf D.} $C_{\sf sparse+latent}$ outperforms $C_{\sf sparse}$: median improvement 0.0016 nats/neuron, $p=5.5\times 10^{-6}$.
-} \label{fig:04}}
-    {\includegraphics{./figures/Figure04.pdf}}
+Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}).
+The respective median differences of the cross-validation loss were 0.018, 0.0025, 0.0029, and 0.0017 nats/cell/bin, significantly greater than zero ($p<10^{-5}$ in each comparison, Wilcoxon signed rank test).
+
+\begin{figure}[!ht]
+    \begin{center}
+    \includegraphics{./figures/src/Fig3.eps}
+    \end{center}
+    \caption{
+    {\bf Cross-validation losses of estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ relative to $C_{\sf sparse+latent}$.}
+    Covariance estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ produced consistently greater validation losses than $C_{\sf sparse+latent}$ ($p<10^{-5}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
+The box plots indicate the $25^{th}$, $50^{th}$, and $75^{th}$ percentiles with the whiskers extending to the minimum and maximum values after excluding the outliers marked with `+'.
+} \label{fig:3}
 \end{figure}
@@
 \begin{equation}\label{eq:quadratic}
-\loss{C,C_{\sf 0}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf 0}^\prime-I)^2,
+\loss{C,C_{\sf sample}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf sample}^\prime-I)^2,
 \end{equation}
@@
 
-We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:05} and Fig.~\ref{fig:06}).
-Although the regularized estimates were similar to the sample correlation matrix (\figref{05}{A and D}), the corresponding partial correlation matrices differed substantially (\figref{05}{B and E}).
-The partial correlation matrix of the regularized estimate was the sum of a sparse component and a low-rank component (\figref{05}{C}).
+We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:4} and Fig.~\ref{fig:5}).
+Although the regularized estimates were similar to the sample correlation matrix (\figref{4}{A and D}), the corresponding partial correlation matrices differed substantially (\figref{4}{B and E}).
+The partial correlation matrix of the regularized estimate was the sum of a sparse component and a low-rank component (\figref{4}{C}).
 
-In the example site (Fig.~\ref{fig:05}), the sparse component had 82.2\% sparsity (or conversely, 17.8\% connectivity: $connectivity=1-sparsity$) with average node degree of 52.9 (\figref{05}{G}).
+In the example site (Fig.~\ref{fig:4}), the sparse component had 89.9\% sparsity (or conversely, 11.1\% connectivity: $connectivity=1-sparsity$) with average node degree of 29.3 (\figref{4}{G}).
 The average node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$, where $p$ is the number of neurons.
-The low-rank component had rank 15, denoting 15 inferred latent units.
-The number of latent units and the average node degrees generally increased with population size, although these values varied widely between the datasets (\figref{06}{A and B}).
-There was an inverse relationship between the number of latent units and the average node degree (\figref{06}{C}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
+The low-rank component had rank 71, denoting 71 inferred latent units.
+The number of latent units and the average node degrees generally increased with population size, although these values varied widely between the datasets (\figref{5}{A and B}).
+There was an inverse relationship between the number of latent units and the average node degree (\figref{5}{C}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
 This variability may be explained by differences in brain states and recording quality and needs further investigation.
 
-The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 31 sites were about 5 times lower than the average sample correlations (\figref{06}{D}).
+The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 31 sites were about 5 times lower than the average sample correlations (\figref{5}{D}).
 This suggests that correlations between neurons build up from large numbers of small interactions with other neurons.
@@
 
-While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (43.3\% in the example site) of negative values (\figref{05}{F}).
-This fraction was consistent across multiple sites but the fraction of negative interactions increased with the inferred connectivity \figref{06}{F}.
+While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (28.7\% in the example site) of negative values (\figref{4}{F}).
+This fraction was consistent across multiple sites but the fraction of negative interactions increased with the inferred connectivity \figref{5}{F}.
 This increase suggests that negative interactions are inferred only after a sufficient density of positive interactions has been uncovered.
@@
     \begin{center}
-        \includegraphics[width=17.35cm]{./figures/Figure05.pdf}
+        \includegraphics[width=17.35cm]{./figures/Figure04.pdf}
     \end{center}
     \caption{{\bf Example of the structure revealed by $C_{\sf sparse+latent}$ the sparse+latent estimator.}
-    {\bf A and D.} The regularized estimate $C_{\sf sparse+latent}$ closely approximates the sample correlation matrix $C_{\sf 0}$. 
-    {\bf B and E.} However, the partial correlation matrices from the two estimates differ substantially.
-    {\bf C.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 82.2\% off-diagonal zeros (bottom-left) and low-rank component of rank 15 (top-right).
-    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations: the gray interval indicates the range of correlations containing 82.2\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
+    {\bf A and B.} The regularized estimate $C_{\sf sparse+latent}$ closely approximates the sample correlation matrix $C_{\sf sample}$. 
+    {\bf C and D.} However, the partial correlation matrices from the two estimates differ substantially.
+    {\bf E.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 89.9\% off-diagonal zeros (bottom-left) and low-rank component of rank 71 (top-right).
+    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations: the gray interval indicates the range of correlations containing 89.9\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
 The significant correlations were outside this interval.
-Yet 58.3\% of the interactions inferred by $C_{\sf sparse+latent}$  linked pairs of neurons whose correlation was below the threshold.
-Only the remaining 41.7\% of the interactions overlapped with sample correlations above the threshold.
+Yet 51.3\% of the interactions inferred by $C_{\sf sparse+latent}$  linked pairs of neurons whose correlation was below the threshold.
+Only the remaining 48.7\% of the interactions overlapped with sample correlations above the threshold.
     {\bf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
@@
 }
-\label{fig:05}
+\label{fig:4}
 \end{FPfigure}
@@
 The networks revealed by the two methods differed substantially.
-In the example site with 17.8\% connectivity inferred by $C_{\sf sparse+latent}$, only 41.7\% of connections coincided with above-threshold sample correlations (\figref{05}{F, H, and I}).
-In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{05}{F}) where high correlations were expected given the rest of the correlation matrix.
-This low overlap between the two types of graphs was consistent across the datasets (\figref{06}{E}).
+In the example site with 11.1\% connectivity inferred by $C_{\sf sparse+latent}$, only 48.7\% of connections coincided with above-threshold sample correlations (\figref{4}{F, H, and I}).
+In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{4}{F}) where high correlations were expected given the rest of the correlation matrix.
+This low overlap between the two types of graphs was consistent across the datasets (\figref{5}{E}).
 
@@
     {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites.} 
-    Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:03} and \ref{fig:05} is circled in blue. 
+    Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:2} and \ref{fig:4} is circled in blue. 
 \\
@@
 }
-\label{fig:06}}
+\label{fig:5}}
 {\includegraphics{./figures/Figure06.pdf}}
@@
 
-We also examined the relationship of the correlation structure to the differences in orientation preference and to the physical distances between the cell pairs (Fig.\;\ref{fig:07}).
-The partial correlations of the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation (\figref{07}{A}) than the sample correlations.
-They also decreased more rapidly with physical distance, both laterally  (\figref{07}{B}) and in depth (\figref{07}{B}).
+We also examined the relationship of the correlation structure to the differences in orientation preference and to the physical distances between the cell pairs (Fig.\;\ref{fig:6}).
+The partial correlations of the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation (\figref{6}{A}) than the sample correlations.
+They also decreased more rapidly with physical distance, both laterally  (\figref{6}{B}) and in depth (\figref{6}{B}).
 
@@
     {\bf A--C.} Average partial correlations (red) estimated by $C_{\sf sparse+latent}$ and average sample correlations (black) averaged across multiple imaged sites.
-At each site the correlations were normalized by the respective average correlation shown in \figref{06}{D}.
+At each site the correlations were normalized by the respective average correlation shown in \figref{5}{D}.
 The number of sites, $n$, that qualified to be included in the analysis is indicated in each panel.
@@
 }
-\label{fig:07}
+\label{fig:6}
 \end{figure}
 
-The positive and negative interactions in the $C_{\sf sparse+latent}$ estimator were organized differently: The density of the positive interactions fell rapidly with differences in the preferred orientation (\figref{07}{D}), lateral displacements (\figref{07}{E}), and displacement in depth (\figref{07}{F}) whereas the negative interactions were much less selective (\figref{07}{D--F}).
+The positive and negative interactions in the $C_{\sf sparse+latent}$ estimator were organized differently: The density of the positive interactions fell rapidly with differences in the preferred orientation (\figref{6}{D}), lateral displacements (\figref{6}{E}), and displacement in depth (\figref{6}{F}) whereas the negative interactions were much less selective (\figref{6}{D--F}).
 
@@
 However, this does not limit the applicability of its conclusions to normal distributions.
-Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:04} and Fig.~S\ref{supp:02}).
+Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:3} and Fig.~S\ref{supp:02}).
 Other probabilistic models, fitted to the same data, would also serve as estimators of the covariance matrix.
@@
 Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.
-Importantly, the inferred functional interactions were substantially different from the network of the most significant sample correlations \figref{05}{F, H, and I}.
-For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}) and may reflect inhibitory circuitry.
+Importantly, the inferred functional interactions were substantially different from the network of the most significant sample correlations \figref{4}{F, H, and I}.
+For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{4}{F}) and may reflect inhibitory circuitry.
 Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity, it may accurately reflect average connectivity profiles.
@@
 For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 14 and 27 minutes (mean 22 minutes).
-Cells that did not have substantial spiking activity (those whose variance was less than 2\% of the $90^{th}$ percentile in each site) were exculded from the analysis.
+Cells that did not have substantial spiking activity (those whose variance was less than 2\% of the $90^{th}$ percentile in each site) were excluded from the analysis.
 
@@
 When the cross-validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.
-This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{02}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:05}--\ref{fig:07}).
+This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:4}--\ref{fig:6}).
 \paragraph{Covariance estimation}
 Within the inner loop of cross-validation, covariance matrix estimation was performed with fixed hyperparameter values provided by the search algorithm.
-The computation of regularized estimators only required the sample covariance matrix $C_{\sf 0}$ of the training dataset.
+The computation of regularized estimators only required the sample covariance matrix $C_{\sf sample}$ of the training dataset.
 
 Estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag})  used two hyperparameters: the covariance shrinkage intensity $\lambda \in [0,1]$ and variance shrinkage intensity $\alpha \in [0,1]$.
-The variances (the diagonal of $C_{\sf 0}$) were shrunk toward (linearly mixed with) their mean value:
+The variances (the diagonal of $C_{\sf sample}$) were shrunk toward (linearly mixed with) their mean value:
 \begin{equation}
-D = (1-\alpha)C_{\sf 0}\circ I + \alpha \frac 1 p \Tr(C_{\sf 0}) I
+D = (1-\alpha)C_{\sf sample}\circ I + \alpha \frac 1 p \Tr(C_{\sf sample}) I
 \end{equation}
@@
 Estimator $C_{\sf factor}$ used two hyperparameters: the number of latent factors $d$ and the shrinkage intensity $\lambda \in [0, 1]$.  
-The matrix $L$ of rank $d$ and the diagonal matrix of individual variances $\Psi$ were computed by solving the minimization problem
+The matrix $L$ of rank $d$ and the diagonal matrix of individual variances $D$ were computed by solving the minimization problem
 \begin{equation}
-(L,\Psi) = \argmin\limits_{\tilde L,\tilde\Psi} \loss{\tilde L + \tilde\Psi,C_{\sf 0}},
+(L,D) = \argmin\limits_{\tilde L,\tilde D} \loss{\tilde L + \tilde D,C_{\sf sample}},
 \end{equation}
 which we solved by an expectation-maximization (EM) algorithm.
-Under our chosen loss function (Eq.~\ref{eq:loss}), this is equivalent to maximum likelihood estimation of $L$ and $\Psi$ under the multivariate Gaussian distribution.
+Under our chosen loss function (Eq.~\ref{eq:loss}), this is equivalent to maximum likelihood estimation of $L$ and $D$ under the multivariate Gaussian distribution.
 The final regularized estimate is obtained by linear shrinkage toward the factor model (Eq.~\ref{eq:c-factor}).
@@
 \begin{equation}
-Z = \argmin\limits_{\tilde Z \succ 0} \loss{{\tilde Z}^{-1},C_{\sf 0}} + \lambda \|\tilde Z \|_1
+Z = \argmin\limits_{\tilde Z \succ 0} \loss{{\tilde Z}^{-1},C_{\sf sample}} + \lambda \|\tilde Z \|_1
 \end{equation}
@@
 \begin{equation}
-S = \argmin\limits_{\tilde S \in Z^\sharp} \loss{\tilde S,C_{\sf 0}},
+S = \argmin\limits_{\tilde S \in Z^\sharp} \loss{\tilde S,C_{\sf sample}},
 \end{equation}
@@
 \begin{equation}
-(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L, C_{\sf 0}} + \lambda\|\tilde Z\|_1
+(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L, C_{\sf sample}} + \lambda\|\tilde Z\|_1
 \end{equation}
@@
 \begin{equation}
-(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L, C_{\sf 0} }
+(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L, C_{\sf sample} }
 \end{equation}
@@
 The partial correlation matrix with both sparse and low-rank components is computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}; it includes the effects of interactions between the visible and latent units.
-This estimate of the partial correlations was used in analyses of average partial correlations (\figref{05}{B and E} and \figref{06}{D}, and \figref{07}{A--C}).
-The sparse component $S$ was used separately in analyses of the connectivity graphs (\figref{05}{F, G, H} and \figref{06}{B, C, E, and F} and \figref{07}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component alone without the low-rank component:
+This estimate of the partial correlations was used in analyses of average partial correlations (\figref{4}{B and E} and \figref{5}{D}, and \figref{6}{A--C}).
+The sparse component $S$ was used separately in analyses of the connectivity graphs (\figref{4}{F, G, H} and \figref{5}{B, C, E, and F} and \figref{6}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component alone without the low-rank component:
 \begin{equation}
@@
 For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units.
-The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{02}{\,Row 2}).
+The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{1}{\,Row 2}).
 Samples were then drawn from multivariate normal distributions with the respective true covariance matrices to be estimated by each of the estimators.
@@
 \caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators with respect to multivariate normal cross-validation loss.}
-\\{\bf A--D.} These panels are identical to those in \figref{04}{A--D}.
-\\{\bf E.} $C_{\sf sparse}$ outperforms $C_{\sf 0}$: median improvement 0.15 nats/neuron, $p=3.9\times 10^{-5}$.
+\\{\bf A--D.} These panels are identical to those in \figref{3}{A--D}.
+\\{\bf E.} $C_{\sf sparse}$ outperforms $C_{\sf sample}$: median improvement 0.15 nats/neuron, $p=3.9\times 10^{-5}$.
 \\{\bf F.} $C_{\sf sparse}$ performs similarly to $C_{\sf diag}$: median improvement $1.0\times10^{-3}$ nats/neuron, $p=0.13$.
 \\{\bf G.} $C_{\sf sparse}$ outperforms $C_{\sf factor}$: median improvement $0.15$ nats/neuron, $p=1.2\times10^{-4}$.
-\\{\bf H.} $C_{\sf factor}$ outperforms $C_{\sf 0}$: median improvement $0.029$ nats/neuron, $p=2.7\times10^{-4}$.
+\\{\bf H.} $C_{\sf factor}$ outperforms $C_{\sf sample}$: median improvement $0.029$ nats/neuron, $p=2.7\times10^{-4}$.
 \\{\bf I.} $C_{\sf factor}$ performs worse than $C_{\sf diag}$: median improvement $-0.16$ nats/neuron, $p=1.6\times10^{-4}$.
-\\{\bf J.} $C_{\sf diag}$ outperforms $C_{\sf 0}$: median improvement $0.14$ nats/neuron, $p=2.7\times10^{-5}$.
+\\{\bf J.} $C_{\sf diag}$ outperforms $C_{\sf sample}$: median improvement $0.14$ nats/neuron, $p=2.7\times10^{-5}$.
 }
@@
 \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-{\caption{{\bf The sparse+latent estimator outperforms the other estimators under the quadratic loss (Eq.~\ref{eq:quadratic}) just as it did under the normal loss (Eq.~\ref{eq:loss}) used in Fig.~\ref{fig:04}.}
-    Histograms of the relative cross-validation loss (Eq.~\ref{eq:rel-cv-loss}) of estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ using the quadratic loss function.
+{\caption{{\bf The sparse+latent estimator outperforms the other estimators under the quadratic loss (Eq.~\ref{eq:quadratic}) just as it did under the normal loss (Eq.~\ref{eq:loss}) used in Fig.~\ref{fig:3}.}
+    Histograms of the cross-validation loss of estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ using the quadratic loss function.
     The histograms were obtained from 31 imaged sites in 15 mice. 
     All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
-    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}.
+    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:2} and Fig.~\ref{fig:4}.
     The $p$-values are for the two-sided Wilcoxon's signed rank test.
-    \\{\bf A.} $C_{\sf sparse+latent}$ outperforms $C_{\sf 0}$: median $\Delta\ell_{C_{\sf 0},C_{\sf sparse+latent}}$ = 0.35, $p=1.7\times 10^{-6}$.
+    \\{\bf A.} $C_{\sf sparse+latent}$ outperforms $C_{\sf sample}$: median $\Delta\ell_{C_{\sf sample},C_{\sf sparse+latent}}$ = 0.35, $p=1.7\times 10^{-6}$.
     \\{\bf B.} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median $\Delta\ell_{C_{\sf diag},C_{\sf sparse+latent}}$ = $2.5\times 10^{-3}$, $p=1.0\times 10^{-4}$.

### 2014-05-05 plos.tex
completed figures 3, 4, 5, and 6.

@@
 where $\circ$ denotes the entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. 
-Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$:
+Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$ \cite{Whittaker:2009}:
 \begin{equation}\label{eq:partial}
@@
 \end{equation}
-
-To see that this is the case, partition the covariance matrix 
-\begin{equation}
-C = \begin{pmatrix}
-C_{11} & C_{12} \\
-C_{12}^\T & C_{22}
-\end{pmatrix}
-\end{equation}
-so that $C_{11}$ is the $2\times 2$ covariance matrix of two cells of interest. Such partitioning can be done for any pair of cells by changing their order. 
-Then the covariance matrix of the conditional distribution of these two cells is \cite{Anderson:2003}:
-\begin{equation}
-\hat C_{11} = C_{11}-C_{12}C^{-1}C_{12}^\T 
-\end{equation}
-The inverse of the conditional covariance matrix can be found in the corresponding partition of the precision matrix $C^{-1}$:
-\begin{equation}
-C^{-1}=\begin{pmatrix}
-\hat C_{11}^{-1} &   -\hat C_{11}^{-1}C_{12}C_{22}^{-1} \\
--C_{22}^{-1}C_{12}^\T\hat C_{11}^{-1}  &  C_{22}^{-1}+C_{22}^{-1}C_{12}^\T\hat C_{11}^{-1}C_{12}C_{22}^{-1}
-\end{pmatrix}
-\end{equation}
-The inverse of a $2\times 2$ correlation matrix reverses the sign of its off-diagonal elements:
-\begin{equation}
-\begin{pmatrix}
-1 & \rho \\
-\rho & 1
-\end{pmatrix}^{-1}
-\propto 
-\begin{pmatrix}
--1 & \rho \\
-\rho & -1 
-\end{pmatrix}
-\end{equation}
-Therefore, scaling $C^{-1}$ to have 1s on the diagonal yields the partial correlation in each of its $2\times 2$ partitions, resulting in a matrix of partial pairwise correlations.
 
@@
     \begin{center}
-        \includegraphics{./figures/Figure02.pdf}
+        \includegraphics{./figures/Figure01.pdf}
     \end{center}
@@
 } \label{fig:2}}
-    {\includegraphics[width=8.3cm]{figures/Figure03.pdf}}
+    {\includegraphics[width=8.3cm]{figures/Figure02.pdf}}
 \end{figure}
@@
 Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}).
-The respective median differences of the cross-validation loss were 0.018, 0.0025, 0.0029, and 0.0017 nats/cell/bin, significantly greater than zero ($p<10^{-5}$ in each comparison, Wilcoxon signed rank test).
+The respective median differences of the cross-validation loss were 0.019, 0.0027, 0.0031, and 0.0022 nats/cell/bin, significantly greater than zero ($p<10^{-4}$ in each comparison, Wilcoxon signed rank test).
 
@@
 \label{fig:5}}
-{\includegraphics{./figures/Figure06.pdf}}
+{\includegraphics{./figures/Figure05.pdf}}
 \end{figure}
@@
     \begin{center}
-        \includegraphics{./figures/Figure07.pdf}
+        \includegraphics{./figures/Figure06.pdf}
     \end{center}
@@
 
-\begin{FPfigure}
+\begin{figure}[!ht]
 \begin{center}
-\includegraphics{./figures/Figure-Supp01.pdf}
+\includegraphics{./figures/src/Supp1.eps}
 \end{center}
-\caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators with respect to multivariate normal cross-validation loss.}
-\\{\bf A--D.} These panels are identical to those in \figref{3}{A--D}.
-\\{\bf E.} $C_{\sf sparse}$ outperforms $C_{\sf sample}$: median improvement 0.15 nats/neuron, $p=3.9\times 10^{-5}$.
-\\{\bf F.} $C_{\sf sparse}$ performs similarly to $C_{\sf diag}$: median improvement $1.0\times10^{-3}$ nats/neuron, $p=0.13$.
-\\{\bf G.} $C_{\sf sparse}$ outperforms $C_{\sf factor}$: median improvement $0.15$ nats/neuron, $p=1.2\times10^{-4}$.
-\\{\bf H.} $C_{\sf factor}$ outperforms $C_{\sf sample}$: median improvement $0.029$ nats/neuron, $p=2.7\times10^{-4}$.
-\\{\bf I.} $C_{\sf factor}$ performs worse than $C_{\sf diag}$: median improvement $-0.16$ nats/neuron, $p=1.6\times10^{-4}$.
-\\{\bf J.} $C_{\sf diag}$ outperforms $C_{\sf sample}$: median improvement $0.14$ nats/neuron, $p=2.7\times10^{-5}$.
-}
+\caption{Comparison of the $C_{\sf sparse}$ estimator to the other estimators with respect to the normal loss function.}
 \label{supp:01}
-\end{FPfigure}
+\end{figure}
 
 \begin{figure}[!ht]
-\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-{\caption{{\bf The sparse+latent estimator outperforms the other estimators under the quadratic loss (Eq.~\ref{eq:quadratic}) just as it did under the normal loss (Eq.~\ref{eq:loss}) used in Fig.~\ref{fig:3}.}
-    Histograms of the cross-validation loss of estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ using the quadratic loss function.
-    The histograms were obtained from 31 imaged sites in 15 mice. 
-    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
-    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:2} and Fig.~\ref{fig:4}.
-    The $p$-values are for the two-sided Wilcoxon's signed rank test.
-    \\{\bf A.} $C_{\sf sparse+latent}$ outperforms $C_{\sf sample}$: median $\Delta\ell_{C_{\sf sample},C_{\sf sparse+latent}}$ = 0.35, $p=1.7\times 10^{-6}$.
-    \\{\bf B.} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median $\Delta\ell_{C_{\sf diag},C_{\sf sparse+latent}}$ = $2.5\times 10^{-3}$, $p=1.0\times 10^{-4}$.
-    \\{\bf C.} $C_{\sf sparse+latent}$ outperforms $C_{\sf factor}$: median $\Delta\ell_{C_{\sf factor},C_{\sf sparse+latent}}$ = 0.27, $p=1.2\times 10^{-6}$.
-    \\{\bf D.} $C_{\sf sparse+latent}$ outperforms $C_{\sf sparse}$: median $\Delta\ell_{C_{\sf sparse},C_{\sf sparse+latent}}$ = $3.7\times 10^{-4}$, $p=2.2\times 10^{-3}$.
-}
-\label{supp:02}}
-{\includegraphics{./figures/Figure-Supp02.pdf}}
+\begin{center}
+\includegraphics{./figures/src/Supp2.eps}
+\end{center}
+\caption{Comparison of the $C_{\sf sparse+latent}$ estimator to the other estimators with respect to the quadratic loss.}
+\label{supp:02}
 \end{figure}

### 2014-05-07 plos.tex
added section on variance estimation

@@
 Indeed, in cross-validation tests, the covariance matrix estimator with this structure consistently outperformed other regularized estimators.
-The sparse component of the estimate expressed a graph of interactions organized with respect to the physical distances and orientation tuning properties of cells: The density of positive \sq{excitatory} interactions decreased rapidly with geometric distances and with differences in orientation preference whereas the negative \sq{inhibitory} interactions were less selective.
-Because of its superior performance, this ``sparse + latent'' estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.
+The sparse component of the estimate defined a graph of interactions.
+These interactions reflected the physical distances and orientation tuning properties of cells: The density of positive \sq{excitatory} interactions decreased rapidly with geometric distances and with differences in orientation preference whereas negative \sq{inhibitory} interactions were less selective.
+Because of its superior performance, this \sq{sparse + latent} estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.
 
@@
 \\
-{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) for the five estimators with respect to the matching estimator.
-Error bars indicate the standard error of the mean based on 30 samples.
+{\bf Row 6:} Cross-validation losses for the five estimators relative to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
     }
@@
 
-The target of estimator $C_{\sf factor}$ is the factor model $L + D$, where $L$ is the $p\times p$ positive \hl{semidefinite} matrix with low rank and $D$ is the diagonal matrix of independent variances.
+Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model $L + D$, where $L$ is the $p\times p$ positive \hl{semidefinite} matrix with low rank and $D$ is a diagonal matrix of independent variances.
 The estimator,
 \begin{equation}\label{eq:c-factor}
-C_{\sf factor} = (1-\lambda) C_{\sf sample} + \lambda (L + D),
+C_{\sf factor} = L + (1-\lambda)D + \lambda \bar D,
 \end{equation}
@@
 This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{2}{C}).
-The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{2}{C and D}).
+The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{2}{C}).
 Only cells that produced detectable calcium activity were included in the analysis (For details, see Methods).
@@
 Each grating was presented for 500 ms, without intervening blanks.
-This stimulus was used to compute the orientation tuning of the recorded cells (\figref{2}{E}).
-To estimate the noise correlation matrix, we presented only 2 distinct directions in some experiments or 5 directions in others with between 100 and 300 repetitions of each direction.
-In this stimulus, each grating lasted 1 second and was followed by a 1-second blank.
-The average stimulus response was subtracted from each trial.
-The traces were then binned into 150 ms intervals for the estimation of the covariance matrix.
-The sample correlation coefficients were largely positive and low (average across sites of 0.02 and \hl{less than 0.04 in any site}) (\figref{2}{G} and \figref{5}{D}).
+This stimulus was used to compute the orientation tuning of the recorded cells (\figref{2}{D}).
+To estimate the noise correlation matrix, we presented only 2 distinct directions in some experiments or 5 directions in others with 100--300 repetitions of each direction.
+Each grating lasted 1 second and was followed by a 1-second blank.
+The traces were then binned into 150 ms intervals aligned to the stimulus onset for the estimation of the covariance matrix.
+The sample correlation coefficients were largely positive and low.
+The average value of the correlation coefficient across sites ranged from 0.0066 to 0.043, mean 0.018 (\figref{2}{F} and \figref{5}{D}).
 
@@
     Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) presented during two-photon recordings of somatic calcium signals using fast 3D random-access microscopy ({\bf B}).
-    {\bf C--G.} Calcium activity data from an example site.
-    {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. The inferred firing rate binned in 150 ms intervals are indicated by red ticks below each trace.
-    {\bf D.} The raster plot of the inferred firing rates, binned in 150 ms intervals, from the entire population from the first (left) and last (right) minute of the entire recording.
-The traces from {\bf C} are highlighted in red.
-    {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site.
+    {\bf C--F.} Calcium activity data from an example site.
+    {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. Spiking activity inferred by non-negative deconvolution is shown below the trace.
+    {\bf D.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site.
 Only spiking cells are shown.
@@
 The gray cells were not significantly tuned.
-    {\bf F.} The sample noise correlation matrix of the activity of the neural population. 
-    {\bf G.} Histogram of noise correlation coefficients. The red line indicates the mean.
+    {\bf E.} The sample noise correlation matrix of the activity of the neural population. 
+    {\bf F.} Histogram of noise correlation coefficients. The red line indicates the mean correlation coefficient of 0.017.
 } \label{fig:2}}
@@
 We therefore hypothesized that covariance matrix estimators that explicitly modeled the partial correlations between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would have a performance advantage.
-However, the observed neurons must also be strongly influenced by global activity fluctuations and by unobserved common inputs, to the advantage of estimators that explicitly model common fluctuations of the entire population: $C_{\sf factor}$ and $C_{\sf sparse+latent}$.
+However, the observed neurons must also be strongly influenced by global activity fluctuations and by unobserved common inputs to the advantage of estimators that explicitly model common fluctuations of the entire population: $C_{\sf factor}$ and $C_{\sf sparse+latent}$.
 If both types of effects are significant, then $C_{\sf sparse+latent}$ should outperform the other estimators.
@@
     \caption{
-    {\bf Cross-validation losses of estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ relative to $C_{\sf sparse+latent}$.}
-    Covariance estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ produced consistently greater validation losses than $C_{\sf sparse+latent}$ ($p<10^{-5}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
+   {\bf Performance of estimator $C_{\sf sparse+latent}$ with respect to the normal loss function (eq.~\ref{eq:loss}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.}
+    Covariance estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ produced consistently greater validation losses than $C_{\sf sparse+latent}$ ($p<10^{-4}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
 The box plots indicate the $25^{th}$, $50^{th}$, and $75^{th}$ percentiles with the whiskers extending to the minimum and maximum values after excluding the outliers marked with `+'.
@@
 
-Furthermore, $C_{\sf sparse}$ outperformed all the other estimators besides $C_{\sf sparse+latent}$ (Fig.~S\ref{supp:01}) suggesting the greater importance of modeling interactions between pairs of neurons than the common fluctuations of activity across the entire population.
+Furthermore, estimator $C_{\sf sparse}$ outperformed estimator $C_{\sf factor}$ (Fig.~S\ref{supp:01}) suggesting the greater importance of modeling interactions between pairs of neurons than the common fluctuations of activity across the entire population.
 
 The same evaluation based on the quadratic loss function,
-\begin{equation}\label{eq:quadratic}
+\begin{equation}\label{eq:quad}
 \loss{C,C_{\sf sample}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf sample}^\prime-I)^2,
@@
     {\bf A.} Average correlations between pairs of neurons tuned to orientation with differences in preferred orientation in the intervals of 0--15$^\circ$, 15--45$^\circ$ and 45--90$^\circ$. 
-    {\bf B.} Average correlations between pairs of neurons located at the same depth ($\pm$25$\mu$m) separated by lateral distances in the intervals of 0--25 $\mu$m, 25--75 $\mu$m, 75--150 $\mu$m, and 150+ $\mu$m.
-    {\bf C.} Average correlations between pairs of neurons displaced laterally by less than 25 $\mu$m separated in depth by distances in the intervals of 0--25 $\mu$m, 25--60 $\mu$m, and 60+ $\mu$m.
+    {\bf B.} Average correlations between pairs of neurons located at the same depth ($\pm$30$\mu$m) separated by lateral distances in the intervals of 0--25 $\mu$m, 25--75 $\mu$m, 75--150 $\mu$m, and 150+ $\mu$m.
+    {\bf C.} Average correlations between pairs of neurons displaced laterally by less than 30 $\mu$m separated in depth by distances in the intervals of 0--25 $\mu$m, 25--60 $\mu$m, and 60+ $\mu$m.
     {\bf D--F.} Normalized connectivity of positive (green) and negative (dark red) interactions from the sparse component obtained from $C_{\sf sparse+latent}$.
@@
 First, the first principal component was subtracted from the traces, which reduced common mode noise related to small cardiovascular movements \cite{Cotton:2013}.
-The resulting traces were low-pass filtered below 0.1 Hz and downsampled to 20 Hz.
+The resulting traces were low-pass filtered below 0.1 Hz and downsampled to 20 Hz (\figref{2}{C}).
 Firing rates were estimated using a fast non-negative deconvolution algorithm \cite{Vogelstein:2010}.
@@
 For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 14 and 27 minutes (mean 22 minutes).
-Cells that did not have substantial spiking activity (those whose variance was less than 2\% of the $90^{th}$ percentile in each site) were excluded from the analysis.
+Cells that did not have substantial spiking activity (those whose variance was less than 1\% of the median across the site) or whose activity was unstable (those whose variance in the least active quarter of the recording did not exceed 1\% of the variance in the most active quarter) were excluded from the analysis.
 
@@
 \begin{equation}
-(L,D) = \argmin\limits_{\tilde L,\tilde D} \loss{\tilde L + \tilde D,C_{\sf sample}},
+(L,D) = \argmin\limits_{\hat L,\hat D} \loss{\hat L + \hat D,C_{\sf sample}},
 \end{equation}
@@
 
-Estimator $C_{\sf sparse}$  has one hyperparameter $\lambda$ to regulate the sparsity of its precision matrix $S$.
-The precision matrix is computed in two steps: First, the zero structure is determined by minimizing the $L_1$-penalized loss:
+Estimator $C_{\sf sparse}$  has one hyperparameter $\lambda$ to regulate the sparsity of its precision matrix $S$. The precision matrix is obtained by minimizing the $L_1$-penalized loss:
 \begin{equation}
-Z = \argmin\limits_{\tilde Z \succ 0} \loss{{\tilde Z}^{-1},C_{\sf sample}} + \lambda \|\tilde Z \|_1
+S = \argmin\limits_{\hat S \succ 0} \loss{{\hat S}^{-1},C_{\sf sample}} + \lambda \|\hat S \|_1
 \end{equation}
-where $\tilde Z\succ 0$ denotes the constraint that $\tilde Z$ be a positive definite matrix and $\|\tilde Z\|_1$ is the element-wise $L_1$ norm of the matrix $\tilde Z$.
+where $\hat S\succ 0$ denotes the constraint that $\hat S$ be a positive definite matrix and $\|\hat S\|_1$ is the element-wise $L_1$ norm of the matrix $\hat S$.
 This problem formulation is known as \emph{graphical lasso} \cite{Meinshausen:2006, Friedman:2008}.
-To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm \cite{Ma:2013}.
-Then, after the zero structure was determined, the remaining coefficients were fitted without penalty:
-\begin{equation}
-S = \argmin\limits_{\tilde S \in Z^\sharp} \loss{\tilde S,C_{\sf sample}},
-\end{equation}
-where $Z^\sharp$ denotes the set of positive-definite matrices with the same zero structure as $Z$.
-This step was also solved by ADMM.
+To solve this minimization problem, we adapted the alternative-direction method of multipliers (ADMM) algorithm \cite{Ma:2013}.
 Then the final estimate is the inverse of $S$ (Eq.~\ref{eq:c-sparse}).
@@
 Rather than searching for the optimal sparse structure of $S_{12}$ and $S_{22}$, an ill-posed problem, we estimated these components together as the low-rank matrix $L$.
-
-The estimate is found in two steps. First, we use the ADMM algorithm to find the zero structure of $S$ by minimizing the $L_1$-penalized loss \cite{Chandrasekaran:2010,Ma:2013}:
+We solved this problem by adapting the ADMM algorithm to minimizing the loss function with penalty on the $L_1$-norm  of $S$ and the nuclear norm of $L$ \cite{Chandrasekaran:2010,Ma:2013}:
 \begin{equation}
-(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L, C_{\sf sample}} + \lambda\|\tilde Z\|_1
-\end{equation}
-Then we find the sparse and low-rank components of the precision matrix by solving (also using ADMM) 
-\begin{equation}
-(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L, C_{\sf sample} }
+(S,L) = \argmin\limits_{\hat S,\hat L} \loss{\hat S-\hat L, C_{\sf sample}} + \alpha\|\hat S\|_1 + \beta \Tr{\hat L}
 \end{equation}
@@
 \section*{Supporting Information}
+\subsection*{Appendix: Estimation of variance}
+One decision that was made in this paper was to model a constant covariance matrix across different conditions.
+In other words, the variances was modeled as a constant across conditions.
+In contrast, it is more common to estimate correlations separately in different conditions and average the correlations coefficients from different conditions.
+In this appendix we justify this decision.
+
+Let's consider two random variables $x_1$ and $x_2$ with zero means and respective variances $v_1^2$ and $v_2^2$ in one condition and $a^2v_1^2$ and $b^2v_2^2$ in the second condition.
+Assume that the correlation coefficient $\rho$ is the same in both conditions.
+
+Let $\langle \cdot \rangle_i$ denote the expectation in condition $i$. Then
+\begin{align*}
+\langle x_1^2 \rangle_1 = v_1^2 && \langle x_2^2 \rangle_1 = v_2^2 && \rho = \frac{\langle x_1 x_2 \rangle_1}{v_1v_2}\\ 
+\langle x_1^2 \rangle_2 = a^2v_1^2 && \langle x_2^2 \rangle_2 = b^2v_2^2  && \rho = \frac{\langle x_1 x_2 \rangle_2}{ab\,v_1v_2} 
+\end{align*}
+
+If we ignore the differences between the variances in the two conditions, assuming equal numbers of trials in each condition, the respective  variances for the two variables will be 
+\begin{align*}
+\tilde v_1^2 = \frac 1 2 \left( v_1^2 + a^2v_1^2 \right) = \frac{1+a^2}{2} v_1^2
+&&
+\tilde v_2^2 = \frac 1 2 \left( v_2^2 + b^2v_2^2 \right) = \frac{1+b^2}{2} v_2^2
+\end{align*}
+The resulting correlation coefficient will be
+\begin{equation*}
+\tilde \rho = \frac{\langle x_1 x_2 \rangle_1 + \langle x_1 x_2 \rangle_2} {2 \tilde v_1 \tilde v_2}
+= \rho \frac{1 + ab}{\sqrt{1+a^2}\sqrt{1+b^2}}
+= \rho \sqrt{\frac{1+2ab+a^2b^2}{1+a^2+b^2+a^2b^2} }
+\end{equation*}
+Note that $\tilde \rho = \rho$ iff $a=b$ and $\tilde \rho<\rho$ when $a\ne b$.
+In other words, when stimulus conditions modulate the variances by different amounts, the assumption of constant variance leads to the underestimation of the correlation between differently tuned pairs of cells.
+This effect may lead to the overestimation of the effect of common tuning on correlations (`limited-range correlations').
+
+\subsection*{Supplementary Figures}
 \setcounter{figure}{0}
@@
 \end{center}
-\caption{Comparison of the $C_{\sf sparse}$ estimator to the other estimators with respect to the normal loss function.}
+\caption{
+{\bf Performance of estimator $C_{\sf sparse}$ with respect to normal loss (Eq.~\ref{eq:loss}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$}.
+Estimator $C_{\sf sparse}$ outperformed $C_{\sf sample}$ ($p=5\cdot 10^{-6}$) and $C_{\sf factor}$ ($p=.01$), but not $C_{\sf diag}$ ($p=.17$).}
 \label{supp:01}
@@
 \end{center}
-\caption{Comparison of the $C_{\sf sparse+latent}$ estimator to the other estimators with respect to the quadratic loss.}
+\caption{
+{\bf Performance of estimator $C_{\sf sparse+latent}$ with respect to \emph{quadratic loss} (Eq.~\ref{eq:quad}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.}
+All comparisons showed significant advantage of estimator $C_{\sf sparse+latent}$ ($p<10^{-3}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).}
 \label{supp:02}

### 2014-05-15 plos.tex

--- 2014-05-03 plos.tex
@@
 The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.
-When a large fraction of the interacting variables are observed, partial correlations more closely reflect the direct causal effects between components of complex systems than correlations.
+When a large fraction of the interacting variables are observed, partial correlations more closely reflect the direct causal effects between components of complex systems than correlations \cite{Whittaker:2009}.
 Partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
@@
 where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.
-The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf sample}$ from the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
+The usual estimate of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf sample}$ of the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
 \begin{equation}\label{eq:sample}
@@
 \end{equation}
-where $\nu$ is the number of degrees of freedom in one neuron's activity within the sample ($\nu=n-1$ if observations are independent).
+where $\nu$ denotes degrees of freedom in one neuron's activity within the sample ($\nu=n-1$ if observations are independent).
 For noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.
@@
 
-Given an arbitrary covariance matrix estimate $C$, the corresponding correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
+Given an covariance matrix estimate, the correlation matrix $R$ is calculated by normalizing the rows and columns of $C$ by the square roots of its diagonal elements to produce unit entries on the diagonal:
 \begin{equation}\label{eq:precision}
-    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2},
+    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
 \end{equation}
-where $\circ$ denotes the entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. 
-Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$:
+Here $\circ$ denotes the entrywise matrix product (Hadamard product) and $I$ is the identity matrix.
+
+\emph{Pairwise partial correlations} are less familiar in the neuroscience community. 
+The partial correlation between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:2009}.
+Partial correlations figure prominantly in probablistic \emph{graphical modelling} wherein the joint distribution is explain by sets of two-way interactions \cite{Whittaker:1990}.
+Under the multivariate Gaussian distribution, zero partial correlations are equivalent to conditional independence of the pair, implying a lack of direct interaction between them \cite{Dempster:1972, Whittaker:1990}.
+In the general case, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1999,Baba:2004}.
+
+Pairwise partial correlations are closely linked to the \emph{precision matrix}, \ie the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1999}.
+Given the covariance estimate $C$, the matrix of partial correlations $P$ is computed by normalizing the rows and columns of the \emph{precision matrix} $C^{-1}$ to produce negative unit entries on the diagonal:
 \begin{equation}\label{eq:partial}

### 2014-05-15 plos.tex
minor

--- 2014-05-07 plos.tex
@@
 One decision that was made in this paper was to model a constant covariance matrix across different conditions.
-In other words, the variances was modeled as a constant across conditions.
-In contrast, it is more common to estimate correlations separately in different conditions and average the correlations coefficients from different conditions.
+In other words, a single variance was estimated for each neuron across all stimulus conditions.
+A more commonplace practice, however, is to estimate the correlations separately in each condition and only then to average the correlations coefficients across all the conditions.
 In this appendix we justify this decision.
 
-Let's consider two random variables $x_1$ and $x_2$ with zero means and respective variances $v_1^2$ and $v_2^2$ in one condition and $a^2v_1^2$ and $b^2v_2^2$ in the second condition.
-Assume that the correlation coefficient $\rho$ is the same in both conditions.
+Let $x_1$ and $x_2$ be random variables with zero means and let their respective variances be $v_1^2$ and $v_2^2$ in condition 1 and $a^2v_1^2$ and $b^2v_2^2$ in condition 2.
+Assume that the correlation coefficient $\rho$ is constant across the two conditions.
 
@@
 
-If we ignore the differences between the variances in the two conditions, assuming equal numbers of trials in each condition, the respective  variances for the two variables will be 
+What is the effect of assuming constant variances across the conditions on estimating the correlation coefficient?
+Let $0<w<1$ be the fraction of trials in condition 1 and $1-w$ be the fraction of trials in condition 2.
+The respective variances for the two variables across both conditions will be
 \begin{align*}
-\tilde v_1^2 = \frac 1 2 \left( v_1^2 + a^2v_1^2 \right) = \frac{1+a^2}{2} v_1^2
-&&
-\tilde v_2^2 = \frac 1 2 \left( v_2^2 + b^2v_2^2 \right) = \frac{1+b^2}{2} v_2^2
+\tilde v_1^2 & = w\langle x_1^2 \rangle_1 + (1-w) \langle x_1^2 \rangle_2 = \left[w+(1-w)a^2\right] v_1^2
+\\
+\tilde v_2^2 & = w\langle x_2^2 \rangle_1 + (1-w) \langle x_2^2 \rangle_2 = \left[w+(1-w)b^2\right] v_2^2
 \end{align*}
@@
 \begin{equation*}
-\tilde \rho = \frac{\langle x_1 x_2 \rangle_1 + \langle x_1 x_2 \rangle_2} {2 \tilde v_1 \tilde v_2}
-= \rho \frac{1 + ab}{\sqrt{1+a^2}\sqrt{1+b^2}}
-= \rho \sqrt{\frac{1+2ab+a^2b^2}{1+a^2+b^2+a^2b^2} }
+\begin{split}
+\tilde \rho & = \frac{w\langle x_1 x_2 \rangle_1 + (1-w)\langle x_1 x_2 \rangle_2} { \tilde v_1 \tilde v_2}
+= \frac{w\rho + (1-w)ab\rho}{\sqrt{w+(1-w)a^2}\sqrt{w+(1-w)b^2}}
+\\
+& = \rho \sqrt{\frac{w^2+2w(1-w)ab+(1-w)^2a^2b^2}{w^2+w(1-w)(a^2+b^2)+(1-w)^2a^2b^2} }
+\end{split}
 \end{equation*}
-Note that $\tilde \rho = \rho$ iff $a=b$ and $\tilde \rho<\rho$ when $a\ne b$.
-In other words, when stimulus conditions modulate the variances by different amounts, the assumption of constant variance leads to the underestimation of the correlation between differently tuned pairs of cells.
-This effect may lead to the overestimation of the effect of common tuning on correlations (`limited-range correlations').
+Since $2ab=a^2+b^2$ when $a=b$ and $2ab < a^2 + b^2$ when $a\ne b$, then $\tilde \rho= \rho$ when $a=b$ and $\tilde \rho < \rho$ when $a\ne b$.
+
+In other words, the assumption of constant variance across stimulus condition results in underestimation of their correlation for cells whose variances have different stimulus tunings but not if they have they are tuned similarly.
+This effect may lead to overestimating of the effect of common tuning on correlations (`limited-range correlations').
+
+However, there are arguments for assuming a constant variance as well.
 

### 2014-05-15 plos.tex

--- 2014-05-07 plos.tex
@@
 
-Given an arbitrary covariance matrix estimate $C$, the corresponding correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
+Given an covariance matrix estimate, the correlation matrix $R$ is calculated by normalizing the rows and columns of $C$ by the square roots of its diagonal elements to produce unit entries on the diagonal:
 \begin{equation}\label{eq:precision}
@@
 Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$ \cite{Whittaker:2009}:
+\begin{equation}\label{eq:partial}
+    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
+>>>>>>> a0b18393448df9cb13967143cee42a00d25a7824
+\end{equation}
+Here $\circ$ denotes the entrywise matrix product (Hadamard product) and $I$ is the identity matrix.
+
+\emph{Pairwise partial correlations} are less familiar in the neuroscience community. 
+The partial correlation between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:2009}.
+Partial correlations figure prominantly in probablistic \emph{graphical modelling} wherein the joint distribution is explain by sets of two-way interactions \cite{Whittaker:1990}.
+Under the multivariate Gaussian distribution, zero partial correlations are equivalent to conditional independence of the pair, implying a lack of direct interaction between them \cite{Dempster:1972, Whittaker:1990}.
+In the general case, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1999,Baba:2004}.
+
+Pairwise partial correlations are closely linked to the \emph{precision matrix}, \ie the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1999}.
+Given the covariance estimate $C$, the matrix of partial correlations $P$ is computed by normalizing the rows and columns of the \emph{precision matrix} $C^{-1}$ to produce negative unit entries on the diagonal:
 \begin{equation}\label{eq:partial}

### 2014-05-17 plos.tex

@@
 On the other hand, a correlation matrix can be specified using the \emph{partial correlations} between pairs of the recorded neurons.
-The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.
-When a large fraction of the interacting variables are observed, partial correlations more closely reflect the direct causal effects between components of complex systems than correlations \cite{Whittaker:2009}.
-Partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
-These opportunities have not yet been explored in neurophysiological studies, as most studies have only considered the distributions of pairwise correlations \cite{Zohary:1994, Bair:2001, Smith:2008, Ecker:2010}.
+The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells \cite{Whittaker:1990}.
+Under some assumptions, partial correlations measure conditional independence between variables and may more directly approximate causal effects between components of complex systems than correlations \cite{Whittaker:1990}.
+For this reason, partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.
+These opportunities have not yet been explored in neurophysiological studies where most analyses have only considered the distributions of pairwise correlations \cite{Zohary:1994, Bair:2001, Smith:2008, Ecker:2010}.
 
@@
 where $\circ$ denotes the entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix.
-Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$ \cite{Whittaker:2009}:
-\begin{equation}\label{eq:partial}
-    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
->>>>>>> a0b18393448df9cb13967143cee42a00d25a7824
-\end{equation}
-Here $\circ$ denotes the entrywise matrix product (Hadamard product) and $I$ is the identity matrix.
 
 \emph{Pairwise partial correlations} are less familiar in the neuroscience community. 
-The partial correlation between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:2009}.
+The partial correlation between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}.
 Partial correlations figure prominantly in probablistic \emph{graphical modelling} wherein the joint distribution is explain by sets of two-way interactions \cite{Whittaker:1990}.
 Under the multivariate Gaussian distribution, zero partial correlations are equivalent to conditional independence of the pair, implying a lack of direct interaction between them \cite{Dempster:1972, Whittaker:1990}.
-In the general case, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1999,Baba:2004}.
+In the general case, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1990,Baba:2004}.
+As neural recordings become increasingly dense, partial correlations may prove useful measures of conditional independence and (lack of functinonal connectivity) between pairs of neurons.
 
-Pairwise partial correlations are closely linked to the \emph{precision matrix}, \ie the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1999}.
+Pairwise partial correlations are closely linked to the elements of the \emph{precision matrix}, \ie the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1990}.
+Zero elements in the precision matrix signify zero partial correlation between the two variables.
 Given the covariance estimate $C$, the matrix of partial correlations $P$ is computed by normalizing the rows and columns of the \emph{precision matrix} $C^{-1}$ to produce negative unit entries on the diagonal:
@@
 \end{equation}
+
+As the size of recorded neuronal populations increases, without substantial increases in recording durations, the \emph{condition number} of the sample covariance matrix also increases, making the partial correlation estimates \emph{ill-conditioned}: small errors in covariance estimates translate into much larger errors in partial correlations \cite{Ledoit:2004,Schafer:2005}.
+With massively multineuronal recordings, partial correlations cannot be estimated without \emph{regularization.}
 
@@
 
-Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model $L + D$, where $L$ is the $p\times p$ positive \hl{semidefinite} matrix with low rank and $D$ is a diagonal matrix of independent variances.
+Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model $L + D$, where $L$ is the $p\times p$ positive semidefinite matrix with low rank and $D$ is a diagonal matrix of independent variances.
 The estimator,
@@
 
-Estimator $C_{\sf sparse+latent}$ has two hyperparameters: the number of latent units $d$ and \hl{hyperparameter $\lambda$ regulating the sparsity level}.
+Estimator $C_{\sf sparse+latent}$ has two hyperparameters: the number of latent units $d$ and hyperparameter $\lambda$ regulating the sparsity level.
 $C_{\sf sparse+latent}$ estimates a larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.

### 2014-05-21 plos.tex
minor

@@
 \begin{equation}\label{eq:sample}
-    C_{\sf sample} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
+	C_{\sf sample} = \frac 1 {\nu-1} \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
 \end{equation}
-where $\nu$ denotes degrees of freedom in one neuron's activity within the sample ($\nu=n-1$ if observations are independent).
-For noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.
+where $\nu$ denotes degrees of freedom in one neuron's activity within the sample ($\nu=n$ when observations are independent).
 In this study, we estimate the true mean, $\mu$, using the sample mean, $\bar x$, but seek a better estimate of the covariance matrix than the sample correlation matrix, $C_{\sf sample}$.
@@
 \\
-{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices shown in Row 2.
+{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=500$ drawn from simulated random processes with respective correlation matrices shown in Row 2.
 The structure of the sample correlation matrix is difficult to discern.
@@
 We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes.
-Even with relatively large sample sizes (\eg $n=1000$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (\figref{1}{\,Row 3}).
+Even with relatively large sample sizes (\eg $n=500$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (\figref{1}{\,Row 3}).
 
@@
 \paragraph{Cross-validation}
-To compare the performance of the estimators against each other, we used conventional 10-fold cross validation.
-Briefly, each recording was split into 30 blocks of equal duration.
-The 30 blocks were then grouped randomly into 10 datasets with 3 blocks in each.
-This splitting of the data was chosen to achieve a high degree of independence between the 10 datasets for cross-validaton purposes.
-Splitting the data into smaller blocks could increase dependencies between them due to temporal correlations thereby violating the assumptions of cross-validation.
-Then, each dataset was used as the testing dataset with the rest of the data used for estimating the covariance matrix.
+To compare the performance of the estimators, we used conventional 10-fold cross validation: Trials were randomly divided into 10 subsets with approximately equal numbers of trials of each condition in each subset.
+Each subset was then used as the testing dataset with the rest of the data used for estimating the covariance matrix.
+The resulting validation losses were averaged to produce the \emph{cross-validation loss} reported throughout the paper.
 
 Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}:  The outer loop evaluated the performance of the estimators with the optimal hyperparameter values estimated within the inner loop.
-Hyperparameters were optimized by a two-phase search algorithm: random search to find a good starting point followed by pattern search to find the global minimum.
+Hyperparameters were optimized by a two-phase search algorithm: random search to find a good starting point for the subsequent pattern search to find the global minimum.
 The inner cross-validation loop subdivided the training dataset from the outer loop to perform 10-fold cross-validation in order to evaluate each choice of the hyperparameter values.
@@
 \section*{Supporting Information}
-\subsection*{Appendix: Estimation of variance}
-One decision that was made in this paper was to model a constant covariance matrix across different conditions.
-In other words, a single variance was estimated for each neuron across all stimulus conditions.
-A more commonplace practice, however, is to estimate the correlations separately in each condition and only then to average the correlations coefficients across all the conditions.
-In this appendix we justify this decision.
+\section*{Appendix: Working with multiple stimulus conditions}
+The definitions provided in the main text describe the case when all observations are made in the same condition with the same mean and variances across all observations.
+In this appendix, we extend these definitions to the case when a common correlation matrix needs to be estimated from measurements collected under several stimulus conditions with potentially different means and variances.
+
+Let $T_c$ be a set of indices of measurements $x(t), t\in T_c$ collected in condition $c$. 
+Then the mean population activity in condition $c$ is the vector
+\begin{equation}
+	\bar x_c = \frac 1 {\sharp T_c} \sum\limits_{t\in T_c} x(t)
+\end{equation}
+Here $\sharp T_c$ denotes the cardinality (number of elements) of set $T_c$.
+
+Similar to Eq.~\ref{eq:sample}, the sample covariance matrix for condition $c$ is 
+\begin{equation}
+	C_c=\frac {\nu_c} {\nu_c - 1}\cdot \frac 1 {\sharp T_c} \sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T,
+\end{equation}
+where $\nu_c$ is the number of degrees of freedom in the signal per neuron. When observations are independent $\nu_c=\sharp T_c$.
+
+The diagonal matrix 
+\begin{equation}
+   V_c = C_c \circ I
+\end{equation}
+contains the sample variances on diagonal.
+
+Then the sample correlation matrix in condition $c$ is 
+\begin{equation}
+	R_c = V_c^{-\frac 1 2}C_cV_c^{-\frac 1 2}
+\end{equation}
+
+Let $u_c$ denote the fraction of measurements in condition $c$:
+\begin{equation}
+	u_c = \frac{\sharp T_c}{\sum\limits_c \sharp T_c}
+\end{equation}
+
+Then the common correlation matrix $\bar R$ is estimated by averaging the condition-specific correlations:
+\begin{equation}
+	\bar R = \sum\limits_c u_c R_c
+\end{equation}
+
+Then the condition-specific covariance matrices are reconstituted with the common correlation matrix:
+\begin{equation}
+	\hat C_c = V_c^{\frac 1 2} \bar R V_c^{\frac 1 2}
+\end{equation}
+
+For evaluation, independent datasets $T_c^\prime$ are used to estimate the testing covariance matrices:
+\begin{equation}
+	C_c^\prime = \frac 1{\sharp T_c^\prime}\sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
+\end{equation}
+Note that the means $\bar x_c$ are carried over from the training dataset and not re-estimated from the testing data.
+
+%The regularization methods applied in this study were originally designed for the estimation of a single covariance matrix with the assumption of constant variances across stimulus conditions.
+In contrast, the variance of the spiking activity of cortical neurons is known to depend on the firing rate and on the stimulus condition \cite{Vogels:1989,Ponce:2013}.
+For simplicity, we assumed constant variances across all conditions and estimated a single covariance matrix.
+This assumption allowed direct application of existing covariance estimation techniques.
+In this appendix, we evaluate the effect of this assumption on the estimate of correlations and propose methods for estimating correlations with variable variances.
+First, we evaluated the effect of assuming a constant variance on the validation loss in the same way that Second, the assumption of constant variance confers a downward bias on estimates of correlations between cells with different stimulus responses but not between cells with similar stimulus responses:
 
@@
 
-However, there are arguments for assuming a constant variance as well.
-
 \subsection*{Supplementary Figures}
@@
 \begin{center}
-\includegraphics{./figures/src/Supp1.eps}
+\includegraphics{./figures/Supp01.pdf}
+\end{center}
+\caption{{\bf Estimation of variances}
+	}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics{./figures/Supp02.pdf}
+\end{center}
+\caption{
+{\bf Optimization of hyperparameters of the $C_{\sf sparse+latent}$ estimator.}
+}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics{./figures/src/Supp03.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics{./figures/src/Supp2.eps}
+\includegraphics{./figures/src/Supp04.eps}
 \end{center}

### 2014-06-09 plos.tex
updated variance estimation

@@
 Functional connectivity reflects local synaptic connections, shared inputs from other regions, and endogenous network activity.
-Although functional connectivity is a phenomenological description without a strict mechanistic interpretation, it can be used to generate hypotheses about the anatomical architecture of the neural circuit and test hypotheses about the processing of information at the population level.
+Although functional connectivity is a phenomenological description without a strict mechanistic interpretation, it can be used to generate hypotheses about the anatomical architecture of the neural circuit and to test hypotheses about the processing of information at the population level.
 
@@
 In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of synaptic connectivity \cite{Gerstein:1964}.
-Interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010}.
+Interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various aspects of circuit organization such as the physical distances between the neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010}.
 
@@
 \item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \ie toward a diagonal covariance matrix.
-\item[$C_{\sf factor}$] -- linear shrinkage of the covariances toward a low-rank matrix, representing inputs from unobserved factors (latent units).
+\item[$C_{\sf factor}$] -- a low-rank approximation of the sample covariance matrix, representing inputs from unobserved shared factors (latent units).
 \item[$C_{\sf sparse}$] -- sparse partial correlations, \ie a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
@@
 We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.
-This estimator revealed a sparse network of partial correlations, which we call \sq{interactions} here, between the observed neurons; it also inferred several latent units exerting linear effects on the observed neurons.
-We analyzed these networks of partial correlations and found the following: Whereas noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.
+This estimator revealed a sparse network of partial correlations (\sq{interactions}), between the observed neurons; it also inferred several latent units exerting linear effects on the observed neurons.
+We analyzed these networks of partial correlations and found the following: Whereas significant noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.
 Moreover, we found that these interactions exhibited a stronger relationship to the physical distances and to the differences in preferred orientations than noise correlations.
@@
 where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.
-The usual estimate of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf sample}$ of the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
+
+Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is an $p\times 1$ vector of firing rates in time bin $t$, and an independent unbiased estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix},
 \begin{equation}\label{eq:sample}
-	C_{\sf sample} = \frac 1 {\nu-1} \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
+	C_{\sf sample} = \left[\mbox{cov}\left(x(t)\right)\right]_{t\in T} =\frac 1 n \sum\limits_{t\in T} (x(t)-\bar x)(x(t)-\bar x)^\T,
 \end{equation}
-where $\nu$ denotes degrees of freedom in one neuron's activity within the sample ($\nu=n$ when observations are independent).
-In this study, we estimate the true mean, $\mu$, using the sample mean, $\bar x$, but seek a better estimate of the covariance matrix than the sample correlation matrix, $C_{\sf sample}$.
+where $n$ is the size of $T$, yields an unbiased estimate of the covariance matrix so that $\E{C_{\sf sample}}=\Sigma$.
 
-Given an covariance matrix estimate, the correlation matrix $R$ is calculated by normalizing the rows and columns of $C$ by the square roots of its diagonal elements to produce unit entries on the diagonal:
+Given any covariance matrix estimate $C$, the corresponding correlation matrix $R$ is calculated by normalizing the rows and columns of $C$ by the square roots of its diagonal elements to produce unit entries on the diagonal:
 \begin{equation}\label{eq:precision}
@@
 The partial correlation between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}.
-Partial correlations figure prominantly in probablistic \emph{graphical modelling} wherein the joint distribution is explain by sets of two-way interactions \cite{Whittaker:1990}.
-Under the multivariate Gaussian distribution, zero partial correlations are equivalent to conditional independence of the pair, implying a lack of direct interaction between them \cite{Dempster:1972, Whittaker:1990}.
+Partial correlations figure prominently in probablistic \emph{graphical modelling} wherein the joint distribution is explained by sets of two-way interactions \cite{Whittaker:1990}.
+For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction between them \cite{Dempster:1972, Whittaker:1990}.
 In the general case, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1990,Baba:2004}.
-As neural recordings become increasingly dense, partial correlations may prove useful measures of conditional independence and (lack of functinonal connectivity) between pairs of neurons.
+As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functinonal connectivity) between pairs of neurons.
 
@@
 
-We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
+We considered four regularized estimators based on distinct families of target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
 In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{1}{\,Row 1}).
@@
 
-Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model $L + D$, where $L$ is the $p\times p$ positive semidefinite matrix with low rank and $D$ is a diagonal matrix of independent variances.
+Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model $L + D$, where $L$ is the $p\times p$ positive semidefinite matrix with low-rank and $D$ is a diagonal matrix of independent variances.
 The estimator,
@@
 \end{equation}
-where, as above, $S$ is a sparse matrix and $L$ is a low rank matrix.
+where, as above, $S$ is a sparse matrix and $L$ is a low-rank matrix.
 The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$.
@@
 In a validation procedure, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf sample}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.
-The estimate $C_{\sf sample}^\prime$ is used to
-compute the \emph{validation loss} $\loss{C,C_{\sf sample}^\prime}$, which approximates the loss  $\loss{C,\Sigma}$.
+The estimate $C_{\sf sample}^\prime$ is used to compute the \emph{validation loss} $\loss{C,C_{\sf sample}^\prime}$, which approximates the loss  $\loss{C,\Sigma}$.
 
@@
 \end{equation*}
-With this property of additivity, the validation loss is an unbiased estimate of the loss function:
+Then, once $C$ is fixed, the validation loss is an unbiased estimate of the loss function:
 \begin{equation*}
-    \E[C_{\sf sample}^\prime]{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E[C_{\sf sample}^\prime]{C_{\sf sample}^\prime}}=\loss{C,\Sigma}.
+    \E{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E{C_{\sf sample}^\prime}}=\loss{C,\Sigma}.
 \end{equation*}
@@
 The sample correlation coefficients were largely positive and low.
-The average value of the correlation coefficient across sites ranged from 0.0066 to 0.043, mean 0.018 (\figref{2}{F} and \figref{5}{D}).
+The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051, mean 0.018 (\figref{2}{F} and \figref{5}{D}).
 
@@
     {\bf C--F.} Calcium activity data from an example site.
-    {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. Spiking activity inferred by non-negative deconvolution is shown below the trace.
-    {\bf D.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site.
+    {\bf C.} Representative calcium signals of seven cells, downsampled to 20 Hz, out of the 292 total recorded cells.
+Spiking activity inferred by non-negative deconvolution is shown by red ticks below the trace.
+    {\bf D.} The spatial arrangement and orientation tuning of the 292 cells from the imaged site.
 Only spiking cells are shown.
@@
     {\bf E.} The sample noise correlation matrix of the activity of the neural population. 
-    {\bf F.} Histogram of noise correlation coefficients. The red line indicates the mean correlation coefficient of 0.017.
+    {\bf F.} Histogram of noise correlation coefficients. The red line indicates the mean correlation coefficient of 0.020.
 } \label{fig:2}}
@@
 Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}).
-The respective median differences of the cross-validation loss were 0.019, 0.0027, 0.0031, and 0.0022 nats/cell/bin, significantly greater than zero ($p<10^{-4}$ in each comparison, Wilcoxon signed rank test).
+The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<10^{-2}$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).
 
@@
    {\bf Performance of estimator $C_{\sf sparse+latent}$ with respect to the normal loss function (eq.~\ref{eq:loss}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.}
-    Covariance estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ produced consistently greater validation losses than $C_{\sf sparse+latent}$ ($p<10^{-4}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
+    Covariance estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ produced consistently greater validation losses than $C_{\sf sparse+latent}$ ($p<10^{-2}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
 The box plots indicate the $25^{th}$, $50^{th}$, and $75^{th}$ percentiles with the whiskers extending to the minimum and maximum values after excluding the outliers marked with `+'.
@@
 \end{figure}
-
-Furthermore, estimator $C_{\sf sparse}$ outperformed estimator $C_{\sf factor}$ (Fig.~S\ref{supp:01}) suggesting the greater importance of modeling interactions between pairs of neurons than the common fluctuations of activity across the entire population.
 
@@
 
-In the example site (Fig.~\ref{fig:4}), the sparse component had 89.9\% sparsity (or conversely, 11.1\% connectivity: $connectivity=1-sparsity$) with average node degree of 29.3 (\figref{4}{G}).
+In the example site (Fig.~\ref{fig:4}), the sparse component had 92.8\% sparsity (or conversely, 7.2\% connectivity: $connectivity=1-sparsity$) with average node degree of 20.9 (\figref{4}{G}).
 The average node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$, where $p$ is the number of neurons.
-The low-rank component had rank 71, denoting 71 inferred latent units.
+The low-rank component had rank 72, denoting 72 inferred latent units.
 The number of latent units and the average node degrees generally increased with population size, although these values varied widely between the datasets (\figref{5}{A and B}).
@@
     {\bf C and D.} However, the partial correlation matrices from the two estimates differ substantially.
-    {\bf E.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 89.9\% off-diagonal zeros (bottom-left) and low-rank component of rank 71 (top-right).
-    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations: the gray interval indicates the range of correlations containing 89.9\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
+    {\bf E.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 92.8\% off-diagonal zeros (bottom-left) and low-rank component of rank 72 (top-right).
+    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations: the gray interval indicates the range of correlations containing 92.8\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
 The significant correlations were outside this interval.
-Yet 51.3\% of the interactions inferred by $C_{\sf sparse+latent}$  linked pairs of neurons whose correlation was below the threshold.
-Only the remaining 48.7\% of the interactions overlapped with sample correlations above the threshold.
+Yet 72.3\% of the interactions inferred by $C_{\sf sparse+latent}$  linked pairs of neurons whose correlation was below the threshold.
+Only the remaining 27.7\% of the interactions overlapped with sample correlations above the threshold.
     {\bf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
@@
 This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:4}--\ref{fig:6}).
+\subsection*{Cross-validation with condition-specific variances}
+The loss function defined in the main text assumes that a single covariance matrix is to be estimated.
+However, neuroscientists are commonly interested in estimating a single noise correlation matrix across multiple stimulus conditions when the response means and variances depend on the stimulus \cite{Vogels:1989, Ponce:2013}.
+
+Perhaps the most obvious approach is to minimize a loss function that directly compares the regularized estimate of the correlation matrix to the sample correlation matrix of the validation dataset.
+Unfortunately, the sample correlation coefficient is not generally unbiased \cite{Fisher:1921}.
+Reducing the discrepancy from a biased estimate is not evidence of approaching the true value of the correlation matrix.
+Instead, we adapted the loss function (Eq.~\ref{eq:loss}) to the case where a single estimate of the correlation matrix is combined with condition-specific variance estimates.
+
+The definitions provided in the main text suit the case when all observations are made in the same condition with the same mean and variances across all observations.
+In this appendix, we extend these definitions to the case when a common correlation matrix needs to be estimated from measurements collected under several stimulus conditions with potentially different means and variances.
+
+Let $\{x(t):t\in T_c\}$ be a set of measurements $x(t), t\in T_c$ collected in condition $c$. 
+Here $n_c$ denotes the number of measurements in condition $c$, \ie the cardinality of $T_c$.
+
+Similar to Eq.~\ref{eq:sample}, the sample covariance matrix for condition $c$ is 
+\begin{equation}
+	C_c = \left[\mbox{cov}(x(t))\right]_{t\in T_c}
+	\equiv \frac 1{n_c} \sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T,
+\end{equation}
+
+Then the sample correlation matrix in condition $c$ is 
+\begin{equation}
+	R_c = V_c^{-\frac 1 2}C_cV_c^{-\frac 1 2},
+\end{equation}
+where $V_c = C_c \circ I$ is the diagonal matrix containing the sample variances.
+
+Let $u_c = \frac{n_c}{\sum\limits_c n_c}$ denote the fraction of measurements in condition $c$.
+Then the common correlation matrix $\bar R$ is estimated by averaging the condition-specific correlations:
+\begin{equation}
+	\bar R = \sum\limits_c u_c R_c 
+	= \sum\limits_c u_c V_c^{-\frac 1 2} C_c V_c^{-\frac 1 2}
+	= \sum\limits_c u_c \frac 1 {n_c} \sum\limits_{t \in T_c} z(t)z(t)^\T
+	= \frac 1 n \sum\limits_{t} z(t)z(t)^\T
+	= \mbox{cov}(z(t))
+\end{equation}
+where $z(t) = V_c^{\frac 1 2} \left(x(t) - \bar x_c\right)$ can be recognized as the z-score signal: signal $x(t)$ normalized by the the sample standard deviation specific to the condition.
+Then $\bar R$ is simply the covariance matrix of $z(t)$ across all conditions.
+This  method for computing neural correlations is most familiar to neuroscientists.
+
+Although we are primarily interested in improved estimation of correlations, the loss function can be improved by improving the estimates of the condition-specific means and variances as well.
+Let $\hat V_c$ and $\hat R$ denote the regularized versions of $V_c$ and $\bar R$.
+
+Then the regularized condition-specific covariance matrices are reconstituted as
+\begin{equation}
+	\hat C_c = \hat V_c^{\frac 1 2} \hat R \hat V_c^{\frac 1 2} 
+\end{equation}
+
+To evaluate the regularized estimate $\hat C_c$, independent sets of measurements $T_c^\prime$ of cardinality $n_c^\prime$ are reserved for the computation the condition-specific testing sample covariance matrices $C_c^\prime$:
+\begin{equation}
+	C_c^\prime = \frac 1{n_c^\prime}\sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
+\end{equation}
+Note that the means $\bar x_c$ are carried over from the training dataset and not re-estimated from the testing data. This obviates the need to adjust the estimate for degrees of freedom.
+
+Then the overall loss function is obtained by averaging the loss functions across all the conditions
+\begin{equation} 
+	\mathcal L = \sum\limits_c u_c \loss{\hat C_c, C_c^\prime}
+\end{equation}
+With the normal loss function (Eq.~\ref{eq:loss}), this becomes 
+\begin{equation} 
+\begin{split}
+\mathcal L 
+=& \frac 1 p \sum\limits_c u_c \left[ \ln\det(\hat C_c) + \Tr(\hat C_c^{-1}C_c^\prime)\right] 
+\\
+=& \frac 1 p \sum\limits_c u_c \left[ \ln\det\left(\hat V_c^{\frac 1 2} \hat R \hat V_c^{\frac 1 2}\right) + \Tr\left(\hat V_c^{-\frac 1 2} \hat R^{-1} \hat V_c^{-\frac 1 2}C_c^\prime\right)\right] 
+\\
+=& \frac 1 p \sum\limits_c u_c \ln \det(\hat V_c) 
++ \frac 1 p \ln \det (\hat R) + \frac 1 p \Tr\left(\hat R^{-1} \sum\limits_c u_c\hat V_c^{-\frac 1 2}C_c^\prime\hat V_c^{-\frac 1 2}\right)
+\\
+=& \frac 1 p \sum\limits_c u_c \Tr(\ln \hat V_c) 
++ \frac 1 p \left[\ln \det (\hat R) + \Tr\left(\hat R^{-1} \hat R^\prime\right)\right] 
+\\
+=& \frac 1 p \sum\limits_c u_c \Tr(\ln \hat V_c) 
++ \loss{\hat R, \hat R^\prime}
+\end{split}
+\end{equation}
+
 \paragraph{Covariance estimation}
@@
 \begin{equation}
-(S,L) = \argmin\limits_{\hat S,\hat L} \loss{\hat S-\hat L, C_{\sf sample}} + \alpha\|\hat S\|_1 + \beta \Tr{\hat L}
+(S,L) = \argmin\limits_{\hat S,\hat L} \loss{\hat S-\hat L, C_{\sf sample}} + \alpha\|\hat S\|_1 + \beta \Tr(\hat L)
 \end{equation}
@@
 \newpage
+
 \section*{Supporting Information}
-\section*{Appendix: Working with multiple stimulus conditions}
-The definitions provided in the main text describe the case when all observations are made in the same condition with the same mean and variances across all observations.
-In this appendix, we extend these definitions to the case when a common correlation matrix needs to be estimated from measurements collected under several stimulus conditions with potentially different means and variances.
-
-Let $T_c$ be a set of indices of measurements $x(t), t\in T_c$ collected in condition $c$. 
-Then the mean population activity in condition $c$ is the vector
-\begin{equation}
-	\bar x_c = \frac 1 {\sharp T_c} \sum\limits_{t\in T_c} x(t)
-\end{equation}
-Here $\sharp T_c$ denotes the cardinality (number of elements) of set $T_c$.
-
-Similar to Eq.~\ref{eq:sample}, the sample covariance matrix for condition $c$ is 
-\begin{equation}
-	C_c=\frac {\nu_c} {\nu_c - 1}\cdot \frac 1 {\sharp T_c} \sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T,
-\end{equation}
-where $\nu_c$ is the number of degrees of freedom in the signal per neuron. When observations are independent $\nu_c=\sharp T_c$.
-
-The diagonal matrix 
-\begin{equation}
-   V_c = C_c \circ I
-\end{equation}
-contains the sample variances on diagonal.
-
-Then the sample correlation matrix in condition $c$ is 
-\begin{equation}
-	R_c = V_c^{-\frac 1 2}C_cV_c^{-\frac 1 2}
-\end{equation}
-
-Let $u_c$ denote the fraction of measurements in condition $c$:
-\begin{equation}
-	u_c = \frac{\sharp T_c}{\sum\limits_c \sharp T_c}
-\end{equation}
-
-Then the common correlation matrix $\bar R$ is estimated by averaging the condition-specific correlations:
-\begin{equation}
-	\bar R = \sum\limits_c u_c R_c
-\end{equation}
-
-Then the condition-specific covariance matrices are reconstituted with the common correlation matrix:
-\begin{equation}
-	\hat C_c = V_c^{\frac 1 2} \bar R V_c^{\frac 1 2}
-\end{equation}
-
-For evaluation, independent datasets $T_c^\prime$ are used to estimate the testing covariance matrices:
-\begin{equation}
-	C_c^\prime = \frac 1{\sharp T_c^\prime}\sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
-\end{equation}
-Note that the means $\bar x_c$ are carried over from the training dataset and not re-estimated from the testing data.
-
-%The regularization methods applied in this study were originally designed for the estimation of a single covariance matrix with the assumption of constant variances across stimulus conditions.
-In contrast, the variance of the spiking activity of cortical neurons is known to depend on the firing rate and on the stimulus condition \cite{Vogels:1989,Ponce:2013}.
-For simplicity, we assumed constant variances across all conditions and estimated a single covariance matrix.
-This assumption allowed direct application of existing covariance estimation techniques.
-In this appendix, we evaluate the effect of this assumption on the estimate of correlations and propose methods for estimating correlations with variable variances.
-First, we evaluated the effect of assuming a constant variance on the validation loss in the same way that Second, the assumption of constant variance confers a downward bias on estimates of correlations between cells with different stimulus responses but not between cells with similar stimulus responses:
-
-Let $x_1$ and $x_2$ be random variables with zero means and let their respective variances be $v_1^2$ and $v_2^2$ in condition 1 and $a^2v_1^2$ and $b^2v_2^2$ in condition 2.
-Assume that the correlation coefficient $\rho$ is constant across the two conditions.
-
-Let $\langle \cdot \rangle_i$ denote the expectation in condition $i$. Then
-\begin{align*}
-\langle x_1^2 \rangle_1 = v_1^2 && \langle x_2^2 \rangle_1 = v_2^2 && \rho = \frac{\langle x_1 x_2 \rangle_1}{v_1v_2}\\ 
-\langle x_1^2 \rangle_2 = a^2v_1^2 && \langle x_2^2 \rangle_2 = b^2v_2^2  && \rho = \frac{\langle x_1 x_2 \rangle_2}{ab\,v_1v_2} 
-\end{align*}
-
-What is the effect of assuming constant variances across the conditions on estimating the correlation coefficient?
-Let $0<w<1$ be the fraction of trials in condition 1 and $1-w$ be the fraction of trials in condition 2.
-The respective variances for the two variables across both conditions will be
-\begin{align*}
-\tilde v_1^2 & = w\langle x_1^2 \rangle_1 + (1-w) \langle x_1^2 \rangle_2 = \left[w+(1-w)a^2\right] v_1^2
-\\
-\tilde v_2^2 & = w\langle x_2^2 \rangle_1 + (1-w) \langle x_2^2 \rangle_2 = \left[w+(1-w)b^2\right] v_2^2
-\end{align*}
-The resulting correlation coefficient will be
-\begin{equation*}
-\begin{split}
-\tilde \rho & = \frac{w\langle x_1 x_2 \rangle_1 + (1-w)\langle x_1 x_2 \rangle_2} { \tilde v_1 \tilde v_2}
-= \frac{w\rho + (1-w)ab\rho}{\sqrt{w+(1-w)a^2}\sqrt{w+(1-w)b^2}}
-\\
-& = \rho \sqrt{\frac{w^2+2w(1-w)ab+(1-w)^2a^2b^2}{w^2+w(1-w)(a^2+b^2)+(1-w)^2a^2b^2} }
-\end{split}
-\end{equation*}
-Since $2ab=a^2+b^2$ when $a=b$ and $2ab < a^2 + b^2$ when $a\ne b$, then $\tilde \rho= \rho$ when $a=b$ and $\tilde \rho < \rho$ when $a\ne b$.
-
-In other words, the assumption of constant variance across stimulus condition results in underestimation of their correlation for cells whose variances have different stimulus tunings but not if they have they are tuned similarly.
-This effect may lead to overestimating of the effect of common tuning on correlations (`limited-range correlations').
-
 \subsection*{Supplementary Figures}
@@
 \end{center}
-\caption{{\bf Estimation of variances}
-	}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics{./figures/Supp02.pdf}
-\end{center}
 \caption{
@@
 \begin{center}
-\includegraphics{./figures/src/Supp03.eps}
-\end{center}
-\caption{
-{\bf Performance of estimator $C_{\sf sparse}$ with respect to normal loss (Eq.~\ref{eq:loss}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$}.
-Estimator $C_{\sf sparse}$ outperformed $C_{\sf sample}$ ($p=5\cdot 10^{-6}$) and $C_{\sf factor}$ ($p=.01$), but not $C_{\sf diag}$ ($p=.17$).}
-\label{supp:01}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics{./figures/src/Supp04.eps}
+\includegraphics{./figures/src/Supp3.eps}
 \end{center}

### 2014-06-16 plos.tex
rev 53

@@
 Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.
-We hypothesized that in these dense recordings, the correlation matrix should be best represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.
+We hypothesized that in these dense recordings, the correlation matrix should be best modeled as the combination of a sparse graph of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.
 Indeed, in cross-validation tests, the covariance matrix estimator with this structure consistently outperformed other regularized estimators.
@@
 But reduced estimation error is not the only benefit of regularization.
-The amount of improvement of estimation is greatest when the structure imposed by regularization limits the solution space, but still represents the dominant dependencies in the data.
-Therefore, by finding the most efficient among many regularized estimators, we learn something about the system itself: the structure of the most efficient estimator is a parsimonious representation of the regularities in the data.
+The estimation improvement is greatest when the structure imposed by regularization limits the solution space while still allowing representation of the dominant dependencies in the data.
+Therefore, finding the most efficient among many regularized estimators leads to insights about the system itself: the structure of the most efficient estimator is a parsimonious representation of the regularities in the data.
 
-The necessity and benefits of regularization increase with the size of the recorded population.
+The advantages due to regularization increase with the size of the recorded population.
 With the advent of  big neural data \cite{Alivisatos:2013}, the search for optimal regularization schemes will become increasingly relevant in any model of population activity.
-Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in large scale neural data will entail the search for optimal regularization schemes that may involve combinations of heuristic rules and numerical techniques specially designed for each neural circuit.
+Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in large-scale neural data will entail the search for optimal regularization schemes.
+Such schemes may involve combinations of heuristic rules and numerical techniques specially designed for given types of neural circuits.
 
@@
 
-We then performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals.
+We then performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for representing the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals.
 In our data, the sample correlation coefficients were largely positive and low.
 We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.
-This estimator revealed a sparse network of partial correlations (\sq{interactions}), between the observed neurons; it also inferred several latent units exerting linear effects on the observed neurons.
+This estimator revealed a sparse network of partial correlations (\sq{interactions}), between the observed neurons; it also inferred latent units exerting linear effects on the observed neurons.
 We analyzed these networks of partial correlations and found the following: Whereas significant noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.
@@
 
-Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is an $p\times 1$ vector of firing rates in time bin $t$, and an independent unbiased estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix},
+Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an independent unbiased estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix},
 \begin{equation}\label{eq:sample}
@@
 where $n$ is the size of $T$, yields an unbiased estimate of the covariance matrix so that $\E{C_{\sf sample}}=\Sigma$.
+
+When the mean is estimated from the same sample, $C_{\sf sample}$ is biased toward zero.
+However, in all the cases when the unbiasedness is required in our study, the mean will be estimated from a separate sample.
 
@@
 Partial correlations figure prominently in probablistic \emph{graphical modelling} wherein the joint distribution is explained by sets of two-way interactions \cite{Whittaker:1990}.
-For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction between them \cite{Dempster:1972, Whittaker:1990}.
+For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}.
 In the general case, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1990,Baba:2004}.
@@
 
-Pairwise partial correlations are closely linked to the elements of the \emph{precision matrix}, \ie the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1990}.
+Pairwise partial correlations are closely related to the elements of the \emph{precision matrix}, \ie the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1990}.
 Zero elements in the precision matrix signify zero partial correlation between the two variables.
@@
 
-As the size of recorded neuronal populations increases, without substantial increases in recording durations, the \emph{condition number} of the sample covariance matrix also increases, making the partial correlation estimates \emph{ill-conditioned}: small errors in covariance estimates translate into much larger errors in partial correlations \cite{Ledoit:2004,Schafer:2005}.
-With massively multineuronal recordings, partial correlations cannot be estimated without \emph{regularization.}
+As the size of recorded neuronal populations increases, without substantial increases in recording durations, the \emph{condition number} of the sample covariance matrix also increases, making the partial correlation estimates \emph{ill-conditioned}: small errors in covariance estimates translate into much larger errors in partial correlations.
+With massively multineuronal recordings, partial correlations cannot be estimated without \emph{regularization} \cite{Ledoit:2004,Schafer:2005}.
 
@@
 \\
-{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros.
+{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with four latent units, {\bf C.}  a correlation matrix with 67\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-3 (\ie three latent units) matrix and a sparse matrix with 76\% off-diagonal zeros.
 \\
@@
 \\
-{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) for the five estimators as a function of sample size.
+{\bf Row 5:} Excess losses (Eq.~\ref{eq:loss}) for the five estimators as a function of sample size.
 The error bars indicate the standard error of the mean based on 30 samples.
@@
 The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (\figref{1}{Row 1, A}).
-If sample correlations arise from spurious effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.
+If a predominant majority of sample correlations are spurious, $C_{\sf diag}$ is expected to be more efficient than other estimators.
 
@@
 The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$.
+See Methods for a more detailed explanation.
 The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (\figref{1}{Row 1, D}) \cite{Chandrasekaran:2010,Ma:2013}.
@@
 We next demonstrate how the most efficient among different regularized estimators can reveal the structure of correlations.
-We constructed four $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (\figref{1}{\,Row 2, A--D}, see Methods for details about how these matrices were generated).
+We constructed four families of $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (\figref{1}{\,Row 2, A--D} and Methods).
 We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes.
@@
 \begin{equation}
-     \loss{X,Y} \equiv -\ln(2\pi) - \frac 2 p L(X|Y)
+	\loss{X,Y} \equiv -\ln(2\pi) - \frac 2 {pn} L(X|Y)
 \end{equation}
-Normalization by $p$ makes the values of the loss function comparable across different population sizes.
+Normalization by $p$ and $n$ makes the values of the loss function comparable across different population sizes and sample sizes.
 
@@
 
-The choice of the normal loss function is motivated by mathematical convenience.
+The choice of the normal loss function is motivated, in part, by mathematical convenience.
 We expect that the main conclusions of our study will not change qualitatively with other well behaved loss functions, such as Stein's entropy loss or quadratic loss \cite{James:1961, Fan:2008, Ledoit:2004, Schafer:2005}.
 
-The \emph{excess loss} function is produced by shifting the loss function to zero at its minimum:
-\begin{equation}\label{eq:excess-loss}
-    \eloss{C,\Sigma} = \loss{C,\Sigma}-\loss{\Sigma,\Sigma}
-\end{equation}
-
-We drew 30 independent samples of sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the excess loss for each of the five estimators.
+We drew 30 independent samples of sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the \emph{excess loss} $\loss{C,\Sigma}-\loss{\Sigma,\Sigma}$ for each of the five estimators.
 The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.
 All the regularized estimators produced better estimates (lower excess loss) than the sample covariance matrix.
-However, estimators whose structure matched the true model typically outperformed the other estimators (\figref{1}{\,row 5}).
-There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$, often outperformed the estimators with the matching structure (\figref{1}{\,Row 5, C and D}).
-Second, $C_{\sf sparse+latent}$ often performed as well as $C_{\sf sparse}$ on data from models without latent units because it reliably inferred the correct number of latent units: zero.
+However, estimators whose structure matched the true model outperformed the other estimators (\figref{1}{\,row 5}).
+Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed equally well to $C_{\sf diag}$ because it correctly identified zero factors and only estimated the individual variances.
+Similarly, when the number of latent units was zero (Column C), $C_{\sf sparse+latent}$ performed equally well to $C_{\sf sparse}$ because it correctly identified zero latent units.
 
@@
 With empirical data, the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly.
-Instead, the loss function must be estimated from the sample through \emph{validation}.
+Instead, the loss function may be estimated from the sample through \emph{validation}.
 In a validation procedure, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf sample}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.
@@
 
-The normal loss (Eq.~\ref{eq:loss}) is particularly convenient because it is additive in its second argument, \ie
+The normal loss (Eq.~\ref{eq:loss}) is particularly suitable because it is additive in its second argument, \ie
 \begin{equation*}\label{eq:additivity}
@@
 The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses; their validation losses are biased estimates of the loss function.
+For these loss functions, cross-validation does not generally evaluate the discrepancy from truth.
+
+The loss function above assumes a single covariance matrix across all conditions in the dataset.
+However, neuroscientists often estimate a common correlation matrix across multiple stimulus conditions when the means and variances of responses are conditioned on the stimulus while correlations are assumed constant across all conditions \cite{Vogels:1989, Ponce:2013}.
+In this case, the stimulus-dependent means and variances are estimated from the training dataset along with a common correlation matrix.
+The adaptation of the normal loss function to this case is described in Methods.
 
@@
 Instead, $K$-fold cross-validation can be used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).
-Then each subset is used as the validation sample with the other $K-1$ serving as the training sample.
+Then each subset is used as the testing sample with the other $K-1$ serving as the training sample.
 The validation losses from all such \sq{folds} are averaged to produce the \emph{cross-validation loss}.
 
-Cross-validation loss accurately reproduced the relationships between the excess losses of the estimators, with somewhat greater variability (\figref{1}{Row 6}).
+Cross-validation loss accurately reproduced the relationships between the excess losses of the estimators (\figref{1}{Row 6}).
 
@@
 \paragraph{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
-We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{2}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
+We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{2}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
 This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{2}{C}).
 The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{2}{C}).
-Only cells that produced detectable calcium activity were included in the analysis (For details, see Methods).
+Only cells that produced detectable calcium activity were included in the analysis (see Methods).
 First, 30 repetitions of full-field drifting gratings of 16 directions were presented in random order.
-Each grating was presented for 500 ms, without intervening blanks.
+Each grating was displayed for 500 ms, without intervening blanks.
 This stimulus was used to compute the orientation tuning of the recorded cells (\figref{2}{D}).
@@
 Each grating lasted 1 second and was followed by a 1-second blank.
-The traces were then binned into 150 ms intervals aligned to the stimulus onset for the estimation of the covariance matrix.
+The traces were then binned into 150 ms intervals aligned on the stimulus onset for the estimation of the covariance matrix.
 The sample correlation coefficients were largely positive and low.
-The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051, mean 0.018 (\figref{2}{F} and \figref{5}{D}).
+The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean of 0.018 (\figref{2}{F} and \figref{5}{D}).
 
@@
     {\bf E.} The sample noise correlation matrix of the activity of the neural population. 
-    {\bf F.} Histogram of noise correlation coefficients. The red line indicates the mean correlation coefficient of 0.020.
+    {\bf F.} Histogram of noise correlation coefficients. The red line indicates the mean correlation coefficient of 0.018.
 } \label{fig:2}}
@@
 
-In these dense populations, direct interactions between cells are likely to influence the patterns of population activity.
+In these densely sampled populations, direct interactions between cells are likely to influence the patterns of population activity.
 We therefore hypothesized that covariance matrix estimators that explicitly modeled the partial correlations between pairs of neurons ($C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would have a performance advantage.
@@
 To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
-The hyperparameters of each estimator were optimized by nested cross-validation (See Methods for details).
+The hyperparameters of each estimator were optimized by nested cross-validation (See Methods).
 Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}).
@@
 We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:4} and Fig.~\ref{fig:5}).
-Although the regularized estimates were similar to the sample correlation matrix (\figref{4}{A and D}), the corresponding partial correlation matrices differed substantially (\figref{4}{B and E}).
-The partial correlation matrix of the regularized estimate was the sum of a sparse component and a low-rank component (\figref{4}{C}).
+Although the regularized estimates were similar to the sample correlation matrix (\figref{4}{A and B}), the corresponding partial correlation matrices differed substantially (\figref{4}{C and D}).
+The partial correlation matrix of the regularized estimate was the sum of a sparse component and a low-rank component (\figref{4}{E}).
+With the sparse+latent estimator, we distinguish the pairwise partial correlations computed according to Eq.~\ref{eq:partial}, which include both the sparse and the latent components, from the \emph{sparse partial correlations}, which are computed from the sparse component (Eq.~\ref{eq:c-sl}) alone and represent the local, direct interactions.
+Thus panels \figref{4}{C and D} show partial correlations whereas panel \figref{4}{E and F} depict the sparse partial correlations.
 
@@
 The low-rank component had rank 72, denoting 72 inferred latent units.
-The number of latent units and the average node degrees generally increased with population size, although these values varied widely between the datasets (\figref{5}{A and B}).
-There was an inverse relationship between the number of latent units and the average node degree (\figref{5}{C}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
-This variability may be explained by differences in brain states and recording quality and needs further investigation.
+The number of latent units increased with population size (\figref{5}{A}) but the average number of connections per cell, \ie \emph{node degree} was highly variable (\figref{5}{B}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
+This variability may be explained by differences in brain states and recording quality and warrants further investigation.
 
-The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 31 sites were about 5 times lower than the average sample correlations (\figref{5}{D}).
+The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 27 sites were about 5 times lower than the average sample correlations (\figref{5}{C}).
 This suggests that correlations between neurons build up from large numbers of small interactions with other neurons.
-Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.41 whereas that of the average partial correlations was 0.35.
+Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29, with larger populations exhibiting more similar values of the average partial correlations than the smaller populations.
 
 While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (28.7\% in the example site) of negative values (\figref{4}{F}).
-This fraction was consistent across multiple sites but the fraction of negative interactions increased with the inferred connectivity \figref{5}{F}.
-This increase suggests that negative interactions are inferred only after a sufficient density of positive interactions has been uncovered.
+The fraction of negative interactions increased with the inferred connectivity \figref{5}{D}, suggesting that negative interactions can be inferred only after a sufficient density of positive interactions has been uncovered.
 
@@
 
-Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013}.
-We therefore compared the interactions inferred by $C_{\sf sparse+latent}$ to those obtained by thresholding the sample correlations with the threshold set to match the connectivity of the $C_{\sf sparse+latent}$ graph.
+Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013, Sadovsky:2014}.
+We therefore compared the interactions in the sparse component of $C_{\sf sparse+latent}$ to those obtained from the sample correlations thresholded to match the connectivity of the sparse component of $C_{\sf sparse+latent}$.
 The networks revealed by the two methods differed substantially.
-In the example site with 11.1\% connectivity inferred by $C_{\sf sparse+latent}$, only 48.7\% of connections coincided with above-threshold sample correlations (\figref{4}{F, H, and I}).
+In the example site with 7.2\% connectivity in $C_{\sf sparse+latent}$, only 27.7\% of the connections coincided with the above-threshold sample correlations (\figref{4}{F, H, and I}).
 In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{4}{F}) where high correlations were expected given the rest of the correlation matrix.
-This low overlap between the two types of graphs was consistent across the datasets (\figref{5}{E}).
 
@@
 \\
-    {\bf A.} The number of inferred latent units vs.~population size.
+    {\bf A.} The number of inferred latent units \emph{vs.}~population size.
     {\bf B.} The average node degree for sparse partial correlations vs.~population size.
-    {\bf C.} The average node degree in the sparse component in each site vs.~the number of latent units described by the low-rank component of the estimate.
-    {\bf D.} The average sample correlations  vs.~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
-    {\bf E.} Percent overlap between the graph of highest sample correlations and the graph of interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie $1-sparsity$) of its sparse component.
-    {\bf F.} The percentage of negative interactions vs.~connectivity in the $C_{\sf sparse+latent}$ estimates.
+    {\bf C.} The average sample correlations \emph{vs.}~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
+    {\bf D.} The percentage of negative interactions vs.~connectivity in the $C_{\sf sparse+latent}$ estimates.
 }
@@
 We also examined the relationship of the correlation structure to the differences in orientation preference and to the physical distances between the cell pairs (Fig.\;\ref{fig:6}).
-The partial correlations of the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation (\figref{6}{A}) than the sample correlations.
+The partial correlations from the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation than the sample correlations (\figref{6}{A}).
 They also decreased more rapidly with physical distance, both laterally (\figref{6}{B}) and in depth (\figref{6}{B}).
+The faster dropoff may be a trivial result of regularization: $L_1$-shrinkage of the inverse covariance reduces small partial correlations by a greater fraction than large partial correlations.
+
+The positive and negative interactions in the $C_{\sf sparse+latent}$ estimator were organized differently: Consistent with known differences in the spatial distribution synaptic connections formed by excitatory and inhibitory cells \cite{Adesnik:2010,Fino:2011,Perin:2011}, the density of the positive interactions decreased rapidly with differences in the preferred orientation (\figref{6}{D}), lateral displacements (\figref{6}{E}), and displacement in depth (\figref{6}{F}) whereas the negative interactions were much less selective to orientation (\figref{6}{D}), did not decrease in density with lateral distance on the scale of 150 $\mu m$ (\figref{6}{E}), but did decrease with changes in depth, indicating primarily within-layer connectivity (\figref{6}{F}).
 
@@
     {\bf A--C.} Average partial correlations (red) estimated by $C_{\sf sparse+latent}$ and average sample correlations (black) averaged across multiple imaged sites.
-At each site the correlations were normalized by the respective average correlation shown in \figref{5}{D}.
+At each site the correlations were normalized by the respective average correlation shown in \figref{5}{C}.
 The number of sites, $n$, that qualified to be included in the analysis is indicated in each panel.
@@
 
-The positive and negative interactions in the $C_{\sf sparse+latent}$ estimator were organized differently: The density of the positive interactions fell rapidly with differences in the preferred orientation (\figref{6}{D}), lateral displacements (\figref{6}{E}), and displacement in depth (\figref{6}{F}) whereas the negative interactions were much less selective (\figref{6}{D--F}).
-
 \section*{Discussion}
@@
 Functional connectivity is often represented as a graph of pairwise interactions.
-The goal of many studies of functional connectivity was to estimate  anatomical connectivity from  observed multineuronal spiking activity.
+The goal of many studies of functional connectivity has been to estimate  anatomical connectivity from  observed multineuronal spiking activity.
 For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
@@
 For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.
-Maximum entropy models constrained by observed pairwise correlations also model pairwise coupling between cells \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.
+Maximum entropy models constrained by observed pairwise correlations are among other models with pairwise coupling between cells \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.
 Assuming that the population response follows a multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
-Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.
+Each probabilistic model, fitted to the same data may reveal a completely different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.
 
@@
 The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to anatomy.
-Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.
+Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013, Sadovsky:2014}.
 However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}.
-Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question and we did not attempt them here.
+Until a physiological interpretation of functional connectivity is established, the physiological relevance of such analyses remains in question and we did not attempt them here.
 
-Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the pairwise interactions on the activity of the other neurons, we must record from all neurons with which the pair interacts.
-Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons.
+Inference of the conditional dependencies also depends on the completeness of the recorded population:  To equate conditional dependencies to direct interaction between neurons, we must record from all neurons with which the pair interacts.
+Unobserved portions of the circuit may manifest as conditional dependencies between observed neurons that do not interact.
 For this reason, statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}.
@@
 \paragraph{Model selection}
+Various model selection criteria have been devised to take into account the quality of the fit of a given model and its complexity and to demonstrate which model serves as the best statistical description for observed data.
+Despite its high computational requirements, cross validation is among the most populate model selection methods due to its minimal assumptions about the data generating process \cite{Arlot:2010}.
+
 We evaluated the covariance matrix estimators using the cross-validated normal log likelihood.
@@
 Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:3} and Fig.~S\ref{supp:02}).
-Other probabilistic models, fitted to the same data, would also serve as estimators of the covariance matrix.
+Other probabilistic models, fitted to the same data, could also serve as estimators of the covariance matrix.
 If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.
@@
 The results of model selection must be interpreted with caution.
-As we demonstrated by simulation, with limited data, an incorrect but more constrained model can produce a better estimate than a more complex model with correct representation of dependencies.
+As we demonstrated by simulation, even models with incorrect forms of dependencies can substantially improve estimates.
 Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not necessarily support the conclusion that it reveals a better representation of dependencies in the data.
-This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary low-dimensional target is guaranteed to produce \emph{some} improvement over the best estimate with many degrees of freedom.
+This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary low-dimensional target can consistently outperform a less constrained estimate.
 
@@
 For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{4}{F}) and may reflect inhibitory circuitry.
-Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity, it may accurately reflect average connectivity profiles.
+Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity, it may accurately reflect average synaptic connectivity profiles \cite{Song:2005, Oswald:2008, Perin:2011, Levy:2012}.
 Future experiments utilizing molecular markers for different cell types and follow-up multiple whole-cell \emph{in vitro} experiments \cite{Hofer:2011, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
@@
 Each stimulus condition was presented between 100 and 300 times.
+
 \paragraph{Data processing}
@@
 When the cross-validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.
-This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:4}--\ref{fig:6}).
-\subsection*{Cross-validation with condition-specific variances}
-The loss function defined in the main text assumes that a single covariance matrix is to be estimated.
-However, neuroscientists are commonly interested in estimating a single noise correlation matrix across multiple stimulus conditions when the response means and variances depend on the stimulus \cite{Vogels:1989, Ponce:2013}.
-
-Perhaps the most obvious approach is to minimize a loss function that directly compares the regularized estimate of the correlation matrix to the sample correlation matrix of the validation dataset.
-Unfortunately, the sample correlation coefficient is not generally unbiased \cite{Fisher:1921}.
-Reducing the discrepancy from a biased estimate is not evidence of approaching the true value of the correlation matrix.
-Instead, we adapted the loss function (Eq.~\ref{eq:loss}) to the case where a single estimate of the correlation matrix is combined with condition-specific variance estimates.
-
-The definitions provided in the main text suit the case when all observations are made in the same condition with the same mean and variances across all observations.
-In this appendix, we extend these definitions to the case when a common correlation matrix needs to be estimated from measurements collected under several stimulus conditions with potentially different means and variances.
-
-Let $\{x(t):t\in T_c\}$ be a set of measurements $x(t), t\in T_c$ collected in condition $c$. 
-Here $n_c$ denotes the number of measurements in condition $c$, \ie the cardinality of $T_c$.
-
-Similar to Eq.~\ref{eq:sample}, the sample covariance matrix for condition $c$ is 
-\begin{equation}
-	C_c = \left[\mbox{cov}(x(t))\right]_{t\in T_c}
-	\equiv \frac 1{n_c} \sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T,
-\end{equation}
-
-Then the sample correlation matrix in condition $c$ is 
-\begin{equation}
-	R_c = V_c^{-\frac 1 2}C_cV_c^{-\frac 1 2},
-\end{equation}
-where $V_c = C_c \circ I$ is the diagonal matrix containing the sample variances.
-
-Let $u_c = \frac{n_c}{\sum\limits_c n_c}$ denote the fraction of measurements in condition $c$.
-Then the common correlation matrix $\bar R$ is estimated by averaging the condition-specific correlations:
-\begin{equation}
-	\bar R = \sum\limits_c u_c R_c 
-	= \sum\limits_c u_c V_c^{-\frac 1 2} C_c V_c^{-\frac 1 2}
-	= \sum\limits_c u_c \frac 1 {n_c} \sum\limits_{t \in T_c} z(t)z(t)^\T
-	= \frac 1 n \sum\limits_{t} z(t)z(t)^\T
-	= \mbox{cov}(z(t))
-\end{equation}
-where $z(t) = V_c^{\frac 1 2} \left(x(t) - \bar x_c\right)$ can be recognized as the z-score signal: signal $x(t)$ normalized by the the sample standard deviation specific to the condition.
-Then $\bar R$ is simply the covariance matrix of $z(t)$ across all conditions.
-This  method for computing neural correlations is most familiar to neuroscientists.
-
-Although we are primarily interested in improved estimation of correlations, the loss function can be improved by improving the estimates of the condition-specific means and variances as well.
-Let $\hat V_c$ and $\hat R$ denote the regularized versions of $V_c$ and $\bar R$.
-
-Then the regularized condition-specific covariance matrices are reconstituted as
-\begin{equation}
-	\hat C_c = \hat V_c^{\frac 1 2} \hat R \hat V_c^{\frac 1 2} 
-\end{equation}
-
-To evaluate the regularized estimate $\hat C_c$, independent sets of measurements $T_c^\prime$ of cardinality $n_c^\prime$ are reserved for the computation the condition-specific testing sample covariance matrices $C_c^\prime$:
-\begin{equation}
-	C_c^\prime = \frac 1{n_c^\prime}\sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
-\end{equation}
-Note that the means $\bar x_c$ are carried over from the training dataset and not re-estimated from the testing data. This obviates the need to adjust the estimate for degrees of freedom.
-
-Then the overall loss function is obtained by averaging the loss functions across all the conditions
-\begin{equation} 
-	\mathcal L = \sum\limits_c u_c \loss{\hat C_c, C_c^\prime}
-\end{equation}
-With the normal loss function (Eq.~\ref{eq:loss}), this becomes 
-\begin{equation} 
-\begin{split}
-\mathcal L 
-=& \frac 1 p \sum\limits_c u_c \left[ \ln\det(\hat C_c) + \Tr(\hat C_c^{-1}C_c^\prime)\right] 
-\\
-=& \frac 1 p \sum\limits_c u_c \left[ \ln\det\left(\hat V_c^{\frac 1 2} \hat R \hat V_c^{\frac 1 2}\right) + \Tr\left(\hat V_c^{-\frac 1 2} \hat R^{-1} \hat V_c^{-\frac 1 2}C_c^\prime\right)\right] 
-\\
-=& \frac 1 p \sum\limits_c u_c \ln \det(\hat V_c) 
-+ \frac 1 p \ln \det (\hat R) + \frac 1 p \Tr\left(\hat R^{-1} \sum\limits_c u_c\hat V_c^{-\frac 1 2}C_c^\prime\hat V_c^{-\frac 1 2}\right)
-\\
-=& \frac 1 p \sum\limits_c u_c \Tr(\ln \hat V_c) 
-+ \frac 1 p \left[\ln \det (\hat R) + \Tr\left(\hat R^{-1} \hat R^\prime\right)\right] 
-\\
-=& \frac 1 p \sum\limits_c u_c \Tr(\ln \hat V_c) 
-+ \loss{\hat R, \hat R^\prime}
-\end{split}
-\end{equation}
+This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:4}--\ref{fig:6}).
 
@@
 We solved this problem by adapting the ADMM algorithm to minimizing the loss function with penalty on the $L_1$-norm  of $S$ and the nuclear norm of $L$ \cite{Chandrasekaran:2010,Ma:2013}:
-\begin{equation}
-(S,L) = \argmin\limits_{\hat S,\hat L} \loss{\hat S-\hat L, C_{\sf sample}} + \alpha\|\hat S\|_1 + \beta \Tr(\hat L)
+\begin{equation}\label{eq:ma}
+(S,L) = \argmin\limits_{\hat S,\hat L} \loss{\hat S-\hat L, C_{\sf sample}} + \alpha\|\hat S\|_1 + \beta\Tr(\hat L)
 \end{equation}
@@
 The MATLAB code for these computations is available at {\tt http://github.com/atlab/cov-est}.
+
+\subsubsection*{Cross-validation with condition-specific variances}
+This study required an unbiased estimate of the discrepancy between various correlation matrix estimates and the true value of the correlation matrix.
+This was accomplished by cross-validation whereby regularized covariance matrix estimates computed from training samples were compared to unbiased covariance matrix estimates computed from independent testing samples.
+Loss functions were defined with respect to the covariance matrix rather than the correlation matrix because no unbiased estimate of the correlation matrix exists for fininte sample sizes in the general case \cite{Fisher:1921}.
+In particular, we showed that with the normal loss function (Eq.~\ref{eq:loss}), the validation loss $\loss{C,C_{\sf sample}^\prime}$ is an unbiased estimate of true loss when $C_{\sf sample}^\prime$ is an unbiased estimate.
+However, this approach only worked under the assumption of uniform variances across all conditions included in the training and testing samples.
+Here, we extend the validation loss to the estimation and evaluation of a common correlation matrix across multiple conditions with different variances:
+
+Let $\{x(t):t\in T_c\}$ be a set of measurements collected in condition $c$ from the training sample and let $\{x(t):t \in T_c^\prime\}$ be a set of measurements in condition $c$ from the testing sample, where $T_c$ and $T_c^\prime$ are sets of time samples.
+
+Similar to Eq.~\ref{eq:sample}, the training and testing sample covariance matrices for condition $c$ are
+\begin{equation}
+	C_{c,{\sf sample}}
+	= \frac 1{n_c} \sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
+\end{equation}
+and
+\begin{equation}
+	C_{c,{\sf sample}}^\prime
+	= \frac 1{n_c^\prime} \sum\limits_{t\in T_c^\prime}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
+\end{equation}
+Here $n_c$ and $n_c^\prime$ denote the sizes of $T_c$ and $T_c^\prime$, respectively.
+
+Since $\bar x_c= \frac 1 {n_c} \sum\limits_{t \in T_c}x(t)$ is estimated from the training sample in both estimates, $C_{c,\sf sample}^\prime$ is an unbiased estimate of the true covariance matrix, $\Sigma$, and can be used for validation whereas $C_{c,\sf sample}$ is biased toward zero and is used as input for regularized covariance estimates.
+
+Then the sample correlation matrix in condition $c$ is 
+\begin{equation}
+	R_{c,\sf sample} = V_{c,\sf sample}^{-\frac 1 2}C_{c,\sf sample}V_{c,\sf sample}^{-\frac 1 2}
+\end{equation}
+where $V_{c,\sf sample} = C_{c,\sf sample} \circ I$ is the diagonal matrix containing the sample variances.
+
+Then the common correlation matrix $R_{\sf sample}$ is estimated by averaging the condition-specific correlations:
+\begin{equation}
+	R_{\sf sample} = \frac 1 n\sum\limits_c n_c R_{c,\sf sample}
+	= \frac 1 n \sum\limits_c \sum\limits_{t \in T_c} z(t)z(t)^\T
+\end{equation}
+where $n=\sum\limits_c n_c$, $z(t) = V_{c,\sf sample}^{-\frac 1 2} \left(x(t) - \bar x_c\right)$ is the z-score signal: signal $x(t)$ normalized by the the sample standard deviation specific to the condition.
+Then $R_{\sf sample}$ is simply the covariance matrix of $z(t)$ across all conditions in the training sample.
+
+With average variances across all conditions $V_{\sf sample} = \frac 1 n \sum\limits_c n_c V_{c,\sf sample}$, the average sample covariance matrix across all conditions is reconstituted from the common sample correlation matrix:
+\begin{equation}
+	C_{\sf sample} = V_{\sf sample}^{\frac 1 2} R_{\sf sample} V_{\sf sample}^{\frac 1 2}
+\end{equation}
+
+The covariance matrix estimators $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$ or $C_{\sf sparse+latent}$ convert $C_{\sf sample}$ into its regularized counterpart denoted here as $C_{\sf reg}$.
+Then the regularized correlation matrix is
+\begin{equation}
+	R_{\sf reg} = V_{\sf sample}^{-\frac 1 2}C_{\sf reg}V_{\sf sample}^{-\frac 1 2}
+\end{equation}
+
+$R_{\sf reg}$ is not a true correlation matrix because its diagonal values may not equal 1.
+
+The condition-specific variances were regularized toward their mean values across all conditions by linear shrinkage with shrinkage intensity $\delta$:
+\begin{equation}
+	V_{c,\sf reg} = \delta V_{\sf sample} + (1-\delta) V_{c,\sf sample}
+\end{equation}
+
+Then the regularized condition-specific covariance matrices were reconstituted as
+\begin{equation}
+	C_{c,\sf reg} = V_{c,\sf reg}^{\frac 1 2} R_{\sf reg} V_{c,\sf reg}^{\frac 1 2} 
+\end{equation}
+
+The overall loss function is obtained by averaging the loss functions across all the conditions.
+With the normal loss function (Eq.~\ref{eq:loss}), this becomes 
+\begin{equation}\label{eq:full-loss}
+\begin{split}
+	&\frac 1 p\sum\limits_c \frac{n_c^\prime}{n^\prime} \loss{C_{c,\sf reg}, C_{c,\sf sample}^\prime}
+\\
+=& \frac 1 p \sum\limits_c \frac{n_c^\prime}{n^\prime} \left[ \ln\det(C_{c,\sf reg}) + \Tr(C_{c,\sf reg}^{-1}C_{c,\sf sample}^\prime)\right] 
+\\
+=& \frac 1 p \sum\limits_c \frac{n_c^\prime}{n^\prime} \left[ \ln\det\left(V_{c,\sf reg}^{-\frac 1 2} R_{\sf reg}^{-1} V_{c,\sf reg}^{-\frac 1 2} \right) 
++ \Tr\left(V_{c,\sf reg}^{\frac 1 2} R_{\sf reg} V_{c,\sf reg}^{\frac 1 2} C_{c,\sf sample}^\prime\right)\right] 
+\\
+=& \frac 1 p \sum\limits_c \frac{n_c^\prime}{n^\prime} \ln \det(V_{c,\sf reg}) 
++ \frac 1 p \ln \det (R_{\sf reg}) + \frac 1 p \Tr\left(R_{\sf reg}^{-1} \sum\limits_c \frac{n_c^\prime}{n^\prime} V_{c,\sf reg}^{-\frac 1 2}C_{c,\sf sample}^\prime V_{c,\sf reg}^{-\frac 1 2}\right)
+\\
+=& \frac 1 p \sum\limits_c \frac{n_c^\prime}{n^\prime} \Tr(\ln V_{c,\sf reg}) 
++ \frac 1 p \left[\ln \det (\hat R) + \Tr\left(R_{\sf reg}^{-1} R_{\sf sample}^\prime\right)\right] 
+\\
+=& \frac 1 p \sum\limits_c \frac{n_c^\prime}{n^\prime} \Tr(\ln V_{c,\sf reg}) 
++ \loss{R_{\sf reg}, R_{\sf sample}^\prime}
+\end{split}
+\end{equation}
+Here $n^\prime = \sum\limits_c n_c^\prime$ and
+\begin{equation}
+R_{\sf sample}^\prime 
+= \frac 1 {n^\prime} \sum\limits_c n_c^\prime V_{c,\sf reg}^{-\frac 1 2}C_{c,\sf sample}^\prime V_{c,\sf reg}^{-\frac 1 2}
+= \frac 1 {n^\prime} \sum\limits_c \sum\limits_{t \in T_c^\prime}z_{\sf reg}(t) z_{\sf reg}(t)^\T
+\end{equation}
+where $z_{\sf reg}(t)=V_{c,\sf reg}^{-\frac 1 2}(x-\bar x)$ is the `z-score' trace computed by normalizing the signals from the testing sample by the regularized variance estimates from the training sample.
+$R_{\sf sample}^\prime$ is not a true correlation matrix because its diagonal elements may not equal 1.
+
+The loss function in Eq.~\ref{eq:full-loss} was used in the main results of this study when the correlation matrix was estimated across multiple stimulus conditions.
+The variance regularization parameter $\delta$ was optimized in the nested cross validation loop together with the covariance estimation hyperparameters.
 
@@
 	{\bf Optimization of hyperparameters of the $C_{\sf sparse+latent}$ estimator.}  
+	{\bf A.} The cross-validated loss function (Eq.~\ref{eq:full-loss}) of the data from the example as a function of the values of hyperparameters $\alpha$ and $\beta$ of the minimization problem in Eq.~\ref{eq:ma}.
+In all panels, the red cross marks the optimal value found by the pattern search algorithm described in Methods.
+	{\bf B.} The degree of connectivity ($1-\mbox{sparsity}$) of the sparse component $S$ of the inverse covariance (Eq.~\ref{eq:c-sl}) as a function of $\alpha$ and $\beta$ for the example site.
+	{\bf C.} The number of latent units or the rank of the low-rank component $L$ of the inverse covariance (Eq.~\ref{eq:c-sl}) as a function of hyperparameters $\alpha$ and $\beta$ for the example site.
+	{\bf D.} The loss function as a function of the the connectivity and the number of latent units for the exapmle site.
 }

### 2014-06-21 plos.tex
updated the statistical tests for Fig. 6.

@@
 \\
-Dimitri Yatsenko,$^{1}$, 
+Dimitri Yatsenko$^{1}$, 
 Kre\v{s}imir Josi\'{c}$^{2}$,
@@
 Importantly, the identity and structure of the most efficient estimator informs about the types of dominant dependencies governing the system.
-
 We sought statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.
@@
 Although the two functions are similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf sample})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf sample}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.
+Minimum expected loss may be attaned at a point far removed from the point of maximum likelihood point given the same sample even when the normal distribution can be assumed.
 
@@
 
-In the example site (Fig.~\ref{fig:4}), the sparse component had 92.8\% sparsity (or conversely, 7.2\% connectivity: $connectivity=1-sparsity$) with average node degree of 20.9 (\figref{4}{G}).
-The average node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$, where $p$ is the number of neurons.
+In the example site (Fig.~\ref{fig:4}), the sparse component had 92.8\% sparsity (or conversely, 7.2\% connectivity: $\mbox{connectivity}=1-\mbox{sparsity}$) with average node degree of 20.9 (\figref{4}{G}).
+The average node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $\mbox{degree} = \mbox{connectivity}\cdot(p-1)$, where $p$ is the number of neurons.
 The low-rank component had rank 72, denoting 72 inferred latent units.
@@
 
-We also examined the relationship of the correlation structure to the differences in orientation preference and to the physical distances between the cell pairs (Fig.\;\ref{fig:6}).
-The partial correlations from the $C_{\sf sparse+latent}$ estimates fell more rapidly with the difference in preferred orientation than the sample correlations (\figref{6}{A}).
-They also decreased more rapidly with physical distance, both laterally (\figref{6}{B}) and in depth (\figref{6}{B}).
-The faster dropoff may be a trivial result of regularization: $L_1$-shrinkage of the inverse covariance reduces small partial correlations by a greater fraction than large partial correlations.
+We examined the relationship of the structure of the $C_{\sf sparse+latent}$ estimates to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:6}).
+Five sites with highest connectivitites were included in the analysis.
+Partial correlations were computed using Eq.~\ref{eq:partial} based on the regularized estimate, including both the sparse and the latent component.
+Connectivity was computed as the fraction of pairs of cells in each categority connected by non-zero elements (interactions) in the sparse component of the estimate, distinguishing between the positive and negative connectivities.
 
-The positive and negative interactions in the $C_{\sf sparse+latent}$ estimator were organized differently: Consistent with known differences in the spatial distribution synaptic connections formed by excitatory and inhibitory cells \cite{Adesnik:2010,Fino:2011,Perin:2011}, the density of the positive interactions decreased rapidly with differences in the preferred orientation (\figref{6}{D}), lateral displacements (\figref{6}{E}), and displacement in depth (\figref{6}{F}) whereas the negative interactions were much less selective to orientation (\figref{6}{D}), did not decrease in density with lateral distance on the scale of 150 $\mu m$ (\figref{6}{E}), but did decrease with changes in depth, indicating primarily within-layer connectivity (\figref{6}{F}).
+First, we compared how correlations and connectivity depended on the difference in preferred orientations of pairs of significantly ($\alpha=0.05$) tuned cells.
+The partial correlations decreased more rapidly with differences in tuning than sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test on the linear regression coefficients).
+Positive connectivity decreased rapidly with difference in the preferred orientations ($p<0.005$ in each of the five sites, $t$-test on the logistic regression parameter) whereas negative connectivity did not increase (\figref{6}{D}): The dependence of connectivity on difference in orientation was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test on the logistic regression parameters).
+
+Second, we compared how correlations and connectivity depended on the physical distance separating pairs of cells.
+We distinguished between lateral displacement, \ie the distance in the \emph{xy} plane between cell pairs located at similar depths $\Delta z < 30 \mu m$, and depth displacement, \ie the difference in depth with lateral displacement under $30\,\mu m$.
+Again, the partial correlations decreased more rapidly with both lateral and depth displacements ($p<10^{-6}$ in each of the five sites, both lateral and depth, two-sample $t$-test on the linear regression coefficients).
+Lateral positive connectivity decreased rapidly with distance ($p<10^{-6}$ in each of the five sites, $t$-test on the logistic regression parameter) (\figref{6}{E}), so that cells separated laterally by less than 25 $\mu m$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu m$.
+The negative connectivity did not significantly change with lateral displacement (\figref{6}{E}): The dependence of connectivity on lateral displacement was significantly higher for positive than for negative interactions ($p<0.05$ in each of the five sites, two-sample $t$-test on the logistic regression parameters).
+Positive connectivity decreased significantly with increased depth displacement of pairs of cells ($p<0.05$ in each of the five sites, $t$-test on the logistic regression parameter).
+Negative connectivity on depth displacement did not reach significance in most sites using the same regression test nor could this dependence be shown to be sufficiently different from the positive connectivity.
 
@@
 Then the diagonal matrix $D$ was used as the target of covariance shrinkage (Eq.~\ref{eq:c-diag}) to produce the final regularized estimate.
-The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean squared error whereas we optimized them with respect to the normal loss function (Eq.~\ref{eq:loss}).
 
@@
 
-\subsubsection*{Cross-validation with condition-specific variances}
+\paragraph{Cross-validation with condition-specific variances}
 This study required an unbiased estimate of the discrepancy between various correlation matrix estimates and the true value of the correlation matrix.
@@
 
-Let $\{x(t):t\in T_c\}$ be a set of measurements collected in condition $c$ from the training sample and let $\{x(t):t \in T_c^\prime\}$ be a set of measurements in condition $c$ from the testing sample, where $T_c$ and $T_c^\prime$ are sets of time samples.
+Let $\{x(t):t\in T_c\}$ be a set of measurements collected in condition $c$ from the training sample and let $\{x(t):t \in T_c^\prime\}$ be a set of measurements in condition $c$ from the testing sample, where $T_c$ and $T_c^\prime$ are time bin indices for the training and testing samples, respectively.
+We marked all quantities estimated from the testing sample with a prime mark: \emph{e.g.}\,$C_{c,\sf sample}^\prime$, $n_c^\prime$.
 
@@
 
-Then the sample correlation matrix in condition $c$ is 
+To apply regularization to the correlation matrix, we first reconstruct the common covariance matrix.
+The common correlation matrix $R_{\sf sample}$ is estimated by averaging the condition-specific correlations:
 \begin{equation}
-	R_{c,\sf sample} = V_{c,\sf sample}^{-\frac 1 2}C_{c,\sf sample}V_{c,\sf sample}^{-\frac 1 2}
+	R_{\sf sample}  
+	= \frac 1 n \sum\limits_c n_c \left(V_{c,\sf sample}^{-\frac 1 2}C_{c,\sf sample}V_{c,\sf sample}^{-\frac 1 2}\right)
+        = \frac 1 n \sum\limits_c \sum\limits_{t \in T_c} z(t)z(t)^\T,
 \end{equation}
-where $V_{c,\sf sample} = C_{c,\sf sample} \circ I$ is the diagonal matrix containing the sample variances.
+where $n=\sum\limits_c n_c$ and $V_{c,\sf sample} = C_{c,\sf sample} \circ I$ is the diagonal matrix containing the sample variances.
+Then $R_{\sf sample}$ is simply the covariance matrix of the $z$-score signal
+$z(t) = V_{c,\sf sample}^{-\frac 1 2} \left(x(t) - \bar x_c\right)$ of the training sample.
 
-Then the common correlation matrix $R_{\sf sample}$ is estimated by averaging the condition-specific correlations:
-\begin{equation}
-	R_{\sf sample} = \frac 1 n\sum\limits_c n_c R_{c,\sf sample}
-	= \frac 1 n \sum\limits_c \sum\limits_{t \in T_c} z(t)z(t)^\T
-\end{equation}
-where $n=\sum\limits_c n_c$, $z(t) = V_{c,\sf sample}^{-\frac 1 2} \left(x(t) - \bar x_c\right)$ is the z-score signal: signal $x(t)$ normalized by the the sample standard deviation specific to the condition.
-Then $R_{\sf sample}$ is simply the covariance matrix of $z(t)$ across all conditions in the training sample.
-
-With average variances across all conditions $V_{\sf sample} = \frac 1 n \sum\limits_c n_c V_{c,\sf sample}$, the average sample covariance matrix across all conditions is reconstituted from the common sample correlation matrix:
+For consistency with prior work, we applied regularization to covariance matrices rather than to correlation matrices.
+The common covariance matrix was estimated by scaling $R_{\sf sample}$ by the average variances across conditions $V_{\sf sample} = \frac 1 n \sum\limits_c n_c V_{c,\sf sample}$:
 \begin{equation}
@@
 \end{equation}
+Note that $C_{\sf sample}$ differs from the sample covariance matrix computed without conditioning the variances on $c$ and this computation helps avoid any biases that would be introduced by ignoring changes in variance.
 
 The covariance matrix estimators $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$ or $C_{\sf sparse+latent}$ convert $C_{\sf sample}$ into its regularized counterpart denoted here as $C_{\sf reg}$.
-Then the regularized correlation matrix is
+
+To evaluate the estimators, we regularized the conditioned variances by linear shrinkrage toward their mean value across all conditions.
+This was accomplished by scaling $C_{\sf reg}$ by the diagonal variance adjustment matrix
 \begin{equation}
-	R_{\sf reg} = V_{\sf sample}^{-\frac 1 2}C_{\sf reg}V_{\sf sample}^{-\frac 1 2}
+	Q_c = \delta I + (1-\delta)V_{\sf sample}^{-1} V_{c,\sf sample}
+\end{equation}
+where $\delta \in [0,1]$ is the variance regularization hyperparameter, which was optimized in the inner loop of cross-validation along with the regularization hyperparameters for $C_{\sf reg}$.
+
+Then the regularized condition-specific covariance matrices became
+\begin{equation}
+	C_{c,\sf reg} = Q_c^{\frac 1 2} C_{\sf reg} Q_c^{\frac 1 2} 
 \end{equation}
 
-$R_{\sf reg}$ is not a true correlation matrix because its diagonal values may not equal 1.
-
-The condition-specific variances were regularized toward their mean values across all conditions by linear shrinkage with shrinkage intensity $\delta$:
-\begin{equation}
-	V_{c,\sf reg} = \delta V_{\sf sample} + (1-\delta) V_{c,\sf sample}
+The overall loss function is obtained by averaging the loss functions across all the conditions:
+\begin{equation}\label{eq:full-loss}
+	\frac 1 {\sum\limits_c n_c^\prime}\sum\limits_c n_c^\prime \loss{C_{c,\sf reg}, C_{c,\sf sample}^\prime} 
 \end{equation}
 
-Then the regularized condition-specific covariance matrices were reconstituted as
-\begin{equation}
-	C_{c,\sf reg} = V_{c,\sf reg}^{\frac 1 2} R_{\sf reg} V_{c,\sf reg}^{\frac 1 2} 
-\end{equation}
-
-The overall loss function is obtained by averaging the loss functions across all the conditions.
-With the normal loss function (Eq.~\ref{eq:loss}), this becomes 
-\begin{equation}\label{eq:full-loss}
-\begin{split}
-	&\frac 1 p\sum\limits_c \frac{n_c^\prime}{n^\prime} \loss{C_{c,\sf reg}, C_{c,\sf sample}^\prime}
-\\
-=& \frac 1 p \sum\limits_c \frac{n_c^\prime}{n^\prime} \left[ \ln\det(C_{c,\sf reg}) + \Tr(C_{c,\sf reg}^{-1}C_{c,\sf sample}^\prime)\right] 
-\\
-=& \frac 1 p \sum\limits_c \frac{n_c^\prime}{n^\prime} \left[ \ln\det\left(V_{c,\sf reg}^{-\frac 1 2} R_{\sf reg}^{-1} V_{c,\sf reg}^{-\frac 1 2} \right) 
-+ \Tr\left(V_{c,\sf reg}^{\frac 1 2} R_{\sf reg} V_{c,\sf reg}^{\frac 1 2} C_{c,\sf sample}^\prime\right)\right] 
-\\
-=& \frac 1 p \sum\limits_c \frac{n_c^\prime}{n^\prime} \ln \det(V_{c,\sf reg}) 
-+ \frac 1 p \ln \det (R_{\sf reg}) + \frac 1 p \Tr\left(R_{\sf reg}^{-1} \sum\limits_c \frac{n_c^\prime}{n^\prime} V_{c,\sf reg}^{-\frac 1 2}C_{c,\sf sample}^\prime V_{c,\sf reg}^{-\frac 1 2}\right)
-\\
-=& \frac 1 p \sum\limits_c \frac{n_c^\prime}{n^\prime} \Tr(\ln V_{c,\sf reg}) 
-+ \frac 1 p \left[\ln \det (\hat R) + \Tr\left(R_{\sf reg}^{-1} R_{\sf sample}^\prime\right)\right] 
-\\
-=& \frac 1 p \sum\limits_c \frac{n_c^\prime}{n^\prime} \Tr(\ln V_{c,\sf reg}) 
-+ \loss{R_{\sf reg}, R_{\sf sample}^\prime}
-\end{split}
-\end{equation}
-Here $n^\prime = \sum\limits_c n_c^\prime$ and
-\begin{equation}
-R_{\sf sample}^\prime 
-= \frac 1 {n^\prime} \sum\limits_c n_c^\prime V_{c,\sf reg}^{-\frac 1 2}C_{c,\sf sample}^\prime V_{c,\sf reg}^{-\frac 1 2}
-= \frac 1 {n^\prime} \sum\limits_c \sum\limits_{t \in T_c^\prime}z_{\sf reg}(t) z_{\sf reg}(t)^\T
-\end{equation}
-where $z_{\sf reg}(t)=V_{c,\sf reg}^{-\frac 1 2}(x-\bar x)$ is the `z-score' trace computed by normalizing the signals from the testing sample by the regularized variance estimates from the training sample.
-$R_{\sf sample}^\prime$ is not a true correlation matrix because its diagonal elements may not equal 1.
-
-The loss function in Eq.~\ref{eq:full-loss} was used in the main results of this study when the correlation matrix was estimated across multiple stimulus conditions.
-The variance regularization parameter $\delta$ was optimized in the nested cross validation loop together with the covariance estimation hyperparameters.
+With the normal loss $\loss{\cdot,\cdot}$ (Eq.~\ref{eq:loss}) and the unbiased testing covariance matrix $C_{c,\sf sample}$, the loss function in Eq.~\ref{eq:full-loss} is an unbiased estimate of the true loss.
+Hence, it was used for evaluations reported in \figref{3}.
 

### 2014-06-23 plos.tex
submission candidate

@@
 \usepackage{amsmath}
+
 % amssymb package, useful for mathematical symbols
@@
 The low-rank component had rank 72, denoting 72 inferred latent units.
-The number of latent units increased with population size (\figref{5}{A}) but the average number of connections per cell, \ie \emph{node degree} was highly variable (\figref{5}{B}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
+The number of latent units increased with population size (\figref{5}{A}) but the connectivity was highly variable (\figref{5}{B}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
 This variability may be explained by differences in brain states and recording quality and warrants further investigation.
@@
     {\bf A.} The number of inferred latent units \emph{vs.}~population size.
-    {\bf B.} The average node degree for sparse partial correlations vs.~population size.
+    {\bf B.} The connectivity of the sparse component of partial correlations as a function of population size.
     {\bf C.} The average sample correlations \emph{vs.}~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
@@
 
-We examined the relationship of the structure of the $C_{\sf sparse+latent}$ estimates to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:6}).
-Five sites with highest connectivitites were included in the analysis.
+We examined how the structure of the $C_{\sf sparse+latent}$ estimates related to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:6}).
+Five sites with highest pairwise connectivities were included in the analysis.
 Partial correlations were computed using Eq.~\ref{eq:partial} based on the regularized estimate, including both the sparse and the latent component.
-Connectivity was computed as the fraction of pairs of cells in each categority connected by non-zero elements (interactions) in the sparse component of the estimate, distinguishing between the positive and negative connectivities.
+Connectivity was computed as the fraction of pairs of cells connected by non-zero elements (interactions) in the sparse component of the estimate, distinguishing between the positive and negative connectivities.
 
-First, we compared how correlations and connectivity depended on the difference in preferred orientations of pairs of significantly ($\alpha=0.05$) tuned cells.
-The partial correlations decreased more rapidly with differences in tuning than sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test on the linear regression coefficients).
-Positive connectivity decreased rapidly with difference in the preferred orientations ($p<0.005$ in each of the five sites, $t$-test on the logistic regression parameter) whereas negative connectivity did not increase (\figref{6}{D}): The dependence of connectivity on difference in orientation was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test on the logistic regression parameters).
+First, we analyzed how correlations and connectivity depended on the difference in preferred orientations ($\Delta ori$) of pairs of significantly ($\alpha=0.05$) tuned cells.
+The partial correlations decayed more rapidly with $\Delta ori$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test on the linear regression coefficients).
+Positive connectivity decreased rapidly with $\Delta ori$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression parameter) whereas negative connectivity did not decrease (\figref{6}{D}): The slope in the logistic model of connectivity with respect to $\Delta ori$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test on the logistic regression parameters).
 
 Second, we compared how correlations and connectivity depended on the physical distance separating pairs of cells.
-We distinguished between lateral displacement, \ie the distance in the \emph{xy} plane between cell pairs located at similar depths $\Delta z < 30 \mu m$, and depth displacement, \ie the difference in depth with lateral displacement under $30\,\mu m$.
-Again, the partial correlations decreased more rapidly with both lateral and depth displacements ($p<10^{-6}$ in each of the five sites, both lateral and depth, two-sample $t$-test on the linear regression coefficients).
-Lateral positive connectivity decreased rapidly with distance ($p<10^{-6}$ in each of the five sites, $t$-test on the logistic regression parameter) (\figref{6}{E}), so that cells separated laterally by less than 25 $\mu m$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu m$.
-The negative connectivity did not significantly change with lateral displacement (\figref{6}{E}): The dependence of connectivity on lateral displacement was significantly higher for positive than for negative interactions ($p<0.05$ in each of the five sites, two-sample $t$-test on the logistic regression parameters).
-Positive connectivity decreased significantly with increased depth displacement of pairs of cells ($p<0.05$ in each of the five sites, $t$-test on the logistic regression parameter).
-Negative connectivity on depth displacement did not reach significance in most sites using the same regression test nor could this dependence be shown to be sufficiently different from the positive connectivity.
+We distinguished between lateral distance, $\Delta x$, in the plane parallel to the pia, and vertical distance, $\Delta z$, orthogonal to the pia.
+When considering the dependence on $\Delta x$, the analysis was limited to cell pairs located at the same depth with $\Delta z < 30\,\mu\mbox{m}$; conversely, when considering the dependence on $\Delta z$, only vertically aligned cell pairs with $\Delta x < 30\,\mu\mbox{m}$ were included.
+Again, the partial correlations decayed more rapidly both laterally and vertically than sample correlations ($p<10^{-6}$ in each of the five sites, for both lateral and vertical distances, two-sample $t$-test on the linear regression coefficients).
+Positive connectivity decayed rapidly with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression parameter) (\figref{6}{E}), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$.
+Although the positive connectivity appeared to decay faster with vertical than with lateral distance, the differences in slopes of the respective logistic regression models was not significant with available data.
+The negative connectivity decayed significantly slower with distance in the available range (\figref{6}{E and F}): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test on the logistic regression parameters).
 
@@
     \end{center}
-    \caption{{\bf Dependence of sample correlations and regularized partial correlations on orientation tuning and physical distance.}
-    {\bf A--C.} Average partial correlations (red) estimated by $C_{\sf sparse+latent}$ and average sample correlations (black) averaged across multiple imaged sites.
-At each site the correlations were normalized by the respective average correlation shown in \figref{5}{C}.
-The number of sites, $n$, that qualified to be included in the analysis is indicated in each panel.
-Sites were included if they had at least 20 pairs of neurons in each of the intervals.
-The error bars indicate the standard error of the mean based on on the number of sites $n$.
-    {\bf A.} Average correlations between pairs of neurons tuned to orientation with differences in preferred orientation in the intervals of 0--15$^\circ$, 15--45$^\circ$ and 45--90$^\circ$. 
-    {\bf B.} Average correlations between pairs of neurons located at the same depth ($\pm$30$\mu$m) separated by lateral distances in the intervals of 0--25 $\mu$m, 25--75 $\mu$m, 75--150 $\mu$m, and 150+ $\mu$m.
-    {\bf C.} Average correlations between pairs of neurons displaced laterally by less than 30 $\mu$m separated in depth by distances in the intervals of 0--25 $\mu$m, 25--60 $\mu$m, and 60+ $\mu$m.
-    {\bf D--F.} Normalized connectivity of positive (green) and negative (dark red) interactions from the sparse component obtained from $C_{\sf sparse+latent}$.
-Normalized connectivity was computed as the fraction of pairs connected by interactions of corresponding signs in each interval divided by the fraction of non-zero interactions across the entire site.
+    \caption{{\bf Dependence of sample correlations, regularized partial correlations, and inferred connectivity of the $C_{\sf sparse+latent}$ estimator on the differences in preferred orientations, $\Delta ori$, and physical distances, $\Delta x$ and $\Delta z$.} The error bars mark the standard errors of the means.
+    {\bf A--C.} Mean sample correlations (black) and mean partial correlations (red) estimated by $C_{\sf sparse+latent}$ across $n=5$ imaged sites.
+The values in each bin are normalized by the average across all pairs of their sites.
+    {\bf A.} Mean sample correlations and partial correlations of the $C_{\sf sparse+latent}$ estimator between pairs of neurons tuned to orientation as a function of $\Delta ori$.
+    {\bf B.} Mean sample correlations and mean partial correlations of the $C_{\sf sparse+latent}$ estimator between horizontally aligned pairs of neurons ($\Delta z<30\,\mu$m) as a function of $\Delta x$.
+    {\bf C.} Mean sample correlations and mean partial correlations of the $C_{\sf sparse+latent}$ estimator between vertically aligned pairs of neurons ($\Delta x<30\,\mu$m) as a function of $\Delta z$.
+    {\bf D--F.} Mean positive connectivity (green) and mean negative connectivity (dark red) inferred by the $C_{\sf sparse+latent}$ estimator in $n=5$ imaged sites.
+Positive connectivity in each condition is computed as the fraction of positive partial correlations in the sparse component of the partial correlation matrix meeting the condition normalized by the connectivity in entire site (See \figref{5}{B}).
+Similarly, negative connectivity is the fraction of negative partial correlations similarly conditioned and normalized.
+    {\bf D.} Positive connectivity decreases with $\Delta ori$ whereas negative connectivity does not.
+    {\bf E.} Positive connectivity decays with $\Delta x$ whereas negative connectivity does not. 
+    {\bf F.} Positive connectivity decays with $\Delta z$. Negative connectivity does not decay for small values of $\Delta z$.
     }
@@
 
-In our study, we sought to infer pairwise functional connectivity networks  in cortical microcircuits.
+In this study, we sought to infer pairwise functional connectivity networks  in cortical microcircuits.
 We hypothesized that partial correlations correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.
@@
 Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
-%\Kcomment{I am not sure about the previous argument.}
 
@@
 For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{4}{F}) and may reflect inhibitory circuitry.
-Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity, it may accurately reflect average synaptic connectivity profiles \cite{Song:2005, Oswald:2008, Perin:2011, Levy:2012}.
-Future experiments utilizing molecular markers for different cell types and follow-up multiple whole-cell \emph{in vitro} experiments \cite{Hofer:2011, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
+
+Distances between cells in physical space and in sensory feature space had a stronger effect on the partial correlations estimated by the $C_{\sf sparse+latent}$ estimator than on sample correlations (\figref{6}{A--C}).
+This difference in decay rates supports the idea that correlations are built up from partial correlations in chains of intermediate cells positioned closer and tuned more similarly to one another, with potentially closer correspondence to anatomical connectivity.
+This effect may be at least partially explained by a trivial effect of regularization: $L_1$ penalty employed by the estimator (Eq.~\ref{eq:ma}) suppresses small partial correlations to greater extent than large partial correlations, enhancing the apparent effect of distance and tuning.
+These distances also had distinctly different effects on the positive and negative connectivities inferred by the estimator (\figref{6}{D--F}); these differences are not explained by trivial numerical effects and may reflect geometric and graphical features of local excitatory and inhibitory networks.
+Indeed, the relationships between patterns of positive and negative connectivities inferred by the estimator mirrored some regularities in studies of excitatory and inhibitory synaptic connectivities with respect to distance, cortical layers, and feature tuning \cite{Song:2005, Oswald:2008, Adesnik:2010, Perin:2011, Fino:2011, Hofer:2011, Isaacson:2011, Levy:2012}.
+For example, while excitatory neurons form synapses within highly specific local cliques \cite{Perin:2011}, inhibitory interneurons form synapses with nearly all excitatory cells within local microcircuits \cite{Fino:2011, Hofer:2011, Packer:2011}.
+To further investigate the link between synaptic connectivity and inferred functional connectivity, future experiments will utilize molecular markers for different cell types with follow-up multiple whole-cell \emph{in vitro} experiments \cite{Hofer:2011, Ko:2013} to directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
 Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation.
@@
 \paragraph{Data processing}
-All data were processed in MATLAB using the DataJoint data processing chain toolbox ({\tt datajoint.github.com}) first developed in our lab.
+All data were processed in MATLAB using the DataJoint data processing chain toolbox \cite{DataJoint:2014}.
 
@@
 To compare the performance of the estimators, we used conventional 10-fold cross validation: Trials were randomly divided into 10 subsets with approximately equal numbers of trials of each condition in each subset.
-Each subset was then used as the testing dataset with the rest of the data used for estimating the covariance matrix.
-The resulting validation losses were averaged to produce the \emph{cross-validation loss} reported throughout the paper.
+Each subset was then used as the testing sample with the rest of the data used as the training sample for estimating the covariance matrix.
+The resulting validation losses were averaged to produce the \emph{validation loss} reported throughout the paper.
 
@@
 Thus the size of the training dataset within the inner loop comprised 81\% of the entire recording.
+Fig.~S1 illustrates the dependence of the validation loss on the hyperparameters of the $C_{\sf sparse+latent}$ estimator for the example site shown in Figures \ref{fig:2} and \ref{fig:4} and the optimal value found by the pattern search algorithm.
 
-When the cross-validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.
-This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:4}--\ref{fig:6}).
+When the validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.
+This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:4}--\ref{fig:6}) once its superior performance has been established.
 
@@
 
-The MATLAB code for these computations is available at {\tt http://github.com/atlab/cov-est}.
+The MATLAB code for these computations is available online \cite{Yatsenko:2014c}.
 
@@
 In all panels, the red cross marks the optimal value found by the pattern search algorithm described in Methods.
-	{\bf B.} The degree of connectivity ($1-\mbox{sparsity}$) of the sparse component $S$ of the inverse covariance (Eq.~\ref{eq:c-sl}) as a function of $\alpha$ and $\beta$ for the example site.
+	{\bf B.} The connectivity ($1-\mbox{sparsity}$) of the sparse component $S$ of the inverse covariance (Eq.~\ref{eq:c-sl}) as a function of $\alpha$ and $\beta$ for the example site.
 	{\bf C.} The number of latent units or the rank of the low-rank component $L$ of the inverse covariance (Eq.~\ref{eq:c-sl}) as a function of hyperparameters $\alpha$ and $\beta$ for the example site.

### 2014-06-25 plos.tex
Minor corrections. This was sent to co-authors for review before resubmission.

@@
 But reduced estimation error is not the only benefit of regularization.
-The estimation improvement is greatest when the structure imposed by regularization limits the solution space while still allowing representation of the dominant dependencies in the data.
-Therefore, finding the most efficient among many regularized estimators leads to insights about the system itself: the structure of the most efficient estimator is a parsimonious representation of the regularities in the data.
+Finding the most efficient among many regularized estimators leads to insights about the system itself: the structure of the most efficient estimator is a parsimonious representation of the regularities in the data.
 
@@
 \begin{equation}\label{eq:sample}
-	C_{\sf sample} = \left[\mbox{cov}\left(x(t)\right)\right]_{t\in T} =\frac 1 n \sum\limits_{t\in T} (x(t)-\bar x)(x(t)-\bar x)^\T,
+	C_{\sf sample} = \frac 1 n \sum\limits_{t\in T} (x(t)-\bar x)(x(t)-\bar x)^\T,
 \end{equation}
@@
 \\
-{\bf Row 6:} Cross-validation losses for the five estimators relative to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
+{\bf Row 6:} Validation losses for the five estimators relative to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
     }
@@
 All the regularized estimators produced better estimates (lower excess loss) than the sample covariance matrix.
-However, estimators whose structure matched the true model outperformed the other estimators (\figref{1}{\,row 5}).
+However, estimators whose structure matched the true model outperformed the other estimators (\figref{1}{\,Row 5}).
 Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed equally well to $C_{\sf diag}$ because it correctly identified zero factors and only estimated the individual variances.
@@
 Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}).
-The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<10^{-2}$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).
+The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).
 
@@
 First, we analyzed how correlations and connectivity depended on the difference in preferred orientations ($\Delta ori$) of pairs of significantly ($\alpha=0.05$) tuned cells.
-The partial correlations decayed more rapidly with $\Delta ori$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test on the linear regression coefficients).
-Positive connectivity decreased rapidly with $\Delta ori$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression parameter) whereas negative connectivity did not decrease (\figref{6}{D}): The slope in the logistic model of connectivity with respect to $\Delta ori$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test on the logistic regression parameters).
+The partial correlations decayed more rapidly with $\Delta ori$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test of the difference of the linear regression coefficients).
+Positive connectivity decreased with $\Delta ori$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression coefficient) whereas negative connectivity did not decrease (\figref{6}{D}): The slope in the logistic model of connectivity with respect to $\Delta ori$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficient).
 
@@
 When considering the dependence on $\Delta x$, the analysis was limited to cell pairs located at the same depth with $\Delta z < 30\,\mu\mbox{m}$; conversely, when considering the dependence on $\Delta z$, only vertically aligned cell pairs with $\Delta x < 30\,\mu\mbox{m}$ were included.
-Again, the partial correlations decayed more rapidly both laterally and vertically than sample correlations ($p<10^{-6}$ in each of the five sites, for both lateral and vertical distances, two-sample $t$-test on the linear regression coefficients).
-Positive connectivity decayed rapidly with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression parameter) (\figref{6}{E}), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$.
-Although the positive connectivity appeared to decay faster with vertical than with lateral distance, the differences in slopes of the respective logistic regression models was not significant with available data.
-The negative connectivity decayed significantly slower with distance in the available range (\figref{6}{E and F}): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test on the logistic regression parameters).
+Again, the partial correlations decayed more rapidly both laterally and vertically than sample correlations ($p<10^{-6}$ in each of the five sites, for both lateral and vertical distances, two-sample $t$-test of the difference of the linear regression coefficients).
+Positive connectivity decayed with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression coefficient) (\figref{6}{E}), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$.
+Although the positive connectivity appeared to decay faster with vertical than with lateral distance, the differences in slopes of the respective logistic regression models were not significant with available data.
+The negative connectivity decayed slower with distance in the available range (\figref{6}{E and F}): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficients).
 
@@
     \caption{{\bf Dependence of sample correlations, regularized partial correlations, and inferred connectivity of the $C_{\sf sparse+latent}$ estimator on the differences in preferred orientations, $\Delta ori$, and physical distances, $\Delta x$ and $\Delta z$.} The error bars mark the standard errors of the means.
-    {\bf A--C.} Mean sample correlations (black) and mean partial correlations (red) estimated by $C_{\sf sparse+latent}$ across $n=5$ imaged sites.
-The values in each bin are normalized by the average across all pairs of their sites.
-    {\bf A.} Mean sample correlations and partial correlations of the $C_{\sf sparse+latent}$ estimator between pairs of neurons tuned to orientation as a function of $\Delta ori$.
-    {\bf B.} Mean sample correlations and mean partial correlations of the $C_{\sf sparse+latent}$ estimator between horizontally aligned pairs of neurons ($\Delta z<30\,\mu$m) as a function of $\Delta x$.
-    {\bf C.} Mean sample correlations and mean partial correlations of the $C_{\sf sparse+latent}$ estimator between vertically aligned pairs of neurons ($\Delta x<30\,\mu$m) as a function of $\Delta z$.
-    {\bf D--F.} Mean positive connectivity (green) and mean negative connectivity (dark red) inferred by the $C_{\sf sparse+latent}$ estimator in $n=5$ imaged sites.
-Positive connectivity in each condition is computed as the fraction of positive partial correlations in the sparse component of the partial correlation matrix meeting the condition normalized by the connectivity in entire site (See \figref{5}{B}).
-Similarly, negative connectivity is the fraction of negative partial correlations similarly conditioned and normalized.
+    {\bf A--C.} Normalized mean sample correlations (black) and normalized mean partial correlations (red) estimated by $C_{\sf sparse+latent}$ across $n=5$ imaged sites.
+The values in each bin are normalized by the means across the entire site, which are shown in \figref{5}{C}, to make the effects more comparable across the sites.
+    {\bf A.} Normalized mean sample correlations and normalized mean partial correlations of the $C_{\sf sparse+latent}$ estimator between pairs of neurons tuned to orientation as a function of $\Delta ori$.
+    {\bf B.} Normalized mean sample correlations and normalized mean partial correlations of the $C_{\sf sparse+latent}$ estimator between horizontally aligned pairs of neurons ($\Delta z<30\,\mu$m) as a function of $\Delta x$.
+    {\bf C.} Normalized mean sample correlations and normalized mean partial correlations of the $C_{\sf sparse+latent}$ estimator between vertically aligned pairs of neurons ($\Delta x<30\,\mu$m) as a function of $\Delta z$.
+    {\bf D--F.} Normalized positive connectivity (green) and normalized negative connectivity (dark red) inferred by the $C_{\sf sparse+latent}$ estimator in $n=5$ imaged sites.
+Normalized positive connectivity in each condition is computed as the fraction of positive partial correlations in the sparse component of $C_{\sf sparse+latent}$ for each category of cell pairs, normalized by the connectivities across the entire sites, which are shown in \figref{5}{B}.
+Similarly, normalized negative connectivity is the fraction of negative partial correlations similarly conditioned and normalized.
+The normalization made the effects more comparable across sites.
     {\bf D.} Positive connectivity decreases with $\Delta ori$ whereas negative connectivity does not.
@@
 Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network and external input.
-Therefore, conditionally independent neurons can be inferred to lack a direct causal influence.
+Therefore, conditional independence can signify the absence of a direct causal influence.
 
@@
 \paragraph{Model selection}
-Various model selection criteria have been devised to take into account the quality of the fit of a given model and its complexity and to demonstrate which model serves as the best statistical description for observed data.
-Despite its high computational requirements, cross validation is among the most populate model selection methods due to its minimal assumptions about the data generating process \cite{Arlot:2010}.
+Various model selection criteria have been devised to select between families of models and the optimal subset of variables in a given model family based on observed data.
+Despite its high computational requirements, cross validation is among the most popular model selection methods due to its minimal assumptions about the data generating process \cite{Arlot:2010}.
 
@@
 Distances between cells in physical space and in sensory feature space had a stronger effect on the partial correlations estimated by the $C_{\sf sparse+latent}$ estimator than on sample correlations (\figref{6}{A--C}).
-This difference in decay rates supports the idea that correlations are built up from partial correlations in chains of intermediate cells positioned closer and tuned more similarly to one another, with potentially closer correspondence to anatomical connectivity.
-This effect may be at least partially explained by a trivial effect of regularization: $L_1$ penalty employed by the estimator (Eq.~\ref{eq:ma}) suppresses small partial correlations to greater extent than large partial correlations, enhancing the apparent effect of distance and tuning.
-These distances also had distinctly different effects on the positive and negative connectivities inferred by the estimator (\figref{6}{D--F}); these differences are not explained by trivial numerical effects and may reflect geometric and graphical features of local excitatory and inhibitory networks.
+These differences support the idea that correlations are built up from partial correlations in chains of intermediate cells positioned closer and tuned more similarly to one another, with potentially closer correspondence to anatomical connectivity.
+These differences may also be at least partially explained by a trivial effect of regularization: $L_1$ penalty employed by the estimator (Eq.~\ref{eq:ma}) suppresses small partial correlations to greater extent than large partial correlations, enhancing the apparent effect of distance and tuning.
+Distances between cells in physical space and in sensory feature space also had distinct effects on the positive and negative connectivities inferred by the estimator (\figref{6}{D--F}); we could not think of a numerical artifact that could produce this effect.
+The distinct positive and negative connectivity patterns  may reflect geometric and graphical features of local excitatory and inhibitory networks.
 Indeed, the relationships between patterns of positive and negative connectivities inferred by the estimator mirrored some regularities in studies of excitatory and inhibitory synaptic connectivities with respect to distance, cortical layers, and feature tuning \cite{Song:2005, Oswald:2008, Adesnik:2010, Perin:2011, Fino:2011, Hofer:2011, Isaacson:2011, Levy:2012}.
 For example, while excitatory neurons form synapses within highly specific local cliques \cite{Perin:2011}, inhibitory interneurons form synapses with nearly all excitatory cells within local microcircuits \cite{Fino:2011, Hofer:2011, Packer:2011}.
-To further investigate the link between synaptic connectivity and inferred functional connectivity, future experiments will utilize molecular markers for different cell types with follow-up multiple whole-cell \emph{in vitro} experiments \cite{Hofer:2011, Ko:2013} to directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
+To further investigate the link between synaptic connectivity and inferred functional connectivity, in future experiments, we will use molecular markers for various cell types with follow-up multiple whole-cell \emph{in vitro} experiments \cite{Hofer:2011, Ko:2013} to directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
 Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation.

### 2014-07-09 plos.tex
edits before submission

@@
 Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.
-We hypothesized that in these dense recordings, the correlation matrix should be best modeled as the combination of a sparse graph of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.
+We hypothesized that in these densely sampled recordings, the correlation matrix should be best modeled as the combination of a sparse graph of pairwise partial correlations representing interactions between the observed neuronal pairs and a low-rank component representing common fluctuations and external inputs.
 Indeed, in cross-validation tests, the covariance matrix estimator with this structure consistently outperformed other regularized estimators.
@@
 These interactions reflected the physical distances and orientation tuning properties of cells: The density of positive \sq{excitatory} interactions decreased rapidly with geometric distances and with differences in orientation preference whereas negative \sq{inhibitory} interactions were less selective.
-Because of its superior performance, this \sq{sparse + latent} estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.
+Because of its superior performance, this \sq{sparse + latent} estimator likely provides a more physiologically relevant representation of the functional connectivity in densely sampled recordings than the sample correlation matrix.
 
@@
 \end{equation}
-where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.
+where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons in a time bin of some duration; and $\mu$ is the vector of expected firing rates.
 
@@
 \end{equation}
-where $n$ is the size of $T$, yields an unbiased estimate of the covariance matrix so that $\E{C_{\sf sample}}=\Sigma$.
+where $n$ is the number of time bins in $T$, yields an unbiased estimate of the covariance matrix so that $\E{C_{\sf sample}}=\Sigma$.
 
-When the mean is estimated from the same sample, $C_{\sf sample}$ is biased toward zero.
+When the mean is estimated from the same sample, $C_{\sf sample}$ becomes biased toward zero.
 However, in all the cases when the unbiasedness is required in our study, the mean will be estimated from a separate sample.
@@
 
-\emph{Pairwise partial correlations} are less familiar in the neuroscience community. 
-The partial correlation between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}.
+The \emph{partial correlation} between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}.
 Partial correlations figure prominently in probablistic \emph{graphical modelling} wherein the joint distribution is explained by sets of two-way interactions \cite{Whittaker:1990}.
 For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}.
-In the general case, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1990,Baba:2004}.
+More generally, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1990,Baba:2004}.
 As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functinonal connectivity) between pairs of neurons.
@@
 
-As the size of recorded neuronal populations increases, without substantial increases in recording durations, the \emph{condition number} of the sample covariance matrix also increases, making the partial correlation estimates \emph{ill-conditioned}: small errors in covariance estimates translate into much larger errors in partial correlations.
+As the size of recorded neuronal populations increases, without substantial increases in recording durations, the \emph{condition number} of the sample covariance matrix also increases, making the partial correlation estimates \emph{ill-conditioned}: small errors in covariance estimates translate into much greater errors in partial correlations after the inversion.
 With massively multineuronal recordings, partial correlations cannot be estimated without \emph{regularization} \cite{Ledoit:2004,Schafer:2005}.
@@
     \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
-    {\bf Row 1.} Graphical representations of the target estimates of the four respective regularized covariance matrix estimators.
-Recorded neurons are represented by green spheres and latent units by light-shaded spheres.
+    {\bf Row 1.} Graphical models of the target estimates of the four respective regularized covariance matrix estimators.
+Recorded neurons are represented by the green spheres and latent units by the lightly shaded spheres.
 Edges represent non-zero partial correlations, \ie \sq{interactions}.
@@
 For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix.
-This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} with several latent units.
+This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} due to several latent units.
 	\\
-{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with four latent units, {\bf C.}  a correlation matrix with 67\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-3 (\ie three latent units) matrix and a sparse matrix with 76\% off-diagonal zeros.
+	{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with four latent units, {\bf C.}  a correlation matrix with 67\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-3 matrix (\ie three latent units) and a sparse matrix with 76\% off-diagonal zeros.
 \\
 {\bf Row 3:} Sample correlation matrices calculated from samples of size $n=500$ drawn from simulated random processes with respective correlation matrices shown in Row 2.
-The structure of the sample correlation matrix is difficult to discern.
+The structure of the sample correlation matrix is difficult to discern by eye.
 \\
-{\bf Row 4:} Correlation matrix estimates computed by estimators with matching structure from the same data as the sample correlation matrices in Row 3.
+{\bf Row 4:} Estimates computed from the same data as in Row 3 using structured estimators of the correct type, optimized by cross-validation.
 The regularized estimates are closer to the truth than the sample correlation matrices.
@@
 Estimators with structure that matches the true model converged to zero faster than the other estimators.
-Some exceptions occur when the sample is too small to reveal the structure in the data.
-In these cases estimators with simpler structures can be more efficient than estimators with true structure.
 \\
-{\bf Row 6:} Validation losses for the five estimators relative to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
+{\bf Row 6:} Validation losses for the five estimators relative to the matching estimator.
+Error bars indicate the standard error of the mean based on 30 samples.
+Differences in validation loss approximate differences in true loss.
     }
@@
 The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (\figref{1}{Row 1, A}).
-If a predominant majority of sample correlations are spurious, $C_{\sf diag}$ is expected to be more efficient than other estimators.
+If sample correlations are largely spurious, $C_{\sf diag}$ is expected to be more efficient than other estimators.
 
-Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model $L + D$, where $L$ is the $p\times p$ positive semidefinite matrix with low-rank and $D$ is a diagonal matrix of independent variances.
+Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model $L + D$, where $L$ is the $p\times p$ positive semidefinite matrix of a low rank and $D$ is a diagonal matrix of independent variances.
+This approximation is the basis for \emph{factor analysis} \cite{Anderson:2003}.
 The estimator,
@@
 \end{equation}
-has two hyperparameters: the rank of $L$ (\ie the number of latent factors) and the shrinkage intensity $\lambda$.
+has two hyperparameters: the rank of $L$ (\ie the number of latent factors) and the shrinkage intensity $\lambda$ to shrink the independent variances toward their mean value $\bar D$.
 The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (\figref{1}{Row 1, B}).
@@
 Although the two functions are similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf sample})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf sample}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.
-Minimum expected loss may be attaned at a point far removed from the point of maximum likelihood point given the same sample even when the normal distribution can be assumed.
+Minimum expected loss may be found at a point that is far removed from the point of maximum likelihood given the same data.
 
@@
 \begin{equation*}\label{eq:additivity}
-    \loss{C,z_1} + \loss{C,z_2} \equiv \loss{C,z_1+z_2}
+    \loss{C,Z_1} + \loss{C,Z_2} \equiv \loss{C,Z_1+Z_2}
 \end{equation*}
-Then, once $C$ is fixed, the validation loss is an unbiased estimate of the loss function:
+Then, with fixed $C$, validation loss is an unbiased estimate of true loss:
 \begin{equation*}
@@
 
-The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses; their validation losses are biased estimates of the loss function.
-For these loss functions, cross-validation does not generally evaluate the discrepancy from truth.
+The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses; their validation losses are biased estimates of the loss function and validation does not exactly evaluate a discrepancy from truth.
 
 The loss function above assumes a single covariance matrix across all conditions in the dataset.
-However, neuroscientists often estimate a common correlation matrix across multiple stimulus conditions when the means and variances of responses are conditioned on the stimulus while correlations are assumed constant across all conditions \cite{Vogels:1989, Ponce:2013}.
+However, neuroscientists often estimate a common correlation matrix across multiple stimulus conditions when the variances of responses are conditioned on the stimulus \cite{Vogels:1989, Ponce:2013}.
 In this case, the stimulus-dependent means and variances are estimated from the training dataset along with a common correlation matrix.
-The adaptation of the normal loss function to this case is described in Methods.
+An adaptation of the normal loss function to this case is described in Methods.
 
@@
 Instead, $K$-fold cross-validation can be used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).
-Then each subset is used as the testing sample with the other $K-1$ serving as the training sample.
+Then each subset is used as the testing sample with the other $K-1$ subsets serving as the training sample.
 The validation losses from all such \sq{folds} are averaged to produce the \emph{cross-validation loss}.
 
-Cross-validation loss accurately reproduced the relationships between the excess losses of the estimators (\figref{1}{Row 6}).
+Cross-validation loss accurately reproduced the relationships between the excess losses of the estimators (\figref{1}{Row 6}), confirming that cross-validation can evaluate true loss.
 
@@
 We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{2}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
-This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{2}{C}).
+This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{2}{C and D}).
 The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{2}{C}).
@@
 Each grating lasted 1 second and was followed by a 1-second blank.
-The traces were then binned into 150 ms intervals aligned on the stimulus onset for the estimation of the covariance matrix.
-The sample correlation coefficients were largely positive and low.
-The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean of 0.018 (\figref{2}{F} and \figref{5}{D}).
+The traces were then binned into 150 ms intervals aligned on the stimulus onset for the estimation of the correlation matrix.
+The sample correlation coefficients were largely positive and low (\figref{2}{E and F}).
+The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean across sites of 0.018 (\figref{5}{D}).
 
@@
     {\bf D.} The spatial arrangement and orientation tuning of the 292 cells from the imaged site.
-Only spiking cells are shown.
 The cells' colors indicate their orientation preferences.
@@
     {\bf E.} The sample noise correlation matrix of the activity of the neural population. 
-    {\bf F.} Histogram of noise correlation coefficients. The red line indicates the mean correlation coefficient of 0.018.
+    {\bf F.} Histogram of noise correlation coefficients in one site. The red line indicates the mean correlation coefficient of 0.020.
 } \label{fig:2}}
@@
    {\bf Performance of estimator $C_{\sf sparse+latent}$ with respect to the normal loss function (eq.~\ref{eq:loss}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.}
-    Covariance estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ produced consistently greater validation losses than $C_{\sf sparse+latent}$ ($p<10^{-2}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
+    Covariance estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ produced consistently greater validation losses than $C_{\sf sparse+latent}$ ($p<0.01$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
 The box plots indicate the $25^{th}$, $50^{th}$, and $75^{th}$ percentiles with the whiskers extending to the minimum and maximum values after excluding the outliers marked with `+'.
@@
 \paragraph{Structure of $C_{\sf sparse+latent}$ estimates}
-The estimator $C_{\sf sparse+latent}$ separates two sources of correlations: a network of linear interactions between pairs of neurons and a set of latent units reflecting common fluctuations across the entire population.
-
 We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:4} and Fig.~\ref{fig:5}).
 Although the regularized estimates were similar to the sample correlation matrix (\figref{4}{A and B}), the corresponding partial correlation matrices differed substantially (\figref{4}{C and D}).
-The partial correlation matrix of the regularized estimate was the sum of a sparse component and a low-rank component (\figref{4}{E}).
-With the sparse+latent estimator, we distinguish the pairwise partial correlations computed according to Eq.~\ref{eq:partial}, which include both the sparse and the latent components, from the \emph{sparse partial correlations}, which are computed from the sparse component (Eq.~\ref{eq:c-sl}) alone and represent the local, direct interactions.
-Thus panels \figref{4}{C and D} show partial correlations whereas panel \figref{4}{E and F} depict the sparse partial correlations.
+The estimates separated two sources of correlations: a network of linear interactions expressed by the sparse component of the inverse and latent units expressed by the low-rank components of the inverse (\figref{4}E).
+The sparse partial correlations revealed a network that differed substantially from the network composed of the greatest coefficients in the sample correlation matrix (\figref{4}{F, G, H, and I}).
 
@@
 The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 27 sites were about 5 times lower than the average sample correlations (\figref{5}{C}).
-This suggests that correlations between neurons build up from large numbers of small interactions with other neurons.
+This suggests that correlations between neurons build up from multiple chains of smaller interactions.
 Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29, with larger populations exhibiting more similar values of the average partial correlations than the smaller populations.
@@
     \end{center}
-    \caption{{\bf Example of the structure revealed by $C_{\sf sparse+latent}$ the sparse+latent estimator.}
-    {\bf A and B.} The regularized estimate $C_{\sf sparse+latent}$ closely approximates the sample correlation matrix $C_{\sf sample}$. 
-    {\bf C and D.} However, the partial correlation matrices from the two estimates differ substantially.
+    \caption{{\bf Structure revealed by $C_{\sf sparse+latent}$.}
+    {\bf A, B.} The regularized estimate $C_{\sf sparse+latent}$ closely approximates the sample correlation matrix $C_{\sf sample}$. 
+    {\bf C, D.} The partial correlation matrices from the two estimates differ substantially.
     {\bf E.} The partial correlation matrix of the regularized estimate is decomposed into a sparse component with 92.8\% off-diagonal zeros (bottom-left) and low-rank component of rank 72 (top-right).
-    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations: the gray interval indicates the range of correlations containing 92.8\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
-The significant correlations were outside this interval.
-Yet 72.3\% of the interactions inferred by $C_{\sf sparse+latent}$  linked pairs of neurons whose correlation was below the threshold.
-Only the remaining 27.7\% of the interactions overlapped with sample correlations above the threshold.
-    {\bf G.} A graphical depiction of the positive (green) and negative (magenta) partial correlations as edges between observed neurons.
-The line density is proportional to the magnitude of the correlation.
-    {\bf H.} A subset of neurons from the center of the cluster shown in {\bf G} showing the regularized partial correlations.
-    {\bf I.} The same subset with sample correlations thresholded to match the sparsity of the regularized interactions.
+    {\bf F.} The sparse component of the regularized partial correlation matrix had little resemblance to the sample correlations: The gray region indicates the range of correlations containing 92.8\% of cells pairs, equal to the fraction of zeros in the sparse partial correlation matrix.
+Correlation coefficients outside this interval formed the network of greatest correlations.
+This network differed from the sparse component of the $C_{\sf sparse+latent}$:  Only 27.7\% of the highest correlations coefficients outside the gray regions coincided with interactions inferred by $C_{\sf sparse+latent}$.
+    {\bf G.} A graphical depiction of the positive (green) and negative (magenta) sparse partial correlations as edges between observed neurons.
+The line density is proportional to the magnitude of the partial correlation.
+    {\bf H.} A subset of neurons from the center of the cluster shown in {\bf G} showing the sparse partial correlations.
+    {\bf I.} The same subset of neurons with edges indicating sample correlations thresholded to match the sparsity of the sparse partial correlation.
+These edges correspond to the sample correlation coefficients outside the gray region in panel F.
 }
@@
 Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013, Sadovsky:2014}.
-We therefore compared the interactions in the sparse component of $C_{\sf sparse+latent}$ to those obtained from the sample correlations thresholded to match the connectivity of the sparse component of $C_{\sf sparse+latent}$.
+We therefore compared the interactions in the sparse component of $C_{\sf sparse+latent}$ to those obtained from the sample correlations thresholded to the same level of connectivity.
 The networks revealed by the two methods differed substantially.
 In the example site with 7.2\% connectivity in $C_{\sf sparse+latent}$, only 27.7\% of the connections coincided with the above-threshold sample correlations (\figref{4}{F, H, and I}).
-In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{4}{F}) where high correlations were expected given the rest of the correlation matrix.
+In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{4}{F}) where high correlations should be expected given the rest of the correlation matrix.
 
@@
 Although the positive connectivity appeared to decay faster with vertical than with lateral distance, the differences in slopes of the respective logistic regression models were not significant with available data.
-The negative connectivity decayed slower with distance in the available range (\figref{6}{E and F}): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficients).
+The negative connectivity decayed slower with distance (\figref{6}{E and F}): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficients).
 
@@
     \end{center}
-    \caption{{\bf Dependence of sample correlations, regularized partial correlations, and inferred connectivity of the $C_{\sf sparse+latent}$ estimator on the differences in preferred orientations, $\Delta ori$, and physical distances, $\Delta x$ and $\Delta z$.} The error bars mark the standard errors of the means.
+    \caption{{\bf Dependence of sample correlations, regularized partial correlations, and connectivity inferred by $C_{\sf sparse+latent}$ on the differences in preferred orientations, $\Delta ori$, and physical distances, $\Delta x$ and $\Delta z$.} The error bars mark the standard errors of the means.
+Five sites with highest connectivity (see \figref{5}{B}) were selected for this analysis.
     {\bf A--C.} Normalized mean sample correlations (black) and normalized mean partial correlations (red) estimated by $C_{\sf sparse+latent}$ across $n=5$ imaged sites.
-The values in each bin are normalized by the means across the entire site, which are shown in \figref{5}{C}, to make the effects more comparable across the sites.
-    {\bf A.} Normalized mean sample correlations and normalized mean partial correlations of the $C_{\sf sparse+latent}$ estimator between pairs of neurons tuned to orientation as a function of $\Delta ori$.
-    {\bf B.} Normalized mean sample correlations and normalized mean partial correlations of the $C_{\sf sparse+latent}$ estimator between horizontally aligned pairs of neurons ($\Delta z<30\,\mu$m) as a function of $\Delta x$.
-    {\bf C.} Normalized mean sample correlations and normalized mean partial correlations of the $C_{\sf sparse+latent}$ estimator between vertically aligned pairs of neurons ($\Delta x<30\,\mu$m) as a function of $\Delta z$.
+The values in each bin are normalized by the means across the entire site, shown in \figref{5}{C}, to make the effects more comparable across the sites.
+    {\bf A.} Mean partial correlations in $C_{\sf sparse+latent}$ depend on $\Delta ori$ more strongly than mean sample correlations.
+    {\bf B.} Mean partial correlations in $C_{\sf sparse+latent}$ depend on $\Delta x$ more strongly than mean sample correlations.
+Only horizontally aligned cell pairs with $\Delta z<30\,\mu m$ were considered.
+    {\bf C.} Mean partial correlations in $C_{\sf sparse+latent}$ depend on $\Delta z$ more strongly than mean sample correlations.
+Only vertically aligned cell pairs with $\Delta x<30\,\mu m$ were considered.
 {\bf D--F.} Normalized positive connectivity (green) and normalized negative connectivity (dark red) inferred by the $C_{\sf sparse+latent}$ estimator in $n=5$ imaged sites.
-Normalized positive connectivity in each condition is computed as the fraction of positive partial correlations in the sparse component of $C_{\sf sparse+latent}$ for each category of cell pairs, normalized by the connectivities across the entire sites, which are shown in \figref{5}{B}.
-Similarly, normalized negative connectivity is the fraction of negative partial correlations similarly conditioned and normalized.
-The normalization made the effects more comparable across sites.
+    Here ``connectivity'' refers to the fraction of the non-zero elements in the sparse component $S$ of the $C_{\sf sparse+latent}$ estimator, which describes inferred direct interactions between specific pairs of the recorded neurons.
+Positive and negative connectivities refer to the fractions of the positive and negative partial correlations computed from  $S$, respectively.
+``Normalized positive connectivity'' is the ratio of the positive connectivity for pairs meeting a given condition, \emph{e.g.}~similar tuning with $\Delta ori <15^{\circ}$, to the average connectivity over the entire site.
+Normalized negative connectivity is computed similarly for the negative connectivity.
+The average connectivity across sites is shown in \figref{5}{B} with only the five most connected sites included in the analysis.
+The normalization made the effects of tuning and distance more comparable across sites.
     {\bf D.} Positive connectivity decreases with $\Delta ori$ whereas negative connectivity does not.
-    {\bf E.} Positive connectivity decays with $\Delta x$ whereas negative connectivity does not. 
+    {\bf E.} Positive connectivity decays with $\Delta x$ whereas negative connectivity does not, within the examined range.
     {\bf F.} Positive connectivity decays with $\Delta z$. Negative connectivity does not decay for small values of $\Delta z$.
@@
 Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity.
-For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.
+For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration, membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.
 Maximum entropy models constrained by observed pairwise correlations are among other models with pairwise coupling between cells \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.
@@
 In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies.
-This may be the reason that partial correlations have not previously been used to describe the functional connectivity in cortical populations until this study.
+Perhaps for this reason, partial correlations have not, until now, been used to describe the functional connectivity in cortical populations.
 
@@
 The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.
-Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
-These techniques may provide more incisive measurements of functional connectivity than classical electrophysiological recordings.
+Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
+These techniques may provide more incisive measurements of functional connectivity than electrophysiological recordings.
 
@@
 Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we \emph{a priori} favored solutions with sparse partial correlations.
-The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution, where the precision matrix specifies the interactions between cell pairs.
-Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
+Under the assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
 
 \paragraph{Functional connectivity as coactivations}
-Another approach to describing the functional connectivity of a circuit is to isolate patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}.
+Another approach to describing the functional connectivity of a circuit is to isolate individual patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}.
 Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}.
@@
 However, an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.
-In our study, the functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.
+In our study, the functional connectivity solely through modes of coactivations was represented by the factor analysis-based estimator $C_{\sf factor}$.
 
@@
 The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
-As techniques have improved to allow recording from larger neuronal populations, experimental neuroscientists have addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.
+As techniques have improved to allow recording from larger neuronal populations, experimental neuroscientists have addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted \cite{Tkacik:2013}.
 However, ambitious projects, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
 Simply increasing recording duration will not be sufficient, and the problem must be addressed by using regularized estimators.
-Regularization biases the solution toward a small subspace to counteract the effect of  sampling noise in the empirical data.
+Regularization biases the solution toward a small subspace in order to counteract the effect of  sampling noise in the empirical data.
 However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hinders interpretation.
@@
 For example, Ganmor et al.
-\cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in retina.
-Generalized linear models typically employ $L_1$ penalty terms to constrain the solution space and to effectively reduce the dimensionality of the solution \cite{Pillow:2008}.
+\cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in the retina.
+As another example of regularization, generalized linear models typically employ $L_1$ penalty terms to constrain the solution space and to effectively reduce the dimensionality of the solution \cite{Pillow:2008}.
 
-In our study, regularization was also accomplished by dimensionality reduction (feature selection) schemes to produce sparse, constrained solutions.
+In our study, regularizations were accomplished by dimensionality reduction (feature selection) schemes to produce sparse, constrained solutions.
+Only the most efficient scheme was considered in the analysis of functional connectivity.
 
 \paragraph{Model selection}
-Various model selection criteria have been devised to select between families of models and the optimal subset of variables in a given model family based on observed data.
-Despite its high computational requirements, cross validation is among the most popular model selection methods due to its minimal assumptions about the data generating process \cite{Arlot:2010}.
+Various model selection criteria have been devised to select between families of models and the optimal subsets of variables in a given model family based on observed data.
+Despite its computational requirements, cross-validation is among the most popular model selection methods due to its minimal assumptions about the data generating process \cite{Arlot:2010}.
 
-We evaluated the covariance matrix estimators using the cross-validated normal log likelihood.
+We evaluated the covariance matrix estimators a loss function based on multivariate normal log likelihood.
 However, this does not limit the applicability of its conclusions to normal distributions.
@@
 The results of model selection must be interpreted with caution.
-As we demonstrated by simulation, even models with incorrect forms of dependencies can substantially improve estimates.
+As we demonstrated by simulation, even models with incorrect forms of dependencies can substantially improve estimates (\figref{1}).
 Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not necessarily support the conclusion that it reveals a better representation of dependencies in the data.
@@
 \paragraph{Data processing}
-All data were processed in MATLAB using the DataJoint data processing chain toolbox \cite{DataJoint:2014}.
+All data were processed in MATLAB using the DataJoint data processing chain toolbox (http://datajoint.github.com).
 
@@
 \paragraph{Cross-validation}
-To compare the performance of the estimators, we used conventional 10-fold cross validation: Trials were randomly divided into 10 subsets with approximately equal numbers of trials of each condition in each subset.
+To compare the performance of the estimators, we used conventional 10-fold cross-validation: Trials were randomly divided into 10 subsets with approximately equal numbers of trials of each condition in each subset.
 Each subset was then used as the testing sample with the rest of the data used as the training sample for estimating the covariance matrix.
@@
 
-The MATLAB code for these computations is available online \cite{Yatsenko:2014c}.
+The MATLAB code for these computations is available online at http://github.com/atlab/cov-est.
 
@@
 This was accomplished by cross-validation whereby regularized covariance matrix estimates computed from training samples were compared to unbiased covariance matrix estimates computed from independent testing samples.
-Loss functions were defined with respect to the covariance matrix rather than the correlation matrix because no unbiased estimate of the correlation matrix exists for fininte sample sizes in the general case \cite{Fisher:1921}.
+Loss functions were defined with respect to the covariance matrix rather than the correlation matrix because no unbiased estimate of the correlation matrix exists for finite sample sizes in the general case \cite{Fisher:1921}.
 In particular, we showed that with the normal loss function (Eq.~\ref{eq:loss}), the validation loss $\loss{C,C_{\sf sample}^\prime}$ is an unbiased estimate of true loss when $C_{\sf sample}^\prime$ is an unbiased estimate.

### 2014-07-10 plos.tex
minor edits

@@
 
-%\paragraph{Conceptual questions and basic approach}
 What structures of correlation matrices best describe the multineuronal activity in specific circuits and in specific brain states?
@@
 % Results and Discussion can be combined.
-\paragraph{Covariance estimation}
+\paragraph{Covariance estimation.}
 The covariance matrix is defined as
@@
 
-\paragraph{Simulation}
+\paragraph{Simulation.}
 We next demonstrate how the most efficient among different regularized estimators can reveal the structure of correlations.
@@
 
-\paragraph{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
+\paragraph{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data.}
 We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{2}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
@@
 
-\paragraph{Structure of $C_{\sf sparse+latent}$ estimates}
+\paragraph{Structure of $C_{\sf sparse+latent}$ estimates.}
 We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:4} and Fig.~\ref{fig:5}).
@@
 
-\paragraph{Relationship of $C_{\sf sparse+latent}$ to orientation tuning and physical distances}
+\paragraph{Relationship of $C_{\sf sparse+latent}$ to orientation tuning and physical distances.}
 
@@
 \section*{Discussion}
-\paragraph{Functional connectivity as a network of pairwise interactions}
+\paragraph{Functional connectivity as a network of pairwise interactions.}
 Functional connectivity is often represented as a graph of pairwise interactions.
@@
 
-\paragraph{Functional connectivity as coactivations}
+\paragraph{Functional connectivity as coactivations.}
 Another approach to describing the functional connectivity of a circuit is to isolate individual patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}.
@@
 
-\paragraph{Combining pairwise interactions and coactivations}
+\paragraph{Combining pairwise interactions and coactivations.}
 In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables \cite{Koster:2013},  high-order correlations \cite{Ganmor:2011}, or global network fluctuations \cite{Tkacik:2013}.
@@
 
-\paragraph{Addressing ill-posedness}
+\paragraph{Addressing ill-posedness.}
 Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.
@@
 
-\paragraph{Model selection}
+\paragraph{Model selection.}
 Various model selection criteria have been devised to select between families of models and the optimal subsets of variables in a given model family based on observed data.
@@
 
-\paragraph{Physiological interpretation and future directions}
+\paragraph{Physiological interpretation and future directions.}
 
-Here we showed that among several models a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix for cortical microcircuits.
+We showed that among several models a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix for cortical microcircuits.
 This finding is valuable in itself: improved estimates of the noise covariance matrix for large datasets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Ecker:2011}
@@
 Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.
-Importantly, the inferred functional interactions were substantially different from the network of the most significant sample correlations \figref{4}{F, H, and I}.
+Importantly, the inferred functional interactions were substantially different from the network of the highest sample correlations \figref{4}{F, H, and I}.
 For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{4}{F}) and may reflect inhibitory circuitry.
@@
 These differences support the idea that correlations are built up from partial correlations in chains of intermediate cells positioned closer and tuned more similarly to one another, with potentially closer correspondence to anatomical connectivity.
-These differences may also be at least partially explained by a trivial effect of regularization: $L_1$ penalty employed by the estimator (Eq.~\ref{eq:ma}) suppresses small partial correlations to greater extent than large partial correlations, enhancing the apparent effect of distance and tuning.
+These differences may also be at least partially explained by a trivial effect of regularization: the $L_1$ penalty applied by the estimator (Eq.~\ref{eq:ma}) suppresses small partial correlations to greater extent than large partial correlations, enhancing the apparent effect of distance and tuning.
 Distances between cells in physical space and in sensory feature space also had distinct effects on the positive and negative connectivities inferred by the estimator (\figref{6}{D--F}); we could not think of a numerical artifact that could produce this effect.
 The distinct positive and negative connectivity patterns  may reflect geometric and graphical features of local excitatory and inhibitory networks.
-Indeed, the relationships between patterns of positive and negative connectivities inferred by the estimator mirrored some regularities in studies of excitatory and inhibitory synaptic connectivities with respect to distance, cortical layers, and feature tuning \cite{Song:2005, Oswald:2008, Adesnik:2010, Perin:2011, Fino:2011, Hofer:2011, Isaacson:2011, Levy:2012}.
+Indeed, the relationships between patterns of positive and negative connectivities inferred by the estimator mirrored the properties of excitatory and inhibitory synaptic connectivities with respect to distance, cortical layers, and feature tuning \cite{Song:2005, Oswald:2008, Adesnik:2010, Perin:2011, Fino:2011, Hofer:2011, Isaacson:2011, Levy:2012}.
 For example, while excitatory neurons form synapses within highly specific local cliques \cite{Perin:2011}, inhibitory interneurons form synapses with nearly all excitatory cells within local microcircuits \cite{Fino:2011, Hofer:2011, Packer:2011}.
-To further investigate the link between synaptic connectivity and inferred functional connectivity, in future experiments, we will use molecular markers for various cell types with follow-up multiple whole-cell \emph{in vitro} experiments \cite{Hofer:2011, Ko:2013} to directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
-Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation.
+To further investigate the link between synaptic connectivity and inferred functional connectivity, in future experiments, we will use molecular markers for various cell types with follow-up multiple whole-cell \emph{in vitro} recordings \cite{Hofer:2011, Ko:2013} to directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
+Finally, the latent units inferred by the estimator can be analyzed for their physiological functions.
 For example, these latent units may be modulated under different brain states (e.g.
@@
 % authors may use "Analysis" 
-\paragraph{Ethics statement}
+\paragraph{Ethics statement.}
 All procedures were conducted in accordance with the ethical guidelines of the National Institutes of Health and were approved by the Baylor College of Medicine IACUC.
 
-\paragraph{Surgery and two-photon imaging}
-The surgical procedures and data acquisition were performed as descrbed in \cite{Cotton:2013}.
+\paragraph{Surgery and two-photon imaging.}
+The surgical procedures and data acquisition were performed as described in \cite{Cotton:2013}.
 Briefly, C57BL/6J mice (aged p40--60) were used.
@@
 Calcium imaging began 1 hour after dye injection.
-All imaging was performed using the 3D-RAMP two-photon microscope as described in \cite{Cotton:2013}.
+All imaging was performed using 3D-RAMP two-photon microscopy \cite{Cotton:2013}.
 First, a 3D stack was acquired and cells were manually segmented.
@@
 
-\paragraph{Visual stimulus}
+\paragraph{Visual stimulus.}
 The visual stimulus consisted of full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s.
 Two sets of stimuli were presented for each imaging site: the first to map directional tuning and the second to estimate noise correlations.
-Directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen equally spaced directions of motion, 500 ms per direction, for 3 min without blanks.
+Directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen directions of motion, 500 ms per direction, without blanks, with 12--30 trials for each direction of motion.
 The data for covariance estimation were collected during presentations of full-field drifting gratings with the same parameters as those used in directional tuning except only two directions (in 9 datasets) or five directions (in 22 datasets) were used and the presentations lasted 1 second and were separated by 1-second blanks.
@@
 
-\paragraph{Data processing}
+\paragraph{Data processing.}
 All data were processed in MATLAB using the DataJoint data processing chain toolbox (http://datajoint.github.com).
 
-The collected fluorescent traces were deconvolved to reconstruct the firing rates for each neuron.
-First, the first principal component was subtracted from the traces, which reduced common mode noise related to small cardiovascular movements \cite{Cotton:2013}.
-The resulting traces were low-pass filtered below 0.1 Hz and downsampled to 20 Hz (\figref{2}{C}).
-Firing rates were estimated using a fast non-negative deconvolution algorithm \cite{Vogelstein:2010}.
+The measured fluorescent traces were deconvolved to reconstruct the firing rates for each neuron: First, the first principal component was subtracted from the traces, which reduced common mode noise related to small cardiovascular movements \cite{Cotton:2013}.
+The resulting traces were high-pass filtered above 0.1 Hz and downsampled to 20 Hz (\figref{2}{C}).
+Then, the firing rates were estimated using a fast non-negative deconvolution algorithm \cite{Vogelstein:2010}.
 
 Orientation tuning was computed by fitting the mean firing rates in response to gratings drifting in directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
-The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times.
-The $p$-value of the fit was computed as the fraction of the permuted datasets for which the $R^2$ value of the tuning function fit exceeded that of the real data.
+The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times; the $p$-values of the fits were computed as the fraction of the permuted datasets for which the $R^2$ value of the tuning function fit equaled or exceeded that of the real data.
 Cells were considered tuned with $p$-values below 0.05.
@@
 
-\paragraph{Cross-validation}
+\paragraph{Cross-validation.}
 To compare the performance of the estimators, we used conventional 10-fold cross-validation: Trials were randomly divided into 10 subsets with approximately equal numbers of trials of each condition in each subset.
@@
 
-Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}:  The outer loop evaluated the performance of the estimators with the optimal hyperparameter values estimated within the inner loop.
+Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}:  The outer loop evaluated the performance of the estimators with the hyperparameter values optimized by cross-validation within the inner loop.
 Hyperparameters were optimized by a two-phase search algorithm: random search to find a good starting point for the subsequent pattern search to find the global minimum.
@@
 
-When the validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.
-This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:4}--\ref{fig:6}) once its superior performance has been established.
+When the validation loss was not required, only the inner loop of cross-validation was used on the entire dataset.
+This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:4}--\ref{fig:6}) once its superior performance has been established.
 
-\paragraph{Covariance estimation}
-Within the inner loop of cross-validation, covariance matrix estimation was performed with fixed hyperparameter values provided by the search algorithm.
-The computation of regularized estimators only required the sample covariance matrix $C_{\sf sample}$ of the training dataset.
+\paragraph{Covariance estimation.}
+Within the inner loop of cross-validation, regularized covariance matrix estimation required only the sample covariance matrix $C_{\sf sample}$ of the training dataset and the hyperparameter values provided by the outer loop.
 
@@
 
-Estimator $C_{\sf factor}$ used two hyperparameters: the number of latent factors $d$ and the shrinkage intensity $\lambda \in [0, 1]$.  
-The matrix $L$ of rank $d$ and the diagonal matrix of individual variances $D$ were computed by solving the minimization problem
+In estimator $C_{\sf factor}$ (Eq.~\ref{eq:c-factor}), the low-rank matrix $L$ and the diagonal matrix $D$ are found by solving the minimization problem
 \begin{equation}
@@
 \end{equation}
-which we solved by an expectation-maximization (EM) algorithm.
-Under our chosen loss function (Eq.~\ref{eq:loss}), this is equivalent to maximum likelihood estimation of $L$ and $D$ under the multivariate Gaussian distribution.
-The final regularized estimate is obtained by linear shrinkage toward the factor model (Eq.~\ref{eq:c-factor}).
+by an expectation-maximization (EM) algorithm.  The final regularized estimate is obtained by linear shrinkage of the diagonal matrix $D$ toward its mean diagonal value.
 
@@
 This problem formulation is known as \emph{graphical lasso} \cite{Meinshausen:2006, Friedman:2008}.
-To solve this minimization problem, we adapted the alternative-direction method of multipliers (ADMM) algorithm \cite{Ma:2013}.
+To solve this minimization problem, we adapted the alternative-direction method of multipliers (ADMM) \cite{Ma:2013}.
 Then the final estimate is the inverse of $S$ (Eq.~\ref{eq:c-sparse}).
@@
 
-Estimator $C_{\sf sparse+latent}$ has two hyperparameters: the number of latent units $d$ and hyperparameter $\lambda$ regulating the sparsity level.
+Estimator $C_{\sf sparse+latent}$ has two hyperparameters: the number of latent units $d$ and sparsity parameter $\lambda$.
 $C_{\sf sparse+latent}$ estimates a larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.
@@
 
-\paragraph{Cross-validation with condition-specific variances}
+\paragraph{Cross-validation with condition-specific variances.}
 This study required an unbiased estimate of the discrepancy between various correlation matrix estimates and the true value of the correlation matrix.
@@
 
-\paragraph{Simulation}
+\paragraph{Simulation.}
 For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units.

### 2014-07-14 plos.tex
minor edits

@@
     {\bf C.} Representative calcium signals of seven cells, downsampled to 20 Hz, out of the 292 total recorded cells.
-Spiking activity inferred by non-negative deconvolution is shown by red ticks below the trace.
+Spiking activity inferred by nonnegative deconvolution is shown by red ticks below the trace.
     {\bf D.} The spatial arrangement and orientation tuning of the 292 cells from the imaged site.
@@
 \paragraph{Surgery and two-photon imaging.}
-The surgical procedures and data acquisition were performed as described in \cite{Cotton:2013}.
-Briefly, C57BL/6J mice (aged p40--60) were used.
+The surgical procedures and data acquisition were performed as described in \cite{Cotton:2013}: C57BL/6J mice (aged p40--60) were used.
 Anesthesia was initiated with isoflurane (3\%) and maintained with the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
@@
 The visual stimulus consisted of full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s.
-Two sets of stimuli were presented for each imaging site: the first to map directional tuning and the second to estimate noise correlations.
-Directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen directions of motion, 500 ms per direction, without blanks, with 12--30 trials for each direction of motion.
-The data for covariance estimation were collected during presentations of full-field drifting gratings with the same parameters as those used in directional tuning except only two directions (in 9 datasets) or five directions (in 22 datasets) were used and the presentations lasted 1 second and were separated by 1-second blanks.
-Each stimulus condition was presented between 100 and 300 times.
+Two types of stimuli were presented for each imaging site: First, directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen directions of motion, 500 ms per direction, without blanks, with 12--30 trials for each direction of motion.
+Second, ro measure correlations, the stimulus was modified to include only two directions of motion (in 9 datasets) or five directions (in 22 datasets) and the gratings were presented for 1 second and were separated by 1-second blanks, with 100--300 trials for each direction of motion.
 
@@
 
-The measured fluorescent traces were deconvolved to reconstruct the firing rates for each neuron: First, the first principal component was subtracted from the traces, which reduced common mode noise related to small cardiovascular movements \cite{Cotton:2013}.
+The measured fluorescent traces were deconvolved to reconstruct the firing rates for each neuron: First, the first principal component was subtracted from the raw traces in order to reduce common mode noise related to small cardiovascular movements \cite{Cotton:2013}.
 The resulting traces were high-pass filtered above 0.1 Hz and downsampled to 20 Hz (\figref{2}{C}).
-Then, the firing rates were estimated using a fast non-negative deconvolution algorithm \cite{Vogelstein:2010}.
+Then, the firing rates were estimated using by nonnegative deconvolution \cite{Vogelstein:2010}.
 
-Orientation tuning was computed by fitting the mean firing rates in response to gratings drifting in directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
-The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times; the $p$-values of the fits were computed as the fraction of the permuted datasets for which the $R^2$ value of the tuning function fit equaled or exceeded that of the real data.
-Cells were considered tuned with $p$-values below 0.05.
+Orientation tuning was computed by fitting the mean firing rates for each direction of motion $\phi$ using two-peaked von Mises tuning functions $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
+The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times; the $p$-values of the fits were computed as the fraction of permutations that yielded $R^2$ equal to or higher than that of the unpermuted data.
+Cells were considered tuned with $p<0.05$.
 
@@
 Each subset was then used as the testing sample with the rest of the data used as the training sample for estimating the covariance matrix.
-The resulting validation losses were averaged to produce the \emph{validation loss} reported throughout the paper.
+The average validation loss over the 10 folds was reported.
 
@@
 When the validation loss was not required, only the inner loop of cross-validation was used on the entire dataset.
-This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:4}--\ref{fig:6}) once its superior performance has been established.
+This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:4}--\ref{fig:6}).
 
@@
 Estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag})  used two hyperparameters: the covariance shrinkage intensity $\lambda \in [0,1]$ and variance shrinkage intensity $\alpha \in [0,1]$.
-The variances (the diagonal of $C_{\sf sample}$) were shrunk toward (linearly mixed with) their mean value:
+The variances (the diagonal of $C_{\sf sample}$) were shrunk linearly toward their mean value $\frac 1 p \Tr(C_{\sf sample})$:
 \begin{equation}
@@
 \end{equation}
-Then the diagonal matrix $D$ was used as the target of covariance shrinkage (Eq.~\ref{eq:c-diag}) to produce the final regularized estimate.
+The $C_{\sf diag}$ estimate was then obtained by shrinking $C_{\sf sample}$ toward $D$ according to Eq.~\ref{eq:c-diag}.
 
@@
 \end{equation}
-by an expectation-maximization (EM) algorithm.  The final regularized estimate is obtained by linear shrinkage of the diagonal matrix $D$ toward its mean diagonal value.
+by an expectation-maximization (EM) algorithm with specified rank of $L$. Additional the diagonal matrix of individual variances is shrunk toward its mean value according to Eq.~\ref{eq:c-factor}.
 
-Estimator $C_{\sf sparse}$  has one hyperparameter $\lambda$ to regulate the sparsity of its precision matrix $S$. The precision matrix is obtained by minimizing the $L_1$-penalized loss:
+In estimator $C_{\sf sparse}$ (Eq.~\ref{eq:c-sparse}), the sparse precision matrix $S$ is found by minimizing the $L_1$-penalized loss with regularization parameter $\lambda$:
 \begin{equation}
@@
 \end{equation}
-where $\hat S\succ 0$ denotes the constraint that $\hat S$ be a positive definite matrix and $\|\hat S\|_1$ is the element-wise $L_1$ norm of the matrix $\hat S$.
+where $\hat S\succ 0$ denotes the constraint that $\hat S$ be a positive-definite matrix and $\|\hat S\|_1$ is the element-wise $L_1$ norm of the matrix $\hat S$.
 This problem formulation is known as \emph{graphical lasso} \cite{Meinshausen:2006, Friedman:2008}.
 To solve this minimization problem, we adapted the alternative-direction method of multipliers (ADMM) \cite{Ma:2013}.
-Then the final estimate is the inverse of $S$ (Eq.~\ref{eq:c-sparse}).
-Unlike $C_{\sf diag}$ and $C_{\sf factor}$, this estimator does not include linear shrinkage: the selection of the sparsity level provides sufficient flexibility to fine tune the regularization level.
+Unlike $C_{\sf diag}$ and $C_{\sf factor}$, this estimator does not include linear shrinkage: the selection of the sparsity level provides sufficient flexibility to fine-tune the regularization level.
 
-Estimator $C_{\sf sparse+latent}$ has two hyperparameters: the number of latent units $d$ and sparsity parameter $\lambda$.
-$C_{\sf sparse+latent}$ estimates a larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.
+Estimator $C_{\sf sparse+latent}$ (Eq.~\ref{eq:c-sl}) estimates a larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.  
 \begin{equation}
@@
 \end{equation}
-where the $p\times p$ partition $S$ corresponds to the visible units and expresses their partial correlation structure, and $S_{12}$ and $S_{22}$ are of size $p\times d$ and $d\times d$, respectively.
+where the $p\times p$ partition $S$ corresponds to the visible units. 
 Then the covariance matrix of the observed population is 
@@
 \end{equation}
-The $p\times p$  matrix $L=S_{12}S_{22}^{-1}S_{12}^\T$ has rank $d$.
-Rather than searching for the optimal sparse structure of $S_{12}$ and $S_{22}$, an ill-posed problem, we estimated these components together as the low-rank matrix $L$.
-We solved this problem by adapting the ADMM algorithm to minimizing the loss function with penalty on the $L_1$-norm  of $S$ and the nuclear norm of $L$ \cite{Chandrasekaran:2010,Ma:2013}:
+The rank of the $p\times p$  matrix $L=S_{12}S_{22}^{-1}S_{12}^\T$ matches the number of the latent units in the joint distribution.
+Rather than finding $S_{12}$ and $S_{22}$ separately, $L$ can be estimated as a low-rank matrix.
+We adapted the ADMM algorithm to minimize the loss function with $L_1$ penalty on $S$ to regulate its sparseness and nuclear norm penalty on $L$ to regulate its rank \cite{Chandrasekaran:2010,Ma:2013}:
 \begin{equation}\label{eq:ma}
@@
 
-The partial correlation matrix with both sparse and low-rank components is computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}; it includes the effects of interactions between the visible and latent units.
-This estimate of the partial correlations was used in analyses of average partial correlations (\figref{4}{B and E} and \figref{5}{D}, and \figref{6}{A--C}).
-The sparse component $S$ was used separately in analyses of the connectivity graphs (\figref{4}{F, G, H} and \figref{5}{B, C, E, and F} and \figref{6}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component alone without the low-rank component:
+The partial correlation matrix (Eq.~\ref{eq:partial}) computed from $C_{\sf sparse+latent}$ includes interactions between the visible and latent units and was used \figref{4}{C and D} and \figref{5}{C}, and \figref{6}{A--C}).
+The partial correlation matrix computed from $S$ alone expresses strengths of pairwise interactions
 \begin{equation}
@@
 \end{equation}
+and were used in \figref{4}{F, G, H}.
 
@@
 
-\paragraph{Cross-validation with condition-specific variances.}
+\paragraph{Cross-validation with conditioned variances.}
 This study required an unbiased estimate of the discrepancy between various correlation matrix estimates and the true value of the correlation matrix.
@@
 However, this approach only worked under the assumption of uniform variances across all conditions included in the training and testing samples.
-Here, we extend the validation loss to the estimation and evaluation of a common correlation matrix across multiple conditions with different variances:
+Here, we extend the validation loss to the estimation and evaluation of a common correlation matrix across multiple conditions with different variances.
 
-Let $\{x(t):t\in T_c\}$ be a set of measurements collected in condition $c$ from the training sample and let $\{x(t):t \in T_c^\prime\}$ be a set of measurements in condition $c$ from the testing sample, where $T_c$ and $T_c^\prime$ are time bin indices for the training and testing samples, respectively.
-We marked all quantities estimated from the testing sample with a prime mark: \emph{e.g.}\,$C_{c,\sf sample}^\prime$, $n_c^\prime$.
+Let $T_c$ and $T_c^\prime$ denote the time bin indices for the training and testing samples, respectively, limited to condition $c$.
+Here, the prime symbol marks quantities estimated from the testing sample.
 
@@
 
-Since $\bar x_c= \frac 1 {n_c} \sum\limits_{t \in T_c}x(t)$ is estimated from the training sample in both estimates, $C_{c,\sf sample}^\prime$ is an unbiased estimate of the true covariance matrix, $\Sigma$, and can be used for validation whereas $C_{c,\sf sample}$ is biased toward zero and is used as input for regularized covariance estimates.
+Note that $\bar x_c= \frac 1 {n_c} \sum\limits_{t \in T_c}x(t)$ is estimated from the training sample but used in both estimates, making $C_{c,\sf sample}^\prime$ an unbiased estimate of the true covariance matrix, $\Sigma$.
+As such, $C_{c,\sf sample}^\prime$ can be used for validation.
 
-To apply regularization to the correlation matrix, we first reconstruct the common covariance matrix.
 The common correlation matrix $R_{\sf sample}$ is estimated by averaging the condition-specific correlations:

### 2014-07-14 plos.tex
minor

@@
 The \emph{partial correlation} between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}.
-Partial correlations figure prominently in probablistic \emph{graphical modelling} wherein the joint distribution is explained by sets of two-way interactions \cite{Whittaker:1990}.
+Partial correlations figure prominently in probabilistic \emph{graphical modeling} wherein the joint distribution is explained by sets of two-way interactions \cite{Whittaker:1990}.
 For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}.
 More generally, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1990,Baba:2004}.
-As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functinonal connectivity) between pairs of neurons.
+As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functional connectivity) between pairs of neurons.
 
@@
 To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
-The hyperparameters of each estimator were optimized by nested cross-validation (See Methods).
+The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods).
 Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}).
@@
 Orientation tuning was computed by fitting the mean firing rates for each direction of motion $\phi$ using two-peaked von Mises tuning functions $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
-The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times; the $p$-values of the fits were computed as the fraction of permutations that yielded $R^2$ equal to or higher than that of the unpermuted data.
+The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times; the $p$-values of the fits were computed as the fraction of permutations that yielded $R^2$ equal to or higher than that of the original data.
 Cells were considered tuned with $p<0.05$.
@@
 
-To evaluate the estimators, we regularized the conditioned variances by linear shrinkrage toward their mean value across all conditions.
-This was accomplished by scaling $C_{\sf reg}$ by the diagonal variance adjustment matrix
+To evaluate the estimators, we regularized the conditioned variances by linear shrinkage toward their mean value across all conditions.
+This was done by scaling $C_{\sf reg}$ by the conditioned variance adjustment matrix $Q_c = \delta I + (1-\delta)V_{\sf sample}^{-1} V_{c,\sf sample}$ to produce the conditioned regularized covariance estimate:
 \begin{equation}
-	Q_c = \delta I + (1-\delta)V_{\sf sample}^{-1} V_{c,\sf sample}
+	C_{c,\sf reg} = Q_c^{\frac 1 2} C_{\sf reg} Q_c^{\frac 1 2},
 \end{equation}
-where $\delta \in [0,1]$ is the variance regularization hyperparameter, which was optimized in the inner loop of cross-validation along with the regularization hyperparameters for $C_{\sf reg}$.
-
-Then the regularized condition-specific covariance matrices became
-\begin{equation}
-	C_{c,\sf reg} = Q_c^{\frac 1 2} C_{\sf reg} Q_c^{\frac 1 2} 
-\end{equation}
+The variance regularization parameter $\delta \in [0,1]$ was optimized in the inner loop of cross-validation along with the regularization hyperparameters for $C_{\sf reg}$.
 
@@
 
-%\section*{Figure Legends}
-\newpage
+\section*{Figure Legends}
 
@@
 	{\bf Optimization of hyperparameters of the $C_{\sf sparse+latent}$ estimator.}  
-	{\bf A.} The cross-validated loss function (Eq.~\ref{eq:full-loss}) of the data from the example as a function of the values of hyperparameters $\alpha$ and $\beta$ of the minimization problem in Eq.~\ref{eq:ma}.
+	{\bf A.} Validation loss (Eq.~\ref{eq:full-loss}) for the example site in \figref{2} and \figref{4} as a function of the hyperparameters $\alpha$ and $\beta$ of the $C_{\sf sparse+latent}$ estimator (Eq.~\ref{eq:c-sl} and Eq.~\ref{eq:ma}).
 In all panels, the red cross marks the optimal value found by the pattern search algorithm described in Methods.
-	{\bf B.} The connectivity ($1-\mbox{sparsity}$) of the sparse component $S$ of the inverse covariance (Eq.~\ref{eq:c-sl}) as a function of $\alpha$ and $\beta$ for the example site.
-	{\bf C.} The number of latent units or the rank of the low-rank component $L$ of the inverse covariance (Eq.~\ref{eq:c-sl}) as a function of hyperparameters $\alpha$ and $\beta$ for the example site.
-	{\bf D.} The loss function as a function of the the connectivity and the number of latent units for the exapmle site.
+	{\bf B.} The connectivity ($1-\mbox{sparsity}$) of the sparse component $S$ as a function of $\alpha$ and $\beta$ for the example site.
+	{\bf C.} The number of latent units, \ie the rank of the low-rank component $L$ as a function of hyperparameters $\alpha$ and $\beta$.
+	{\bf D.} The loss function as a function of the connectivity and the number of latent units.
 }

### 2014-07-14 manuscript.tex
added manuscript.tex in latest latex template

--- 2014-07-14 plos.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
+% Based on Template for PLoS
+% Version 2.0 July 2014
+%
+%
+% -- FIGURES AND TABLES
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - tabs/spacing/line breaks within cells to alter layout
+% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
+% - colors, shading, or graphic objects
+% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
+%
+% For sideways tables, use the {rotating} package and use \begin{sidewaystable} instead of \begin{table} in the appropriate section. PLOS guidelines do not accomodate sideways figures.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
+%
+% Please be sure to include all portions of an equation in the math environment, and for any superscripts or subscripts also include the base number/text.
+For example, use $mathrm{mm}^2$ instead of mm$^2$ (do not use \textsuperscript command).
+%
+% DO NOT USE the \rm command to render mathmode characters in roman font, instead use $\mathrm{}$
+% For bolding characters in mathmode, please use $\mathbf{}$ 
+%
+% Please add line breaks to long equations when possible in order to fit our 2-column layout. 
+%
+% For inline equations, please do not include punctuation within the math environment unless this is part of the equation.
+%
+% For spaces within the math environment please use the \; or \: commands, even within \text{} (do not use smaller spacing as this does not convert well).
+%
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 \usepackage{amsmath}
-
 % amssymb package, useful for mathematical symbols
 \usepackage{amssymb}
-
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
 
@@
 
-\usepackage{color} 
-\usepackage[CaptionAfterwards]{fltpage}
-\usepackage{floatrow}
+\usepackage{hyperref}
+
+% line numbers
+\usepackage{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% rotating package for sideways tables
+%\usepackage{rotating}
+
+% If you wish to include algorithms, please use one of the packages below.
+Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
+%\usepackage{algorithmic}
+%\usepackage{algorithmicx}
+
 % Use doublespacing - comment out for single spacing
@@
 %\doublespacing
-
-\usepackage{soul}
 
@@
 
-% Use the PLoS provided bibtex style
+% Use the PLoS provided BiBTeX style
 \bibliographystyle{plos2009}
@@
 \pagestyle{myheadings}
-%% ** EDIT HERE **
 
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-
-\newcommand{\Kcomment}[1]{{\color{blue}{[KJ: #1]}}}
-\newcommand{\Acomment}[1]{{\color{red}{[AE: #1]}}}
-
+%% Include all macros below. Please limit the use of macros.
 \DeclareMathOperator{\Tr}{tr}
-\newcommand{\sq}[1]{\lq#1\rq}
-\newcommand{\mcond}{\,\middle\vert\,}
-\newcommand{\cond}{\,\vert\,}
-\newcommand{\figref}[2]{Fig.\;\ref{fig:#1}\,#2}
 \newcommand{\loss}[1]{\mathcal L\left(#1\right)}
-\newcommand{\eloss}[1]{\mathcal L_0\left(#1\right)}
 \newcommand{\T}{{\sf T}}
 \newcommand{\E}[2][]{\mathbb E_{#1}\left[ #2\right]}    % expected value
-\newcommand{\ie}{\emph{i.e.}\;}
-\newcommand{\eg}{\emph{e.g.}\;}
-
 \DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator{\rank}{rank}
-
 %% END MACROS SECTION
@@
 \begin{document}
+
 % Title must be 150 characters or less
@@
 {\Large
-Improved estimation and interpretation of correlations in neural circuits
+\textbf{Improved estimation and interpretation of correlations in neural circuits}
 }
@@
 
+% Please keep the abstract between 250 and 300 words
 \section*{Abstract}
-% Please keep the abstract between 250 and 300 words
 Ambitious projects aim to record the activity of ever larger and denser neuronal populations \emph{in vivo}.
@@
 However, estimating and interpreting large correlation matrices is statistically challenging.
-Estimation can be improved by regularization, \ie by imposing a structure on the estimate.
+Estimation can be improved by regularization, \emph{i.e.}\;by imposing a structure on the estimate.
 The amount of improvement depends on how closely the assumed structure represents dependencies in the data.
@@
 The sparse component of the estimate defined a graph of interactions.
-These interactions reflected the physical distances and orientation tuning properties of cells: The density of positive \sq{excitatory} interactions decreased rapidly with geometric distances and with differences in orientation preference whereas negative \sq{inhibitory} interactions were less selective.
-Because of its superior performance, this \sq{sparse + latent} estimator likely provides a more physiologically relevant representation of the functional connectivity in densely sampled recordings than the sample correlation matrix.
+These interactions reflected the physical distances and orientation tuning properties of cells: The density of positive `excitatory' interactions decreased rapidly with geometric distances and with differences in orientation preference whereas negative `inhibitory' interactions were less selective.
+Because of its superior performance, this `sparse + latent' estimator likely provides a more physiologically relevant representation of the functional connectivity in densely sampled recordings than the sample correlation matrix.
 
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
 \section*{Author Summary}
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.
-
 It is now possible to record the spiking activity of hundreds of neurons at the same time.
-A meaningful statistical description of the collective activity of these neural populations -- their \sq{functional connectivity} -- is a forefront challenge in neuroscience.
+A meaningful statistical description of the collective activity of these neural populations -- their `functional connectivity' -- is a forefront challenge in neuroscience.
 We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.
 Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.
-In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (\sq{interactions}) combined with a low-rank component.
-The inferred interactions were both positive (\sq{excitatory}) and negative (\sq{inhibitory}) and reflected the spatial organization and orientation preferences of the interacting cells.
+In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (`interactions') combined with a low-rank component.
+The inferred interactions were both positive (`excitatory') and negative (`inhibitory') and reflected the spatial organization and orientation preferences of the interacting cells.
+We propose that  the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.
+
+\section*{Introduction}
+It is now possible to record the spiking activity of hundreds of neurons at the same time.
+A meaningful statistical description of the collective activity of these neural populations -- their `functional connectivity' -- is a forefront challenge in neuroscience.
+We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.
+Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.
+In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (`interactions') combined with a low-rank component.
+The inferred interactions were both positive (`excitatory') and negative (`inhibitory') and reflected the spatial organization and orientation preferences of the interacting cells.
 We propose that  the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.
@@
 Pearson correlations between the spiking activity of pairs of neurons are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.
-In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
+In particular, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}.
 In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of synaptic connectivity \cite{Gerstein:1964}.
@@
 The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.
-This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of the largest eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
+This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of the largest eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.
 The \emph{sample correlation matrix} is an unbiased estimate of the true correlations but its many free parameters make it sensitive to sampling noise.
@@
 Estimation can be improved through \emph{regularization},  the technique of deliberately imposing a structure on an estimate in order to reduce its estimation error \cite{Schafer:2005, Bickel:2006}.
-To \sq{impose a structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation  with fewer free parameters, the \emph{target estimate}.
+To `impose a structure' on an estimate means to bias (`shrink') it toward a reduced representation  with fewer free parameters, the \emph{target estimate}.
 The optimal target estimate and the optimal amount of shrinkage can be obtained from the data sample either analytically \cite{Ledoit:2003, Ledoit:2004, Schafer:2005}  or by cross-validation \cite{Friedman:1989}.
@@
 \item[$C_{\sf sample}$] -- sample covariance matrix, the unbiased, unregularized estimator.
-\item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \ie toward a diagonal covariance matrix.
+\item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \emph{i.e.}\;toward a diagonal covariance matrix.
 \item[$C_{\sf factor}$] -- a low-rank approximation of the sample covariance matrix, representing inputs from unobserved shared factors (latent units).
-\item[$C_{\sf sparse}$] -- sparse partial correlations, \ie a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
+\item[$C_{\sf sparse}$] -- sparse partial correlations, \emph{i.e.}\;a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
 \item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons \emph{and} linear interactions with a number of latent units.
@@
 We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.
-This estimator revealed a sparse network of partial correlations (\sq{interactions}), between the observed neurons; it also inferred latent units exerting linear effects on the observed neurons.
+This estimator revealed a sparse network of partial correlations (`{interactions'), between the observed neurons; it also inferred latent units exerting linear effects on the observed neurons.
 We analyzed these networks of partial correlations and found the following: Whereas significant noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.
@@
 
+% Results and Discussion can be combined.
+% We only support three levels of headings, please do not create a heading level below \subsubsection.
 \section*{Results}
-% Results and Discussion can be combined.
-\paragraph{Covariance estimation.}
+\subsection*{Covariance estimation}
 The covariance matrix is defined as
@@
 
-Pairwise partial correlations are closely related to the elements of the \emph{precision matrix}, \ie the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1990}.
+Pairwise partial correlations are closely related to the elements of the \emph{precision matrix}, \emph{i.e.}\;the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1990}.
 Zero elements in the precision matrix signify zero partial correlation between the two variables.
@@
 We considered four regularized estimators based on distinct families of target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$.
-In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{1}{\,Row 1}).
-
-\begin{FPfigure}
-    \begin{center}
-        \includegraphics{./figures/Figure01.pdf}
-    \end{center}
-    \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
-    {\bf Row 1.} Graphical models of the target estimates of the four respective regularized covariance matrix estimators.
-Recorded neurons are represented by the green spheres and latent units by the lightly shaded spheres.
-Edges represent non-zero partial correlations, \ie \sq{interactions}.
-        {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear dependencies. 
-        {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units.
-        {\bf  Row 1, C}.
-For estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix.
-This approximation describes systems in which correlations arise from a sparse set of  linear associations between the observed units.
-        {\bf  Row 1, D}.
-For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix.
-This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} due to several latent units.
-	\\
-	{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with four latent units, {\bf C.}  a correlation matrix with 67\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-3 matrix (\ie three latent units) and a sparse matrix with 76\% off-diagonal zeros.
-\\
-{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=500$ drawn from simulated random processes with respective correlation matrices shown in Row 2.
-The structure of the sample correlation matrix is difficult to discern by eye.
-\\
-{\bf Row 4:} Estimates computed from the same data as in Row 3 using structured estimators of the correct type, optimized by cross-validation.
-The regularized estimates are closer to the truth than the sample correlation matrices.
-\\
-{\bf Row 5:} Excess losses (Eq.~\ref{eq:loss}) for the five estimators as a function of sample size.
-The error bars indicate the standard error of the mean based on 30 samples.
-Estimators with structure that matches the true model converged to zero faster than the other estimators.
-\\
-{\bf Row 6:} Validation losses for the five estimators relative to the matching estimator.
-Error bars indicate the standard error of the mean based on 30 samples.
-Differences in validation loss approximate differences in true loss.
-    }
-    \label{fig:1}
-\end{FPfigure}
+In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (Fig.~\ref{fig:1} Row 1).
 
@@
 \end{equation}
-The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (\figref{1}{Row 1, A}).
+The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (Fig.~\ref{fig:1} Row 1, A).
 If sample correlations are largely spurious, $C_{\sf diag}$ is expected to be more efficient than other estimators.
@@
 \end{equation}
-has two hyperparameters: the rank of $L$ (\ie the number of latent factors) and the shrinkage intensity $\lambda$ to shrink the independent variances toward their mean value $\bar D$.
-The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (\figref{1}{Row 1, B}).
+has two hyperparameters: the rank of $L$ (\emph{i.e.}\;the number of latent factors) and the shrinkage intensity $\lambda$ to shrink the independent variances toward their mean value $\bar D$.
+The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (Fig.~\ref{fig:1} Row 1, B).
 
@@
 \end{equation}
-Here $S$ is a sparse matrix, \ie one in which a large fraction of off-diagonal elements are set to zero.
+Here $S$ is a sparse matrix, \emph{i.e.}\;one in which a large fraction of off-diagonal elements are set to zero.
 The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$.
 The problem of finding the optimal set of non-zero elements of the precision matrix is known as \emph{covariance selection} \cite{Dempster:1972}.
-The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{1}{Row 1, C}).
+The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (`interactions') between some pairs of neurons (Fig.~\ref{fig:1}} Row 1, C).
 
@@
 See Methods for a more detailed explanation.
-The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (\figref{1}{Row 1, D}) \cite{Chandrasekaran:2010,Ma:2013}.
+The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (Fig.~\ref{fig:1} Row 1, D) \cite{Chandrasekaran:2010,Ma:2013}.
 
-We refer to the sparse partial correlations in estimators $C_{\sf sparse}$ and $C_{\sf sparse+latent}$ as \sq{interactions}.
+We refer to the sparse partial correlations in estimators $C_{\sf sparse}$ and $C_{\sf sparse+latent}$ as `interactions'.
 
-\paragraph{Simulation.}
+\subsection*{Simulation}
 We next demonstrate how the most efficient among different regularized estimators can reveal the structure of correlations.
-We constructed four families of $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (\figref{1}{\,Row 2, A--D} and Methods).
+We constructed four families of $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (Fig.~\ref{fig:1} Row 2, A--D and Methods).
 We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes.
-Even with relatively large sample sizes (\eg $n=500$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (\figref{1}{\,Row 3}).
+Even with relatively large sample sizes (\emph{e.g.}\;$n=500$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (Fig.~\ref{fig:1} Row 3).
 
@@
 All the regularized estimators produced better estimates (lower excess loss) than the sample covariance matrix.
-However, estimators whose structure matched the true model outperformed the other estimators (\figref{1}{\,Row 5}).
+However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Row 5).
 Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed equally well to $C_{\sf diag}$ because it correctly identified zero factors and only estimated the individual variances.
@@
 
-The normal loss (Eq.~\ref{eq:loss}) is particularly suitable because it is additive in its second argument, \ie
+The normal loss (Eq.~\ref{eq:loss}) is particularly suitable because it is additive in its second argument, \emph{i.e.} 
 \begin{equation*}\label{eq:additivity}
@@
 Then each subset is used as the testing sample with the other $K-1$ subsets serving as the training sample.
-The validation losses from all such \sq{folds} are averaged to produce the \emph{cross-validation loss}.
+The validation losses from all such `folds' are averaged to produce the \emph{cross-validation loss}.
 
-Cross-validation loss accurately reproduced the relationships between the excess losses of the estimators (\figref{1}{Row 6}), confirming that cross-validation can evaluate true loss.
+Cross-validation loss accurately reproduced the relationships between the excess losses of the estimators (Fig.~\ref{fig:1} Row 6), confirming that cross-validation can evaluate true loss.
 
@@
 
-\paragraph{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data.}
-We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{2}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
-This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{2}{C and D}).
-The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{2}{C}).
+\subsection*{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
+We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:2} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
+This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:2} C and D).
+The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:2} C).
 Only cells that produced detectable calcium activity were included in the analysis (see Methods).
@@
 Each grating was displayed for 500 ms, without intervening blanks.
-This stimulus was used to compute the orientation tuning of the recorded cells (\figref{2}{D}).
+This stimulus was used to compute the orientation tuning of the recorded cells (Fig.~\ref{fig:2} D).
 To estimate the noise correlation matrix, we presented only 2 distinct directions in some experiments or 5 directions in others with 100--300 repetitions of each direction.
@@
 The traces were then binned into 150 ms intervals aligned on the stimulus onset for the estimation of the correlation matrix.
-The sample correlation coefficients were largely positive and low (\figref{2}{E and F}).
-The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean across sites of 0.018 (\figref{5}{D}).
-
-\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-    {\caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
-    Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) presented during two-photon recordings of somatic calcium signals using fast 3D random-access microscopy ({\bf B}).
-    {\bf C--F.} Calcium activity data from an example site.
-    {\bf C.} Representative calcium signals of seven cells, downsampled to 20 Hz, out of the 292 total recorded cells.
-Spiking activity inferred by nonnegative deconvolution is shown by red ticks below the trace.
-    {\bf D.} The spatial arrangement and orientation tuning of the 292 cells from the imaged site.
-The cells' colors indicate their orientation preferences.
-The gray cells were not significantly tuned.
-    {\bf E.} The sample noise correlation matrix of the activity of the neural population. 
-    {\bf F.} Histogram of noise correlation coefficients in one site. The red line indicates the mean correlation coefficient of 0.020.
-} \label{fig:2}}
-    {\includegraphics[width=8.3cm]{figures/Figure02.pdf}}
-\end{figure}
+The sample correlation coefficients were largely positive and low (Fig.~\ref{fig:2} E and F).
+The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean across sites of 0.018 (Fig.~\ref{fig:5} D).
 
@@
 To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
-The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods).
+The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~\ref{supp:1} and  Methods).
 Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}).
 The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).
-
-\begin{figure}[!ht]
-    \begin{center}
-    \includegraphics{./figures/src/Fig3.eps}
-    \end{center}
-    \caption{
-   {\bf Performance of estimator $C_{\sf sparse+latent}$ with respect to the normal loss function (eq.~\ref{eq:loss}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.}
-    Covariance estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ produced consistently greater validation losses than $C_{\sf sparse+latent}$ ($p<0.01$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
-The box plots indicate the $25^{th}$, $50^{th}$, and $75^{th}$ percentiles with the whiskers extending to the minimum and maximum values after excluding the outliers marked with `+'.
-} \label{fig:3}
-\end{figure}
 
@@
 
-\paragraph{Structure of $C_{\sf sparse+latent}$ estimates.}
+\subsection*{Structure of $C_{\sf sparse+latent}$ estimates}
 We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:4} and Fig.~\ref{fig:5}).
-Although the regularized estimates were similar to the sample correlation matrix (\figref{4}{A and B}), the corresponding partial correlation matrices differed substantially (\figref{4}{C and D}).
-The estimates separated two sources of correlations: a network of linear interactions expressed by the sparse component of the inverse and latent units expressed by the low-rank components of the inverse (\figref{4}E).
-The sparse partial correlations revealed a network that differed substantially from the network composed of the greatest coefficients in the sample correlation matrix (\figref{4}{F, G, H, and I}).
+Although the regularized estimates were similar to the sample correlation matrix (Fig.~\ref{fig:4} A and B), the corresponding partial correlation matrices differed substantially (Fig.~\ref{fig:4} C and D).
+The estimates separated two sources of correlations: a network of linear interactions expressed by the sparse component of the inverse and latent units expressed by the low-rank components of the inverse (Fig.~\ref{fig:4} E).
+The sparse partial correlations revealed a network that differed substantially from the network composed of the greatest coefficients in the sample correlation matrix (Fig.~\ref{fig:4} F, G, H, and I).
 
-In the example site (Fig.~\ref{fig:4}), the sparse component had 92.8\% sparsity (or conversely, 7.2\% connectivity: $\mbox{connectivity}=1-\mbox{sparsity}$) with average node degree of 20.9 (\figref{4}{G}).
-The average node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $\mbox{degree} = \mbox{connectivity}\cdot(p-1)$, where $p$ is the number of neurons.
+In the example site (Fig.~\ref{fig:4}), the sparse component had 92.8\% sparsity (or conversely, 7.2\% connectivity: $\mbox{connectivity}=1-\mbox{sparsity}$) with average node degree of 20.9 (Fig.~\ref{fig:4} G).
+The average node degree, \emph{i.e.}\;the average number of interactions linking each neuron, is related to connectivity as $\mbox{degree} = \mbox{connectivity}\cdot(p-1)$, where $p$ is the number of neurons.
 The low-rank component had rank 72, denoting 72 inferred latent units.
-The number of latent units increased with population size (\figref{5}{A}) but the connectivity was highly variable (\figref{5}{B}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
+The number of latent units increased with population size (Fig.~\ref{fig:5} A) but the connectivity was highly variable (Fig.~\ref{fig:5} B): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
 This variability may be explained by differences in brain states and recording quality and warrants further investigation.
 
-The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 27 sites were about 5 times lower than the average sample correlations (\figref{5}{C}).
+The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 27 sites were about 5 times lower than the average sample correlations (Fig.~\ref{fig:5} C).
 This suggests that correlations between neurons build up from multiple chains of smaller interactions.
@@
 
-While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (28.7\% in the example site) of negative values (\figref{4}{F}).
-The fraction of negative interactions increased with the inferred connectivity \figref{5}{D}, suggesting that negative interactions can be inferred only after a sufficient density of positive interactions has been uncovered.
+While the sample correlations were mostly positive, the sparse component of the partial correlations (`interactions') had a high fraction (28.7\% in the example site) of negative values (Fig.~\ref{fig:4} F).
+The fraction of negative interactions increased with the inferred connectivity (Fig.~\ref{fig:5} D), suggesting that negative interactions can be inferred only after a sufficient density of positive interactions has been uncovered.
 
-\begin{FPfigure}
-    \begin{center}
-        \includegraphics[width=17.35cm]{./figures/Figure04.pdf}
-    \end{center}
+Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013, Sadovsky:2014}.
+We therefore compared the interactions in the sparse component of $C_{\sf sparse+latent}$ to those obtained from the sample correlations thresholded to the same level of connectivity.
+The networks revealed by the two methods differed substantially.
+In the example site with 7.2\% connectivity in $C_{\sf sparse+latent}$, only 27.7\% of the connections coincided with the above-threshold sample correlations (Fig.~\ref{fig:4} F, H, and I).
+In particular, most of the inferred negative interactions corresponded to low sample correlations (Fig.~\ref{fig:4} F) where high correlations should be expected given the rest of the correlation matrix.
+
+\subsubsection*{Relationship of $C_{\sf sparse+latent}$ to orientation tuning and physical distances}
+
+We examined how the structure of the $C_{\sf sparse+latent}$ estimates related to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:6}).
+Five sites with highest pairwise connectivities were included in the analysis.
+Partial correlations were computed using Eq.~\ref{eq:partial} based on the regularized estimate, including both the sparse and the latent component.
+Connectivity was computed as the fraction of pairs of cells connected by non-zero elements (interactions) in the sparse component of the estimate, distinguishing between the positive and negative connectivities.
+
+First, we analyzed how correlations and connectivity depended on the difference in preferred orientations ($\Delta ori$) of pairs of significantly ($\alpha=0.05$) tuned cells.
+The partial correlations decayed more rapidly with $\Delta ori$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test of the difference of the linear regression coefficients).
+Positive connectivity decreased with $\Delta ori$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression coefficient) whereas negative connectivity did not decrease (Fig.~\ref{fig:6} D): The slope in the logistic model of connectivity with respect to $\Delta ori$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficient).
+
+Second, we compared how correlations and connectivity depended on the physical distance separating pairs of cells.
+We distinguished between lateral distance, $\Delta x$, in the plane parallel to the pia, and vertical distance, $\Delta z$, orthogonal to the pia.
+When considering the dependence on $\Delta x$, the analysis was limited to cell pairs located at the same depth with $\Delta z < 30\,\mu\mbox{m}$; conversely, when considering the dependence on $\Delta z$, only vertically aligned cell pairs with $\Delta x < 30\,\mu\mbox{m}$ were included.
+Again, the partial correlations decayed more rapidly both laterally and vertically than sample correlations ($p<10^{-6}$ in each of the five sites, for both lateral and vertical distances, two-sample $t$-test of the difference of the linear regression coefficients).
+Positive connectivity decayed with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression coefficient) (Fig.~\ref{fig:6} E), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$.
+Although the positive connectivity appeared to decay faster with vertical than with lateral distance, the differences in slopes of the respective logistic regression models were not significant with available data.
+The negative connectivity decayed slower with distance (Fig.~\ref{fig:6} E and F): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficients).
+
+\section*{Discussion}
+
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis" 
+\section*{Materials and Methods}
+
+% Do NOT remove this, even if you are not including acknowledgments.
+
+\section*{Acknowledgments}
+
+%\section*{References}
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% OR
+%
+% Compile your BiBTeX database using our plos2009.bst
+% style file and paste the contents of your .bbl file
+% here.
+% 
+\bibliography{references.bib}
+
+\section*{Figure Legends}
+% This section is for figure legends only, do not include
+% graphics in your manuscript file.
+
+\begin{figure}
+\caption{
+{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
+     {\bf Row 1.} Graphical models of the target estimates of the four respective regularized covariance matrix estimators.
+Recorded neurons are represented by the green spheres and latent units by the lightly shaded spheres.
+Edges represent non-zero partial correlations, \emph{i.e.}\;`interactions'.
+     {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear dependencies.
+     {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units.
+     {\bf  Row 1, C}.
+For estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix.
+This approximation describes systems in which correlations arise from a sparse set of  linear associations between the observed units.
+     {\bf  Row 1, D}.
+For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix.
+This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} due to several latent units.
+     \\
+     {\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with four latent units, {\bf C.}  a correlation matrix with 67\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-3 matrix (\emph{i.e.}\;three latent units) and a sparse matrix with 76\% off-diagonal zeros.
+     \\
+{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=500$ drawn from simulated random processes with respective correlation matrices shown in Row 2.
+The structure of the sample correlation matrix is difficult to discern by eye.
+     \\
+{\bf Row 4:} Estimates computed from the same data as in Row 3 using structured estimators of the correct type, optimized by cross-validation.
+The regularized estimates are closer to the truth than the sample correlation matrices.
+     \\
+{\bf Row 5:} Excess losses (Eq.~\ref{eq:loss}) for the five estimators as a function of sample size.
+The error bars indicate the standard error of the mean based on 30 samples.
+Estimators with structure that matches the true model converged to zero faster than the other estimators.
+     \\
+{\bf Row 6:} Validation losses for the five estimators relative to the matching estimator.
+Error bars indicate the standard error of the mean based on 30 samples.
+Differences in validation loss approximate differences in true loss.
+}
+\label{fig:1}
+\end{figure}
+
+\begin{figure}
+\caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
+Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) presented during two-photon recordings of somatic calcium signals using fast 3D random-access microscopy ({\bf B}).
+{\bf C--F.} Calcium activity data from an example site.
+{\bf C.} Representative calcium signals of seven cells, downsampled to 20 Hz, out of the 292 total recorded cells.
+Spiking activity inferred by nonnegative deconvolution is shown by red ticks below the trace.
+{\bf D.} The spatial arrangement and orientation tuning of the 292 cells from the imaged site. The cells' colors indicate their orientation preferences. The gray cells were not significantly tuned.
+{\bf E.} The sample noise correlation matrix of the activity of the neural population.
+{\bf F.} Histogram of noise correlation coefficients in one site. The red line indicates the mean correlation coefficient of 0.020.
+} \label{fig:2}
+\end{figure}
+
+\begin{figure}
+\caption{
+{\bf Performance of estimator $C_{\sf sparse+latent}$ with respect to the normal loss function (eq.~\ref{eq:loss}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.}
+Covariance estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ produced consistently greater validation losses than $C_{\sf sparse+latent}$ ($p<0.01$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
+The box plots indicate the $25^{th}$, $50^{th}$, and $75^{th}$ percentiles with the whiskers extending to the minimum and maximum values after excluding the outliers marked with `+'.
+} \label{fig:3}
+\end{figure}
+
+\begin{figure}
 \caption{{\bf Structure revealed by $C_{\sf sparse+latent}$.}
@@
 \label{fig:4}
-\end{FPfigure}
+\end{figure}
 
-Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013, Sadovsky:2014}.
-We therefore compared the interactions in the sparse component of $C_{\sf sparse+latent}$ to those obtained from the sample correlations thresholded to the same level of connectivity.
-The networks revealed by the two methods differed substantially.
-In the example site with 7.2\% connectivity in $C_{\sf sparse+latent}$, only 27.7\% of the connections coincided with the above-threshold sample correlations (\figref{4}{F, H, and I}).
-In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{4}{F}) where high correlations should be expected given the rest of the correlation matrix.
-
-\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
-    {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites.} 
+\begin{figure}
+\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites.}
 Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:2} and \ref{fig:4} is circled in blue.
@@
 }
-\label{fig:5}}
-{\includegraphics{./figures/Figure05.pdf}}
+\label{fig:5}
 \end{figure}
 
-\paragraph{Relationship of $C_{\sf sparse+latent}$ to orientation tuning and physical distances.}
-
-We examined how the structure of the $C_{\sf sparse+latent}$ estimates related to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:6}).
-Five sites with highest pairwise connectivities were included in the analysis.
-Partial correlations were computed using Eq.~\ref{eq:partial} based on the regularized estimate, including both the sparse and the latent component.
-Connectivity was computed as the fraction of pairs of cells connected by non-zero elements (interactions) in the sparse component of the estimate, distinguishing between the positive and negative connectivities.
-
-First, we analyzed how correlations and connectivity depended on the difference in preferred orientations ($\Delta ori$) of pairs of significantly ($\alpha=0.05$) tuned cells.
-The partial correlations decayed more rapidly with $\Delta ori$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test of the difference of the linear regression coefficients).
-Positive connectivity decreased with $\Delta ori$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression coefficient) whereas negative connectivity did not decrease (\figref{6}{D}): The slope in the logistic model of connectivity with respect to $\Delta ori$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficient).
-
-Second, we compared how correlations and connectivity depended on the physical distance separating pairs of cells.
-We distinguished between lateral distance, $\Delta x$, in the plane parallel to the pia, and vertical distance, $\Delta z$, orthogonal to the pia.
-When considering the dependence on $\Delta x$, the analysis was limited to cell pairs located at the same depth with $\Delta z < 30\,\mu\mbox{m}$; conversely, when considering the dependence on $\Delta z$, only vertically aligned cell pairs with $\Delta x < 30\,\mu\mbox{m}$ were included.
-Again, the partial correlations decayed more rapidly both laterally and vertically than sample correlations ($p<10^{-6}$ in each of the five sites, for both lateral and vertical distances, two-sample $t$-test of the difference of the linear regression coefficients).
-Positive connectivity decayed with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression coefficient) (\figref{6}{E}), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$.
-Although the positive connectivity appeared to decay faster with vertical than with lateral distance, the differences in slopes of the respective logistic regression models were not significant with available data.
-The negative connectivity decayed slower with distance (\figref{6}{E and F}): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficients).
-
-\begin{figure}[!ht]
-    \begin{center}
-        \includegraphics{./figures/Figure06.pdf}
-    \end{center}
+\begin{figure}
 \caption{{\bf Dependence of sample correlations, regularized partial correlations, and connectivity inferred by $C_{\sf sparse+latent}$ on the differences in preferred orientations, $\Delta ori$, and physical distances, $\Delta x$ and $\Delta z$.} The error bars mark the standard errors of the means.
-Five sites with highest connectivity (see \figref{5}{B}) were selected for this analysis.
+Five sites with highest connectivity (see Fig.~\ref{fig:5} B) were selected for this analysis.
 {\bf A--C.} Normalized mean sample correlations (black) and normalized mean partial correlations (red) estimated by $C_{\sf sparse+latent}$ across $n=5$ imaged sites.
-The values in each bin are normalized by the means across the entire site, shown in \figref{5}{C}, to make the effects more comparable across the sites.
+The values in each bin are normalized by the means across the entire site, shown in Fig.~\ref{fig:5} C, to make the effects more comparable across the sites.
 {\bf A.} Mean partial correlations in $C_{\sf sparse+latent}$ depend on $\Delta ori$ more strongly than mean sample correlations.
@@
 Normalized negative connectivity is computed similarly for the negative connectivity.
-The average connectivity across sites is shown in \figref{5}{B} with only the five most connected sites included in the analysis.
+The average connectivity across sites is shown in Fig.~\ref{fig:5} B with only the five most connected sites included in the analysis.
 The normalization made the effects of tuning and distance more comparable across sites.
@@
 
-\section*{Discussion}
-\paragraph{Functional connectivity as a network of pairwise interactions.}
-Functional connectivity is often represented as a graph of pairwise interactions.
-The goal of many studies of functional connectivity has been to estimate  anatomical connectivity from  observed multineuronal spiking activity.
-For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
-Such signatures are ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.
-With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.
-Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network and external input.
-Therefore, conditional independence can signify the absence of a direct causal influence.
+%\section*{Tables}
+% 
+% See introductory notes if you wish to include sideways tables.
+%
+% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements.
+Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
+%
+%\begin{table}[!ht]
+%\caption{
+%\bf{Table title}}
+%\begin{tabular}{|c|c|c|}
+%table information
+%\end{tabular}
+%\begin{flushleft}Table caption
+%\end{flushleft}
+%\label{tab:label}
+% \end{table}
 
-Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity.
-For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration, membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.
-Maximum entropy models constrained by observed pairwise correlations are among other models with pairwise coupling between cells \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.
-Assuming that the population response follows a multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
-Each probabilistic model, fitted to the same data may reveal a completely different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.
-
-It is not yet clear which approach provides the best correspondence with anatomical connectivity.
-Little experimental evidence is available to answer this question.
-The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to anatomy.
-Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013, Sadovsky:2014}.
-However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}.
-Until a physiological interpretation of functional connectivity is established, the physiological relevance of such analyses remains in question and we did not attempt them here.
-
-Inference of the conditional dependencies also depends on the completeness of the recorded population:  To equate conditional dependencies to direct interaction between neurons, we must record from all neurons with which the pair interacts.
-Unobserved portions of the circuit may manifest as conditional dependencies between observed neurons that do not interact.
-For this reason, statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}.
-In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies.
-Perhaps for this reason, partial correlations have not, until now, been used to describe the functional connectivity in cortical populations.
-
-Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.
-While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}.
-However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.
-Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.
-The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.
-Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
-These techniques may provide more incisive measurements of functional connectivity than electrophysiological recordings.
-
-The low temporal resolution of calcium signals limits the use of functional connectivity methods that rely on millisecond-scale binning of signals (cross-correlograms, some GLMs, and binary maximum entropy models).
-Hence, most studies of functional connectivity have relied on instantaneous sample correlations \cite{Greenberg:2008, Golshani:2009, Hofer:2011, Malmersjo:2013} .
-Although some investigators have interpreted such correlations as indicators of (chemical or electrical) synaptic connectivity, most used them as more general indicators of functional connectivity without relating them to underlying mechanisms.
-
-In this study, we sought to infer pairwise functional connectivity networks  in cortical microcircuits.
-We hypothesized that partial correlations correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.
-Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we \emph{a priori} favored solutions with sparse partial correlations.
-Under the assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
-
-\paragraph{Functional connectivity as coactivations.}
-Another approach to describing the functional connectivity of a circuit is to isolate individual patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}.
-Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}.
-Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \ie groups of tightly interconnected groups of cells involved in a common computation.
-Coactivation patterns could also result from shared input from unobserved parts of the circuit, or global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.
-
-Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity.
-However, an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.
-In our study, the functional connectivity solely through modes of coactivations was represented by the factor analysis-based estimator $C_{\sf factor}$.
-
-\paragraph{Combining pairwise interactions and coactivations.}
-In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables \cite{Koster:2013},  high-order correlations \cite{Ganmor:2011}, or global network fluctuations \cite{Tkacik:2013}.
-
-In our study, we combined pairwise interactions with collective coactivations by applying the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables \cite{Chandrasekaran:2010, Ma:2013}.
-The resulting estimator, $C_{\sf sparse+latent}$, effectively decomposed the functional connectivity into a sparse network of pairwise interactions and coactivation mode vectors.
-
-\paragraph{Addressing ill-posedness.}
-Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.
-The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
-As techniques have improved to allow recording from larger neuronal populations, experimental neuroscientists have addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted \cite{Tkacik:2013}.
-However, ambitious projects, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
-Simply increasing recording duration will not be sufficient, and the problem must be addressed by using regularized estimators.
-Regularization biases the solution toward a small subspace in order to counteract the effect of  sampling noise in the empirical data.
-However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hinders interpretation.
-
-Several strategies have been developed to limit the model space in order to improve the quality of the estimate.
-For example, Ganmor et al.
-\cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in the retina.
-As another example of regularization, generalized linear models typically employ $L_1$ penalty terms to constrain the solution space and to effectively reduce the dimensionality of the solution \cite{Pillow:2008}.
-
-In our study, regularizations were accomplished by dimensionality reduction (feature selection) schemes to produce sparse, constrained solutions.
-Only the most efficient scheme was considered in the analysis of functional connectivity.
-
-\paragraph{Model selection.}
-Various model selection criteria have been devised to select between families of models and the optimal subsets of variables in a given model family based on observed data.
-Despite its computational requirements, cross-validation is among the most popular model selection methods due to its minimal assumptions about the data generating process \cite{Arlot:2010}.
-
-We evaluated the covariance matrix estimators a loss function based on multivariate normal log likelihood.
-However, this does not limit the applicability of its conclusions to normal distributions.
-Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:3} and Fig.~S\ref{supp:02}).
-Other probabilistic models, fitted to the same data, could also serve as estimators of the covariance matrix.
-If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.
-
-The results of model selection must be interpreted with caution.
-As we demonstrated by simulation, even models with incorrect forms of dependencies can substantially improve estimates (\figref{1}).
-Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not necessarily support the conclusion that it reveals a better representation of dependencies in the data.
-This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary low-dimensional target can consistently outperform a less constrained estimate.
-
-\paragraph{Physiological interpretation and future directions.}
-
-We showed that among several models a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix for cortical microcircuits.
-This finding is valuable in itself: improved estimates of the noise covariance matrix for large datasets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Ecker:2011}
-
-Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.
-Importantly, the inferred functional interactions were substantially different from the network of the highest sample correlations \figref{4}{F, H, and I}.
-For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{4}{F}) and may reflect inhibitory circuitry.
-
-Distances between cells in physical space and in sensory feature space had a stronger effect on the partial correlations estimated by the $C_{\sf sparse+latent}$ estimator than on sample correlations (\figref{6}{A--C}).
-These differences support the idea that correlations are built up from partial correlations in chains of intermediate cells positioned closer and tuned more similarly to one another, with potentially closer correspondence to anatomical connectivity.
-These differences may also be at least partially explained by a trivial effect of regularization: the $L_1$ penalty applied by the estimator (Eq.~\ref{eq:ma}) suppresses small partial correlations to greater extent than large partial correlations, enhancing the apparent effect of distance and tuning.
-Distances between cells in physical space and in sensory feature space also had distinct effects on the positive and negative connectivities inferred by the estimator (\figref{6}{D--F}); we could not think of a numerical artifact that could produce this effect.
-The distinct positive and negative connectivity patterns  may reflect geometric and graphical features of local excitatory and inhibitory networks.
-Indeed, the relationships between patterns of positive and negative connectivities inferred by the estimator mirrored the properties of excitatory and inhibitory synaptic connectivities with respect to distance, cortical layers, and feature tuning \cite{Song:2005, Oswald:2008, Adesnik:2010, Perin:2011, Fino:2011, Hofer:2011, Isaacson:2011, Levy:2012}.
-For example, while excitatory neurons form synapses within highly specific local cliques \cite{Perin:2011}, inhibitory interneurons form synapses with nearly all excitatory cells within local microcircuits \cite{Fino:2011, Hofer:2011, Packer:2011}.
-To further investigate the link between synaptic connectivity and inferred functional connectivity, in future experiments, we will use molecular markers for various cell types with follow-up multiple whole-cell \emph{in vitro} recordings \cite{Hofer:2011, Ko:2013} to directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
-Finally, the latent units inferred by the estimator can be analyzed for their physiological functions.
-For example, these latent units may be modulated under different brain states (e.g.
-slow-wave sleep, attention) and stimulus conditions (e.g.
-certain types of stimuli may engage feedback connections).
-
-\section*{Methods}
-% You may title this section "Methods" or "Models". 
-% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
-% authors may use "Analysis" 
-\paragraph{Ethics statement.}
-All procedures were conducted in accordance with the ethical guidelines of the National Institutes of Health and were approved by the Baylor College of Medicine IACUC.
-
-\paragraph{Surgery and two-photon imaging.}
-The surgical procedures and data acquisition were performed as described in \cite{Cotton:2013}: C57BL/6J mice (aged p40--60) were used.
-Anesthesia was initiated with isoflurane (3\%) and maintained with the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
-A craniotomy was performed over the right primary visual cortex.
-Membrane-permeant calcium indicator Oregon Green 488 BAPTA-1 AM (OGB-1, Invitrogen) was loaded by bolus injection.
-The craniotomy was sealed using a glass coverslip secured with dental cement.
-
-Calcium imaging began 1 hour after dye injection.
-All imaging was performed using 3D-RAMP two-photon microscopy \cite{Cotton:2013}.
-First, a 3D stack was acquired and cells were manually segmented.
-Then calcium signal were collected by sampling in the center of each cell at rates of 100 Hz or higher, depending on the number of cells.
-
-\paragraph{Visual stimulus.}
-The visual stimulus consisted of full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s.
-Two types of stimuli were presented for each imaging site: First, directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen directions of motion, 500 ms per direction, without blanks, with 12--30 trials for each direction of motion.
-Second, ro measure correlations, the stimulus was modified to include only two directions of motion (in 9 datasets) or five directions (in 22 datasets) and the gratings were presented for 1 second and were separated by 1-second blanks, with 100--300 trials for each direction of motion.
-
-\paragraph{Data processing.}
-All data were processed in MATLAB using the DataJoint data processing chain toolbox (http://datajoint.github.com).
-
-The measured fluorescent traces were deconvolved to reconstruct the firing rates for each neuron: First, the first principal component was subtracted from the raw traces in order to reduce common mode noise related to small cardiovascular movements \cite{Cotton:2013}.
-The resulting traces were high-pass filtered above 0.1 Hz and downsampled to 20 Hz (\figref{2}{C}).
-Then, the firing rates were estimated using by nonnegative deconvolution \cite{Vogelstein:2010}.
-
-Orientation tuning was computed by fitting the mean firing rates for each direction of motion $\phi$ using two-peaked von Mises tuning functions $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
-The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times; the $p$-values of the fits were computed as the fraction of permutations that yielded $R^2$ equal to or higher than that of the original data.
-Cells were considered tuned with $p<0.05$.
-
-For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 14 and 27 minutes (mean 22 minutes).
-Cells that did not have substantial spiking activity (those whose variance was less than 1\% of the median across the site) or whose activity was unstable (those whose variance in the least active quarter of the recording did not exceed 1\% of the variance in the most active quarter) were excluded from the analysis.
-
-\paragraph{Cross-validation.}
-To compare the performance of the estimators, we used conventional 10-fold cross-validation: Trials were randomly divided into 10 subsets with approximately equal numbers of trials of each condition in each subset.
-Each subset was then used as the testing sample with the rest of the data used as the training sample for estimating the covariance matrix.
-The average validation loss over the 10 folds was reported.
-
-Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}:  The outer loop evaluated the performance of the estimators with the hyperparameter values optimized by cross-validation within the inner loop.
-Hyperparameters were optimized by a two-phase search algorithm: random search to find a good starting point for the subsequent pattern search to find the global minimum.
-The inner cross-validation loop subdivided the training dataset from the outer loop to perform 10-fold cross-validation in order to evaluate each choice of the hyperparameter values.
-Thus the size of the training dataset within the inner loop comprised 81\% of the entire recording.
-Fig.~S1 illustrates the dependence of the validation loss on the hyperparameters of the $C_{\sf sparse+latent}$ estimator for the example site shown in Figures \ref{fig:2} and \ref{fig:4} and the optimal value found by the pattern search algorithm.
-
-When the validation loss was not required, only the inner loop of cross-validation was used on the entire dataset.
-This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (\figref{1}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:4}--\ref{fig:6}).
-
-\paragraph{Covariance estimation.}
-Within the inner loop of cross-validation, regularized covariance matrix estimation required only the sample covariance matrix $C_{\sf sample}$ of the training dataset and the hyperparameter values provided by the outer loop.
-
-Estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag})  used two hyperparameters: the covariance shrinkage intensity $\lambda \in [0,1]$ and variance shrinkage intensity $\alpha \in [0,1]$.
-The variances (the diagonal of $C_{\sf sample}$) were shrunk linearly toward their mean value $\frac 1 p \Tr(C_{\sf sample})$:
-\begin{equation}
-D = (1-\alpha)C_{\sf sample}\circ I + \alpha \frac 1 p \Tr(C_{\sf sample}) I
-\end{equation}
-The $C_{\sf diag}$ estimate was then obtained by shrinking $C_{\sf sample}$ toward $D$ according to Eq.~\ref{eq:c-diag}.
-
-In estimator $C_{\sf factor}$ (Eq.~\ref{eq:c-factor}), the low-rank matrix $L$ and the diagonal matrix $D$ are found by solving the minimization problem
-\begin{equation}
-(L,D) = \argmin\limits_{\hat L,\hat D} \loss{\hat L + \hat D,C_{\sf sample}},
-\end{equation}
-by an expectation-maximization (EM) algorithm with specified rank of $L$. Additional the diagonal matrix of individual variances is shrunk toward its mean value according to Eq.~\ref{eq:c-factor}.
-
-In estimator $C_{\sf sparse}$ (Eq.~\ref{eq:c-sparse}), the sparse precision matrix $S$ is found by minimizing the $L_1$-penalized loss with regularization parameter $\lambda$:
-\begin{equation}
-S = \argmin\limits_{\hat S \succ 0} \loss{{\hat S}^{-1},C_{\sf sample}} + \lambda \|\hat S \|_1
-\end{equation}
-where $\hat S\succ 0$ denotes the constraint that $\hat S$ be a positive-definite matrix and $\|\hat S\|_1$ is the element-wise $L_1$ norm of the matrix $\hat S$.
-This problem formulation is known as \emph{graphical lasso} \cite{Meinshausen:2006, Friedman:2008}.
-To solve this minimization problem, we adapted the alternative-direction method of multipliers (ADMM) \cite{Ma:2013}.
-Unlike $C_{\sf diag}$ and $C_{\sf factor}$, this estimator does not include linear shrinkage: the selection of the sparsity level provides sufficient flexibility to fine-tune the regularization level.
-
-Estimator $C_{\sf sparse+latent}$ (Eq.~\ref{eq:c-sl}) estimates a larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.  
-\begin{equation}
-S^\ast=
-\begin{pmatrix}
-S & S_{12} \\
-S_{12}^\T & S_{22}
-\end{pmatrix},
-\end{equation}
-where the $p\times p$ partition $S$ corresponds to the visible units. 
-Then the covariance matrix of the observed population is 
-\begin{equation}
-C_{\sf sparse+latent} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
-\end{equation}
-The rank of the $p\times p$  matrix $L=S_{12}S_{22}^{-1}S_{12}^\T$ matches the number of the latent units in the joint distribution.
-Rather than finding $S_{12}$ and $S_{22}$ separately, $L$ can be estimated as a low-rank matrix.
-We adapted the ADMM algorithm to minimize the loss function with $L_1$ penalty on $S$ to regulate its sparseness and nuclear norm penalty on $L$ to regulate its rank \cite{Chandrasekaran:2010,Ma:2013}:
-\begin{equation}\label{eq:ma}
-(S,L) = \argmin\limits_{\hat S,\hat L} \loss{\hat S-\hat L, C_{\sf sample}} + \alpha\|\hat S\|_1 + \beta\Tr(\hat L)
-\end{equation}
-
-The partial correlation matrix (Eq.~\ref{eq:partial}) computed from $C_{\sf sparse+latent}$ includes interactions between the visible and latent units and was used \figref{4}{C and D} and \figref{5}{C}, and \figref{6}{A--C}).
-The partial correlation matrix computed from $S$ alone expresses strengths of pairwise interactions
-\begin{equation}
-P_{\sf sparse} = -(S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
-\end{equation}
-and were used in \figref{4}{F, G, H}.
-
-The MATLAB code for these computations is available online at http://github.com/atlab/cov-est.
-
-\paragraph{Cross-validation with conditioned variances.}
-This study required an unbiased estimate of the discrepancy between various correlation matrix estimates and the true value of the correlation matrix.
-This was accomplished by cross-validation whereby regularized covariance matrix estimates computed from training samples were compared to unbiased covariance matrix estimates computed from independent testing samples.
-Loss functions were defined with respect to the covariance matrix rather than the correlation matrix because no unbiased estimate of the correlation matrix exists for finite sample sizes in the general case \cite{Fisher:1921}.
-In particular, we showed that with the normal loss function (Eq.~\ref{eq:loss}), the validation loss $\loss{C,C_{\sf sample}^\prime}$ is an unbiased estimate of true loss when $C_{\sf sample}^\prime$ is an unbiased estimate.
-However, this approach only worked under the assumption of uniform variances across all conditions included in the training and testing samples.
-Here, we extend the validation loss to the estimation and evaluation of a common correlation matrix across multiple conditions with different variances.
-
-Let $T_c$ and $T_c^\prime$ denote the time bin indices for the training and testing samples, respectively, limited to condition $c$.
-Here, the prime symbol marks quantities estimated from the testing sample.
-
-Similar to Eq.~\ref{eq:sample}, the training and testing sample covariance matrices for condition $c$ are
-\begin{equation}
-	C_{c,{\sf sample}}
-	= \frac 1{n_c} \sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
-\end{equation}
-and
-\begin{equation}
-	C_{c,{\sf sample}}^\prime
-	= \frac 1{n_c^\prime} \sum\limits_{t\in T_c^\prime}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
-\end{equation}
-Here $n_c$ and $n_c^\prime$ denote the sizes of $T_c$ and $T_c^\prime$, respectively.
-
-Note that $\bar x_c= \frac 1 {n_c} \sum\limits_{t \in T_c}x(t)$ is estimated from the training sample but used in both estimates, making $C_{c,\sf sample}^\prime$ an unbiased estimate of the true covariance matrix, $\Sigma$.
-As such, $C_{c,\sf sample}^\prime$ can be used for validation.
-
-The common correlation matrix $R_{\sf sample}$ is estimated by averaging the condition-specific correlations:
-\begin{equation}
-	R_{\sf sample}  
-	= \frac 1 n \sum\limits_c n_c \left(V_{c,\sf sample}^{-\frac 1 2}C_{c,\sf sample}V_{c,\sf sample}^{-\frac 1 2}\right)
-        = \frac 1 n \sum\limits_c \sum\limits_{t \in T_c} z(t)z(t)^\T,
-\end{equation}
-where $n=\sum\limits_c n_c$ and $V_{c,\sf sample} = C_{c,\sf sample} \circ I$ is the diagonal matrix containing the sample variances.
-Then $R_{\sf sample}$ is simply the covariance matrix of the $z$-score signal
-$z(t) = V_{c,\sf sample}^{-\frac 1 2} \left(x(t) - \bar x_c\right)$ of the training sample.
-
-For consistency with prior work, we applied regularization to covariance matrices rather than to correlation matrices.
-The common covariance matrix was estimated by scaling $R_{\sf sample}$ by the average variances across conditions $V_{\sf sample} = \frac 1 n \sum\limits_c n_c V_{c,\sf sample}$:
-\begin{equation}
-	C_{\sf sample} = V_{\sf sample}^{\frac 1 2} R_{\sf sample} V_{\sf sample}^{\frac 1 2}
-\end{equation}
-Note that $C_{\sf sample}$ differs from the sample covariance matrix computed without conditioning the variances on $c$ and this computation helps avoid any biases that would be introduced by ignoring changes in variance.
-
-The covariance matrix estimators $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$ or $C_{\sf sparse+latent}$ convert $C_{\sf sample}$ into its regularized counterpart denoted here as $C_{\sf reg}$.
-
-To evaluate the estimators, we regularized the conditioned variances by linear shrinkage toward their mean value across all conditions.
-This was done by scaling $C_{\sf reg}$ by the conditioned variance adjustment matrix $Q_c = \delta I + (1-\delta)V_{\sf sample}^{-1} V_{c,\sf sample}$ to produce the conditioned regularized covariance estimate:
-\begin{equation}
-	C_{c,\sf reg} = Q_c^{\frac 1 2} C_{\sf reg} Q_c^{\frac 1 2},
-\end{equation}
-The variance regularization parameter $\delta \in [0,1]$ was optimized in the inner loop of cross-validation along with the regularization hyperparameters for $C_{\sf reg}$.
-
-The overall loss function is obtained by averaging the loss functions across all the conditions:
-\begin{equation}\label{eq:full-loss}
-	\frac 1 {\sum\limits_c n_c^\prime}\sum\limits_c n_c^\prime \loss{C_{c,\sf reg}, C_{c,\sf sample}^\prime} 
-\end{equation}
-
-With the normal loss $\loss{\cdot,\cdot}$ (Eq.~\ref{eq:loss}) and the unbiased testing covariance matrix $C_{c,\sf sample}$, the loss function in Eq.~\ref{eq:full-loss} is an unbiased estimate of the true loss.
-Hence, it was used for evaluations reported in \figref{3}.
-
-\paragraph{Simulation.}
-For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units.
-The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{1}{\,Row 2}).
-Samples were then drawn from multivariate normal distributions with the respective true covariance matrices to be estimated by each of the estimators.
-
-\section*{Acknowledgments}
-We thank Genevera Allen for a helpful discussion, and Eftychios Pnevmatikakis for helpful suggestions and feedback on the manuscript.
-This work was supported by grants NEI R01-EY018847, NEI P30-EY002520-33, NEI T32-EY07001-37, NIDA R01DA028525, the NIH-Pioneer award DP1-OD008301 to AST, the McKnight Scholar Award to AST; the Arnold \& Beckman Foundation Young Investigator Award to AST; and NSF grants DMS-0817649, DMS-1122094 to KJ.
-% The bibtex filename
-\bibliography{references.bib}
-
-\section*{Figure Legends}
-
-\section*{Supporting Information}
-\subsection*{Supplementary Figures}
-\setcounter{figure}{0}
-\renewcommand{\figurename}{Figure S}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics{./figures/Supp01.pdf}
-\end{center}
-\caption{
-	{\bf Optimization of hyperparameters of the $C_{\sf sparse+latent}$ estimator.}  
-	{\bf A.} Validation loss (Eq.~\ref{eq:full-loss}) for the example site in \figref{2} and \figref{4} as a function of the hyperparameters $\alpha$ and $\beta$ of the $C_{\sf sparse+latent}$ estimator (Eq.~\ref{eq:c-sl} and Eq.~\ref{eq:ma}).
-In all panels, the red cross marks the optimal value found by the pattern search algorithm described in Methods.
-	{\bf B.} The connectivity ($1-\mbox{sparsity}$) of the sparse component $S$ as a function of $\alpha$ and $\beta$ for the example site.
-	{\bf C.} The number of latent units, \ie the rank of the low-rank component $L$ as a function of hyperparameters $\alpha$ and $\beta$.
-	{\bf D.} The loss function as a function of the connectivity and the number of latent units.
-}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics{./figures/src/Supp3.eps}
-\end{center}
-\caption{
-{\bf Performance of estimator $C_{\sf sparse+latent}$ with respect to \emph{quadratic loss} (Eq.~\ref{eq:quad}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.}
-All comparisons showed significant advantage of estimator $C_{\sf sparse+latent}$ ($p<10^{-3}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).}
-\label{supp:02}
-\end{figure}
+\section*{Supporting Information Legends}
+%
+% Please enter your Supporting Information captions below in the following format:
+%\item{\bf Figure SX. Enter mandatory title here.} Enter optional descriptive information here.
+% 
+%\begin{description}
+%\item {\bf}
+%\item {\bf}
+%\end{description}
 

### 2014-07-15 manuscript.tex
submitted

@@
 To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
-The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~\ref{supp:1} and  Methods).
+The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods).
 Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}).
@@
 \end{equation}
-instead of the normal loss function reproduced the same relationship between the estimators (Fig.~S\ref{supp:02}).
+instead of the normal loss function reproduced the same relationship between the estimators (Fig.~S2).
 This suggests that the results of this study are robust to the choice of the loss function and do not depend on the assumption of gaussianity.
@@
 This suggests that correlations between neurons build up from multiple chains of smaller interactions.
-Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29, with larger populations exhibiting more similar values of the average partial correlations than the smaller populations.
+Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29, with larger populations exhibiting greater uniformity of average partial correlations than the smaller populations.
 
@@
 
-\subsubsection*{Relationship of $C_{\sf sparse+latent}$ to orientation tuning and physical distances}
+\subsection*{Relationship of $C_{\sf sparse+latent}$ to orientation tuning and physical distances}
 
@@
 \section*{Discussion}
+\subsection*{Functional connectivity as a network of pairwise interactions}
+Functional connectivity is often represented as a graph of pairwise interactions.
+The goal of many studies of functional connectivity has been to estimate  anatomical connectivity from  observed multineuronal spiking activity.
+For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
+Such signatures are ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.
+With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.
+Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network and external input.
+Therefore, conditional independence can signify the absence of a direct causal influence.
+
+Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity.
+For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration, membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.
+Maximum entropy models constrained by observed pairwise correlations are among other models with pairwise coupling between cells \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.
+Assuming that the population response follows a multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.
+Each probabilistic model, fitted to the same data may reveal a completely different network of `interactions',  \emph{i.e.}\;conditional dependencies between pairs of cells.
+
+It is not yet clear which approach provides the best correspondence with anatomical connectivity.
+Little experimental evidence is available to answer this question.
+The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to anatomy.
+Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013, Sadovsky:2014}.
+However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}.
+Until a physiological interpretation of functional connectivity is established, the physiological relevance of such analyses remains in question and we did not attempt them here.
+
+Inference of the conditional dependencies also depends on the completeness of the recorded population:  To equate conditional dependencies to direct interaction between neurons, we must record from all neurons with which the pair interacts.
+Unobserved portions of the circuit may manifest as conditional dependencies between observed neurons that do not interact.
+For this reason, statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}.
+In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies.
+Perhaps for this reason, partial correlations have not, until now, been used to describe the functional connectivity in cortical populations.
+
+Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.
+While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}.
+However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.
+Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.
+The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.
+Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.
+These techniques may provide more incisive measurements of functional connectivity than electrophysiological recordings.
+
+The low temporal resolution of calcium signals limits the use of functional connectivity methods that rely on millisecond-scale binning of signals (cross-correlograms, some GLMs, and binary maximum entropy models).
+Hence, most studies of functional connectivity have relied on instantaneous sample correlations \cite{Greenberg:2008, Golshani:2009, Hofer:2011, Malmersjo:2013} .
+Although some investigators have interpreted such correlations as indicators of (chemical or electrical) synaptic connectivity, most used them as more general indicators of functional connectivity without relating them to underlying mechanisms.
+
+In this study, we sought to infer pairwise functional connectivity networks  in cortical microcircuits.
+We hypothesized that partial correlations correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.
+Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we \emph{a priori} favored solutions with sparse partial correlations.
+Under the assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
+
+\subsection*{Functional connectivity as coactivations}
+Another approach to describing the functional connectivity of a circuit is to isolate individual patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}.
+Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}.
+Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \emph{i.e.}\;groups of tightly interconnected groups of cells involved in a common computation.
+Coactivation patterns could also result from shared input from unobserved parts of the circuit, or global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.
+
+Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity.
+However, an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.
+In our study, the functional connectivity solely through modes of coactivations was represented by the factor analysis-based estimator $C_{\sf factor}$.
+
+\subsection*{Combining pairwise interactions and coactivations}
+In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables \cite{Koster:2013},  high-order correlations \cite{Ganmor:2011}, or global network fluctuations \cite{Tkacik:2013}.
+
+In our study, we combined pairwise interactions with collective coactivations by applying the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables \cite{Chandrasekaran:2010, Ma:2013}.
+The resulting estimator, $C_{\sf sparse+latent}$, effectively decomposed the functional connectivity into a sparse network of pairwise interactions and coactivation mode vectors.
+
+\subsection*{Addressing ill-posedness}
+Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.
+The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
+As techniques have improved to allow recording from larger neuronal populations, experimental neuroscientists have addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted \cite{Tkacik:2013}.
+However, ambitious projects, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
+Simply increasing recording duration will not be sufficient, and the problem must be addressed by using regularized estimators.
+Regularization biases the solution toward a small subspace in order to counteract the effect of  sampling noise in the empirical data.
+However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hinders interpretation.
+
+Several strategies have been developed to limit the model space in order to improve the quality of the estimate.
+For example, Ganmor et al.
+\cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in the retina.
+As another example of regularization, generalized linear models typically employ $L_1$ penalty terms to constrain the solution space and to effectively reduce the dimensionality of the solution \cite{Pillow:2008}.
+
+In our study, regularizations were accomplished by dimensionality reduction (feature selection) schemes to produce sparse, constrained solutions.
+Only the most efficient scheme was considered in the analysis of functional connectivity.
+
+\subsection*{Model selection}
+Various model selection criteria have been devised to select between families of models and the optimal subsets of variables in a given model family based on observed data.
+Despite its computational requirements, cross-validation is among the most popular model selection methods due to its minimal assumptions about the data generating process \cite{Arlot:2010}.
+
+We evaluated the covariance matrix estimators a loss function based on multivariate normal log likelihood.
+However, this does not limit the applicability of its conclusions to normal distributions.
+Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:3} and Fig.~S2).
+Other probabilistic models, fitted to the same data, could also serve as estimators of the covariance matrix.
+If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.
+
+The results of model selection must be interpreted with caution.
+As we demonstrated by simulation, even models with incorrect forms of dependencies can substantially improve estimates (Fig.~\ref{fig:1}).
+Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not necessarily support the conclusion that it reveals a better representation of dependencies in the data.
+This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary low-dimensional target can consistently outperform a less constrained estimate.
+
+\subsection*{Physiological interpretation and future directions}
+
+We showed that among several models a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix for cortical microcircuits.
+This finding is valuable in itself: improved estimates of the noise covariance matrix for large datasets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Ecker:2011}
+
+Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.
+Importantly, the inferred functional interactions were substantially different from the network of the highest sample correlations.
+For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (Fig.~\ref{fig:4} F) and may reflect inhibitory circuitry.
+
+Distances between cells in physical space and in sensory feature space had a stronger effect on the partial correlations estimated by the $C_{\sf sparse+latent}$ estimator than on sample correlations (Fig.~\ref{fig:6} A--C).
+These differences support the idea that correlations are built up from partial correlations in chains of intermediate cells positioned closer and tuned more similarly to one another, with potentially closer correspondence to anatomical connectivity.
+These differences may also be at least partially explained by a trivial effect of regularization: the $L_1$ penalty applied by the estimator (Eq.~\ref{eq:ma}) suppresses small partial correlations to greater extent than large partial correlations, enhancing the apparent effect of distance and tuning.
+Distances between cells in physical space and in sensory feature space also had distinct effects on the positive and negative connectivities inferred by the estimator (Fig.~\ref{fig:6} D--F); we could not think of a numerical artifact that could produce this effect.
+The distinct positive and negative connectivity patterns  may reflect geometric and graphical features of local excitatory and inhibitory networks.
+Indeed, the relationships between patterns of positive and negative connectivities inferred by the estimator mirrored the properties of excitatory and inhibitory synaptic connectivities with respect to distance, cortical layers, and feature tuning \cite{Song:2005, Oswald:2008, Adesnik:2010, Perin:2011, Fino:2011, Hofer:2011, Isaacson:2011, Levy:2012}.
+For example, while excitatory neurons form synapses within highly specific local cliques \cite{Perin:2011}, inhibitory interneurons form synapses with nearly all excitatory cells within local microcircuits \cite{Fino:2011, Hofer:2011, Packer:2011}.
+To further investigate the link between synaptic connectivity and inferred functional connectivity, in future experiments, we will use molecular markers for various cell types with follow-up multiple whole-cell \emph{in vitro} recordings \cite{Hofer:2011, Ko:2013} to directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry.
+Finally, the latent units inferred by the estimator can be analyzed for their physiological functions.
+For example, these latent units may be modulated under different brain states (e.g.
+slow-wave sleep, attention) and stimulus conditions (e.g.
+certain types of stimuli may engage feedback connections).
 
@@
 
+\subsection*{Ethics statement}
+All procedures were conducted in accordance with the ethical guidelines of the National Institutes of Health and were approved by the Baylor College of Medicine IACUC.
+
+\subsection*{Surgery and two-photon imaging}
+The surgical procedures and data acquisition were performed as described in \cite{Cotton:2013}: C57BL/6J mice (aged p40--60) were used.
+Anesthesia was initiated with isoflurane (3\%) and maintained with the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
+A craniotomy was performed over the right primary visual cortex.
+Membrane-permeant calcium indicator Oregon Green 488 BAPTA-1 AM (OGB-1, Invitrogen) was loaded by bolus injection.
+The craniotomy was sealed using a glass coverslip secured with dental cement.
+
+Calcium imaging began 1 hour after dye injection.
+All imaging was performed using 3D-RAMP two-photon microscopy \cite{Cotton:2013}.
+First, a 3D stack was acquired and cells were manually segmented.
+Then calcium signal were collected by sampling in the center of each cell at rates of 100 Hz or higher, depending on the number of cells.
+
+\subsection*{Visual stimulus}
+The visual stimulus consisted of full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s.
+Two types of stimuli were presented for each imaging site: First, directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen directions of motion, 500 ms per direction, without blanks, with 12--30 trials for each direction of motion.
+Second, ro measure correlations, the stimulus was modified to include only two directions of motion (in 9 datasets) or five directions (in 22 datasets) and the gratings were presented for 1 second and were separated by 1-second blanks, with 100--300 trials for each direction of motion.
+
+\subsection*{Data processing}
+All data were processed in MATLAB using the DataJoint data processing chain toolbox (http://datajoint.github.com).
+
+The measured fluorescent traces were deconvolved to reconstruct the firing rates for each neuron: First, the first principal component was subtracted from the raw traces in order to reduce common mode noise related to small cardiovascular movements \cite{Cotton:2013}.
+The resulting traces were high-pass filtered above 0.1 Hz and downsampled to 20 Hz (Fig.~\ref{fig:2} C).
+Then, the firing rates were estimated using by nonnegative deconvolution \cite{Vogelstein:2010}.
+
+Orientation tuning was computed by fitting the mean firing rates for each direction of motion $\phi$ using two-peaked von Mises tuning functions $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction.
+The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times; the $p$-values of the fits were computed as the fraction of permutations that yielded $R^2$ equal to or higher than that of the original data.
+Cells were considered tuned with $p<0.05$.
+
+For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 14 and 27 minutes (mean 22 minutes).
+Cells that did not have substantial spiking activity (those whose variance was less than 1\% of the median across the site) or whose activity was unstable (those whose variance in the least active quarter of the recording did not exceed 1\% of the variance in the most active quarter) were excluded from the analysis.
+
+\subsection*{Cross-validation}
+To compare the performance of the estimators, we used conventional 10-fold cross-validation: Trials were randomly divided into 10 subsets with approximately equal numbers of trials of each condition in each subset.
+Each subset was then used as the testing sample with the rest of the data used as the training sample for estimating the covariance matrix.
+The average validation loss over the 10 folds was reported.
+
+Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}:  The outer loop evaluated the performance of the estimators with the hyperparameter values optimized by cross-validation within the inner loop.
+Hyperparameters were optimized by a two-phase search algorithm: random search to find a good starting point for the subsequent pattern search to find the global minimum.
+The inner cross-validation loop subdivided the training dataset from the outer loop to perform 10-fold cross-validation in order to evaluate each choice of the hyperparameter values.
+Thus the size of the training dataset within the inner loop comprised 81\% of the entire recording.
+Fig.~S1 illustrates the dependence of the validation loss on the hyperparameters of the $C_{\sf sparse+latent}$ estimator for the example site shown in Figures \ref{fig:2} and \ref{fig:4} and the optimal value found by the pattern search algorithm.
+
+When the validation loss was not required, only the inner loop of cross-validation was used on the entire dataset.
+This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (Fig.~\ref{fig:1} Rows 4 and 5) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:4}--\ref{fig:6}).
+
+\subsection*{Covariance estimation}
+Within the inner loop of cross-validation, regularized covariance matrix estimation required only the sample covariance matrix $C_{\sf sample}$ of the training dataset and the hyperparameter values provided by the outer loop.
+
+Estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag})  used two hyperparameters: the covariance shrinkage intensity $\lambda \in [0,1]$ and variance shrinkage intensity $\alpha \in [0,1]$.
+The variances (the diagonal of $C_{\sf sample}$) were shrunk linearly toward their mean value $\frac 1 p \Tr(C_{\sf sample})$:
+\begin{equation}
+D = (1-\alpha)C_{\sf sample}\circ I + \alpha \frac 1 p \Tr(C_{\sf sample}) I
+\end{equation}
+The $C_{\sf diag}$ estimate was then obtained by shrinking $C_{\sf sample}$ toward $D$ according to Eq.~\ref{eq:c-diag}.
+
+In estimator $C_{\sf factor}$ (Eq.~\ref{eq:c-factor}), the low-rank matrix $L$ and the diagonal matrix $D$ are found by solving the minimization problem
+\begin{equation}
+(L,D) = \argmin\limits_{\hat L,\hat D} \loss{\hat L + \hat D,C_{\sf sample}},
+\end{equation}
+by an expectation-maximization (EM) algorithm with specified rank of $L$. Additional the diagonal matrix of individual variances is shrunk toward its mean value according to Eq.~\ref{eq:c-factor}.
+
+In estimator $C_{\sf sparse}$ (Eq.~\ref{eq:c-sparse}), the sparse precision matrix $S$ is found by minimizing the $L_1$-penalized loss with regularization parameter $\lambda$:
+\begin{equation}
+S = \argmin\limits_{\hat S \succ 0} \loss{{\hat S}^{-1},C_{\sf sample}} + \lambda \|\hat S \|_1
+\end{equation}
+where $\hat S\succ 0$ denotes the constraint that $\hat S$ be a positive-definite matrix and $\|\hat S\|_1$ is the element-wise $L_1$ norm of the matrix $\hat S$.
+This problem formulation is known as \emph{graphical lasso} \cite{Meinshausen:2006, Friedman:2008}.
+To solve this minimization problem, we adapted the alternative-direction method of multipliers (ADMM) \cite{Ma:2013}.
+Unlike $C_{\sf diag}$ and $C_{\sf factor}$, this estimator does not include linear shrinkage: the selection of the sparsity level provides sufficient flexibility to fine-tune the regularization level.
+
+Estimator $C_{\sf sparse+latent}$ (Eq.~\ref{eq:c-sl}) estimates a larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.
+\begin{equation}
+S^\ast=
+\begin{pmatrix}
+S & S_{12} \\
+S_{12}^\T & S_{22}
+\end{pmatrix},
+\end{equation}
+where the $p\times p$ partition $S$ corresponds to the visible units.
+Then the covariance matrix of the observed population is
+\begin{equation}
+C_{\sf sparse+latent} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
+\end{equation}
+The rank of the $p\times p$  matrix $L=S_{12}S_{22}^{-1}S_{12}^\T$ matches the number of the latent units in the joint distribution.
+Rather than finding $S_{12}$ and $S_{22}$ separately, $L$ can be estimated as a low-rank matrix.
+We adapted the ADMM algorithm to minimize the loss function with $L_1$ penalty on $S$ to regulate its sparseness and nuclear norm penalty on $L$ to regulate its rank \cite{Chandrasekaran:2010,Ma:2013}:
+\begin{equation}\label{eq:ma}
+(S,L) = \argmin\limits_{\hat S,\hat L} \loss{\hat S-\hat L, C_{\sf sample}} + \alpha\|\hat S\|_1 + \beta\Tr(\hat L)
+\end{equation}
+
+The partial correlation matrix (Eq.~\ref{eq:partial}) computed from $C_{\sf sparse+latent}$ includes interactions between the visible and latent units and was used in Fig.~\ref{fig:4} C and D and Fig.~\ref{fig:5} C, and Fig.~\ref{fig:6} A--C).
+The partial correlation matrix computed from $S$ alone expresses strengths of pairwise interactions
+\begin{equation}
+P_{\sf sparse} = -(S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
+\end{equation}
+and were used in Fig.~\ref{fig:4} F, G, H.
+
+The MATLAB code for these computations is available online at http://github.com/atlab/cov-est.
+\subsection*{Cross-validation with conditioned variances}
+This study required an unbiased estimate of the discrepancy between various correlation matrix estimates and the true value of the correlation matrix.
+This was accomplished by cross-validation whereby regularized covariance matrix estimates computed from training samples were compared to unbiased covariance matrix estimates computed from independent testing samples.
+Loss functions were defined with respect to the covariance matrix rather than the correlation matrix because no unbiased estimate of the correlation matrix exists for finite sample sizes in the general case \cite{Fisher:1921}.
+In particular, we showed that with the normal loss function (Eq.~\ref{eq:loss}), the validation loss $\loss{C,C_{\sf sample}^\prime}$ is an unbiased estimate of true loss when $C_{\sf sample}^\prime$ is an unbiased estimate.
+However, this approach only worked under the assumption of uniform variances across all conditions included in the training and testing samples.
+Here, we extend the validation loss to the estimation and evaluation of a common correlation matrix across multiple conditions with different variances.
+
+Let $T_c$ and $T_c^\prime$ denote the time bin indices for the training and testing samples, respectively, limited to condition $c$.
+Here, the prime symbol marks quantities estimated from the testing sample.
+
+Similar to Eq.~\ref{eq:sample}, the training and testing sample covariance matrices for condition $c$ are
+\begin{equation}
+    C_{c,{\sf sample}}
+    = \frac 1{n_c} \sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
+\end{equation}
+and
+\begin{equation}
+    C_{c,{\sf sample}}^\prime
+    = \frac 1{n_c^\prime} \sum\limits_{t\in T_c^\prime}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
+\end{equation}
+Here $n_c$ and $n_c^\prime$ denote the sizes of $T_c$ and $T_c^\prime$, respectively.
+
+Note that $\bar x_c= \frac 1 {n_c} \sum\limits_{t \in T_c}x(t)$ is estimated from the training sample but used in both estimates, making $C_{c,\sf sample}^\prime$ an unbiased estimate of the true covariance matrix, $\Sigma$.
+As such, $C_{c,\sf sample}^\prime$ can be used for validation.
+
+The common correlation matrix $R_{\sf sample}$ is estimated by averaging the condition-specific correlations:
+\begin{equation}
+    R_{\sf sample}  
+    = \frac 1 n \sum\limits_c n_c \left(V_{c,\sf sample}^{-\frac 1 2}C_{c,\sf sample}V_{c,\sf sample}^{-\frac 1 2}\right)
+    = \frac 1 n \sum\limits_c \sum\limits_{t \in T_c} z(t)z(t)^\T,
+\end{equation}
+where $n=\sum\limits_c n_c$ and $V_{c,\sf sample} = C_{c,\sf sample} \circ I$ is the diagonal matrix containing the sample variances.
+Then $R_{\sf sample}$ is simply the covariance matrix of the $z$-score signal $z(t) = V_{c,\sf sample}^{-\frac 1 2} \left(x(t) - \bar x_c\right)$ of the training sample.
+
+For consistency with prior work, we applied regularization to covariance matrices rather than to correlation matrices.
+The common covariance matrix was estimated by scaling $R_{\sf sample}$ by the average variances across conditions $V_{\sf sample} = \frac 1 n \sum\limits_c n_c V_{c,\sf sample}$:
+\begin{equation}
+    C_{\sf sample} = V_{\sf sample}^{\frac 1 2} R_{\sf sample} V_{\sf sample}^{\frac 1 2}
+\end{equation}
+Note that $C_{\sf sample}$ differs from the sample covariance matrix computed without conditioning the variances on $c$ and this computation helps avoid any biases that would be introduced by ignoring changes in variance.
+
+The covariance matrix estimators $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$ or $C_{\sf sparse+latent}$ convert $C_{\sf sample}$ into its regularized counterpart denoted here as $C_{\sf reg}$.
+
+To evaluate the estimators, we regularized the conditioned variances by linear shrinkage toward their mean value across all conditions.
+This was done by scaling $C_{\sf reg}$ by the conditioned variance adjustment matrix $Q_c = \delta I + (1-\delta)V_{\sf sample}^{-1} V_{c,\sf sample}$ to produce the conditioned regularized covariance matrix estimate:
+\begin{equation}
+    C_{c,\sf reg} = Q_c^{\frac 1 2} C_{\sf reg} Q_c^{\frac 1 2}
+\end{equation}
+
+The variance regularization parameter $\delta \in [0,1]$ was optimized in the inner loop of cross-validation along with the other hyperparameters.
+
+The overall loss function is obtained by averaging the loss functions across all the conditions:
+\begin{equation}\label{eq:full-loss}
+    \frac 1 {\sum\limits_c n_c^\prime}\sum\limits_c n_c^\prime \loss{C_{c,\sf reg}, C_{c,\sf sample}^\prime} 
+\end{equation}
+
+With the normal loss $\loss{\cdot,\cdot}$ (Eq.~\ref{eq:loss}) and the unbiased testing covariance matrix $C_{c,\sf sample}$, the loss function in Eq.~\ref{eq:full-loss} is an unbiased estimate of the true loss.
+Hence, it was used for evaluations reported in Fig.~\ref{fig:3}.
+
+\subsection*{Simulation}
+For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units.
+The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (Fig.~\ref{fig:1} Row 2).
+Samples were then drawn from multivariate normal distributions with the respective true covariance matrices to be estimated by each of the estimators.
+
 % Do NOT remove this, even if you are not including acknowledgments.
@@
 \section*{Acknowledgments}
+We thank Genevera Allen for a helpful discussion, and Eftychios Pnevmatikakis for helpful suggestions and feedback on the manuscript.
+This work was supported by grants NEI R01-EY018847, NEI P30-EY002520-33, NEI T32-EY07001-37, NIDA R01DA028525, the NIH-Pioneer award DP1-OD008301 to AST, the McKnight Scholar Award to AST; the Arnold \& Beckman Foundation Young Investigator Award to AST; and NSF grants DMS-0817649, DMS-1122094 to KJ.
 
@@
 
+\newpage
 \section*{Figure Legends}
@@
 
-\begin{figure}
+\begin{figure}[h!]
 \caption{
@@
 
+\newpage
 \section*{Supporting Information Legends}
@@
 % 
-%\begin{description}
-%\item {\bf}
-%\item {\bf}
-%\end{description}
+\begin{description}
+\item {\bf Figure S1.
+Optimization of hyperparameters of the $C_{\sf sparse+latent}$ estimator.} {\bf A.} Validation loss (Eq.~\ref{eq:full-loss}) for the example site in Fig.~\ref{fig:2} and \ref{fig:4} as a function of the hyperparameters $\alpha$ and $\beta$ of the $C_{\sf sparse+latent}$ estimator (Eq.~\ref{eq:c-sl} and Eq.~\ref{eq:ma}).
+In all panels, the red cross marks the optimal value found by the pattern search algorithm described in Methods.
+{\bf B.} The connectivity ($1-\mbox{sparsity}$) of the sparse component $S$ as a function of $\alpha$ and $\beta$ for the example site.
+{\bf C.} The number of latent units, \emph{i.e.}~the rank of the low-rank component $L$, as a function of hyperparameters $\alpha$ and $\beta$.
+{\bf D.} The loss function as a function of the connectivity and the number of latent units.
+
+\item {\bf Figure S2.
+Performance of estimator $C_{\sf sparse+latent}$ with respect to \emph{quadratic loss} (Eq.~\ref{eq:quad}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.} All comparisons showed significant advantage of estimator $C_{\sf sparse+latent}$ ($p<10^{-3}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
+\end{description}
 

### 2014-07-16 manuscript.tex
resubmission after suggested corrrections

@@
 \section*{Author Summary}
-It is now possible to record the spiking activity of hundreds of neurons at the same time.
-A meaningful statistical description of the collective activity of these neural populations -- their `functional connectivity' -- is a forefront challenge in neuroscience.
-We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.
-Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.
-In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (`interactions') combined with a low-rank component.
-The inferred interactions were both positive (`excitatory') and negative (`inhibitory') and reflected the spatial organization and orientation preferences of the interacting cells.
-We propose that  the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.
-
-\section*{Introduction}
 It is now possible to record the spiking activity of hundreds of neurons at the same time.
@@
 We thank Genevera Allen for a helpful discussion, and Eftychios Pnevmatikakis for helpful suggestions and feedback on the manuscript.
-This work was supported by grants NEI R01-EY018847, NEI P30-EY002520-33, NEI T32-EY07001-37, NIDA R01DA028525, the NIH-Pioneer award DP1-OD008301 to AST, the McKnight Scholar Award to AST; the Arnold \& Beckman Foundation Young Investigator Award to AST; and NSF grants DMS-0817649, DMS-1122094 to KJ.
 

### 2014-10-29 manuscript.tex
added Ising model

@@
 \begin{equation}\label{eq:precision}
-    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2},
+    R = (\mbox{diag}(C))^{-\frac 1 2} C (\mbox{diag}(C))^{-\frac 1 2},
 \end{equation}
-where $\circ$ denotes the entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix.
+where $\mbox{diag}(C)$ denotes the diagonal matrix with the diagonal elements from $C$.
 
@@
 \begin{equation}\label{eq:partial}
-    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
+    P = -\left(\mbox{diag}(C^{-1})\right)^{-\frac 1 2} C^{-1} \left(\mbox{diag}(C^{-1})\right)^{-\frac 1 2}
 \end{equation}
@@
 \end{equation}
-where, as above, $S$ is a sparse matrix and $L$ is a low-rank matrix.
+where $S$ is a sparse matrix and $L$ is a low-rank matrix.
 The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$.
@@
 We constructed four families of $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (Fig.~\ref{fig:1} Row 2, A--D and Methods).
-We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes.
-Even with relatively large sample sizes (\emph{e.g.}\;$n=500$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (Fig.~\ref{fig:1} Row 3).
+We used these covariance matrices as the ground truth in multivariate normal (MVN) distributions with zero means and drew samples of various sizes.
+The sample correlation matrices from finite samples (\emph{e.g.}\ $n=500$ in Fig.~\ref{fig:1} Row 3) were contaminated with sampling noise and their underlying structures were difficult to discern.
 
@@
 
-Cross-validation loss accurately reproduced the relationships between the excess losses of the estimators (Fig.~\ref{fig:1} Row 6), confirming that cross-validation can evaluate true loss.
+Cross-validation loss accurately reproduced the ranking of estimators (Fig.~\ref{fig:1} Row 6), confirming that cross-validation can evaluate the true loss of the estimators.
+
+To demonstrate that estimator rankings did not depend on the assumption of gaussianity, we repeated the cross-validated evaluation with a non-Gaussian data generating model, non-Gaussian loss function, or both:
+
+First, we replaced the data-generating model with the pairwise Ising model \cite{Hertz:2011}, which has equivalent mathematical form to the MVN model,
+\begin{equation}
+    x\sim \frac 1 {Z(S,h)} \exp\left(-\frac 1 2 x^\T S x + h^\T x \right),
+\end{equation}
+but defined on the binary domain $x\in\{-1,+1\}^p$ rather than the continuous domain of the MVN model. The partition function $Z(S,h)$ normalizes the distribution on the models' respective domains. 
+For the MVN, the matrix $S$ is the precision matrix or the inverse of the covariance matrix; and the mean values are $\mu=-S^{-1}h$.
+In the Ising model, $-S$ is the matrix of pairwise interactions and $h$ is the vector of intrinsic activity drives of individual cells but $S$ and $h$ do not have a closed-form relationship to the mean activity and the covariance matrix of the distribution.
+However, both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $S$ indicate conditional independence between the corresponding pair of neurons.
+
+First, replacing the normal loss function with the quadratic loss function, 
+\begin{equation}\label{eq:quad}
+\loss{C,C_{\sf sample}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf sample}^\prime-I)^2
+\end{equation}
+did not affect estimator rankings (Fig.~S1 A--D).
 
@@
 To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
-The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods).
+The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S2 and  Methods).
 Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}).
@@
 The same evaluation based on the quadratic loss function,
-\begin{equation}\label{eq:quad}
-\loss{C,C_{\sf sample}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf sample}^\prime-I)^2,
-\end{equation}
-instead of the normal loss function reproduced the same relationship between the estimators (Fig.~S2).
+instead of the normal loss function reproduced the same relationship between the estimators (Fig.~S3).
+
 This suggests that the results of this study are robust to the choice of the loss function and do not depend on the assumption of gaussianity.
@@
 However, this does not limit the applicability of its conclusions to normal distributions.
-Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:3} and Fig.~S2).
+Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:3} and Fig.~S3).
 Other probabilistic models, fitted to the same data, could also serve as estimators of the covariance matrix.
@@
 Thus the size of the training dataset within the inner loop comprised 81\% of the entire recording.
-Fig.~S1 illustrates the dependence of the validation loss on the hyperparameters of the $C_{\sf sparse+latent}$ estimator for the example site shown in Figures \ref{fig:2} and \ref{fig:4} and the optimal value found by the pattern search algorithm.
+Fig.~S2 illustrates the dependence of the validation loss on the hyperparameters of the $C_{\sf sparse+latent}$ estimator for the example site shown in Figures \ref{fig:2} and \ref{fig:4} and the optimal value found by the pattern search algorithm.
 
@@
 \begin{equation}
-D = (1-\alpha)C_{\sf sample}\circ I + \alpha \frac 1 p \Tr(C_{\sf sample}) I
+D = (1-\alpha)\mbox{diag}(C_{\sf sample}) + \alpha \frac 1 p \Tr(C_{\sf sample}) I
 \end{equation}
@@
 \begin{equation}
-P_{\sf sparse} = -(S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
+P_{\sf sparse} = -(\mbox{diag}(S))^{-\frac 1 2} S  (\mbox{diag}(S))^{-\frac 1 2}
 \end{equation}
@@
 \end{equation}
-where $n=\sum\limits_c n_c$ and $V_{c,\sf sample} = C_{c,\sf sample} \circ I$ is the diagonal matrix containing the sample variances.
+where $n=\sum\limits_c n_c$ and $V_{c,\sf sample} = \mbox{diag}(C_{c,\sf sample})$ is the diagonal matrix containing the sample variances.
 Then $R_{\sf sample}$ is simply the covariance matrix of the $z$-score signal $z(t) = V_{c,\sf sample}^{-\frac 1 2} \left(x(t) - \bar x_c\right)$ of the training sample.
@@
 % 
-\bibliography{references.bib}
+\bibliography{references}
 
@@
 \begin{description}
-\item {\bf Figure S1.
+\item {\bf Figure S2.
 Optimization of hyperparameters of the $C_{\sf sparse+latent}$ estimator.} {\bf A.} Validation loss (Eq.~\ref{eq:full-loss}) for the example site in Fig.~\ref{fig:2} and \ref{fig:4} as a function of the hyperparameters $\alpha$ and $\beta$ of the $C_{\sf sparse+latent}$ estimator (Eq.~\ref{eq:c-sl} and Eq.~\ref{eq:ma}).
@@
 
-\item {\bf Figure S2.
+\item {\bf Figure S3.
 Performance of estimator $C_{\sf sparse+latent}$ with respect to \emph{quadratic loss} (Eq.~\ref{eq:quad}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.} All comparisons showed significant advantage of estimator $C_{\sf sparse+latent}$ ($p<10^{-3}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).

### 2014-10-30 manuscript.tex
removed plos.tex and plos.pdf, updated manuscript.tex

@@
 
-To demonstrate that estimator rankings did not depend on the assumption of gaussianity, we repeated the cross-validated evaluation with a non-Gaussian data generating model, non-Gaussian loss function, or both:
-
-First, we replaced the data-generating model with the pairwise Ising model \cite{Hertz:2011}, which has equivalent mathematical form to the MVN model,
+To demonstrate that estimator rankings did not depend on the assumption of gaussianity, we repeated the cross-validated evaluation with a non-Gaussian data-generating model, the pairwise Ising model (Fig.~S1).
+The Ising model \cite{Hertz:2011} has equivalent mathematical form to the MVN model,
 \begin{equation}
@@
 \end{equation}
-but defined on the binary domain $x\in\{-1,+1\}^p$ rather than the continuous domain of the MVN model. The partition function $Z(S,h)$ normalizes the distribution on the models' respective domains. 
-For the MVN, the matrix $S$ is the precision matrix or the inverse of the covariance matrix; and the mean values are $\mu=-S^{-1}h$.
-In the Ising model, $-S$ is the matrix of pairwise interactions and $h$ is the vector of intrinsic activity drives of individual cells but $S$ and $h$ do not have a closed-form relationship to the mean activity and the covariance matrix of the distribution.
-However, both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $S$ indicate conditional independence between the corresponding pair of neurons.
-
-First, replacing the normal loss function with the quadratic loss function, 
-\begin{equation}\label{eq:quad}
-\loss{C,C_{\sf sample}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf sample}^\prime-I)^2
-\end{equation}
-did not affect estimator rankings (Fig.~S1 A--D).
+but is defined on the multivariate binary domain $x\in\{-1,+1\}^p$ rather than the continuous domain of the MVN model. 
+The partition function $Z(S,h)$ normalizes the distribution on the models' respective domains. 
+In the MVN model, the matrix $S$ is the precision matrix or the inverse of the covariance matrix; and the mean values are $\mu=-S^{-1}h$.  
+For the Ising model, $-S$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives.
+In contrast to the MVN model,  $S$ and $h$ do not have a closed-form relationship to the mean firing rates or their covariance matrix.
+Both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $S$ indicate conditional independence between the corresponding pair of neurons.
 
@@
 The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).
-
-The same evaluation based on the quadratic loss function,
-instead of the normal loss function reproduced the same relationship between the estimators (Fig.~S3).
-
-This suggests that the results of this study are robust to the choice of the loss function and do not depend on the assumption of gaussianity.
 
@@
 However, this does not limit the applicability of its conclusions to normal distributions.
-Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:3} and Fig.~S3).
 Other probabilistic models, fitted to the same data, could also serve as estimators of the covariance matrix.
@@
 \begin{description}
+\item {\bf Figure S1. Performance of covariance estimators when the ground truth is the  Ising model.} 
+{\bf A.} With the ground truth having a diagonal matrix of dependencies (no pairwise dependencies),
+
 \item {\bf Figure S2.
@@
 {\bf D.} The loss function as a function of the connectivity and the number of latent units.
-
-\item {\bf Figure S3.
-Performance of estimator $C_{\sf sparse+latent}$ with respect to \emph{quadratic loss} (Eq.~\ref{eq:quad}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.} All comparisons showed significant advantage of estimator $C_{\sf sparse+latent}$ ($p<10^{-3}$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
 \end{description}
-
 \end{document}

### 2014-11-05 manuscript.tex
completed resubmission 2

@@
 We constructed four families of $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (Fig.~\ref{fig:1} Row 2, A--D and Methods).
-We used these covariance matrices as the ground truth in multivariate normal (MVN) distributions with zero means and drew samples of various sizes.
+We used these covariance matrices as the ground truth in multivariate Gaussian distributions with zero means and drew samples of various sizes.
 The sample correlation matrices from finite samples (\emph{e.g.}\ $n=500$ in Fig.~\ref{fig:1} Row 3) were contaminated with sampling noise and their underlying structures were difficult to discern.
@@
 
-Cross-validation loss accurately reproduced the ranking of estimators (Fig.~\ref{fig:1} Row 6), confirming that cross-validation can evaluate the true loss of the estimators.
+Cross-validation accurately reproduced relative values of the loss and the rankings of estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).
 
-To demonstrate that estimator rankings did not depend on the assumption of gaussianity, we repeated the cross-validated evaluation with a non-Gaussian data-generating model, the pairwise Ising model (Fig.~S1).
-The Ising model \cite{Hertz:2011} has equivalent mathematical form to the MVN model,
-\begin{equation}
-    x\sim \frac 1 {Z(S,h)} \exp\left(-\frac 1 2 x^\T S x + h^\T x \right),
+To demonstrate that estimator rankings did not strongly depend on the gaussianity of the data-generating process, we repeated the cross-validated evaluation after replacing the Gaussian data-generating models with pairwise Ising models.
+Ising models have been used to infer functional connectivity from neuronal spike trains \cite{Hertz:2011}. 
+Conveniently, the Ising model has equivalent mathematical form to the Gaussian distribution,
+\begin{equation}\label{eq:ising}
+    x\sim \frac 1 {Z(S,h)} \exp\left(-\frac 1 2 x^\T S x + h^\T x \right)
 \end{equation}
-but is defined on the multivariate binary domain $x\in\{-1,+1\}^p$ rather than the continuous domain of the MVN model. 
+but the Ising model is defined on the multivariate binary domain $x\in\{-1,+1\}^p$ rather than the continuous domain of the Gaussian  model. 
+Both models are maximum-entropy models constrained by the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
 The partition function $Z(S,h)$ normalizes the distribution on the models' respective domains. 
-In the MVN model, the matrix $S$ is the precision matrix or the inverse of the covariance matrix; and the mean values are $\mu=-S^{-1}h$.  
-For the Ising model, $-S$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives.
-In contrast to the MVN model,  $S$ and $h$ do not have a closed-form relationship to the mean firing rates or their covariance matrix.
+In the Gaussian model, the matrix $S$ is the precision matrix or the inverse of the covariance matrix; and the mean values are $\mu=-S^{-1}h$.  
+For the Ising model, $-S$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives but they do not have a simple relationship to the means and the covariance matrix.
 Both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $S$ indicate conditional independence between the corresponding pair of neurons. 
+Despite the departure from strictly linear dependencies between the variables characteristic of the Gaussian distribution, the Ising model yielded the same rankings of estimators as the Gaussian model in cross-validation (Fig.~S1).
 
@@
 
-First, we analyzed how correlations and connectivity depended on the difference in preferred orientations ($\Delta ori$) of pairs of significantly ($\alpha=0.05$) tuned cells.
-The partial correlations decayed more rapidly with $\Delta ori$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test of the difference of the linear regression coefficients).
-Positive connectivity decreased with $\Delta ori$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression coefficient) whereas negative connectivity did not decrease (Fig.~\ref{fig:6} D): The slope in the logistic model of connectivity with respect to $\Delta ori$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficient).
+First, we analyzed how correlations and connectivity depended on the difference in preferred orientations ($\Delta \mbox{ori}$) of pairs of significantly ($\alpha=0.05$) tuned cells.
+The partial correlations decayed more rapidly with $\Delta\mbox{ori}$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test of the difference of the linear regression coefficients).
+Positive connectivity decreased with $\Delta\mbox{ori}$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression coefficient) whereas negative connectivity did not decrease (Fig.~\ref{fig:6} D): The slope in the logistic model of connectivity with respect to $\Delta\mbox{ori}$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficient).
 
@@
 The rank of the $p\times p$  matrix $L=S_{12}S_{22}^{-1}S_{12}^\T$ matches the number of the latent units in the joint distribution.
-Rather than finding $S_{12}$ and $S_{22}$ separately, $L$ can be estimated as a low-rank matrix.
-We adapted the ADMM algorithm to minimize the loss function with $L_1$ penalty on $S$ to regulate its sparseness and nuclear norm penalty on $L$ to regulate its rank \cite{Chandrasekaran:2010,Ma:2013}:
+Rather than finding $S_{12}$ and $S_{22}$ separately, $L$ can be estimated as a low-rank positive semidefinite matrix.
+To simultaneously optimize the sparse component $S$ and the low-rank component $L$, we adapted the loss function with $L_1$ penalty on $S$ combined with a penalty on the trace of $L$ \cite{Chandrasekaran:2010,Ma:2013}:
 \begin{equation}\label{eq:ma}
@@
 \end{equation}
+The trace of a symmetric semidefinite matrix equals the sum of the absolute values of its eigenvalues, \emph{i.e.} its \emph{nuclear norm}; penalty on $\Tr(L)$ favors solutions with few non-zero eigenvalues or, equivalently, low-rank solutions while keeping the convexity of the overall optimization problem \cite{Fazel:2002,Recht:2010}.
+This allows using convex optimization algorithm such as ADMM to be applied with great computational efficiency \cite{Ma:2013}.
 
@@
 \begin{figure}
-\caption{{\bf Dependence of sample correlations, regularized partial correlations, and connectivity inferred by $C_{\sf sparse+latent}$ on the differences in preferred orientations, $\Delta ori$, and physical distances, $\Delta x$ and $\Delta z$.} The error bars mark the standard errors of the means.
+\caption{
+{\bf Dependence of sample correlations, regularized partial correlations, and connectivity inferred by $C_{\sf sparse+latent}$ on the differences in preferred orientations, $\Delta \mbox{ori}$, and physical distances: horizontal $\Delta x$ and depth $\Delta z$.}
 Five sites with highest connectivity (see Fig.~\ref{fig:5} B) were selected for this analysis.
-{\bf A--C.} Normalized mean sample correlations (black) and normalized mean partial correlations (red) estimated by $C_{\sf sparse+latent}$ across $n=5$ imaged sites.
-The values in each bin are normalized by the means across the entire site, shown in Fig.~\ref{fig:5} C, to make the effects more comparable across the sites.
-{\bf A.} Mean partial correlations in $C_{\sf sparse+latent}$ depend on $\Delta ori$ more strongly than mean sample correlations.
-{\bf B.} Mean partial correlations in $C_{\sf sparse+latent}$ depend on $\Delta x$ more strongly than mean sample correlations.
-Only horizontally aligned cell pairs with $\Delta z<30\,\mu m$ were considered.
-{\bf C.} Mean partial correlations in $C_{\sf sparse+latent}$ depend on $\Delta z$ more strongly than mean sample correlations.
-Only vertically aligned cell pairs with $\Delta x<30\,\mu m$ were considered.
-{\bf D--F.} Normalized positive connectivity (green) and normalized negative connectivity (dark red) inferred by the $C_{\sf sparse+latent}$ estimator in $n=5$ imaged sites.
-Here ``connectivity'' refers to the fraction of the non-zero elements in the sparse component $S$ of the $C_{\sf sparse+latent}$ estimator, which describes inferred direct interactions between specific pairs of the recorded neurons.
-Positive and negative connectivities refer to the fractions of the positive and negative partial correlations computed from  $S$, respectively.
-``Normalized positive connectivity'' is the ratio of the positive connectivity for pairs meeting a given condition, \emph{e.g.}~similar tuning with $\Delta ori <15^{\circ}$, to the average connectivity over the entire site.
-Normalized negative connectivity is computed similarly for the negative connectivity.
-The average connectivity across sites is shown in Fig.~\ref{fig:5} B with only the five most connected sites included in the analysis.
-The normalization made the effects of tuning and distance more comparable across sites.
-{\bf D.} Positive connectivity decreases with $\Delta ori$ whereas negative connectivity does not.
-{\bf E.} Positive connectivity decays with $\Delta x$ whereas negative connectivity does not, within the examined range.
-{\bf F.} Positive connectivity decays with $\Delta z$. Negative connectivity does not decay for small values of $\Delta z$.
+\\
+{\bf A--C.} Mean sample correlations in relation to $\Delta\mbox{ori}$,  $\Delta x$ and $\Delta z$, respectively.
+For $\Delta x$ averages, only horizontally aligned cell pairs with $\Delta z<30\,\mu m$ were considered.
+Similarly, for $\Delta z$ averages, only vertically aligned cell pairs with $\Delta x<30\,\mu m$ were considered.
+\\
+{\bf D--F.} Mean partial correlations regularized by the $C_{\sf sparse+latent}$ estimator binned the same way as the sample correlations above.
+The partial correlations exhibit stronger dependence on $\Delta\mbox{ori}$, $\Delta x$, and $\Delta z$ than sample correlations.
+\\
+{\bf G--I.} Positive connectivity (green) and negative connectivity (red) inferred by the $C_{\sf sparse+latent}$ estimator. 
+Positive and negative connectivities refer to the fractions of the positive and negative partial correlations computed from the sparse component $S$ of $C_{\sf sparse+latent}$.  
+Positive connectivity decreases with $\Delta \mbox{ori}$, $\Delta x$, and $\Delta z$. 
+Negative connectivity does not decrease with $\Delta \mbox{ori}$, $\Delta x$ within the examined range, and with $\Delta z$ for small values of $\Delta z<60\,\mu m$.
 }
@@
 \begin{description}
-\item {\bf Figure S1. Performance of covariance estimators when the ground truth is the  Ising model.} 
-{\bf A.} With the ground truth having a diagonal matrix of dependencies (no pairwise dependencies),
+\item {\bf Figure S1. Performance of covariance estimators on samples drawn from Ising models.} 
+{\bf A--D} Validation losses of covariance matrix estimators relative to the estimator whose structure matches the ground truth.
+The calculation is performed identically to Fig.~\ref{fig:1} Row 6 except Ising models are used in place of Gaussian models for generating data.
 

### 2014-11-21 manuscript.tex
fixed error in equation 17
Ready for second resubmission

@@
 \DeclareMathOperator{\Tr}{tr}
-\newcommand{\loss}[1]{\mathcal L\left(#1\right)}
+\newcommand{\loss}[1]{\mathcal L(#1)}
 \newcommand{\T}{{\sf T}}
@@
 
-The evaluation of any covariance matrix estimator, $C$, is performed with respect to a \emph{loss function} $\loss{C,\Sigma}$ to quantify its discrepancy from truth, $\Sigma$.
-The loss function must attain its minimum when $C=\Sigma$.
+The evaluation of any covariance matrix estimator, $C$, is performed with respect to a \emph{loss function} $\ell(C,\Sigma)$ to quantify its discrepancy from the truth, $\Sigma$.
+The loss function is chosen to attain its minimum when $C=\Sigma$.
+Here, in the role of the loss function we adopted the Kullback-Leibler divergence between multivariate normal distributions with equal means, scaled by $\frac 1 p$ to make its values comparable across different population sizes:
+\begin{equation}\label{eq:loss}
+    \ell(C,\Sigma) = 
+    \frac 1 p D_{KL}\left(\mathcal N_\Sigma\,\Vert\,\mathcal N_C\right) = 
+    \frac 1 {2p} \left[\Tr(C^{-1}\Sigma) + \ln\det C - \ln\det\Sigma - p\right] 
+\end{equation}
+Therefore, $\ell(C,\Sigma)$ is expressed in nats/neuron per time bin.
 
-In this study, we adopted the \emph{normal loss} function defined as
-\begin{equation}\label{eq:loss}
-    \loss{C,\Sigma} = \frac 1 p\left[\ln \det C + \Tr(C^{-1}\Sigma)\right]
+We drew 30 independent samples with sample sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the loss $\ell(C,\Sigma)$ for each of the five estimators.  
+The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  
+All the regularized estimators produced better estimates (lower loss) than the sample covariance matrix.  
+However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Row 5).
+
+Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed nearly equally well to $C_{\sf diag}$ because it correctly identified zero factors and only estimated the individual variances.
+Similarly, when the number of latent units was zero (Column C), $C_{\sf sparse+latent}$ performed equally well to $C_{\sf sparse}$ because it correctly inferred zero latent units.
+With increasing sample sizes, all estimators converged to the truth (zero loss).  
+However, the discriminability between their performances only increased with sample size (data not shown).
+
+In more realistic conditions, when the ground truth is not accessible, the loss cannot be computed directly but may be estimated from data through \emph{validation}.
+In a validation procedure, a validation sample covariance matrix $C_{\sf sample}^\prime$ is computed from a testing data set that is independent from the data used for computing $C$.
+Then the \emph{validation loss} function $\loss{C,C^\prime_{\sf sample}}$ measures the discrepancy of $C$ from $C^\prime_{\sf sample}$. 
+Here, in the role of validation loss, we adopted the negative multivariate normal log likelihood of $C$ given $C^\prime_{\sf sample}$, also scaled by $\frac 1 p$ and omitting the constant term: 
+\begin{equation}\label{eq:vloss}
+    \loss{C,C^\prime_{\sf sample}} = \frac 1 {2p} \left[\Tr(C^{-1}C^\prime_{\sf sample})+\ln\det C \right]
 \end{equation}
 
-This loss function is related to the multivariate normal log-likelihood function $L(\Sigma|C_{\sf sample})$,
+Since $\loss{\cdot,\cdot}$ is additive in its second argument and $C^\prime_{\sf sample}$ is an unbiased estimate of $\Sigma$, then, for given $C$ and $\Sigma$, validation loss is an unbiased estimate of true loss up to a constant:
 \begin{equation}
-    \loss{X,Y} \equiv -\ln(2\pi) - \frac 2 {pn} L(X|Y)
+    \E{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E{C_{\sf sample}^\prime}}=\loss{C,\Sigma} = \ell(C,\Sigma) + \mbox{const}.
 \end{equation}
-Normalization by $p$ and $n$ makes the values of the loss function comparable across different population sizes and sample sizes.
+Therefore, the validation procedure allows comparing the relative values of loss of different covariance estimators.
 
-Although the two functions are similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf sample})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf sample}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.
-Minimum expected loss may be found at a point that is far removed from the point of maximum likelihood given the same data.
+Indeed, validation loss computed by 10-fold cross-validation (see Methods) accurately reproduces the relative values of the loss and the rankings of the covariance estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).
 
-The choice of the normal loss function is motivated, in part, by mathematical convenience.
-We expect that the main conclusions of our study will not change qualitatively with other well behaved loss functions, such as Stein's entropy loss or quadratic loss \cite{James:1961, Fan:2008, Ledoit:2004, Schafer:2005}.
+In the example above, the data were sampled from  multivariate normal random variables.
+In such models, partial correlations perfectly characterize the conditional dependencies between variables and the graphical models of partial correlations exactly correspond to the conditional dependencies in the data.
+To demonstrate that estimator rankings were robust to deviations from Gaussian models, as long as dependencies between variables included a strong linear component, we repeated the same cross-validated evaluation using pairwise Ising models to generate the data.
 
-We drew 30 independent samples of sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the \emph{excess loss} $\loss{C,\Sigma}-\loss{\Sigma,\Sigma}$ for each of the five estimators.
-The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.
-All the regularized estimators produced better estimates (lower excess loss) than the sample covariance matrix.
-However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Row 5).
-Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed equally well to $C_{\sf diag}$ because it correctly identified zero factors and only estimated the individual variances.
-Similarly, when the number of latent units was zero (Column C), $C_{\sf sparse+latent}$ performed equally well to $C_{\sf sparse}$ because it correctly identified zero latent units.
-
-With increasing sample sizes, all estimators converged to the truth (zero excess loss).  However,
-the discriminability between their performances increased with sample size (data not shown).
-
-With empirical data, the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly.
-Instead, the loss function may be estimated from the sample through \emph{validation}.
-In a validation procedure, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf sample}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.
-The estimate $C_{\sf sample}^\prime$ is used to compute the \emph{validation loss} $\loss{C,C_{\sf sample}^\prime}$, which approximates the loss  $\loss{C,\Sigma}$.
-
-The normal loss (Eq.~\ref{eq:loss}) is particularly suitable because it is additive in its second argument, \emph{i.e.} 
-\begin{equation*}\label{eq:additivity}
-    \loss{C,Z_1} + \loss{C,Z_2} \equiv \loss{C,Z_1+Z_2}
-\end{equation*}
-Then, with fixed $C$, validation loss is an unbiased estimate of true loss:
-\begin{equation*}
-    \E{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E{C_{\sf sample}^\prime}}=\loss{C,\Sigma}.
-\end{equation*}
-
-The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses; their validation losses are biased estimates of the loss function and validation does not exactly evaluate a discrepancy from truth.
-
-The loss function above assumes a single covariance matrix across all conditions in the dataset.
-However, neuroscientists often estimate a common correlation matrix across multiple stimulus conditions when the variances of responses are conditioned on the stimulus \cite{Vogels:1989, Ponce:2013}.
-In this case, the stimulus-dependent means and variances are estimated from the training dataset along with a common correlation matrix.
-An adaptation of the normal loss function to this case is described in Methods.
-
-Acquiring a separate testing sample is not practical.
-Instead, $K$-fold cross-validation can be used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).
-Then each subset is used as the testing sample with the other $K-1$ subsets serving as the training sample.
-The validation losses from all such `folds' are averaged to produce the \emph{cross-validation loss}.
-
-Cross-validation accurately reproduced relative values of the loss and the rankings of estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).
-
-To demonstrate that estimator rankings did not strongly depend on the gaussianity of the data-generating process, we repeated the cross-validated evaluation after replacing the Gaussian data-generating models with pairwise Ising models.
 Ising models have been used to infer functional connectivity from neuronal spike trains \cite{Hertz:2011}. 
@@
 \begin{equation}\label{eq:ising}
-    x\sim \frac 1 {Z(S,h)} \exp\left(-\frac 1 2 x^\T S x + h^\T x \right)
+    x\sim \frac 1 {Z(J,h)} \exp\left(\frac 1 2 x^\T J x + h^\T x \right)
 \end{equation}
-but the Ising model is defined on the multivariate binary domain $x\in\{-1,+1\}^p$ rather than the continuous domain of the Gaussian  model. 
-Both models are maximum-entropy models constrained by the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
-The partition function $Z(S,h)$ normalizes the distribution on the models' respective domains. 
-In the Gaussian model, the matrix $S$ is the precision matrix or the inverse of the covariance matrix; and the mean values are $\mu=-S^{-1}h$.  
-For the Ising model, $-S$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives but they do not have a simple relationship to the means and the covariance matrix.
-Both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $S$ indicate conditional independence between the corresponding pair of neurons. 
-Despite the departure from strictly linear dependencies between the variables characteristic of the Gaussian distribution, the Ising model yielded the same rankings of estimators as the Gaussian model in cross-validation (Fig.~S1).
+but the Ising model is defined on the multivariate binary domain rather than the continuous domain. 
+Both models are maximum-entropy models constrained match the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
+The partition function $Z(J,h)$ normalizes the distribution on the models' respective domains. 
+In the Gaussian model, the matrix $-J^{-1}$ is the covariance matrix; and the mean values are $\mu=J^{-1}h$.  
+For the Ising model, $J$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives although they do not have a simple relationship to the means and the covariance matrix.
+Both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $J$ indicate conditional independence between the corresponding pair of neurons.
 
-This simulation study demonstrated that, with sufficiently large sample sizes, imposing the correct type of structure on the estimate leads to greater estimation improvement than imposing other structures.
+Indeed, despite their considerable departure from strictly linear conditional dependencies, Ising models yielded the same relationships between the performances covariance estimators as the Gaussian models in cross-validation (Fig.~\ref{fig:2}).
+Identical interaction matrices $J$  of the joint distributions over the observable and latent variables were used for both the Gaussian and the Ising models.
+
+This simulation study demonstrated that cross-validated evaluation of regularized estimators of the covariance matrix of population activity can discriminate between structures of dependencies in the population.
+The identity of the most efficient covariance estimator for a particular neural circuit is therefore an empirical finding characterizing the nature of circuit  interactions.
 
 \subsection*{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
-We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:2} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
-This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:2} C and D).
-The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:2} C).
+We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:3} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
+This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:3} C and D).
+The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:3} C).
 Only cells that produced detectable calcium activity were included in the analysis (see Methods).
@@
 Each grating was displayed for 500 ms, without intervening blanks.
-This stimulus was used to compute the orientation tuning of the recorded cells (Fig.~\ref{fig:2} D).
+This stimulus was used to compute the orientation tuning of the recorded cells (Fig.~\ref{fig:3} D).
 To estimate the noise correlation matrix, we presented only 2 distinct directions in some experiments or 5 directions in others with 100--300 repetitions of each direction.
@@
 The traces were then binned into 150 ms intervals aligned on the stimulus onset for the estimation of the correlation matrix.
-The sample correlation coefficients were largely positive and low (Fig.~\ref{fig:2} E and F).
-The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean across sites of 0.018 (Fig.~\ref{fig:5} D).
+The sample correlation coefficients were largely positive and low (Fig.~\ref{fig:3} E and F).
+The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean across sites of 0.018 (Fig.~\ref{fig:6} D).
 
@@
 To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
-The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S2 and  Methods).
-Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}).
+The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods).
+Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:4}).
 The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).
@@
 \subsection*{Structure of $C_{\sf sparse+latent}$ estimates}
-We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:4} and Fig.~\ref{fig:5}).
-Although the regularized estimates were similar to the sample correlation matrix (Fig.~\ref{fig:4} A and B), the corresponding partial correlation matrices differed substantially (Fig.~\ref{fig:4} C and D).
-The estimates separated two sources of correlations: a network of linear interactions expressed by the sparse component of the inverse and latent units expressed by the low-rank components of the inverse (Fig.~\ref{fig:4} E).
-The sparse partial correlations revealed a network that differed substantially from the network composed of the greatest coefficients in the sample correlation matrix (Fig.~\ref{fig:4} F, G, H, and I).
+We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:5} and Fig.~\ref{fig:6}).
+Although the regularized estimates were similar to the sample correlation matrix (Fig.~\ref{fig:5} A and B), the corresponding partial correlation matrices differed substantially (Fig.~\ref{fig:5} C and D).
+The estimates separated two sources of correlations: a network of linear interactions expressed by the sparse component of the inverse and latent units expressed by the low-rank components of the inverse (Fig.~\ref{fig:5} E).
+The sparse partial correlations revealed a network that differed substantially from the network composed of the greatest coefficients in the sample correlation matrix (Fig.~\ref{fig:5} F, G, H, and I).
 
-In the example site (Fig.~\ref{fig:4}), the sparse component had 92.8\% sparsity (or conversely, 7.2\% connectivity: $\mbox{connectivity}=1-\mbox{sparsity}$) with average node degree of 20.9 (Fig.~\ref{fig:4} G).
+In the example site (Fig.~\ref{fig:5}), the sparse component had 92.8\% sparsity (or conversely, 7.2\% connectivity: $\mbox{connectivity}=1-\mbox{sparsity}$) with average node degree of 20.9 (Fig.~\ref{fig:5} G).
 The average node degree, \emph{i.e.}\;the average number of interactions linking each neuron, is related to connectivity as $\mbox{degree} = \mbox{connectivity}\cdot(p-1)$, where $p$ is the number of neurons.
 The low-rank component had rank 72, denoting 72 inferred latent units.
-The number of latent units increased with population size (Fig.~\ref{fig:5} A) but the connectivity was highly variable (Fig.~\ref{fig:5} B): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
+The number of latent units increased with population size (Fig.~\ref{fig:6} A) but the connectivity was highly variable (Fig.~\ref{fig:6} B): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions.
 This variability may be explained by differences in brain states and recording quality and warrants further investigation.
 
-The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 27 sites were about 5 times lower than the average sample correlations (Fig.~\ref{fig:5} C).
+The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 27 sites were about 5 times lower than the average sample correlations (Fig.~\ref{fig:6} C).
 This suggests that correlations between neurons build up from multiple chains of smaller interactions.
@@
 
-While the sample correlations were mostly positive, the sparse component of the partial correlations (`interactions') had a high fraction (28.7\% in the example site) of negative values (Fig.~\ref{fig:4} F).
-The fraction of negative interactions increased with the inferred connectivity (Fig.~\ref{fig:5} D), suggesting that negative interactions can be inferred only after a sufficient density of positive interactions has been uncovered.
+While the sample correlations were mostly positive, the sparse component of the partial correlations (`interactions') had a high fraction (28.7\% in the example site) of negative values (Fig.~\ref{fig:5} F).
+The fraction of negative interactions increased with the inferred connectivity (Fig.~\ref{fig:6} D), suggesting that negative interactions can be inferred only after a sufficient density of positive interactions has been uncovered.
 
@@
 The networks revealed by the two methods differed substantially.
-In the example site with 7.2\% connectivity in $C_{\sf sparse+latent}$, only 27.7\% of the connections coincided with the above-threshold sample correlations (Fig.~\ref{fig:4} F, H, and I).
-In particular, most of the inferred negative interactions corresponded to low sample correlations (Fig.~\ref{fig:4} F) where high correlations should be expected given the rest of the correlation matrix.
+In the example site with 7.2\% connectivity in $C_{\sf sparse+latent}$, only 27.7\% of the connections coincided with the above-threshold sample correlations (Fig.~\ref{fig:5} F, H, and I).
+In particular, most of the inferred negative interactions corresponded to low sample correlations (Fig.~\ref{fig:5} F) where high correlations should be expected given the rest of the correlation matrix.
 
@@
 
-We examined how the structure of the $C_{\sf sparse+latent}$ estimates related to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:6}).
+We examined how the structure of the $C_{\sf sparse+latent}$ estimates related to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:7}).
 Five sites with highest pairwise connectivities were included in the analysis.
@@
 The partial correlations decayed more rapidly with $\Delta\mbox{ori}$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test of the difference of the linear regression coefficients).
-Positive connectivity decreased with $\Delta\mbox{ori}$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression coefficient) whereas negative connectivity did not decrease (Fig.~\ref{fig:6} D): The slope in the logistic model of connectivity with respect to $\Delta\mbox{ori}$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficient).
+Positive connectivity decreased with $\Delta\mbox{ori}$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression coefficient) whereas negative connectivity did not decrease (Fig.~\ref{fig:7} D): The slope in the logistic model of connectivity with respect to $\Delta\mbox{ori}$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficient).
 
@@
 Again, the partial correlations decayed more rapidly both laterally and vertically than sample correlations ($p<10^{-6}$ in each of the five sites, for both lateral and vertical distances, two-sample $t$-test of the difference of the linear regression coefficients).
-Positive connectivity decayed with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression coefficient) (Fig.~\ref{fig:6} E), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$.
+Positive connectivity decayed with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression coefficient) (Fig.~\ref{fig:7} E), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$.
 Although the positive connectivity appeared to decay faster with vertical than with lateral distance, the differences in slopes of the respective logistic regression models were not significant with available data.
-The negative connectivity decayed slower with distance (Fig.~\ref{fig:6} E and F): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficients).
+The negative connectivity decayed slower with distance (Fig.~\ref{fig:7} E and F): The slope in the respective logistic models with respect to the lateral distance was significantly higher for positive than for negative connectivities ($p<0.05$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficients).
 
@@
 Importantly, the inferred functional interactions were substantially different from the network of the highest sample correlations.
-For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (Fig.~\ref{fig:4} F) and may reflect inhibitory circuitry.
+For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (Fig.~\ref{fig:5} F) and may reflect inhibitory circuitry.
 
-Distances between cells in physical space and in sensory feature space had a stronger effect on the partial correlations estimated by the $C_{\sf sparse+latent}$ estimator than on sample correlations (Fig.~\ref{fig:6} A--C).
+Distances between cells in physical space and in sensory feature space had a stronger effect on the partial correlations estimated by the $C_{\sf sparse+latent}$ estimator than on sample correlations (Fig.~\ref{fig:7} A--C).
 These differences support the idea that correlations are built up from partial correlations in chains of intermediate cells positioned closer and tuned more similarly to one another, with potentially closer correspondence to anatomical connectivity.
 These differences may also be at least partially explained by a trivial effect of regularization: the $L_1$ penalty applied by the estimator (Eq.~\ref{eq:ma}) suppresses small partial correlations to greater extent than large partial correlations, enhancing the apparent effect of distance and tuning.
-Distances between cells in physical space and in sensory feature space also had distinct effects on the positive and negative connectivities inferred by the estimator (Fig.~\ref{fig:6} D--F); we could not think of a numerical artifact that could produce this effect.
+Distances between cells in physical space and in sensory feature space also had distinct effects on the positive and negative connectivities inferred by the estimator (Fig.~\ref{fig:7} D--F); we could not think of a numerical artifact that could produce this effect.
 The distinct positive and negative connectivity patterns  may reflect geometric and graphical features of local excitatory and inhibitory networks.
@@
 The measured fluorescent traces were deconvolved to reconstruct the firing rates for each neuron: First, the first principal component was subtracted from the raw traces in order to reduce common mode noise related to small cardiovascular movements \cite{Cotton:2013}.
-The resulting traces were high-pass filtered above 0.1 Hz and downsampled to 20 Hz (Fig.~\ref{fig:2} C).
+The resulting traces were high-pass filtered above 0.1 Hz and downsampled to 20 Hz (Fig.~\ref{fig:3} C).
 Then, the firing rates were estimated using by nonnegative deconvolution \cite{Vogelstein:2010}.
@@
 Thus the size of the training dataset within the inner loop comprised 81\% of the entire recording.
-Fig.~S2 illustrates the dependence of the validation loss on the hyperparameters of the $C_{\sf sparse+latent}$ estimator for the example site shown in Figures \ref{fig:2} and \ref{fig:4} and the optimal value found by the pattern search algorithm.
+Fig.~S1 illustrates the dependence of the validation loss on the hyperparameters of the $C_{\sf sparse+latent}$ estimator for the example site shown in Figures \ref{fig:3} and \ref{fig:5} and the optimal value found by the pattern search algorithm.
 
 When the validation loss was not required, only the inner loop of cross-validation was used on the entire dataset.
-This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (Fig.~\ref{fig:1} Rows 4 and 5) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:4}--\ref{fig:6}).
+This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (Fig.~\ref{fig:1} Rows 4 and 5) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:5}--\ref{fig:7}).
 
@@
 
-The partial correlation matrix (Eq.~\ref{eq:partial}) computed from $C_{\sf sparse+latent}$ includes interactions between the visible and latent units and was used in Fig.~\ref{fig:4} C and D and Fig.~\ref{fig:5} C, and Fig.~\ref{fig:6} A--C).
+The partial correlation matrix (Eq.~\ref{eq:partial}) computed from $C_{\sf sparse+latent}$ includes interactions between the visible and latent units and was used in Fig.~\ref{fig:5} C and D and Fig.~\ref{fig:6} C, and Fig.~\ref{fig:7} A--C).
 The partial correlation matrix computed from $S$ alone expresses strengths of pairwise interactions
@@
 \end{equation}
-and were used in Fig.~\ref{fig:4} F, G, H.
+and were used in Fig.~\ref{fig:5} F, G, H.
 
@@
 \subsection*{Cross-validation with conditioned variances}
-This study required an unbiased estimate of the discrepancy between various correlation matrix estimates and the true value of the correlation matrix.
-This was accomplished by cross-validation whereby regularized covariance matrix estimates computed from training samples were compared to unbiased covariance matrix estimates computed from independent testing samples.
-Loss functions were defined with respect to the covariance matrix rather than the correlation matrix because no unbiased estimate of the correlation matrix exists for finite sample sizes in the general case \cite{Fisher:1921}.
-In particular, we showed that with the normal loss function (Eq.~\ref{eq:loss}), the validation loss $\loss{C,C_{\sf sample}^\prime}$ is an unbiased estimate of true loss when $C_{\sf sample}^\prime$ is an unbiased estimate.
-However, this approach only worked under the assumption of uniform variances across all conditions included in the training and testing samples.
-Here, we extend the validation loss to the estimation and evaluation of a common correlation matrix across multiple conditions with different variances.
+Special attention was given to estimating the variances.  
+All evaluations and optimization in this study were defined with respect to the covariance matrices.  
+However, neuroscientists often estimate a common correlation matrix across multiple stimulus conditions when the variances of responses are conditioned on the stimulus \cite{Vogels:1989, Ponce:2013}.
+In this study, we too conditioned the variances on the stimulus but estimated a single correlation matrix across all conditions.
+Here we describe the computation of the validation loss (Eq.~\ref{eq:vloss}) when the variances were allowed to vary with the stimulus condition.
 
@@
 
-The overall loss function is obtained by averaging the loss functions across all the conditions:
+The overall validation loss is obtained by averaging the validation losses across all conditions:
 \begin{equation}\label{eq:full-loss}
@@
 
-With the normal loss $\loss{\cdot,\cdot}$ (Eq.~\ref{eq:loss}) and the unbiased testing covariance matrix $C_{c,\sf sample}$, the loss function in Eq.~\ref{eq:full-loss} is an unbiased estimate of the true loss.
-Hence, it was used for evaluations reported in Fig.~\ref{fig:3}.
+With negative normal log-likelihood as the validation loss  (Eq.~\ref{eq:vloss}) and the unbiased validation covariance matrix $C_{c,\sf sample}$, the loss function in Eq.~\ref{eq:full-loss} is an unbiased estimate of the true loss.
+Hence, it was used for evaluations reported in Fig.~\ref{fig:4}.
 
@@
 Recorded neurons are represented by the green spheres and latent units by the lightly shaded spheres.
-Edges represent non-zero partial correlations, \emph{i.e.}\;`interactions'.
+Edges represent conditional dependencies, \emph{i.e.}\;`interactions'.
      {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear dependencies.
@@
      \\
-{\bf Row 5:} Excess losses (Eq.~\ref{eq:loss}) for the five estimators as a function of sample size.
-The error bars indicate the standard error of the mean based on 30 samples.
+{\bf Row 5:} True loss (Eq.~\ref{eq:loss}) for the five estimators as a function of sample size.
+The error bars indicate the standard deviation of the mean.
 Estimators with structure that matches the true model converged to zero faster than the other estimators.
      \\
-{\bf Row 6:} Validation losses for the five estimators relative to the matching estimator.
-Error bars indicate the standard error of the mean based on 30 samples.
+{\bf Row 6:} Validation loss (Eq.~\ref{eq:vloss}) for the five estimators relative to the matching estimators for each type of ground truth.
+Error bars indicate the standard deviation of the mean.
 Differences in validation loss approximate differences in true loss.
@@
 \label{fig:1}
+\end{figure}
+
+\begin{figure} 
+\caption{
+{\bf Performance of covariance estimators on samples drawn from Ising models.} 
+{\bf A--D} Validation losses of covariance matrix estimators relative to the estimator whose structure matches the ground truth.
+The calculation is performed identically to Fig.~\ref{fig:1} Row 6 except Ising models are used as ground truth.
+}
+\label{fig:2}
 \end{figure}
@@
 {\bf F.} Histogram of noise correlation coefficients in one site. The red line indicates the mean correlation coefficient of 0.020.
-} \label{fig:2}
+} \label{fig:3}
 \end{figure}
@@
 \caption{
-{\bf Performance of estimator $C_{\sf sparse+latent}$ with respect to the normal loss function (eq.~\ref{eq:loss}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.}
+{\bf Performance of estimator $C_{\sf sparse+latent}$ expressed as validation loss (eq.~\ref{eq:vloss}) relative to the other estimators: $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$.}
 Covariance estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ produced consistently greater validation losses than $C_{\sf sparse+latent}$ ($p<0.01$ in each comparison, Wilcoxon signed rank test, $n=27$ sites in 14 mice).
 The box plots indicate the $25^{th}$, $50^{th}$, and $75^{th}$ percentiles with the whiskers extending to the minimum and maximum values after excluding the outliers marked with `+'.
-} \label{fig:3}
+} \label{fig:4}
 \end{figure}
@@
 }
-\label{fig:4}
+\label{fig:5}
 \end{figure}
@@
 \caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites.}
-Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:2} and \ref{fig:4} is circled in blue.
+Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:3} and \ref{fig:5} is circled in blue.
 \\
@@
 }
-\label{fig:5}
+\label{fig:6}
 \end{figure}
@@
 {\bf Dependence of sample correlations, regularized partial correlations, and connectivity inferred by $C_{\sf sparse+latent}$ on the differences in preferred orientations, $\Delta \mbox{ori}$, and physical distances: horizontal $\Delta x$ and depth $\Delta z$.}
-Five sites with highest connectivity (see Fig.~\ref{fig:5} B) were selected for this analysis.
+Five sites with highest connectivity (see Fig.~\ref{fig:6} B) were selected for this analysis.
 \\
@@
 }
-\label{fig:6}
+\label{fig:7}
 \end{figure}
@@
 \begin{description}
-\item {\bf Figure S1. Performance of covariance estimators on samples drawn from Ising models.} 
-{\bf A--D} Validation losses of covariance matrix estimators relative to the estimator whose structure matches the ground truth.
-The calculation is performed identically to Fig.~\ref{fig:1} Row 6 except Ising models are used in place of Gaussian models for generating data.
-
-\item {\bf Figure S2.
-Optimization of hyperparameters of the $C_{\sf sparse+latent}$ estimator.} {\bf A.} Validation loss (Eq.~\ref{eq:full-loss}) for the example site in Fig.~\ref{fig:2} and \ref{fig:4} as a function of the hyperparameters $\alpha$ and $\beta$ of the $C_{\sf sparse+latent}$ estimator (Eq.~\ref{eq:c-sl} and Eq.~\ref{eq:ma}).
+\item {\bf Figure S1.
+Optimization of hyperparameters of the $C_{\sf sparse+latent}$ estimator.} {\bf A.} Validation loss (Eq.~\ref{eq:full-loss}) for the example site in Fig.~\ref{fig:3} and \ref{fig:5} as a function of the hyperparameters $\alpha$ and $\beta$ of the $C_{\sf sparse+latent}$ estimator (Eq.~\ref{eq:c-sl} and Eq.~\ref{eq:ma}).
 In all panels, the red cross marks the optimal value found by the pattern search algorithm described in Methods.

### 2014-11-21 manuscript.tex
minor

@@
 
-Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed nearly equally well to $C_{\sf diag}$ because it correctly identified zero factors and only estimated the individual variances.
-Similarly, when the number of latent units was zero (Column C), $C_{\sf sparse+latent}$ performed equally well to $C_{\sf sparse}$ because it correctly inferred zero latent units.
+Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed equally well to $C_{\sf diag}$ because it correctly inferred zero factors and only estimated the individual variances.
+Similarly, when the number of latent units was zero (Column C), $C_{\sf sparse+latent}$ performed nearly equally well to $C_{\sf sparse}$ because it correctly inferred zero latent units.
 With increasing sample sizes, all estimators converged to the truth (zero loss).  
@@
 In a validation procedure, a validation sample covariance matrix $C_{\sf sample}^\prime$ is computed from a testing data set that is independent from the data used for computing $C$.
-Then the \emph{validation loss} function $\loss{C,C^\prime_{\sf sample}}$ measures the discrepancy of $C$ from $C^\prime_{\sf sample}$. 
+Then the \emph{validation loss} $\loss{C,C^\prime_{\sf sample}}$ measures the discrepancy of $C$ from $C^\prime_{\sf sample}$. 
 Here, in the role of validation loss, we adopted the negative multivariate normal log likelihood of $C$ given $C^\prime_{\sf sample}$, also scaled by $\frac 1 p$ and omitting the constant term: 
@@
 
-Since $\loss{\cdot,\cdot}$ is additive in its second argument and $C^\prime_{\sf sample}$ is an unbiased estimate of $\Sigma$, then, for given $C$ and $\Sigma$, validation loss is an unbiased estimate of true loss up to a constant:
+Since $\loss{\cdot,\cdot}$ is additive in its second argument and $C^\prime_{\sf sample}$ is an unbiased estimate of $\Sigma$, then, for given $C$ and $\Sigma$, the validation loss is an unbiased estimate of the true loss, up to a constant:
 \begin{equation}
@@
 
-Indeed, validation loss computed by 10-fold cross-validation (see Methods) accurately reproduces the relative values of the loss and the rankings of the covariance estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).
+Indeed, the validation loss computed by 10-fold cross-validation (see Methods) accurately reproduces the relative values of the true loss and the rankings of the covariance estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).
 
@@
 In such models, partial correlations perfectly characterize the conditional dependencies between variables and the graphical models of partial correlations exactly correspond to the conditional dependencies in the data.
-To demonstrate that estimator rankings were robust to deviations from Gaussian models, as long as dependencies between variables included a strong linear component, we repeated the same cross-validated evaluation using pairwise Ising models to generate the data.
+To demonstrate that estimator rankings were robust to deviations from Gaussian models, we repeated the same cross-validated evaluation using pairwise Ising models to generate the data.
 
@@
 
-To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
+To test this hypothesis, we computed the relative validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
 The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods).
 Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:4}).
-The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).
+The respective median differences of the validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).
 
@@
 
-We evaluated the covariance matrix estimators a loss function based on multivariate normal log likelihood.
+We evaluated the covariance matrix estimators using a loss function derived from the normal distribution.
 However, this does not limit the applicability of its conclusions to normal distributions.
@@
 When the validation loss was not required, only the inner loop of cross-validation was used on the entire dataset.
-This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (Fig.~\ref{fig:1} Rows 4 and 5) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:5}--\ref{fig:7}).
+This approach was used to compute the covariance matrix estimates and their true loss in the simulation study (Fig.~\ref{fig:1} Rows 4 and 5) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:5}--\ref{fig:7}).
 

### 2014-11-24 manuscript.tex
Alex's changes.

@@
 In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of synaptic connectivity \cite{Gerstein:1964}.
-Interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various aspects of circuit organization such as the physical distances between the neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010}.
+Interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various aspects of circuit organization such as the physical distances between the neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2014, Renart:2010}.
 
@@
 Improved estimation of the correlation matrix is beneficial in itself.
-For example, improved estimates can be used to optimize  decoding of the population activity \cite{Friedman:1989}.
+For example, improved estimates can be used to optimize  decoding of the population activity \cite{Friedman:1989, Berens:2012}.
 But reduced estimation error is not the only benefit of regularization.
@@
 
-Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an independent unbiased estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix},
+Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an unbiased estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix} is
 \begin{equation}\label{eq:sample}
@@
     \end{equation}
-where $n$ is the number of time bins in $T$, yields an unbiased estimate of the covariance matrix so that $\E{C_{\sf sample}}=\Sigma$.
-
+where $n$ is the number of time bins in $T$.
+The sample covariance matrix yields an unbiased estimate of the covariance matrix ($\E{C_{\sf sample}}=\Sigma$) if the mean, $\bar x$, is independent of the data sample, $\{x(t): t\in T$\}.
 When the mean is estimated from the same sample, $C_{\sf sample}$ becomes biased toward zero.
-However, in all the cases when the unbiasedness is required in our study, the mean will be estimated from a separate sample.
+However, in all the cases where unbiasedness is required, we will estimate the mean from a separate sample.
 
@@
 but the Ising model is defined on the multivariate binary domain rather than the continuous domain. 
-Both models are maximum-entropy models constrained match the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
+Both models are maximum-entropy models constrained to match the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
 The partition function $Z(J,h)$ normalizes the distribution on the models' respective domains. 
 In the Gaussian model, the matrix $-J^{-1}$ is the covariance matrix; and the mean values are $\mu=J^{-1}h$.  
-For the Ising model, $J$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives although they do not have a simple relationship to the means and the covariance matrix.
+For the Ising model, $J$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives, although they do not have a simple relationship to the means and the covariance matrix.
 Both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $J$ indicate conditional independence between the corresponding pair of neurons.
 
-Indeed, despite their considerable departure from strictly linear conditional dependencies, Ising models yielded the same relationships between the performances covariance estimators as the Gaussian models in cross-validation (Fig.~\ref{fig:2}).
+Indeed, despite their considerable departure from strictly linear conditional dependencies, Ising models yielded the same relationships between the performances of the covariance estimators as the Gaussian models in cross-validation (Fig.~\ref{fig:2}).
 Identical interaction matrices $J$  of the joint distributions over the observable and latent variables were used for both the Gaussian and the Ising models.
@@
 \subsection*{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
-We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:3} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
+We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of lightly sedated mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:3} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
 This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:3} C and D).
@@
 This stimulus was used to compute the orientation tuning of the recorded cells (Fig.~\ref{fig:3} D).
-To estimate the noise correlation matrix, we presented only 2 distinct directions in some experiments or 5 directions in others with 100--300 repetitions of each direction.
+To estimate the noise correlation matrix, we presented only two distinct directions in some experiments or five directions in others with 100--300 repetitions of each direction.
 Each grating lasted 1 second and was followed by a 1-second blank.
@@
 
-Inference of the conditional dependencies also depends on the completeness of the recorded population:  To equate conditional dependencies to direct interaction between neurons, we must record from all neurons with which the pair interacts.
+Inference of the conditional dependencies also depends on the completeness of the recorded population:  To equate conditional dependency to direct interaction between two neurons, we must record from all neurons with which the pair interacts.
 Unobserved portions of the circuit may manifest as conditional dependencies between observed neurons that do not interact.
@@
 However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.
-Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.
+Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013, Theis:2014}.
 The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.
@@
 Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \emph{i.e.}\;groups of tightly interconnected groups of cells involved in a common computation.
-Coactivation patterns could also result from shared input from unobserved parts of the circuit, or global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.
+Coactivation patterns could also result from shared input from unobserved parts of the circuit, or global network fluctuations modulating the activity of the local circuit \cite{Okun:2012, Ecker:2014}.
 
@@
 \subsection*{Combining pairwise interactions and coactivations}
-In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables \cite{Koster:2013},  high-order correlations \cite{Ganmor:2011}, or global network fluctuations \cite{Tkacik:2013}.
+In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables, higher-order correlations, or global network fluctuations \cite{Ganmor:2011, Tkacik:2013, Pfau:2013, Koster:2014, Ecker:2014}.
 
@@
 We showed that among several models a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix for cortical microcircuits.
-This finding is valuable in itself: improved estimates of the noise covariance matrix for large datasets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Ecker:2011}
+This finding is valuable in itself: improved estimates of the noise covariance matrix for large datasets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Ecker:2011}
 
@@
 The surgical procedures and data acquisition were performed as described in \cite{Cotton:2013}: C57BL/6J mice (aged p40--60) were used.
-Anesthesia was initiated with isoflurane (3\%) and maintained with the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
+For surgery, animals were initially anesthetized with isoflurane (3\%).
+During the experiments, animals were sedated with a mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
 A craniotomy was performed over the right primary visual cortex.
@@
 Two types of stimuli were presented for each imaging site: First, directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen directions of motion, 500 ms per direction, without blanks, with 12--30 trials for each direction of motion.
-Second, ro measure correlations, the stimulus was modified to include only two directions of motion (in 9 datasets) or five directions (in 22 datasets) and the gratings were presented for 1 second and were separated by 1-second blanks, with 100--300 trials for each direction of motion.
+Second, to measure correlations, the stimulus was modified to include only two directions of motion (in 9 datasets) or five directions (in 22 datasets) and the gratings were presented for 1 second and were separated by 1-second blanks, with 100--300 trials for each direction of motion.
 
@@
 \end{equation}
-by an expectation-maximization (EM) algorithm with specified rank of $L$. Additional the diagonal matrix of individual variances is shrunk toward its mean value according to Eq.~\ref{eq:c-factor}.
+by an expectation-maximization (EM) algorithm with specified rank of $L$. In addition, the diagonal matrix of individual variances is shrunk toward its mean value according to Eq.~\ref{eq:c-factor}.
 

### 2014-11-28 manuscript.tex
corrections based on co-authors' inputs

@@
 where $n$ is the number of time bins in $T$. 
-The sample covariance matrix yields an unbiased estimate of the covariance matrix ($\E{C_{\sf sample}}=\Sigma$) if the mean, $\bar x$, is independent of the data sample, $\{x(t): t\in T$\}.
-When the mean is estimated from the same sample, $C_{\sf sample}$ becomes biased toward zero.
-However, in all the cases where unbiasedness is required, we will estimate the mean from a separate sample.
+The sample covariance matrix yields an unbiased estimate of the true covariance matrix ($\E{C_{\sf sample}}=\Sigma$) if the mean activity, $\bar x$, is estimated independently, in a separate data sample.
+In all cases when the unbiasedness of the sample covariance matrix matters in this paper, the means are estimated independently.
 
@@
 
-As the size of recorded neuronal populations increases, without substantial increases in recording durations, the \emph{condition number} of the sample covariance matrix also increases, making the partial correlation estimates \emph{ill-conditioned}: small errors in covariance estimates translate into much greater errors in partial correlations after the inversion.
+Increasing the number of recorded neurons results in a higher \emph{condition number} of the covariance matrix \cite{Ledoit:2004} making the partial correlation estimates \emph{ill-conditioned}: small errors in the covariance estimates translate into much greater errors in the estimates of partial correlations.
 With massively multineuronal recordings, partial correlations cannot be estimated without \emph{regularization} \cite{Ledoit:2004,Schafer:2005}.
@@
 Similarly, when the number of latent units was zero (Column C), $C_{\sf sparse+latent}$ performed nearly equally well to $C_{\sf sparse}$ because it correctly inferred zero latent units.
-With increasing sample sizes, all estimators converged to the truth (zero loss).  
-However, the discriminability between their performances only increased with sample size (data not shown).
+With increasing sample sizes, all estimators converged to the ground truth (zero loss) but the estimators with correct structure outperformed the others even for large samples.
 
@@
 
-Indeed, the validation loss computed by 10-fold cross-validation (see Methods) accurately reproduces the relative values of the true loss and the rankings of the covariance estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).
+Indeed, the validation loss computed by 10-fold cross-validation (see Methods) accurately reproduced the relative values of the true loss and the rankings of the covariance estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).
 
@@
 \subsection*{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
-We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of lightly sedated mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:3} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
+We recorded the calcium activity of densely sampled populations of neurons in layers 2/3 and upper layer 4 in primary visual cortex of sedated mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:3} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
 This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:3} C and D).
@@
 This suggests that correlations between neurons build up from multiple chains of smaller interactions.
-Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29, with larger populations exhibiting greater uniformity of average partial correlations than the smaller populations.
+Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29, with larger populations exhibiting greater uniformity of average partial correlations than the smaller populations ($p=0.002$ Brown-Forsythe test).
 
@@
 \subsection*{Combining pairwise interactions and coactivations}
-In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables, higher-order correlations, or global network fluctuations \cite{Ganmor:2011, Tkacik:2013, Pfau:2013, Koster:2014, Ecker:2014}.
+In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables, higher-order correlations, or global network fluctuations \cite{Ganmor:2011, Tkacik:2013, Pfau:2013, Koster:2013, Ecker:2014}.
 
@@
 \subsection*{Addressing ill-posedness}
-Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.
+Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data can produce very large errors in the inferred network of dependencies.
 The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.
@@
 However, ambitious projects, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
-Simply increasing recording duration will not be sufficient, and the problem must be addressed by using regularized estimators.
+Simply increasing recording duration will not be practical or sufficient, and the problem must be addressed by using regularized estimators.
 Regularization biases the solution toward a small subspace in order to counteract the effect of  sampling noise in the empirical data.
@@
 These differences may also be at least partially explained by a trivial effect of regularization: the $L_1$ penalty applied by the estimator (Eq.~\ref{eq:ma}) suppresses small partial correlations to greater extent than large partial correlations, enhancing the apparent effect of distance and tuning.
-Distances between cells in physical space and in sensory feature space also had distinct effects on the positive and negative connectivities inferred by the estimator (Fig.~\ref{fig:7} D--F); we could not think of a numerical artifact that could produce this effect.
-The distinct positive and negative connectivity patterns  may reflect geometric and graphical features of local excitatory and inhibitory networks.
-Indeed, the relationships between patterns of positive and negative connectivities inferred by the estimator mirrored the properties of excitatory and inhibitory synaptic connectivities with respect to distance, cortical layers, and feature tuning \cite{Song:2005, Oswald:2008, Adesnik:2010, Perin:2011, Fino:2011, Hofer:2011, Isaacson:2011, Levy:2012}.
+Still, the distinct positive and negative connectivity patterns (Fig.\ \ref{fig:7} D--F)  may reflect geometric and graphical features of local excitatory and inhibitory networks. 
+Indeed, the relationships between patterns of positive and negative connectivities inferred by the estimator resembled the properties of excitatory and inhibitory synaptic connectivities with respect to distance, cortical layers, and feature tuning \cite{Song:2005, Oswald:2008, Adesnik:2010, Perin:2011, Fino:2011, Hofer:2011, Isaacson:2011, Levy:2012}.
 For example, while excitatory neurons form synapses within highly specific local cliques \cite{Perin:2011}, inhibitory interneurons form synapses with nearly all excitatory cells within local microcircuits \cite{Fino:2011, Hofer:2011, Packer:2011}.
@@
 slow-wave sleep, attention) and stimulus conditions (e.g.
-certain types of stimuli may engage feedback connections).
+certain types of stimuli may engage feedback connections) \cite{Reimer:2014,Fu:2014}.
 
@@
 \begin{equation}\label{eq:ma}
-(S,L) = \argmin\limits_{\hat S,\hat L} \loss{\hat S-\hat L, C_{\sf sample}} + \alpha\|\hat S\|_1 + \beta\Tr(\hat L)
+	(S,L) = \argmin\limits_{\hat S,\hat L} \loss{(\hat S-\hat L)^{-1}, C_{\sf sample}} + \alpha\|\hat S\|_1 + \beta\Tr(\hat L)
 \end{equation}

### 2014-12-07 manuscript.tex

--- 2014-11-28 manuscript.tex
@@
 We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.
-Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.
+Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations arising from linear interactions between pairs of neurons are well approximated by their pairwise \emph{partial} correlations.
 In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (`interactions') combined with a low-rank component.
@@
 \begin{description}
-\item[$C_{\sf sample}$] -- sample covariance matrix, the unbiased, unregularized estimator.
+\item[$C_{\sf sample}$] -- the sample covariance matrix, the unbiased estimator.
 \item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \emph{i.e.}\;toward a diagonal covariance matrix.
@@
 We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.
-This estimator revealed a sparse network of partial correlations (`{interactions'), between the observed neurons; it also inferred latent units exerting linear effects on the observed neurons.
+This estimator revealed a sparse network of partial correlations (`interactions'), between the observed neurons; it also inferred latent units exerting linear effects on the observed neurons.
 We analyzed these networks of partial correlations and found the following: Whereas significant noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.
-Moreover, we found that these interactions exhibited a stronger relationship to the physical distances and to the differences in preferred orientations than noise correlations.
+Moreover, the average partial correlations and the inferred positive interactions in particular exhibited a stronger relationship to the physical distances and to the differences in preferred orientations than noise correlations.
 In contrast, the inferred negative interactions were less selective.
@@
 
-Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an unbiased estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix} is
+Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix} is
 \begin{equation}\label{eq:sample}
@@
 The \emph{partial correlation} between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}.
-Partial correlations figure prominently in probabilistic \emph{graphical modeling} wherein the joint distribution is explained by sets of two-way interactions \cite{Whittaker:1990}.
+Partial correlations figure prominently in probabilistic \emph{graphical modeling} wherein the joint distribution is explained by sets of pairwise interactions \cite{Whittaker:1990}.
 For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}.
-More generally, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1990,Baba:2004}.
+More generally, partial correlations can serve as a measure of conditional independence under the assumption that dependencies in the system are close to linear effects \cite{Whittaker:1990,Baba:2004}.
 As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functional connectivity) between pairs of neurons.
@@
 Pairwise partial correlations are closely related to the elements of the \emph{precision matrix}, \emph{i.e.}\;the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1990}.
-Zero elements in the precision matrix signify zero partial correlation between the two variables.
+Zero elements in the precision matrix signify zero partial correlations between the corresponding pairs of variables.
 Given the covariance estimate $C$, the matrix of partial correlations $P$ is computed by normalizing the rows and columns of the \emph{precision matrix} $C^{-1}$ to produce negative unit entries on the diagonal:
@@
 
-Increasing the number of recorded neurons results in a higher \emph{condition number} of the covariance matrix \cite{Ledoit:2004} making the partial correlation estimates \emph{ill-conditioned}: small errors in the covariance estimates translate into much greater errors in the estimates of partial correlations.
+Increasing the number of recorded neurons results in a higher \emph{condition number} of the sample covariance matrix \cite{Ledoit:2004} making the partial correlation estimates more \emph{ill-conditioned}: small errors in the covariance estimates translate into greater errors in the estimates of the partial correlations.
 With massively multineuronal recordings, partial correlations cannot be estimated without \emph{regularization} \cite{Ledoit:2004,Schafer:2005}.
@@
 The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons' variances.
-Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf sample}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
+Regularization is achieved by linear \emph{shrinkage} of the sample covariance matrix $C_{\sf sample}$ toward $D$ as controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
 \begin{equation}\label{eq:c-diag}
@@
 The problem of finding the optimal set of non-zero elements of the precision matrix is known as \emph{covariance selection} \cite{Dempster:1972}. 
-The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (`interactions') between some pairs of neurons (Fig.~\ref{fig:1}} Row 1, C).
+The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (`interactions') between some pairs of neurons (Fig.~\ref{fig:1} Row 1, C).
 
@@
 \subsection*{Simulation}
-We next demonstrate how the most efficient among different regularized estimators can reveal the structure of correlations.
+We next demonstrated how the most efficient among different regularized estimators can reveal the structure of correlations.
 We constructed four families of $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (Fig.~\ref{fig:1} Row 2, A--D and Methods).
@@
 The loss function is chosen to attain its minimum when $C=\Sigma$.
-Here, in the role of the loss function we adopted the Kullback-Leibler divergence between multivariate normal distributions with equal means, scaled by $\frac 1 p$ to make its values comparable across different population sizes:
+Here, in the role of the loss function we adopted the Kullback-Leibler divergence between multivariate normal distributions with equal means, scaled by $\frac 2 p$ to make its values comparable across different population sizes:
 \begin{equation}\label{eq:loss}
     \ell(C,\Sigma) = 
-    \frac 1 p D_{KL}\left(\mathcal N_\Sigma\,\Vert\,\mathcal N_C\right) = 
-    \frac 1 {2p} \left[\Tr(C^{-1}\Sigma) + \ln\det C - \ln\det\Sigma - p\right] 
+    \frac 2 p D_{KL}\left(\mathcal N_\Sigma\,\Vert\,\mathcal N_C\right) = 
+    \frac 1 {p} \left[\Tr(C^{-1}\Sigma) + \ln\det C - \ln\det\Sigma - p\right] 
 \end{equation}
-Therefore, $\ell(C,\Sigma)$ is expressed in nats/neuron per time bin.
+Thus $\ell(C,\Sigma)$ is expressed in nats/neuron per time bin.
 
-We drew 30 independent samples with sample sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the loss $\ell(C,\Sigma)$ for each of the five estimators.  
-The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  
-All the regularized estimators produced better estimates (lower loss) than the sample covariance matrix.  
-However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Row 5).
-
-Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed equally well to $C_{\sf diag}$ because it correctly inferred zero factors and only estimated the individual variances.
-Similarly, when the number of latent units was zero (Column C), $C_{\sf sparse+latent}$ performed nearly equally well to $C_{\sf sparse}$ because it correctly inferred zero latent units.
-With increasing sample sizes, all estimators converged to the ground truth (zero loss) but the estimators with correct structure outperformed the others even for large samples.
-
-In more realistic conditions, when the ground truth is not accessible, the loss cannot be computed directly but may be estimated from data through \emph{validation}.
+When the ground truth is not accessible, the loss cannot be computed directly but may be estimated from data through \emph{validation}.
 In a validation procedure, a validation sample covariance matrix $C_{\sf sample}^\prime$ is computed from a testing data set that is independent from the data used for computing $C$.
 Then the \emph{validation loss} $\loss{C,C^\prime_{\sf sample}}$ measures the discrepancy of $C$ from $C^\prime_{\sf sample}$. 
-Here, in the role of validation loss, we adopted the negative multivariate normal log likelihood of $C$ given $C^\prime_{\sf sample}$, also scaled by $\frac 1 p$ and omitting the constant term: 
+Here, in the role of validation loss, we adopted the negative multivariate normal log likelihood of $C$ given $C^\prime_{\sf sample}$, also scaled by $\frac 2 p$ and omitting the constant term: 
 \begin{equation}\label{eq:vloss}
-    \loss{C,C^\prime_{\sf sample}} = \frac 1 {2p} \left[\Tr(C^{-1}C^\prime_{\sf sample})+\ln\det C \right]
+    \loss{C,C^\prime_{\sf sample}} = \frac 1 p \left[\Tr(C^{-1}C^\prime_{\sf sample})+\ln\det C \right]
 \end{equation}
@@
 \end{equation}
-Therefore, the validation procedure allows comparing the relative values of loss of different covariance estimators.
+Therefore, validation procedure allow comparing the relative values of the average loss attained by different covariance matrix estimators.
 
-Indeed, the validation loss computed by 10-fold cross-validation (see Methods) accurately reproduced the relative values of the true loss and the rankings of the covariance estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).
+We drew 30 independent samples with sample sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the loss $\ell(C,\Sigma)$ for each of the five estimators.  
+The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  
+All the regularized estimators produced better estimates (lower loss) than the sample covariance matrix.  
+However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Row 5).
+The validation loss computed by 10-fold cross-validation (see Methods) accurately reproduced the relative values of the true loss and the rankings of the covariance estimators even without access to the ground truth (Fig.~\ref{fig:1} Row 6).
 
-In the example above, the data were sampled from  multivariate normal random variables.
-In such models, partial correlations perfectly characterize the conditional dependencies between variables and the graphical models of partial correlations exactly correspond to the conditional dependencies in the data.
+Note that when the ground truth had zero correlations (Column A), $C_{\sf factor}$ performed equally well to $C_{\sf diag}$ because it correctly inferred zero factors and only estimated the individual variances.
+Similarly, when the number of latent units was zero (Column C), $C_{\sf sparse+latent}$ performed nearly equally well to $C_{\sf sparse}$ because it correctly inferred zero latent units.
+With increasing sample sizes, all estimators converged to the ground truth (zero loss) but the estimators with correct structure outperformed the others even for large samples.
+
+In Gaussian models, partial correlations perfectly characterize the conditional dependencies between variables and the graphical models of partial correlations exactly correspond to the conditional dependencies in the data.
 To demonstrate that estimator rankings were robust to deviations from Gaussian models, we repeated the same cross-validated evaluation using pairwise Ising models to generate the data.
-
 Ising models have been used to infer functional connectivity from neuronal spike trains \cite{Hertz:2011}. 
@@
 This suggests that correlations between neurons build up from multiple chains of smaller interactions.
-Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29, with larger populations exhibiting greater uniformity of average partial correlations than the smaller populations ($p=0.002$ Brown-Forsythe test).
+Furthermore, the average partial correlations were less variable ($p=0.002$ Brown-Forsythe test): the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29.
 
@@
 \subsection*{Addressing ill-posedness}
-Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data can produce very large errors in the inferred network of dependencies.
+Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data can produce very large errors in the inferred network of dependencies (Fig.~\ref{fig:5} C and D).
 The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.

### 2014-12-17 final.tex
final revision in final.tex

--- 2014-12-07 manuscript.tex
@@
 \item[$C_{\sf factor}$] -- a low-rank approximation of the sample covariance matrix, representing inputs from unobserved shared factors (latent units).
-\item[$C_{\sf sparse}$] -- sparse partial correlations, \emph{i.e.}\;a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
+\item[$C_{\sf sparse}$] -- sparse partial correlations, \emph{i.e.}\;all but a small fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
 \item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons \emph{and} linear interactions with a number of latent units.
@@
 In our data, the sample correlation coefficients were largely positive and low.
-We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.
-This estimator revealed a sparse network of partial correlations (`interactions'), between the observed neurons; it also inferred latent units exerting linear effects on the observed neurons.
+We found that the most efficient estimate of the correlation matrix in these data was $C_{\sf sparse+latent}$.
+This estimator revealed a sparse network of partial correlations (`interactions'), between the observed neurons; it also inferred a number of latent units interacting with the observed neurons.
 We analyzed these networks of partial correlations and found the following: Whereas significant noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.
-Moreover, the average partial correlations and the inferred positive interactions in particular exhibited a stronger relationship to the physical distances and to the differences in preferred orientations than noise correlations.
+Moreover, the inferred positive interactions exhibited a substantially stronger relationship to the physical distances and to the differences in preferred orientations than noise correlations.
 In contrast, the inferred negative interactions were less selective.
@@
     \end{equation}
-    where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons in a time bin of some duration; and $\mu$ is the vector of expected firing rates.
-
-Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix} is
+    where the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons in a time bin of some duration, $\E{\cdot}$ denotes expectation, and $\mu$ is the vector of expected firing rates.
+
+Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ contains observed firing rates in time bin $t$, and an independent estimate of the mean firing rates $\bar x$, the \emph{sample covariance matrix},
 \begin{equation}\label{eq:sample}
@@
     \end{equation}
-where $n$ is the number of time bins in $T$. 
-The sample covariance matrix yields an unbiased estimate of the true covariance matrix ($\E{C_{\sf sample}}=\Sigma$) if the mean activity, $\bar x$, is estimated independently, in a separate data sample.
-In all cases when the unbiasedness of the sample covariance matrix matters in this paper, the means are estimated independently.
+where $n$ is the number of time bins in $T$, is an unbiased estimate of the true covariance matrix, \emph{i.e.} $\E{C_{\sf sample}}=\Sigma$. 
+In all cases when the unbiasedness of the sample covariance matrix matters in this paper, the mean activity is estimated independently from a separate sample.
 
@@
 Partial correlations figure prominently in probabilistic \emph{graphical modeling} wherein the joint distribution is explained by sets of pairwise interactions \cite{Whittaker:1990}.
-For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}.
+For multivariate Gaussian distributions, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}.
 More generally, partial correlations can serve as a measure of conditional independence under the assumption that dependencies in the system are close to linear effects \cite{Whittaker:1990,Baba:2004}.
@@
 
-Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model $L + D$, where $L$ is the $p\times p$ positive semidefinite matrix of a low rank and $D$ is a diagonal matrix of independent variances.
-This approximation is the basis for \emph{factor analysis} \cite{Anderson:2003}.
-The estimator,
+Estimator $C_{\sf factor}$ approximates the covariance matrix by the factor model, The estimator,
 \begin{equation}\label{eq:c-factor}
-C_{\sf factor} = L + (1-\lambda)D + \lambda \bar D,
-\end{equation}
-has two hyperparameters: the rank of $L$ (\emph{i.e.}\;the number of latent factors) and the shrinkage intensity $\lambda$ to shrink the independent variances toward their mean value $\bar D$.
+C_{\sf factor} = L + D,
+\end{equation}
+where $L$ is a $p\times p$ symmetric positive semidefinite matrix with low rank and $D$ is a diagonal matrix. 
+This approximation is the basis for \emph{factor analysis} \cite{Anderson:2003}, where matrix $L$ represents covariances arising from latent factors.
+The rank of $L$ corresponds to the number of latent factors.
+Matrix $D$ contains the variances of the cells' independent activity from the latent factors.
+The estimator is regularized by selecting the rank of $L$ and by shrinking the independent variances in $D$ toward their mean.
 The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (Fig.~\ref{fig:1} Row 1, B).
@@
 Here $S$ is a sparse matrix, \emph{i.e.}\;one in which a large fraction of off-diagonal elements are set to zero.  
-The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$. 
-The problem of finding the optimal set of non-zero elements of the precision matrix is known as \emph{covariance selection} \cite{Dempster:1972}. 
+The estimator is regularized by adjusting the sparsity (fraction of off-diagonal zeros) of $S$. 
+The problem of finding the optimal set of non-zero elements in $S$ is known as \emph{covariance selection} \cite{Dempster:1972}. 
 The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (`interactions') between some pairs of neurons (Fig.~\ref{fig:1} Row 1, C).
 
-Estimator $C_{\sf sparse+latent}$ is obtained by approximating the sample covariance matrix by a matrix whose inverse is the sum of a sparse component and a low-rank component:
+Estimator $C_{\sf sparse+latent}$ is obtained by approximating the sample covariance matrix by a matrix whose inverse is the difference of a sparse component and a low-rank component:
 \begin{equation}\label{eq:c-sl}
@@
 where $S$ is a sparse matrix and $L$ is a low-rank matrix.
-The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$.
-See Methods for a more detailed explanation.
+The estimator is regularized by adjusting the sparsity of $S$ and the rank of $L$.
+See Methods for more detailed explanations.
 The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (Fig.~\ref{fig:1} Row 1, D) \cite{Chandrasekaran:2010,Ma:2013}.
@@
 \end{equation}
-Therefore, validation procedure allow comparing the relative values of the average loss attained by different covariance matrix estimators.
+Therefore, validation procedures allow comparing the relative values of the loss attained by different covariance matrix estimators.
 
@@
 All the regularized estimators produced better estimates (lower loss) than the sample covariance matrix.  
-However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Row 5).
-The validation loss computed by 10-fold cross-validation (see Methods) accurately reproduced the relative values of the true loss and the rankings of the covariance estimators even without access to the ground truth (Fig.~\ref{fig:1} Row 6).
+However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Rows 4 and 5).
+The validation loss computed by 10-fold cross-validation (see Methods) accurately reproduced the relative values of the true loss and the rankings of the estimators even without access to the ground truth (Fig.~\ref{fig:1} Row 6).
 
@@
 Both models are maximum-entropy models constrained to match the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
-The partition function $Z(J,h)$ normalizes the distribution on the models' respective domains. 
+The partition function $Z(J,h)$ normalizes the distributions on the models' respective domains. 
 In the Gaussian model, the matrix $-J^{-1}$ is the covariance matrix; and the mean values are $\mu=J^{-1}h$.  
@@
 
-This simulation study demonstrated that cross-validated evaluation of regularized estimators of the covariance matrix of population activity can discriminate between structures of dependencies in the population.
-The identity of the most efficient covariance estimator for a particular neural circuit is therefore an empirical finding characterizing the nature of circuit  interactions.
+This simulation study demonstrated that cross-validated evaluation of regularized estimators of the covariance matrices of population activity can discriminate between structures of dependencies in the population.
+The selection of the most efficient covariance estimators for particular neural circuits is therefore an empirical finding characteristic of the nature of circuit interactions.
 
@@
 We recorded the calcium activity of densely sampled populations of neurons in layers 2/3 and upper layer 4 in primary visual cortex of sedated mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:3} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
-This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:3} C and D).
-The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:3} C).
+This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in $200\times200\times100$ $\mu$m$^3$ volumes of cortical tissue (Fig.~\ref{fig:3} C and D).
+The instantaneous firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:3} C).
 Only cells that produced detectable calcium activity were included in the analysis (see Methods).
 First, 30 repetitions of full-field drifting gratings of 16 directions were presented in random order.
-Each grating was displayed for 500 ms, without intervening blanks.
+Each grating was played for 500 ms, without intervening blanks.
 This stimulus was used to compute the orientation tuning of the recorded cells (Fig.~\ref{fig:3} D).
@@
 The sample correlation coefficients were largely positive and low (Fig.~\ref{fig:3} E and F).
-The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean across sites of 0.018 (Fig.~\ref{fig:6} D).
+The average value of the correlation coefficient across sites ranged from 0.0065 to 0.051 with the mean across sites of 0.018.
 
@@
 
-To test this hypothesis, we computed the relative validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
+To test this hypothesis, we computed the validation loss of estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.
 The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods).
 Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:4}).
-The respective median differences of the validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).
+The respective median differences of the validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, Wilcoxon signed rank test).
 
 \subsection*{Structure of $C_{\sf sparse+latent}$ estimates}
-We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:5} and Fig.~\ref{fig:6}).
+We examined the composition of the $C_{\sf sparse+latent}$ estimates for each imaged site (Fig.~\ref{fig:5} and Fig.~\ref{fig:6}).
 Although the regularized estimates were similar to the sample correlation matrix (Fig.~\ref{fig:5} A and B), the corresponding partial correlation matrices differed substantially (Fig.~\ref{fig:5} C and D).
@@
 In the example site with 7.2\% connectivity in $C_{\sf sparse+latent}$, only 27.7\% of the connections coincided with the above-threshold sample correlations (Fig.~\ref{fig:5} F, H, and I).
-In particular, most of the inferred negative interactions corresponded to low sample correlations (Fig.~\ref{fig:5} F) where high correlations should be expected given the rest of the correlation matrix.
+In particular, most of the inferred negative interactions corresponded to low sample correlations (Fig.~\ref{fig:5} F) where high correlations are expected given the rest of the correlation matrix.
 
@@
 
-We examined how the structure of the $C_{\sf sparse+latent}$ estimates related to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:7}).
+We then examined how the structure of the $C_{\sf sparse+latent}$ estimates related to the differences in orientation preference and to the physical distances separating pairs of cells (Fig.\;\ref{fig:7}).
 Five sites with highest pairwise connectivities were included in the analysis.
 Partial correlations were computed using Eq.~\ref{eq:partial} based on the regularized estimate, including both the sparse and the latent component.
-Connectivity was computed as the fraction of pairs of cells connected by non-zero elements (interactions) in the sparse component of the estimate, distinguishing between the positive and negative connectivities.
-
-First, we analyzed how correlations and connectivity depended on the difference in preferred orientations ($\Delta \mbox{ori}$) of pairs of significantly ($\alpha=0.05$) tuned cells.
-The partial correlations decayed more rapidly with $\Delta\mbox{ori}$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test of the difference of the linear regression coefficients).
+Connectivity was computed as the fraction of pairs of cells connected by non-zero elements (interactions) in the sparse component of the estimate, segregated into positive and negative connectivities.
+
+First, we analyzed how correlations and connectivity depended on the differences in preferred orientations ($\Delta \mbox{ori}$) of pairs of significantly ($\alpha=0.05$) tuned cells.
+The partial correlations decayed more rapidly with $\Delta\mbox{ori}$ than did sample correlations ($p<10^{-9}$ in each of the five sites, two-sample $t$-test of the difference of the linear regression coefficients in normalized data).
 Positive connectivity decreased with $\Delta\mbox{ori}$ ($p<0.005$ in each of the five sites, $t$-test on the logistic regression coefficient) whereas negative connectivity did not decrease (Fig.~\ref{fig:7} D): The slope in the logistic model of connectivity with respect to $\Delta\mbox{ori}$ was significantly higher for positive than for negative interactions ($p<0.04$ in each of the five sites, two-sample $t$-test of the difference of the logistic regression coefficient).
@@
 Again, the partial correlations decayed more rapidly both laterally and vertically than sample correlations ($p<10^{-6}$ in each of the five sites, for both lateral and vertical distances, two-sample $t$-test of the difference of the linear regression coefficients).
-Positive connectivity decayed with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression coefficient) (Fig.~\ref{fig:7} E), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$.
+Positive connectivity decayed with distance ($p<10^{-6}$ in each of the five sites for positive interactions and $p<0.05$ for negative interactions, $t$-test on the logistic regression coefficient in normalized data) (Fig.~\ref{fig:7} E), so that cells separated laterally by less than 25 $\mu\mbox{m}$ were 3.2 times more likely to be connected than cells separated laterally by more than 150 $\mu\mbox{m}$.
 Although the positive connectivity appeared to decay faster with vertical than with lateral distance, the differences in slopes of the respective logistic regression models were not significant with available data.
@@
 The goal of many studies of functional connectivity has been to estimate  anatomical connectivity from  observed multineuronal spiking activity.
-For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
+For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.
 Such signatures are ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.
@@
 Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network and external input.
-Therefore, conditional independence can signify the absence of a direct causal influence.
+Therefore, conditional independence shown statistically can signify the absence of a direct causal influence.
 
@@
 However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}.
-Until a physiological interpretation of functional connectivity is established, the physiological relevance of such analyses remains in question and we did not attempt them here.
+Until a physiological interpretation of functional connectivity is established, the physiological relevance of such analyses remains in question and we did not attempt applying graph-theoretical analyses to our results.
 
@@
 Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.
-While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}.
+While the temporal resolution of calcium signals is limited by the calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}.
 However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.
@@
 Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we \emph{a priori} favored solutions with sparse partial correlations.
-Under the assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.
+Under the assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can better represent the  functional dependencies in the circuit than correlations.
 
 \subsection*{Functional connectivity as coactivations}
-Another approach to describing the functional connectivity of a circuit is to isolate individual patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}.
-Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}.
-Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \emph{i.e.}\;groups of tightly interconnected groups of cells involved in a common computation.
+Another approach to describing the functional connectivity of a circuit is to isolate individual patterns of multineuronal coactivations.
+Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps} \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}.
+Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies, \emph{i.e.}\ groups of tightly interconnected groups of cells involved in a common computation \cite{Gerstein:1989, Ch:2010}.
 Coactivation patterns could also result from shared input from unobserved parts of the circuit, or global network fluctuations modulating the activity of the local circuit \cite{Okun:2012, Ecker:2014}.
@@
 As techniques have improved to allow recording from larger neuronal populations, experimental neuroscientists have addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted \cite{Tkacik:2013}.
-However, ambitious projects, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
+However, ambitious projects already underway, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations.
 Simply increasing recording duration will not be practical or sufficient, and the problem must be addressed by using regularized estimators.
-Regularization biases the solution toward a small subspace in order to counteract the effect of  sampling noise in the empirical data.
+Regularization biases the solution toward a small subspace in order to counteract the effects of  sampling noise in the empirical data.
 However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hinders interpretation.
@@
 Various model selection criteria have been devised to select between families of models and the optimal subsets of variables in a given model family based on observed data.
-Despite its computational requirements, cross-validation is among the most popular model selection methods due to its minimal assumptions about the data generating process \cite{Arlot:2010}.
+Despite its high computational demands, cross-validation is among the most popular model selection approaches due to its minimal assumptions about the data-generating process \cite{Arlot:2010}.
 
@@
 \subsection*{Visual stimulus}
-The visual stimulus consisted of full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s.
+The visual stimulus consisted of full-field drifting gratings with 90\% contrast, 10 cd/m$^2$ luminance, 0.08 cycles/degree spatial frequency, and 2 cycles/s temporal frequency.
 Two types of stimuli were presented for each imaging site: First, directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen directions of motion, 500 ms per direction, without blanks, with 12--30 trials for each direction of motion.
@@
 The variances (the diagonal of $C_{\sf sample}$) were shrunk linearly toward their mean value $\frac 1 p \Tr(C_{\sf sample})$:
-\begin{equation}
+\begin{equation}\label{eq:shrink}
 D = (1-\alpha)\mbox{diag}(C_{\sf sample}) + \alpha \frac 1 p \Tr(C_{\sf sample}) I
@@
 
-In estimator $C_{\sf factor}$ (Eq.~\ref{eq:c-factor}), the low-rank matrix $L$ and the diagonal matrix $D$ are found by solving the minimization problem
+In estimator $C_{\sf factor}$ (Eq.~\ref{eq:c-factor}), the low-rank matrix $L$ and the diagonal matrix $D$ were found by solving the minimization problem
 \begin{equation}
@@
 \end{equation}
-by an expectation-maximization (EM) algorithm with specified rank of $L$. In addition, the diagonal matrix of individual variances is shrunk toward its mean value according to Eq.~\ref{eq:c-factor}.
-
-In estimator $C_{\sf sparse}$ (Eq.~\ref{eq:c-sparse}), the sparse precision matrix $S$ is found by minimizing the $L_1$-penalized loss with regularization parameter $\lambda$:
+using an expectation-maximization (EM) algorithm for a specified rank of $L$. After that, the diagonal of $D$ was linearly shrunk toward the its mean diagonal value similar to Eq.~\ref{eq:shrink}.
+
+In estimator $C_{\sf sparse}$ (Eq.~\ref{eq:c-sparse}), the sparse precision matrix $S$ was found by minimizing the $L_1$-penalized loss with regularization parameter $\lambda$:
 \begin{equation}
@@
 Rather than finding $S_{12}$ and $S_{22}$ separately, $L$ can be estimated as a low-rank positive semidefinite matrix.
-To simultaneously optimize the sparse component $S$ and the low-rank component $L$, we adapted the loss function with $L_1$ penalty on $S$ combined with a penalty on the trace of $L$ \cite{Chandrasekaran:2010,Ma:2013}:
+To simultaneously optimize the sparse component $S$ and the low-rank component $L$, we adapted the loss function with an $L_1$ penalty on $S$ and another penalty on the trace of $L$ \cite{Chandrasekaran:2010,Ma:2013}:
 \begin{equation}\label{eq:ma}
@@
 
-Let $T_c$ and $T_c^\prime$ denote the time bin indices for the training and testing samples, respectively, limited to condition $c$.
-Here, the prime symbol marks quantities estimated from the testing sample.
+Let $T_c$ and $T_c^\prime$ denote the sets of time bin indices for the training and testing samples, respectively, limited to condition $c$.
 
@@
 The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (Fig.~\ref{fig:1} Row 2).
-Samples were then drawn from multivariate normal distributions with the respective true covariance matrices to be estimated by each of the estimators.
+Samples were then drawn from multivariate normal distributions models with the respective true covariance matrices to be estimated by each of the estimators.
+For Ising models, the negative inverse of the true covariance matrix was used as the matrix of coupling coefficients and the sampling was performed by the Metropolis-Hastings algorithm.
 
@@
 % 
-\bibliography{references}
+\begin{thebibliography}{100}
+\providecommand{\url}[1]{\texttt{#1}}
+\providecommand{\urlprefix}{URL }
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
+  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
+  \urlstyle{rm}\Url}\fi
+\providecommand{\bibAnnoteFile}[1]{%
+  \IfFileExists{#1}{\begin{quotation}\noindent\textsc{Key:} #1\\
+  \textsc{Annotation:}\ \input{#1}\end{quotation}}{}}
+\providecommand{\bibAnnote}[2]{%
+  \begin{quotation}\noindent\textsc{Key:} #1\\
+  \textsc{Annotation:}\ #2\end{quotation}}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{Averbeck:2006}
+Averbeck BB, Latham PE, Pouget A (2006) Neural correlations, population coding
+  and computation.
+\newblock Nat Rev Neurosci 7: 358--366.
+\bibAnnoteFile{Averbeck:2006}
+
+\bibitem{Zohary:1994}
+Zohary E, Shadlen MN, Newsome WT (1994) Correlated neuronal discharge rate and
+  its implications for psychophysical performance.
+\newblock Nature 370: 140--143.
+\bibAnnoteFile{Zohary:1994}
+
+\bibitem{Kohn:2005}
+Kohn A, Smith MA (2005) Stimulus dependence of neuronal correlation in primary
+  visual cortex of the macaque.
+\newblock J Neurosci 25: 3661-73.
+\bibAnnoteFile{Kohn:2005}
+
+\bibitem{Bair:2001}
+Bair W, Zohary E, Newsome WT (2001) Correlated firing in macaque visual area
+  mt: time scales and relationship to behavior.
+\newblock The journal of Neuroscience 21: 1676--1697.
+\bibAnnoteFile{Bair:2001}
+
+\bibitem{Ecker:2010}
+Ecker AS, Berens P, Keliris GA, Bethge M, Logothetis NK, et~al. (2010)
+  Decorrelated neuronal firing in cortical microcircuits.
+\newblock Science 327: 584-587.
+\bibAnnoteFile{Ecker:2010}
+
+\bibitem{Abbott:1999}
+Abbott L, Dayan P (1999) The effect of correlated variability on the accuracy
+  of a population code.
+\newblock Neural computation 11: 91--101.
+\bibAnnoteFile{Abbott:1999}
+
+\bibitem{Sompolinsky:2001}
+Sompolinsky H, Yoon H, Kang K, Shamir M (2001) Population coding in neuronal
+  systems with correlated noise.
+\newblock Physical Review E 64: 051904.
+\bibAnnoteFile{Sompolinsky:2001}
+
+\bibitem{Nirenberg:2003}
+Nirenberg S, Latham PE (2003) Decoding neuronal spike trains: How important are
+  correlations?
+\newblock Proceedings of the National Academy of Sciences 100: 7348--7353.
+\bibAnnoteFile{Nirenberg:2003}
+
+\bibitem{Josic:2009}
+Josic K, Shea-Brown E, Doiron B, de~la Rocha J (2009) Stimulus-dependent
+  correlations and population codes.
+\newblock Neural computation 21: 2774--2804.
+\bibAnnoteFile{Josic:2009}
+
+\bibitem{Berens:2011}
+Berens P, Ecker AS, Gerwinn S, Tolias AS, Bethge M (2011) Reassessing optimal
+  neural population codes with neurometric functions.
+\newblock Proc Natl Acad Sci U S A 108: 4423-8.
+\bibAnnoteFile{Berens:2011}
+
+\bibitem{Ecker:2011}
+Ecker AS, Berens P, Tolias AS, Bethge M (2011) The effect of noise correlations
+  in populations of diversely tuned neurons.
+\newblock The Journal of Neuroscience 31: 14272--14283.
+\bibAnnoteFile{Ecker:2011}
+
+\bibitem{Gerstein:1964}
+Gerstein G, Clark W (1964) Simultaneous studies of firing patterns in several
+  neurons.
+\newblock Science 143: 1325--1327.
+\bibAnnoteFile{Gerstein:1964}
+
+\bibitem{Smith:2008}
+Smith MA, Kohn A (2008) Spatial and temporal scales of neuronal correlation in
+  primary visual cortex.
+\newblock J Neurosci 28: 12591--12603.
+\bibAnnoteFile{Smith:2008}
+
+\bibitem{Denman:2013}
+Denman DJ, Contreras D (2013) The structure of pairwise correlation in mouse
+  primary visual cortex reveals functional organization in the absence of an
+  orientation map.
+\newblock Cereb Cortex .
+\bibAnnoteFile{Denman:2013}
+
+\bibitem{Ko:2011}
+Ko H, Hofer SB, Pichler B, Buchanan KA, Sj{\"o}str{\"o}m PJ, et~al. (2011)
+  Functional specificity of local synaptic connections in neocortical networks.
+\newblock Nature 473: 87--91.
+\bibAnnoteFile{Ko:2011}
+
+\bibitem{Arieli:1995}
+Arieli A, Shoham D, Hildesheim R, Grinvald A (1995) Coherent spatiotemporal
+  patterns of ongoing activity revealed by real-time optical imaging coupled
+  with single-unit recording in the cat visual cortex.
+\newblock Journal of Neurophysiology 73: 2072--2093.
+\bibAnnoteFile{Arieli:1995}
+
+\bibitem{Chiu:2002}
+Chiu C, Weliky M (2002) Relationship of correlated spontaneous activity to
+  functional ocular dominance columns in the developing visual cortex.
+\newblock Neuron 35: 1123--1134.
+\bibAnnoteFile{Chiu:2002}
+
+\bibitem{Kenet:2003}
+Kenet T, Bibitchkov D, Tsodyks M, Grinvald A, Arieli A (2003) Spontaneously
+  emerging cortical representations of visual attributes.
+\newblock Nature 425: 954--956.
+\bibAnnoteFile{Kenet:2003}
+
+\bibitem{Cohen:2008}
+Cohen MR, Newsome WT (2008) Context-dependent changes in functional circuitry
+  in visual area mt.
+\newblock Neuron 60: 162--173.
+\bibAnnoteFile{Cohen:2008}
+
+\bibitem{Cohen:2009}
+Cohen MR, Maunsell JH (2009) Attention improves performance primarily by
+  reducing interneuronal correlations.
+\newblock Nature neuroscience 12: 1594--1600.
+\bibAnnoteFile{Cohen:2009}
+
+\bibitem{Rothschild:2010}
+Rothschild G, Nelken I, Mizrahi A (2010) Functional organization and population
+  dynamics in the mouse primary auditory cortex.
+\newblock Nature neuroscience 13: 353--360.
+\bibAnnoteFile{Rothschild:2010}
+
+\bibitem{Smith:2013b}
+Smith MA, Sommer MA (2013) Spatial and temporal scales of neuronal correlation
+  in visual area v4.
+\newblock The Journal of Neuroscience 33: 5422--5432.
+\bibAnnoteFile{Smith:2013b}
+
+\bibitem{Hofer:2011}
+Hofer SB, Ko H, Pichler B, Vogelstein J, Ros H, et~al. (2011) Differential
+  connectivity and response dynamics of excitatory and inhibitory neurons in
+  visual cortex.
+\newblock Nature neuroscience 14: 1045--1052.
+\bibAnnoteFile{Hofer:2011}
+
+\bibitem{Hansen:2012}
+Hansen BJ, Chelaru MI, Dragoi V (2012) Correlated variability in laminar
+  cortical circuits.
+\newblock Neuron 76: 590--602.
+\bibAnnoteFile{Hansen:2012}
+
+\bibitem{Smith:2013}
+Smith MA, Jia X, Zandvakili A, Kohn A (2013) Laminar dependence of neuronal
+  correlations in visual cortex.
+\newblock Journal of neurophysiology 109: 940--947.
+\bibAnnoteFile{Smith:2013}
+
+\bibitem{Golshani:2009}
+Golshani P, Gon{\c c}alves JT, Khoshkhoo S, Mostany R, Smirnakis S, et~al.
+  (2009) Internally mediated developmental desynchronization of neocortical
+  network activity.
+\newblock J Neurosci 29: 10890-9.
+\bibAnnoteFile{Golshani:2009}
+
+\bibitem{Gu:2011}
+Gu Y, Liu S, Fetsch CR, Yang Y, Fok S, et~al. (2011) Perceptual learning
+  reduces interneuronal correlations in macaque visual cortex.
+\newblock Neuron 71: 750--761.
+\bibAnnoteFile{Gu:2011}
+
+\bibitem{Ko:2013}
+Ko H, Cossell L, Baragli C, Antolik J, Clopath C, et~al. (2013) The emergence
+  of functional microcircuits in visual cortex.
+\newblock Nature 496: 96--100.
+\bibAnnoteFile{Ko:2013}
+
+\bibitem{Greenberg:2008}
+Greenberg DS, Houweling AR, Kerr JN (2008) Population imaging of ongoing
+  neuronal activity in the visual cortex of awake rats.
+\newblock Nature neuroscience 11: 749--751.
+\bibAnnoteFile{Greenberg:2008}
+
+\bibitem{Goard:2009}
+Goard M, Dan Y (2009) Basal forebrain activation enhances cortical coding of
+  natural scenes.
+\newblock Nat Neurosci 12: 1444-9.
+\bibAnnoteFile{Goard:2009}
+
+\bibitem{Kohn:2009}
+Kohn A, Zandvakili A, Smith MA (2009) Correlations and brain states: from
+  electrophysiology to functional imaging.
+\newblock Curr Opin Neurobiol 19: 434-8.
+\bibAnnoteFile{Kohn:2009}
+
+\bibitem{Ecker:2014}
+Ecker AS, Berens P, Cotton RJ, Subramaniyan M, Denfield GH, et~al. (2014) State
+  dependence of noise correlations in macaque primary visual cortex.
+\newblock Neuron 82: 235--248.
+\bibAnnoteFile{Ecker:2014}
+
+\bibitem{Renart:2010}
+Renart A, de~la Rocha J, Bartho P, Hollender L, Parga N, et~al. (2010) The
+  asynchronous state in cortical circuits.
+\newblock Science 327: 587-90.
+\bibAnnoteFile{Renart:2010}
+
+\bibitem{Perkel:1967}
+Perkel DH, Gerstein GL, Moore GP (1967) Neuronal spike trains and stochastic
+  point processes: {II}. simultaneous spike trains.
+\newblock Biophysical journal 7: 419--440.
+\bibAnnoteFile{Perkel:1967}
+
+\bibitem{Moore:1970}
+Moore GP, Segundo JP, Perkel DH, Levitan H (1970) Statistical signs of synaptic
+  interaction in neurons.
+\newblock Biophysical Journal 10: 876--900.
+\bibAnnoteFile{Moore:1970}
+
+\bibitem{Shadlen:1998}
+Shadlen MN, Newsome WT (1998) The variable discharge of cortical neurons:
+  implications for connectivity, computation, and information coding.
+\newblock J Neurosci 18: 3870-96.
+\bibAnnoteFile{Shadlen:1998}
+
+\bibitem{Salinas:2001}
+Salinas E, Sejnowski TJ (2001) Correlated neuronal activity and the flow of
+  neural information.
+\newblock Nature Reviews Neuroscience 2: 539--550.
+\bibAnnoteFile{Salinas:2001}
+
+\bibitem{Ostojic:2009}
+Ostojic S, Brunel N, Hakim V (2009) How connectivity, background activity, and
+  synaptic properties shape the cross-correlation between spike trains.
+\newblock The Journal of Neuroscience 29: 10234--10253.
+\bibAnnoteFile{Ostojic:2009}
+
+\bibitem{Rosenbaum:2011}
+Rosenbaum R, Josi{\'c} K (2011) Mechanisms that modulate the transfer of
+  spiking correlations.
+\newblock Neural computation 23: 1261--1305.
+\bibAnnoteFile{Rosenbaum:2011}
+
+\bibitem{Whittaker:1990}
+Whittaker J (1990) Graphical models in applied multivariate statistics.
+\newblock Wiley Publishing.
+\bibAnnoteFile{Whittaker:1990}
+
+\bibitem{Schafer:2005}
+Sch{\"a}fer J, Strimmer K, et~al. (2005) A shrinkage approach to large-scale
+  covariance matrix estimation and implications for functional genomics.
+\newblock Statistical applications in genetics and molecular biology 4: 32.
+\bibAnnoteFile{Schafer:2005}
+
+\bibitem{Peng:2009}
+Peng J, Wang P, Zhou N, Zhu J (2009) Partial correlation estimation by joint
+  sparse regression models.
+\newblock Journal of the American Statistical Association 104.
+\bibAnnoteFile{Peng:2009}
+
+\bibitem{Varoquaux:2012}
+Varoquaux G, Gramfort A, Poline JB, Thirion B (2012) Markov models for fmri
+  correlation structure: is brain functional connectivity small world, or
+  decomposable into networks?
+\newblock Journal of Physiology-Paris 106: 212--221.
+\bibAnnoteFile{Varoquaux:2012}
+
+\bibitem{Ryali:2012}
+Ryali S, Chen T, Supekar K, Menon V (2012) Estimation of functional
+  connectivity in fmri data using stability selection-based sparse partial
+  correlation with elastic net penalty.
+\newblock Neuroimage 59: 3852--3861.
+\bibAnnoteFile{Ryali:2012}
+
+\bibitem{Ledoit:2004}
+Ledoit O, Wolf M (2004) A well-conditioned estimator for large-dimensional
+  covariance matrices.
+\newblock Journal of multivariate analysis 88: 365--411.
+\bibAnnoteFile{Ledoit:2004}
+
+\bibitem{Bickel:2006}
+Bickel PJ, Li B, Tsybakov AB, van~de Geer SA, Yu B, et~al. (2006)
+  Regularization in statistics.
+\newblock Test 15: 271--344.
+\bibAnnoteFile{Bickel:2006}
+
+\bibitem{Ledoit:2003}
+Ledoit O, Wolf M (2003) Improved estimation of the covariance matrix of stock
+  returns with an application to portfolio selection.
+\newblock Journal of Empirical Finance 10: 603--621.
+\bibAnnoteFile{Ledoit:2003}
+
+\bibitem{Friedman:1989}
+Friedman JH (1989) Regularized discriminant analysis.
+\newblock Journal of the American statistical association 84: 165--175.
+\bibAnnoteFile{Friedman:1989}
+
+\bibitem{Berens:2012}
+Berens P, Ecker AS, Cotton RJ, Ma WJ, Bethge M, et~al. (2012) A fast and simple
+  population code for orientation in primate v1.
+\newblock The Journal of Neuroscience 32: 10618--10626.
+\bibAnnoteFile{Berens:2012}
+
+\bibitem{Alivisatos:2013}
+Alivisatos AP, Chun M, Church GM, Deisseroth K, Donoghue JP, et~al. (2013) The
+  brain activity map.
+\newblock Science 339: 1284--1285.
+\bibAnnoteFile{Alivisatos:2013}
+
+\bibitem{Anderson:2003}
+Anderson T (2003) An introduction to multivariate statistical analysis.
+\newblock Wiley series in probability and statistics .
+\bibAnnoteFile{Anderson:2003}
+
+\bibitem{Dempster:1972}
+Dempster A (1972) Covariance selection.
+\newblock Biometrics : 157--175.
+\bibAnnoteFile{Dempster:1972}
+
+\bibitem{Baba:2004}
+Baba K, Shibata R, Sibuya M (2004) Partial correlation and conditional
+  correlation as measures of conditional independence.
+\newblock Australian \& New Zealand Journal of Statistics 46: 657--664.
+\bibAnnoteFile{Baba:2004}
+
+\bibitem{Chandrasekaran:2010}
+Chandrasekaran V, Parrilo PA, Willsky AS (2010) Latent variable graphical model
+  selection via convex optimization.
+\newblock In: Communication, Control, and Computing (Allerton), 2010 48th
+  Annual Allerton Conference on. IEEE, pp. 1610--1613.
+\bibAnnoteFile{Chandrasekaran:2010}
+
+\bibitem{Ma:2013}
+Ma S, Xue L, Zou H (2013) Alternating direction methods for latent variable
+  gaussian graphical model selection.
+\newblock Neural computation : 1--27.
+\bibAnnoteFile{Ma:2013}
+
+\bibitem{Hertz:2011}
+Hertz J, Roudi Y, Tyrcha J (2013) Ising models for inferring network structure
+  from spike data.
+\newblock In: Quiroga RQ, Panzeri S, editors, Principles of neural coding, CRC
+  Press.
+\bibAnnoteFile{Hertz:2011}
+
+\bibitem{Jaynes:1957}
+Jaynes ET (1957) Information theory and statistical mechanics.
+\newblock Phys Rev 106: 620--630.
+\bibAnnoteFile{Jaynes:1957}
+
+\bibitem{Reddy:2005}
+Reddy GD, Saggau P (2005) Fast three-dimensional laser scanning scheme using
+  acousto-optic deflectors.
+\newblock J Biomed Opt 10: 064038.
+\bibAnnoteFile{Reddy:2005}
+
+\bibitem{Katona:2012}
+Katona G, Szalay G, Ma{\'a}k P, Kasz{\'a}s A, Veress M, et~al. (2012) Fast
+  two-photon in vivo imaging with three-dimensional random-access scanning in
+  large tissue volumes.
+\newblock Nat Methods .
+\bibAnnoteFile{Katona:2012}
+
+\bibitem{Cotton:2013}
+Cotton RJ, Froudarakis E, Storer P, Saggau P, Tolias AS (2013)
+  Three-dimensional mapping of microcircuit correlation structure.
+\newblock Frontiers in Neural Circuits 7: 151.
+\bibAnnoteFile{Cotton:2013}
+
+\bibitem{Vogelstein:2010}
+Vogelstein JT, Packer AM, Machado TA, Sippy T, Babadi B, et~al. (2010) Fast
+  nonnegative deconvolution for spike train inference from population calcium
+  imaging.
+\newblock Journal of neurophysiology 104: 3691--3704.
+\bibAnnoteFile{Vogelstein:2010}
+
+\bibitem{Feldt:2011}
+Feldt S, Bonifazi P, Cossart R (2011) Dissecting functional connectivity of
+  neuronal microcircuits: experimental and theoretical insights.
+\newblock Trends in neurosciences 34: 225--236.
+\bibAnnoteFile{Feldt:2011}
+
+\bibitem{Malmersjo:2013}
+Malmersj{\"o} S, Rebellato P, Smedler E, Planert H, Kanatani S, et~al. (2013)
+  Neural progenitors organize in small-world networks to promote cell
+  proliferation.
+\newblock Proceedings of the National Academy of Sciences 110: E1524--E1532.
+\bibAnnoteFile{Malmersjo:2013}
+
+\bibitem{Sadovsky:2014}
+Sadovsky AJ, MacLean JN (2014) Mouse visual neocortex supports multiple
+  stereotyped patterns of microcircuit activity.
+\newblock The Journal of Neuroscience 34: 7769-7777.
+\bibAnnoteFile{Sadovsky:2014}
+
+\bibitem{Alonso:1998}
+Alonso JM, Martinez LM (1998) Functional connectivity between simple cells and
+  complex cells in cat striate cortex.
+\newblock Nature neuroscience 1: 395--403.
+\bibAnnoteFile{Alonso:1998}
+
+\bibitem{Aertsen:1989}
+Aertsen A, Gerstein G, Habib M, Palm G (1989) Dynamics of neuronal firing
+  correlation: modulation of ``effective connectivity".
+\newblock Journal of neurophysiology 61: 900--917.
+\bibAnnoteFile{Aertsen:1989}
+
+\bibitem{Pillow:2008}
+Pillow JW, Shlens J, Paninski L, Sher A, Litke AM, et~al. (2008)
+  Spatio-temporal correlations and visual signalling in a complete neuronal
+  population.
+\newblock Nature 454: 995--999.
+\bibAnnoteFile{Pillow:2008}
+
+\bibitem{Schneidman:2006}
+Schneidman E, Berry MJ 2nd, Segev R, Bialek W (2006) Weak pairwise correlations
+  imply strongly correlated network states in a neural population.
+\newblock Nature 440: 1007-12.
+\bibAnnoteFile{Schneidman:2006}
+
+\bibitem{Tkacik:2006}
+Tkacik G, Schneidman E, Berry I, Michael J, Bialek W (2006) Ising models for
+  networks of real neurons.
+\newblock arXiv preprint q-bio/0611072 .
+\bibAnnoteFile{Tkacik:2006}
+
+\bibitem{Yu:2008}
+Yu S, Huang D, Singer W, Nikolic D (2008) A small world of neuronal synchrony.
+\newblock Cereb Cortex 18: 2891-901.
+\bibAnnoteFile{Yu:2008}
+
+\bibitem{Tang:2008}
+Tang A, Jackson D, Hobbs J, Chen W, Smith JL, et~al. (2008) A maximum entropy
+  model applied to spatial and temporal correlations from cortical networks in
+  vitro.
+\newblock J Neurosci 28: 505-18.
+\bibAnnoteFile{Tang:2008}
+
+\bibitem{Shlens:2009}
+Shlens J, Field GD, Gauthier JL, Greschner M, Sher A, et~al. (2009) The
+  structure of large-scale synchronized firing in primate retina.
+\newblock J Neurosci 29: 5022-31.
+\bibAnnoteFile{Shlens:2009}
+
+\bibitem{Zalesky:2012}
+Zalesky A, Fornito A, Bullmore E (2012) On the use of correlation as a measure
+  of network connectivity.
+\newblock Neuroimage 60: 2096--2106.
+\bibAnnoteFile{Zalesky:2012}
+
+\bibitem{Grewe:2010}
+Grewe BF, Langer D, Kasper H, Kampa BM, Helmchen F (2010) High-speed in vivo
+  calcium imaging reveals neuronal network activity with near-millisecond
+  precision.
+\newblock Nat Meth 7: 399--405.
+\bibAnnoteFile{Grewe:2010}
+
+\bibitem{Theis:2014}
+Theis L, Berens P, Froudarakis E, Reimer J, Roman-Roson M, et~al. (2014)
+  Supervised learning sets benchmark for robust spike detection from calcium
+  imaging signals.
+\newblock bioRxiv : 010777.
+\bibAnnoteFile{Theis:2014}
+
+\bibitem{Leung:2013}
+Leung LC, Wang GX, Mourrain P (2013) Imaging zebrafish neural circuitry from
+  whole brain to synapse.
+\newblock Frontiers in neural circuits 7.
+\bibAnnoteFile{Leung:2013}
+
+\bibitem{Ahrens:2013}
+Ahrens MB, Orger MB, Robson DN, Li JM, Keller PJ (2013) Whole-brain functional
+  imaging at cellular resolution using light-sheet microscopy.
+\newblock Nature methods 10: 413--420.
+\bibAnnoteFile{Ahrens:2013}
+
+\bibitem{Perin:2011}
+Perin R, Berger TK, Markram H (2011) A synaptic organizing principle for
+  cortical neuronal groups.
+\newblock Proc Natl Acad Sci U S A 108: 5419-24.
+\bibAnnoteFile{Perin:2011}
+
+\bibitem{Gerstein:1989}
+Gerstein GL, Bedenbaugh P, Aertsen AM (1989) Neuronal assemblies.
+\newblock Biomedical Engineering, IEEE Transactions on 36: 4--14.
+\bibAnnoteFile{Gerstein:1989}
+
+\bibitem{Chapin:1999}
+Chapin JK, Nicolelis MA (1999) Principal component analysis of neuronal
+  ensemble activity reveals multidimensional somatosensory representations.
+\newblock Journal of neuroscience methods 94: 121--140.
+\bibAnnoteFile{Chapin:1999}
+
+\bibitem{Peyrache:2010}
+Peyrache A, Benchenane K, Khamassi M, Wiener SI, Battaglia FP (2010) Principal
+  component analysis of ensemble recordings reveals cell assemblies at high
+  temporal resolution.
+\newblock Journal of computational neuroscience 29: 309--325.
+\bibAnnoteFile{Peyrache:2010}
+
+\bibitem{Ch:2010}
+Ch'Ng YH, Reid RC (2010) Cellular imaging of visual cortex reveals the spatial
+  and functional organization of spontaneous activity.
+\newblock Frontiers in integrative neuroscience 4.
+\bibAnnoteFile{Ch:2010}
+
+\bibitem{Lopes:2011}
+Lopes-dos Santos V, Conde-Ocazionez S, Nicolelis MA, Ribeiro ST, Tort AB (2011)
+  Neuronal assembly detection and cell membership specification by principal
+  component analysis.
+\newblock PloS one 6: e20996.
+\bibAnnoteFile{Lopes:2011}
+
+\bibitem{Lopes:2013}
+Lopes-dos Santos V, Ribeiro S, Tort AB (2013) Detecting cell assemblies in
+  large neuronal populations.
+\newblock Journal of Neuroscience Methods 220: 149 -- 166.
+\bibAnnoteFile{Lopes:2013}
+
+\bibitem{Okun:2012}
+Okun M, Yger P, Marguet SL, Gerard-Mercier F, Benucci A, et~al. (2012)
+  Population rate dynamics and multineuron firing patterns in sensory cortex.
+\newblock J Neurosci 32: 17108-19.
+\bibAnnoteFile{Okun:2012}
+
+\bibitem{Ganmor:2011}
+Ganmor E, Segev R, Schneidman E (2011) Sparse low-order interaction network
+  underlies a highly correlated and learnable neural population code.
+\newblock Proc Natl Acad Sci U S A 108: 9679-84.
+\bibAnnoteFile{Ganmor:2011}
+
+\bibitem{Tkacik:2013}
+Tka{\v{c}}ik G, Marre O, Amodei D, Schneidman E, Bialek W, et~al. (2013)
+  Searching for collective behavior in a network of real neurons.
+\newblock arXiv preprint arXiv:13063061 .
+\bibAnnoteFile{Tkacik:2013}
+
+\bibitem{Pfau:2013}
+Pfau D, Pnevmatikakis EA, Paninski L (2013) Robust learning of low-dimensional
+  dynamics from large neural ensembles.
+\newblock In: Advances in Neural Information Processing Systems. pp.
+  2391--2399.
+\bibAnnoteFile{Pfau:2013}
+
+\bibitem{Koster:2013}
+K{\"o}ster U, Sohl-Dickstein J, Gray CM, Olshausen BA (2013) Higher order
+  correlations within cortical layers dominate functional connectivity in
+  microcolumns.
+\newblock arXiv preprint arXiv:13010050 .
+\bibAnnoteFile{Koster:2013}
+
+\bibitem{Roudi:2009}
+Roudi Y, Nirenberg S, Latham PE (2009) Pairwise maximum entropy models for
+  studying large biological systems: when they can work and when they can't.
+\newblock PLoS computational biology 5: e1000380.
+\bibAnnoteFile{Roudi:2009}
+
+\bibitem{Arlot:2010}
+Arlot S, Celisse A (2010) A survey of cross-validation procedures for model
+  selection.
+\newblock Statistics Surveys 4: 40--79.
+\bibAnnoteFile{Arlot:2010}
+
+\bibitem{Efron:1977}
+Efron B, Morris CN (1977) Stein's paradox in statistics.
+\newblock WH Freeman.
+\bibAnnoteFile{Efron:1977}
+
+\bibitem{Song:2005}
+Song S, Sj{\"o}str{\"o}m PJ, Reigl M, Nelson S, Chklovskii DB (2005) Highly
+  nonrandom features of synaptic connectivity in local cortical circuits.
+\newblock PLoS biology 3: e68.
+\bibAnnoteFile{Song:2005}
+
+\bibitem{Oswald:2008}
+Oswald AMM, Reyes AD (2008) Maturation of intrinsic and synaptic properties of
+  layer 2/3 pyramidal neurons in mouse auditory cortex.
+\newblock Journal of neurophysiology 99: 2998.
+\bibAnnoteFile{Oswald:2008}
+
+\bibitem{Adesnik:2010}
+Adesnik H, Scanziani M (2010) Lateral competition for cortical space by
+  layer-specific horizontal circuits.
+\newblock Nature 464: 1155--1160.
+\bibAnnoteFile{Adesnik:2010}
+
+\bibitem{Fino:2011}
+Fino E, Yuste R (2011) Dense inhibitory connectivity in neocortex.
+\newblock Neuron 69: 1188-203.
+\bibAnnoteFile{Fino:2011}
+
+\bibitem{Isaacson:2011}
+Isaacson JS, Scanziani M (2011) How inhibition shapes cortical activity.
+\newblock Neuron 72: 231--243.
+\bibAnnoteFile{Isaacson:2011}
+
+\bibitem{Levy:2012}
+Levy RB, Reyes AD (2012) Spatial profile of excitatory and inhibitory synaptic
+  connectivity in mouse primary auditory cortex.
+\newblock J Neurosci 32: 5609-19.
+\bibAnnoteFile{Levy:2012}
+
+\bibitem{Packer:2011}
+Packer AM, Yuste R (2011) Dense, unspecific connectivity of neocortical
+  parvalbumin-positive interneurons: a canonical microcircuit for inhibition?
+\newblock The Journal of Neuroscience 31: 13260--13271.
+\bibAnnoteFile{Packer:2011}
+
+\bibitem{Reimer:2014}
+Reimer J, Froudarakis E, Cadwell CR, Yatsenko D, Denfield GH, et~al. (2014)
+  Pupil fluctuations track fast switching of cortical states during quiet
+  wakefulness.
+\newblock Neuron 84: 355--362.
+\bibAnnoteFile{Reimer:2014}
+
+\bibitem{Fu:2014}
+Fu Y, Tucciarone JM, Espinosa JS, Sheng N, Darcy DP, et~al. (2014) A cortical
+  circuit for gain control by behavioral state.
+\newblock Cell 156: 1139--1152.
+\bibAnnoteFile{Fu:2014}
+
+\bibitem{Meinshausen:2006}
+Meinshausen N, B{\"u}hlmann P (2006) High-dimensional graphs and variable
+  selection with the lasso.
+\newblock The Annals of Statistics 34: 1436--1462.
+\bibAnnoteFile{Meinshausen:2006}
+
+\bibitem{Friedman:2008}
+Friedman J, Hastie T, Tibshirani R (2008) Sparse inverse covariance estimation
+  with the graphical lasso.
+\newblock Biostatistics 9: 432--441.
+\bibAnnoteFile{Friedman:2008}
+
+\bibitem{Fazel:2002}
+Fazel M (2002) Matrix rank minimization with applications.
+\newblock Ph.D. thesis, Stanford University.
+\bibAnnoteFile{Fazel:2002}
+
+\bibitem{Recht:2010}
+Recht B, Fazel M, Parrilo PA (2010) Guaranteed minimum-rank solutions of linear
+  matrix equations via nuclear norm minimization.
+\newblock SIAM review 52: 471--501.
+\bibAnnoteFile{Recht:2010}
+
+\bibitem{Vogels:1989}
+Vogels R, Spileers W, Orban G (1989) The response variability of striate
+  cortical neurons in the behaving monkey.
+\newblock Experimental brain research 77: 432--436.
+\bibAnnoteFile{Vogels:1989}
+
+\bibitem{Ponce:2013}
+Ponce-Alvarez A, Thiele A, Albright TD, Stoner GR, Deco G (2013)
+  Stimulus-dependent variability and noise correlations in cortical mt neurons.
+\newblock Proceedings of the National Academy of Sciences 110: 13162--13167.
+\bibAnnoteFile{Ponce:2013}
+
+\end{thebibliography}
 

### 2015-02-24 manuscript.tex

--- 2014-12-06 manuscript.tex
@@
 In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of synaptic connectivity \cite{Gerstein:1964}.
-Interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various aspects of circuit organization such as the physical distances between the neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2014, Renart:2010}.
+Interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various aspects of circuit organization such as the physical distances between the neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010}.
 
@@
 Improved estimation of the correlation matrix is beneficial in itself.
-For example, improved estimates can be used to optimize  decoding of the population activity \cite{Friedman:1989, Berens:2012}.
+For example, improved estimates can be used to optimize  decoding of the population activity \cite{Friedman:1989}.
 But reduced estimation error is not the only benefit of regularization.
@@
 
-Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an unbiased estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix} is
+Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an independent unbiased estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix},
 \begin{equation}\label{eq:sample}
@@
     \end{equation}
-where $n$ is the number of time bins in $T$.
-The sample covariance matrix yields an unbiased estimate of the covariance matrix ($\E{C_{\sf sample}}=\Sigma$) if the mean, $\bar x$, is independent of the data sample, $\{x(t): t\in T$\}.
+where $n$ is the number of time bins in $T$, yields an unbiased estimate of the covariance matrix so that $\E{C_{\sf sample}}=\Sigma$.
+
 When the mean is estimated from the same sample, $C_{\sf sample}$ becomes biased toward zero.
-However, in all the cases where unbiasedness is required, we will estimate the mean from a separate sample.
+However, in all the cases when the unbiasedness is required in our study, the mean will be estimated from a separate sample.
 
@@
 Ising models have been used to infer functional connectivity from neuronal spike trains \cite{Hertz:2011}. 
-Conveniently, the Ising model has equivalent mathematical form to the Gaussian distribution,
+Conveniently, the Ising model is described by a probability distribution equivalent in form to the Gaussian distribution,
 \begin{equation}\label{eq:ising}
-    x\sim \frac 1 {Z(J,h)} \exp\left(\frac 1 2 x^\T J x + h^\T x \right)
+    x\sim \frac 1 {Z(J,h)} \exp\left(\frac 1 2 x^\T J x + h^\T x \right).
 \end{equation}
-but the Ising model is defined on the multivariate binary domain rather than the continuous domain. 
-Both models are maximum-entropy models constrained to match the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
+However, the Ising model is defined on the multivariate binary domain rather than the continuous domain. 
+Both models are maximum-entropy models constrained to match mean firing rates and the covariance matrix \cite{Jaynes:1957}.
 The partition function $Z(J,h)$ normalizes the distribution on the models' respective domains. 
-In the Gaussian model, the matrix $-J^{-1}$ is the covariance matrix; and the mean values are $\mu=J^{-1}h$.  
-For the Ising model, $J$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives, although they do not have a simple relationship to the means and the covariance matrix.
+In the Gaussian model, the matrix $-J^{-1}$ is the covariance matrix, and the mean values are $\mu=J^{-1}h$.  
+For the Ising model, $J$ is the matrix of pairwise interactions, and $h$ is the vector of the cells' individual activity drives.
+Neither $J$ nor $h$ have a simple relationship to the means and the covariance matrix.
 Both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $J$ indicate conditional independence between the corresponding pair of neurons.
 
-Indeed, despite their considerable departure from strictly linear conditional dependencies, Ising models yielded the same relationships between the performances of the covariance estimators as the Gaussian models in cross-validation (Fig.~\ref{fig:2}).
+Indeed, despite their considerable departure from strictly linear conditional dependencies, when we sampled data from an Ising models we observed similar relationships between the covariance estimators as in the Gaussian models in cross-validation (Fig.~\ref{fig:2}).
 Identical interaction matrices $J$  of the joint distributions over the observable and latent variables were used for both the Gaussian and the Ising models.
@@
 \subsection*{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
-We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of lightly sedated mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:3} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
+We recorded the calcium activity of densely sampled populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (Fig.~\ref{fig:3} A--B) \cite{Reddy:2005, Katona:2012, Cotton:2013}.
 This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (Fig.~\ref{fig:3} C and D).
@@
 This stimulus was used to compute the orientation tuning of the recorded cells (Fig.~\ref{fig:3} D).
-To estimate the noise correlation matrix, we presented only two distinct directions in some experiments or five directions in others with 100--300 repetitions of each direction.
+To estimate the noise correlation matrix, we presented only 2 distinct directions in some experiments or 5 directions in others with 100--300 repetitions of each direction.
 Each grating lasted 1 second and was followed by a 1-second blank.
@@
 
-Inference of the conditional dependencies also depends on the completeness of the recorded population:  To equate conditional dependency to direct interaction between two neurons, we must record from all neurons with which the pair interacts.
+Inference of the conditional dependencies also depends on the completeness of the recorded population:  To equate conditional dependencies to direct interaction between neurons, we must record from all neurons with which the pair interacts.
 Unobserved portions of the circuit may manifest as conditional dependencies between observed neurons that do not interact.
@@
 However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.
-Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013, Theis:2014}.
+Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.
 The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.
@@
 Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \emph{i.e.}\;groups of tightly interconnected groups of cells involved in a common computation.
-Coactivation patterns could also result from shared input from unobserved parts of the circuit, or global network fluctuations modulating the activity of the local circuit \cite{Okun:2012, Ecker:2014}.
+Coactivation patterns could also result from shared input from unobserved parts of the circuit, or global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.
 
@@
 \subsection*{Combining pairwise interactions and coactivations}
-In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables, higher-order correlations, or global network fluctuations \cite{Ganmor:2011, Tkacik:2013, Pfau:2013, Koster:2013, Ecker:2014}.
+In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables \cite{Koster:2013},  high-order correlations \cite{Ganmor:2011}, or global network fluctuations \cite{Tkacik:2013}.
 
@@
 We showed that among several models a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix for cortical microcircuits.
-This finding is valuable in itself: improved estimates of the noise covariance matrix for large datasets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Ecker:2011}
+This finding is valuable in itself: improved estimates of the noise covariance matrix for large datasets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Ecker:2011}
 
@@
 The surgical procedures and data acquisition were performed as described in \cite{Cotton:2013}: C57BL/6J mice (aged p40--60) were used.
-For surgery, animals were initially anesthetized with isoflurane (3\%).
-During the experiments, animals were sedated with a mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
+Anesthesia was initiated with isoflurane (3\%) and maintained with the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.
 A craniotomy was performed over the right primary visual cortex.
@@
 Two types of stimuli were presented for each imaging site: First, directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen directions of motion, 500 ms per direction, without blanks, with 12--30 trials for each direction of motion.
-Second, to measure correlations, the stimulus was modified to include only two directions of motion (in 9 datasets) or five directions (in 22 datasets) and the gratings were presented for 1 second and were separated by 1-second blanks, with 100--300 trials for each direction of motion.
+Second, ro measure correlations, the stimulus was modified to include only two directions of motion (in 9 datasets) or five directions (in 22 datasets) and the gratings were presented for 1 second and were separated by 1-second blanks, with 100--300 trials for each direction of motion.
 
@@
 \end{equation}
-by an expectation-maximization (EM) algorithm with specified rank of $L$. In addition, the diagonal matrix of individual variances is shrunk toward its mean value according to Eq.~\ref{eq:c-factor}.
+by an expectation-maximization (EM) algorithm with specified rank of $L$. Additional the diagonal matrix of individual variances is shrunk toward its mean value according to Eq.~\ref{eq:c-factor}.
 
